### 2022-06-09 A - Model Parameterization /Online Supplement A - Model Parameterization for Aggregated and Individual Models.Rmd
All documents except .rds

---
title             : "Online Supplement A: Model Parameterization for Aggregated and Individual Models"
author            : "" 
bibliography      : ["r-references.bib"]
header-includes   :
   - \usepackage{bm}
   - \usepackage{amsmath}
   - \usepackage{setspace}
   - \usepackage{xcolor}
   - \definecolor{mypink}{RGB}{255, 230, 255}
   - \usepackage{todonotes}
   - \usepackage{graphicx}
   - \usepackage{float}
   - \renewcommand{\figurename}{Figure A}
   - \makeatletter
   - \def\fnum@figure{\figurename\thefigure}
   - \makeatother

csl               : apa6.csl
output            : pdf_document
---
\spacing{1.5}
\fontsize{12}{12}

```{r setup, include=FALSE}
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

knitr::opts_chunk$set(tidy.opts=list(width.cutoff=40), tidy=TRUE, echo = TRUE, message = FALSE, warning = FALSE, fig.pos = "H")
```

\scriptsize

```{R loadpackages, message = FALSE, warning = FALSE, echo = FALSE}
# Load the packages
library("papaja")
library("LaplacesDemon")
library("rstan")
library("brms")
library("plyr")
library("lattice")
library("ggplot2")
library("dplyr")
library("readr")
library("rmarkdown")
library("Rmisc")
library("devtools")
library("gghalves") 
library("bayesplot")
library("bridgesampling") 
library("gridExtra") 
library("tibble") 
library("kableExtra") 
library("truncnorm")
library("ggbeeswarm")
```
\normalsize

This appendix illustrates the difference between hierarchical models and analyses of individual and aggregated data by showing the parameterization for the latter two. In case of an analysis of aggregated data where the *k*th observation/trial for the *i*th participant in the *j*th condition comes from a normal distribution with the mean equal to the intercept $\gamma$, the effect of the side (i.e., smaller or greater than 5, $\beta$) and the effect of the distance from 5 (i.e., symbolic distance effect, $\delta$’s), and the variance $\sigma^2$, the equation is as follows:

\begin{equation}
\tag{A1}
Y_{ijk} \sim \text{Normal}(\gamma + x_j \beta_{side} + u_j \delta_{7} + v_j \delta_{6} + w_j \delta_{4} + z_j \delta_{3}, \sigma^2), 
\end{equation}

where

<!-- \[ -->
<!-- x_j =  -->
<!-- \begin{cases} -->
<!-- \frac{1}{2} & j < 5\\ -->
<!-- - \frac{1}{2} & j > 5, \\ -->
<!-- \end{cases} -->

<!-- \; \; \; -->

<!-- \[ -->
<!-- u_j =  -->
<!-- \begin{cases} -->
<!-- 1 & j = 7\\ -->
<!-- 0 & \text{Otherwise}, -->
<!-- \end{cases} -->

<!-- \; \; \; -->

<!-- v_j =  -->
<!-- \begin{cases} -->
<!-- 1 & j = 6\\ -->
<!-- 0 & \text{Otherwise}, -->
<!-- \end{cases} -->
<!-- \] -->

\begin{align*}
&
\begin{aligned}
x_j = 
\begin{cases}
\frac{1}{2} & j < 5\\
- \frac{1}{2} & j > 5, \\
\end{cases}
\end{aligned}
&
\begin{aligned}
u_j = 
\begin{cases}
1 & j = 7\\
0 & \text{Otherwise},
\end{cases}
\end{aligned}
&
\begin{aligned}
v_j = 
\begin{cases}
1 & j = 6\\
0 & \text{Otherwise}.
\end{cases}
\end{aligned}
\end{align*}

<!-- \[ -->
<!-- w_j =  -->
<!-- \begin{cases} -->
<!-- 1 & j = 4\\ -->
<!-- 0 & \text{Otherwise}, -->
<!-- \end{cases} -->

<!-- \; \; \; -->

<!-- z_j =  -->
<!-- \begin{cases} -->
<!-- 1 & j = 3\\ -->
<!-- 0 & \text{Otherwise}. -->
<!-- \end{cases} -->
<!-- \] -->

\begin{align*}
&
\begin{aligned}
w_j = 
\begin{cases}
1 & j = 4\\
0 & \text{Otherwise},
\end{cases}
\end{aligned}
&
\begin{aligned}
z_j = 
\begin{cases}
1 & j = 3\\
0 & \text{Otherwise}.
\end{cases}
\end{aligned}
\end{align*}

The equation shows that, for instance, for participant *i*=1 and digit *j*=7, the mean RT is $\gamma - \frac{1}{2}\beta + \delta_{7}$. This equation shows that there is no individual effect of the condition. Some individual RTs could differ from the general pattern of mean RTs.

When it is decided to investigate individual RTs instead, the overall response time is ignored, resulting in the following equation (in case of the normal model) for *i*=1:

\begin{equation}
\tag{A2}
Y_{i=1, jk} \sim \text{Normal}(\gamma_{i = 1} + x_j \beta_{i = 1} + u_j \delta_{7, i = 1} + v_j \delta_{6, i = 1} + w_j \delta_{4, i = 1} + z_j \delta_{3, i = 1}, \sigma^2_{i = 1}), 
\end{equation}

The equation is very similar to the one shown before. However, notice that the intercept $\gamma_i$, the effects $\beta_{i}$ and $\delta$’s, and the variance $\sigma^2_{i}$ are now individual specific.

However, a hierarchical analysis matches the symbolic distance effect data the best. We have observations per person per condition. Therefore, we can investigate a general trend and individual deviations from this trend.

### 2022-06-09 B - Default Priors Rstan/Online Supplement B - Default Priors Rstan.Rmd
All documents except .rds

---
title             : "Online Supplement B: Default Priors Rstan"
author            : "" 
bibliography      : ["r-references.bib"]
header-includes   :
   - \usepackage{bm}
   - \usepackage{amsmath}
   - \usepackage{setspace}
   - \usepackage{xcolor}
   - \definecolor{mypink}{RGB}{255, 230, 255}
   - \usepackage{todonotes}
   - \usepackage{graphicx}
   - \usepackage{float}
   - \renewcommand{\figurename}{Figure B}
   - \makeatletter
   - \def\fnum@figure{\figurename\thefigure}
   - \makeatother

csl               : apa6.csl
output            : pdf_document
---
\spacing{1.5}
\fontsize{12}{12}

```{r setup, include=FALSE}
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

knitr::opts_chunk$set(tidy.opts=list(width.cutoff=40), tidy=TRUE, echo = TRUE, message = FALSE, warning = FALSE, fig.pos = "H")
```

\scriptsize

```{R loadpackages, message = FALSE, warning = FALSE, echo = FALSE}
# Load the packages
library("papaja")
library("LaplacesDemon")
library("rstan")
library("brms")
library("plyr")
library("lattice")
library("ggplot2")
library("dplyr")
library("readr")
library("rmarkdown")
library("Rmisc")
library("devtools")
library("gghalves") 
library("bayesplot")
library("bridgesampling") 
library("gridExtra") 
library("tibble") 
library("kableExtra") 
library("truncnorm")
library("ggbeeswarm")
```
\normalsize

Software packages have default prior settings that ensure estimation of the results even if the users do not choose priors themselves. It is important to realize that many of these default priors do not take the complexity of the hierarchical models into account (i.e., the hierarchical structure and the resulting number of parameters), therefore, it is *not* recommended to use the default priors for Bayesian hierarchical modeling. Here, we will provide details on the default priors for `rstan`, show predictions for data from these priors, and highlight resulting issues.

### Rstan 
When no prior is specified in `rstan` [@R-rstan], a uniform distribution is applied using the defined range of the variable [@carpenter2017stan]. The uniform distribution is shown in the figure below.

```{r uniformdistributionplot, fig.cap="Different uniform distributions that vary in range (i.e., minimum and maximum).", fig.align = "left", fig.height = 3, fig.width = 6, echo = FALSE}
# all set 1, 1

parametersunidistrplot <- c("Min = 0, Max = 1" = "#D95F02", "Min = 0, Max = 3" = "#1B9E77", "Min = 0, Max = 10" = "#E7298A")

unidistriplot <- ggplot(data = data.frame(x = c(-5, 5)), aes(x)) +
  stat_function(fun = dunif, n = 1001, args = list(min = 0, max = 1), linetype = 2, geom = "area", alpha = 0.3, position = "identity", aes(color = "Min = 0, Max = 1", fill = "Min = 0, Max = 1")) +  ylab("Density") +
  stat_function(fun = dunif, n = 1001, args = list(0, 3), linetype = 3, geom = "area", alpha = 0.3, position = "identity", aes(color = "Min = 0, Max = 3", fill = "Min = 0, Max = 3")) + 
  stat_function(fun = dunif, n = 1001, args = list(0, 10), geom = "area", alpha = 0.3, position = "identity", aes(color = "Min = 0, Max = 10", fill = "Min = 0, Max = 10")) + geom_hline(yintercept=0, colour="white", size=1) + scale_color_manual(name = "Settings", values = parametersunidistrplot, aesthetics = c("colour", "fill")) + xlim(-1, 10) + theme(legend.title = element_text(size = 8), legend.key.width=unit(0.45,"cm"), legend.key.height=unit(0.5,"cm")) + xlab("Parameter value")

theme_set(theme_apa(base_size = 9))
unidistriplot

```

The figure shows the uniform distribution for different settings of the minimum and maximum. The minimum and maximum define the range of the distribution where the minimum represents the lower bound and the maximum the upper bound. The figure illustrates that increasing the range, would result in wider distribution (i.e., the pink distribution is much wider than the green and orange distribution). In contrast, decreasing the range would result in a very specific prior with the probability spread out over a narrow range of values (i.e., the orange distribution is very narrow compared to the green and pink distribution).

However, if the range of the variable is not defined in `Stan`, the bounds will revert to $-\infty$ and $\infty$, resulting in an improper prior. This means that the probability distribution is not proper: integrating over a range of $-\infty$ to $\infty$ does not result in 1.

What would happen if these default priors were applied to the symbolic distance effect? In our scenario, we do not have a specified range for the parameters. Only the lower bound of the variances is specified as 0 (i.e., only positive variances are possible). Therefore, $\sigma^2$ and the hyperpriors can be defined in the following way:

\begin{equation}
\begin{aligned}
\mu_{\gamma}, \mu_{\beta}, \mu_{\delta}, \sigma^2, \sigma^2_{\gamma}, \sigma^2_{\beta}, \sigma^2_{\delta}, &\sim \text{U}(-\infty, \infty), \\ 
\sigma^2, \sigma^2_{\gamma}, \sigma^2_{\beta}, \sigma^2_{\delta}, &\sim \text{U}(0, \infty). 
\end{aligned}
\end{equation}

All mean parameters have a flat prior, ranging from minus  $-\infty$ to $\infty$, while the prior on the variances ranges from zero to $\infty$. Since at least one of the bounds (i.e., minimum or maximum) is not defined, the priors are improper. We cannot sample from $-\infty$ to $\infty$. As a result, prior visualization and prior prediction is impossible. However, `rstan` would still provide results as it does not identify improper priors [@stan2018a, section 24.3].

```{r priorvisualizationstan, fig.cap="Visualization of \\textit{rstan} default prior setting for our model. On the left, for every parameter the prior distribution is presented. On the right, the prior distribution for prior parameters of the parameters is shown.", fig.align = "left", fig.height=6, fig.width=8, warning=FALSE, message=FALSE, eval = FALSE, echo = FALSE}

# Plots 
x <- seq(-3, 3, by=.01)
# color codes: https://flaviocopes.com/rgb-color-codes/

# Instead do this, take integral  
Nplotst <- 1000000
mugammaplotst <- runif(Nplotst, -Inf, Inf)
vargammaplotst <- runif(Nplotst, 0, Inf)
betaplot2st <- rnorm(Nplotst, mugammaplotst, sqrt(vargammaplotst))

# library(ggridges)
pgamma2datast <- data.frame(betaplot2st)
pgamma2st <- ggplot(pgamma2datast, aes(x = betaplot2st)) + geom_density_line(color = "#0072B2", fill = "#0072B2", alpha = 0.1, position = "identity") + ylab("Density") + xlab("x") + labs(title = expression(~gamma[i])) + theme(plot.title = element_text(size = 18)) + xlim(-5, 5)

parametersgammast <- c("Mean" = "#D95F02", "Variance" = "#1B9E77")

ptogethergamma2st <- ggplot(data = data.frame(x = c(-5, 5)), aes(x)) +
  stat_function(fun = dunif, n = 1001, args = list(min = 0, max = 1), aes(color = "Mean", fill = "Mean"), geom = "area", alpha = 0.3) + ylab("Density") +
  stat_function(fun = dunif, n = 1001, args = list(min = 0, max = 1), geom = "area", aes(color = "Variance", fill = "Variance"), alpha = 0.3, position = "identity") + geom_hline(yintercept=0, colour="white", size=1) + scale_color_manual(name = "Prior parameters", values = parametersgammast, aesthetics = c("colour", "fill"), labels = c(expression(~mu[~gamma]), expression(~sigma[~gamma]^2))) + 
  theme(legend.text=element_text(size = 9), legend.title = element_text(size = 7), legend.key.width=unit(0.45,"cm"), legend.key.height=unit(0.5,"cm"))

# Beta 
mubetaplotst <- runif(Nplotst, 0, 1)
varbetaplotst <- runif(Nplotst, 0, 1)
betaplot22st <- rnorm(Nplotst, mubetaplotst, sqrt(varbetaplotst))

pbeta2datast <- data.frame(betaplot22st)
pbeta2st <- ggplot(pbeta2datast, aes(x = betaplot22st)) + geom_density_line(color = "#0072B2", fill = "#0072B2", alpha = 0.1, position = "identity") + ylab("Density") + xlab("x") + labs(title = expression(~beta["side, i"])) + theme(plot.title = element_text(size = 18)) + xlim(-5, 5)

ptogetherbeta2st <- ggplot(data = data.frame(x = c(-5, 5)), aes(x)) +
  stat_function(fun = dunif, n = 1001, args = list(min = 0, max = 1), aes(color = "Mean", fill = "Mean"), geom = "area", alpha = 0.3) + ylab("Density") +
  stat_function(fun = dunif, n = 1001, args = list(min = 0, max = 1), geom = "area", aes(color = "Variance", fill = "Variance"), alpha = 0.3, position = "identity") + geom_hline(yintercept=0, colour="white", size=1) + scale_color_manual(name = "Prior parameters", values = parametersgamma, aesthetics = c("colour", "fill"), labels = c(expression(~mu[~beta]), expression(~sigma[~beta]^2))) + 
  theme(legend.text=element_text(size = 9), legend.title = element_text(size = 7), legend.key.width=unit(0.45,"cm"), legend.key.height=unit(0.5,"cm"))

# Delta's

mudeltaplotst <- runif(Nplotst, 0, 1)
vardeltaplotst <- runif(Nplotst, 0, 1)
deltaplot2st <- rnorm(Nplotst, mudeltaplotst, sqrt(vardeltaplotst))

pdelta2datast <- data.frame(deltaplot2st)
pdelta2st <- ggplot(pdelta2datast, aes(x = deltaplot2st)) + geom_density_line(color = "#0072B2", fill = "#0072B2", alpha = 0.1, position = "identity") + ylab("Density") + xlab("x") + labs(title = expression(~delta[i])) + theme(plot.title = element_text(size = 18)) + xlim(-5, 5)

ptogetherdelta2st <- ggplot(data = data.frame(x = c(-5, 5)), aes(x)) +
  stat_function(fun = dunif, n = 1001, args = list(min = 0, max = 1), aes(color = "Mean", fill = "Mean"), geom = "area", alpha = 0.3) + ylab("Density") +
  stat_function(fun = dunif, n = 1001, args = list(min = 0, max = 1), geom = "area", aes(color = "Variance", fill = "Variance"), alpha = 0.3, position = "identity") + geom_hline(yintercept=0, colour="white", size=1) + scale_color_manual(name = "Prior parameters", values = parametersgammast, aesthetics = c("colour", "fill"), labels = c(expression(~mu[~delta]), expression(~sigma[~delta]^2))) + 
  theme(legend.text=element_text(size = 9), legend.title = element_text(size = 7), legend.key.width=unit(0.45,"cm"), legend.key.height=unit(0.5,"cm"))

theme_set(theme_apa(base_size = 9))
#ptogether

# Final figure 
# grid.arrange(pgamma2st, ptogethergamma2st, pbeta2st, ptogetherbeta2st, pdelta2st, ptogetherdelta2st, nrow = 3, ncol = 2)

```

```{r priorpredictionsim4stan, warning=FALSE, message=FALSE, cache=TRUE, cache.lazy=FALSE, eval=FALSE, echo = FALSE}

# Without for loop 
set.seed(123)  # keep so people can reproduces results?

datalistpp3st <- list()
for (i in 1:100){
# Hyperpriors: These are not individual specific right? So only sample 1? 
mugammappst <- runif(1, 0, 1)
vargammappst <- runif(1, 0, 1)
mubetappst <- runif(1, 0, 1)
varbetappst <- runif(1, 0, 1)
mudeltappst <- runif(1, 0, 1)
vardeltappst <- runif(1, 0, 1)

# Priors: for each person sample prior 1 time 
gammaippst <- rnorm(Ipp, mugammappst, vargammappst)
betaippst <-  rnorm(Ipp, mubetappst, varbetappst)
delta1ppst <- rnorm(Ipp, mudeltappst, vardeltappst)
delta2ppst <- rnorm(Ipp, mudeltappst, vardeltappst)
delta3ppst <- rnorm(Ipp, mudeltappst, vardeltappst)
delta4ppst <- rnorm(Ipp, mudeltappst, vardeltappst)

sigmappst <- runif(Ipp, 0, 1)  # Only 1 because not individual specific?

# Or start for loop here? Or do we want to vary persons as well? 
# data
Yppst <- rnorm(Ipp*Jpp*Kpp, gammaippst[exppp] + xjpp * betaippst[exppp] + 
              ujpp * delta1ppst[exppp] + vjpp * delta2ppst[exppp] +
              wjpp * delta3ppst[exppp] + zjpp * delta4ppst[exppp],
              sqrt(sigmappst))
datartppst <- data.frame(exppp, condpp, subpp, repetition = rep(i, length(exppp)), gammaippst[exppp], betaippst[exppp], delta1ppst[exppp], delta2ppst[exppp], delta3ppst[exppp], delta4ppst[exppp], Yppst)
datalistpp3st[[i]] <- datartppst
}

# Add datasets under each other 
bigdatapp3st <- do.call(rbind, datalistpp3st)
# hist(bigdatapp3st$Ypp)

# histpp <- NA
# for (i in 1:100){
#  histpp[i] <- hist(bigdatapp3$Ypp[bigdatapp3$repetition==i])
# }
```

# References 
\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
&nbsp;

### 2022-06-09 C - Tutorial Normal Model Rstan/Online Supplement C - Bayesian Hierarchical Modeling in rstan.Rmd
All documents except .rds

---
title             : "Online Supplement C: Bayesian Hierarchical Modeling in rstan"
author            : "Myrthe Veenman" 
bibliography      : ["r-references.bib"]
header-includes   :
   - \usepackage{bm}
   - \usepackage{amsmath}
   - \usepackage{setspace}
   - \usepackage{xcolor}
   - \definecolor{mypink}{RGB}{255, 230, 255}
   - \usepackage{todonotes}
   - \usepackage{graphicx}
   - \usepackage{float}

csl               : apa6.csl
output            : pdf_document
---
\spacing{1.5}
\fontsize{12}{12}

```{r setup, include=FALSE}
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

knitr::opts_chunk$set(tidy.opts=list(width.cutoff=40), tidy=TRUE, echo = TRUE, message = FALSE, warning = FALSE, fig.pos = "H")
```

|    This online supplement provides a tutorial on Bayesian Hierarchical models (BHM) in `rstan` [@R-rstan]. In addition, it contains information about the Number Classification Task that could not be included in the paper. Only the fit of the full model will be shown as the fitting the others should follow logically from this tutorial. Technical details, such as explanations of statistical terms (e.g., $\hat{R}$), will not be discussed here but can be found in the paper. This tutorial will cover the following:

1. Required packages 
2. Input  
  2.1 Stan file   
  2.2 Data 
3. Fit model 
4. Output   
  4.1 General Effects   
  4.2 Individual Effects 
5. Model comparison 
6. Additional resources

## Required Packages

|    In this practical, the following packages will be used:

- For the document layout: *rmarkdown* [Version 1.16; @R-rmarkdown], *papaja* [Version 0.1.0.9842; @R-papaja], *knitr* [Version 1.25; @R-knitr], *kableExtra* [Version 1.1.0; @R-kableExtra]
- For data structuring: *LaplacesDemon* [Version 16.1.4; @R-LaplacesDemon], *plyr* [Version 1.8.4; @R-plyr], *dplyr* [Version 0.8.3; @R-dplyr], *readr* [Version 1.3.1; @R-readr]
- To fit the model: *rstan* [Version 2.19.2; @R-rstan]
- For the visualization of the results: *lattice* [Version 0.20.38; @R-lattice], *ggplot2* [Version 3.3.2; @R-ggplot2], *Rmisc* [Version 1.5; @R-Rmisc], *devtools* [Version 2.2.1; @R-devtools], *gghalves* [Version 0.1.0; @R-gghalves], *bayesplot* [Version 1.7.1; @R-bayesplot], *brms* [Version 2.13.0; @R-brms_a; @R-brms_b], *gridExtra* [Version 2.3; @R-gridExtra], *tibble* [Version 3.0.2; @R-tibble]
- For model comparison: *bridgesamping* [Version 1.0.0; @R-bridgesampling]

&nbsp;

\scriptsize

```{R loadpackages, message = FALSE, warning = FALSE}
# Load the packages
library("papaja")
library("LaplacesDemon")
library("rstan")
library("brms")
library("plyr")
library("lattice")
library("ggplot2")
library("dplyr")
library("readr")
library("rmarkdown")
library("Rmisc")
library("devtools")
library("gghalves") 
library("bayesplot")
library("bridgesampling") 
library("gridExtra") 
library("tibble") 
library("kableExtra") 
library("truncnorm")
library("ggbeeswarm")
```
\normalsize

## Input

|    To fit a BHM in *rstan*, we use the *stan* function, as shown below. 
&nbsp;

\scriptsize 
```{r modelfitpreview, eval=FALSE, echo=TRUE}
model_fit <- stan(file = "./myModel.stan",  # Stan file with model  
                  data = myData_list,       # List with observed data and constants
                  iter = 4000,              # Number (Nr.) of iterations per chain
                  chains = 4,               # Nr. of chains 
                  warmup = 1000)            # Nr. iterations for warmup per chain
```
\normalsize

We can see from the code above that the function requires the following input:

- Stan file 
- A list with the data

The other options do not require coding and are discussed in the paper.

### Stan file

|  The stan file `myModel.stan` is used to fit the full model and can be found under R objects > rstan > linear model. It contains descriptions for every line of code. The stan file is divided in three sections (in our case):

1. Data 
2. Parameters 
3. Model

|    As the name says, the Data section specifies the variables we will provide with data. It contains the number of total observations, the number of groups (in our case the number of individuals), the response times, the condition for every response time (side of the digit and the digit indicator) and the prior settings. There are different types of variables. A vector indicates that the variable consists of several numbers and between `[]` the length of the vector is given. An integer indicates that the variable is one full number (i.e., no half numbers such as 0.3, 1.5, 6.33). Real means that the number can be a half number, so it allows numbers as 0.5, 1.3, 5.66, etc. With `<lower>` and `<upper>` the lower and upper bound of the parameter can be specified. For instance, for a variance it could be indicated that this variable cannot be below 0 with `<lower = 0>`.     
|    In the Parameters section, all the parameters that need to be estimated are given. These are the general and individual effects. The specification of the parameters works in the same way as the variable specification in the Data section. Therefore, the general effects are reals, while the individual parameters are vectors with a length of the number of groups (in our case individuals, this means that there will be an estimate for every individual).  
|    The last section, Model, contains the formula to obtain the estimates. We use the target specification as it saves the constants, necessary for model comparison later on. For example, `target += normal_lpdf(delta2 | mu4, sqrt(g4))` basically means that "delta2" is normally distributed with a mean of "mu4" and a standard deviation of "g4". Note that we want to estimate the variances, not the standard deviations. However, in the normal distribution function, the standard deviation is asked for. Therefore, we computed the square root of the variance parameters. Furthermore, it is important that you end every line with `;`, and end the stan file with an empty line. With `//` you can add a comment to the code.

### Data

|    In the function, we put a list that contains all the information corresponding to the Data section of the stan file. This means that every parameter set in the Data section of the stan file, should have an element in the list we put in the function. In our case, this should result in 23 elements in the list. 
|  First, we will load the data and clean it according to the requirements in the paper. 
&nbsp;

\scriptsize 
```{r readdata}
indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
colnames(indat)=c('sub','block','trial','stim','resp','rt','error')

## Cleaning the data according to criteria discussed in paper  
# (code retrieved from Julia Haaf:
# https://github.com/PerceptionAndCognitionLab/bf-order/blob/public/papers/submission/R-scripts/ld5.R)
clean=function()
{
  indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
  colnames(indat)=c('sub','block','trial','stim','resp','rt','error')

  bad1=indat$sub%in%c(34,43)
  bad2=indat$rt<250 | indat$rt>2000
  bad3=indat$err==1
  bad4=indat$block==0 & indat$trial<20
  bad5=indat$trial==0

  bad=bad1 | bad2 | bad3 |bad4 |bad5
  dat=indat[!bad,]
  return(dat)
}

indat1 <- clean()  # Final dataset 
```
\normalsize

|    Now, we will construct the list we provide the `stan` function. The first element of the list is N, corresponding to the number of groups in the data. In our case these are the number of individuals. The second element is the number of total observations in the data. Then, we specify the dependent variable/observations. In our case these are the response times (RTs). Next, we add an indicator to the list that specifies which observation belongs to which individual (or group). 
&nbsp;

\scriptsize 
```{r}
# Data into list and fill the list with necessary information 
# Number of groups (individuals)
myData_list <- list(N = length(unique(indat1$sub)))

# Number of total observations 
myData_list$All <- nrow(indat1)

# Observations 
myData_list$y <- indat1$rt/1000  # rt in seconds instead of milliseconds

# Need a variable with group number that are chronological (that doesn't skip numbers)
for (j in 1:nrow(indat1)){
  if (j == 1) {
    indat1$sub1[j] <- 1}
  else if (indat1$sub[j] == indat1$sub[j-1]) {
    myData_list$group_inds[j] <- indat1$sub1[j-1]}
  else myData_list$group_inds[j] <- indat1$sub1[j-1] + 1
}

# myData_list$group_inds <- indat1$sub1  # Add variable with indicator person to the list used for the analysis

```
\normalsize

|    The next step is to specify the indicators. These are the variables that specify per observation of a variable applies or not. For the side parameter this means that the side indicator is $\frac{1}{2}$ for observations where the digit was smaller than 5, and  $-\frac{1}{2}$ for observations where the digit was greater than 5. For the digit parameters it means that the digit indicator is 1 when for the observation the digit was equal to 7, 6, 4 or 3, and 0 when it was not. 
&nbsp;

\scriptsize 
```{r digitparameters}
### Add variable with information of the side of the digit to the list, for parameter beta    
# j = condition, if j > 5 than x = -1/2, if j < 5 than x = 1/2 
# indat1$stim 0 = 2, 1=3, 2=4, 3=6, 4=7, 5=8 
# so if bigger than 2 than 1/2, else than -1/2 
for (j in 1:nrow(indat1)){
  if (indat1$stim[j] < 3) {
    myData_list$side[j] <- 1/2}
  else myData_list$side[j] <- -1/2
}

### Add variable with information about difference between digits to the list, for parameters delta's 
#### Difference 8 and 7 
for (j in 1:nrow(indat1)){ 
  if (indat1$stim[j] == 5) {  # 5 = 8 , reference: https://github.com/PerceptionCognitionLab/data0/blob/master/lexDec-dist5/ld5.txt
    myData_list$dif1[j] <- 0}
  else if (indat1$stim[j] == 4) {
    myData_list$dif1[j] <- 1}
  else myData_list$dif1[j] <- 0
}

#### Difference 7 and 6 
for (j in 1:nrow(indat1)){ 
  if (indat1$stim[j] == 4) {  # 4 = 7 , reference: https://github.com/PerceptionCognitionLab/data0/blob/master/lexDec-dist5/ld5.txt
    myData_list$dif2[j] <- 0}
  else if (indat1$stim[j] == 3) {  # 3 = 6 
    myData_list$dif2[j] <- 1}
  else myData_list$dif2[j] <- 0
}

#### Difference 4 and 3 
for (j in 1:nrow(indat1)){ 
  if (indat1$stim[j] == 1) {  # 1 = 3, 2 = 4 , reference: https://github.com/PerceptionCognitionLab/data0/blob/master/lexDec-dist5/ld5.txt
    myData_list$dif3[j] <- 0}
  else if (indat1$stim[j] == 2) {
    myData_list$dif3[j] <- 1}
  else myData_list$dif3[j] <- 0
}

#### Difference 3 and 2 
for (j in 1:nrow(indat1)){ 
  if (indat1$stim[j] == 0) {  # 0 = 2, 1 = 3 , reference: https://github.com/PerceptionCognitionLab/data0/blob/master/lexDec-dist5/ld5.txt
    myData_list$dif4[j] <- 0}
  else if (indat1$stim[j] == 1) {
    myData_list$dif4[j] <- 1}
  else myData_list$dif4[j] <- 0
}

# datareal$dif1 <- indat1$dif1
# datareal$dif2 <- indat1$dif2
# datareal$dif3 <- indat1$dif3 
# datareal$dif4 <- indat1$dif4

```
\normalsize

|  Finally, we specify the priors. The prior specification is discussed in the paper.
&nbsp;

\scriptsize 
```{r priors}
### Set the priors and add them to the list 
a <- 0.5        # mean for mu, mu is the mean of theta  
b <- 1          # variance for mu 
c <- 3          # mean for sigma 
d <- 0.7        # variance for sigma, sigma is the variance of observations  
e <- 3          # mean for g, g is the variance of theta   
f <- 0.7        # variance for g

#### Priors for beta 
a2 <- 0        # mean for mean of beta
b2 <- 0.09     # variance for mean of beta, used to be 1 
e2 <- 3        # mean for variance of beta, used to be .7, adjusted 
f2 <- 0.5      # variance for variance of beta, used to be .2., adjusted

#### Prior for deltas, same for every delta 
a3 <- 0        # mean for mean of delta 
b3 <- 0.09     # sd for mean of delta, 0.5, 0.3  
dd3 <- 3       # mean for variance of delta, adjusted
f3 <- 0.5      # variance for variance of delta, adjusted

myData_list$a <- a
myData_list$b <- b
myData_list$c <- c
myData_list$d <- d
myData_list$dd <- e
myData_list$f <- f
myData_list$a2 <- a2
myData_list$b2 <- b2
myData_list$dd2 <- e2
myData_list$f2 <- f2
myData_list$a3 <- a3
myData_list$b3 <- b3
myData_list$dd3 <- dd3
myData_list$f3 <- f3
```
\normalsize

As described in the paper, we could check whether these priors are reasonable considering our expectations by performing prior prediction.

```{r priorprediction, echo = FALSE, fig.cap="Predicted mean RT difference between conditions per trial. The blue dot represents the aggregated effect. ", fig.align = "left", fig.height=4, fig.width = 5}
set.seed(123)

Ipp <- 10  # repeat 10 times 
Jpp <- 6   # digit 2, 3, 4, 6, 7, and 8 
Kpp <- 60  # 60 observations per condition per person

exppp <- rep(1:Ipp, each = Jpp * Kpp)  # which row belongs to which participant 
condpp <- rep(0:5, Ipp, each = Kpp)  # to which condition row belongs  
subpp <- rep(1:Kpp, Ipp * Jpp)  # to which observation per condition per parcticipant belongs row

# add parameters x, u, v, w, z 
xjpp <- NA 
ujpp <- NA 
vjpp <- NA 
wjpp <- NA 
zjpp <- NA 
# side effect, greater or smaller than 5  
for (j in 1:length(condpp)){  # 3 = 6
  if (condpp[j] < 3) {
    xjpp[j] <- 1/2}
  else xjpp[j] <- -1/2
}

### Add variable with information about difference between digits, for parameters delta's 
#### Difference 8 and 7 
for (j in 1:length(condpp)){ 
  if (condpp[j] == 5) {  # 5 = 8 , 
    ujpp[j] <- 0}
  else if (condpp[j] == 4) {
    ujpp[j] <- 1}
  else ujpp[j] <- 0
}

# Difference 7 and 6 
for (j in 1:length(condpp)){ 
  if (condpp[j] == 4) {  # 4 = 7, 3 = 6  
    vjpp[j] <- 0}
  else if (condpp[j] == 3) {
    vjpp[j] <- 1}
  else vjpp[j] <- 0
}

# Difference 4 and 3 
for (j in 1:length(condpp)){ 
  if (condpp[j] == 1) {  # 1 = 3, 2 = 4 
    wjpp[j] <- 0}
  else if (condpp[j] == 2) {
    wjpp[j] <- 1}
  else wjpp[j] <- 0
}

# Difference 3 and 2 
for (j in 1:length(condpp)){ 
  if (condpp[j] == 0) {  # 0 = 2, 1 = 3
    zjpp[j] <- 0}
  else if (condpp[j] == 1) {
    zjpp[j] <- 1}
  else zjpp[j] <- 0
}

datalistpp3 <- list()
for (i in 1:1000){
  # Hyperpriors: These are not individual specific right? So only sample 1? 
  mugammapp <- rtruncnorm(1, a = 0, mean = 0.5, sd = sqrt(1))
  vargammapp <- rinvgamma(1, e, f)
  mubetapp <- rnorm(1, a2, sqrt(b2))
  varbetapp <- rinvgamma(1, e2, f2)
  mudeltapp <- rnorm(1, a3, sqrt(b3))
  vardeltapp <- rinvgamma(1, dd3, f3)

  # Priors: for each person sample prior 1 time 
  gammaipp <- rnorm(Ipp, mugammapp, sqrt(vargammapp))
  betaipp <-  rnorm(Ipp, mubetapp, sqrt(varbetapp))
  delta1pp <- rnorm(Ipp, mudeltapp, sqrt(vardeltapp))
  delta2pp <- rnorm(Ipp, mudeltapp, sqrt(vardeltapp))
  delta3pp <- rnorm(Ipp, mudeltapp, sqrt(vardeltapp))
  delta4pp <- rnorm(Ipp, mudeltapp, sqrt(vardeltapp))

  sigmapp <- rinvgamma(1, c, d)  # Only 1 because not individual specific?

  # Or start for loop here? Or do we want to vary persons as well? 
  # data
  Ypp <- rnorm(Ipp*Jpp*Kpp, gammaipp[exppp] + xjpp * betaipp[exppp] + 
                 ujpp * delta1pp[exppp] + vjpp * delta2pp[exppp] +
                 wjpp * delta3pp[exppp] + zjpp * delta4pp[exppp],
               sqrt(sigmapp))
  datartpp <- data.frame(exppp, condpp, subpp, repetition = rep(i, length(exppp)), gammaipp[exppp], betaipp[exppp], delta1pp[exppp], delta2pp[exppp], delta3pp[exppp], delta4pp[exppp], Ypp)
  datalistpp3[[i]] <- datartpp
}

# Add datasets under each other 
bigdatapp3_rstan <- do.call(rbind, datalistpp3)

########### Create figure function

## Or plot mean rt per condition per trial 
  priorpredfigurecreate <- function(bigdatapp3brms){
    plotsimpp <- aggregate(bigdatapp3brms$Ypp, list(bigdatapp3brms$condpp, bigdatapp3brms$repetition), mean)
    meanrtsimpp <- aggregate(bigdatapp3brms$Ypp, list(bigdatapp3brms$condpp), mean)  # mean RT per condition

    # Mean difference per digits per trial and side effect (mean(2, 3, 4) - mean (6, 7, 8))
    plotsimpp <- aggregate(bigdatapp3brms$Ypp, list(bigdatapp3brms$condpp, bigdatapp3brms$repetition), mean)
    # meanrtsimpp <- aggregate(bigdatapp3$Ypp, list(bigdatapp3$condpp), mean)

    sideapp3 <- NA 
    diff1app3 <- NA   # length 1000 
    diff2app3 <- NA 
    diff3app3 <- NA
    diff4app3 <- NA 
    repapp3 <- NA

    # Group.2 is the trial number (1000 in total per condition)
    for (i in 1:length(unique(plotsimpp$Group.2))){
      repapp3[i] <- i 
      sideapp3[i] <- ((plotsimpp$x[plotsimpp$Group.1==0 & plotsimpp$Group.2 == i] + plotsimpp$x[plotsimpp$Group.1==1 & plotsimpp$Group.2 == i] + plotsimpp$x[plotsimpp$Group.1==2 & plotsimpp$Group.2 == i]) - 
                     (plotsimpp$x[plotsimpp$Group.1==3 & plotsimpp$Group.2 == i] + plotsimpp$x[plotsimpp$Group.1==4 & plotsimpp$Group.2 == i] + plotsimpp$x[plotsimpp$Group.1==5 & plotsimpp$Group.2 == i]))
      diff1app3[i] <- (plotsimpp$x[plotsimpp$Group.1==4 & plotsimpp$Group.2 == i]- plotsimpp$x[plotsimpp$Group.1==5 & plotsimpp$Group.2 == i])  # 4 = 7 and 5 = 8 
      diff2app3[i] <- (plotsimpp$x[plotsimpp$Group.1==3 & plotsimpp$Group.2 == i]- plotsimpp$x[plotsimpp$Group.1==5 & plotsimpp$Group.2 == i])  # 3 = 6, 4 = 7

      diff3app3[i] <- (plotsimpp$x[plotsimpp$Group.1==2 & plotsimpp$Group.2 == i]- plotsimpp$x[plotsimpp$Group.1==0 & plotsimpp$Group.2 == i])  # 2 = 4, 1 = 3

      diff4app3[i] <- (plotsimpp$x[plotsimpp$Group.1==1 & plotsimpp$Group.2 == i]- plotsimpp$x[plotsimpp$Group.1==0 & plotsimpp$Group.2 == i])  # 1 = 3, 0 = 2

    }
    #diffdatapp3 <- data.frame(repapp3, diff1app3, diff2app3, diff3app3, diff4app3)  # length 1000 
    diffdatapp33 <- data.frame(trial = rep(repapp3, 5), condition = rep(1:5, each = length(repapp3)), difference =  c(sideapp3, diff4app3, diff3app3, diff2app3, diff1app3))

    # difference aggregated 
    diffmeandatapp33 <- data.frame(condition = 1:5, 
                                   difference =  c(((meanrtsimpp$x[meanrtsimpp$Group.1==0] + meanrtsimpp$x[meanrtsimpp$Group.1==1] + meanrtsimpp$x[meanrtsimpp$Group.1==2])-(meanrtsimpp$x[meanrtsimpp$Group.1==3] + meanrtsimpp$x[meanrtsimpp$Group.1==4] + meanrtsimpp$x[meanrtsimpp$Group.1==5])), 
                                                   (meanrtsimpp$x[meanrtsimpp$Group.1==1] - meanrtsimpp$x[meanrtsimpp$Group.1==0]), 
                                                   (meanrtsimpp$x[meanrtsimpp$Group.1==2] - meanrtsimpp$x[meanrtsimpp$Group.1==1]), 
                                                   (meanrtsimpp$x[meanrtsimpp$Group.1==3] - meanrtsimpp$x[meanrtsimpp$Group.1==4]), 
                                                   (meanrtsimpp$x[meanrtsimpp$Group.1==4] - meanrtsimpp$x[meanrtsimpp$Group.1==5])))

    # Old figure 
    diffrtplotsimpp <- ggplot(diffdatapp33, aes(y = difference, x = as.factor(condition))) + 
      geom_line(aes(group = trial), alpha = .014) + 
      geom_point(data = diffdatapp33, aes(group = trial), alpha = 0.014) + 
      geom_line(data = diffmeandatapp33, aes(group = 1), color='blue') + 
      geom_point(data = diffmeandatapp33, aes(group = 1), color='blue') + 
      labs(x = "Digits", y = "Response Time Difference (Seconds)") + 
      theme_bw() + 
      theme(axis.line = element_line(colour = "black"), panel.border = element_blank(), panel.background = element_blank()) + 
      scale_x_discrete(breaks = 1:5, labels = c("side", "3-2", "4-3", "6-7", "7-8"))   # + ylim(-0.1,0.1) then you see the difference better but missing a lot of trials 
      theme_set(theme_apa(base_size = 9))

    priorpredictionplot1 <- diffrtplotsimpp + ylim(-15,15)

    # Try new figure

    # Create violin plot 
    priorpredictionplotnew_1 <- ggplot(diffdatapp33, aes(y = difference, x = as.factor(condition))) + 
      geom_violin(width = 1, alpha = 0.1) + 
      geom_quasirandom(alpha = 0.014, width = 0.2, dodge.width=1) + 
      geom_point(data = diffmeandatapp33, aes(group = 1), color='blue') + 
      labs(x = "Effects", y = "Response Time Difference (Seconds)") + 
      theme_bw() + 
      theme(axis.line = element_line(colour = "black"), panel.border = element_blank(), panel.background = element_blank()) + 
      scale_x_discrete(breaks = 1:5, labels = c("Side", "3-2", "4-2", "6-8", "7-8"))   # + ylim(-0.1,0.1) then you see the difference better but missing a lot of trials

  return(priorpredictionplotnew_1)
}

########### End figure function

# Create figure 
priorpredictionplotnew_2 <- priorpredfigurecreate(bigdatapp3_rstan)

theme_set(theme_apa(base_size = 9))
priorpredictionplotnew_2
```

## Fit the Model

|    Now we have the required input, we can fit the model. 
&nbsp;

\scriptsize 
``` {r modelfit, eval = FALSE}
library(rstan) 
model_fit <- stan(file = "myPath/rewritestanmodel3_julia.stan", 
                  data = myData_list, 
                  iter = 4000, 
                  chains = 4,
                  warmup = 1000, 
                  cores = 4)
```
\normalsize

We have already fitted the model and saved it in the R object `model_fit`. In case you do not want to wait until the model has been fitted but continue immediately with the rest of this tutorial, you can load this R object that contains the model estimates.  
&nbsp;

\scriptsize 
```{r loadmodelfit, eval = FALSE}
model_fit <- readRDS(file = "myPath/rstan_model_fit.rds")
```
\normalsize 
```{r loadmodelfit2, echo = FALSE}
setwd('..')
model_fit <- readRDS(file = "R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
```

## Output

|    *Brms* offers the *launch_shinystan* function which will load a shiny app with the results of the model fit. It is very extensive. It, for instance, includes model diagnostics as trace plots, posterior plots, prediction plots and way more. 
&nbsp;

\scriptsize 
```{r shinyrstan, eval = FALSE}
launch_shinystan(model_fit)
```
\normalsize

It also has the possibility to save plots so you do not have to code them yourself. However, if you want the full control over the layout, it might be better to code them yourself. Therefore, we will show you how the program the plots presented in the paper.

|    The *summary* function provides an overview of the model fit. From this function, we can get a lot of information. As the output is structured in a specific way, it sometimes is hard to figure out how to isolate specific outcomes. [This document](https://cran.r-project.org/web/packages/rstan/vignettes/stanfit-objects.html) nicely explains how to obtain specific output from the model fit.

### Trace plot

|    The trace plot shows whether the posterior distribution of the parameter has converged. With the package's *plot* function it is possible to obtain the trace plots per parameter by setting the option *plotfun* to *trace*. By default, it will show the trace plots of the first ten parameters. With the *pars* function you can specify for which parameters you want to inspect the trace plots. The trace plot will automatically remove the warm-up (burn-in) period from the chain (which is desirable), however, if you would like to see that as well you have to set the option *inc_warmup* to *TRUE*. As the *plot* option returns a *ggplot* object, it is possible to manually adjust the plot using *ggplot* functions. We created the function *traceplotfunc* to obtain the trace plot for the parameter you are interested in. In the figure below, we show the trace plots for all the general effects. 
&nbsp;

\scriptsize 
```{r traceplotB, fig.cap="Trace plots of the general effect parameters.", fig.align = "left", fig.height = 6, fig.width = 8}

# Trace plot for parameter mu4
trdif2rstan <- plot(model_fit,     # Model fit 
                    plotfun = "trace",   # Specify that you want trace plots 
                    pars = c("mu4"))     # Select parameter, 
#it is also possible to select more than one at the same time, such as c("mu2", "mu3","mu4")

# In case you want to see the warmup period as well 
trdif2rstanwarmup <- plot(model_fit,    # Model fit 
                          plotfun = "trace",  # Specify that you want trace plots 
                          pars = c("mu4"),    # Select parameter
                          inc_warmup = TRUE)  # Show warmup period as well

# We created a function, so you can easily check all our parameters 
traceplotfunc <- function(parameter, title){

# Trace plot for parameter mu4
trdif2rstan <- plot(model_fit,     # Model fit 
                    plotfun = "trace",   # Specify that you want trace plots 
                    pars = c(parameter))     # Select parameter, 
# it is also possible to select more than one at the same time, such as c("mu2", "mu3","mu4")

# Adjust the plot (without warmup) with ggplot 
trdif2rstan2 <- trdif2rstan +   # Object returned from plot function 
                ylab(title) + 
                theme(axis.title.y=element_text(angle = 0, 
                                                vjust = 0.5, 
                                                hjust = -0.5, 
                                                margin = margin(0, 0.7, 0, 0, "cm"), size = 14), 
                      #legend.position = "none",  
                      axis.line.x  = element_line(size = 0.6), 
                      axis.line.y  = element_line(size = 0.6), 
                      plot.margin = unit(c(0.3,0.1,0.3,0.05),"cm"), 
                      plot.subtitle = element_text(size = 10.5), 
                      axis.ticks = element_line(size = 0.6), 
                      axis.text=element_text(size=8.5), 
                       axis.ticks.length=unit(.1, "cm")) + 
                labs(subtitle = "") + 
                #ylim(c(0.01, 0.07)) + 
                xlab("Iteration")

return(trdif2rstan2)
}
# Show trace plot for one digit parameter 
tr2 <- traceplotfunc(parameter = "mu2", title = expression(~mu[~beta]))
tr2 <- tr2 + theme(legend.position = "none")
tr3 <- traceplotfunc(parameter = "mu3", title = expression(~mu[~delta[~7]]))
tr3 <- tr3 + theme(legend.position = "none")
tr4 <- traceplotfunc(parameter = "mu4", title = expression(~mu[~delta[~6]]))
tr4 <- tr4 + theme(legend.position = "none")
tr5 <- traceplotfunc(parameter = "mu5", title = expression(~mu[~delta[~4]]))
tr5 <- tr5 + theme(legend.position = "none")
tr6 <- traceplotfunc(parameter = "mu6", title = expression(~mu[~delta[~3]]))
legendtr <- cowplot::get_legend(tr6)
tr6 <- tr6 + theme(legend.position = "none")

# Plot general effects together 
grid.arrange(tr2, tr3, tr4, tr5, tr6, legendtr, nrow = 2, ncol = 3)
```
\normalsize

```{r traceplotsd, echo = FALSE}
# Show trace plot for one digit parameter 
tr2 <- traceplotfunc(parameter = "g2", title = expression(~sigma[~beta]^2))
tr2 <- tr2 + theme(legend.position = "none")
tr3 <- traceplotfunc(parameter = "g3", title = expression(~sigma[~delta[~7]]^2))
tr3 <- tr3 + theme(legend.position = "none")
tr4 <- traceplotfunc(parameter = "g4", title = expression(~sigma[~delta[~6]]^2))
tr4 <- tr4 + theme(legend.position = "none")
tr5 <- traceplotfunc(parameter = "g5", title = expression(~sigma[~delta[~4]]^2))
tr5 <- tr5 + theme(legend.position = "none")
tr6 <- traceplotfunc(parameter = "g6", title = expression(~sigma[~delta[~3]]^2))
legendtr <- cowplot::get_legend(tr6)
tr6 <- tr6 + theme(legend.position = "none")

# Plot general effects together 
grid.arrange(tr2, tr3, tr4, tr5, tr6, legendtr, nrow = 2, ncol = 3)
```

Because there are so many parameters (each for every individual), it is hard to inspect them all individually. A more efficient way to check if the parameters have converged is by checking the $\hat{R}$ and the number of effective samples.

### Rhat & Number of Effective Samples

|    As for the trace plots, the standard *plot* function offers the possibility to plot the $\hat{R}$ (Rhat) and the ratio of the effective sample size to the total posterior sample size (this is different from what is plotted in the paper). 
&nbsp;

\scriptsize 
```{r rhatneffstandard, eval = FALSE}
# Plot frequency of rhat 
plot(model_fit, plotfun= "rhat")

# Plot frequency of the ratio 
plot(model_fit, plotfun = "ess")
```
\normalsize

|    However, you can also manually do this. To obtain the plots from the paper, the following code can be run. Instead of the ratio of the effective sample size to the total posterior sample size, we obtain the frequency of the numbers of effective samples (as shown in the paper). 
&nbsp;

\scriptsize 
```{r rhatneff, fig.cap="The frequency of the $\\hat{R}$ statistic and the number of effective samples. A: The frequency of the $\\hat{R}$. B: The frequency of the number of effective samples where the dashed lines represent the total number of iterations.", fig.align = "left", fig.height = 4, fig.width = 8, eval = FALSE, echo = FALSE}
# Obtain rhat for every parameter from model fit 
rhatrstan <- summary(model_fit)$summary[,"Rhat"]

# Obtain number of effective samples for every paramter from model fit 
neffrstan <- summary(model_fit)$summary[,"n_eff"]

# Save as dataframe so we can plot it with ggplot
rstanrhatneff <- data.frame(rhat = rhatrstan, neff = neffrstan)

# Figures 
# Rhat histogram 
rhatplotrstan <- ggplot(rstanrhatneff, aes(rhat)) + 
  geom_histogram(binwidth = 0.0001, alpha = 0.5, position = "identity", fill = 
                   "darkgreen") +   # Histogram, with binwidth the width of the frequency bars is set, 
  #with alpha the color is made more transparant, 
  #with position identity the bars are not set on top of each other in case of plotting multiple groups 
  xlim(0.998, 1.025) +         # Works for this example but can be left out 
  xlab(expression(hat(R))) +   # Set x axis title 
  ylab("Frequency") +          # Set y axis title 
  theme_classic() +            # Removes grid 
  labs(subtitle = "A") +       # Adds subtitle 
  theme(legend.position = "none")  # Removes legend

# Neff histogram 
neffplotrstan <- ggplot(rstanrhatneff, aes(neff)) + 
  geom_histogram(binwidth = 100, alpha = 0.5, position = "identity", fill = 
                   "darkgreen") + 
  xlab("Number of effective samples") + 
  ylab("") + 
  theme_classic() + 
  guides(fill=guide_legend(title="Package")) + 
  labs(subtitle = "B") + 
  geom_vline(xintercept = 12000, linetype = "dashed", color = "darkgreen", size = .3)  # Line representing otal number of iterations

# Plot together 
grid.arrange(rhatplotrstan, neffplotrstan, nrow = 1, ncol = 2)

```
\normalsize

\scriptsize 
```{r rhatneff2, fig.cap="The frequency of the $\\hat{R}$ statistic and the number of effective samples. A: The frequency of the $\\hat{R}$. B: The frequency of the number of effective samples where the dashed lines represent the total number of iterations.", fig.align = "left", fig.height = 4, fig.width = 8}
# Obtain rhat for every parameter from model fit 
rhatrstan <- summary(model_fit)$summary[,"Rhat"]
fixedrhatrstan <- rhatrstan[1:13]  # for general effects 
deltameanrrstan <- mean(fixedrhatrstan[3:6])  # calculate the mean for delta parameters

# Obtain number of effective samples for every parameter from model fit 
neffrstan <- summary(model_fit)$summary[,"n_eff"]
fixedneffrstan <- neffrstan[1:13]  # for general effects 
fixedneffrstanmean <- mean(fixedneffrstan[3:6])  # calculate the mean for delta parameters

# Save as dataframe so we can plot it with ggplot
rstanrhatneff <- data.frame(rhat = rhatrstan, neff = neffrstan)

# Combine rhat into one data.frame 
rhatrstan_trans <- t(rhatrstan)
rhatrstan_trans2 <- as.data.frame(rhatrstan_trans)

# remove lp and fixed effects 
# rhat 
rhatrstan2 <- rhatrstan[-c(1:13, 326)]

rhatrstan4 <- t(rhatrstan2)
rhatrstan5 <- as.data.frame(rhatrstan4)
rhatrstan6 <- unlist(rhatrstan5)
rstanrhatneff_v2 <- data.frame(package = rep("Rstan", 312), 
                                 parameter = rep(c("Gamma", "Beta", "Delta", "Delta", 
                                                   "Delta", "Delta", "Gamma", "Beta", 
                                                   "Delta", "Delta", "Delta", "Delta"), 
                                                 each = 52), 
                                 rhat = rhatrstan6)

# Create violin plot 
  figrhatnew <- ggplot(rstanrhatneff_v2, aes(x = factor(parameter, level = c("Gamma", "Beta", "Delta")), y = rhat, colour = package)) + 
    geom_violin(width = 1) + 
    geom_quasirandom(alpha = 0.1, width = 0.2, dodge.width=1) + 
    geom_point(aes(x=1, y= fixedrhatrstan[1]), shape = 8, colour="darkgreen", size = 2) + 
    geom_point(aes(x=2, y= fixedrhatrstan[2]), shape = 8, colour="darkgreen", size = 2) + 
    geom_point(aes(x=3, y= deltameanrrstan), shape = 8, colour="darkgreen", size = 2) + 
    xlab("Parameter") + ylab(expression(hat(R))) +
    scale_colour_manual(values = "darkgreen") + 
    theme_classic() + 
    theme(legend.position = "none") + 
    scale_x_discrete(labels= c( 
      expression(~gamma), 
      expression(~beta), 
      expression(~delta))) + 
    labs(subtitle = "C") + 
    theme(axis.title.y = element_text(angle = 0, 
                                      vjust = 0.5, 
                                      hjust = -0.5, 
                                      margin = margin(0, 1.1, 0, 0, "cm")))

  # neff
  # Combine neff into one data.frame 
  neffrstan2 <- data.frame(neff = neffrstan)
  neffrstan3 <- neffrstan2$neff[-c(1:13, 326)]

  neffrstanbrms2 <- data.frame(package = rep("Rstan", 312), 
                               parameter = rep(c("Gamma", "Beta", "Delta", "Delta", 
                                                 "Delta", "Delta", "Gamma", "Beta", 
                                                 "Delta", "Delta", "Delta", "Delta"), each = 52),
                               neff = neffrstan3)

  figneffnew <- ggplot(neffrstanbrms2, aes(x = factor(parameter, level = c("Gamma", "Beta", "Delta")), y = neff, colour = package)) + 
    geom_violin(width = 1) + 
    geom_quasirandom(alpha = 0.1, width = 0.2, dodge.width=1) + 
    geom_point(aes(x = 1, y = fixedneffrstan[1]), shape = 8, colour="darkgreen", size = 2) + 
    geom_point(aes(x = 2, y = fixedneffrstan[2]), shape = 8, colour="darkgreen", size = 2) + 
    geom_point(aes(x = 3, y = fixedneffrstanmean), shape = 8, colour="darkgreen", size = 2) + 
    xlab("Parameter") + ylab("Number of effective samples") +
    scale_colour_manual(values = "darkgreen") + 
    theme_classic() + 
    theme(legend.position = "none") + 
    scale_x_discrete(labels= c( 
      expression(~gamma), 
      expression(~beta), 
      expression(~delta))) + 
    labs(subtitle = "D") + 
    geom_hline(yintercept = 12000, linetype = "dashed", color = "darkgreen", size = .3)

grid.arrange(figrhatnew, figneffnew, nrow = 1, ncol = 2)

```
\normalsize

### General effects

|    The general effects are represented by $\mu_{\delta_{7}}$, $\mu_{\delta_{6}}$, $\mu_{\delta_{4}}$, $\mu_{\delta_{3}}$, and $\mu_{\beta}$. To investigate whether there is a digit and side effect, we will inspect the estimates for these parameters.

#### Posterior Distribution

|    With the *plot* function, the posterior distribution per parameter can be displayed. 
&nbsp;

\scriptsize 
```{r posteriorstandard, eval = FALSE}
# Seperate plot per parameter of posterior density 
plot(model_fit, 
     plotfun = "dens")  
# without specifying pars, the posterior distributions of the first ten parameters are displayed 
```
\normalsize

|    Another possibility is to use the *bayesplot* package to plot all the posterior densities together in one plot. This makes it easier to compare the posterior distributions of each parameter. 
&nbsp;

\scriptsize 
```{r posteriormanual, fig.cap="The posterior distributions for the general effects. The middle line within the distributions represents the mean posterior. The shaded area within the distributions represents 95\\% of the probability mass.", fig.align = "left", fig.height = 4, fig.width = 5}
postc <- as.array(model_fit)  # change structure of model fit

# With dimnames(postc) we could check the names of parameters 
# The ones useful for us are "mu2", "mu3", "mu4", "mu5", and "mu6"

color_scheme_set("green")  # plot the distributions in green

pairsc <- mcmc_areas(
  postc,                                        # Structured model fit 
  pars = c("mu2", "mu3", "mu4", "mu5", "mu6"),  # Select parameters 
  prob = 0.95,                                   # Plot 80% credible intervals   
  prob_outer = 0.99,                            # 99% outer interval            
  point_est = "mean"                            # Plot line in middle    distribution representing the mean 
)
# Returns ggplot object

# Adjust ggplot object further 
pairsc2 <- pairsc +                             # ggplot object 
            ggtitle("Posterior Distribution") +                  # Add title 
            scale_x_continuous(breaks = c(-0.025, 0, 0.025, 0.05, 0.075)) +   # specify the breaks and labels on x-axis, this is analysis dependent 
            theme_apa() +                       # Select theme 
            scale_y_discrete(labels= c(         # Customize labels to y axis 
                                       expression(~mu[~beta]), 
                                       expression(~mu[~delta[~7]]), 
                                       expression(~mu[~delta[~6]]), 
                                       expression(~mu[~delta[~4]]), 
                                       expression(~mu[~delta[~3]])))

pairsc2
```
\normalsize

|    We can present more information about the distribution by providing a table with the estimated mean of the posterior distribution, the standard error of this mean, the lower and upper bound of the 95% credible interval, the $\hat{R}$ and the number of effective samples. 
&nbsp;

\scriptsize 
```{r geneftable, results = "asis", cache = TRUE}
# Obtain the mean, standard error, 95% credible interval, rhat and number of effective samples for the intercept, and the digit and side parameters 
fitrstanhiertable <- summary(model_fit,  # model fit 
                             # Select parameters 
                             pars = c("mu", "mu2", "mu3", "mu4", "mu5", 
                                      "mu6"))$summary[,c("mean", "se_mean", 
                                                         "2.5%", "97.5%", 
                                                         "n_eff", "Rhat")]

# The results are transformed to a dataframe that allows us to plot the results nicely in a table 
fitrstanhiertable2 <- as.data.frame(fitrstanhiertable)

# We manually set the row names of the table using latex for mathematical symbols 
rownamesfitrstanhiertable <-  c("$\\mu_{\\gamma}$","$\\mu_{\\beta}$", "$\\mu_{\\delta_{7}}$", "$\\mu_{\\delta_{6}}$", "$\\mu_{\\delta_{4}}$", "$\\mu_{\\delta_{3}}$")

# We manually set the column names of the table 
colnames(fitrstanhiertable2) <- c("Mean", "SE", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")

# The column with row names is added to the table 
fitrstanhiertable3 <- add_column(fitrstanhiertable2, 
                                 Parameters = rownamesfitrstanhiertable, 
                                 .before = "Mean")

# Using the apa_table function from the papaja package the results are presented in an apa table 
apa_table(fitrstanhiertable3,   # Data frame with results 
          row.names = FALSE,    # Default row names are deleted (we set them ourself)
          caption = "Posterior Mean, Standard Error (SE) of the Mean, Lower and 
          Upper Bound of the 95\\% Credible Interval, the Number of Effective 
          Samples, and the $\\hat{R}$ of the General Effect Parameters as 
          estimated by Rstan.", 
          align = c("l", "c", "c", "c", "c", "c", "c"),   # How should results be presented, in the midle of the table or outlined left 
          digits = 3,       # The number of digits behind the ".". 
          placement = "h",  # Place after code 
          escape = FALSE)   # This indicates that the results contain latex code that should be read as latex code. It is important to do this in combination with putting the results to "asis" in the {R} section above.

```
\normalsize

### Individual Effects

|    Next, we look at the individual deviations from the general effects. Therefore, we inspect the variance and the individual estimates of the digit and side effects.

#### Variance Estimates

|    First, we can show a table with the posterior distribution estimates of the variance parameters. This indicates the variation of the general effect. We called our variances *g* (*g* until *g6*). 
&nbsp;

\scriptsize 
```{r inddevtable, results = "asis", cache = TRUE}
fitrstanhiertable2 <- summary(model_fit,   # model fit 
                              # Select variance parameters (we called them g)
                              pars = c("g", "g2", "g3", "g4", "g5", 
                                       "g6"))$summary[,c("mean", "se_mean", 
                                                         "2.5%", "97.5%", 
                                                         "n_eff", "Rhat")]

# The results are transformed to a dataframe that allows us to plot the results nicely in a table 
fitrstanhiertable22 <- as.data.frame(fitrstanhiertable2)

# We manually set the row names of the table using latex for mathematical symbols 
rownamesfitrstanhiertable2 <-  c("$\\sigma_{\\gamma}^2$","$\\sigma_{\\beta}^2$", "$\\sigma_{\\delta_{7}}^2$", "$\\sigma_{\\delta_{6}}^2$", "$\\sigma_{\\delta_{4}}^2$", "$\\sigma_{\\delta_{3}}^2$")

# The column with row names is added to the table 
colnames(fitrstanhiertable22) <- c("Mean", "SE", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")

# The column with row names is added to the table 
fitrstanhiertable32 <- add_column(fitrstanhiertable22, 
                                  Parameters = rownamesfitrstanhiertable2, 
                                  .before = "Mean")

# Using the apa_table function from the papaja package the results are presented in an apa table 
apa_table(fitrstanhiertable32,  # Data frame with results
          row.names = FALSE,    # Default row names are deleted (we set them ourself)
          caption = "Posterior Variance, Standard Error (SE) of the Variance, 
          Lower and Upper Bound of the 95\\% Credible Interval, the Number of 
          Effective Samples, and the $\\hat{R}$ of the Variance Parameters           as estimated by Rstan.", 
          align = c("l", "c", "c", "c", "c", "c", "c"),  # How should results be presented, in the midle of the table or outlined left 
          digits = 3,        # The number of digits behind the ".". 
          place = "h",       # Place after code 
          escape = FALSE)    # This indicates that the results contain latex code that should be read as latex code. It is important to do this in combination with putting the results to "asis" in the {R} section above.

```
\normalsize

#### Individual Estimates

|    Next, we can look at the individual estimates for the digit and side effect parameters in two different ways. First, we plot the individual estimates and display their 95% credible interval. The credible interval is displayed in pink when it contains zero, and displayed in blue when it does not contain zero. In addition, the dashed line represent the general effect estimate. 
|    The code below shows you how to obtain this code for one parameter. The other parameters follow logically from this example and would result in the figure below. 
&nbsp;

\scriptsize 
```{r randefplotexample}
# Gamma 
## Create object with parameter names 
gamnamesstan <- paste("gammai", "[",1:52,"]", sep="")

# Extract the parameter estimates using the names from before 
gamstan <- as.data.frame(summary(model_fit, 
                                 pars = c( gamnamesstan))$summary)

# Extract the general effect estimate of the parameter 
gamstanmu <- as.data.frame(summary(model_fit, 
                                   pars = c("mu"))$summary)

# Add the parameter names to the data.frame 
gamstan$parameters <- c(gamnamesstan)

# Add parameter to dataframe. When the credible interval contains zero, the parameter is set pink, otherwise to blue 
for (i in 1:nrow(gamstan)){  # 1 if CI contains zero, 0 otherwise 
  if (gamstan$`2.5%`[i]*gamstan$`97.5%`[i] < 0) {
    gamstan$zero2[i] <- "pink"
  } else (gamstan$zero2[i] <- "blue") 
}

# Order the estimates ascending 
gamstan2 <- gamstan[order(gamstan$mean),]
gamstan2$order <- c(1:52)  # Add parameter to data frame with the order

# Plot the individual estimates for gamma (intercept)
gammarandomplot <- ggplot(gamstan2, aes(x = order, y = mean, ymin = `2.5%`, ymax = `97.5%`)) + 
  geom_linerange(aes(color = "blue"), color = "#00BFC4") + 
  geom_point() + 
  geom_hline(yintercept = gamstanmu$mean, linetype = "dashed", alpha = .3) + 
  labs(title = expression(~gamma[~i]), y = "Estimation") + 
  theme(legend.position = "none", 
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```
\normalsize

```{r indvestirstan, echo = FALSE, fig.cap="The posterior means with 95\\% credible interval in pink and red as estimated by \\textit{rstan} for every individual shown in increasing order. The dashed line represents the general posterior mean. A pink interval means that the interval contains zero, a blue interval represents an interval that does not contain zero.", fig.align = "left", fig.height = 6, fig.width = 8}
# Documentation on how to retrieve objects from stan output: https://cran.r-project.org/web/packages/rstan/vignettes/stanfit-objects.html

# Gamma 
## Create object with parameter names 
gamnamesstan <- paste("gammai", "[",1:52,"]", sep="")

# Extract the parameter estimates using the names from before 
gamstan <- as.data.frame(summary(model_fit, 
                                 pars = c( gamnamesstan))$summary)

# Extract the general effect estimate of the parameter 
gamstanmu <- as.data.frame(summary(model_fit, 
                                   pars = c("mu"))$summary)

# Add the parameter names to the data.frame 
gamstan$parameters <- c(gamnamesstan)

# Add parameter to dataframe. When the credible interval contains zero, the parameter is set pink, otherwise to blue 
for (i in 1:nrow(gamstan)){  # 1 if CI contains zero, 0 otherwise 
  if (gamstan$`2.5%`[i]*gamstan$`97.5%`[i] < 0) {
    gamstan$zero2[i] <- "pink"
  } else (gamstan$zero2[i] <- "blue") 
}

# Order the estimates ascending 
gamstan2 <- gamstan[order(gamstan$mean),]
gamstan2$order <- c(1:52)  # Add parameter to data frame with the order

# Plot the individual estimates for gamma (intercept)
gammarandomplot <- ggplot(gamstan2, aes(x = order, y = mean, ymin = `2.5%`, ymax = `97.5%`)) + 
  geom_linerange(aes(color = "blue"), color = "#00BFC4") + 
  geom_point() + 
  geom_hline(yintercept = gamstanmu$mean, linetype = "dashed", alpha = .3) + 
  labs(title = expression(~gamma[~i]), y = "Estimation") + 
  theme(legend.position = "none", 
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

# Beta  
betnamesstan <- paste("betai", "[",1:52,"]", sep="")
betstan <- as.data.frame(summary(model_fit, pars = c( betnamesstan))$summary)
betstanmu <- as.data.frame(summary(model_fit, pars = c("mu2"))$summary)
betstan$parameters <- c(betnamesstan)

for (i in 1:nrow(betstan)){  # 1 if CI contains zero, 0 otherwise 
  if (betstan$`2.5%`[i]*betstan$`97.5%`[i] < 0) {
    betstan$zero2[i] <- "blue"
  } else (betstan$zero2[i] <- "pink") 
}

betstan2 <- betstan[order(betstan$mean),]
betstan2$order <- c(1:52)

betarandomplot <- ggplot(betstan2, aes(x = order, y = mean, ymin = `2.5%`, ymax = `97.5%`)) + 
  geom_linerange(aes(color = zero2)) + 
  geom_point() + 
  geom_hline(yintercept = betstanmu$mean, linetype = "dashed", alpha = .3) + 
  labs(title = expression(~beta[~"side,i"]), y = "") + 
  geom_hline(yintercept = 0, color = "red", alpha = .3) +  
  theme(legend.position = "none", 
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

# Delta1 
del1namesstan <- paste("delta1", "[",1:52,"]", sep="")
delt1stan <- as.data.frame(summary(model_fit, pars = c( del1namesstan))$summary)
del1stanmu <- as.data.frame(summary(model_fit, pars = c("mu3"))$summary)
delt1stan$parameters <- c(del1namesstan)

for (i in 1:nrow(delt1stan)){  # 1 if CI contains zero, 0 otherwise 
  if (delt1stan$`2.5%`[i]*delt1stan$`97.5%`[i] < 0) {
    delt1stan$zero2[i] <- "blue"
  } else (delt1stan$zero2[i] <- "pink") 
}

delt1stan2 <- delt1stan[order(delt1stan$mean),]
delt1stan2$order <- c(1:52)

delta1randomplot <- ggplot(delt1stan2, aes(x = order, y = mean, ymin = `2.5%`, ymax = `97.5%`)) + 
  geom_linerange(aes(color = zero2)) + 
  geom_point() + 
  geom_hline(yintercept = del1stanmu$mean, linetype = "dashed", alpha = .3) + 
  labs(title = expression(~delta[~"7-8,i"]), y = "") + 
  geom_hline(yintercept = 0, color = "red", alpha = .3) +  
  theme(legend.position = "none", 
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

# Delta 2 
del2namesstan <- paste("delta2", "[",1:52,"]", sep="")
delt2stan <- as.data.frame(summary(model_fit, pars = c( del2namesstan))$summary)
del2stanmu <- as.data.frame(summary(model_fit, pars = c("mu4"))$summary)
delt2stan$parameters <- c(del2namesstan)

for (i in 1:nrow(delt2stan)){  # 1 if CI contains zero, 0 otherwise 
  if (delt2stan$`2.5%`[i]*delt2stan$`97.5%`[i] < 0) {
    delt2stan$zero2[i] <- "blue"
  } else (delt2stan$zero2[i] <- "pink") 
}

delt2stan2 <- delt2stan[order(delt2stan$mean),]
delt2stan2$order <- c(1:52)

delta2randomplot <- ggplot(delt2stan2, aes(x = order, y = mean, ymin = `2.5%`, ymax = `97.5%`)) + 
  geom_linerange(aes(color = zero2)) + 
  geom_point() + 
  geom_hline(yintercept = del2stanmu$mean, linetype = "dashed", alpha = .3) + 
  labs(title = expression(~delta[~"6,i"]), y = "Estimation") + 
  geom_hline(yintercept = 0, color = "red", alpha = .3) +  
  theme(legend.position = "none", 
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

# Delta 3 
del3namesstan <- paste("delta3", "[",1:52,"]", sep="")
delt3stan <- as.data.frame(summary(model_fit, pars = c( del3namesstan))$summary)
del3stanmu <- as.data.frame(summary(model_fit, pars = c("mu5"))$summary)
delt3stan$parameters <- c(del3namesstan)

for (i in 1:nrow(delt3stan)){  # 1 if CI contains zero, 0 otherwise 
  if (delt3stan$`2.5%`[i]*delt3stan$`97.5%`[i] < 0) {
    delt3stan$zero2[i] <- "blue"
  } else (delt3stan$zero2[i] <- "pink") 
}

delt3stan2 <- delt3stan[order(delt3stan$mean),]
delt3stan2$order <- c(1:52)

delta3randomplot <- ggplot(delt3stan2, aes(x = order, y = mean, ymin = `2.5%`, ymax = `97.5%`)) + 
  geom_linerange(aes(color = zero2)) + 
  geom_point() + 
  geom_hline(yintercept = del3stanmu$mean, linetype = "dashed", alpha = .3) + 
  labs(title = expression(~delta[~"4-3,i"]), y = "") + 
  geom_hline(yintercept = 0, color = "red", alpha = .3) +  
  theme(legend.position = "none", 
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

# Delta 4 
del4namesstan <- paste("delta4", "[",1:52,"]", sep="")
delt4stan <- as.data.frame(summary(model_fit, pars = c( del4namesstan))$summary)
del4stanmu <- as.data.frame(summary(model_fit, pars = c("mu6"))$summary)
delt4stan$parameters <- c(del4namesstan)

for (i in 1:nrow(delt4stan)){  # 1 if CI contains zero, 0 otherwise 
  if (delt4stan$`2.5%`[i]*delt4stan$`97.5%`[i] < 0) {
    delt4stan$zero2[i] <- "blue"
  } else (delt4stan$zero2[i] <- "pink") 
}

delt4stan2 <- delt4stan[order(delt4stan$mean),]
delt4stan2$order <- c(1:52)

delta4randomplot <- ggplot(delt4stan2, aes(x = order, y = mean, ymin = `2.5%`, ymax = `97.5%`)) + 
  geom_linerange(aes(color = zero2)) + 
  geom_point() + 
  geom_hline(yintercept = del4stanmu$mean, linetype = "dashed", alpha = .3) + 
  labs(title = expression(~delta[~"3-2,i"]), y = "") + 
  geom_hline(yintercept = 0, color = "red", alpha = .3) +  
  theme(legend.position = "none", 
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),)

# Plot together 
grid.arrange(gammarandomplot, betarandomplot, delta1randomplot, delta2randomplot, delta3randomplot, delta4randomplot, nrow = 2, ncol = 3)
```

|    The second way is the plot the individual estimates and display the variation of these individual estimates. For this figure, we used the design and code from @j_van_langen_2020_3715576. 
&nbsp;

\scriptsize 
```{r modelestimatesrstanfigure3, fig.cap="Model estimates for digit effect parameters. The points represent the mean parameter estimate for each individual. The violin plot on each right side shows the variance of the individual parameter estimates.", fig.align = "left", fig.height = 4, fig.width = 5}
## Delta's 
## Create dataframe 
parameter_m7 <- rep(c("Delta 1", "Delta 2", "Delta 3", "Delta 4"), each = 52 )

### Obtain all individual estimates delta 
hier_modeld <- model_fit  # model fit

# Obtain individual estimates 
delta1_summary <- summary(hier_modeld, pars = c("delta1"), probs = c(0.1, 0.9))$summary

delta2_summary <- summary(hier_modeld, pars = c("delta2"), probs = c(0.1, 0.9))$summary

delta3_summary <- summary(hier_modeld, pars = c("delta3"), probs = c(0.1, 0.9))$summary

delta4_summary <- summary(hier_modeld, pars = c("delta4"), probs = c(0.1, 0.9))$summary

parest_m7 <- rbind(delta1_summary, delta2_summary, delta3_summary, delta4_summary)
datafr_m7 <- data.frame(parameter_m7, parest_m7)

## Plot 
set.seed(321)
datafr_m7$x <- rep(c(1, 2, 3, 4), each = 52)
datafr_m7$xj <- jitter(datafr_m7$x, amount = 0.09)
datafr_m7$id <- rep(c(1:52), 4)

plotmodelest_m7 <- ggplot(data=datafr_m7, aes(y=mean)) +

  #Add geom_() objects
  geom_point(data = datafr_m7 %>% filter(x=="1"), aes(x=xj), color = 'dodgerblue', size = 1.5,
             alpha = .6) +
  geom_point(data = datafr_m7 %>% filter(x=="2"), aes(x=xj), color = 'darkgreen', size = 1.5,
             alpha = .6) +
  geom_point(data = datafr_m7 %>% filter(x=="3"), aes(x=xj), color = 'darkorange', size = 1.5, 
             alpha = .6) +
  geom_point(data = datafr_m7 %>% filter(x=="4"), aes(x=xj), color = 'gold2', size = 1.5, 
             alpha = .6) +

  geom_line(aes(x=xj, group=id), color = 'lightgray', alpha = .3) +

  geom_line(data=data.frame(x=-1:4.5,y=0), aes(x = x, y = y), linetype = "dashed", size = .3) +

  geom_half_violin(
    data = datafr_m7 %>% filter(x=="1"),aes(x = x, y = mean), position = position_nudge(x = 3.2), 
    side = "r", fill = 'dodgerblue', alpha = .5, color = "dodgerblue", trim = TRUE) +

  geom_half_violin(
    data = datafr_m7 %>% filter(x=="2"),aes(x = x, y = mean), position = position_nudge(x = 2.2), 
    side = "r", fill = "darkgreen", alpha = .5, color = "darkgreen", trim = TRUE) +

  geom_half_violin(
    data = datafr_m7 %>% filter(x=="3"),aes(x = x, y = mean), position = position_nudge(x = 1.2), 
    side = "r", fill = "darkorange", alpha = .5, color = "darkorange", trim = TRUE) +

  geom_half_violin(
    data = datafr_m7 %>% filter(x=="4"),aes(x = x, y = mean), position = position_nudge(x = 0.2), 
    side = "r", fill = "gold2", alpha = .5, color = "gold2", trim = TRUE) +

  #Define additional settings
  scale_x_continuous(breaks=c(1,2,3,4), labels=c(expression(~delta[~7]), expression(~delta[~6]), expression(~delta[~4]), expression(~delta[~3]))) +
  xlab("") + ylab("Value") +
  ggtitle('') +
  theme_classic() + 
  coord_cartesian(xlim = c(0.75, 5), ylim = c(-0.05, 0.12))

plotmodelest_m7
```
\normalsize

## Model comparison

|    In case we have different models, we can compare them to check under which model the data is most likely. This can be done with the Bayes factor. In *rstan* the Bayes factor can be obtained by using the *bridgesampling* package. To be able to use this package, it is important that the *.stan* file is written in a way that it saves the constants (i.e., `target += normal_lpdf(mu4 | a3, sqrt(b3));`).   
|    First, we fit all the models (except for the full model that we already fitted. 
&nbsp;

\scriptsize 
```{r modelcomparison, eval = FALSE, echo = TRUE}
# Fit the other models 
## Null model 
BFnullmodel <- stan(file = "myPath/rewritestanmodel3h0.stan", 
                    data = myData_list, 
                    iter = 4000, 
                    chains = 4,  
                    warmup = 1000, 
                    cores = 4)

## Side model 
BFsidemodel <- stan(file = "myPath/rewritestanmodel3hside.stan", 
                       data = myData_list, 
                       iter = 4000, 
                       chains = 4, 
                       warmup = 1000, 
                       cores = 4)

## Digit model 
BFdigitmodel <- stan(file = "myPath/rewritestanmodel3hdigit.stan", 
                         data = myData_list, 
                         iter = 4000, 
                         chains = 4, 
                         warmup = 1000, 
                         cores = 4)
```
\normalsize

|    Then, we use the `bridge_sampler` function to obtain the marginal likelihoods. In case you loaded the full model from the provided R object, you will need an extra step as described in [this issue](https://github.com/quentingronau/bridgesampling/issues/7). 
&nbsp;

\scriptsize 
```{r bfpart2, eval = FALSE, echo = TRUE}
# Bridge sampling 
# library(bridgesampling)
bridge_Hnull <- bridge_sampler(BFnullmodel)
bridge_Hside <- bridge_sampler(BFsidemodel)
bridge_Hdigit <- bridge_sampler(BFdigitmodel)
bridge_Hfull <- bridge_sampler(model_fit)  # in case you just fitted the model yourself

## In case you loaded the full model from the R object, you will need some preparation (https://github.com/quentingronau/bridgesampling/issues/7)

### Fit model again but do not run any chains 
new_mod <- stan(file = "myPath/rewritestanmodel3_julia.stan", 
                data = myData_list, 
                chains = 0)

### Now do the bridgesampling 
bridge_Hfull <- bridge_sampler(model_fit, new_mod)
```
\normalsize

|    Finally, we can obtain the Bayes factor. You obtain the evidence in favor of the first model that you specify in the `bf` function. 
&nbsp;

\scriptsize 
```{r bfpart3, eval = FALSE, echo = TRUE}

# Obtain BF (in this way evidence in favor of the null hypothesis)
bridgesampling::bf(bridge_Hnull, bridge_Hfull)   # Null against full

# Obtain BF (in this way evidence in favor of the full hypothesis)
bridgesampling::bf(bridge_Hfull, bridge_Hnull)   # Full against null
bridgesampling::bf(bridge_Hfull, bridge_Hside)   # Full against side 
bridgesampling::bf(bridge_Hfull, bridge_Hdigit)  # Full against digit 
```
\normalsize

## Additional Resources

|    To learn more about fitting a Bayesian Hierarchical Model, we recommend the following resources:

- [Stan forum](https://discourse.mc-stan.org/c/interfaces/rstan) 
- [Accessing the contents of a stanfit object](https://cran.r-project.org/web/packages/rstan/vignettes/stanfit-objects.html)

\newpage

# References 
\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
&nbsp;

### 2022-06-09 D - Tutorial Normal Model brms/Online Supplement D - Bayesian Hierarchical Modeling in brms.Rmd
All documents except .rds

---
title             : "Online Supplement D: Bayesian Hierarchical Modeling in brms"
author            : "Myrthe Veenman"
bibliography      : ["r-references.bib"]

header-includes   :
   - \usepackage{bm}
   - \usepackage{amsmath}
   - \usepackage{setspace}
   - \usepackage{xcolor}
   - \definecolor{mypink}{RGB}{255, 230, 255}
   - \usepackage{todonotes}
   - \usepackage{graphicx}
   - \usepackage{float}

csl               : apa6.csl
output            : pdf_document
---
\spacing{1.5}
\fontsize{12}{12}

```{r setup, include=FALSE}
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

knitr::opts_chunk$set(tidy.opts=list(width.cutoff=40), tidy=TRUE, echo = TRUE, message = FALSE, warning = FALSE, fig.pos = "H")
```

|    This online supplement provides a tutorial on Bayesian Hierarchical models (BHM) in `brms` [@R-brms_a; @R-brms_b]. In addition, it contains information about the Number Classification Task that could not be included in the paper. As `brms` runs on `Stan` [@carpenter2017stan], it is very similar to `rstan` [@R-rstan]. However, there are some differences that will be discussed in this tutorial. Only the fit of the full model will be shown as the fitting the others should follow logically from this tutorial. Technical details, such as explanations of statistical terms (e.g., $\hat{R}$), will not be discussed here but can be found in the paper. In concrete, the tutorial will cover the following:

1. Required packages 
2. Input  
  2.1 Formula  
  2.2 Family  
  2.3 Data  
  2.4 Priors 
3. Fit model 
4. Output   
  4.1 General Effects   
  4.2 Individual Effects 
5. Model comparison 
6. Additional resources

## Required Packages

In this practical, the following packages will be used:

- For the document layout: *rmarkdown* [Version 1.16; @R-rmarkdown], *papaja* [Version 0.1.0.9842; @R-papaja], *knitr* [Version 1.25; @R-knitr], *kableExtra* [Version 1.1.0; @R-kableExtra]
- For data structuring: *LaplacesDemon* [Version 16.1.4; @R-LaplacesDemon], *plyr* [Version 1.8.4; @R-plyr], *dplyr* [Version 0.8.3; @R-dplyr], *readr* [Version 1.3.1; @R-readr], *truncnorm* [Version 1.0.8; @truncnorm]
- To fit the model and for model comparison: *brms* [Version 2.13.0; @R-brms_a; @R-brms_b]
- For the visualization of the results: *lattice* [Version 0.20.38; @R-lattice], *ggplot2* [Version 3.3.2; @R-ggplot2], *Rmisc* [Version 1.5; @R-Rmisc], *devtools* [Version 2.2.1; @R-devtools], *gghalves* [Version 0.1.0; @R-gghalves], *bayesplot* [Version 1.7.1; @R-bayesplot], *gridExtra* [Version 2.3; @R-gridExtra], *ggridges* [Version 0.5.3; @ggridges], *ggbeeswarm* [Version 0.6.0; @ggbeeswarm], *tibble* [Version 3.0.2; @R-tibble]

&nbsp;

\scriptsize 
```{R loadpackages, message = FALSE, warning = FALSE}
# Load the packages
library("papaja")
library("LaplacesDemon")
library("rstan")
library("brms")
library("plyr")
library("lattice")
library("ggplot2")
library("dplyr")
library("readr")
library("rmarkdown")
library("Rmisc")
library("devtools")
library("gghalves") 
library("bayesplot")
library("bridgesampling") 
library("gridExtra") 
library("tibble") 
library("kableExtra") 
library("bayestestR")
library("ggridges")
library("truncnorm")
library("ggbeeswarm")
```
\normalsize

## Input

|    To fit a BHM in *brms*, we use the *brm* function, as shown below. 
&nbsp;

\scriptsize 
```{r modelfitpreview, eval=FALSE, echo=TRUE}
model_fit <- brm(formula = formula,                      # Model formula 
                 data = myData_dataframe,                # Dataset (parameters, observations) 
                 family = gaussian(),                    # Family, distribution dependent variable 
                 prior = priorsmodel,                    # Specification of the priors 
                 warmup = 1000,                          # Nr. iterations for warmup (per chain)
                 iter = 4000,                            # Nr. iterations (per chain)
                 chains = 4,                             # Nr. chains 
                 core = 4)                               # Nr. of cores for parallel estimation

```
\normalsize

We can see from the code above that the function requires the following input:

- Formula 
- Family 
- Data 
- Priors

The other options do not require coding and are discussed in the paper.

### Formula

|    The formula is an object that specifies the dependent variable as a function of the independent variables. The `~` (tilde) separates the dependent variable (`rt`) on the left side from the independent variables on the right side. The formula distinguishes between general effects and individual deviations. The first part of the equation on the predictor side of the formula shown below (i.e., `1 + side + dif1 + dif2 + dif3 + dif4`) represents the general effects. The `1` represents the intercept, in the case of the symbolic distance effect $\mu_{\gamma}$, `side`, and `dif1` to `dif4` represent the general effects of side and distance ($\mu_{\beta}$ and $\mu_{\delta}$). Within the round brackets individual deviations and grouping variables are specified. In this case the grouping variable (`ind`) indicates that each participant has an individual effect. In other cases effects might vary per item, block, or even the combination of participant and item. The double bar (`||`) indicates that the correlation between the parameters should not be modeled. When using a single vertical bar (`|`), correlations across all individual effects (intercept, side and distance) are modeled. This approach, however, is a bit more difficult to interpret and would also not correspond to the implemented `Stan` model. 
&nbsp;

\scriptsize 
```{r formulafit}
formula <- rt2 ~ 1 + side + dif1 + dif2 + dif3 + dif4 +   # General effects 
          (1 + side + dif1 + dif2 + dif3 + dif4 || ind)   # Individual effects

```
\normalsize

However, for the digit classification task, we specified the formula in a slightly different way that allowed us to use a different prior on the intercept compared to the distance and side effects. In the first part, we specify that the dependent variable (in our case response times) are predicted by `a` and `b`. In the second part, we define `a`. It contains the general and individual distance and side effects. Also note that we replaced `1` by `0`, indicating that we do not want the default intercept. In the third and final part, we specify `b`, representing the general and individual intercepts by the variable `inter`.

\scriptsize
```{r formulafit2, eval = FALSE}
formula <- brms::bf(rt2 ~ a + b, nl = TRUE,  # part 1 
                  lf(a ~ 0 + side + dif1 + dif2 + dif3 + dif4 + (0 + side + dif1 + dif2 + dif3 + dif4 || ind), center = TRUE),   # part 2 
                  lf(b ~ 0 + inter + (0 + inter || ind), cmc = FALSE)  # part 3 
)
```
\normalsize

### Family

|    The argument `family` in the `brm` function specifies the exponential family of distributions according to which the dependent variable is assumed to be distributed. For the normal model, this argument takes the value `gaussian`.

### Data

|    The `brm` function requires a data frame (here called `myData_dataframe`) that contains all variables mentioned in the formula. This means that every parameter in the formula should have an element in the data frame that we put in the function. In our case, this should result in seven elements in the data frame (i.e., `rt2`, `side`, `dif1`, `dif2`, `dif3`, `dif4`, and `ind`).  
|  First, we will load the data and clean it according to the requirements in the paper.  
&nbsp;

\scriptsize 
```{r readdata}
indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
colnames(indat)=c('sub','block','trial','stim','resp','rt','error')

## Cleaning the data according to criteria discussed in paper 
# (code retrieved from Julia Haaf:
# https://github.com/PerceptionAndCognitionLab/bf-order/blob/public/papers/submission/R-scripts/ld5.R)
clean=function()
{
  indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
  colnames(indat)=c('sub','block','trial','stim','resp','rt','error')

  bad1=indat$sub%in%c(34,43)
  bad2=indat$rt<250 | indat$rt>2000
  bad3=indat$err==1
  bad4=indat$block==0 & indat$trial<20
  bad5=indat$trial==0

  bad=bad1 | bad2 | bad3 |bad4 |bad5
  dat=indat[!bad,]
  return(dat)
}

indat1 <- clean()  
myData_dataframe <- indat1  # Final dataset 
```
\normalsize

|    Now, we will expand the data.frame we provide the *brm* function. We specify our dependent variable. In our case, these are the RTs. The RTs are given in milliseconds, but we would like them in seconds. Therefore, we transform the original RT (i.e., divide them by 1000) and save this in our new dependent variable `rt2`. 
&nbsp;

\scriptsize 
```{r dependentvariable}
### Response time from milliseconds to seconds 
myData_dataframe$rt2 <- indat1$rt/1000
```
\normalsize

|    The next step is to specify the intercept (`inter`) and indicators (`side`, `dif1`, `dif2`, `dif3`, and `dif4`). These are the variables that specify per observation of a variable applies or not. For the side parameter this means that the side indicator is $\frac{1}{2}$ for observations where the digit was smaller than 5, and  $-\frac{1}{2}$ for observations where the digit was greater than 5. For the digit parameters it means that the digit indicator is 1 when for the observation the digit was equal to 7, 6, 4 or 3, and 0 when it was not. 
&nbsp;

\scriptsize 
```{r digitparameters}

# create variable inter for the intercept gamma 
myData_dataframe$inter <- rep(1, nrow(indat1))

### Add variable with information of the side of the digit to the dataframe, for parameter beta    
# j = condition, if j > 5 than x = -1/2, if j < 5 than x = 1/2 
# indat1$stim 0 = 2, 1=3, 2=4, 3=6, 4=7, 5=8 
# so if bigger than 2 than 1/2, else than -1/2 
for (j in 1:nrow(indat1)){
  if (indat1$stim[j] < 3) {
    myData_dataframe$side[j] <- 1/2}
  else myData_dataframe$side[j] <- -1/2
}

### Add variable with information about difference between digits to the dataframe, for parameters delta's 
#### Difference 8 and 7 
for (j in 1:nrow(indat1)){ 
  if (indat1$stim[j] == 5) {  # 5 = 8 , reference: https://github.com/PerceptionCognitionLab/data0/blob/master/lexDec-dist5/ld5.txt
    myData_dataframe$dif1[j] <- 0}
  else if (indat1$stim[j] == 4) {
    myData_dataframe$dif1[j] <- 1}
  else myData_dataframe$dif1[j] <- 0
}

#### Difference 7 and 6 
for (j in 1:nrow(indat1)){ 
  if (indat1$stim[j] == 4) {  # 4 = 7 , reference: https://github.com/PerceptionCognitionLab/data0/blob/master/lexDec-dist5/ld5.txt
    myData_dataframe$dif2[j] <- 0}
  else if (indat1$stim[j] == 3) {  # 3 = 6 
    myData_dataframe$dif2[j] <- 1}
  else myData_dataframe$dif2[j] <- 0
}

#### Difference 4 and 3 
for (j in 1:nrow(indat1)){ 
  if (indat1$stim[j] == 1) {  # 1 = 3, 2 = 4 , reference: https://github.com/PerceptionCognitionLab/data0/blob/master/lexDec-dist5/ld5.txt
    myData_dataframe$dif3[j] <- 0}
  else if (indat1$stim[j] == 2) {
    myData_dataframe$dif3[j] <- 1}
  else myData_dataframe$dif3[j] <- 0
}

#### Difference 3 and 2 
for (j in 1:nrow(indat1)){ 
  if (indat1$stim[j] == 0) {  # 0 = 2, 1 = 3 , reference: https://github.com/PerceptionCognitionLab/data0/blob/master/lexDec-dist5/ld5.txt
    myData_dataframe$dif4[j] <- 0}
  else if (indat1$stim[j] == 1) {
    myData_dataframe$dif4[j] <- 1}
  else myData_dataframe$dif4[j] <- 0
}
```
\normalsize

|    Lastly, we add an indicator to the data frame (`ind`) that specifies which observation belongs to which individual (or group).  
&nbsp;

\scriptsize 
```{r participantnumber, eval = FALSE, echo = FALSE}
## Data add necessary information 
### Add participant number 
for (j in 1:nrow(indat1)){
  if (j == 1) {
    myData_dataframe$ind[j] <- 1}
  else if (indat1$sub[j] == indat1$sub[j-1]) {
    myData_dataframe$ind[j] <- indat1$sub1[j-1]}
  else myData_dataframe$ind[j] <- indat1$sub1[j-1] + 1
}
```
\normalsize

\scriptsize 
```{r participantnumber2}
## Data add necessary information 
### Add participant number 
myData_dataframe$ind <- indat1$sub + 1

```
\normalsize

### Priors

|    Finally, we specify the priors. In `brms`, priors can be set on the general effects represented by `class = b` and on the individual variation represented by `class = sd`. It is important to note that priors can only be set on the specific parameters that are part of the brms model specification. For example, `brms` only allows setting priors on standard deviations and not on variances.
Therefore, for the normal model of the symbolic distance effect, we will use the priors for standard deviations specified in Equation 7. To set the priors, we use the `set_prior` function, as shown below. In this function, first, the distribution is specified, using the `Stan` programming language. Then, the type of parameter to which the prior applies is defined by `class`. The parameter to which this prior applies is specified by `coef`. For the priors on the individual effects, an additional element has to be specified, namely, the grouping variable (i.e., `group`). For example, the prior distributions on the general side effect, $\mu_{\beta}$, and on its variability between individuals, $\sigma_{\beta}$, can be specified as follows: 
&nbsp;

\scriptsize 
```{r priorsillusbrms2, eval = FALSE, echo = TRUE}
set_prior("normal(0,0.3)", class = "b", coef="side")  # general effect 
set_prior("inv_gamma(13.8, 5.3)", class = "sd",       # individual variation 
          coef="side", group = "ind") 
```
\normalsize

For the digit classification task, we want to specify a truncated normal prior on the intercept, because response times cannot be negative. This means that we are setting a different prior on the intercept than on the side and digit effects. To accomplish this, we split up the formula in two different parts. For each variable, we specify to which part of the formula they belong with `nlpar`. In addition, for the intercept, we specify that the prior has a lower bound of zero with `lb`. All priors are saved in an R `data.frame` that is used in the `brm` function.

\scriptsize 
```{r priors}
priorsmodel <- c(set_prior("normal(0.5,1)", class = "b", lb = 0, nlpar = "b"),  
                  set_prior("inv_gamma(13.8, 6.3)", class = "sd", coef="inter", group = "ind", nlpar = "b"), 
                  set_prior("normal(0,0.3)", class = "b", coef="side",  nlpar = "a"),  
                  set_prior("inv_gamma(13.8,5.3)", class = "sd", coef="side", group = "ind",  nlpar = "a"), 
                  set_prior("normal(0,0.3)", class = "b", coef="dif1",  nlpar = "a"), 
                  set_prior("inv_gamma(13.8,5.3)", class = "sd", coef="dif1", group = "ind",  nlpar = "a"),
                  set_prior("normal(0,0.3)", class = "b", coef="dif2",  nlpar = "a"),  
                  set_prior("inv_gamma(13.8,5.3)", class = "sd", coef="dif2", group = "ind",  nlpar = "a"),
                  set_prior("normal(0,0.3)", class = "b", coef="dif3",  nlpar = "a"), 
                  set_prior("inv_gamma(13.8,5.3)", class = "sd", coef="dif3", group = "ind",  nlpar = "a"),
                  set_prior("normal(0,0.3)", class = "b", coef="dif4",  nlpar = "a"),  
                  set_prior("inv_gamma(13.8,5.3)", class = "sd", coef="dif4", group = "ind",  nlpar = "a"),
                  set_prior("inv_gamma(13.8, 6.3)", class = "sigma"))

```
\normalsize

|    As described in the paper, we could check whether these priors are reasonable considering our expectations by performing prior prediction.
&nbsp;

\scriptsize 
```{r priorprediction, echo = FALSE, fig.cap="Predicted mean RT difference between conditions per trial. The blue line represents the aggregated trend. ", fig.align = "left", fig.height=4, fig.width=5}
set.seed(123)

Ipp <- 10  # repeat 10 times 
Jpp <- 6   # digit 2, 3, 4, 6, 7, and 8 
Kpp <- 60  # 60 observations per condition per person

exppp <- rep(1:Ipp, each = Jpp * Kpp)  # which row belongs to which participant 
condpp <- rep(0:5, Ipp, each = Kpp)  # to which condition row belongs  
subpp <- rep(1:Kpp, Ipp * Jpp)  # to which observation per condition per parcticipant belongs row

# add parameters x, u, v, w, z 
xjpp <- NA 
ujpp <- NA 
vjpp <- NA 
wjpp <- NA 
zjpp <- NA 
# side effect, greater or smaller than 5  
for (j in 1:length(condpp)){  # 3 = 6
  if (condpp[j] < 3) {
    xjpp[j] <- 1/2}
  else xjpp[j] <- -1/2
}

### Add variable with information about difference between digits, for parameters delta's 
#### Difference 8 and 7 
for (j in 1:length(condpp)){ 
  if (condpp[j] == 5) {  # 5 = 8 , 
    ujpp[j] <- 0}
  else if (condpp[j] == 4) {
    ujpp[j] <- 1}
  else ujpp[j] <- 0
}

# Difference 7 and 6 
for (j in 1:length(condpp)){ 
  if (condpp[j] == 4) {  # 4 = 7, 3 = 6  
    vjpp[j] <- 0}
  else if (condpp[j] == 3) {
    vjpp[j] <- 1}
  else vjpp[j] <- 0
}

# Difference 4 and 3 
for (j in 1:length(condpp)){ 
  if (condpp[j] == 1) {  # 1 = 3, 2 = 4 
    wjpp[j] <- 0}
  else if (condpp[j] == 2) {
    wjpp[j] <- 1}
  else wjpp[j] <- 0
}

# Difference 3 and 2 
for (j in 1:length(condpp)){ 
  if (condpp[j] == 0) {  # 0 = 2, 1 = 3
    zjpp[j] <- 0}
  else if (condpp[j] == 1) {
    zjpp[j] <- 1}
  else zjpp[j] <- 0
}

datalistpp3 <- list()
for (i in 1:1000){
# Hyperpriors: These are not individual specific so only sample 1  
mugammapp <- rtruncnorm(1, a = 0, mean = 0.5, sd = sqrt(1))
vargammapp <- rinvgamma(1, 3, 0.7)
mubetapp <- rnorm(1, 0, 0.3)
varbetapp <- rinvgamma(1, 3, 0.5)
mudeltapp <- rnorm(1, 0, 0.3)
vardeltapp <- rinvgamma(1, 3, 0.5)

# Priors: for each person sample prior 1 time 
gammaipp <- rnorm(Ipp, mugammapp, sqrt(vargammapp))
betaipp <-  rnorm(Ipp, mubetapp, sqrt(varbetapp))
delta1pp <- rnorm(Ipp, mudeltapp, sqrt(vardeltapp))
delta2pp <- rnorm(Ipp, mudeltapp, sqrt(vardeltapp))
delta3pp <- rnorm(Ipp, mudeltapp, sqrt(vardeltapp))
delta4pp <- rnorm(Ipp, mudeltapp, sqrt(vardeltapp))

sigmapp <- rinvgamma(1, 3, 0.7)  # Only 1 because not individual specific

# data
Ypp <- rnorm(Ipp*Jpp*Kpp, gammaipp[exppp] + xjpp * betaipp[exppp] + 
              ujpp * delta1pp[exppp] + vjpp * delta2pp[exppp] +
              wjpp * delta3pp[exppp] + zjpp * delta4pp[exppp],
              sqrt(sigmapp))  # model formula 
datartpp <- data.frame(exppp,  # data frame of results and conditions 
                       condpp, 
                       subpp, 
                       repetition = rep(i, length(exppp)), 
                       gammaipp[exppp], betaipp[exppp], delta1pp[exppp],
                       delta2pp[exppp], delta3pp[exppp], delta4pp[exppp], Ypp)
datalistpp3[[i]] <- datartpp
}

# Add datasets under each other 
bigdatapp3 <- do.call(rbind, datalistpp3)

newbigdatapp3 <- bigdatapp3[bigdatapp3$exppp == 1,]

theme_set(theme_apa(base_size = 9))

########### Create figure function

## Or plot mean rt per condition per trial 
priorpredfigurecreate <- function(bigdatapp3brms){
  plotsimpp <- aggregate(bigdatapp3brms$Ypp, list(bigdatapp3brms$condpp, bigdatapp3brms$repetition), mean)
  meanrtsimpp <- aggregate(bigdatapp3brms$Ypp, list(bigdatapp3brms$condpp), mean)  # mean RT per condition

  # Mean difference per digits per trial and side effect (mean(2, 3, 4) - mean (6, 7, 8))
  plotsimpp <- aggregate(bigdatapp3brms$Ypp, list(bigdatapp3brms$condpp, bigdatapp3brms$repetition), mean)
  # meanrtsimpp <- aggregate(bigdatapp3$Ypp, list(bigdatapp3$condpp), mean)

  sideapp3 <- NA 
  diff1app3 <- NA   # length 1000 
  diff2app3 <- NA 
  diff3app3 <- NA
  diff4app3 <- NA 
  repapp3 <- NA

  # Group.2 is the trial number (1000 in total per condition)
  for (i in 1:length(unique(plotsimpp$Group.2))){
    repapp3[i] <- i 
    sideapp3[i] <- ((plotsimpp$x[plotsimpp$Group.1==0 & plotsimpp$Group.2 == i] + plotsimpp$x[plotsimpp$Group.1==1 & plotsimpp$Group.2 == i] + plotsimpp$x[plotsimpp$Group.1==2 & plotsimpp$Group.2 == i]) - 
                   (plotsimpp$x[plotsimpp$Group.1==3 & plotsimpp$Group.2 == i] + plotsimpp$x[plotsimpp$Group.1==4 & plotsimpp$Group.2 == i] + plotsimpp$x[plotsimpp$Group.1==5 & plotsimpp$Group.2 == i]))
    diff1app3[i] <- (plotsimpp$x[plotsimpp$Group.1==4 & plotsimpp$Group.2 == i]- plotsimpp$x[plotsimpp$Group.1==5 & plotsimpp$Group.2 == i])  # 4 = 7 and 5 = 8 
    diff2app3[i] <- (plotsimpp$x[plotsimpp$Group.1==3 & plotsimpp$Group.2 == i]- plotsimpp$x[plotsimpp$Group.1==4 & plotsimpp$Group.2 == i])  # 3 = 6, 4 = 7

    diff3app3[i] <- (plotsimpp$x[plotsimpp$Group.1==2 & plotsimpp$Group.2 == i]- plotsimpp$x[plotsimpp$Group.1==1 & plotsimpp$Group.2 == i])  # 2 = 4, 1 = 3

    diff4app3[i] <- (plotsimpp$x[plotsimpp$Group.1==1 & plotsimpp$Group.2 == i]- plotsimpp$x[plotsimpp$Group.1==0 & plotsimpp$Group.2 == i])  # 1 = 3, 0 = 2

  }
  #diffdatapp3 <- data.frame(repapp3, diff1app3, diff2app3, diff3app3, diff4app3)  # length 1000 
  diffdatapp33 <- data.frame(trial = rep(repapp3, 5), condition = rep(1:5, each = length(repapp3)), difference =  c(sideapp3, diff4app3, diff3app3, diff2app3, diff1app3))

  # difference aggregated 
  diffmeandatapp33 <- data.frame(condition = 1:5, 
                                 difference =  c(((meanrtsimpp$x[meanrtsimpp$Group.1==0] + meanrtsimpp$x[meanrtsimpp$Group.1==1] + meanrtsimpp$x[meanrtsimpp$Group.1==2])-(meanrtsimpp$x[meanrtsimpp$Group.1==3] + meanrtsimpp$x[meanrtsimpp$Group.1==4] + meanrtsimpp$x[meanrtsimpp$Group.1==5])), 
                                                 (meanrtsimpp$x[meanrtsimpp$Group.1==1] - meanrtsimpp$x[meanrtsimpp$Group.1==0]), 
                                                 (meanrtsimpp$x[meanrtsimpp$Group.1==2] - meanrtsimpp$x[meanrtsimpp$Group.1==1]), 
                                                 (meanrtsimpp$x[meanrtsimpp$Group.1==3] - meanrtsimpp$x[meanrtsimpp$Group.1==4]), 
                                                 (meanrtsimpp$x[meanrtsimpp$Group.1==4] - meanrtsimpp$x[meanrtsimpp$Group.1==5])))

  # Old figure 
  diffrtplotsimpp <- ggplot(diffdatapp33, aes(y = difference, x = as.factor(condition))) + 
    geom_line(aes(group = trial), alpha = .014) + 
    geom_point(data = diffdatapp33, aes(group = trial), alpha = 0.014) + 
    geom_line(data = diffmeandatapp33, aes(group = 1), color='blue') + 
    geom_point(data = diffmeandatapp33, aes(group = 1), color='blue') + 
    labs(x = "Digits", y = "Response Time Difference (Seconds)") + 
    theme_bw() + 
    theme(axis.line = element_line(colour = "black"), panel.border = element_blank(), panel.background = element_blank()) + 
    scale_x_discrete(breaks = 1:5, labels = c("side", "3-2", "4-3", "6-7", "7-8"))   # + ylim(-0.1,0.1) then you see the difference better but missing a lot of trials 
    theme_set(theme_apa(base_size = 9))

  priorpredictionplot1 <- diffrtplotsimpp + ylim(-15,15)

  # Try new figure

  # Create violin plot 
  priorpredictionplotnew_1 <- ggplot(diffdatapp33, aes(y = difference, x = as.factor(condition))) + 
    geom_violin(width = 1, alpha = 0.1) + 
    geom_quasirandom(alpha = 0.014, width = 0.2, dodge.width=1) + 
    geom_point(data = diffmeandatapp33, aes(group = 1), color='blue') + 
    labs(x = "Digits", y = "Response Time Difference (Seconds)") + 
    theme_bw() + 
    theme(axis.line = element_line(colour = "black"), panel.border = element_blank(), panel.background = element_blank()) + 
    scale_x_discrete(breaks = 1:5, labels = c("Side", "3-2", "4-3", "6-7", "7-8"))   # + ylim(-0.1,0.1) then you see the difference better but missing a lot of trials

  return(priorpredictionplotnew_1)
}

########### End figure function

priorpredfigurecreate(bigdatapp3)

```
\normalsize

## Fit the Model

|    Now that we have the required input, we can estimate the model.  
&nbsp;

\scriptsize 
``` {r modelfit, eval = FALSE}
model_fit <- brm(formula = formula,                      # Model formula 
                 data = myData_dataframe,                # Dataset (parameters, observations) 
                 family = gaussian(),                    # Family, distribution dependent variable 
                 prior = priorsmodel,                    # Specification of the priors 
                 # sample_prior = "yes",                 # To calculate Bayes factors for point hypotheses  
                 warmup = 1000,                          # Nr. iterations for warmup (per chain)
                 iter = 4000,                            # Nr. iterations (per chain)
                 chains = 4,                             # Nr. chains 
                 core = 4)                               # Nr. of cores for parallel estimation

```
\normalsize

We have already estimated the model and saved it in the R object `brmsmodelfit_final`. In case you do not want to wait until the model has been fitted but continue immediately with the rest of this tutorial, you can load this R object that contains the model estimations.  
&nbsp;

\scriptsize 
```{r loadmodelfit, eval = FALSE}
model_fit <- readRDS(file = "myPath/brmsmodelfit_final.rds")

# For model comparison - sample_prior was set to "yes" 
model_fit_bf <- readRDS("myPath/brmsmodelfit_BF_final.rds")

```
\normalsize

\scriptsize 
```{r loadmodelfits, show = FALSE}
model_fit <- readRDS(file = "/Users/myrtheveenman/Documents/GitHub/multilevel-tutorial-paper/R objects/brms/Truncated/BRMSmodelfit18_03012022.rds")

# For model comparison - sample_prior was set to "yes" 
model_fit_bf <- readRDS("/Users/myrtheveenman/Documents/GitHub/multilevel-tutorial-paper/R objects/brms/Truncated/BRMSmodelfit18_BF_03012022.rds")

```
\normalsize

## Output 
|    The `summary` function provides an overview of the model fit. From this function, we can get a lot of information.

\scriptsize 
```{r modelsummary}
summary(model_fit)
```
\normalsize

|    `brms` offers the `launch_shinystan` function which will load a shiny app with the results of the model fit. It is very extensive. It, for instance, includes model diagnostics as trace plots, posterior plots, prediction plots and way more. 
&nbsp;

\scriptsize 
```{r shinyrstan, eval = FALSE}
launch_shinystan(model_fit)
```
\normalsize

It also has the possibility to save plots. However, if you want the full control over the layout, it might be better to code them yourself. Therefore, we will show how to create the figures presented in the paper.

### Trace plot

|    The trace plot shows whether the posterior distribution of the parameter has converged. With the package `mcmc_plot` function (used to be `stanplot`) it is possible to obtain the trace plots per parameter by setting the option `type` to `trace`. With the `variable` argument, you can specify for which parameters you want to inspect the trace plots. With `variables`, you can check the names of the parameters in the `brms` output.

\scriptsize 
```{r checkvarnames, eval = FALSE}
variables(model_fit)
```
\normalsize

The trace plot will automatically remove the warm-up (burn-in) period from the chain (which is desirable), however, the x axis will indicate that it is showing iteration 0 to 4000 (in our case) while it is actually showing the 1000th to 5000th iteration. As the `plot` option returns a `ggplot` object, it is possible to manually adjust the plot using `ggplot` functions. We created the function `traceplotfunc` to obtain the trace plot for the parameter you are interested in. In the figure below, we show the trace plots for all the general effects and variances. 
&nbsp;

\scriptsize 
```{r traceplotB, fig.cap="Trace plots of the general effect parameters.", fig.align = "left", fig.height = 6, fig.width = 8}

# Default trace plot for parameter mu4
# I think warmup already deleted, but then have to adjust labels x axis 
trdif2brms <- mcmc_plot(model_fit,       # Model fit 
                       variable = "b_a_dif2",  # Select parameter, 
                       # it is also possible to select more than one at the same time, 
                       # such as c("mu2", "mu3","mu4")
                       type = "trace")   # Specify that you want trace plots

# Function 
traceplotfunc <- function(parameter, title){
trdif2brms <- mcmc_plot(model_fit,       # Model fit 
                       variable = parameter,  # Select parameter, 
                       type = "trace")   # Specify that you want trace plots

trdif2brms2 <- trdif2brms + 
               scale_x_continuous(breaks=c(1000, 3000, 5000), 
                                  labels=c(2000, 4000, 6000)) + 
                theme_classic() +              
                theme(axis.title.y=element_text(angle = 0, 
                                                vjust = 0.5, 
                                                hjust = -0.5, 
                                                margin = margin(0, 0.7, 0, 0, "cm"), size = 14), 
                      axis.line.x = element_line(color="black"), 
                     axis.line.y = element_line(color="black"), 
                     #legend.position = "none",  
                     plot.margin = unit(c(0.3,0.1,0.4,0),"cm")) + 
               scale_color_manual(values=c("#E66101", "#998EC3", "#542788", "#F1A340")) + 
               ylab(title) + 
               xlab("Iteration") + 
               labs(subtitle = "") #+ 
               #ylim(c(0.01, 0.07)) + 
               #scale_y_continuous(limits = c(0.01, 0.07), 
                                  #breaks = c(0.02, 0.04, 0.06), 
                                  #labels = c(0.02, 0.04, 0.06))

return(trdif2brms2)
}

# Show trace plot for one digit parameter 
tr2 <- traceplotfunc(parameter = "b_a_side", title = expression(~mu[~beta]))
tr2 <- tr2 + theme(legend.position = "none")
tr3 <- traceplotfunc(parameter = "b_a_dif1", title = expression(~mu[~delta[~7]]))
tr3 <- tr3 + theme(legend.position = "none")
tr4 <- traceplotfunc(parameter = "b_a_dif2", title = expression(~mu[~delta[~6]]))
tr4 <- tr4 + theme(legend.position = "none")
tr5 <- traceplotfunc(parameter = "b_a_dif3", title = expression(~mu[~delta[~4]]))
tr5 <- tr5 + theme(legend.position = "none")
tr6 <- traceplotfunc(parameter = "b_a_dif4", title = expression(~mu[~delta[~3]]))
legendtr <- cowplot::get_legend(tr6)
tr6 <- tr6 + theme(legend.position = "none")

# Plot general effects together 
grid.arrange(tr2, tr3, tr4, tr5, tr6, legendtr, nrow = 2, ncol = 3)
```
\normalsize

\scriptsize 
```{r traceplotsd, }
# Show trace plot for one sd parameter 
tr2 <- traceplotfunc(parameter = "sd_sub1__a_side", title = expression(~sigma[~beta]))
tr2 <- tr2 + theme(legend.position = "none")
tr3 <- traceplotfunc(parameter = "sd_sub1__a_dif1", title = expression(~sigma[~delta[~7]]))
tr3 <- tr3 + theme(legend.position = "none")
tr4 <- traceplotfunc(parameter = "sd_sub1__a_dif2", title = expression(~sigma[~delta[~6]]))
tr4 <- tr4 + theme(legend.position = "none")
tr5 <- traceplotfunc(parameter = "sd_sub1__a_dif3", title = expression(~sigma[~delta[~4]]))
tr5 <- tr5 + theme(legend.position = "none")
tr6 <- traceplotfunc(parameter = "sd_sub1__a_dif4", title = expression(~sigma[~delta[~3]]))
legendtr <- cowplot::get_legend(tr6)
tr6 <- tr6 + theme(legend.position = "none")

# Plot general effects together 
grid.arrange(tr2, tr3, tr4, tr5, tr6, legendtr, nrow = 2, ncol = 3)
```
\normalsize 
Because there are many parameters (each for every individual), it is hard to inspect them all individually. A more efficient way to check if the parameters have converged is by checking the $\hat{R}$ and the number of effective samples.

### Rhat & Number of Effective Samples

|    As for the trace plots, the `mcmc_plot` function offers the possibility to plot the $\hat{R}$ (Rhat) and the ratio of effective sample size to the total posterior sample size (this is different from what is plotted in the paper). 
&nbsp;

\scriptsize 
```{r rhatneffstandard, eval = FALSE}
# Plot frequency of rhat 
mcmc_plot(model_fit, type = "rhat_hist")   # Specify that you want trace plots

# Plot frequency of the ratio 
mcmc_plot(model_fit, type = "neff_hist") 
```
\normalsize

However, these figures can be manually adjusted. To obtain the figures from the paper, the following steps can be followed. First, we save the values for $\hat{R}$ and parameter names.

\scriptsize 
```{r rhatexamp}
# Obtain rhat and neff brms 
## Rhat 
rhatbrms <- mcmc_plot(model_fit, type = "rhat_hist")$data$value  # save rhat values 
rhatbrms_para <- mcmc_plot(model_fit, type = "rhat_hist")$data$parameter  # save parameter names 
rhatbrmsratio_comb <- data.frame(parameter = rhatbrms_para, value = rhatbrms)  # combine values and parameters names into one data.frame 
```
\normalsize

Then, we obtain the $\hat{R}$ of the general effects.

\scriptsize 
```{r rhatexamp2}
# Obtain rhats of general effects 
rhatbrms_v2 <- mcmc_plot(model_fit, type = "rhat_hist")$data
fixedrhatbrms_v2 <- c(rhatbrms_v2$value[rhatbrms_v2$parameter == "b_b_inter"], 
                      rhatbrms_v2$value[rhatbrms_v2$parameter == "b_a_side"], 
                      rhatbrms_v2$value[rhatbrms_v2$parameter == "b_a_dif1"], 
                      rhatbrms_v2$value[rhatbrms_v2$parameter == "b_a_dif2"], 
                      rhatbrms_v2$value[rhatbrms_v2$parameter == "b_a_dif3"], 
                      rhatbrms_v2$value[rhatbrms_v2$parameter == "b_a_dif4"])

deltameanrrbrms <- mean(fixedrhatbrms_v2[3:6])  # compute mean of delta parameters (digit effects)

```
\normalsize

We also save the $\hat{R}$ of the individual effects.

\scriptsize 
```{r rhatexamp3}
# Obtain rhats of all individual effects 
rhatbrmsratio3 <- c(rhatbrmsratio_comb[grep(",inter]", rhatbrmsratio_comb$parameter),]$value,
                      rhatbrmsratio_comb[grep(",side]", rhatbrmsratio_comb$parameter),]$value,
                      rhatbrmsratio_comb[grep(",dif1]", rhatbrmsratio_comb$parameter),]$value,
                      rhatbrmsratio_comb[grep(",dif2]", rhatbrmsratio_comb$parameter),]$value,
                      rhatbrmsratio_comb[grep(",dif3]", rhatbrmsratio_comb$parameter),]$value,
                      rhatbrmsratio_comb[grep(",dif4]", rhatbrmsratio_comb$parameter),]$value)

```
\normalsize

Next, the data of the individual effects is restructured to the format used in *ggplot*.

\scriptsize 
```{r rhatexamp4}
# Restructure in a data frame we can use to create a figure with ggplot 
rstanrhatneff_v2 <- data.frame(package = rep("Brms", 312), 
                                 parameter = rep(c("Gamma", "Beta", "Delta",                                                          "Delta", "Delta", "Delta"), 
                                                 each = 52), 
                                 rhat = rhatbrmsratio3)

```
\normalsize

We follow the same steps for the number of effective samples (Neff).

\scriptsize 
```{r neffexamp5}
neffbrmsratio_value <- mcmc_plot(model_fit, type = "neff_hist")$data$value   # save neff values, these are the values of the ration Neff/N, N representing the total number of iterations
neffbrmsratio_para <- mcmc_plot(model_fit, type = "neff_hist")$data$parameter  # save parameter names 
neffbrmsratio_comb <- data.frame(parameter = neffbrmsratio_para, value = neffbrmsratio_value)  # combine values and parameter names into data frame

# Neff of general effects 
fixedneffbrms2 <- c(neffbrmsratio_comb[grep("b_inter", neffbrmsratio_comb$parameter),]$value[1],
                      neffbrmsratio_comb[grep("b_a_side", neffbrmsratio_comb$parameter),]$value[1],
                      neffbrmsratio_comb[grep("b_a_dif1", neffbrmsratio_comb$parameter),]$value[1],
                      neffbrmsratio_comb[grep("b_a_dif2", neffbrmsratio_comb$parameter),]$value[1],
                      neffbrmsratio_comb[grep("b_a_dif3", neffbrmsratio_comb$parameter),]$value[1],
                      neffbrmsratio_comb[grep("b_a_dif4", neffbrmsratio_comb$parameter),]$value[1])

fixedneffbrms3 <- fixedneffbrms2 * 12000  # we multiply values with total number of iterations (12000) to get Neff (used to be ration Neff/N, N representing the total number of iterations)
fixedneffbrmsmean <- mean(fixedneffbrms3[3:6])  # calculate mean Neff of delta parameters

# Neff of individual effects  
neffbrmsratio3 <- c(neffbrmsratio_comb[grep(",inter]", neffbrmsratio_comb$parameter),]$value,
                      neffbrmsratio_comb[grep(",side]", neffbrmsratio_comb$parameter),]$value,
                      neffbrmsratio_comb[grep(",dif1]", neffbrmsratio_comb$parameter),]$value,
                      neffbrmsratio_comb[grep(",dif2]", neffbrmsratio_comb$parameter),]$value,
                      neffbrmsratio_comb[grep(",dif3]", neffbrmsratio_comb$parameter),]$value,
                      neffbrmsratio_comb[grep(",dif4]", neffbrmsratio_comb$parameter),]$value)
neffbrms2 <- neffbrmsratio3  * 12000  # we multiply values with total number of iterations (12000) to get Neff (used to be ration Neff/N, N representing the total number of iterations)

# Restructure in a data frame we can use to create a figure with ggplot 
neffrstanbrms2 <- data.frame(package = rep("Brms", 312), 
                               parameter = rep(c("Gamma", "Beta", "Delta", "Delta",                                                  "Delta", "Delta"), each = 52),
                               neff = c(neffbrms2))

```
\normalsize

Finally, we visualize the $\hat{R}$ and Neff using violin plots. 
&nbsp;

\scriptsize 
```{r rhatneff, fig.cap="The frequency of the $\\hat{R}$ statistic and the number of effective samples. A: The frequency of the $\\hat{R}$. B: The frequency of the number of effective samples where the dashed lines represent the total number of iterations.", fig.align = "left", fig.height = 4, fig.width = 8}
# Create violin plots 
## Rhat 
  figrhatnew <- ggplot(rstanrhatneff_v2, aes(x = factor(parameter, level = c("Gamma", "Beta", "Delta")), y = rhat, colour = package)) + 
    geom_violin(width = 1) + 
    geom_quasirandom(alpha = 0.1, width = 0.2, dodge.width=1) + 
    geom_point(aes(x = 1, y = fixedrhatbrms_v2[1]), shape = 8, colour="darkorange", size = 2) + 
    geom_point(aes(x = 2, y = fixedrhatbrms_v2[2]), shape = 8, colour="darkorange", size = 2) + 
    geom_point(aes(x = 3, y = deltameanrrbrms ), shape = 8, colour="darkorange", size = 2) + 
    xlab("Parameter") + ylab(expression(hat(R))) +
    scale_colour_manual(values = "darkorange") + 
    theme_classic() + 
    theme(legend.position = "none") + 
    scale_x_discrete(labels= c( 
      expression(~gamma), 
      expression(~beta), 
      expression(~delta))) + 
    labs(subtitle = "C") + 
    theme(axis.title.y = element_text(angle = 0, 
                                      vjust = 0.5, 
                                      hjust = -0.5, 
                                      margin = margin(0, 1.1, 0, 0, "cm")))

## Neff 
figneffnew <- ggplot(neffrstanbrms2, aes(x = factor(parameter, level = c("Gamma", "Beta", "Delta")), y = neff, colour = package)) + 
    geom_violin(width = 1) + 
    geom_quasirandom(alpha = 0.1, width = 0.2, dodge.width=1) + 
    geom_point(aes(x=1, y= fixedneffbrms3[1]), shape = 8, colour="darkorange", size = 2) + 
    geom_point(aes(x=2, y= fixedneffbrms3[2]), shape = 8, colour="darkorange", size = 2) + 
    geom_point(aes(x=3, y= fixedneffbrmsmean), shape = 8, colour="darkorange", size = 2) + 
    xlab("Parameter") + ylab("Number of effective samples") +
    scale_colour_manual(values = "darkorange") + 
    theme_classic() + 
    theme(legend.position = "none") + 
    scale_x_discrete(labels= c( 
      expression(~gamma), 
      expression(~beta), 
      expression(~delta))) + 
    labs(subtitle = "D") + 
    geom_hline(yintercept = 20000, linetype = "dashed", color = "darkorange", size = .3)

# Plot together 
grid.arrange(figrhatnew, figneffnew, nrow = 1, ncol = 2)

```
\normalsize

### General effects

|    The general effects are represented by $\mu_{\delta_{7}}$, $\mu_{\delta_{6}}$, $\mu_{\delta_{4}}$, $\mu_{\delta_{3}}$, and $\mu_{\beta}$. To investigate whether there is a digit and side effect, we will inspect the estimates for these parameters.

#### Posterior Distribution

|    With the `mcmc_plot` function, the posterior distribution per parameter can be displayed. 
&nbsp;

\scriptsize 
```{r posteriorstandard, eval = FALSE}
# Seperate plot per parameter of posterior density 
mcmc_plot(model_fit, 
     type = "dens")  
```
\normalsize

|    Another possibility is to use the `bayesplot` package to display all the posterior densities together in one plot. This makes it easier to compare the posterior distributions of each parameter. 
&nbsp;

\scriptsize 
```{r posteriormanual, fig.cap="The posterior distributions for the general effects. The middle line within the distributions represents the mean posterior. The shaded area within the distributions represents 95\\% of the probability mass.", fig.align = "left", fig.height = 4, fig.width = 5}
post8 <- as.array(model_fit)  # change structure of model fit

color_scheme_set("orange")  # plot the distributions in orange

pairs8 <- mcmc_areas(
  post8,    # Structured model fit 
  pars = c("b_a_side", "b_a_dif1", "b_a_dif2", "b_a_dif3", "b_a_dif4"),  # Select parameters 
  prob = 0.95, # 80% intervals  # Plot 95% credible intervals
  prob_outer = 0.99, # 99%     # 99% outer interval
  point_est = "mean"           # Plot line in the middle 
)
# Returns ggplot object

# Adjust ggplot object further   
pairs82 <- pairs8 + 
           theme_apa() + 
           ggtitle("Brms") + 
           xlim(-0.05, 0.095) + 
           #scale_x_continuous(breaks = c(-0.025, 0, 0.025, 0.05, 0.075)) + 
           scale_y_discrete(labels= c( 
                                              expression(~mu[~beta]), 
                                              expression(~mu[~delta[~7]]), 
                                              expression(~mu[~delta[~6]]), 
                                              expression(~mu[~delta[~4]]), 
                                              expression(~mu[~delta[~3]])))

pairs82
```
\normalsize

|    We can present more information about the distribution by providing a table with the estimated mean of the posterior distribution, the standard error of this mean, the lower and upper bound of the 95% credible interval, the $\hat{R}$ and the number of effective samples. 
&nbsp;

\scriptsize 
```{r geneftable, results = "asis", cache = TRUE}
# Obtain the mean, standard error, 95% credible interval, rhat and number of effective samples for the intercept, and the digit and side parameters 
fitbrmshiertable <- summary(model_fit)[[14]][,c("Estimate", "Est.Error", "l-95% CI", "u-95% CI", "Bulk_ESS", "Rhat")]  # variance

# The results are transformed to a dataframe that allows us to plot the results nicely in a table 
fitbrmshiertable2 <- as.data.frame(fitbrmshiertable)
fitbrmshiertable2 <- rbind(fitbrmshiertable2[6,], fitbrmshiertable2[1:5,])

# We manually set the row names of the table using latex for mathematical symbols 
rownamesfitbrmshiertable <-  c("$\\mu_{\\gamma}$","$\\mu_{\\beta}$", "$\\mu_{\\delta_{7}}$", 
                               "$\\mu_{\\delta_{6}}$", "$\\mu_{\\delta_{4}}$", "$\\mu_{\\delta_{3}}$")

# We manually set the column names of the table 
colnames(fitbrmshiertable2) <- c("Mean", "SE", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")

# The column with row names is added to the table 
fitbrmshiertable3 <- add_column(fitbrmshiertable2, 
                                 Parameters = rownamesfitbrmshiertable, 
                                 .before = "Mean")

# Using the apa_table function from the papaja package the results are presented in an apa table 
apa_table(fitbrmshiertable3,   # Data frame with results 
          row.names = FALSE,    # Default row names are deleted (we set them ourself)
          caption = "Posterior Mean, Standard Error (SE) of the Mean, Lower and 
          Upper Bound of the 95\\% Credible Interval, the Number of Effective 
          Samples, and the $\\hat{R}$ of the General Effect Parameters. ", 
          align = c("l", "c", "c", "c", "c", "c", "c"),   # How should results be presented, in the midle of the table or outlined left 
          digits = 3,       # The number of digits behind the ".". 
          placement = "h",  # Place after code 
          escape = FALSE)   # This indicates that the results contain latex code that should be read as latex code. It is important to do this in combination with putting the results to "asis" in the {R} section above.

```
\normalsize

### Individual Effects

\    Next, we look at the individual deviations from the general effects. Therefore, we inspect the standard deviation and the individual estimates of the digit and side effects.

#### Standard Deviation Estimates

|    First, we can show a table with the posterior distribution estimates of the standard deviation parameters. Note that this is different from *rstan* where we estimated the variances instead of the standard deviations. 
&nbsp;

\scriptsize 
```{r inddevtable, results = "asis", cache = TRUE}
# Obtain standard deviation estimates 
fitbrmshiertable2 <- summary(model_fit)[[17]]$sub1[,c("Estimate", "Est.Error", "l-95% CI", "u-95% CI", "Bulk_ESS", "Rhat")]

# The results are transformed to a dataframe that allows us to plot the results nicely in a table 
fitbrmshiertable22 <- as.data.frame(fitbrmshiertable2)
fitbrmshiertable22 <- rbind(fitbrmshiertable22[6,], fitbrmshiertable22[1:5,])

# We manually set the row names of the table using latex for mathematical symbols 
rownamesfitbrmshiertable2 <-  c("$\\sigma_{\\gamma}$","$\\sigma_{\\beta}$", "$\\sigma_{\\delta_{7}}$", 
                                "$\\sigma_{\\delta_{6}}$", "$\\sigma_{\\delta_{4}}$", "$\\sigma_{\\delta_{3}}$")

# The column with row names is added to the table 
colnames(fitbrmshiertable22) <- c("Mean", "SE", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")

# The column with row names is added to the table 
fitbrmshiertable32 <- add_column(fitbrmshiertable22, 
                                  Parameters = rownamesfitbrmshiertable2, 
                                  .before = "Mean")

# Using the apa_table function from the papaja package the results are presented in an apa table 
apa_table(fitbrmshiertable32,  # Data frame with results
          row.names = FALSE,    # Default row names are deleted (we set them ourself)
          caption = "Posterior Standard Deviation, Standard Error (SE) of the Standard Deviation, 
          Lower and Upper Bound of the 95\\% Credible Interval, the Number of 
          Effective Samples, and the $\\hat{R}$ of the Standard Deviation Parameters           as estimated by Brms", 
          align = c("l", "c", "c", "c", "c", "c", "c"),  # How should results be presented, in the midle of the table or outlined left 
          digits = 3,        # The number of digits behind the ".". 
          place = "h",       # Place after code 
          escape = FALSE)    # This indicates that the results contain latex code that should be read as latex code. It is important to do this in combination with putting the results to "asis" in the {R} section above.

```
\normalsize

#### Individual Estimates

|    Next, we can look at the individual estimates for the digit and side effect parameters in two different ways. First, we plot the individual estimates and display their 95% credible interval. The credible interval is displayed in pink when it contains zero, and displayed in blue when it does not contain zero. In addition, the dashed line represent the general effect estimate. 
|    The code below shows you how to obtain this code for one parameter. The other parameters follow logically from this example and would result in the figure below. 
&nbsp;

\scriptsize 
```{r randefplotexample}
# Gamma 
# Extract the parameter estimates 
brmsgammar <- coef(model_fit)$sub1[,,"b_inter"][,c("Estimate", "Q2.5", "Q97.5")]
brmsgammar <- as.data.frame(brmsgammar)

# Extract the general effect estimate of the parameter 
fixbremsgammar <- fixef(model_fit)["b_inter","Estimate"]

# Order the estimates ascending 
brmsgammar2 <- brmsgammar[order(brmsgammar$Estimate),]
brmsgammar2$order <- c(1:52)

# Add parameter to dataframe. When the credible interval contains zero, the parameter is set pink, otherwise to blue 
for (i in 1:nrow(brmsgammar2)){  # 1 if CI contains zero, 0 otherwise 
  if (brmsgammar2$Q2.5[i]*brmsgammar2$Q97.5[i] < 0) {
    brmsgammar2$zero2[i] <- "#F8766D"
  } else (brmsgammar2$zero2[i] <- "#00BFC4") 
}

# Plot the individual estimates for gamma (intercept)
gammarandomplotbrms <- ggplot(brmsgammar2, aes(x = order, y = Estimate, ymin = Q2.5, ymax = Q97.5)) + 
  geom_linerange(color = brmsgammar2$zero2) + 
  geom_point() + 
  geom_hline(yintercept = 0, linetype = "dashed", alpha = .3) + 
  labs(title = expression(~gamma[~i]), y = "") + 
  geom_hline(yintercept = fixbremsgammar, color = "red", alpha = .3) +  
  theme(legend.position = "none", 
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),)

```
\normalsize

```{r indvestirstan, echo = FALSE, fig.cap="The posterior means with 95\\% credible interval in pink and red as estimated by \\textit{brms} for every individual shown in increasing order. The dashed line represents the general posterior mean. A pink interval means that the interval contains zero, a blue interval represents an interval that does not contain zero.", fig.align = "left", fig.height = 6, fig.width = 8}
# Gamma 
brmsgammar <- coef(model_fit)$sub1[,,"b_inter"][,c("Estimate", "Q2.5", "Q97.5")]
brmsgammar <- as.data.frame(brmsgammar)
fixbremsgammar <- fixef(model_fit)["b_inter","Estimate"]

brmsgammar2 <- brmsgammar[order(brmsgammar$Estimate),]
brmsgammar2$order <- c(1:52)

for (i in 1:nrow(brmsgammar2)){  # 1 if CI contains zero, 0 otherwise 
  if (brmsgammar2$Q2.5[i]*brmsgammar2$Q97.5[i] < 0) {
    brmsgammar2$zero2[i] <- "#F8766D"
  } else (brmsgammar2$zero2[i] <- "#00BFC4") 
}

gammarandomplotbrms <- ggplot(brmsgammar2, aes(x = order, y = Estimate, ymin = Q2.5, ymax = Q97.5)) + 
  geom_linerange(color = brmsgammar2$zero2) + 
  geom_point() + 
  geom_hline(yintercept = 0, linetype = "dashed", alpha = .3) + 
  labs(title = expression(~gamma[~i]), y = "Estimation") + 
  geom_hline(yintercept = fixbremsgammar, color = "red", alpha = .3) +  
  theme(legend.position = "none", 
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),)

# Beta 
brmsbetar <- coef(model_fit)$sub1[,,"a_side"][,c("Estimate", "Q2.5", "Q97.5")]
brmsbetar <- as.data.frame(brmsbetar)
fixbremsbetar <- fixef(model_fit)["a_side","Estimate"]

brmsbetar2 <- brmsbetar[order(brmsbetar$Estimate),]
brmsbetar2$order <- c(1:52)

for (i in 1:nrow(brmsbetar2)){  # 1 if CI contains zero, 0 otherwise 
  if (brmsbetar2$Q2.5[i]*brmsbetar2$Q97.5[i] < 0) {
    brmsbetar2$zero2[i] <- "#F8766D"
  } else (brmsbetar2$zero2[i] <- "#00BFC4") 
}

betarandomplotbrms <- ggplot(brmsbetar2, aes(x = order, y = Estimate, ymin = Q2.5, ymax = Q97.5)) + 
  geom_linerange(color = brmsbetar2$zero2) + 
  geom_point() + 
  geom_hline(yintercept = 0, linetype = "dashed", alpha = .3) + 
  labs(title = expression(~beta[~"side,i"]), y = "") + 
  geom_hline(yintercept = fixbremsbetar, color = "red", alpha = .3) +  
  theme(legend.position = "none", 
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),)

# Delta 1 
brmsdel1r <- coef(model_fit)$sub1[,,"a_dif1"][,c("Estimate", "Q2.5", "Q97.5")]
brmsdel1r <- as.data.frame(brmsdel1r)
fixbremsdif2r <- fixef(model_fit)["a_dif1","Estimate"]

brmsdel1r2 <- brmsdel1r[order(brmsdel1r$Estimate),]
brmsdel1r2$order <- c(1:52)

for (i in 1:nrow(brmsdel1r2)){  # 1 if CI contains zero, 0 otherwise 
  if (brmsdel1r2$Q2.5[i]*brmsdel1r2$Q97.5[i] < 0) {
    brmsdel1r2$zero2[i] <- "#F8766D"
  } else (brmsdel1r2$zero2[i] <- "#00BFC4") 
}

dif1randomplotbrms <- ggplot(brmsdel1r2, aes(x = order, y = Estimate, ymin = Q2.5, ymax = Q97.5)) + 
  geom_linerange(color = brmsdel1r2$zero2) + 
  geom_point() + 
  geom_hline(yintercept = 0, linetype = "dashed", alpha = .3) + 
  labs(title = expression(~delta[~"7,i"]), y = "") + 
  geom_hline(yintercept = fixbremsdif2r, color = "red", alpha = .3) +  
  theme(legend.position = "none", 
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),)

# Delta 2 
brmsdel2r <- coef(model_fit)$sub1[,,"a_dif2"][,c("Estimate", "Q2.5", "Q97.5")]
brmsdel2r <- as.data.frame(brmsdel2r)
fixbremsdif22r <- fixef(model_fit)["a_dif2","Estimate"]

brmsdel2r2 <- brmsdel2r[order(brmsdel2r$Estimate),]
brmsdel2r2$order <- c(1:52)

for (i in 1:nrow(brmsdel2r2)){  # 1 if CI contains zero, 0 otherwise 
  if (brmsdel2r2$Q2.5[i]*brmsdel2r2$Q97.5[i] < 0) {
    brmsdel2r2$zero2[i] <- "#F8766D"
  } else (brmsdel2r2$zero2[i] <- "#00BFC4") 
}

dif2randomplotbrms <- ggplot(brmsdel2r2, aes(x = order, y = Estimate, ymin = Q2.5, ymax = Q97.5)) + 
  geom_linerange(color = brmsdel2r2$zero2) + 
  geom_point() + 
  geom_hline(yintercept = 0, linetype = "dashed", alpha = .3) + 
  labs(title = expression(~delta[~"6,i"]), y = "Estimation") + 
  geom_hline(yintercept = fixbremsdif22r, color = "red", alpha = .3) +  
  theme(legend.position = "none", 
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),)

# Delta 3 
brmsdel3r <- coef(model_fit)$sub1[,,"a_dif3"][,c("Estimate", "Q2.5", "Q97.5")]
brmsdel3r <- as.data.frame(brmsdel3r)
fixbremsdif3r <- fixef(model_fit)["a_dif3","Estimate"]

brmsdel3r2 <- brmsdel3r[order(brmsdel3r$Estimate),]
brmsdel3r2$order <- c(1:52)

for (i in 1:nrow(brmsdel3r2)){  # 1 if CI contains zero, 0 otherwise 
  if (brmsdel3r2$Q2.5[i]*brmsdel3r2$Q97.5[i] < 0) {
    brmsdel3r2$zero2[i] <- "#F8766D"
  } else (brmsdel3r2$zero2[i] <- "#00BFC4") 
}

dif3randomplotbrms <- ggplot(brmsdel3r2, aes(x = order, y = Estimate, ymin = Q2.5, ymax = Q97.5)) + 
  geom_linerange(color = brmsdel3r2$zero2) + 
  geom_point() + 
  geom_hline(yintercept = 0, linetype = "dashed", alpha = .3) + 
  labs(title = expression(~delta[~"4,i"]), y = "") + 
  geom_hline(yintercept = fixbremsdif3r, color = "red", alpha = .3) +  
  theme(legend.position = "none", 
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),)

# Delta 4 
brmsdel4r <- coef(model_fit)$sub1[,,"a_dif4"][,c("Estimate", "Q2.5", "Q97.5")]
brmsdel4r <- as.data.frame(brmsdel4r)
fixbremsdif4r <- fixef(model_fit)["a_dif4","Estimate"]

brmsdel4r2 <- brmsdel4r[order(brmsdel4r$Estimate),]
brmsdel4r2$order <- c(1:52)

for (i in 1:nrow(brmsdel4r2)){  # 1 if CI contains zero, 0 otherwise 
  if (brmsdel4r2$Q2.5[i]*brmsdel4r2$Q97.5[i] < 0) {
    brmsdel4r2$zero2[i] <- "#F8766D"
  } else (brmsdel4r2$zero2[i] <- "#00BFC4") 
}

dif4randomplotbrms <- ggplot(brmsdel4r2, aes(x = order, y = Estimate, ymin = Q2.5, ymax = Q97.5)) + 
  geom_linerange(color = brmsdel4r2$zero2) + 
  geom_point() + 
  geom_hline(yintercept = 0, linetype = "dashed", alpha = .3) + 
  labs(title = expression(~delta[~"3,i"]), y = "") + 
  geom_hline(yintercept = fixbremsdif4r, color = "red", alpha = .3) +  
  theme(legend.position = "none", 
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),)

# Plot together 
theme_set(theme_apa(base_size = 9))
grid.arrange(gammarandomplotbrms, betarandomplotbrms, dif1randomplotbrms, dif2randomplotbrms, dif3randomplotbrms, dif4randomplotbrms, nrow = 2, ncol = 3)

```

|    The second way is the plot the individual estimates and display the variation of these individual estimates. For this figure, we used the design and code from @j_van_langen_2020_3715576. 
&nbsp;

\scriptsize 
```{r modelestimatesrstanfigure3, fig.cap="Model estimates for digit effect parameters. The points represent the mean parameter estimate for each individual. The violin plot on each right side shows the variance of the individual parameter estimates.", fig.align = "left", fig.height = 4, fig.width = 5}
# Brms
## Create data.frame

### Delta's individual estimates 
estimatesbrms <- coef(model_fit)$sub1
#### Delta 1 - difference 8 and 7 
del1estimatesbrms <- estimatesbrms[,,"a_dif1"]

#### Delta 2 - difference 7 and 6 
del2estimatesbrms <- estimatesbrms[,,"a_dif2"]

#### Delta 3 - difference 4 and 3
del3estimatesbrms <- estimatesbrms[,,"a_dif3"]

#### Delta 4 - difference 3 and 2 
del4estimatesbrms <- estimatesbrms[,,"a_dif4"]

#### create dataframe 
parameter_m7 <- rep(c("Delta 1", "Delta 2", "Delta 3", "Delta 4"), each = 52 )

#### gamma_summary2 not using for plot 
#### beta_summary2 not using for plot 
parest_m7brms <- rbind(del1estimatesbrms[,1], del2estimatesbrms[,1], del3estimatesbrms[,1], del4estimatesbrms[,1])
parest_m7brms2 <- t(parest_m7brms)
parest_m7brms3 <- as.data.frame(parest_m7brms2)
parest_m7brms4 <- c(parest_m7brms3$V1, parest_m7brms3$V2, parest_m7brms3$V3, parest_m7brms3$V4)
datafr_m7brms <- data.frame(parameter_m7, parest_m7brms4)

# Plot 
set.seed(321)
datafr_m7brms$x <- rep(c(1, 2, 3, 4), each = 52)
datafr_m7brms$xj <- jitter(datafr_m7brms$x, amount = 0.09)
datafr_m7brms$id <- rep(c(1:52), 4)

plotmodelest_m7brms <- ggplot(data=datafr_m7brms, aes(y=parest_m7brms4)) +

  #Add geom_() objects
  geom_point(data = datafr_m7brms %>% filter(x=="1"), aes(x=xj), color = 'dodgerblue', size = 1.5,
             alpha = .6) +
  geom_point(data = datafr_m7brms %>% filter(x=="2"), aes(x=xj), color = 'darkgreen', size = 1.5,
             alpha = .6) +
  geom_point(data = datafr_m7brms %>% filter(x=="3"), aes(x=xj), color = 'darkorange', size = 1.5, 
             alpha = .6) +
  geom_point(data = datafr_m7brms %>% filter(x=="4"), aes(x=xj), color = 'gold2', size = 1.5, 
             alpha = .6) +

  geom_line(aes(x=xj, group=id), color = 'lightgray', alpha = .3) +

  geom_line(data=data.frame(x=-1:4.5,y=0), aes(x = x, y = y), linetype = "dashed", size = .3) +

  geom_half_violin(
    data = datafr_m7brms %>% filter(x=="1"),aes(x = x, y = parest_m7brms4), position = position_nudge(x = 3.2), 
    side = "r", fill = 'dodgerblue', alpha = .5, color = "dodgerblue", trim = TRUE) +

  geom_half_violin(
    data = datafr_m7brms %>% filter(x=="2"),aes(x = x, y = parest_m7brms4), position = position_nudge(x = 2.2), 
    side = "r", fill = "darkgreen", alpha = .5, color = "darkgreen", trim = TRUE) +

  geom_half_violin(
    data = datafr_m7brms %>% filter(x=="3"),aes(x = x, y = parest_m7brms4), position = position_nudge(x = 1.2), 
    side = "r", fill = "darkorange", alpha = .5, color = "darkorange", trim = TRUE) +

  geom_half_violin(
    data = datafr_m7brms %>% filter(x=="4"),aes(x = x, y = parest_m7brms4), position = position_nudge(x = 0.2), 
    side = "r", fill = "gold2", alpha = .5, color = "gold2", trim = TRUE) +

  #Define additional settings
  scale_x_continuous(breaks=c(1,2,3,4), labels=c(expression(~delta[~7]), expression(~delta[~6]), expression(~delta[~4]), expression(~delta[~3]))) +
  xlab("Parameter") + ylab("Value") +
  ggtitle('Brms') +
  theme_classic() + 
  coord_cartesian(xlim = c(0.75, 5), ylim = c(-0.1, 0.17))

plotmodelest_m7brms
```
\normalsize

## Model comparison

|    In case we have different models, we can compare them to check under which model the data is most likely. This can be done with the Bayes factor, as explained in the paper.

### Savage-Dickey Density Ratio

In `brms` the Bayes factor, in the form of the Savage-Dickey density ratio, can be obtained by using the `hypothesis` function. To be able to use this function, it is important that the `sample_prior` option in the `brm` function when fitting the model is set to `TRUE`. 
|    For every parameter we want to investigate, we compare the posterior against the null hypothesis. In the `hypothesis` function the null hypothesis is specified by the `hypothesis` option. There, the parameter of interest is set to the null hypothesis, in this case, that the effect will be equal to zero. 
&nbsp;

\scriptsize  
```{r modelcomparison, eval = FALSE, echo = TRUE}
# You have to set sample_prior = TRUE when fitting the model 
# In case you haven't set sample_prior to TRUE, you could load the following model fit where sample_prior was set to TRUE 
brmmodel8bf <- readRDS("myPath/brmsmodelfit_BF_final.rds")

# Side effect 
b1brms <- hypothesis(x = model_fit_bf,           # Model fit 
                     hypothesis = "a_side = 0")     # Null hypothesis

```
\normalsize

|    Then, we can retrieve the Savage-Dickey density ratio by extracting the `Evid.Ratio`.   
&nbsp;

\scriptsize 
```{r bfpart2, eval = FALSE, echo = TRUE}
b1brms$hypothesis$Evid.Ratio 
```
\normalsize

|    Finally, we can plot the prior and posterior with the standard `plot` function. At the point of interest, in our case 0, we can check the height of the prior and posterior distribution. The ratio of these two represents the Savage-Dickey ratio. The function returns a `ggplot` object, that can then be manually adjusted. 
&nbsp;

\scriptsize 
```{r bfpart3, eval = FALSE, echo = TRUE}
plotsdratiob1 <- plot(b1brms,                      # Result hypothesis 
                      plot = F,                    # Does not plot directly 
                      theme = theme_get())[[1]]    # Theme used in plot

plotsdratiob1 + xlim(-0.07, 0.07)                  # Adjust x axis plot 
```
\normalsize

We also created a function that provides this figure. The figure below shows the prior and posterior with the Savage-Dickey density ratio for the general side and digit effects. 
&nbsp;

```{r bfpart4, echo = FALSE, fig.cap="The prior and posterior distribution of the side and digit parameters estimated by the brms package. The dots represent the height of the distributions at zero. The ratio of these dots indicates the Savage-Dickey density ratio. The Bayes factor represents the evidence for the hypothesis that the effect equals zero. A Bayes factor of zero indicates that the evidence is in favor of the alternative hypothesis (the effect does not equal zero).", fig.align = "left", fig.height = 6, fig.width = 10}
# Obtain Savage-Dickey ratio 
# Plot function 
savdicratioplot <- function(plotsdratio){  # put in plot.hypothesisbrms object 
  plotsratio3data <- plotsdratio$data

  #density(plotsratio3data$values[plotsratio3data$Type == "Posterior"])
  xpriorsd <- c(density(plotsratio3data$values[plotsratio3data$Type == "Posterior"])[1]$x, density(plotsratio3data$values[plotsratio3data$Type == "Prior"])[1]$x)  # x 
  ypriorsd <- c(density(plotsratio3data$values[plotsratio3data$Type == "Posterior"])[2]$y, density(plotsratio3data$values[plotsratio3data$Type == "Prior"])[2]$y)  # y 
  xypriorsd <- data.frame(x = xpriorsd, y = ypriorsd, Type = rep(c("Posterior", "Prior"), each = 512))
  maxprior <- max(density(plotsratio3data$values[plotsratio3data$Type == "Prior"])[2]$y)

sdrdif3plot2 <- ggplot(xypriorsd, aes(x, y, color = Type, fill = Type)) + 
  geom_line() + 
  geom_area(position = "identity", alpha = 0.3) + 
  geom_point(aes(x = 0 , y = maxprior), size = 1) + 
  xlim(-0.07, 0.07) + 
  ylab("Density") + 
  xlab("Value") + 
  ggtitle(expression(~delta[~"3,i"])) + 
  scale_fill_manual(values = c("darkorange", "darkgreen")) + 
  scale_color_manual(values = c("darkorange", "darkgreen")) + 
  theme_classic() + 
  theme(legend.position = "none")

return(sdrdif3plot2)

}
b1brms <- hypothesis(x = model_fit_bf, hypothesis = "a_side = 0")  
plotsdratiob1 <- plot(b1brms, plot = F, theme = theme_get())[[1]]
sdrbetplot <- savdicratioplot(plotsdratiob1) 
sdrbetplot <- sdrbetplot + ggtitle(expression(~mu[~beta])) + xlim(-0.1,0.1) + geom_point(aes(x = 0, y = 25), color = "darkorange", size = 1)
sdrbetplot <- sdrbetplot + 
  geom_segment(aes(x = 0, y = 1.5, xend = 0, yend = 24.5), size=0.1, color = "black", linetype = "dashed") + 
  annotate("text", x = c(0.03, 0.05), xend = 0.03, y = 10, yend = 10.4, label = c(expression(paste("BF"["01"], " = ")), round(b1brms$hypothesis$Evid.Ratio, 2)), size = 2)

# Dif 1 
h1brms <- hypothesis(model_fit_bf, "a_dif1 = 0")  # Evid.Ratio is Bayes factor BF01, have to set sample_prior = TRUE when fitting model 
plotsdratio1 <- plot(h1brms,plot = F, theme = theme_get())[[1]]

sdrdif1plot <- savdicratioplot(plotsdratio1) 
sdrdif1plot <- sdrdif1plot + 
  ggtitle(expression(~mu[~delta[~"7"]])) + xlim(-0.1,0.1) + 
  geom_point(aes(x = 0, y = 20), color = "darkorange", size = 1) + 
  geom_segment(aes(x = 0, y = 1.5, xend = 0, yend = 19), size=0.1, color = "black", linetype = "dashed") + 
  annotate("text", x = c(0.035, 0.05), xend = 0.03, y = 10, yend = 10.4, label = c(expression(paste("BF"["01"], " = ")), round(h1brms$hypothesis$Evid.Ratio, 2)), size = 2) + 
  ylab("")

# Dif 2 
h2brms <- hypothesis(model_fit_bf, "a_dif2 = 0")  # Evid.Ratio is Bayes factor BF01, have to set sample_prior = TRUE when fitting model 
plotsdratio2 <- plot(h2brms,plot = F, theme = theme_get())[[1]]

sdrdif2plot <- savdicratioplot(plotsdratio2) 
sdrdif2plot <- sdrdif2plot + 
  ggtitle(expression(~mu[~delta[~"6"]])) + 
  xlim(-0.1,0.1) + 
  geom_point(aes(x = 0, y = 0), color = "darkorange", size = 1) + 
  geom_segment(aes(x = 0, y = 1.5, xend = 0, yend = 0), size=0.1, color = "black", linetype = "dashed") + 
  annotate("text", x = c(0.0, 0.02), xend = 0.03, y = 10, yend = 10.4, label = c(expression(paste("BF"["01"], " = ")), round(h2brms$hypothesis$Evid.Ratio, 2)), size = 2)

# Dif 3 
h3brms <- hypothesis(model_fit_bf, "a_dif3 = 0")  # Evid.Ratio is Bayes factor BF01, have to set sample_prior = TRUE when fitting model 
plotsdratio3 <- plot(h3brms,plot = F, theme = theme_get())[[1]]

# Get data from plot 
plotsratio3data <- plotsdratio3$data

sdrdif3plot <- savdicratioplot(plotsdratio3) 
sdrdif3plot <- sdrdif3plot + 
  ggtitle(expression(~mu[~delta[~"4"]])) + 
  xlim(-0.1,0.1) + 
  geom_point(aes(x = 0, y = 0), color = "darkorange", size = 1) + 
  geom_segment(aes(x = 0, y = 1.5, xend = 0, yend = 0), size=0.1, color = "black", linetype = "dashed") + 
  annotate("text", x = c(0.0, 0.02), xend = 0.03, y = 10, yend = 10.4, label = c(expression(paste("BF"["01"], " = ")), round(h3brms$hypothesis$Evid.Ratio, 2)), size = 2) + 
  ylab("")

# Dif 4 
h4brms <- hypothesis(model_fit_bf, "a_dif4 = 0")  # Evid.Ratio is Bayes factor BF01, have to set sample_prior = TRUE when fitting model 
plotsdratio4 <- plot(h4brms,plot = F, theme = theme_get())[[1]]

# Get data from plot 
plotsratio4data <- plotsdratio4$data

# Get density from data 
# density(plotsratio4data$values[plotsratio4data$Type == "Posterior"])
xpriorsd <- c(density(plotsratio4data$values[plotsratio4data$Type == "Posterior"])[1]$x, density(plotsratio4data$values[plotsratio4data$Type == "Prior"])[1]$x)  # x 
ypriorsd <- c(density(plotsratio4data$values[plotsratio4data$Type == "Posterior"])[2]$y, density(plotsratio4data$values[plotsratio4data$Type == "Prior"])[2]$y)  # y 
xypriorsd <- data.frame(x = xpriorsd, y = ypriorsd, Type = rep(c("Posterior", "Prior"), each = 512))
# max(density(plotsratio4data$values[plotsratio4data$Type == "Prior"])[2]$y)  # 1.343088

# Self made plot with dots  
sdrdif4plot2 <- ggplot(xypriorsd, aes(x, y, color = Type, fill = Type)) + 
  geom_line() + 
  geom_area(position = "identity", alpha = 0.3) + 
  geom_point(aes(x = 0 , y = 1.343088), size = 1) + 
  geom_point(aes(x = 0, y = 9), color = "darkorange", size = 1) + 
  xlim(-0.1, 0.1) + 
  ylab("Density") + 
  xlab("Value") + 
  theme_classic() + 
  ggtitle(expression(~mu[~delta[~"3"]])) + 
  scale_fill_manual(values = c("darkorange", "darkgreen")) + 
  scale_color_manual(values = c("darkorange", "darkgreen")) + 
  geom_segment(aes(x = 0, y = 1.5, xend = 0, yend = 8.5), size=0.1, color = "black", linetype = "dashed") + 
  annotate("text", x = c(-0.01, 0.007), xend = 0.03, y = 10, yend = 10.4, label = c(expression(paste("BF"["01"], " = ")), round(h4brms$hypothesis$Evid.Ratio, 2)), size = 2) + 
  guides(color = guide_legend(override.aes = list(linetype = 0, shape = NA)))

# Extract legend 
library(cowplot)
legendsdr <- cowplot::get_legend(sdrdif4plot2)

sdrdif4plot2 <- sdrdif4plot2 + theme(legend.position = "none") + ylab("")

# the provided plot 
sdrdif4plot <- plotsdratio4 + 
                  xlim(-0.07, 0.07) + 
                  geom_point(aes(x=0, y=0), colour="blue") + 
                  geom_point(aes(x = 0 , y = 1.343088), colour = "green") + 
                  geom_vline(xintercept = 0) + 
                  geom_point(data = plotsratio4data, y = plotsratio4data$values, x = 0, color="red")

# Plot all together 
grid.arrange(sdrbetplot, sdrdif1plot, legendsdr, sdrdif2plot, sdrdif3plot, sdrdif4plot2, nrow = 2, ncol = 3)

```

### Bridge Sampling 
To compare models that differ in more than one parameter, we can use bridge sampling. We could, for instance, compare the side model to the full model (for more details on these models see the paper). First, we need to estimate the models. We already estimated the full model. Next, we estimate the side model:

\scriptsize 
```{r fitsidemodel, eval = FALSE}
# Formula 
formula3 <- brms::bf(rt2 ~ a + b, nl = TRUE, 
                     lf(a ~ 0 + side + (0 + side || sub1), center = TRUE),  
                     lf(b ~ 0 + inter + (0 + inter || sub1), cmc = FALSE)
)

# Priors 
priorsmodel3 <- c(set_prior("normal(0.5,1)", class = "b", lb = 0, nlpar = "b"),  
                  set_prior("inv_gamma(13.8, 6.3)", class = "sd", coef="inter", group = "sub1", nlpar = "b"), 
                  set_prior("normal(0,0.3)", class = "b", coef="side",  nlpar = "a"),  
                  set_prior("inv_gamma(13.8,5.3)", class = "sd", coef="side", group = "sub1",  nlpar = "a"), 
                  set_prior("inv_gamma(13.8, 6.3)", class = "sigma"))

# Estimate side model 
side_model_fit <- brm(formula = formula3,         # Model formula 
                data = myData_dataframe,              # Data frame with variables + dependent variable 
                family = gaussian(),        # Response distribution 
                prior = priorsmodel3,       # Priors 
                # sample_prior = TRUE, need to add this for model comparison  
                warmup = 1000,              # Iterations used for warmup 
                iter = 6000,                # Total iterations per chain 
                chains = 4,                 # Number of chains 
                core = 4,                   # Cores for parallel estimation 
                control = list(adapt_delta = 0.97, max_treedepth = 15),  # Control sampler's behavior, this avoided problems of convergement
            save_pars = save_pars(all = TRUE))

```
\normalsize

Using `bayes_factor` function from the the `bridgesampling` package, we will estimate the Bayes factor of the full model against the side model.

\scriptsize 
```{r bridgesamplingexample, eval = FALSE}
bridgesampling::bayes_factor(model_fit, side_model_fit)[1]$bf  # BF of the full model against the side model 
```
\normalsize

To check the stability of the estimated Bayes factor, we could repeat this process from model estimation to model comparison using bridge sampling a few times like we did in the paper.

## Additional Resources

|    To learn more about fitting a Bayesian Hierarchical Model, we recommend the following resources:

- [Stan forum](https://discourse.mc-stan.org/c/interfaces/brms/36) 
- [Bayes factors with brms](https://vuorre.netlify.app/post/2017/03/21/bayes-factors-with-brms/ )

\newpage

# References 
\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
&nbsp;

### 2022-06-09 E - Tutorial Log-Normal Model Rstan/Online Supplement E - Bayesian Hierarchical Modeling in rstan - Log Model.Rmd
All documents except .rds

--- 2022-06-09 C - Tutorial Normal Model Rstan/Online Supplement C - Bayesian Hierarchical Modeling in rstan.Rmd
@@
 ---
-title             : "Online Supplement C: Bayesian Hierarchical Modeling in rstan"
+title             : "Online Supplement E: Bayesian Hierarchical Modeling in rstan - Log Model"
 author            : "Myrthe Veenman" 
@@
 
-|    This online supplement provides a tutorial on Bayesian Hierarchical models (BHM) in `rstan` [@R-rstan].
+|    This online supplement provides a tutorial on log-normal Bayesian Hierarchical models (BHM) in `rstan` [@R-rstan].
 In addition, it contains information about the Number Classification Task that could not be included in the paper.
@@
 
-|    To fit a BHM in *rstan*, we use the *stan* function, as shown below. 
+|    To fit a log-normal BHM in *rstan*, we use the *stan* function, as shown below. 
 &nbsp;
@@
 ```{r modelfitpreview, eval=FALSE, echo=TRUE}
-model_fit <- stan(file = "./myModel.stan",  # Stan file with model  
+model_fit <- stan(file = "./myLogModel.stan",  # Stan file with model  
                   data = myData_list,       # List with observed data and constants
@@
 
-|  The stan file `myModel.stan` is used to fit the full model and can be found under R objects > rstan > linear model.
+|  The stan file `myLogModel.stan` is used to fit the full model and can be found under R objects > rstan > linear model.
 It contains descriptions for every line of code.
@@
 ### Set the priors and add them to the list 
-a <- 0.5        # mean for mu, mu is the mean of theta  
+a <- -0.5        # mean for mu, mu is the mean of theta  
 b <- 1          # variance for mu 
 c <- 3          # mean for sigma 
-d <- 0.7        # variance for sigma, sigma is the variance of observations  
+d <- 0.3        # variance for sigma, sigma is the variance of observations  
 e <- 3          # mean for g, g is the variance of theta   
-f <- 0.7        # variance for g
+f <- 0.3        # variance for g
 
@@
 a2 <- 0        # mean for mean of beta
-b2 <- 0.09     # variance for mean of beta, used to be 1 
+b2 <- 0.005     # variance for mean of beta, used to be 1 
 e2 <- 3        # mean for variance of beta, used to be .7, adjusted 
-f2 <- 0.5      # variance for variance of beta, used to be .2., adjusted
+f2 <- 0.01      # variance for variance of beta, used to be .2., adjusted
 
@@
 a3 <- 0        # mean for mean of delta 
-b3 <- 0.09     # sd for mean of delta, 0.5, 0.3  
+b3 <- 0.005     # sd for mean of delta, 0.5, 0.3  
 dd3 <- 3       # mean for variance of delta, adjusted
-f3 <- 0.5      # variance for variance of delta, adjusted
+f3 <- 0.01      # variance for variance of delta, adjusted
 
@@
 
+### Set the priors and add them to the list 
+a <- -0.5  # mean for mu, mu is the mean of theta  
+b <- 1  # variance for mu 
+c <- 3  # mean for sigma 
+d <- .3  # variance for sigma, sigma is the variance of observations  
+e <- 3  # mean for g, g is the variance of theta   
+f <- .3  # variance for g
+
+#### Priors for beta 
+a2 <- 0  # mean for mean of beta
+b2 <- 0.005  # sd for mean of beta, used to be 1 
+e2 <- 3  # mean for sd of beta, used to be .7 
+f2 <- 0.01  # sd for sd of beta, used to be .2.
+
+#### Prior for deltas, same for every delta 
+a3 <- 0  # mean for mean of delta 
+b3 <- 0.005  # sd for mean of delta, 0.005 pretty good, 0,01 worse 
+dd3 <- 3  # mean for sd of delta 
+f3 <- 0.01  # sd for sd of delta, 0.01 good, 0.05 worse
+
 datalistpp3 <- list()
-for (i in 1:1000){
+delta1pp2 <- list()
+gammaipp2 <- list()
+betaipp2 <- list()
+sigmapp2 <- c()
+for (i in 1:1000){  # with 10000 my computer cannot handle it 
   # Hyperpriors: These are not individual specific right? So only sample 1? 
-  mugammapp <- rtruncnorm(1, a = 0, mean = 0.5, sd = sqrt(1))
+  mugammapp <- rnorm(1, -0.5, sqrt(1))
   vargammapp <- rinvgamma(1, e, f)
@@
   # data
-  Ypp <- rnorm(Ipp*Jpp*Kpp, gammaipp[exppp] + xjpp * betaipp[exppp] + 
+  Ypp <- rlnorm(Ipp*Jpp*Kpp, gammaipp[exppp] + xjpp * betaipp[exppp] + 
                   ujpp * delta1pp[exppp] + vjpp * delta2pp[exppp] +
@@
                 sqrt(sigmapp))
+  gammaipp2[[i]] <- gammaipp
+  betaipp2[[i]] <- betaipp
+  delta1pp2[[i]] <- delta1pp
+  sigmapp2[i] <- sigmapp
   datartpp <- data.frame(exppp, condpp, subpp, repetition = rep(i, length(exppp)), gammaipp[exppp], betaipp[exppp], delta1pp[exppp], delta2pp[exppp], delta3pp[exppp], delta4pp[exppp], Ypp)
@@
 # Add datasets under each other 
-bigdatapp3_rstan <- do.call(rbind, datalistpp3)
+bigdatapp3 <- do.call(rbind, datalistpp3)
 
@@
 # Create figure 
-priorpredictionplotnew_2 <- priorpredfigurecreate(bigdatapp3_rstan)
+priorpredictionplotnew_2 <- priorpredfigurecreate(bigdatapp3)
 
@@
 library(rstan) 
-model_fit <- stan(file = "myPath/rewritestanmodel3_julia.stan", 
+model_fit <- stan(file = "myPath/myLogModel.stan", 
                   data = myData_list, 
@@
 
-We have already fitted the model and saved it in the R object `model_fit`.
+We have already fitted the model and saved it in the R object `logmodel_fit`.
 In case you do not want to wait until the model has been fitted but continue immediately with the rest of this tutorial, you can load this R object that contains the model estimates.
@@
 ```{r loadmodelfit, eval = FALSE}
-model_fit <- readRDS(file = "myPath/rstan_model_fit.rds")
+logmodel_fit <- readRDS(file = "myPath/rstan_logmodel_fit.rds")
 ```
@@
 setwd('..')
-model_fit <- readRDS(file = "R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
+logmodel_fit <- readRDS(file = "R objects/rstan/log-normal model/hier_logmodelc_adj13012022.rds")
 ```
@@
 
-|    *Brms* offers the *launch_shinystan* function which will load a shiny app with the results of the model fit.
+|    `Brms` offers the `launch_shinystan` function which will load a shiny app with the results of the model fit.
 It is very extensive.
@@
 ```{r shinyrstan, eval = FALSE}
-launch_shinystan(model_fit)
+launch_shinystan(logmodel_fit)
 ```
@@
 
-|    The *summary* function provides an overview of the model fit.
+|    The `summary` function provides an overview of the model fit.
 From this function, we can get a lot of information.
@@
 |    The trace plot shows whether the posterior distribution of the parameter has converged.
-With the package's *plot* function it is possible to obtain the trace plots per parameter by setting the option *plotfun* to *trace*.
+With the package's `plot` function it is possible to obtain the trace plots per parameter by setting the option `plotfun` to `trace`.
 By default, it will show the trace plots of the first ten parameters.
-With the *pars* function you can specify for which parameters you want to inspect the trace plots.
-The trace plot will automatically remove the warm-up (burn-in) period from the chain (which is desirable), however, if you would like to see that as well you have to set the option *inc_warmup* to *TRUE*.
-As the *plot* option returns a *ggplot* object, it is possible to manually adjust the plot using *ggplot* functions.
-We created the function *traceplotfunc* to obtain the trace plot for the parameter you are interested in.
+With the `pars` function you can specify for which parameters you want to inspect the trace plots.
+The trace plot will automatically remove the warm-up (burn-in) period from the chain (which is desirable), however, if you would like to see that as well you have to set the option `inc_warmup` to `TRUE`.
+As the `plot` option returns a `ggplot` object, it is possible to manually adjust the plot using `ggplot` functions.
+We created the function `traceplotfunc` to obtain the trace plot for the parameter you are interested in.
 In the figure below, we show the trace plots for all the general effects.
@@
 # Trace plot for parameter mu4
-trdif2rstan <- plot(model_fit,     # Model fit 
+trdif2rstan <- plot(logmodel_fit,     # Model fit 
                     plotfun = "trace",   # Specify that you want trace plots 
@@
 # In case you want to see the warmup period as well 
-trdif2rstanwarmup <- plot(model_fit,    # Model fit 
+trdif2rstanwarmup <- plot(logmodel_fit,    # Model fit 
                           plotfun = "trace",  # Specify that you want trace plots 
@@
 # Plot frequency of rhat 
-plot(model_fit, plotfun= "rhat")
+plot(logmodel_fit, plotfun= "rhat")
 
 # Plot frequency of the ratio 
-plot(model_fit, plotfun = "ess")
+plot(logmodel_fit, plotfun = "ess")
 ```
@@
 # Obtain rhat for every parameter from model fit 
-rhatrstan <- summary(model_fit)$summary[,"Rhat"]
+rhatrstan <- summary(logmodel_fit)$summary[,"Rhat"]
 
 # Obtain number of effective samples for every paramter from model fit 
-neffrstan <- summary(model_fit)$summary[,"n_eff"]
+neffrstan <- summary(logmodel_fit)$summary[,"n_eff"]
 
@@
 # Obtain rhat for every parameter from model fit 
-rhatrstan <- summary(model_fit)$summary[,"Rhat"]
+rhatrstan <- summary(logmodel_fit)$summary[,"Rhat"]
 fixedrhatrstan <- rhatrstan[1:13]  # for general effects 
@@
 # Obtain number of effective samples for every parameter from model fit 
-neffrstan <- summary(model_fit)$summary[,"n_eff"]
+neffrstan <- summary(logmodel_fit)$summary[,"n_eff"]
 fixedneffrstan <- neffrstan[1:13]  # for general effects 
@@
 
-|    With the *plot* function, the posterior distribution per parameter can be displayed. 
+|    With the `plot` function, the posterior distribution per parameter can be displayed. 
 &nbsp;
@@
 # Seperate plot per parameter of posterior density 
-plot(model_fit, 
+plot(logmodel_fit, 
      plotfun = "dens")  
@@
 
-|    Another possibility is to use the *bayesplot* package to plot all the posterior densities together in one plot. This makes it easier to compare the posterior distributions of each parameter. 
+|    Another possibility is to use the `bayesplot` package to plot all the posterior densities together in one plot. This makes it easier to compare the posterior distributions of each parameter. 
 &nbsp;
@@
 The shaded area within the distributions represents 95\\% of the probability mass.", fig.align = "left", fig.height = 4, fig.width = 5}
-postc <- as.array(model_fit)  # change structure of model fit
+postc <- as.array(logmodel_fit)  # change structure of model fit
 
@@
 # Obtain the mean, standard error, 95% credible interval, rhat and number of effective samples for the intercept, and the digit and side parameters 
-fitrstanhiertable <- summary(model_fit,  # model fit 
+fitrstanhiertable <- summary(logmodel_fit,  # model fit 
                              # Select parameters 
@@
 
-|    First, we can show a table with the posterior distribution estimates of the variance parameters. This indicates the variation of the general effect. We called our variances *g* (*g* until *g6*). 
+|    First, we can show a table with the posterior distribution estimates of the variance parameters. This indicates the variation of the general effect. We called our variances `g` (`g` until `g6`). 
 &nbsp;
@@
 ```{r inddevtable, results = "asis", cache = TRUE}
-fitrstanhiertable2 <- summary(model_fit,   # model fit 
+fitrstanhiertable2 <- summary(logmodel_fit,   # model fit 
                               # Select variance parameters (we called them g)
@@
 # Extract the parameter estimates using the names from before 
-gamstan <- as.data.frame(summary(model_fit, 
+gamstan <- as.data.frame(summary(logmodel_fit, 
                                  pars = c( gamnamesstan))$summary)
@@
 # Extract the general effect estimate of the parameter 
-gamstanmu <- as.data.frame(summary(model_fit, 
+gamstanmu <- as.data.frame(summary(logmodel_fit, 
                                    pars = c("mu"))$summary)
@@
 betnamesstan <- paste("betai", "[",1:52,"]", sep="")
-betstan <- as.data.frame(summary(model_fit, pars = c( betnamesstan))$summary)
-betstanmu <- as.data.frame(summary(model_fit, pars = c("mu2"))$summary)
+betstan <- as.data.frame(summary(logmodel_fit, pars = c( betnamesstan))$summary)
+betstanmu <- as.data.frame(summary(logmodel_fit, pars = c("mu2"))$summary)
 betstan$parameters <- c(betnamesstan)
@@
 del1namesstan <- paste("delta1", "[",1:52,"]", sep="")
-delt1stan <- as.data.frame(summary(model_fit, pars = c( del1namesstan))$summary)
-del1stanmu <- as.data.frame(summary(model_fit, pars = c("mu3"))$summary)
+delt1stan <- as.data.frame(summary(logmodel_fit, pars = c( del1namesstan))$summary)
+del1stanmu <- as.data.frame(summary(logmodel_fit, pars = c("mu3"))$summary)
 delt1stan$parameters <- c(del1namesstan)
@@
 del2namesstan <- paste("delta2", "[",1:52,"]", sep="")
-delt2stan <- as.data.frame(summary(model_fit, pars = c( del2namesstan))$summary)
-del2stanmu <- as.data.frame(summary(model_fit, pars = c("mu4"))$summary)
+delt2stan <- as.data.frame(summary(logmodel_fit, pars = c( del2namesstan))$summary)
+del2stanmu <- as.data.frame(summary(logmodel_fit, pars = c("mu4"))$summary)
 delt2stan$parameters <- c(del2namesstan)
@@
 del3namesstan <- paste("delta3", "[",1:52,"]", sep="")
-delt3stan <- as.data.frame(summary(model_fit, pars = c( del3namesstan))$summary)
-del3stanmu <- as.data.frame(summary(model_fit, pars = c("mu5"))$summary)
+delt3stan <- as.data.frame(summary(logmodel_fit, pars = c( del3namesstan))$summary)
+del3stanmu <- as.data.frame(summary(logmodel_fit, pars = c("mu5"))$summary)
 delt3stan$parameters <- c(del3namesstan)
@@
 del4namesstan <- paste("delta4", "[",1:52,"]", sep="")
-delt4stan <- as.data.frame(summary(model_fit, pars = c( del4namesstan))$summary)
-del4stanmu <- as.data.frame(summary(model_fit, pars = c("mu6"))$summary)
+delt4stan <- as.data.frame(summary(logmodel_fit, pars = c( del4namesstan))$summary)
+del4stanmu <- as.data.frame(summary(logmodel_fit, pars = c("mu6"))$summary)
 delt4stan$parameters <- c(del4namesstan)
@@
 ### Obtain all individual estimates delta 
-hier_modeld <- model_fit  # model fit
+hier_modeld <- logmodel_fit  # model fit
 
@@
 This can be done with the Bayes factor.
-In *rstan* the Bayes factor can be obtained by using the *bridgesampling* package.
-To be able to use this package, it is important that the *.stan* file is written in a way that it saves the constants (i.e., `target += normal_lpdf(mu4 | a3, sqrt(b3));`).
+In `rstan` the Bayes factor can be obtained by using the `bridgesampling` package.
+To be able to use this package, it is important that the `.stan` file is written in a way that it saves the constants (i.e., `target += normal_lpdf(mu4 | a3, sqrt(b3));`).
 |    First, we fit all the models (except for the full model that we already fitted. 
@@
 ## Null model 
-BFnullmodel <- stan(file = "myPath/rewritestanmodel3h0.stan", 
+BFnulllogmodel <- stan(file = "myPath/logstanmodelh0.stan", 
                     data = myData_list, 
@@
 ## Side model 
-BFsidemodel <- stan(file = "myPath/rewritestanmodel3hside.stan", 
+BFsidelogmodel <- stan(file = "myPath/logstanmodelhside.stan", 
                        data = myData_list, 
@@
 ## Digit model 
-BFdigitmodel <- stan(file = "myPath/rewritestanmodel3hdigit.stan", 
+BFdigitlogmodel <- stan(file = "myPath/logstanmodelhdigit.stan", 
                          data = myData_list, 
@@
 # library(bridgesampling)
-bridge_Hnull <- bridge_sampler(BFnullmodel)
-bridge_Hside <- bridge_sampler(BFsidemodel)
-bridge_Hdigit <- bridge_sampler(BFdigitmodel)
-bridge_Hfull <- bridge_sampler(model_fit)  # in case you just fitted the model yourself
+bridge_Hnulllog <- bridge_sampler(BFnulllogmodel)
+bridge_Hsidelog <- bridge_sampler(BFsidelogmodel)
+bridge_Hdigitlog <- bridge_sampler(BFdigitlogmodel)
+bridge_Hfulllog <- bridge_sampler(logmodel_fit)  # in case you just fitted the model yourself
 
@@
 ### Fit model again but do not run any chains 
-new_mod <- stan(file = "myPath/rewritestanmodel3_julia.stan", 
+new_logmod <- stan(file = "myPath/myLogModel.stan", 
                 data = myData_list, 
@@
 ### Now do the bridgesampling 
-bridge_Hfull <- bridge_sampler(model_fit, new_mod)
+bridge_Hfulllog <- bridge_sampler(logmodel_fit, new_logmod)
 ```
@@
 # Obtain BF (in this way evidence in favor of the null hypothesis)
-bridgesampling::bf(bridge_Hnull, bridge_Hfull)   # Null against full
+bridgesampling::bf(bridge_Hnulllog, bridge_Hfulllog)   # Null against full
 
 # Obtain BF (in this way evidence in favor of the full hypothesis)
-bridgesampling::bf(bridge_Hfull, bridge_Hnull)   # Full against null
-bridgesampling::bf(bridge_Hfull, bridge_Hside)   # Full against side 
-bridgesampling::bf(bridge_Hfull, bridge_Hdigit)  # Full against digit 
+bridgesampling::bf(bridge_Hfulllog, bridge_Hnulllog)   # Full against null
+bridgesampling::bf(bridge_Hfulllog, bridge_Hsidelog)   # Full against side 
+bridgesampling::bf(bridge_Hfulllog, bridge_Hdigitlog)  # Full against digit 
 ```

### 2022-06-09 F - Tutorial Log-Normal Model brms/Online Supplement F - Bayesian Hierarchical Modeling in brms - Log Model.Rmd
All documents except .rds

--- 2022-06-09 D - Tutorial Normal Model brms/Online Supplement D - Bayesian Hierarchical Modeling in brms.Rmd
@@
 ---
-title             : "Online Supplement D: Bayesian Hierarchical Modeling in brms"
+title             : "Online Supplement F: Bayesian Hierarchical Modeling in brms - Log Model"
 author            : "Myrthe Veenman"
@@
 
-|    This online supplement provides a tutorial on Bayesian Hierarchical models (BHM) in `brms` [@R-brms_a; @R-brms_b].
+|    This online supplement provides a tutorial on log-normal Bayesian Hierarchical models (BHM) in `brms` [@R-brms_a; @R-brms_b].
 In addition, it contains information about the Number Classification Task that could not be included in the paper.
@@
 
-|    To fit a BHM in *brms*, we use the *brm* function, as shown below. 
+|    To fit a log-normal BHM in *brms*, we use the *brm* function, as shown below. 
 &nbsp;
@@
                  data = myData_dataframe,                # Dataset (parameters, observations) 
-                 family = gaussian(),                    # Family, distribution dependent variable 
+                 family = lognormal(),                    # Family, distribution dependent variable 
                  prior = priorsmodel,                    # Specification of the priors 
@@
 
-However, for the digit classification task, we specified the formula in a slightly different way that allowed us to use a different prior on the intercept compared to the distance and side effects.
-In the first part, we specify that the dependent variable (in our case response times) are predicted by `a` and `b`.
-In the second part, we define `a`.
-It contains the general and individual distance and side effects.
-Also note that we replaced `1` by `0`, indicating that we do not want the default intercept.
-In the third and final part, we specify `b`, representing the general and individual intercepts by the variable `inter`.
-
-\scriptsize
-```{r formulafit2, eval = FALSE}
-formula <- brms::bf(rt2 ~ a + b, nl = TRUE,  # part 1 
-                  lf(a ~ 0 + side + dif1 + dif2 + dif3 + dif4 + (0 + side + dif1 + dif2 + dif3 + dif4 || ind), center = TRUE),   # part 2 
-                  lf(b ~ 0 + inter + (0 + inter || ind), cmc = FALSE)  # part 3 
-)
-```
-\normalsize
-
 ### Family
@@
 |    The argument `family` in the `brm` function specifies the exponential family of distributions according to which the dependent variable is assumed to be distributed.
-For the normal model, this argument takes the value `gaussian`.
+For the log-normal model, this argument takes the value `lognormal`.
 
@@
 For example, `brms` only allows setting priors on standard deviations and not on variances.
-Therefore, for the normal model of the symbolic distance effect, we will use the priors for standard deviations specified in Equation 7.
+Therefore, for the normal model of the symbolic distance effect, we will use the priors for standard deviations specified in Equation 9.
 To set the priors, we use the `set_prior` function, as shown below.
@@
 ```{r priorsillusbrms2, eval = FALSE, echo = TRUE}
-set_prior("normal(0,0.3)", class = "b", coef="side")  # general effect 
-set_prior("inv_gamma(13.8, 5.3)", class = "sd",       # individual variation 
+set_prior("normal(-0.5, 1)", class = "b", coef="side")  # general effect 
+set_prior("inv_gamma(13.8, 4.1)", class = "sd",       # individual variation 
           coef="side", group = "ind") 
@@
 
-For the digit classification task, we want to specify a truncated normal prior on the intercept, because response times cannot be negative.
-This means that we are setting a different prior on the intercept than on the side and digit effects.
-To accomplish this, we split up the formula in two different parts.
-For each variable, we specify to which part of the formula they belong with `nlpar`.
-In addition, for the intercept, we specify that the prior has a lower bound of zero with `lb`.
 All priors are saved in an R `data.frame` that is used in the `brm` function.
@@
 ```{r priors}
-priorsmodel <- c(set_prior("normal(0.5,1)", class = "b", lb = 0, nlpar = "b"),  
-                  set_prior("inv_gamma(13.8, 6.3)", class = "sd", coef="inter", group = "ind", nlpar = "b"), 
-                  set_prior("normal(0,0.3)", class = "b", coef="side",  nlpar = "a"),  
-                  set_prior("inv_gamma(13.8,5.3)", class = "sd", coef="side", group = "ind",  nlpar = "a"), 
-                  set_prior("normal(0,0.3)", class = "b", coef="dif1",  nlpar = "a"), 
-                  set_prior("inv_gamma(13.8,5.3)", class = "sd", coef="dif1", group = "ind",  nlpar = "a"),
-                  set_prior("normal(0,0.3)", class = "b", coef="dif2",  nlpar = "a"),  
-                  set_prior("inv_gamma(13.8,5.3)", class = "sd", coef="dif2", group = "ind",  nlpar = "a"),
-                  set_prior("normal(0,0.3)", class = "b", coef="dif3",  nlpar = "a"), 
-                  set_prior("inv_gamma(13.8,5.3)", class = "sd", coef="dif3", group = "ind",  nlpar = "a"),
-                  set_prior("normal(0,0.3)", class = "b", coef="dif4",  nlpar = "a"),  
-                  set_prior("inv_gamma(13.8,5.3)", class = "sd", coef="dif4", group = "ind",  nlpar = "a"),
-                  set_prior("inv_gamma(13.8, 6.3)", class = "sigma"))
+priorslogmodel <- c(set_prior("normal(-0.5, 1)", class = "Intercept"),  
+                  set_prior("inv_gamma(13.8, 4.1)", class = "sd", coef="Intercept", group = "sub1"), 
+                  set_prior("normal(0, 0.07)", class = "b", coef="side"),  
+                  set_prior("inv_gamma(13.8, 0.7)", class = "sd", coef="side", group = "sub1"), 
+                  set_prior("normal(0, 0.07)", class = "b", coef="dif1"), 
+                  set_prior("inv_gamma(13.8, 0.7)", class = "sd", coef="dif1", group = "sub1"),
+                  set_prior("normal(0, 0.07)", class = "b", coef="dif2"),  
+                  set_prior("inv_gamma(13.8, 0.7)", class = "sd", coef="dif2", group = "sub1"),
+                  set_prior("normal(0, 0.07)", class = "b", coef="dif3"), 
+                  set_prior("inv_gamma(13.8, 0.7)", class = "sd", coef="dif3", group = "sub1"),
+                  set_prior("normal(0, 0.07)", class = "b", coef="dif4"),  
+                  set_prior("inv_gamma(13.8, 0.7)", class = "sd", coef="dif4", group = "sub1"),
+                  set_prior("inv_gamma(13.8, 4.1)", class = "sigma"))
 
@@
 datalistpp3 <- list()
-for (i in 1:1000){
-# Hyperpriors: These are not individual specific so only sample 1  
-mugammapp <- rtruncnorm(1, a = 0, mean = 0.5, sd = sqrt(1))
-vargammapp <- rinvgamma(1, 3, 0.7)
-mubetapp <- rnorm(1, 0, 0.3)
-varbetapp <- rinvgamma(1, 3, 0.5)
-mudeltapp <- rnorm(1, 0, 0.3)
-vardeltapp <- rinvgamma(1, 3, 0.5)
+delta1pp2 <- list()
+gammaipp2 <- list()
+betaipp2 <- list()
+sigmapp2 <- c()
+for (i in 1:1000){  # with 10000 my computer cannot handle it 
+  # Hyperpriors: These are not individual specific right? So only sample 1? 
+  mugammapp <- rnorm(1, -0.5, sqrt(1))
+  vargammapp <- rinvgamma(1, 13.8, 4.1)
+  mubetapp <- rnorm(1, 0, 0.07)
+  varbetapp <- rinvgamma(1, 13.8, 0.7)
+  mudeltapp <- rnorm(1, 0, 0.07)
+  vardeltapp <- rinvgamma(1, 13.8, 0.7)
 
@@
 
-sigmapp <- rinvgamma(1, 3, 0.7)  # Only 1 because not individual specific
-
+  sigmapp <- rinvgamma(1, 13.8, 4.1)  # Only 1 because not individual specific?
+
+  # Or start for loop here? Or do we want to vary persons as well? 
   # data
-Ypp <- rnorm(Ipp*Jpp*Kpp, gammaipp[exppp] + xjpp * betaipp[exppp] + 
+  Ypp <- rlnorm(Ipp*Jpp*Kpp, gammaipp[exppp] + xjpp * betaipp[exppp] + 
                   ujpp * delta1pp[exppp] + vjpp * delta2pp[exppp] +
                   wjpp * delta3pp[exppp] + zjpp * delta4pp[exppp],
-              sqrt(sigmapp))  # model formula 
-datartpp <- data.frame(exppp,  # data frame of results and conditions 
-                       condpp, 
-                       subpp, 
-                       repetition = rep(i, length(exppp)), 
-                       gammaipp[exppp], betaipp[exppp], delta1pp[exppp],
-                       delta2pp[exppp], delta3pp[exppp], delta4pp[exppp], Ypp)
+                sqrt(sigmapp))
+  gammaipp2[[i]] <- gammaipp
+  betaipp2[[i]] <- betaipp
+  delta1pp2[[i]] <- delta1pp
+  sigmapp2[i] <- sigmapp
+  datartpp <- data.frame(exppp, condpp, subpp, repetition = rep(i, length(exppp)), gammaipp[exppp], betaipp[exppp], delta1pp[exppp], delta2pp[exppp], delta3pp[exppp], delta4pp[exppp], Ypp)
   datalistpp3[[i]] <- datartpp
@@
 bigdatapp3 <- do.call(rbind, datalistpp3)
-
-newbigdatapp3 <- bigdatapp3[bigdatapp3$exppp == 1,]
 
@@
 ``` {r modelfit, eval = FALSE}
-model_fit <- brm(formula = formula,                      # Model formula 
+logmodel_fit <- brm(formula = logformula,                      # Model formula 
                  data = myData_dataframe,                # Dataset (parameters, observations) 
-                 family = gaussian(),                    # Family, distribution dependent variable 
-                 prior = priorsmodel,                    # Specification of the priors 
+                 family = lognormal(),                    # Family, distribution dependent variable 
+                 prior = priorslogmodel,                    # Specification of the priors 
                  # sample_prior = "yes",                 # To calculate Bayes factors for point hypotheses  
@@
 
-We have already estimated the model and saved it in the R object `brmsmodelfit_final`.
+We have already estimated the model and saved it in the R object `brmslogmodelfit_final`.
 In case you do not want to wait until the model has been fitted but continue immediately with the rest of this tutorial, you can load this R object that contains the model estimations.
@@
 ```{r loadmodelfit, eval = FALSE}
-model_fit <- readRDS(file = "myPath/brmsmodelfit_final.rds")
+logmodel_fit <- readRDS(file = "myPath/brmslogmodelfit_final.rds")
 
 # For model comparison - sample_prior was set to "yes" 
-model_fit_bf <- readRDS("myPath/brmsmodelfit_BF_final.rds")
+logmodel_fit_bf <- readRDS("myPath/brmslogmodelfit_BF_final.rds")
 
@@
 ```{r loadmodelfits, show = FALSE}
-model_fit <- readRDS(file = "/Users/myrtheveenman/Documents/GitHub/multilevel-tutorial-paper/R objects/brms/Truncated/BRMSmodelfit18_03012022.rds")
+logmodel_fit <- readRDS(file = "/Users/myrtheveenman/Documents/GitHub/multilevel-tutorial-paper/R objects/brms/non-linear/brmmodel8log13012022.rds")
 
 # For model comparison - sample_prior was set to "yes" 
-model_fit_bf <- readRDS("/Users/myrtheveenman/Documents/GitHub/multilevel-tutorial-paper/R objects/brms/Truncated/BRMSmodelfit18_BF_03012022.rds")
+logmodel_fit_bf <- readRDS("/Users/myrtheveenman/Documents/GitHub/multilevel-tutorial-paper/R objects/brms/BF/brmmodel8log26052022_bf.rds")
 
@@
 ```{r modelsummary}
-summary(model_fit)
+summary(logmodel_fit)
 ```
@@
 ```{r checkvarnames, eval = FALSE}
-variables(model_fit)
+variables(logmodel_fit)
 ```
@@
 # I think warmup already deleted, but then have to adjust labels x axis 
-trdif2brms <- mcmc_plot(model_fit,       # Model fit 
-                       variable = "b_a_dif2",  # Select parameter, 
+trdif2brms <- mcmc_plot(logmodel_fit,       # Model fit 
+                       variable = "b_dif2",  # Select parameter, 
                        # it is also possible to select more than one at the same time, 
@@
 traceplotfunc <- function(parameter, title){
-trdif2brms <- mcmc_plot(model_fit,       # Model fit 
+trdif2brms <- mcmc_plot(logmodel_fit,       # Model fit 
                        variable = parameter,  # Select parameter, 
@@
 # Show trace plot for one digit parameter 
-tr2 <- traceplotfunc(parameter = "b_a_side", title = expression(~mu[~beta]))
+tr2 <- traceplotfunc(parameter = "b_side", title = expression(~mu[~beta]))
 tr2 <- tr2 + theme(legend.position = "none")
-tr3 <- traceplotfunc(parameter = "b_a_dif1", title = expression(~mu[~delta[~7]]))
+tr3 <- traceplotfunc(parameter = "b_dif1", title = expression(~mu[~delta[~7]]))
 tr3 <- tr3 + theme(legend.position = "none")
-tr4 <- traceplotfunc(parameter = "b_a_dif2", title = expression(~mu[~delta[~6]]))
+tr4 <- traceplotfunc(parameter = "b_dif2", title = expression(~mu[~delta[~6]]))
 tr4 <- tr4 + theme(legend.position = "none")
-tr5 <- traceplotfunc(parameter = "b_a_dif3", title = expression(~mu[~delta[~4]]))
+tr5 <- traceplotfunc(parameter = "b_dif3", title = expression(~mu[~delta[~4]]))
 tr5 <- tr5 + theme(legend.position = "none")
-tr6 <- traceplotfunc(parameter = "b_a_dif4", title = expression(~mu[~delta[~3]]))
+tr6 <- traceplotfunc(parameter = "b_dif4", title = expression(~mu[~delta[~3]]))
 legendtr <- cowplot::get_legend(tr6)
@@
 # Show trace plot for one sd parameter 
-tr2 <- traceplotfunc(parameter = "sd_sub1__a_side", title = expression(~sigma[~beta]))
+tr2 <- traceplotfunc(parameter = "sd_sub1__side", title = expression(~sigma[~beta]))
 tr2 <- tr2 + theme(legend.position = "none")
-tr3 <- traceplotfunc(parameter = "sd_sub1__a_dif1", title = expression(~sigma[~delta[~7]]))
+tr3 <- traceplotfunc(parameter = "sd_sub1__dif1", title = expression(~sigma[~delta[~7]]))
 tr3 <- tr3 + theme(legend.position = "none")
-tr4 <- traceplotfunc(parameter = "sd_sub1__a_dif2", title = expression(~sigma[~delta[~6]]))
+tr4 <- traceplotfunc(parameter = "sd_sub1__dif2", title = expression(~sigma[~delta[~6]]))
 tr4 <- tr4 + theme(legend.position = "none")
-tr5 <- traceplotfunc(parameter = "sd_sub1__a_dif3", title = expression(~sigma[~delta[~4]]))
+tr5 <- traceplotfunc(parameter = "sd_sub1__dif3", title = expression(~sigma[~delta[~4]]))
 tr5 <- tr5 + theme(legend.position = "none")
-tr6 <- traceplotfunc(parameter = "sd_sub1__a_dif4", title = expression(~sigma[~delta[~3]]))
+tr6 <- traceplotfunc(parameter = "sd_sub1__dif4", title = expression(~sigma[~delta[~3]]))
 legendtr <- cowplot::get_legend(tr6)
@@
 # Plot frequency of rhat 
-mcmc_plot(model_fit, type = "rhat_hist")   # Specify that you want trace plots
+mcmc_plot(logmodel_fit, type = "rhat_hist")   # Specify that you want trace plots
 
 # Plot frequency of the ratio 
-mcmc_plot(model_fit, type = "neff_hist") 
+mcmc_plot(logmodel_fit, type = "neff_hist") 
 ```
@@
 ## Rhat 
-rhatbrms <- mcmc_plot(model_fit, type = "rhat_hist")$data$value  # save rhat values 
-rhatbrms_para <- mcmc_plot(model_fit, type = "rhat_hist")$data$parameter  # save parameter names 
+rhatbrms <- mcmc_plot(logmodel_fit, type = "rhat_hist")$data$value  # save rhat values 
+rhatbrms_para <- mcmc_plot(logmodel_fit, type = "rhat_hist")$data$parameter  # save parameter names 
 rhatbrmsratio_comb <- data.frame(parameter = rhatbrms_para, value = rhatbrms)  # combine values and parameters names into one data.frame 
@@
 # Obtain rhats of general effects 
-rhatbrms_v2 <- mcmc_plot(model_fit, type = "rhat_hist")$data
-fixedrhatbrms_v2 <- c(rhatbrms_v2$value[rhatbrms_v2$parameter == "b_b_inter"], 
-                      rhatbrms_v2$value[rhatbrms_v2$parameter == "b_a_side"], 
-                      rhatbrms_v2$value[rhatbrms_v2$parameter == "b_a_dif1"], 
-                      rhatbrms_v2$value[rhatbrms_v2$parameter == "b_a_dif2"], 
-                      rhatbrms_v2$value[rhatbrms_v2$parameter == "b_a_dif3"], 
-                      rhatbrms_v2$value[rhatbrms_v2$parameter == "b_a_dif4"])
+rhatbrms_v2 <- mcmc_plot(logmodel_fit, type = "rhat_hist")$data
+fixedrhatbrms_v2 <- c(rhatbrms_v2$value[rhatbrms_v2$parameter == "b_Intercept"], 
+                      rhatbrms_v2$value[rhatbrms_v2$parameter == "b_side"], 
+                      rhatbrms_v2$value[rhatbrms_v2$parameter == "b_dif1"], 
+                      rhatbrms_v2$value[rhatbrms_v2$parameter == "b_dif2"], 
+                      rhatbrms_v2$value[rhatbrms_v2$parameter == "b_dif3"], 
+                      rhatbrms_v2$value[rhatbrms_v2$parameter == "b_dif4"])
 
@@
 # Obtain rhats of all individual effects 
-rhatbrmsratio3 <- c(rhatbrmsratio_comb[grep(",inter]", rhatbrmsratio_comb$parameter),]$value,
+rhatbrmsratio3 <- c(rhatbrmsratio_comb[grep(",Intercept]", rhatbrmsratio_comb$parameter),]$value,
                       rhatbrmsratio_comb[grep(",side]", rhatbrmsratio_comb$parameter),]$value,
@@
 ```{r neffexamp5}
-neffbrmsratio_value <- mcmc_plot(model_fit, type = "neff_hist")$data$value   # save neff values, these are the values of the ration Neff/N, N representing the total number of iterations
-neffbrmsratio_para <- mcmc_plot(model_fit, type = "neff_hist")$data$parameter  # save parameter names 
+neffbrmsratio_value <- mcmc_plot(logmodel_fit, type = "neff_hist")$data$value   # save neff values, these are the values of the ration Neff/N, N representing the total number of iterations
+neffbrmsratio_para <- mcmc_plot(logmodel_fit, type = "neff_hist")$data$parameter  # save parameter names 
 neffbrmsratio_comb <- data.frame(parameter = neffbrmsratio_para, value = neffbrmsratio_value)  # combine values and parameter names into data frame
@@
 # Neff of general effects 
-fixedneffbrms2 <- c(neffbrmsratio_comb[grep("b_inter", neffbrmsratio_comb$parameter),]$value[1],
-                      neffbrmsratio_comb[grep("b_a_side", neffbrmsratio_comb$parameter),]$value[1],
-                      neffbrmsratio_comb[grep("b_a_dif1", neffbrmsratio_comb$parameter),]$value[1],
-                      neffbrmsratio_comb[grep("b_a_dif2", neffbrmsratio_comb$parameter),]$value[1],
-                      neffbrmsratio_comb[grep("b_a_dif3", neffbrmsratio_comb$parameter),]$value[1],
-                      neffbrmsratio_comb[grep("b_a_dif4", neffbrmsratio_comb$parameter),]$value[1])
+fixedneffbrms2 <- c(neffbrmsratio_comb[grep("b_Intercept", neffbrmsratio_comb$parameter),]$value[1],
+                      neffbrmsratio_comb[grep("b_side", neffbrmsratio_comb$parameter),]$value[1],
+                      neffbrmsratio_comb[grep("b_dif1", neffbrmsratio_comb$parameter),]$value[1],
+                      neffbrmsratio_comb[grep("b_dif2", neffbrmsratio_comb$parameter),]$value[1],
+                      neffbrmsratio_comb[grep("b_dif3", neffbrmsratio_comb$parameter),]$value[1],
+                      neffbrmsratio_comb[grep("b_dif4", neffbrmsratio_comb$parameter),]$value[1])
 
@@
 # Neff of individual effects  
-neffbrmsratio3 <- c(neffbrmsratio_comb[grep(",inter]", neffbrmsratio_comb$parameter),]$value,
+neffbrmsratio3 <- c(neffbrmsratio_comb[grep(",Intercept]", neffbrmsratio_comb$parameter),]$value,
                       neffbrmsratio_comb[grep(",side]", neffbrmsratio_comb$parameter),]$value,
@@
 # Seperate plot per parameter of posterior density 
-mcmc_plot(model_fit, 
+mcmc_plot(logmodel_fit, 
      type = "dens")  
@@
 The shaded area within the distributions represents 95\\% of the probability mass.", fig.align = "left", fig.height = 4, fig.width = 5}
-post8 <- as.array(model_fit)  # change structure of model fit
+post8 <- as.array(logmodel_fit)  # change structure of model fit
 
@@
   post8,    # Structured model fit 
-  pars = c("b_a_side", "b_a_dif1", "b_a_dif2", "b_a_dif3", "b_a_dif4"),  # Select parameters 
+  pars = c("b_side", "b_dif1", "b_dif2", "b_dif3", "b_dif4"),  # Select parameters 
   prob = 0.95, # 80% intervals  # Plot 95% credible intervals
@@
 # Obtain the mean, standard error, 95% credible interval, rhat and number of effective samples for the intercept, and the digit and side parameters 
-fitbrmshiertable <- summary(model_fit)[[14]][,c("Estimate", "Est.Error", "l-95% CI", "u-95% CI", "Bulk_ESS", "Rhat")]  # variance
+fitbrmshiertable <- summary(logmodel_fit)[[14]][,c("Estimate", "Est.Error", "l-95% CI", "u-95% CI", "Bulk_ESS", "Rhat")]  # variance
 
@@
 fitbrmshiertable2 <- as.data.frame(fitbrmshiertable)
-fitbrmshiertable2 <- rbind(fitbrmshiertable2[6,], fitbrmshiertable2[1:5,])
 
@@
 # Obtain standard deviation estimates 
-fitbrmshiertable2 <- summary(model_fit)[[17]]$sub1[,c("Estimate", "Est.Error", "l-95% CI", "u-95% CI", "Bulk_ESS", "Rhat")]
+fitbrmshiertable2 <- summary(logmodel_fit)[[17]]$sub1[,c("Estimate", "Est.Error", "l-95% CI", "u-95% CI", "Bulk_ESS", "Rhat")]
 
@@
 fitbrmshiertable22 <- as.data.frame(fitbrmshiertable2)
-fitbrmshiertable22 <- rbind(fitbrmshiertable22[6,], fitbrmshiertable22[1:5,])
 
@@
 # Extract the parameter estimates 
-brmsgammar <- coef(model_fit)$sub1[,,"b_inter"][,c("Estimate", "Q2.5", "Q97.5")]
+brmsgammar <- coef(logmodel_fit)$sub1[,,"Intercept"][,c("Estimate", "Q2.5", "Q97.5")]
 brmsgammar <- as.data.frame(brmsgammar)
@@
 # Extract the general effect estimate of the parameter 
-fixbremsgammar <- fixef(model_fit)["b_inter","Estimate"]
+fixbremsgammar <- fixef(logmodel_fit)["Intercept","Estimate"]
 
@@
 # Gamma 
-brmsgammar <- coef(model_fit)$sub1[,,"b_inter"][,c("Estimate", "Q2.5", "Q97.5")]
+brmsgammar <- coef(logmodel_fit)$sub1[,,"Intercept"][,c("Estimate", "Q2.5", "Q97.5")]
 brmsgammar <- as.data.frame(brmsgammar)
-fixbremsgammar <- fixef(model_fit)["b_inter","Estimate"]
+fixbremsgammar <- fixef(logmodel_fit)["Intercept","Estimate"]
 
@@
 # Beta 
-brmsbetar <- coef(model_fit)$sub1[,,"a_side"][,c("Estimate", "Q2.5", "Q97.5")]
+brmsbetar <- coef(logmodel_fit)$sub1[,,"side"][,c("Estimate", "Q2.5", "Q97.5")]
 brmsbetar <- as.data.frame(brmsbetar)
-fixbremsbetar <- fixef(model_fit)["a_side","Estimate"]
+fixbremsbetar <- fixef(logmodel_fit)["side","Estimate"]
 
@@
 # Delta 1 
-brmsdel1r <- coef(model_fit)$sub1[,,"a_dif1"][,c("Estimate", "Q2.5", "Q97.5")]
+brmsdel1r <- coef(logmodel_fit)$sub1[,,"dif1"][,c("Estimate", "Q2.5", "Q97.5")]
 brmsdel1r <- as.data.frame(brmsdel1r)
-fixbremsdif2r <- fixef(model_fit)["a_dif1","Estimate"]
+fixbremsdif2r <- fixef(logmodel_fit)["dif1","Estimate"]
 
@@
 # Delta 2 
-brmsdel2r <- coef(model_fit)$sub1[,,"a_dif2"][,c("Estimate", "Q2.5", "Q97.5")]
+brmsdel2r <- coef(logmodel_fit)$sub1[,,"dif2"][,c("Estimate", "Q2.5", "Q97.5")]
 brmsdel2r <- as.data.frame(brmsdel2r)
-fixbremsdif22r <- fixef(model_fit)["a_dif2","Estimate"]
+fixbremsdif22r <- fixef(logmodel_fit)["dif2","Estimate"]
 
@@
 # Delta 3 
-brmsdel3r <- coef(model_fit)$sub1[,,"a_dif3"][,c("Estimate", "Q2.5", "Q97.5")]
+brmsdel3r <- coef(logmodel_fit)$sub1[,,"dif3"][,c("Estimate", "Q2.5", "Q97.5")]
 brmsdel3r <- as.data.frame(brmsdel3r)
-fixbremsdif3r <- fixef(model_fit)["a_dif3","Estimate"]
+fixbremsdif3r <- fixef(logmodel_fit)["dif3","Estimate"]
 
@@
 # Delta 4 
-brmsdel4r <- coef(model_fit)$sub1[,,"a_dif4"][,c("Estimate", "Q2.5", "Q97.5")]
+brmsdel4r <- coef(logmodel_fit)$sub1[,,"dif4"][,c("Estimate", "Q2.5", "Q97.5")]
 brmsdel4r <- as.data.frame(brmsdel4r)
-fixbremsdif4r <- fixef(model_fit)["a_dif4","Estimate"]
+fixbremsdif4r <- fixef(logmodel_fit)["dif4","Estimate"]
 
@@
 ### Delta's individual estimates 
-estimatesbrms <- coef(model_fit)$sub1
+estimatesbrms <- coef(logmodel_fit)$sub1
 #### Delta 1 - difference 8 and 7 
-del1estimatesbrms <- estimatesbrms[,,"a_dif1"]
+del1estimatesbrms <- estimatesbrms[,,"dif1"]
 
 #### Delta 2 - difference 7 and 6 
-del2estimatesbrms <- estimatesbrms[,,"a_dif2"]
+del2estimatesbrms <- estimatesbrms[,,"dif2"]
 
 #### Delta 3 - difference 4 and 3
-del3estimatesbrms <- estimatesbrms[,,"a_dif3"]
+del3estimatesbrms <- estimatesbrms[,,"dif3"]
 
 #### Delta 4 - difference 3 and 2 
-del4estimatesbrms <- estimatesbrms[,,"a_dif4"]
+del4estimatesbrms <- estimatesbrms[,,"dif4"]
 
@@
 # In case you haven't set sample_prior to TRUE, you could load the following model fit where sample_prior was set to TRUE 
-brmmodel8bf <- readRDS("myPath/brmsmodelfit_BF_final.rds")
+brmmodel8bf <- readRDS("myPath/brmslogmodelfit_BF_final.rds")
 
 # Side effect 
-b1brms <- hypothesis(x = model_fit_bf,           # Model fit 
+b1brms <- hypothesis(x = logmodel_fit_bf,           # Model fit 
                      hypothesis = "a_side = 0")     # Null hypothesis
@@
 }
-b1brms <- hypothesis(x = model_fit_bf, hypothesis = "a_side = 0")  
+b1brms <- hypothesis(x = logmodel_fit_bf, hypothesis = "side = 0")  
 plotsdratiob1 <- plot(b1brms, plot = F, theme = theme_get())[[1]]
 sdrbetplot <- savdicratioplot(plotsdratiob1) 
-sdrbetplot <- sdrbetplot + ggtitle(expression(~mu[~beta])) + xlim(-0.1,0.1) + geom_point(aes(x = 0, y = 25), color = "darkorange", size = 1)
+sdrbetplot <- sdrbetplot + ggtitle(expression(~mu[~beta])) + xlim(-0.1,0.1) + geom_point(aes(x = 0, y = 30), color = "darkorange", size = 1)
 sdrbetplot <- sdrbetplot + 
-  geom_segment(aes(x = 0, y = 1.5, xend = 0, yend = 24.5), size=0.1, color = "black", linetype = "dashed") + 
-  annotate("text", x = c(0.03, 0.05), xend = 0.03, y = 10, yend = 10.4, label = c(expression(paste("BF"["01"], " = ")), round(b1brms$hypothesis$Evid.Ratio, 2)), size = 2)
+  geom_segment(aes(x = 0, y = 5.7, xend = 0, yend = 30), size=0.1, color = "black", linetype = "dashed") + 
+  annotate("text", x = c(0.03, 0.037), xend = 0.03, y = 10, yend = 10.4, label = c(expression(paste("BF"["01"], " = ")), round(b1brms$hypothesis$Evid.Ratio, 2)), size = 2)
 
 # Dif 1 
-h1brms <- hypothesis(model_fit_bf, "a_dif1 = 0")  # Evid.Ratio is Bayes factor BF01, have to set sample_prior = TRUE when fitting model 
+h1brms <- hypothesis(logmodel_fit_bf, "dif1 = 0")  # Evid.Ratio is Bayes factor BF01, have to set sample_prior = TRUE when fitting model 
 plotsdratio1 <- plot(h1brms,plot = F, theme = theme_get())[[1]]
@@
   ggtitle(expression(~mu[~delta[~"7"]])) + xlim(-0.1,0.1) + 
-  geom_point(aes(x = 0, y = 20), color = "darkorange", size = 1) + 
-  geom_segment(aes(x = 0, y = 1.5, xend = 0, yend = 19), size=0.1, color = "black", linetype = "dashed") + 
-  annotate("text", x = c(0.035, 0.05), xend = 0.03, y = 10, yend = 10.4, label = c(expression(paste("BF"["01"], " = ")), round(h1brms$hypothesis$Evid.Ratio, 2)), size = 2) + 
+  geom_point(aes(x = 0, y = 1.7), color = "darkorange", size = 1) + 
+  geom_segment(aes(x = 0, y = 5.7, xend = 0, yend = 2), size=0.1, color = "black", linetype = "dashed") + 
+  annotate("text", x = c(-0.02, -0.012), xend = 0.03, y = 10, yend = 10.4, label = c(expression(paste("BF"["01"], " = ")), round(h1brms$hypothesis$Evid.Ratio, 2)), size = 2) + 
   ylab("")
@@
 # Dif 2 
-h2brms <- hypothesis(model_fit_bf, "a_dif2 = 0")  # Evid.Ratio is Bayes factor BF01, have to set sample_prior = TRUE when fitting model 
+h2brms <- hypothesis(logmodel_fit_bf, "dif2 = 0")  # Evid.Ratio is Bayes factor BF01, have to set sample_prior = TRUE when fitting model 
 plotsdratio2 <- plot(h2brms,plot = F, theme = theme_get())[[1]]
@@
   geom_point(aes(x = 0, y = 0), color = "darkorange", size = 1) + 
-  geom_segment(aes(x = 0, y = 1.5, xend = 0, yend = 0), size=0.1, color = "black", linetype = "dashed") + 
-  annotate("text", x = c(0.0, 0.02), xend = 0.03, y = 10, yend = 10.4, label = c(expression(paste("BF"["01"], " = ")), round(h2brms$hypothesis$Evid.Ratio, 2)), size = 2)
+  geom_segment(aes(x = 0, y = 5.7, xend = 0, yend = 0), size=0.1, color = "black", linetype = "dashed") + 
+  annotate("text", x = c(0.0, 0.005), xend = 0.03, y = 10, yend = 10.4, label = c(expression(paste("BF"["01"], " = ")), round(h2brms$hypothesis$Evid.Ratio, 2)), size = 2)
 
 # Dif 3 
-h3brms <- hypothesis(model_fit_bf, "a_dif3 = 0")  # Evid.Ratio is Bayes factor BF01, have to set sample_prior = TRUE when fitting model 
+h3brms <- hypothesis(logmodel_fit_bf, "dif3 = 0")  # Evid.Ratio is Bayes factor BF01, have to set sample_prior = TRUE when fitting model 
 plotsdratio3 <- plot(h3brms,plot = F, theme = theme_get())[[1]]
@@
   geom_point(aes(x = 0, y = 0), color = "darkorange", size = 1) + 
-  geom_segment(aes(x = 0, y = 1.5, xend = 0, yend = 0), size=0.1, color = "black", linetype = "dashed") + 
-  annotate("text", x = c(0.0, 0.02), xend = 0.03, y = 10, yend = 10.4, label = c(expression(paste("BF"["01"], " = ")), round(h3brms$hypothesis$Evid.Ratio, 2)), size = 2) + 
+  geom_segment(aes(x = 0, y = 5.7, xend = 0, yend = 0), size=0.1, color = "black", linetype = "dashed") + 
+  annotate("text", x = c(0.0, 0.005), xend = 0.03, y = 10, yend = 10.4, label = c(expression(paste("BF"["01"], " = ")), round(h3brms$hypothesis$Evid.Ratio, 2)), size = 2) + 
   ylab("")
@@
 # Dif 4 
-h4brms <- hypothesis(model_fit_bf, "a_dif4 = 0")  # Evid.Ratio is Bayes factor BF01, have to set sample_prior = TRUE when fitting model 
+h4brms <- hypothesis(logmodel_fit_bf, "dif4 = 0")  # Evid.Ratio is Bayes factor BF01, have to set sample_prior = TRUE when fitting model 
 plotsdratio4 <- plot(h4brms,plot = F, theme = theme_get())[[1]]
@@
   geom_area(position = "identity", alpha = 0.3) + 
-  geom_point(aes(x = 0 , y = 1.343088), size = 1) + 
-  geom_point(aes(x = 0, y = 9), color = "darkorange", size = 1) + 
+  geom_point(aes(x = 0 , y = 5.6), size = 1) + 
+  geom_point(aes(x = 0, y = 0), color = "darkorange", size = 1) + 
   xlim(-0.1, 0.1) + 
@@
   scale_color_manual(values = c("darkorange", "darkgreen")) + 
-  geom_segment(aes(x = 0, y = 1.5, xend = 0, yend = 8.5), size=0.1, color = "black", linetype = "dashed") + 
-  annotate("text", x = c(-0.01, 0.007), xend = 0.03, y = 10, yend = 10.4, label = c(expression(paste("BF"["01"], " = ")), round(h4brms$hypothesis$Evid.Ratio, 2)), size = 2) + 
+  geom_segment(aes(x = 0, y = 5, xend = 0, yend = 0), size=0.1, color = "black", linetype = "dashed") + 
+  annotate("text", x = c(-0.01, -0.005), xend = 0.03, y = 10, yend = 10.4, label = c(expression(paste("BF"["01"], " = ")), round(h4brms$hypothesis$Evid.Ratio, 2)), size = 2) + 
   guides(color = guide_legend(override.aes = list(linetype = 0, shape = NA)))
@@
 # Formula 
-formula3 <- brms::bf(rt2 ~ a + b, nl = TRUE, 
-                     lf(a ~ 0 + side + (0 + side || sub1), center = TRUE),  
-                     lf(b ~ 0 + inter + (0 + inter || sub1), cmc = FALSE)
-)
+formula_side <- rt2 ~ 1 + side +   # General effects 
+  (1 + side || sub1)
 
 # Priors 
-priorsmodel3 <- c(set_prior("normal(0.5,1)", class = "b", lb = 0, nlpar = "b"),  
-                  set_prior("inv_gamma(13.8, 6.3)", class = "sd", coef="inter", group = "sub1", nlpar = "b"), 
-                  set_prior("normal(0,0.3)", class = "b", coef="side",  nlpar = "a"),  
-                  set_prior("inv_gamma(13.8,5.3)", class = "sd", coef="side", group = "sub1",  nlpar = "a"), 
-                  set_prior("inv_gamma(13.8, 6.3)", class = "sigma"))
+priorsmodellog_side <- c(set_prior("normal(-0.5,1)", class = "Intercept"),  
+                  set_prior("inv_gamma(13.75136,4.102063)", class = "sd", coef="Intercept", group = "sub1"), 
+                  set_prior("normal(0,0.07071068)", class = "b", coef="side"),  
+                  set_prior("inv_gamma(13.75208,0.7489703)", class = "sd", coef="side", group = "sub1"), 
+                  set_prior("inv_gamma(13.75136, 4.102063)", class = "sigma"))
 
 # Estimate side model 
-side_model_fit <- brm(formula = formula3,         # Model formula 
+side_logmodel_fit <- brm(formula = formula_side,         # Model formula 
                 data = myData_dataframe,              # Data frame with variables + dependent variable 
-                family = gaussian(),        # Response distribution 
-                prior = priorsmodel3,       # Priors 
+                family = lognormal(),        # Response distribution 
+                prior = priorsmodellog_side,       # Priors 
                 # sample_prior = TRUE, need to add this for model comparison  
                 warmup = 1000,              # Iterations used for warmup 
-                iter = 6000,                # Total iterations per chain 
+                iter = 4000,                # Total iterations per chain 
                 chains = 4,                 # Number of chains 
@@
 ```{r bridgesamplingexample, eval = FALSE}
-bridgesampling::bayes_factor(model_fit, side_model_fit)[1]$bf  # BF of the full model against the side model 
+bridgesampling::bayes_factor(logmodel_fit, side_logmodel_fit)[1]$bf  # BF of the full model against the side model 
 ```

### 2022-06-09 H - Posterior Distributions Tables/Online Supplement H - Tables with Information on Posterior Distributions.Rmd
All documents except .rds

---
title             : "Online Supplement H: Tables with Information on Posterior Distributions"
author            : " "
bibliography      : ["r-references.bib"]

header-includes   :
   - \usepackage{bm}
   - \usepackage{amsmath}
   - \usepackage{setspace}
   - \usepackage{xcolor}
   - \definecolor{mypink}{RGB}{255, 230, 255}
   - \usepackage{todonotes}
   - \usepackage{graphicx}
   - \usepackage{float}
   - \renewcommand{\tablename}{Table G}
   - \makeatletter
   - \def\tnum@table{\tablename\thetable}
   - \makeatother

csl               : apa6.csl
output            : pdf_document
---
\spacing{1.5}
\fontsize{12}{12}

```{r setup, include=FALSE}
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

knitr::opts_chunk$set(tidy.opts=list(width.cutoff=40), tidy=TRUE, echo = TRUE, message = FALSE, warning = FALSE, fig.pos = "H")
```

\scriptsize 
```{R loadpackages, message = FALSE, warning = FALSE, echo = FALSE}
# Load the packages
library("papaja")
library("LaplacesDemon")
library("rstan")
library("brms")
library("plyr")
library("lattice")
library("ggplot2")
library("dplyr")
library("readr")
library("rmarkdown")
library("Rmisc")
library("devtools")
library("gghalves") 
library("bayesplot")
library("bridgesampling") 
library("gridExtra") 
library("tibble") 
library("kableExtra") 
library("bayestestR")
library("ggridges")
library("truncnorm")
library("ggbeeswarm")
```
\normalsize

|    This Online Supplement contains information on the posterior distributions of the general effects and individual variation estimated by `rstan` and `brms` for the normal and log-normal models. In the first table, the estimated mean of the posterior distribution, the standard error of this mean, the lower and upper bound of the 95% credible interval, the $\hat{R}$ and the number of effective samples are provided. In the second table, the estimated variance of the posterior distribution, the standard error of this variance, the lower and upper bound of the 95% credible interval, the $\hat{R}$ and the number of effective samples are provided. 
&nbsp;

## Normal model 
### Rstan 
\normalsize 
```{r loadmodelfit1, echo = FALSE}
setwd('..')
model_fit <- readRDS(file = "R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
```

```{r geneftable1, results = "asis", cache = TRUE, echo = FALSE}
# Obtain the mean, standard error, 95% credible interval, rhat and number of effective samples for the intercept, and the digit and side parameters 
fitrstanhiertable <- summary(model_fit,  # model fit 
                             # Select parameters 
                             pars = c("mu", "mu2", "mu3", "mu4", "mu5", 
                                      "mu6"))$summary[,c("mean", "se_mean", 
                                                         "2.5%", "97.5%", 
                                                         "n_eff", "Rhat")]

# The results are transformed to a dataframe that allows us to plot the results nicely in a table 
fitrstanhiertable2 <- as.data.frame(fitrstanhiertable)

# We manually set the row names of the table using latex for mathematical symbols 
rownamesfitrstanhiertable <-  c("$\\mu_{\\gamma}$","$\\mu_{\\beta}$", "$\\mu_{\\delta_{7}}$", "$\\mu_{\\delta_{6}}$", "$\\mu_{\\delta_{4}}$", "$\\mu_{\\delta_{3}}$")

# We manually set the column names of the table 
colnames(fitrstanhiertable2) <- c("Mean", "SE", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")

# The column with row names is added to the table 
fitrstanhiertable3 <- add_column(fitrstanhiertable2, 
                                 Parameters = rownamesfitrstanhiertable, 
                                 .before = "Mean")

# Using the apa_table function from the papaja package the results are presented in an apa table 
apa_table(fitrstanhiertable3,   # Data frame with results 
          row.names = FALSE,    # Default row names are deleted (we set them ourself)
          caption = "Posterior Mean, Standard Error (SE) of the Mean, Lower and 
          Upper Bound of the 95\\% Credible Interval, the Number of Effective 
          Samples, and the $\\hat{R}$ of the General Effect Parameters as 
          estimated by Rstan.", 
          align = c("l", "c", "c", "c", "c", "c", "c"),   # How should results be presented, in the midle of the table or outlined left 
          digits = 3,       # The number of digits behind the ".". 
          placement = "h",  # Place after code 
          escape = FALSE)   # This indicates that the results contain latex code that should be read as latex code. It is important to do this in combination with putting the results to "asis" in the {R} section above.

```

```{r inddevtable1, results = "asis", cache = TRUE, echo = FALSE}
fitrstanhiertable2 <- summary(model_fit,   # model fit 
                              # Select variance parameters (we called them g)
                              pars = c("g", "g2", "g3", "g4", "g5", 
                                       "g6"))$summary[,c("mean", "se_mean", 
                                                         "2.5%", "97.5%", 
                                                         "n_eff", "Rhat")]

# The results are transformed to a dataframe that allows us to plot the results nicely in a table 
fitrstanhiertable22 <- as.data.frame(fitrstanhiertable2)

# We manually set the row names of the table using latex for mathematical symbols 
rownamesfitrstanhiertable2 <-  c("$\\sigma_{\\gamma}^2$","$\\sigma_{\\beta}^2$", "$\\sigma_{\\delta_{7}}^2$", "$\\sigma_{\\delta_{6}}^2$", "$\\sigma_{\\delta_{4}}^2$", "$\\sigma_{\\delta_{3}}^2$")

# The column with row names is added to the table 
colnames(fitrstanhiertable22) <- c("Mean", "SE", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")

# The column with row names is added to the table 
fitrstanhiertable32 <- add_column(fitrstanhiertable22, 
                                  Parameters = rownamesfitrstanhiertable2, 
                                  .before = "Mean")

# Using the apa_table function from the papaja package the results are presented in an apa table 
apa_table(fitrstanhiertable32,  # Data frame with results
          row.names = FALSE,    # Default row names are deleted (we set them ourself)
          caption = "Posterior Variance, Standard Error (SE) of the Variance, 
          Lower and Upper Bound of the 95\\% Credible Interval, the Number of 
          Effective Samples, and the $\\hat{R}$ of the Variance Parameters           as estimated by Rstan.", 
          align = c("l", "c", "c", "c", "c", "c", "c"),  # How should results be presented, in the midle of the table or outlined left 
          digits = 3,        # The number of digits behind the ".". 
          place = "h",       # Place after code 
          escape = FALSE)    # This indicates that the results contain latex code that should be read as latex code. It is important to do this in combination with putting the results to "asis" in the {R} section above.

```
\newpage

### brms 
```{r loadmodelfits2, show = FALSE, echo = FALSE}
setwd('..')
model_fit <- readRDS(file = "R objects/brms/Truncated/BRMSmodelfit18_03012022.rds")

```

\scriptsize 
```{r geneftable2, results = "asis", cache = TRUE, echo = FALSE}
# Obtain the mean, standard error, 95% credible interval, rhat and number of effective samples for the intercept, and the digit and side parameters 
fitbrmshiertable <- summary(model_fit)[[14]][,c("Estimate", "Est.Error", "l-95% CI", "u-95% CI", "Bulk_ESS", "Rhat")]  # variance

# The results are transformed to a dataframe that allows us to plot the results nicely in a table 
fitbrmshiertable2 <- as.data.frame(fitbrmshiertable)
fitbrmshiertable2 <- rbind(fitbrmshiertable2[6,], fitbrmshiertable2[1:5,])

# We manually set the row names of the table using latex for mathematical symbols 
rownamesfitbrmshiertable <-  c("$\\mu_{\\gamma}$","$\\mu_{\\beta}$", "$\\mu_{\\delta_{7}}$", 
                               "$\\mu_{\\delta_{6}}$", "$\\mu_{\\delta_{4}}$", "$\\mu_{\\delta_{3}}$")

# We manually set the column names of the table 
colnames(fitbrmshiertable2) <- c("Mean", "SE", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")

# The column with row names is added to the table 
fitbrmshiertable3 <- add_column(fitbrmshiertable2, 
                                 Parameters = rownamesfitbrmshiertable, 
                                 .before = "Mean")

# Using the apa_table function from the papaja package the results are presented in an apa table 
apa_table(fitbrmshiertable3,   # Data frame with results 
          row.names = FALSE,    # Default row names are deleted (we set them ourself)
          caption = "Posterior Mean, Standard Error (SE) of the Mean, Lower and 
          Upper Bound of the 95\\% Credible Interval, the Number of Effective 
          Samples, and the $\\hat{R}$ of the General Effect Parameters as estimated by brms. ", 
          align = c("l", "c", "c", "c", "c", "c", "c"),   # How should results be presented, in the midle of the table or outlined left 
          digits = 3,       # The number of digits behind the ".". 
          placement = "h",  # Place after code 
          escape = FALSE)   # This indicates that the results contain latex code that should be read as latex code. It is important to do this in combination with putting the results to "asis" in the {R} section above.

```
\normalsize

\scriptsize 
```{r inddevtable2, results = "asis", cache = TRUE, echo = FALSE}
# Obtain standard deviation estimates 
fitbrmshiertable2 <- summary(model_fit)[[17]]$sub1[,c("Estimate", "Est.Error", "l-95% CI", "u-95% CI", "Bulk_ESS", "Rhat")]

# The results are transformed to a dataframe that allows us to plot the results nicely in a table 
fitbrmshiertable22 <- as.data.frame(fitbrmshiertable2)
fitbrmshiertable22 <- rbind(fitbrmshiertable22[6,], fitbrmshiertable22[1:5,])

# We manually set the row names of the table using latex for mathematical symbols 
rownamesfitbrmshiertable2 <-  c("$\\sigma_{\\gamma}$","$\\sigma_{\\beta}$", "$\\sigma_{\\delta_{7}}$", 
                                "$\\sigma_{\\delta_{6}}$", "$\\sigma_{\\delta_{4}}$", "$\\sigma_{\\delta_{3}}$")

# The column with row names is added to the table 
colnames(fitbrmshiertable22) <- c("Mean", "SE", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")

# The column with row names is added to the table 
fitbrmshiertable32 <- add_column(fitbrmshiertable22, 
                                  Parameters = rownamesfitbrmshiertable2, 
                                  .before = "Mean")

# Using the apa_table function from the papaja package the results are presented in an apa table 
apa_table(fitbrmshiertable32,  # Data frame with results
          row.names = FALSE,    # Default row names are deleted (we set them ourself)
          caption = "Posterior Standard Deviation, Standard Error (SE) of the Standard Deviation, 
          Lower and Upper Bound of the 95\\% Credible Interval, the Number of 
          Effective Samples, and the $\\hat{R}$ of the Standard Deviation Parameters           as estimated by brms", 
          align = c("l", "c", "c", "c", "c", "c", "c"),  # How should results be presented, in the midle of the table or outlined left 
          digits = 3,        # The number of digits behind the ".". 
          place = "h",       # Place after code 
          escape = FALSE)    # This indicates that the results contain latex code that should be read as latex code. It is important to do this in combination with putting the results to "asis" in the {R} section above.

```
\normalsize 
\newpage

## Log normal model 
### Rstan 
```{r loadmodelfit3, echo = FALSE}
setwd('..')
logmodel_fit <- readRDS(file = "R objects/rstan/log-normal model/hier_logmodelc_adj13012022.rds")
```

\scriptsize 
```{r geneftable3, results = "asis", cache = TRUE, echo = FALSE}
# Obtain the mean, standard error, 95% credible interval, rhat and number of effective samples for the intercept, and the digit and side parameters 
fitrstanhiertable <- summary(logmodel_fit,  # model fit 
                             # Select parameters 
                             pars = c("mu", "mu2", "mu3", "mu4", "mu5", 
                                      "mu6"))$summary[,c("mean", "se_mean", 
                                                         "2.5%", "97.5%", 
                                                         "n_eff", "Rhat")]

# The results are transformed to a dataframe that allows us to plot the results nicely in a table 
fitrstanhiertable2 <- as.data.frame(fitrstanhiertable)

# We manually set the row names of the table using latex for mathematical symbols 
rownamesfitrstanhiertable <-  c("$\\mu_{\\gamma}$","$\\mu_{\\beta}$", "$\\mu_{\\delta_{7}}$", "$\\mu_{\\delta_{6}}$", "$\\mu_{\\delta_{4}}$", "$\\mu_{\\delta_{3}}$")

# We manually set the column names of the table 
colnames(fitrstanhiertable2) <- c("Mean", "SE", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")

# The column with row names is added to the table 
fitrstanhiertable3 <- add_column(fitrstanhiertable2, 
                                 Parameters = rownamesfitrstanhiertable, 
                                 .before = "Mean")

# Using the apa_table function from the papaja package the results are presented in an apa table 
apa_table(fitrstanhiertable3,   # Data frame with results 
          row.names = FALSE,    # Default row names are deleted (we set them ourself)
          caption = "Posterior Mean, Standard Error (SE) of the Mean, Lower and 
          Upper Bound of the 95\\% Credible Interval, the Number of Effective 
          Samples, and the $\\hat{R}$ of the General Effect Parameters as 
          estimated by Rstan.", 
          align = c("l", "c", "c", "c", "c", "c", "c"),   # How should results be presented, in the midle of the table or outlined left 
          digits = 3,       # The number of digits behind the ".". 
          placement = "h",  # Place after code 
          escape = FALSE)   # This indicates that the results contain latex code that should be read as latex code. It is important to do this in combination with putting the results to "asis" in the {R} section above.

```
\normalsize

\scriptsize 
```{r inddevtable3, results = "asis", cache = TRUE, echo = FALSE}
fitrstanhiertable2 <- summary(logmodel_fit,   # model fit 
                              # Select variance parameters (we called them g)
                              pars = c("g", "g2", "g3", "g4", "g5", 
                                       "g6"))$summary[,c("mean", "se_mean", 
                                                         "2.5%", "97.5%", 
                                                         "n_eff", "Rhat")]

# The results are transformed to a dataframe that allows us to plot the results nicely in a table 
fitrstanhiertable22 <- as.data.frame(fitrstanhiertable2)

# We manually set the row names of the table using latex for mathematical symbols 
rownamesfitrstanhiertable2 <-  c("$\\sigma_{\\gamma}^2$","$\\sigma_{\\beta}^2$", "$\\sigma_{\\delta_{7}}^2$", "$\\sigma_{\\delta_{6}}^2$", "$\\sigma_{\\delta_{4}}^2$", "$\\sigma_{\\delta_{3}}^2$")

# The column with row names is added to the table 
colnames(fitrstanhiertable22) <- c("Mean", "SE", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")

# The column with row names is added to the table 
fitrstanhiertable32 <- add_column(fitrstanhiertable22, 
                                  Parameters = rownamesfitrstanhiertable2, 
                                  .before = "Mean")

# Using the apa_table function from the papaja package the results are presented in an apa table 
apa_table(fitrstanhiertable32,  # Data frame with results
          row.names = FALSE,    # Default row names are deleted (we set them ourself)
          caption = "Posterior Variance, Standard Error (SE) of the Variance, 
          Lower and Upper Bound of the 95\\% Credible Interval, the Number of 
          Effective Samples, and the $\\hat{R}$ of the Variance Parameters           as estimated by Rstan.", 
          align = c("l", "c", "c", "c", "c", "c", "c"),  # How should results be presented, in the midle of the table or outlined left 
          digits = 3,        # The number of digits behind the ".". 
          place = "h",       # Place after code 
          escape = FALSE)    # This indicates that the results contain latex code that should be read as latex code. It is important to do this in combination with putting the results to "asis" in the {R} section above.

```
\normalsize 
\newpage

### brms

\scriptsize 
```{r loadmodelfits4, show = FALSE, echo = FALSE}
# Log model brms 
setwd('..')
logmodel_fit <- readRDS(file = "R objects/brms/log-normal/brmmodel8log13012022.rds")

```
\normalsize

\scriptsize 
```{r geneftable4, results = "asis", cache = TRUE, echo = FALSE}
# Obtain the mean, standard error, 95% credible interval, rhat and number of effective samples for the intercept, and the digit and side parameters 
fitbrmshiertable <- summary(logmodel_fit)[[14]][,c("Estimate", "Est.Error", "l-95% CI", "u-95% CI", "Bulk_ESS", "Rhat")]  # variance

# The results are transformed to a dataframe that allows us to plot the results nicely in a table 
fitbrmshiertable2 <- as.data.frame(fitbrmshiertable)

# We manually set the row names of the table using latex for mathematical symbols 
rownamesfitbrmshiertable <-  c("$\\mu_{\\gamma}$","$\\mu_{\\beta}$", "$\\mu_{\\delta_{7}}$", 
                               "$\\mu_{\\delta_{6}}$", "$\\mu_{\\delta_{4}}$", "$\\mu_{\\delta_{3}}$")

# We manually set the column names of the table 
colnames(fitbrmshiertable2) <- c("Mean", "SE", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")

# The column with row names is added to the table 
fitbrmshiertable3 <- add_column(fitbrmshiertable2, 
                                 Parameters = rownamesfitbrmshiertable, 
                                 .before = "Mean")

# Using the apa_table function from the papaja package the results are presented in an apa table 
apa_table(fitbrmshiertable3,   # Data frame with results 
          row.names = FALSE,    # Default row names are deleted (we set them ourself)
          caption = "Posterior Mean, Standard Error (SE) of the Mean, Lower and Upper Bound of the 95\\% Credible Interval, Number of Effective Samples, and $\\hat{R}$ of the General Effect Parameters as estimated by brms. ", 
          align = c("l", "c", "c", "c", "c", "c", "c"),   # How should results be presented, in the midle of the table or outlined left 
          digits = 3,       # The number of digits behind the ".". 
          placement = "h",  # Place after code 
          escape = FALSE)   # This indicates that the results contain latex code that should be read as latex code. It is important to do this in combination with putting the results to "asis" in the {R} section above.

```
\normalsize

\scriptsize 
```{r inddevtable4, results = "asis", cache = TRUE, echo = FALSE}
# Obtain standard deviation estimates 
fitbrmshiertable2 <- summary(logmodel_fit)[[17]]$sub1[,c("Estimate", "Est.Error", "l-95% CI", "u-95% CI", "Bulk_ESS", "Rhat")]

# The results are transformed to a dataframe that allows us to plot the results nicely in a table 
fitbrmshiertable22 <- as.data.frame(fitbrmshiertable2)

# We manually set the row names of the table using latex for mathematical symbols 
rownamesfitbrmshiertable2 <-  c("$\\sigma_{\\gamma}$","$\\sigma_{\\beta}$", "$\\sigma_{\\delta_{7}}$", 
                                "$\\sigma_{\\delta_{6}}$", "$\\sigma_{\\delta_{4}}$", "$\\sigma_{\\delta_{3}}$")

# The column with row names is added to the table 
colnames(fitbrmshiertable22) <- c("Mean", "SE", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")

# The column with row names is added to the table 
fitbrmshiertable32 <- add_column(fitbrmshiertable22, 
                                  Parameters = rownamesfitbrmshiertable2, 
                                  .before = "Mean")

# Using the apa_table function from the papaja package the results are presented in an apa table 
apa_table(fitbrmshiertable32,  # Data frame with results
          row.names = FALSE,    # Default row names are deleted (we set them ourself)
          caption = "Posterior Variance, Standard Error (SE) of the Standard Deviation, Lower and Upper Bound of the 95\\% Credible Interval, Number of Effective Samples, and $\\hat{R}$ of the Variance Parameters as estimated by brms.", 
          align = c("l", "c", "c", "c", "c", "c", "c"),  # How should results be presented, in the midle of the table or outlined left 
          digits = 3,        # The number of digits behind the ".". 
          place = "h",       # Place after code 
          escape = FALSE)    # This indicates that the results contain latex code that should be read as latex code. It is important to do this in combination with putting the results to "asis" in the {R} section above.

```
\normalsize

### 2022-06-09 manuscript.Rmd
All documents except .rds

---
title             : "Bayesian Hierarchical Modeling: An Introduction and Reassessment"
shorttitle        : "Bayesian Hierarchical Modeling"

author: 
  - name          : "Myrthe Veenman"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : " "
    email         : "myrthe.veenman@gmail.com"
  - name          : "Angelika Stefan"
    affiliation   : "1"
    corresponding : no    # Define only one corresponding author
  - name          : "Julia Haaf"
    affiliation   : "1"
    corresponding : no    # Define only one corresponding author

affiliation:
  - id            : "1"
    institution   : "University of Amsterdam"

authornote: |

  This paper is written in R Markdown. The R Markdown file, other R code, and additional information on this study can be found in the Online Supplement.

abstract: |
  With the recent development of easy-to-use tools for Bayesian analysis, psychologists have started to embrace Bayesian hierarchical modeling. Bayesian hierarchical models provide an intuitive account of inter- and intraindividual variability and are particularly suited for the evaluation of repeated-measures designs. Here, we provide guidance for model specificaton and interpretation in Bayesian hierarchical modeling and describe common pitfalls that can arise in the process of model fitting and evaluation. Our introduction gives particular emphasis to prior specification and prior sensitivity, as well as to the calculation of Bayes factors for model comparisons. We illustrate the use of state-of-the-art software programs `Stan` and `brms`. The result is an overview over best practices in Bayesian hierarchical modeling that, as we hope, will help psychologists in making the best use of Bayesian hierarchical modeling.

keywords          : "Multilevel, Tutorial, rstan, brms, Bayes factor"

bibliography      : ["r-references.bib"]

header-includes   :
   - \usepackage{bm}
   - \usepackage{amsmath}
   - \usepackage{setspace}
   - \usepackage{xcolor}
   - \definecolor{mypink}{RGB}{255, 230, 255}
   - \usepackage{todonotes}
   - \newcommand{\Julia}[1]{\todo[inline, color=mypink]{#1}}
   - \usepackage{graphicx}
   - \usepackage{float}
   - \usepackage{tabularx}
   - \usepackage{hyperref}

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
figcaption        : yes 
indent            : true

documentclass     : "apa6"
classoption       : "doc"
output            : papaja::apa6_pdf
csl               : apa6.csl
---

```{r setup, include = FALSE}
library(rmarkdown)
library("papaja")
library(knitr)
library(plyr)
library(dplyr)
library(ggplot2)
library(rstan)
library(brms)
library(LaplacesDemon)
library(gridExtra)
library(ggridges)
library(BayesFactor)
library(crch)
library(foreign)
library(tibble)
library(bridgesampling)
library(cowplot)
library(kableExtra)
library(readr)
library(lattice)
library(Rmisc)
library(devtools)
library(gghalves)
library(bayesplot)
library(ggmcmc)
library(lme4)
library(logspline)
library(truncnorm)
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed, fig.pos = "H")
```

\newpage

# Introduction

Imagine a typical scenario in the life of an experimental psychology graduate student: You have just collected a large data set, either online or in the lab. Each participant has spent one valuable hour completing many trials per experimental condition. Now it is time to analyze the data, and your supervisor suggests to run a within-subjects ANOVA. This procedure requires you to first aggregate the data by condition and participant. Doing so reduces your data set from 10,000 rows to just 400. Reducing the data by aggregation makes you a bit queasy: It feels like your efforts in data collection and the participants' time are not valued this way. You know there needs to be a better way. And there is: Hierarchical modeling.

Just as in the scenario above, experiments in psychology are often implemented in a repeated-measures design where participants respond to several items, stimuli, or conditions. From a data analysis perspective, such a design implies that observations are nested within participants. The statistical and the psychological literature agree that the optimal analysis accounting for such a data structure is hierarchical modeling [e.g., @efron1977stein; Lee, -@lee2011cognitive; Rouder & Lu, -@rouder2005introduction]. And yet, it is still not the norm to apply hierarchical modeling to data from psychological experiments. One reason for this is that hierarchical modeling is more difficult than familiar procedures such as ANOVA.

With recent software development, hierarchical modeling has become more and more accessible. In addition, many researchers have argued that hierarchical modeling is easier and the interpretation is more intuitive when it is done in the Bayesian statistical framework [e.g., Lynch, -@lynch2007introduction; Rouder & Lu, 2005; Rouder, Morey, & Pratte, -@rouder2013hierarchical]. 
However, there are challenges to the Bayesian way of hierarchical modeling (Rouder & Lu, 2005). These challenges, such as the choice of priors, the need for model comparisons of highly complex models, and programming skills required to use state-of-the-art software solutions, can make researchers hesitant to use this approach. We believe that existing introductions to Bayesian hierarchical modeling do not adequately address these challenges. For example, they use outdated software (Rouder et al., 2013), do not cover model comparison (Rouder & Lu, 2005), or do not discuss prior specification [@shiffrin2008survey]. This paper aims to make Bayesian hierarchical modeling available to a wider public by providing:

- A comparison of and guidance for the use of two software packages,  
- Guidance on prior specification, 
- Practical instructions for Bayes factor model comparison.

Our tutorial is directed at researchers who have a basic understanding of Bayesian inference and parameter estimation and are planning to perform more complex Bayesian analyses.

The application of Bayesian hierarchical analysis will be illustrated using a digit classification task collected by @rouder2005hierarchical. In an experimental task participants repeatedly indicated whether a target digit (i.e., 2, 3, 4, 6, 7, or 8) was greater or smaller than 5. The participants in the task performed multiple trials per condition (i.e., digit). Therefore, observations are nested in participants. Our analyses focus on two effects that can potentially occur in this digit classification task: The digit effect and the side effect. According to the symbolic distance hypothesis [Moyer & Landauer, -@moyer1967time] processing of numbers is analogue. If numbers are closer on the number line, we may confuse them more, and comparing them is more difficult. In line with this hypothesis, the digit effect postulates that response times (RTs) are slower when digits are closer to 5. Additionally, we may hypothesize that RTs are affected by whether the target digit was smaller or greater than 5, yet it is unclear which side of 5 would lead to faster or slower RTs. The digit classification task is a standard task that usually would be analyzed in a 3 (digit) by 2 (side) within-subjects ANOVA. In this tutorial, we show how inferences can be improved using a hierarchical Bayesian model instead.

```{r data, include = FALSE, eval = FALSE}
# maybe add object here where I save the link 
indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
colnames(indat)=c('sub','block','trial','stim','resp','rt','error')

hist(indat$rt[indat$sub==0])
table(indat$rt[indat$sub==0])
nrow(indat[indat$sub==0,])

unique(indat$stim)
for (i in unique(indat$stim)){
hist(indat$rt[indat$stim==i])
}

clean=function()
{
  indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
  colnames(indat)=c('sub','block','trial','stim','resp','rt','error')

  bad1=indat$sub%in%c(34,43)
  bad2=indat$rt<250 | indat$rt>2000
  bad3=indat$err==1
  bad4=indat$block==0 & indat$trial<20
  bad5=indat$trial==0

  bad=bad1 | bad2 | bad3 |bad4 |bad5
  dat=indat[!bad,]
  return(dat)
}

# x axis digit 
# y axis: mean RT pp 
indat1 <- clean()

meanrtindat <- aggregate(indat1$rt, list(indat1$stim, indat1$sub), mean)  # mean response time per condition per participant 
meanrt <- aggregate(indat1$rt, list(indat1$stim), mean)
for (i in 1:nrow(meanrtindat)){
  if (meanrtindat$Group.1[i] == 0) {meanrtindat$Group.1[i] <- 2 
  } else if(meanrtindat$Group.1[i] == 1) {meanrtindat$Group.1[i] <- 3 
  } else if(meanrtindat$Group.1[i] == 2) {meanrtindat$Group.1[i] <- 4
  } else if(meanrtindat$Group.1[i] == 3) {meanrtindat$Group.1[i] <- 6 
  } else if(meanrtindat$Group.1[i] == 4) {meanrtindat$Group.1[i] <- 7 
  } else {meanrtindat$Group.1[i] <- 8} 
}

for (i in 1:nrow(meanrt)){
  if (meanrt$Group.1[i] == 0) {meanrt$Group.1[i] <- 2 
  } else if(meanrt$Group.1[i] == 1) {meanrt$Group.1[i] <- 3 
  } else if(meanrt$Group.1[i] == 2) {meanrt$Group.1[i] <- 4
  } else if(meanrt$Group.1[i] == 3) {meanrt$Group.1[i] <- 6 
  } else if(meanrt$Group.1[i] == 4) {meanrt$Group.1[i] <- 7 
  } else {meanrt$Group.1[i] <- 8} 
}
```

```{r figure1code, echo=FALSE, fig.align='left', fig.cap="Response times in milliseconds in two conditions (i.e., digit 3 and 4). The grey lines represent the individual response time effects, and the blue line the mean aggregated response time effect.", fig.height=4.5, fig.width=6, eval = FALSE}
theme_set(theme_apa(base_size = 9))
meanrtplot <- ggplot(meanrtindat, aes(y = x, x = as.factor(Group.1))) + geom_line(aes(group = Group.2), alpha = .15) + 
  geom_line(data = meanrt, aes(group = 1), color='blue') + geom_point(data = meanrt, aes(group = 1), color='blue') + 
  labs(x = "Digit", y = "Response Time (Ms)") + theme_bw() + theme(axis.line = element_line(colour = "black"), panel.border = element_blank(), panel.background = element_blank())

# meanrtplot

# Difference 3 and 4 
dif34indat <- meanrtindat[which(meanrtindat$Group.1 == 3 | meanrtindat$Group.1 == 4),]
meandif34 <- meanrt[which(meanrt$Group.1 == 3 | meanrt$Group.1 == 4),]

meanrtplot2 <- ggplot(dif34indat, aes(y = x, x = as.factor(Group.1))) + geom_line(aes(group = Group.2), alpha = .15) + geom_point(alpha = .15) + 
  geom_line(data = meandif34, aes(group = 1), color='blue', size = 1) + geom_point(data = meandif34, aes(group = 1), color='blue', shape = 21, size = 2, stroke = 1) + 
  labs(x = "Digit", y = "Response Time (Ms)") + theme_bw() + theme(axis.line = element_line(colour = "black"), panel.border = element_blank(), panel.background = element_blank())

meanrtplot2 
```

```{r figure1, echo=FALSE, fig.align='left', fig.cap="Panel A shows the response times in seconds for two conditions of the number classification task (i.e., digits 3 and 4). The grey lines represent the individual effects, and the blue line the aggregated effect. Panel B shows how individual estimates are shrunken towards the general trend. Sample represents the observed effect per participant in seconds, while Model represents the estimated effects per participant. ", fig.height=3, fig.width=6, out.width ="95%"}
knitr::include_graphics("Images/Fig1_meanrtandhierplot_v2.png")
```

```{r figure1b, echo=FALSE, fig.align='left', fig.cap="RTs per condition (i.e., digit). The grey lines represent the individual RTs, and the blue line the mean aggregated response time.", fig.height=4, fig.width=6, eval = FALSE}

meanrtindat$Group.3 <- ifelse(meanrtindat$Group.1 > 5, 1, 0)
meanside <- with(meanrtindat, tapply(x, Group.3, mean))
meanside2 <- as.data.frame(t(meanside))
meanside3 <- data.frame(y = c(meanside2$`0`, meanside2$`1`), type = c("Smaller", "Larger"))

indside <- with(meanrtindat, tapply(x, list(Group.3, Group.2), mean))
indside2 <- as.data.frame(t(indside))
indside3 <- data.frame(y = c(indside2$`0`,indside2$`1`), type = rep(c("Smaller", "Larger"), each = nrow(indside2)), group = rep(1:52, 2))

# matplot(indside, type = "b", lty = 1, col = adjustcolor(1, .3), pch = 20
#         , ylim = c(300, 1200), xlim = c(0.8, 2.2), xaxt = "n")
# axis(1, 1:2, labels = c("smaller", "larger"))
# lines(meanside, lwd = 2, type = "b", col = "darkred")

level_order <- c("Smaller", "Larger")
theme_set(theme_apa(base_size = 9))
meanrtplot2 <- ggplot(indside3, aes(y = y, x = factor(type, level_order))) + 
  geom_line(aes(group = group), alpha = .15) + geom_point() + 
  geom_line(data = meanside3, aes(group = 1), color='blue', size = 1) + 
  geom_point(data = meanside3, 
             aes(group = 1), color='blue', shape = 21, size = 2, stroke = 1) + 
  labs(x = "", y = "Response Time (Ms)") + 
  theme_bw() + 
  theme(axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank())

# meanrtplot2 
```

## Why Multilevel Modeling?

Before diving into hierarchical modeling, let us consider two alternative approaches: Analyzing aggregated data and analyzing the data separately for each participant. Analyses of aggregated data such as the ANOVA from the scenario above examine a general effect that is assumed to be consistent across subjects. For instance, when investigating whether there is an effect of digit on RT, the mean RT for each participant is computed for each digit-condition, and a repeated-measures ANOVA is conducted on these aggregated scores. Analyses of aggregated data cannot investigate whether a trend is consistent across participants. Therefore, when this option is applied and individual variability is ignored, the general trend can be an inaccurate representation of the true general trend [Rouder, Morey, & Pratte, 2013; Haaf & Rouder, 2017; van Doorn, Aust, Haaf, Stefan, and Wagenmakers, -@van2021bayes]. This is shown in Panel A of Figure \@ref(fig:figure1) where some of the individual trends (i.e., slopes of the grey lines) differ from the general trend (i.e., slope of the blue line). If every participant takes part in each condition, the person variability affects all conditions, thus inducing a correlation across conditions. For example, participants in a symbolic distance effect experiment repeatedly respond to target digits inducing a correlation between the responses to the target digits, such as between target digit 3 and target digit 4. Therefore, the observations are not independent as we assume if we perform analysis on aggregated data. As a result, the type I error rate of a test of the effect increases: there is a higher probability of concluding that there is an effect if there is no effect (Rouder & Lu, 2005).

A second alternative analysis method is to conduct an analysis on each participant's data independently. For example, in the digit classification task, we would estimate one digit effect per person (cf. the grey lines in Figure 1A), but no overall effect (cf. the blue line in Figure 1A). Therefore, when analyzing participant data independently, it is hard to draw any general conclusions.

In hierarchical analyses, the general or overall effect is called fixed effect, and the individual-specific deviations from this effect are called random effects (Rouder & Lu, 2005). We can also estimate individual effects, not as deviations from an overarching effect, but as effects for each individual, as obtained from a per-participant analysis. In this tutorial, we will use the terms general effect, individual deviation, and individual effects instead of fixed and random effects, because we believe these terms are better suited for the interpretation of our analyses.

If estimated for each participant independently, observed individual effects are quite variable because they are perturbed by sample noise. In hierarchical modeling, these individual effects are optimally and automatically corrected towards the general trend [@efron1977stein]. This phenomenon, also called "shrinkage", is illustrated in Panel B of Figure \@ref(fig:figure1). The rightmost observation in the figure is in sample quite large compared to the other observed effects. The estimated effect (green point) is drastically corrected towards the mean. Conversely, the general trend is only influenced slightly by individuals showing a highly divergent trend. By adding the individual variability to the model, we no longer assume independence between observations, but the overall effect in the model leaves the possibility to generalize the effect. Therefore, hierarchical analysis can be considered a middle ground between an independent analysis and an analysis of aggregated data. 
```{r figure2, echo=FALSE, eval = FALSE, fig.cap="Sample represents the observed response time in milliseconds per individual, while model represents the individual response time estimates. Individual estimates are shrunken towards the general trend. ", fig.align = "left", fig.height = 2, fig.width = 3, out.width="70%"}
knitr::include_graphics("Images/hierplotefmor.png")
```

The Bayesian framework offers an intuitive approach for hierarchical modeling (Rouder & Lu, 2005; Lynch, 2007). The flexibility of the Bayesian approach facilitates the implementation of complex hierarchical models (Rouder, Morey, & Pratte, 2013). It also allows for the inclusion of prior knowledge into the analysis, as well as the monitoring of evidence for specific hypotheses as the data accumulate [@wagenmakers2018bayesian]. A complete discussion of the Bayesion approach is beyond the scope of this manuscript. We refer the interested reader to Etz and Vandekerckhove [-@etz2018introduction] and Wagenmakers and colleagues [-@wagenmakers2018bayesian].

# Model Specification 
In the following section, we specify two models that can be used for our application example: a normal and a log-normal model.^[Online Supplement A provides the parameterization for models of aggregated and individual data, to illustrate the difference with hierarchical models. ] We are aware that many psychologists do not engage with equations on a daily basis. In this section, however, we try to accessibly summarize the key considerations of model specification. We encourage everyone to read on and not skip the section. Additionally, we provide a non-technical verbal summary.

Throughout the section, we assume that each observation $Y_{ijk}$ is the response time from participant $i$ in trial $k$ of condition $j$. The condition $j$ is the digit presented to the participant, and can take the values 2, 3, 4, 6, 7, and 8. We further assume that the digit side ($x$) is recorded using effect coding where a value of -0.5 indicates that the digit is smaller than 5, and a value of 0.5 indicates that the digit is larger than 5. Additionally, we assume that the digits presented in the trial are dummy-coded in the data using four variables ($u$, $v$, $w$, $z$). If digits 2 and 8 were presented all four variables are zero. The dummy variables represent the digits 3 ($u$), 4 ($v$), 6 ($w$), and 7 ($z$), and each take the value 1 if the respective digit was presented in the trial, and 0 if it was not.

## Normal Model

In the normal model, the observation $Y_{ijk}$ is assumed to come from a normal distribution. The mean of this normal distribution is predicted by the person-specific intercept $\gamma$, the effect of the side (i.e., smaller or greater than 5, $\beta$) and the effect of the distance from 5 (i.e., symbolic distance effect, $\delta$’s). The variance is represented by $\sigma^2$. This is captured in the following equation:

\begin{equation}
Y_{ijk} \sim \text{Normal}(\gamma_{i} + x_j \beta_{i} + u_j \delta_{7i} + v_j \delta_{6 i} + w_j \delta_{4i} + z_j \delta_{3i}, \sigma^2), (\#eq:datalevel)
\end{equation}

where

<!-- \[ -->
<!-- x_j =  -->
<!-- \begin{cases} -->
<!-- \frac{1}{2} & j < 5\\ -->
<!-- - \frac{1}{2} & j > 5, \\ -->
<!-- \end{cases} -->

<!-- \; \; \; -->

<!-- \[ -->
<!-- u_j =  -->
<!-- \begin{cases} -->
<!-- 1 & j = 7\\ -->
<!-- 0 & \text{Otherwise}, -->
<!-- \end{cases} -->

<!-- \; \; \; -->

<!-- v_j =  -->
<!-- \begin{cases} -->
<!-- 1 & j = 6\\ -->
<!-- 0 & \text{Otherwise}, -->
<!-- \end{cases} -->
<!-- \] -->

\begin{align*}
&
\begin{aligned}
x_j = 
\begin{cases}
\frac{1}{2} & j < 5\\
- \frac{1}{2} & j > 5, \\
\end{cases}
\end{aligned}
&
\begin{aligned}
u_j = 
\begin{cases}
1 & j = 7\\
0 & \text{Otherwise},
\end{cases}
\end{aligned}
&
\begin{aligned}
v_j = 
\begin{cases}
1 & j = 6\\
0 & \text{Otherwise},
\end{cases}
\end{aligned}
\end{align*}

<!-- \[ -->
<!-- w_j =  -->
<!-- \begin{cases} -->
<!-- 1 & j = 4\\ -->
<!-- 0 & \text{Otherwise}, -->
<!-- \end{cases} -->

<!-- \; \; \; -->

<!-- z_j =  -->
<!-- \begin{cases} -->
<!-- 1 & j = 3\\ -->
<!-- 0 & \text{Otherwise}. -->
<!-- \end{cases} -->
<!-- \] -->

\begin{align*}
&
\begin{aligned}
w_j = 
\begin{cases}
1 & j = 4\\
0 & \text{Otherwise},
\end{cases}
\end{aligned}
&
\begin{aligned}
z_j = 
\begin{cases}
1 & j = 3\\
0 & \text{Otherwise}.
\end{cases}
\end{aligned}
\end{align*}
The model equation shows that, for instance, for participant *i* = 1 and digit *j* = 7, the mean RT is modeled by: $\gamma_{i = 1} - \frac{1}{2} \beta_{i = 1} + \delta_{7,i = 1}$. If all $\delta$s are positive, then the model equation indicates that individuals respond fastest when the digits 2 and 8 are presented, as in this situation there are no digit effects in the equation.

The intercept $\gamma_i$, the side effect $\beta$, and the digit effects $\delta$ are person-specific. This means that for these parameters, the model will produce one estimate for each participant. The multilevel model assumes that these individual-level effects in turn follow a normal distribution:

\begin{equation}
\begin{aligned}
\gamma_{i} &\sim \text{Normal}(\mu_{\gamma}, \sigma^2_{\gamma}), (\#eq:fixedvsrandom)\\
\beta_{i} &\sim \text{Normal}(\mu_{\beta}, \sigma^2_{\beta}), \\
\delta_{7i} &\sim \text{Normal}(\mu_{\delta_{7}}, \sigma^2_{\delta_{7}}), \\
\delta_{6i} &\sim \text{Normal}(\mu_{\delta_{6}}, \sigma^2_{\delta_{6}}), \\
\delta_{4i} &\sim \text{Normal}(\mu_{\delta_{4}}, \sigma^2_{\delta_{4}}),\\
\delta_{3i} &\sim \text{Normal}(\mu_{\delta_{3}}, \sigma^2_{\delta_{3}}).\\
\end{aligned}
\end{equation}
The means ($\mu_{\gamma}$, $\mu_{\beta}$, $\mu_{\delta}$) of these normal distributions represent the general expected effect, for example, $\mu_{\beta}$ represents the expected effect of digit side across individuals. The variances ($\sigma^2_{\gamma}$, $\sigma^2_{\beta}$, $\sigma^2_{\delta}$) represent the individual variation. Therefore, the individual level depends on another level, the overall effect level, containing $\mu_{\gamma}$, $\mu_{\beta}$, $\mu_{\delta}$, $\sigma^2_{\gamma}$, $\sigma^2_{\beta}$, and $\sigma^2_{\delta}$.

## Log-normal Model 
A normal model may not be appropriate in case of more complex theories, hypotheses, or data structures. For example, the data in the example are response times. The normal model, however, can yield negative values, which seems very problematic. In addition, RT distributions are right-skewed. A better representation of the data at hand is therefore given by a log-normal model [@schramm2019reaction]. This means that Equation \@ref(eq:datalevel) has to be adjusted:

\begin{equation}
Y_{ijk} \sim \text{LogNormal}(\gamma_{i} + x_j \beta_{i} + u_j \delta_{7i} + v_j \delta_{6i} + w_j \delta_{4i} + z_j \delta_{3i}, \sigma^2).  
\end{equation}
Note that the only change in the model formulation is the change of the likelihood function from a normal distribution to a log-normal distribution. The location parameter of the log-normal distribution is described through the same combination of parameters as the mean of the normal model. However, it has a different interpretation. For example, for a location parameter of $\mu = 0$ and a scale parameter of $\sigma^2 = 1$, the mode of the log-normal distribution is $0.368$, and the mean is $1.65$.

## Summary

In summary, the models are placed directly on the raw RT data. The only difference between the normal and the log-normal models is the likelihood distribution. The normal model just assumes a symmetric normal distibution that can take on any value. The log-normal model assumes a log-normal distribution, which is restricted to positive values and has a right skew.

Both the normal and the log-normal models contain the following parameters: $\gamma_i$ corresponds to the participant's overall response time, $\beta_i$ corresponds to the participant's side effect, and the four $\delta_{\cdot i}$ correspond to the effects between the digits furthest away from 5 (2 or 8) and the other digits. These person-specific effects all come from parent distributions. This results in the hierarchical structure shown in Figure \@ref(fig:hierarchyplot). The figure highlights the relationship between the parameters and the data.

In Bayesian modeling, prior distributions are needed to complete model specification. In the next section, we will take you through the process of setting these priors, followed by sections on Bayesian estimation and hypothesis testing for both the normal and log-normal model. In every section, we will first explain the normal model, followed by an illustration of the log-normal model.

```{r hierarchyplot, echo=FALSE, fig.cap="Hierarchical structure of the normal and lognormal models for the number classification task.", fig.align = "left", fig.height = 3, fig.width = 4, out.width="70%"}
knitr::include_graphics("Images/Fig2_Hierarchical structure.png")
```

# Prior Settings 
Bayesian analysis requires the specification of prior distributions. Prior distributions are probability distributions on model parameters that specify beliefs about the relative plausibility of parameter values before seeing the data [@wagenmakers2018bayesian]. These prior beliefs can then be updated with the data to obtain the posterior beliefs about the model parameters, following Bayes' rule [@jeffreys1961theory].

For our models, priors are needed on the general effects ($\mu_{\gamma}$, $\mu_{\beta}$, $\mu_{\delta}$) and on the variances ($\sigma_{\gamma}^2$, $\sigma_{\beta}^2$, $\sigma_{\delta}^2$, and $\sigma^2$), as shown in Equation \@ref(eq:fixedvsrandom). For each of these parameters, we specify our beliefs as a probability distribution. For example, if we expect small side effects in the digit classification example, the prior distribution on $\mu_{\beta}$ should have its peak close to zero. Large values for the variance parameters indicate large individual differences, whereas values close to zero indicate that effects are similar across individuals [@haaf2017developing]. If we expect little variability of side effects between people, the prior assigned to $\sigma_{\beta}$ should also have its peak close to zero. The width of prior distributions expresses how certain we are about the parameters before seeing the data. For example, if we want to express that we are certain that the side effect is zero, we can formulate a prior that allocates all probability mass to a parameter value of zero.

## Fear of Commitment  
In many cases substantive researchers fear committing to prior settings because they are unsure about their choices for the type and settings of the distribution. Therefore, they tend to choose wide priors to indicate little prior knowledge about the parameter values [@aczel2018expert]. These include default priors offered by programming packages, such as `rstan` (Stan Development Team, 2019a; illustrated in Online Supplement B) and `brms` [@R-brms_a; @R-brms_b]. Wide priors are priors that allocate (approximately) equal probability density to a great range of values, such as a uniform distribution without bounds [@gelman2006prior]. However, wide priors can be problematic. First, since the distribution is spread out across a wide range of values, the probability density on any specific parameter value is very low. Therefore, any effect is unlikely under the prior as a wide range of effects is deemed plausible. This influences model comparisons: the support for the null hypothesis becomes increasingly large. This is also referred to as the Jeffreys-Lindley paradox [@lindley1957statistical].

Another important issue is that wide priors are often improper, that is, the probability density does not integrate to one [@hobert1996effect]. An example of an improper prior is a uniform prior distribution ranging from minus infinity to infinity. In case of improper priors, an estimation of the general and individual effects cannot be provided. However, even proper priors can result in improper posteriors [@hobert1996effect]. Although estimation should not be possible in this situation, statistical software might still provide results without notifying the user that the posterior distribution does not exist. If users are not aware of this problem, this can result in misleading conclusions. Therefore, it is important to check if *(a)* the priors are proper and *(b)* the priors result in proper posteriors. There is no fail-safe method that guarantees proper posteriors for hierarchical models. The best approach is to choose well-reasoned informative priors for the model, for instance, by using the proposed four-step procedure below.

## Procedure 
We propose four steps to choose suitable priors for Bayesian hierarchical models. We first present the steps and then apply them to choose suitable priors for the parameters $\mu_{\gamma}$, $\mu_{\beta}$, $\mu_{\delta}$, $\sigma_{\gamma}^2$, $\sigma_{\beta}^2$, $\sigma_{\delta}^2$ and $\sigma^2$ in our example.

```{r distributionsteps, echo=FALSE, fig.cap = "Prior specification steps.", fig.align='center', warning=FALSE, message=FALSE, fig.height=3, fig.width=4, cache = TRUE}
# https://jokergoo.github.io/circlize_book/book/graphics.html 
library(circlize)
library(unikn)   # https://cran.r-project.org/web/packages/unikn/vignettes/colors.html

# circos.initialize(letters[1:4], xlim = c(0, 1))
# col = rand_color(4)
# tail = c("point", "normal", "point", "normal")
# circos.track(ylim = c(0, 1), panel.fun = function(x, y) {
#     circos.arrow(x1 = 0, x2 = 1, y = 0.5, width = 0.4, 
#         arrow.head.width = 0.6, arrow.head.length = ux(1, "cm"), 
#         col = col[CELL_META$sector.numeric.index], 
#         tail = tail[CELL_META$sector.numeric.index])
# }, bg.border = NA, track.height = 0.4)

circos.clear()
cell_cycle = data.frame(phase = c("1. Distribution family", "2. Distribution settings", "3. Prior predictions", "4. Adjust"), levels = c("Distribution type", "Distribution settings", "Prior predictions", "Adjust"),
                        hour = c(8, 8, 8, 8))
color = c(NA)
circos.par(start.degree = 90)
circos.initialize(cell_cycle$phase, xlim = cbind(rep(0, 4), cell_cycle$hour))
circos.track(ylim = c(0, 1), panel.fun = function(x, y) {
    circos.arrow(CELL_META$xlim[1], CELL_META$xlim[2], 
        arrow.head.width = CELL_META$yrange*0.8, arrow.head.length = ux(0.5, "cm"),
        col = color[CELL_META$sector.numeric.index])
    circos.text(CELL_META$xcenter, CELL_META$ycenter, CELL_META$sector.index, 
        facing = "bending.inside", niceFacing = TRUE, cex = 0.6)
}, bg.border = NA, track.height = 0.4)
```

## Step 1: Distribution Family  
The first step in specifying a prior distribution is to find a suitable distribution family. For instance, the prior plausibility of parameter values can be expressed through a normal, an inverse gamma, or a Student's *t*-distribution. These distributional families have different shapes and a different support, as illustrated in Figure \@ref(fig:distritypeplot).  For example, the inverse gamma distribution is limited to non-negative parameter values and has a positive skew. The normal and Student's *t*-distribution both have support over the real line, however, the Student's *t*-distribution has wider tails, that is, it assigns a higher plausibility to values that are on the extremes of the distribution.

The support of the prior distribution should cover all theoretically possible parameter values. For example, if a parameter cannot have negative values, such as a standard deviation, the inverse gamma distribution would be a suitable choice. Conversely, if a parameter can take positive and negative values, the normal and Student *t*-distribution would be suitable.

Note that although distributions within a distributional family share the same functional form, they can differ a lot in their shape and location. The exact appearance of a prior distribution is controlled by a set of hyperparameters, for example, the mean and variance for the location and scale of a normal prior. In the next section, we will give some guidance for specifying these hyperparameters.

```{r distritypeplot, echo=FALSE, fig.cap="Common probability density distributions that may be used as priors. Panel A shows normal distributions varying in mean and standard deviation. Panel B shows inverse gamma distributions that differ in the shape and scale parameter. Panel C shows Student's \\textit{t}-distributions that differ in the location and scale parameter. For all three \\textit{t}-distributions, the degrees of freedom are set to 3.", fig.align = "left", fig.height = 8, fig.width = 12, out.width="100%"}
knitr::include_graphics("Images/Fig4_distrtypefigure_v3.png")
```

## Step 2: Set Hyperparameters  
Location, scale, and sometimes shape of the selected distribution can be adjusted by choosing adequate hyperparameters. The goal is to adjust the prior to reflect which parameter values we deem most plausible and how much uncertainty there is. For instance, panel A of Figure \@ref(fig:distritypeplot) illustrates, that increasing the mean in the normal distribution leads to a shift of location: higher parameter values are deemed more plausible according to the pink distribution compared to the green distribution. Panel B shows how uncertainty can be specified, for example, by decreasing the shape of the inverse gamma distribution while keeping the scale constant: the orange distribution assigns probability to a wider range of values than the green distribution.

For this step, it is important to consider the scale on which variables are measured. For instance, RTs could be measured in seconds or in milliseconds. This means that if a researcher expects a response time of one second, they will formulate a prior distribution with a mean of 1 if RT is measured in seconds, or formulate a prior distribution with a mean of 1000 if RT is measured in milliseconds. Similarly, the measurement scale will also affect the width of the specified prior.

## Step 3: Prior Predictions 
A key advantage of model specification with priors is that we can obtain preditions on what the data would look like that are generated from the model. These prior predictions allow us to make the implications of our specified priors more concrete. Researchers can look at relevant statistic that summarize the simulated data and decide if they correspond to reasonable observations for their specific research context. If the simulated data are in line with the researchers' expectations, the priors can be used for the analysis. If the simulated data do not match the expectations, the priors have to be adjusted (Step 4).

```{r priorpredfigurepanel, echo=FALSE, fig.cap="Side and digit effects from predicted data. Violins represent the distribution of predicted effects across 1000 simulation runs. The blue points are the average effects across simultions. Panel A shows the prior predictions for the normal model based on the default priors of \\textit{brms}. Panel B shows the prior predictions for the normal model based on the selected priors. Panel C shows the prior predictions for the lognormal model based on the priors chosen for the normal model. Panel D shows the prior predictions for the lognormal model based on the selected priors.", fig.align = "left", fig.height = 3, fig.width = 4, out.width="100%"}
knitr::include_graphics("Images/Fig5_priorpredfigure_v2.png")
```

The simulation proceeds as follows: First, one random draw from each prior distribution is taken. This is plugged into the individual effect distributions (Equation (2)). Then, the likelihood of the model, such as specified in Equations (1) and (3), is used to simulate observations per condition. This means that for *n* participants, *n* random values for the model parameters are generated based on the priors, which will result in the *k* observations per *j* conditions. This process is repeated *m* times, where *m* is the number of simulations drawn. Note that the more simulations are conducted, the closer one gets to the full picture because enough samples from the prior distribution are needed. Finally, we can visualize the simulated data. Figure \@ref(fig:priorpredfigurepanel) shows prior predictions for the key effects from the number classification task. For this task, we mainly care about the side effect and digit effects corresponding to the $\beta$s and the $\delta$s from the model. Here, the prior predictions for the average side effect and the average digit effect across participants are depicted.

For step 3, the intention is not to compare the actual data of interest to the data predicted by the priors. Instead, the researchers' *expectations* of plausible data are compared to the prior predictions. Therefore, prior predictive checks are conducted *before* looking at the actual data. If the prior predictions match the expectations, the priors can be used for the analysis. If they do not match the expectations, the priors need to be adjusted.

## Step 4: Adjust 
Based on the results of the prior predictions, it might be necessary to adjust the priors. Typically, the focus of adjustments will be on finding adequate parameter values for the prior distribution, that is, on repeating step 2 and the following steps of the prior specification process.

In the following two sections, we will provide a walk-through of all four steps of prior specification for our two models of the symbolic distance effect.

## Normal Model

In the symbolic distance model, we have to specify priors for the parameters $\mu_{\gamma}$, $\mu_{\beta}$, $\mu_{\delta}$, $\sigma^2_{\gamma}$, $\sigma^2_{\beta}$, and $\sigma^2_{\delta}$. The parameters $\mu_{\gamma}$, $\mu_{\beta}$, and $\mu_{\delta}$ can be positive or negative. All variance terms ($\sigma_{\gamma}^2$, $\sigma_{\beta}^2$, $\sigma_{\delta}^2$, and $\sigma^2$) are constrained to be positive. If there is no information from previous studies, researchers might be inclined to use the default priors in `brms` as a starting point.

In `brms`, different default priors are set on different types of parameters [@R-brms_a]. The prior on the group-level regression weights ($\mu_{\beta}$, $\mu_{\delta}$) is a normal distribution with a mean of 0 and a standard deviation of 1. The prior on the intercept $\mu_{\gamma}$ is a scaled and shifted Student's *t*-distribution with 3 degrees of freedom, a location parameter of 1, and a scale parameter of 10. `brms` specifies priors on standard deviations ($\sqrt{\sigma^2}$) instead of variances. For the normal model, this means that priors are set on $\sigma_{\gamma}$, $\sigma_{\beta}$, $\sigma_{\delta}$ and $\sigma$, instead of on $\sigma_{\gamma}^2$, $\sigma_{\beta}^2$, $\sigma_{\delta}^2$ and $\sigma^2$. The default prior on the standard deviations is a central, scaled Student's *t*-distribution with 3 degrees of freedom and a scale parameter of 10 that is truncated at zero. To summarize, the default priors would be set the following way:

\begin{equation}
\begin{aligned}
\mu_{\gamma} &\sim \text{Student's t}(3, 1, 10), (\#eq:chosenpriors1) \\ 
\mu_{\beta}, \mu_{\delta} &\sim \text{Normal}(0, 1), \\ 
\sigma, \sigma_{\gamma}, \sigma_{\beta}, \sigma_{\delta} &\sim \text{Truncated-Student's t}_{+}(3, 0, 10). 
\end{aligned}
\end{equation}
The prior distributions `brms` specifies on the standard deviations can be transformed to prior distributions on variances. Prior distributions on variances have a smaller scale than priors on standard deviation. Note, however, that the transformation of the prior distributions is not as simple as the transformation of the parameter itself. In our Online Supplement, we provide R-code with a numeric solution to the transformation. For the `brms` priors, we obtain the following implied prior distributions on variances:

\begin{equation}
\begin{aligned}
\sigma_{\gamma}^2, \sigma_{\beta}^2, \sigma_{\delta}^2, \sigma^2 &\sim \text{Student's t}(3, 0, 1.48). 
\end{aligned}
\end{equation}

The priors are visualized on the right side of Figure \@ref(fig:priorvisualizationbrms2). Note that these prior distributions are on the group-level parameters and represent the uncertainty about the general effect. The uncertainty about individual effects is represented by the marginal prior distribution on the left. This distribution takes the prior variability of the mean and variance as well as the implied individual variability into account.

```{r priorvisualizationbrms2, fig.cap="Visualization of \\textit{brms} default prior settings for the normal model. On the right, the prior distributions on general effects parameters are shown. On the left, the implied distribution of the individual effects for every parameter based on these priors are presented. ", fig.align = "left", message=FALSE, warning=FALSE, cache = TRUE}
# library(crch)  for truncated student's t

# Plots 
x <- seq(-3, 3, by=.01)
# color codes: https://flaviocopes.com/rgb-color-codes/

# Instead do this, take integral  
Nplotst <- 1000000
mugammaplotbrms <- rst(Nplotst, nu = 3, mu = 1, sigma = 10)
vargammaplotbrms <- rtt(Nplotst, df = 3, location = 0, scale = 10, left = 0, right = Inf)  # should be truncated 
betaplot2brms <- rnorm(Nplotst, mugammaplotbrms, sqrt(vargammaplotbrms))

pgamma2databrms <- data.frame(betaplot2brms)
pgamma2brms <- ggplot(pgamma2databrms, aes(x = betaplot2brms)) + geom_density_line(color = "#0072B2", fill = "#0072B2", alpha = 0.1, position = "identity") + ylab("Density") + xlab("") + labs(title = expression(~gamma[i])) + theme(plot.title = element_text(size = 14)) + xlim(-20, 20)

parametersgammabrms <- c("Mean" = "#D95F02", "Variance" = "#1B9E77")

ptogethergamma2brms <- ggplot(data = data.frame(x = c(-20, 20)), aes(x)) +
  stat_function(fun = dst, n = 1001, args = list(nu = 3, mu = 1, sigma = 10), aes(color = "Mean", fill = "Mean"), geom = "area", alpha = 0.3) + ylab("") + xlab("") + 
  stat_function(fun = dtt, n = 1001, args = list(df = 3, location = 0, scale = 10, left = 0, right = Inf), geom = "area", aes(color = "Variance", fill = "Variance"), alpha = 0.3, position = "identity") + geom_hline(yintercept=0, colour="white", size=1) + scale_color_manual(name = "Prior on", values = parametersgammabrms, aesthetics = c("colour", "fill"), labels = c(expression(~mu[~gamma]), expression(~sigma[~gamma]))) + 
  theme(legend.text=element_text(size = 9), legend.title = element_text(size = 7), legend.key.width=unit(0.45,"cm"), legend.key.height=unit(0.5,"cm"))

# Beta 
mubetaplotbrms <- rnorm(Nplotst, 0, 1)
varbetaplotbrms <- rtt(Nplotst, df = 3, location = 0, scale = 10, left = 0, right = Inf)  
betaplot22brms <- rnorm(Nplotst, mubetaplotbrms, sqrt(varbetaplotbrms))

pbeta2databrms <- data.frame(betaplot22brms)
pbeta2brms <- ggplot(pbeta2databrms, aes(x = betaplot22brms)) + geom_density_line(color = "#0072B2", fill = "#0072B2", alpha = 0.1, position = "identity") + ylab("Density") + xlab("") + labs(title = expression(~beta["i"])) + theme(plot.title = element_text(size = 14)) + xlim(-10, 10)

ptogetherbeta2brms <- ggplot(data = data.frame(x = c(-20, 20)), aes(x)) +
  stat_function(fun = dnorm, n = 1001, args = list(0, 1), aes(color = "Mean", fill = "Mean"), geom = "area", alpha = 0.3) + ylab("") + xlab("") + 
  stat_function(fun = dtt, n = 1001, args = list(df = 3, location = 0, scale = 10, left = 0, right = Inf), geom = "area", aes(color = "Variance", fill = "Variance"), alpha = 0.3, position = "identity") + geom_hline(yintercept=0, colour="white", size=1) + scale_color_manual(name = "Prior on", values = parametersgammabrms, aesthetics = c("colour", "fill"), labels = c(expression(~mu[~beta]), expression(~sigma[~beta]))) + 
  theme(legend.text=element_text(size = 9), legend.title = element_text(size = 7), legend.key.width=unit(0.45,"cm"), legend.key.height=unit(0.5,"cm"))

# Delta's

mudeltaplotbrms <- rnorm(Nplotst, 0, 1)
vardeltaplotbrms <- rtt(Nplotst, df = 3, location = 0, scale = 10, left = 0, right = Inf)  
deltaplot2brms <- rnorm(Nplotst, mudeltaplotbrms, sqrt(vardeltaplotbrms))

pdelta2databrms <- data.frame(deltaplot2brms)
pdelta2brms <- ggplot(pdelta2databrms, aes(x = deltaplot2brms)) + geom_density_line(color = "#0072B2", fill = "#0072B2", alpha = 0.1, position = "identity") + ylab("Density") + xlab("Individual effect") + labs(title = expression(~delta[i])) + theme(plot.title = element_text(size = 14)) + xlim(-10, 10)

ptogetherdelta2brms <- ggplot(data = data.frame(x = c(-20, 20)), aes(x)) +
  stat_function(fun = dnorm, n = 1001, args = list(0, 1), aes(color = "Mean", fill = "Mean"), geom = "area", alpha = 0.3) + ylab("") + xlab("Parameter value") + 
  stat_function(fun = dtt, n = 1001, args = list(df = 3, location = 0, scale = 10, left = 0, right = Inf)  , geom = "area", aes(color = "Variance", fill = "Variance"), alpha = 0.3, position = "identity") + geom_hline(yintercept=0, colour="white", size=1) + scale_color_manual(name = "Prior on", values = parametersgammabrms, aesthetics = c("colour", "fill"), labels = c(expression(~mu[~delta]), expression(~sigma[~delta]))) + 
  theme(legend.text=element_text(size = 9), legend.title = element_text(size = 7), legend.key.width=unit(0.45,"cm"), legend.key.height=unit(0.5,"cm"))

theme_set(theme_apa(base_size = 9))
#ptogether

# Final figure 
grid.arrange(pgamma2brms, ptogethergamma2brms, pbeta2brms, ptogetherbeta2brms, pdelta2brms, ptogetherdelta2brms, nrow = 3, ncol = 2, widths = c(0.4, 0.6))
```

We will continue with step 3 of the prior specification process and check what the expected data would look like according to the selected priors. The results of the prior prediction for the symbolic distance effect with the priors from Equation \@ref(eq:chosenpriors1) with the number of participants $n = 10$, the number of trials per condition $k = 60$, and the number of simulations $m = 1000$ are shown in panel A of Figure \@ref(fig:priorpredfigurepanel). Code for the simulations can be found in our Online Supplement.

The figure shows the difference in RTs between digits for every repetition averaged across subjects (i.e., 1000 grey dots per effect). The blue dots represent the mean effect. Using the default prior from `brms`, the mean effects are expected to be around zero, with a mean effect of zero. However, the range around this mean is very wide, ranging from -10 to 10 second effects. Most of the trials are in a smaller range of -5 to 5 second effects. This means that according to our priors, we would expect the effects to fall in this range. For this task, where participants have to respond as fast as possible, such large RT effects are unlikely. Therefore, it seems reasonable to adjust the priors.

```{r priorvisualization2, fig.cap="Visualization of the adjusted prior settings for the normal model. On the right, the prior distributions on general effects parameters are shown. On the left, the implied distribution of the individual effects for every parameter based on these priors are presented. ", fig.align = "left", fig.height = 6, fig.width = 8, warning=FALSE, message=FALSE, cache = TRUE}
### Set the priors 
a <- 0.5  # mean for mu, mu is the mean of theta  
b <- 1  # variance for mu 
c <- 3  # mean for sigma 
d <- .7  # variance for sigma, sigma is the variance of observations  
e <- 3  # mean for g, g is the variance of theta   
f <- .7  # variance for g

#### Priors for beta 
a2 <- 0  # mean for mean of beta
b2 <- 0.09  # variance for mean of beta, used to be 1 
e2 <- 3  # mean for variance of beta, used to be .7 
f2 <- 0.5  # variance for variance of beta, used to be .2.

#### Prior for deltas, same for every delta 
a3 <- 0  # mean for mean of delta 
b3 <- 0.09  # sd for mean of delta, 0.5, 0.3  
dd3 <- 3  # mean for variance of delta 
f3 <- 0.5   # variance for variance of delta

# Plots 
x <- seq(-3, 3, by=.01)
# color codes: https://flaviocopes.com/rgb-color-codes/

## Gamma 
mu <- rnorm(10, a, b)
g <- rinvgamma(10, e, f)
gammaplot <- dnorm(x, mu, g)  # I think this is not possible. Get a value for mu and g for each x --> 601

# Instead do this, take integral  
Nplot <- 1000000
mugammaplot <- rtruncnorm(Nplot, a = 0, mean = 0.5, sd = 1)
vargammaplot <- rinvgamma(Nplot, e, f)
betaplot2 <- rnorm(Nplot, mugammaplot, sqrt(vargammaplot))
# plot(density(betaplot2, na.rm = TRUE))

pgamma2data <- data.frame(betaplot2)
pgamma2 <- ggplot(pgamma2data, aes(x = betaplot2)) + geom_density_line(color = "#0072B2", fill = "#0072B2", alpha = 0.1, position = "identity") + ylab("Density") + xlab("") + labs(title = expression(~gamma[i])) + theme(plot.title = element_text(size = 14)) + xlim(-10, 10)

parametersgamma <- c("Mean" = "#D95F02", "Variance" = "#1B9E77")

ptogethergamma <- ggplot(data = data.frame(x = c(-5, 5)), aes(x)) +
  stat_function(fun = dnorm, n = 1001, args = list(mean = a, sd = b), geom = "area", color = "#D95F02", fill = "#D95F02", alpha = 0.3) + ylab("") +
  stat_function(fun = dinvgamma, n = 1001, args = list(shape = e, scale = f), geom = "area", color = "#1B9E77", fill = "#1B9E77", alpha = 0.3, position = "identity") + geom_hline(yintercept=0, colour="white", size=1) + scale_color_manual(name = "Prior parameters gamma", values = parametersgamma) + 
  scale_fill_manual(name = "Bar",values = parametersgamma)

ptogethergamma2 <- ggplot(data = data.frame(x = c(-5, 5)), aes(x)) +
  stat_function(fun = dtruncnorm, n = 1001, args = list(a = 0, mean = 0.5, sd = 1), aes(color = "Mean", fill = "Mean"), geom = "area", alpha = 0.3) + ylab("") + xlab("") + 
  stat_function(fun = dinvgamma, n = 1001, args = list(shape = e, scale = f), geom = "area", aes(color = "Variance", fill = "Variance"), alpha = 0.3, position = "identity") + geom_hline(yintercept=0, colour="white", size=1) + scale_color_manual(name = "Prior on", values = parametersgamma, aesthetics = c("colour", "fill"), labels = c(expression(~mu[~gamma]), expression(~sigma[~gamma]^2))) + 
  theme(legend.text=element_text(size = 9), legend.title = element_text(size = 7), legend.key.width=unit(0.45,"cm"), legend.key.height=unit(0.5,"cm")) + xlim(-1, 1)

# Beta 
mu2 <- dnorm(x, a2, b2)
g2 <- dinvgamma(x, e2, f2)
betaplot <- dnorm(x, mu2, g2)

mubetaplot <- rnorm(Nplot, a2, b2)
varbetaplot <- rnorm(Nplot, e2, f2)
betaplot22 <- rnorm(Nplot, mubetaplot, sqrt(varbetaplot))
# plot(density(betaplot22, na.rm = TRUE))

pbeta2data <- data.frame(betaplot22)
pbeta2 <- ggplot(pbeta2data, aes(x = betaplot22)) + geom_density_line(color = "#0072B2", fill = "#0072B2", alpha = 0.1, position = "identity") + ylab("Density") + xlab("") + labs(title = expression(~beta["i"])) + theme(plot.title = element_text(size = 14)) + xlim(-10, 10)

ptogetherbeta2 <- ggplot(data = data.frame(x = c(-5, 5)), aes(x)) +
  stat_function(fun = dnorm, n = 1001, args = list(mean = a2, sd = b2), aes(color = "Mean", fill = "Mean"), geom = "area", alpha = 0.3) + ylab("") + xlab("") + 
  stat_function(fun = dinvgamma, n = 1001, args = list(shape = e2, scale = f2), geom = "area", aes(color = "Variance", fill = "Variance"), alpha = 0.3, position = "identity") + geom_hline(yintercept=0, colour="white", size=1) + scale_color_manual(name = "Prior on", values = parametersgamma, aesthetics = c("colour", "fill"), labels = c(expression(~mu[~beta]), expression(~sigma[~beta]^2))) + 
  theme(legend.text=element_text(size = 9), legend.title = element_text(size = 7), legend.key.width=unit(0.45,"cm"), legend.key.height=unit(0.5,"cm")) + xlim(-1, 1)

# Delta's 
mu3 <- dnorm(x, a3, b3)
g3 <- dinvgamma(x, dd3, f3)
deltaplot <- dnorm(x, mu3, g3)

mudeltaplot <- rnorm(Nplot, a3, b3)
vardeltaplot <- rnorm(Nplot, dd3, f3)
deltaplot2 <- rnorm(Nplot, mudeltaplot, sqrt(vardeltaplot))

pdelta2data <- data.frame(deltaplot2)
pdelta2 <- ggplot(pdelta2data, aes(x = deltaplot2)) + geom_density_line(color = "#0072B2", fill = "#0072B2", alpha = 0.1, position = "identity") + ylab("Density") + xlab("Individual Effect") + labs(title = expression(~delta[i])) + theme(plot.title = element_text(size = 14)) + xlim(-10, 10)

ptogetherdelta2 <- ggplot(data = data.frame(x = c(-5, 5)), aes(x)) +
  stat_function(fun = dnorm, n = 1001, args = list(mean = a3, sd = b3), aes(color = "Mean", fill = "Mean"), geom = "area", alpha = 0.3) + ylab("") + xlab("Parameter Value") + 
  stat_function(fun = dinvgamma, n = 1001, args = list(shape = dd3, scale = f3), geom = "area", aes(color = "Variance", fill = "Variance"), alpha = 0.3, position = "identity") + geom_hline(yintercept=0, colour="white", size=1) + scale_color_manual(name = "Prior on", values = parametersgamma, aesthetics = c("colour", "fill"), labels = c(expression(~mu[~delta]), expression(~sigma[~delta]^2))) + 
  theme(legend.text=element_text(size = 9), legend.title = element_text(size = 7), legend.key.width=unit(0.45,"cm"), legend.key.height=unit(0.5,"cm")) + xlim(-1, 1)

theme_set(theme_apa(base_size = 9))
# ptogether

# Final figure 
grid.arrange(pgamma2, ptogethergamma2, pbeta2, ptogetherbeta2, pdelta2, ptogetherdelta2, nrow = 3, ncol = 2, widths = c(0.4, 0.6))

```

The normal distribution on $\mu_{\beta}$ and $\mu_{\delta}$ seems reasonable if we assume that a two-sided hypothesis about the symbolic distance effect is tested. However, we expect less variance around this mean. Therefore, we decrease the setting for the variation. The wide tails of the Student's *t*-distribution could be one of the reasons for the high range in variation. We change the prior for $\mu_{\gamma}$ to a normal distribution that has more narrow tails. As RTs cannot be negative, we adjust the prior to be positive only by truncating the normal distribution using a lower bound of zero. Finally, we would like to use a more common conjugate prior for the variances to ensure obtaining a proper posterior distribution. ^[Conjugate priors result in posterior distributions that belong to the same distribution family as the prior distribution [@Fink1997]. For a normal likelihood, these are the normal distribution for the mean and the inverse gamma distribution for the variance.] Therefore, we choose an inverse gamma distribution. We can also be more specific about the expected RTs. Based on our knowledge of these types of tasks [e.g., @haaf2017developing], it seems highly plausible that participants will respond within a second or two. We would expect effects in the range of 10 milliseconds to 100 milliseconds. We adjust our priors using these new distribution types and settings, evaluate the prior predictions (shown in panel B of Figure \@ref(fig:priorpredfigurepanel)) and arrive at the following priors:^[Because response times cannot be negative the prior for the overall baseline $\mu_\gamma$ is a truncated normal with a lower bound of zero.]

\begin{equation}
\begin{aligned}
\mu_{\gamma} &\sim \text{Truncated-Normal}_{+}(0.5, 1), (\#eq:chosenpriors2var) \\
\mu_{\beta} &\sim \text{Normal}(0, 0.09), \\
\mu_{\delta} &\sim \text{Normal}(0, 0.09), \\
\sigma^2 &\sim \text{Inverse-Gamma}(3, 0.7),  \\
\sigma^2_{\gamma} &\sim \text{Inverse-Gamma}(3, 0.7), \\   
\sigma^2_{\beta} &\sim \text{Inverse-Gamma}(3, 0.5), \\   
\sigma^2_{\delta} &\sim \text{Inverse-Gamma}(3, 0.5).    
\end{aligned}
\end{equation}
Setting equivalent prior distributions on the standard deviations instead of on the variances results in the following priors:

\begin{equation}
\begin{aligned}
\mu_{\gamma} &\sim \text{Truncated-Normal}_{+}(0.5, 1), (\#eq:chosenpriors2sd) \\
\mu_{\beta} &\sim \text{Normal}(0, 0.3), \\ 
\mu_{\delta} &\sim \text{Normal}(0, 0.3), \\  
\sigma &\sim \text{Inverse-Gamma}(13.8, 6.3),  \\ 
\sigma_{\gamma} &\sim \text{Inverse-Gamma}(13.8, 6.3), \\   
\sigma_{\beta} &\sim \text{Inverse-Gamma}(13.8, 5.3), \\    
\sigma_{\delta} &\sim \text{Inverse-Gamma}(13.8, 5.3).     
\end{aligned}
\end{equation}
The adjusted priors are visualized in Figure \@ref(fig:priorvisualization2). The figure shows that the adjusted priors assign their highest probability density to a smaller range of values and are, therefore, more specific than `brms`'s default priors presented in Figure \@ref(fig:priorvisualizationbrms2). In addition, the shape of the distribution is different for the individual variation, resulting in a higher probability for values closer to zero. Panel B of Figure \@ref(fig:priorpredfigurepanel) shows the predictions of the normal model with adjusted prior distributions. Most effects now range between -0.5 and 0.5 seconds. This aligns better with our expectations based on existing literature, but still allows for a considerable uncertainty about the true size of the effects.

## Log-normal model

We follow the same procedure for prior specification in the log-normal model. As before, we prefer to apply conjugate priors. The conjugate priors that we applied to the normal model are also conjugate priors for the log-normal model. As a starting point, we can therefore use the adjusted priors of the normal model and perform a prior prediction, shown in panel C of Figure \@ref(fig:priorpredfigurepanel). These priors result in a slightly higher variance in RTs than we would expect, and, importantly, the mean response time differences are not centered around zero in all conditions. This pattern is inconsistent with our expectations, which means that prior distributions for the log-normal model need to be adjusted.

```{r priorpredfigurepanellog, echo=FALSE, fig.cap="Predicted mean RT difference between digits per repetition. The blue line represents the aggregated trend. Panel A represents the prior prediction based on the priors of the normal model. Panel B shows the prior prediction based on the selected prior for the log-normal model.", fig.align = "left", fig.height = 3, fig.width = 4, out.width="100%", eval = FALSE}
knitr::include_graphics("Images/priorpredfigure_lognormal_v2.png")
```
The transformation of the prior distributions is not as simple as the transformation of the parameter itself. Panel C shows that with the same priors as for the normal model, we obtain different data for the log-normal model compared to the normal model (panel B). Therefore, we adjusted the priors until the simulated data matched our expectations and looked like panel B. This yields the following priors that result in the prior prediction shown in panel D of Figure \@ref(fig:priorpredfigurepanel):^[For the overall baseline $\mu_\gamma$ we adjusted the prior based on our expectation of the expected overall response times. A mean parameter of -0.5 for the log-normal likelihood corresponds to a mean response time of 1.]

\begin{equation}
\begin{aligned}
\mu_{\gamma} &\sim \text{Normal}(-0.5, 1), (\#eq:chosenpriorslogvar) \\
\mu_{\beta} &\sim \text{Normal}(0, 0.005), \\
\mu_{\delta} &\sim \text{Normal}(0, 0.005), \\
\sigma^2 &\sim \text{Inverse-Gamma}(3, 0.3),  \\
\sigma^2_{\gamma} &\sim \text{Inverse-Gamma}(3, 0.3), \\   
\sigma^2_{\beta} &\sim \text{Inverse-Gamma}(3, 0.01), \\   
\sigma^2_{\delta} &\sim \text{Inverse-Gamma}(3, 0.01).    
\end{aligned}
\end{equation}
Setting the same prior distribution on the standard deviations instead of on the variances yields the following priors:

\begin{equation}
\begin{aligned}
\mu_{\gamma} &\sim \text{Normal}(-0.5, 1), (\#eq:chosenpriorslogsd) \\
\mu_{\beta} &\sim \text{Normal}(0, 0.07), \\
\mu_{\delta} &\sim \text{Normal}(0, 0.07), \\
\sigma &\sim \text{Inverse-Gamma}(13.8, 4.1),  \\
\sigma_{\gamma} &\sim \text{Inverse-Gamma}(13.8, 4.1), \\   
\sigma_{\beta} &\sim \text{Inverse-Gamma}(13.8, 0.7), \\   
\sigma_{\delta} &\sim \text{Inverse-Gamma}(13.8, 0.7).    
\end{aligned}
\end{equation}

## Prior sensitivity 
In the previous sections, we decided on one set of prior distributions for each of the proposed models. However, researchers may find it difficult to settle on a single set of prior distributions in practice. This raises the question: What happens if you cannot decide on one prior? If several possibilities seem plausible, it is possible to use them all separately for the analysis and investigate their influence on the analysis results. This is called a prior sensitivity analysis [@roos2015sensitivity]. Generally, when it comes to prior specification, it is important to be *transparent* and to provide a justification for the selected prior distributions [@Stefan_etal2021]. Prior distributions are rarely inadmissible. However, as we demonstrated earlier, their usefulness in constraining model predictions can vary. Therefore, providing a justification of prior distributions in a paper can help readers gauge the generative quality of a presented model. When multiple justifiable prior distributions are applied to the same analysis, it is important to be transparent about this to avoid the impression of cherry-picking priors. If priors are chosen for their best fit to the data, this can lead to overfitting and spurious results. Prior sensitivity analyses and transparency about the prior specification process more generally, can therefore help to increase trust in the research results.

# Bayesian Parameter Estimation

In Bayesian statistics, parameter estimates are obtained from the posterior distribution. The posterior distribution captures the uncertainty about the parameter after seeing the data, $\bf{Y}$. The mean of this distribution can be used as a point estimate for the parameter. The posterior distribution is calculated by updating the prior distribution using Bayes rule. For the symbolic distance effect, we have a model with many parameters, so the posterior distribution is also multidimensional. Instead of writing out all the parameters, we replace all the model parameters with the vector $\bm{\theta}$, resulting in the joint posterior:

\begin{equation}
\underbrace{P(\bm{\theta}|\bf{Y})}_{\text{Posterior}} = \underbrace{P(\bm{\theta})}_{\text{Prior}}\times\underbrace{\frac{P(\bf{Y}|\bm{\theta})}{P(\bf{Y})}}_{\text{Updating factor}}.
(\#eq:Bayesrule)
\end{equation}

Typically, we are not interested in the distribution of all parameters together, but in the posterior of one parameter independent of all the others. This posterior distribution is called the marginal posterior, and it is computed by integrating out all other parameters of the joint posterior. The marginal posterior distribution for the side effect $\mu_{\beta}$, for instance, is:

\begin{equation}
P(\mu_{\beta}| \bf{Y}) = \int P(\bf{Y} | \mu_{\beta}, \bm{\theta})P(\mu_{\beta})P(\bm{\theta})) d\bm{\theta}  
\end{equation}
where $\bm{\theta}$ now represents all parameters except for $\mu_{\beta}$. With an increasing number of parameters and an increasingly complex model structure, it is no longer possible to find analytical solutions for the multidimensional integrals involved in the posterior distributions. A solution to this issue is to sample from the joint posterior distribution. Using algorithms such as Markov chain Monte Carlo sampling [MCMC; for an introduction see  @van2018simple], Gibbs sampling [@chib1995marginal] or Hamiltonian Monte Carlo sampling [@betancourt2015hamiltonian]. It is out of the scope of this paper to go into detail on the algorithms, but we refer the interested reader to a tutorial paper on MCMC by van Ravenzwaaij et al. [-@van2018simple], a conceptual introduction to Hamiltonian Monte Carlo by Betancourt [-@betancourt2018conceptual], and a summary of samplers by Green et al. [-@green2015bayesian].

An approximation for the marginal posterior distribution can be obtained by assessing the empirical distribution of MCMC samples for the parameter of interest disregarding the other parameters. Since the frequency of combinations of samples in the joint posterior is determined by their posterior plausibility, this procedure is equivalent to a multidimensional integration across all other parameters.

Every sampling algorithm requires the specification of the number of samples (also called iterations) that are drawn from the posterior distribution. Furthermore, all MCMC sampling methods have in common that results depend (somewhat) on chosen starting values, that is, initial parameter values for which the unnormalized posterior (i.e., likelihood times prior) is evaluated. This dependency can be assessed if the sampling algorithm is repeated several times (i.e., chains) with different starting values. Since the first few iterations of a chain---often called warm-up or burn-in---are somewhat dependent on the starting value, these samples are typically discarded.

Before interpreting the estimation results, it has to be checked whether the posterior distribution of the parameters has converged, meaning that a stationary posterior distribution has been reached [@Vehtari_2021]. We will explain these convergence checks in more detail in the section Model Diagnostics.

## Software 
There are many different programs available to fit Bayesian hierarchical models, that is, to obtain samples from the joint posterior distribution of parameters. In this tutorial, we focus on two of the most commonly used `R` packages, `rstan` [@R-rstan] and `brms` [@R-brms_a; @R-brms_b]. We assume that readers are somewhat familiar with `R`. The package `rstan` provides the most modern algorithms and flexibility in model set-up. The package `brms` was created as an overlay to `rstan` and is based on the well-known `lme4` syntax [@R-lme4]. The goal of `brms` is to ease the transition to Bayesian hierarchical modeling for novice users of Bayesian statistics. The packages work similarly for the normal and log-normal model. In this section, we explain how a model can be fitted in each package using code snippets. In addition, we offer a separate Rmarkdown file for each package in our Online Supplement that explains in more depth how to analyze the full symbolic distance effect model with the specific package.

### Rstan

`Rstan` allows the application of `Stan` [@carpenter2017stan], a probabilistic programming language, in `R`. The package uses a version of the Hamiltonian Monte Carlo (HMC) algorithm, No-U-turn sampler [NUTS; Hoffman & Gelman, -@hoffman2014no], for sampling. `Rstan` requires two files: a `.stan` file containing the model specification and an R script containing the code for the model fitting. In the `.stan` file, the parameters, priors, and the likelihood are specified. An example `.stan` file of a simple, non-hierarchical normal model, where the variance is already known and the mean has to be estimated, is shown below [adapted example from @nicenboim2021introduction, Chapter 10]. The `.stan` file for the symbolic distance effect, including explanations, can be found in the Online Supplement. The file is divided into three sections: `data`, `parameters`, and `model`.

\scriptsize
```{r stanfileexample, eval = FALSE, echo = TRUE}
data {
  int<lower = 1> N;             // Total number of trials 
  vector[N] y;                  // Score in each trial 
  real<lower = 0> sigma;        // The variance of the normal distribution 
}

parameters {
  real mu; 
}

model {
  // Priors: 
  target += normal_lpdf(mu | 0, 20);
  // Likelihood: 
  for(i in 1:N)
    target += normal_lpdf(y[i] | mu, sigma);
}
```
\normalsize

The `data` section specifies the variables in the dataset, as well as constant values in the model. For the symbolic distance effect model, the `data` section contains the total number of observations, the participant index variable, the response time variable in the dataset, the condition for every response time (side of the digit and the digit indicator, i.e., variables $x_j$, $u_j$, $v_j$, $w_j$ and $z_j$ from \@ref(eq:datalevel)) and the prior specification according to Equations \@ref(eq:chosenpriors2var) and \@ref(eq:chosenpriorslogvar). The `parameters` section specifies all parameters that are estimated. These are the general and individual effects. The last section, `model`, contains the likelihood function and priors, as specified earlier in this article.

In the corresponding `R` file, the model is fitted to the data using the `stan` function, as shown below. In this function, it is possible to specify the number of iterations, chains and warm-up samples in the Monte Carlo sampling procedure. Next to these specifications, the function requires a directory path to the `.stan` file and an `R` object of type list containing the data (here called `myData_list`). Every element of the `data` section in the `.stan` file needs to be specified with the same name in the data object. 
&nbsp;

\scriptsize
```{r stanfitexample, eval = FALSE, echo = TRUE}
model_fit <- stan(file = "./myModel.stan",  # Stan file with model  
                  data = myData_list,       # List with observed data and constants
                  iter = 4000,              # Number (Nr.) of iterations per chain
                  chains = 4,               # Nr. of chains 
                  warmup = 1000)            # Nr. iterations for warmup per chain
```
\normalsize

The package `rstan` offers a lot of flexibility. It is possible to specify the likelihood and priors in the exact way as is wished. However, this comes with the price of some technical knowledge. The commented code for fitting the full symbolic distance effect model using `rstan`, can be found in [Online Supplement C: *Bayesian Hierarchical Modeling in Rstan*](https://www.dropbox.com/s/06f6vubmnhnm326/Appendix-C--Fit-BHM-in-rstan.pdf?dl=0).

### Brms

The `R` package `brms` also allows fitting Bayesian models using `Stan`. However, model specification and model fitting can be achieved using a much simplified `R`-script. Model specification is based on a formula syntax that is similar to the notation in the popular `lme4` package for frequentist hierarchical modeling [@R-lme4]. The formula is an object that specifies the dependent variable as a function of the independent variables. The `~` (tilde) separates the dependent variable (`rt`) on the left side from the independent variables on the right side. The formula distinguishes between general effects and individual deviations. The first part of the equation on the predictor side of the formula shown below (i.e., `1 + side + dif1 + dif2 + dif3 + dif4`) represents the general effects. The `1` represents the intercept, in the case of the symbolic distance effect $\mu_{\gamma}$, `side`, and `dif1` to `dif4` represent the general effects of side and distance ($\mu_{\beta}$ and $\mu_{\delta}$). Within the round brackets individual deviations and grouping variables are specified. In this case the grouping variable (`ind`) indicates that each participant has an individual effect. In other cases effects might vary per item, block, or even the combination of participant and item. The double bar (`||`) indicates that the correlation between the parameters should not be modeled. When using a single ba (`|`) then correlations across all individual effects (intercept, side and distance) are modeled. This approach, however, is a bit more difficult to interpret and would also not correspond to the implemented `Stan` model.

\scriptsize
```{r brmsfitexample, eval = FALSE, echo = TRUE}
formula <- rt ~ 1 + side + dif1 + dif2 + dif3 + dif4 +   # General effects 
          (1 + side + dif1 + dif2 + dif3 + dif4 || ind)  # Individual effects

model_fit <- brm(formula = formula,                      # Model formula 
                 data = myData_dataframe,                # Dataset (parameters, observations) 
                 family = gaussian(),                    # Family, distribution dependent variable 
                 prior = priorsmodel,                    # Specification of the priors 
                 warmup = 1000,                          # Nr. iterations for warmup (per chain)
                 iter = 4000,                            # Nr. iterations (per chain)
                 chains = 4,                             # Nr. chains 
                 core = 4)                               # Nr. of cores for parallel estimation
```

```{r formulafit, eval = FALSE, echo = FALSE}
formula <- rt ~ 1 + side + dif1 + dif2 + dif3 + dif4 +    # General effects 
          (1 + side + dif1 + dif2 + dif3 + dif4 || ind)   # Individual effects

```

\normalsize The `brm` function requires a data frame (here called `myData_dataframe`) that contains all variables mentioned in the formula. The argument `family` in the `brm` function specifies the exponential family of distributions according to which the dependent variable is assumed to be distributed. For the normal model, this argument takes the value `gaussian`; for the log-normal model, it needs to be specified as `lognormal`.

Finally, we specify the priors. In `brms`, priors can be set on the general effects represented by `class = b` and on the individual variation represented by `class = sd`. It is important to note that priors can only be set on the specific parameters that are part of the brms model specification. For example, `brms` only allows setting priors on standard deviations and not on variances.
Therefore, for the normal model of the symbolic distance effect, we will use the priors for standard deviations specified in Equation \@ref(eq:chosenpriors2sd). To set the priors, we use the `set_prior` function, as shown below. In this function, first, the distribution is specified, using the `Stan` programming language. Then, the type of parameter to which the prior applies is defined by `class`. The parameter to which this prior applies is specified by `coef`. For the priors on the individual effects, an additional element has to be specified, namely, the grouping variable (i.e., `group`). For example, the prior distributions on the general side effect, $\mu_{\beta}$, and on its variability between individuals, $\sigma_{\beta}$, can be specified as follows: 
&nbsp;

\scriptsize 
```{r priorsillusbrms2, eval = FALSE, echo = TRUE}
set_prior("normal(0,0.3)", class = "b", coef="side")  # general effect 
set_prior("inv_gamma(13.8, 5.3)", class = "sd",       # individual variation 
          coef="side", group = "ind") 
```
\normalsize

\scriptsize 
```{r priorsillusbrms, eval = FALSE, echo = FALSE}
priorsmodel <- c(set_prior("normal(0,1)", class = "Intercept"),  
                 set_prior("inv_gamma(13.75136, 4.102063)", class = "sd", coef="Intercept", group = "sub1"), 
                 set_prior("normal(0,0.3)", class = "b", coef="xj"),  
                 set_prior("inv_gamma(13.75134,5.295736)", class = "sd", coef="xj", group = "sub1"), 
                 set_prior("normal(0,0.3)", class = "b", coef="dif1"), 
                 set_prior("inv_gamma(13.75134,5.295736)", class = "sd", coef="dif1", group = "sub1"),
                 set_prior("normal(0,0.3)", class = "b", coef="dif2"),  
                 set_prior("inv_gamma(13.75134,5.295736)", class = "sd", coef="dif2", group = "sub1"),
                 set_prior("normal(0,0.3)", class = "b", coef="dif3"), 
                 set_prior("inv_gamma(13.75134,5.295736)", class = "sd", coef="dif3", group = "sub1"),
                 set_prior("normal(0,0.3)", class = "b", coef="dif4"),  
                 set_prior("inv_gamma(13.75134,5.295736)", class = "sd", coef="dif4", group = "sub1"),
                 set_prior("inv_gamma(13.75136, 4.102063)", class = "sigma"))

```

\normalsize If you do not specify a prior for a certain parameter, the default prior as explained in the prior section will be applied. It is possible to save all priors in an R `data.frame` and provide these in the `brm` function, this and the full code with explanation for fitting the full symbolic distance effect model using `brms` is shown in our [Online Supplement D: *Bayesian Hierarchical Modeling in brms*](https://www.dropbox.com/s/mthphl3ce2kk0xx/Appendix-D--Fit-BHM-in-brms.pdf?dl=0).^[In our example of the Number Classification Task we are applying a truncated normal prior distribution for only one parameter. This requires a slightly different `brms` formula and prior specification, explained in the Online Supplement.]

# Estimation of the Number Classification Task
## Data 
For our example, we will use the data set by Rouder et al. (2005). The data are available under a free license on GitHub [https://github.com/PerceptionCognitionLab/data0/tree/master/lexDec-dist5](https://github.com/PerceptionCognitionLab/data0/tree/master/lexDec-dist5). The data set contains observations from 54 participants who took part in a number classification task. Each participant performed 360 trials in six blocks. For the analysis, only the correct responses are included. To account for practice effects, the first 25 trials of the first block are removed from the analysis. The trials after the breaks between the blocks are also removed from the analysis to correct for lack of concentration. In addition, trials with very fast RTs (< 250 MS) and very slow RTs (> 2000 MS) are excluded. Two participants were excluded because they did not respond properly to the task. Similar criteria were applied by @haaf2019capturing. This results in 52 participants and a total of 17031 observations. In the following two sections we will describe the model fitting and parameter estimation procedure for the normal and log-normal model, respectively.

## Normal Model

```{r traceplots3p, echo=FALSE, fig.align = "left", fig.height = 11, fig.width = 4, out.width="100%", fig.cap="Panel A and B show the trace plot of the digit parameter $\\mu_{\\delta_{6}}$ for both packages. Panel C shows the estimated $\\hat{R}$ for all individual effects per parameter group. The stars represent the $\\hat{R}$ of the general effects. Panel D illustrates the variation in the estimated number of effective samples per parameter group. The dashed lines represent the total number of iterations for each package. The stars represent the number of effective samples of the general effects."}
knitr::include_graphics("Images/Fig8_traceplots3pnew_v2.png")
```

### Model Fitting 
The model is fitted using the packages `rstan` and `brms`. For `rstan`, four chains are run with 4000 iterations per chain, of which 1000 serve as warm-up. This results in a total of 12000 samples from the joint posterior distribution. For `brms`, four chains are run with 6000 iterations per chain, of which 1000 serve as warm-up. This results in a total of 20000 draws from the joint posterior distribution.

### Model Diagnostics   
Before we can interpret the results, we have to evaluate whether they are reliable. Specifically, it has to be checked whether the sampling chains of the parameters have converged, meaning that a stationary posterior distribution has been reached [@Vehtari_2021]. If convergence is not achieved, the posterior estimates may change substantially if the sampling algorithm was run with more iterations or different starting values. The most common convergence measures are trace plots and the $\hat{R}$ statistic. We will also discuss an additional check provided by `rstan` and `brms`, namely, the number of effective samples. For a critical evaluation of currently used convergence statistics, we recommend the interested reader to consult Vehtari et al. (2019).

#### Trace plot

Trace plots show the sampled parameter value at each iteration. If the distributions have converged, the plot should look like a hairy caterpillar [Lee & Wagenmakers, -@lee2013bayesian, p. 99]: the iterations move up and down, but have much overlap in the middle. If multiple chains are used, iterations are shown per chain. In this case, the chains should overlap. If the posterior distribution did not converge, one chain could, for instance, be found on top of the figure, while all other chains are on the bottom [for an illustration see @vehtari2019rank]. Or, perhaps the first half of the chain is on the top of the graph and the second half on the bottom. The trace plot does not include burn-in samples, since these will not be used for parameter estimation.

When inspecting the trace plots of the symbolic distance effect parameters in the normal model, the posterior distributions of the parameters appear to have converged as the iterations come together and look similar to a caterpillar. The trace plot of one of the parameters, $\delta_{6, i}$, is shown in panel A for `rstan` and panel B for `brms` of Figure \@ref(fig:traceplots3p). The trace plot in panel B shows less overlap in the middle, indicating that `brms` had a bit more dependency within chains. This might be a sign of slight issues. However, not that much to be too concerned. Trace plots for other parameters can be found in the Online Supplement.

#### $\hat{\bm{R}}$
Next to a visual inspection of convergence, a numerical inspection is usually performed as well. A common approach is to check $\hat{R}$, also called the Gelman-Rubin Diagnostic [@gelman1992inference]. $\hat{R}$ is the ratio of between-chain variance and within-chain variance [@sorensen2015bayesian]. If the chains diverge, the between-chain variance will be higher than the within-chain variance, resulting in a $\hat{R}$ greater than 1. This indicates that the chains have not converged. There are different types of $\hat{R}$, such as the split- and rank-based $\hat{R}$ [@gelman2013bayesian; @vehtari2019rank]. Even though researchers tend to trust convergence statistics more than visual inspection, these statistics are also not fool-proof. Therefore, the development of better convergence statistics is a current topic in Bayesian analysis methods.

The $\hat{R}$ values for all individual and general symbolic distance effect parameters computed by `rstan` and `brms` are shown in panel C of Figure \@ref(fig:traceplots3p). The figure shows that $\hat{R}$'s are not substantially greater than 1. Therefore, it seems that the distributions have converged.

#### Number of Effective Samples 
Another popular convergence statistic is the number of effective samples. This statistic is concerned with the dependence of posterior samples. In MCMC algorithms, the samples are to some degree dependent on one another: the parameter values at iteration *i* are similar to the parameter values at iteration *i-1*. The number of effective samples represents the estimated total number of *independent* draws from the posterior for every model parameter [@stan2018b, section 15.4]. If the number of effective samples is low, this indicates a high dependency of iterations in the sampling procedure. Therefore, the number of effective samples can be seen as a measure of the amount of new information about the posterior distribution that is provided by the total number of samples drawn from the posterior. A common rule for interpretation is that the number should be close to or even greater than the number of total iterations [@vehtari2019rank; @neff2018].

For the symbolic distance effect, the estimated number of effective samples is provided in panel D of Figure \@ref(fig:traceplots3p). For the package `rstan` the number is generally high, much higher than the actual number of samples depicted by the green dashed line. The `brms` package shows much more variation. The lowest effective sample sizes can be found for the intercept parameters. Table 1 also displays the number of effective samples for the general effects parameters in the `rstan` model. If the number of effective samples is too low, `brms` and `rstan` will return a warning. Here, this was not case, and all other diagnostics were good, so we continue with the interpretation of the estimates.

### Estimation Results
Considering good/acceptable convergence diagnostics, we can now interpret the estimation results. First, we present the general effects, then we visualize the individual variability of these effects.

```{r modelposteriorsmufigure3, echo=FALSE, fig.cap="The posterior distributions for the general effects estimated by the two packages. The middle line within the distributions represents the posterior mean. The shaded area within the distributions represent 95\\% credible intervals.", fig.align = "left", fig.height = 11, fig.width = 1, out.width="100%"}
knitr::include_graphics("Images/Fig9_posteriormu_v2.png")
```

#### General Effects 
For the symbolic distance effect, there are two main questions we want to answer. First, we want to know whether there is a digit effect and, second, we want to know whether there is a side effect. Therefore, the key model parameters for general effects are $\mu_{\beta}$, $\mu_{\delta_{3}}$, $\mu_{\delta_4}$, $\mu_{\delta_6}$, and $\mu_{\delta_{7}}$. In the following section, we will focus our reporting on these parameters.

The posterior distributions for the focal parameters in the normal model are presented in Figure \@ref(fig:modelposteriorsmufigure3). The posterior means estimated by `rstan` and `brms` can be found in the Online Supplement. The credible interval displayed in the figure is the Bayesian version of a confidence interval and provides a measure of uncertainty about the parameter value [@wagenmakers2018bayesian].

```{r geneftable, results = "hide", cache = TRUE, warning = FALSE}
# example - version 1 with SE 
rstanfithier8 <- readRDS(file = "R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
fitrstanhiertable <- summary(rstanfithier8, pars = c("mu", "mu2", "mu3", "mu4", "mu5", "mu6"))$summary[,c("mean", "se_mean", "2.5%", "97.5%", "n_eff", "Rhat")]
# as.data.frame(fitrstanhiertable)
fitrstanhiertable2 <- as.data.frame(fitrstanhiertable)
rownamesfitrstanhiertable <-  c("$\\mu_{\\gamma}$","$\\mu_{\\beta}$", "$\\mu_{\\delta_{7}}$", "$\\mu_{\\delta_{6}}$", "$\\mu_{\\delta_{4}}$", "$\\mu_{\\delta_{3}}$")
colnames(fitrstanhiertable2) <- c("Mean", "SE", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")

fitrstanhiertable3 <- add_column(fitrstanhiertable2, Parameters = rownamesfitrstanhiertable, .before = "Mean")
#apa_table(fitrstanhiertable3, row.names = FALSE, caption = "Posterior Mean, Standar Error (SE) of the Mean, Lower and Upper Bound of the 95\\% Credible Interval, the Number of Effective Samples, and the $\\hat{R}$ of the General Effect Parameters as estimated by Rstan.", align = c("l", "c", "c", "c", "c", "c", "c"), digits = 3, escape = FALSE)

# example - version 2 without SE 
rstanfithier8 <- readRDS(file = "R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
fitrstanhiertable <- summary(rstanfithier8, pars = c("mu", "mu2", "mu3", "mu4", "mu5", "mu6"))$summary[,c("mean", "2.5%", "97.5%", "n_eff", "Rhat")]
# as.data.frame(fitrstanhiertable)
fitrstanhiertable2 <- as.data.frame(fitrstanhiertable)
rownamesfitrstanhiertable <-  c("$\\mu_{\\gamma}$","$\\mu_{\\beta}$", "$\\mu_{\\delta_{7}}$", "$\\mu_{\\delta_{6}}$", "$\\mu_{\\delta_{4}}$", "$\\mu_{\\delta_{3}}$")
colnames(fitrstanhiertable2) <- c("Mean", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")

fitrstanhiertable3 <- add_column(fitrstanhiertable2, Parameters = rownamesfitrstanhiertable, .before = "Mean")
apa_table(fitrstanhiertable3, row.names = FALSE, caption = "Posterior Mean, Lower and Upper Bound of the 95\\% Credible Interval, Number of Effective Samples, and $\\hat{R}$ of the General Effect Parameters as estimated by rstan.", align = c("l", "c", "c", "c", "c", "c"), digits = 3, placement = "h", escape = FALSE)
```
The parameter estimates produced by the two packages look very similar. The `brms` estimates seem to be slightly higher with slightly more narrow posterior distributions compared to `rstan`. All parameters are presented on a scale of seconds. As expected, the effects are small. For instance, $\mu_{\delta_{7}}$ has a posterior mean of `r printnum(fitrstanhiertable3["mu3", "Mean"], digits = 3)`, corresponding to a 11 millisecond effect.

For the digit effects, positive parameter estimates indicate that the response is slower for the non-baseline digits, that is, digits that are closer to 5. The parameters $\mu_{\delta_{6}}$ and $\mu_{\delta_{4}}$ are distributed around higher values and their credible intervals do not contain zero. Note that this result is consistent with the symbolic distance hypothesis that the digits 4 and 6 have the highest response times because they are closest to the 5, the comparison value. The 95% credible intervals of $\mu_{\delta_{7}}$ and $\mu_{\delta_{3}}$ contain zero. Therefore, it is plausible that the response times do not differ substantially from the baseline condition. The posterior distribution of the side effect $\mu_{\beta}$ is centered around zero with a posterior mean of `r printnum(fitrstanhiertable3["mu2", "Mean"], digits = 3)` and a 95% credible interval ranging from `r printnum(fitrstanhiertable3["mu2", "Lower Bound"], digits = 3)` to `r printnum(fitrstanhiertable3["mu2", "Upper Bound"], digits = 3)`. This indicates a small to non-existent side effect.

#### Individual effects

```{r inddevtable, results = "asis", cache = TRUE}
# example - With SE 
# rstanfithier8 <- readRDS(file = "R objects/rstanmodelfit_final_kopie.rds")
fitrstanhiertable2 <- summary(rstanfithier8, pars = c("g", "g2", "g3", "g4", "g5", "g6"))$summary[,c("mean", "se_mean", "2.5%", "97.5%", "n_eff", "Rhat")]
# as.data.frame(fitrstanhiertable)
fitrstanhiertable22 <- as.data.frame(fitrstanhiertable2)
rownamesfitrstanhiertable2 <-  c("$\\sigma_{\\gamma}^2$","$\\sigma_{\\beta}^2$", "$\\sigma_{\\delta_{7}}^2$", "$\\sigma_{\\delta_{6}}^2$", "$\\sigma_{\\delta_{4}}^2$", "$\\sigma_{\\delta_{3}}^2$")
colnames(fitrstanhiertable22) <- c("Mean", "SE", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")

fitrstanhiertable32 <- add_column(fitrstanhiertable22, Parameters = rownamesfitrstanhiertable2, .before = "Mean")
# apa_table(fitrstanhiertable32, row.names = FALSE, caption = "Posterior Variance, Standar Error (SE) of the Variance, Lower and Upper Bound of the 95\\% Credible Interval, the Number of Effective Samples, and the $\\hat{R}$ of the General Effect Parameters as estimated by Rstan.", align = c("l", "c", "c", "c", "c", "c", "c"), digits = 3, escape = FALSE)

# example - without SE 
# rstanfithier8 <- readRDS(file = "R objects/rstanmodelfit_final_kopie.rds")
fitrstanhiertable2 <- summary(rstanfithier8, pars = c("g", "g2", "g3", "g4", "g5", "g6"))$summary[,c("mean", "2.5%", "97.5%", "n_eff", "Rhat")]
# as.data.frame(fitrstanhiertable)
fitrstanhiertable22 <- as.data.frame(fitrstanhiertable2)
rownamesfitrstanhiertable2 <-  c("$\\sigma_{\\gamma}^2$","$\\sigma_{\\beta}^2$", "$\\sigma_{\\delta_{7}}^2$", "$\\sigma_{\\delta_{6}}^2$", "$\\sigma_{\\delta_{4}}^2$", "$\\sigma_{\\delta_{3}}^2$")
colnames(fitrstanhiertable22) <- c("Mean", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")

fitrstanhiertable32 <- add_column(fitrstanhiertable22, Parameters = rownamesfitrstanhiertable2, .before = "Mean")
apa_table(fitrstanhiertable32, row.names = FALSE, caption = "Posterior Variance, Lower and Upper Bound of the 95\\% Credible Interval, Number of Effective Samples, and $\\hat{R}$ of the Variance Parameters as estimated by Rstan.", align = c("l", "c", "c", "c", "c", "c"), digits = 3, placement = "h", escape = FALSE)
```
Thanks to the hierarchical structure of the model, we can investigate whether the general effects hold for all the individuals by inspecting the individual variation and the individual parameter estimates. The individual variation is represented by the variance or standard deviation parameters in the model. The summaries of the posterior distributions of the variance parameters as estimated by `rstan` are shown in Table \@ref(tab:inddevtable). The estimates for the standard deviations by `brms` are smaller compared to `rstan` and can be found in Online Supplement D.

Posterior means of the variances for the digit and side effects are all very close to `r printnum(fitrstanhiertable32["g3", "Mean"], digits = 3)` which corresponds to a standard deviation of `r printnum(sqrt(fitrstanhiertable32["g3", "Mean"]), digits = 3)` seconds, that is, 141 milliseconds. Additionally, the 95% credible intervals do not contain zero. If we consider the effect of digit 7 wit an 11 millisecond general effect this is a substantial amount of individual variation. Therefore, we conclude that there are individual differences in the symbolic distance effect.

```{r randomparfigure, echo=FALSE, fig.cap="The posterior means for individual effect parameters with 95\\% credible intervals as estimated by \\textit{rstan}, shown in increasing order. The dashed line represents the general posterior mean. Pink intervals contain zero, blue interval do not contain zero.", fig.align = "left", fig.height = 1, fig.width = 0.5, out.width="100%"}
knitr::include_graphics("Images/Fig10_randompar_v2.png")

```

Figure \@ref(fig:randomparfigure) shows the posterior means and 95% credible intervals for the individual effects of all 52 participants. Estimates are printed in increasing order, from lowest to highest. Pink intervals indicate that the 95% credible interval contains zero, while blue intervals are either entirely above or below zero. A corresponding figure with individual effect estimates for `brms` can be found in Online Supplement D.

The largest variation across participants can be seen for the intercept parameter. This means that participants vary considerably in their baseline response times. There is also considerable individual variation in the digit effects ($\delta$'s). However, most of the intervals contain zero which implies that the effects might be small or non-existent. The variability is most pronounced for the effect parameters of the digits close to 5. For these digits, there are also more intervals that do not contain zero, indicating that an effect might be present on the subject level. The posterior mean of the individual side effect ($\beta_{i}$) varies between -0.05 and 0.05. Again, all but four 95% credible intervals include zero indicating uncertainty about the existence of an effect for anyone. The credible intervals of the side effect and digit effects $\delta_{7}$ and $\delta_{3}$ indicate that different subjects may show opposite effects [@haaf2019some].

Finally, Figure \@ref(fig:modelestimatesfigure3) shows how the relationship between individuals' digit effects.^[The design and code for this figure are by @j_van_langen_2020_3715576.] The posterior means of individual effects are represented by the dots. The right side of each plot represents the distribution of the individual estimates. Note that this distribution is not the posterior distribution of the general effect. It rather corresponds to the left side of Figure 6. The two panels of Figure \@ref(fig:modelestimatesfigure3) show that the estimates are almost identical between `Stan` and `brms`. This means that differences in the parametrization between the packages did not influence the results of the parameter estimation.

```{r modelestimatesfigure3, echo=FALSE, fig.cap="Model estimates for digit effect parameters in \\textit{rstan}. The points represent the mean parameter estimates for each individual. The violin plots on the right side show the variance of the individual parameter estimates. The size of effects is not very constistent within a person.", fig.align = "left", fig.height = 11, fig.width = 1, out.width="75%"}
knitr::include_graphics("Images/Fig11_modelestimatefigure2_onlyrstan_v2.png")
```

## Log-normal Model 
### Model Estimation 
The model is fitted using the packages `rstan` and `brms`. For `rstan`, we use the priors as specified in Equation \@ref(eq:chosenpriorslogvar). For `brms`, we use the priors in Equation \@ref(eq:chosenpriorslogsd). The rest of the settings are equivalent for both packages. We ran four chains with 4000 iterations per chain, including 1000 warm-up iterations. This results in a total of 12000 samples from the posterior distribution.

### Model Diagnostics 
Before interpreting the results, we have to check if the posterior distributions of the parameters have converged. The trace plot of $\mu_{\delta_{6}}$ is shown panel A for `rstan` and B for `brms` of Figure \@ref(fig:traceplotlog). Trace plots for all other parameters can be found in the Online Supplement. All plots look like a hairy caterpillar. Therefore, they indicate that the posterior distributions of the parameters have converged. $\hat{R}$ values shown in Figure \@ref(fig:traceplotlog)C for all parameters are all close to 1 and most of the number of effective samples shown in Figure \@ref(fig:traceplotlog)D exceed the total number of iterations. The number of effective samples for the general effects are comparatively low but still sufficient. In combination with the other diagnostics, we conclude that the posterior distributions of the parameters have converged and we proceed with the interpretation of the results.

(ref:captiontraceplotlog) Panel A and B show the trace plot of the digit parameter $\\mu_{\\delta_{6}}$ in the log-normal model for both packages. Panel C shows the estimated $\\hat{R}$ for all individual effects per parameter group. The stars represent the $\\hat{R}$ of the general effects. Panel D illustrates the variation in the estimated number of effective samples per parameter group. The dashed lines represent the total number of iterations for each package. The stars represent the number of effective samples of the general effects.

```{r traceplotlog, echo=FALSE, fig.cap="(ref:captiontraceplotlog)", fig.align = "left", fig.height = 11, fig.width = 4, out.width="100%"}
knitr::include_graphics("Images/Fig12_traceplots3plog_v2.png")
```

### Results
#### General effects  
We inspect the posterior distributions for the general effect parameters presented in Figure \@ref(fig:postplotlog). 
```{r postplotlog, echo=FALSE, fig.cap="The posterior distributions for the general effects in the log-normal model estimated by the two packages. The middle line within the distributions represents the posterior mean. The shaded area within the distributions represent 95\\% credible intervals.", fig.align = "left", fig.height = 11, fig.width = 1, out.width="100%"}
knitr::include_graphics("Images/Fig13_posteriormulog_v2.png")
```

```{r geneftablelog, results = "hide", cache = TRUE, warning = F}
# example - with SE 
rstanfithier8log <- readRDS(file = "R objects/rstan/log-normal model/hier_logmodelc_adj13012022.rds")
fitrstanhiertablelog <- summary(rstanfithier8log, pars = c("mu", "mu2", "mu3", "mu4", "mu5", "mu6"))$summary[,c("mean", "se_mean", "2.5%", "97.5%", "n_eff", "Rhat")]
# as.data.frame(fitrstanhiertable)
fitrstanhiertable2log <- as.data.frame(fitrstanhiertablelog)
rownamesfitrstanhiertablelog <-  c("$\\mu_{\\gamma}$","$\\mu_{\\beta}$", "$\\mu_{\\delta_{7}}$", "$\\mu_{\\delta_{6}}$", "$\\mu_{\\delta_{4}}$", "$\\mu_{\\delta_{3}}$")
colnames(fitrstanhiertable2log) <- c("Mean", "SE", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")

fitrstanhiertable3log <- add_column(fitrstanhiertable2log, Parameters = rownamesfitrstanhiertablelog, .before = "Mean")
# apa_table(fitrstanhiertable3log, row.names = FALSE, caption = "Posterior Mean, Standar Error (SE) of the Mean, Lower and Upper Bound of the 95\\% Credible Interval, the Number of Effective Samples, and the $\\hat{R}$ of the General Effect Parameters as estimated by Rstan.", align = c("l", "c", "c", "c", "c", "c", "c"), digits = 3, escape = FALSE)

# example - without SE 
rstanfithier8log <- readRDS(file = "R objects/rstan/log-normal model/hier_logmodelc_adj13012022.rds")
fitrstanhiertablelog <- summary(rstanfithier8log, pars = c("mu", "mu2", "mu3", "mu4", "mu5", "mu6"))$summary[,c("mean", "2.5%", "97.5%", "n_eff", "Rhat")]
# as.data.frame(fitrstanhiertable)
fitrstanhiertable2log <- as.data.frame(fitrstanhiertablelog)
rownamesfitrstanhiertablelog <-  c("$\\mu_{\\gamma}$","$\\mu_{\\beta}$", "$\\mu_{\\delta_{7}}$", "$\\mu_{\\delta_{6}}$", "$\\mu_{\\delta_{4}}$", "$\\mu_{\\delta_{3}}$")
colnames(fitrstanhiertable2log) <- c("Mean", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")

fitrstanhiertable3log <- add_column(fitrstanhiertable2log, Parameters = rownamesfitrstanhiertablelog, .before = "Mean")
apa_table(fitrstanhiertable3log, row.names = FALSE, caption = "Posterior Mean, Lower and Upper Bound of the 95\\% Credible Interval, Number of Effective Samples, and $\\hat{R}$ of the General Effect Parameters as estimated by Rstan.", align = c("l", "c", "c", "c", "c", "c"), digits = 3, placement = "h", escape = FALSE)
```

```{r modelfitlog, eval = FALSE}
# Based on model estimates 
> exp(-0.56 - (1/2)*0.01 + 0.08)
[1] 0.6156972
> exp(-0.56 - (1/2)*0.01 + 0.03)
[1] 0.5856693
> 0.6156972 - 0.5856693
[1] 0.0300279
> exp(-.56 + .5  (-.01) + .03) - exp(-.56 + .5  (-.01))
Error: attempt to apply non-function
> exp(-.56 + .5 * (-.01) + .03) - exp(-.56 + .5  * (-.01))  # difference 3 and 2 
[1] 0.01730914
> exp(-.56 + .5 * (-.01) + .08) - exp(-.56 + .5  * (-.01))  # difference 4 and 2? 
[1] 0.04733705
```
We note that the values of the parameter estimates for the log-normal model cannot be interpreted in the same way as in normal model. The reason is that the mean parameter of the log-normal distribution does not correspond to its expected value. To better understand the size of an effect that corresponds to the parameter values, some calculations are required. For instance, if we would like to know the median difference in response times for digit 3 and 2, we calculate the difference between the exponential of the RT estimate for the condition digit = 3 and the exponential of the RT estimate for the condition digit = 2:

\begin{equation}
\begin{aligned}
&e^{(\mu_{\gamma} + x_j \mu_{\beta} + w_j \mu_{\delta_{3}})} - e^{(\mu_{\gamma} +  x_j \mu_{\beta})} \\
&= e^{(-.56 + .5 \times (-.01) + 1 \times 0.03)} - e^{(-.56 + .5 \times (-.01))} \\ 
&= 0.017 
\end{aligned}
\end{equation}
This value corresponds to the median estimate of $\mu_{\delta_{3}}$ from the normal model. Calculating the mean estimate of $\mu_{\delta_{3}}$ would be similar, but more complicated as the mean of a log-normal distribution also depends on its variance.

One property of these parameter estimates we can immediately interpret is their sign (i.e., positive, negative or zero). A positive estimate indicates that the RT increases, whereas a negative effect indicates a decrease. When inspecting the general digit effect parameters, they all appear to positively influence the RT. This is in line with our expectations since the baseline RT in the model is set equal to the RT for the digits furthest away from 5 and the symbolic distance effect postulates that RT should increase as the digits become closer to five. The side effect is close to zero, with a posterior mean of `r printnum(fitrstanhiertable3log["mu2", "Mean"], digits = 3)` and a 95% credible interval from `r printnum(fitrstanhiertable3log["mu2", "Lower Bound"], digits = 3)` to `r printnum(fitrstanhiertable3log["mu2", "Upper Bound"], digits = 3)`. Thus, the general effect of side seems small or negligible.

#### Individual effects 
Next, we inspect the individual deviations from the general effects. Table \@ref(tab:inddevtablelog) presents the estimates of the variance parameters by `rstan`. Results for brms can be found in the Online Supplement. 
```{r inddevtablelog, results = "asis", cache = TRUE}
# example - with SE 
# rstanfithier8log <- readRDS(file = "R objects/hier_logmodelc_adj.rds")
fitrstanhiertable2log <- summary(rstanfithier8log, pars = c("g", "g2", "g3", "g4", "g5", "g6"))$summary[,c("mean", "se_mean", "2.5%", "97.5%", "n_eff", "Rhat")]
# as.data.frame(fitrstanhiertable)
fitrstanhiertable2log2 <- as.data.frame(fitrstanhiertable2log)
rownamesfitrstanhiertable2log2 <-  c("$\\sigma_{\\gamma}^2$","$\\sigma_{\\beta}^2$", "$\\sigma_{\\delta_{7}}^2$", "$\\sigma_{\\delta_{6}}^2$", "$\\sigma_{\\delta_{4}}^2$", "$\\sigma_{\\delta_{3}}^2$")
colnames(fitrstanhiertable2log2) <- c("Mean", "SE", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")

fitrstanhiertable32log <- add_column(fitrstanhiertable2log2, Parameters = rownamesfitrstanhiertable2log2, .before = "Mean")
# apa_table(fitrstanhiertable32log, row.names = FALSE, caption = "Posterior Variance, Lower and Upper Bound of the 95\\% Credible Interval, Number of Effective Samples, and $\\hat{R}$ of the General Effect Parameters as estimated by Rstan", align = c("l", "c", "c", "c", "c", "c", "c"), digits = 3, escape = FALSE)

# example - without SE 
# rstanfithier8log <- readRDS(file = "R objects/hier_logmodelc_adj.rds")
fitrstanhiertable2log <- summary(rstanfithier8log, pars = c("g", "g2", "g3", "g4", "g5", "g6"))$summary[,c("mean", "2.5%", "97.5%", "n_eff", "Rhat")]
# as.data.frame(fitrstanhiertable)
fitrstanhiertable2log2 <- as.data.frame(fitrstanhiertable2log)
rownamesfitrstanhiertable2log2 <-  c("$\\sigma_{\\gamma}^2$","$\\sigma_{\\beta}^2$", "$\\sigma_{\\delta_{7}}^2$", "$\\sigma_{\\delta_{6}}^2$", "$\\sigma_{\\delta_{4}}^2$", "$\\sigma_{\\delta_{3}}^2$")
colnames(fitrstanhiertable2log2) <- c("Mean", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")

fitrstanhiertable32log <- add_column(fitrstanhiertable2log2, Parameters = rownamesfitrstanhiertable2log2, .before = "Mean")
apa_table(fitrstanhiertable32log, row.names = FALSE, caption = "Posterior Variance, Lower and Upper Bound of the 95\\% Credible Interval, the Number of Effective Samples, and the $\\hat{R}$ of the Variance Parameters as estimated by Rstan.", align = c("l", "c", "c", "c", "c", "c"), digits = 3, placement = "h", escape = FALSE)
```
The estimated parameter values for the variance terms are very small, but different from zero. This indicates that individuals may vary slightly with respect to the general effect parameters. The estimates for the standard deviations by `brms` are similar compared to the estimates by `rstan`, but slightly smaller. The exact estimates by `brms` can be found in Online Supplement F.
Figure \@ref(fig:randomparplotlog) shows the individual estimates for the digit and side effects. The pink lines indicate that the 95% credible interval of the individual effect contains zero. The estimated general effect is represented by the dashed line. For $\beta$, $\delta_{7}$, and $\delta_{3}$, most 95% credible intervals contain zero. For $\delta_{6}$ and $\delta_{4}$, most credible intervals are entirely larger than zero, indicating an increase in RT for most participants.

```{r randomparplotlog, echo=FALSE, fig.cap="The posterior means for individual effect parameters in the log-normal model with 95\\% credible intervals as estimated by \\textit{rstan}, shown in increasing order. The dashed line represents the general posterior mean. Pink intervals contain zero, blue interval do not contain zero.", fig.align = "left", fig.height = 1, fig.width = 0.5, out.width="100%"}
knitr::include_graphics("Images/Fig14_randomparlog_v2.png")
```
We can further inspect the individual estimates for the digit effects in Figure \@ref(fig:modeldigitestplotlog). The figure shows the individual estimates for the digit effects with the distribution of the variability of the individual estimates (not the posterior distribution of the general effect). The figure illustrates that there is considerable variance in the point estimates of the individual digit effects. However, compared to the normal model there is considerably more hierarchical shrinkage.

# Model Comparison 
The symbolic distance effect postulates that response times (RTs) are slower when digits are closer to 5 (digit effects), and that RTs may be influenced by whether the target digit is smaller or greater than 5 (side effect). Using Bayesian estimation, we are not able to directly test these hypotheses. Credible intervals often contain zero, but they always also contain lots of other values that have a chance of being the "true" parameter value, making them unsuitable to test hypotheses. Therefore, our goal for the following section is to set up a Bayesian hypothesis test for the parameters of interest in our model. Specifically, we will achieve this by comparing the predictive accuracy of a null model where we set one or some parameters to zero, and and effects models where these parameters are unconstrained.

```{r modeldigitestplotlog, echo=FALSE, fig.cap="Model estimates for digit effect parameters in the log-normal model in \\textit{rstan}. The points represent the mean parameter estimates for each individual. The violin plots on the right side show the variance of the individual parameter estimates.", fig.align = "left", fig.height = 11, fig.width = 1, out.width="75%"}
knitr::include_graphics("Images/Fig15_inddiflogplot_onlyrstan_v2.png")
```

## Bayes factor
The Bayes factor (BF) is a measure of how well one model predicts the data compared to another [@jeffreys1961theory; @kass1995bayes; @rouder2018teaching]. The Bayes factor is defined as the probability of the data under one model compared to the probability of the data under another model. For instance, the Bayes factor comparing a null model ($\mathcal{M}_0$) and a full model ($\mathcal{M}_{\text{F}}$) can be obtained by:

\begin{equation}
\text{BF}_{0\text{F}} = \frac{P(\bm{Y} | \mathcal{M}_0)}{P(\bm{Y} | \mathcal{M}_{\text{F}})}. 
\end{equation}

The Bayes factor can also be interpreted as the relative evidence of two competing models. Bayes factors have a few very convenient characteristics. One of them is that Bayes factors are transitive, that is, evidence for the full model over the null model can be obtained by flipping numerator and denominatior in equation 13. For a full overview of the advantages of Bayesian inference with Bayes factor, see Wagenmakers et al. [-@wagenmakers2018bayesian].

Our models consist of several parameters. Therefore, computing the Bayes factor of, for example, the normal model versus the log-normal model becomes somewhat difficult. Let the vector of all parameters of the normal model be $\bm{\theta_\text{n}}$ and the vector of all parameters for the log-normal model be $\bm{\theta_{\text{l}}}$. With the prior distributions for the two models, $f(\bm{\theta_\text{n}})$ and $f(\bm{\theta_\text{l}})$, the Bayes factor can be calculated by the following ratio of marginal likelihoods:

\begin{equation}
\text{BF}_{\text{nl}} = \frac{\int_{-\infty}^{\infty}P(\bm{Y} | \mathcal{M}_\text{n}, \bm{\theta_{\text{n}}})f(\bm{\theta_{\text{n}}})d\bm{\theta_{\text{n}}}}{\int_{-\infty}^{\infty}P(\bm{Y} | \mathcal{M}_\text{l}, \bm{\theta_{\text{l}}})f(\bm{\theta_{\text{l}}})d\bm{\theta_{\text{l}}}}. 
\end{equation}

Imagine this results in $\text{BF}_{\text{nl}} = 100$. This means that the data are 100 times more likely under the normal model than under the log-normal model, clear evidence in favor of the normal model. But how do we know whether a Bayes factor of 100-to-1, or 10-to-1 or 3-to-1. As a reminder, Bayes factors are the relative predictive accuracy of two models for the observed data. Therefore, Bayes factors are odds or ratios. Odds themselves are directly interpretable without the need for decisions or cutoffs. For example, if a presidential candidate is favored 10-to-1 over another one, then these are just the odds, and it is beside the point whether these are large or not. According to this line of thought, Rouder et al. [-@rouder2018theories] argue that researchers should take a similar approach when interpreting and reporting Bayes factors. While there exist rules of thumb to categorize Bayes factors verbally, we suggest researchers use their research context to form substantive conclusions based on Bayes factors, not arbitrary rules of thumb.

Same as the posterior distribution, Bayes factors can rarely be computed analytically due to the multidimensional integrals involved. The difficulty to compute Bayes factors is one of the major drawbacks of Bayesian inference. However, several algorithms are available to estimate Bayes factor from posterior samples. In the following section we will present two methods to obtain Bayes factors for the symbolic distance effect.

### Savage-Dickey Density Ratio  
One method to compute Bayes factors is the Savage-Dickey density ratio [SDD-ratio; Dickey & Lientz, -@dickey1970weighted; Morey, Rouder, Pratte, & Speckman, -@morey2011using; Wagenmakers, Lodewyckx, Kuriyal, & Grasman, -@wagenmakers2010bayesian]. This approach can be readily used to estimate the Bayes factor between two nested models that differ in only one parameter. A prime example for such a case is comparing an effect model with a null model. For example, we may compare a model with an overall side effect (effect model) to a model with no overall side effect (null model). In this case, the Bayes factor in favor of the side effect can be approximated by the ratio of the posterior density of the side effect and the prior density of the side effect at the point zero. More generally, the Savage Dickey approach works whenever one model fixes a parameter to a constant and another model assigns a prior distribution to the same parameter.

Figure \@ref(fig:sdratioplotall) shows a visualization of the SDD-ratio for the side parameter in the normal model of the symbolic distance effect. The posterior distribution is depicted in orange, the prior in green. The Bayes factor can be computed as a ratio of the posterior and prior density at the parameter value of zero, as depicted by the dashed line connecting the two points on the density functions.

```{r sdratioplotall, echo=FALSE, fig.cap="The prior and posterior distribution of the side parameter estimated by the \\textit{brms} package. The dots represent the height of the distributions at zero. The ratio of these dots indicates the Savage-Dickey density ratio. The Bayes factor represents the evidence for the hypothesis that the effect equals zero. A Bayes factor above one indicates that the evidence is in favor of the null hypothesis.", fig.align = "left", fig.height = 6, fig.width = 4, out.width="60%"}
knitr::include_graphics("Images/Fig16_sdrplotside_v2.png")
```
With `rstan`, the SDD-ratio can be computed by fitting a density function to the posterior distribution obtained from the MCMC samples, and comparing the value of this density function at $\mu_{\beta} = 0$ to the value of the prior distribution at $\mu_{\beta} = 0$. This is illustrated in the code block below. First, we extract the posterior samples for the parameter $\mu_{\beta}$, the general effect of side. We apply the `logspline` function [@logspline] to the samples to obtain an estimate of the log-density of the posterior distribution. The density at the zero is calculated with the `dlogspline` function. Finally, we compare the estimated density of the posterior to the density of the prior.

\scriptsize
```{r codesnippetbfrstansdratio, eval=FALSE, echo=TRUE}
library(logspline)
# Get posterior samples for the side parameter mu2
samples_side <- rstan::extract(model_fit)$mu2   
log.posterior <- logspline(samples_side)                 # estimate of log-density post. distr. 
posterior_at_zero <- dlogspline(0, log.posterior)        # compute posterior density 
prior_at_zero <- dnorm(0, 0, 0.3)                        # computer prior density 
BF01 <- posterior_at_zero/prior_at_zero                  # compare estimated density posterior to prior at point 0 
```
\normalsize
In `brms`, the SDD-ratio can be computed using the built-in `hypothesis` function, as shown in the code snippet below. The resulting Bayes factor, also depicted in Figure \@ref(fig:sdratioplotall), is $BF_{01} = 18.28$, indicating evidence for the null model and against a side effect.
&nbsp;

\scriptsize
```{r codesnippetbfbrms, eval=FALSE, echo=TRUE}
# We will use the model fit from before, specified as model_fit
hypothesis_fit <- hypothesis(model_fit,    # Object containing model fit 
                             "mu2 = 0")   # The parameter we want to test with the null hypothesis 
```
\normalsize

The Savage-Dickey approach has a number of drawbacks [see, for example, Heck, -@heck2019caveat]. First, one issue is the quality of approximation for the SSD-ratio. If either the prior or, even more problematically, the posterior have a really low density at the test value (in this case zero), it becomes difficult to estimate the SSD-ratio from the samples. This issue often occurs when the effect is large, moving the posterior distribution away from zero. In many cases there might be no samples around zero, which means that the density estimate will be very inaccurate. Therefore, we would recommend only to use the SSD-ratio if you are confident that enough posterior samples are available from the tails of the distribution. One way of ensuring that is to drastically increase the number of drawn samples.

Second, the Savage-Dickey approach only applies for a limited set of to-be-compared models. In the example used here, we compared a model where the overall side effect is allowed to vary with a model where it is zero. Both models, however, still allow for individual variability of the side effect. For the null model, individual variability implies that some individuals are expected to have an effect favoring larger numbers and other individuals are expected to have an effect favoring smaller numbers, but, on average, both groups perfectly balance out at zero. To us, this model seems nonsensical, and, more importantly, it is not what researchers have in mind if they want to test a model without side effect. We would prefer to use a null model where none of the participants have a side effect. Yet, the Savage-Dickey density ratio is not available for such a model compared to the effects model.

### Bridge Sampling  
Bridge sampling offers a more flexible way to obtain Bayes factors [@kass1995bayes]. It allows the computation of Bayes factors for comparing models that differ in more than one parameter. Specifically, bridge sampling is a sampling-based algorithmic method to obtain the marginal likelihood of the data under a given model from the posterior samples [@meng1996simulating; @gronau2017tutorial].

Using bridge sampling we are able to estimate Bayes factors between any models of interest for the classification task data. We identify four models of interest:

1. Full model: This is the model we have been using throughout this paper. It states that there is an effect of side and an effect of digits for all participants. 
2. Side model: This model specifies that participants show a side effect but no digit effect. All digit parameters $\delta_{3}$ to $\delta_{7}$ are set to zero. 
3. Digit model: This model states that there are no side effects but that participants exhibit digits effects. All side parameters $\beta$ are set to zero. 
4. Null model: In this model, there is no side or digit effect on the RT. The digit parameters $\delta_{3}$ to $\delta_{7}$ and side parameters $\beta$ are set to zero.

Note that models models 2-4 are nested in the most complex full model, and the null model is nested in models 3 and 4, but models 3 and 4 are not nested. Additionally, the models differ by several parameters. For example, compared to the full model the Digit model restricts the side parameter for all participants to zero. Using bridge sampling, we can estimate the marginal likelihood for each model separately, and then compute Bayes factors as their ratios (Equation 14).

To obtain the marginal likelihood for a specific model in `rstan` with bridge sampling [@gronau2017tutorial], the `bridgesampling` package can be used [@R-bridgesampling]. To obtain the Bayes factor comparing the side model and the full model, we first have to fit both models. Then, we use the `bridge_sampler` function from the `bridgesampling` package to obtain the marginal likelihood of the data under each of the models. Finally, the models are compared using the `bf` function from the `bridgesampling` package. 
&nbsp;

\scriptsize
```{r codesnippetrstanestBF, eval = FALSE, echo = TRUE}

# We will use the model fit of both models (called sidemodel and fullmodel)
# (how to perform this is illustrated in the model estimation section)

# Bridge sampling 
library(bridgesampling)
ML_Hside <- bridge_sampler(sidemodel)  # compute the marginal likelihood under the side model 
ML_Hfull <- bridge_sampler(fullmodel)  # compute the marginal likelihood under the full model

# Obtain Bayes factor (in this way evidence in favor of the side hypothesis)
bridgesampling::bf(ML_Hside, ML_Hfull)

```
\normalsize
In the `brms` package, Bayes factors based on bridge sampling can be obtained with the function `bayes_factor` as shown in the code snippet below. The function takes two fitted models as input. Therefore, both models first have to be fitted as explained in Model Estimation section of this manuscript.

\scriptsize
```{r codesnippetbrmsestBF, eval = FALSE, echo = TRUE}
bayes_factor(fitsidemodel, fitfullmodel)$bf
```
\normalsize

Even though bridge sampling is a flexible method for obtaining Bayes factors, it is still an estimation method based on posterior samples. Therefore, as with any Bayes factor estimation method, we advice to take two precautions with bridge sampling. First, it is always wise to increase the number of iterations when performing model comparison. The reason is that convergence of estimation of posterior distributions and convergence of estimation of the marginal likelihood can differ. Therefore, a rule of thumb is to use ten times the number of iterations than we would usually use for estimation. The second precaution is to conduct a stability analysis. Both the posterior samples from the model estimation procedure and the estimate of the marginal likelihood from the bridge sampling can be variable. Therefore, stability may best be assessed by repeating the entire process of fitting each model and applying the bridge sampling function several times.

We assess the stability of the obtained Bayes factor by repeating the estimating process ten times with `rstan` and `brms`, respectively. The results are shown in Figure \@ref(fig:bsbfestimatesbrmsrstan). The figure shows the evidence in favor of the null, digit and side model against the full model. All Bayes factors are larger than one. Therefore, it can be concluded that the data are more likely under less complex models. There is almost no variation in the computed Bayes factors per model, indicating that the Bayes factor estimate is stable. However, the Bayes factor estimates in `rstan` generally show stronger evidence than Bayes factor estimates based on `brms`. It is important to note that this difference is not due to the bridge sampling function, but a result of a difference in parametrization. However, for both packages, the conclusion is the same: the data are least likely to have occurred under the full model. In case the estimates are unstable, it is recommended to further increase the number of posterior samples.

```{r bsbfestimatesbrmsrstan, echo=FALSE, fig.cap="The BF estimates of the other models (0; null, side and digit model) against the full model (1). The green triangles represent the estimates using \\textit{rstan}. The orange dots represent the \\textit{brms} estimates. The dashed line represents a $\\text{BF}_{01}$ of 1. ", fig.align = "left", fig.height = 11, fig.width = 4, out.width="70%"}
knitr::include_graphics("Images/Fig17_normalbfplot01.png")
```
Finally, it is time to answer our research question: Is there a symbolic distance effect? When considering all Bayes factor estimates from Figure \@ref(fig:sdratioplotall) and \@ref(fig:bsbfestimatesbrmsrstan), the null model ($\text{BF}_{max} = 232 * 10^{105}$ for `rstan`) performs best compared to the side model ($\text{BF}_{max} = 458 * 10^{68}$ for `rstan`), the digit model ($\text{BF}_{max} = 269 * 10^{31}$ for `rstan`). Therefore, the data from Rouder et al. (2005) show evidence against the symbolic distance effect.

## Sensitivity Analysis 
The full model including the side and digit effects performs worst compared to all other models in the Bayes factor model comparison. However, we did find non-zero digit effects in the model estimation. How is this possible?

A possible explanation is a lack of individual variability in the digit and side effects. The full model assumes that participants vary in the digit and side effect, and we specified our expectations about this size of the variability in the prior. However, if there is a lack of individual variability in the effects, the full model is punished for the complexity added by the individual variation (illustrated by Rouder et al., 2021). To investigate this idea, we can reformulate the prior distributions on the variance parameters to indicate a smaller individual variation or, even more extreme, remove these individual effects from the model entirely. Then, we can calculate the posterior model probabilities for all specified models.

Specifically, we will compare models from three scenarios: scenario 1 specifies large individual differences equivalent to Equation \@ref(eq:chosenpriors2var). For the models in scenario 2, we adjust the priors on $\sigma^2_{\beta}$ and $\sigma^2_{\delta}$ in the following way:

\begin{equation}
\begin{aligned}
\sigma^2_{\beta} &\sim \text{Inverse-Gamma}(3, 0.05), \\   
\sigma^2_{\delta} &\sim \text{Inverse-Gamma}(3, 0.05). 
\end{aligned}
\end{equation}
The smaller value for the scale parameter of the inverse gamma distribution (i.e., 0.05 instead of 0.5) results in a shift of the peak of the prior towards smaller values for the individual variance. This means that the variation of the individual effects is expected to be much smaller than before.

In scenario 3, we remove individual variability from the side, digit, and full model. This results in the following full model, from which the side and digit model follow logically:

\begin{equation}
Y_{ijk} \sim \text{Normal}(\gamma_{i} + x_j \mu_\beta + u_j \mu_{\delta{7}} + v_j \mu_{\delta{6}} + w_j \mu_{\delta{4}} + z_j \mu_{\delta{3}}, \sigma^2). (\#eq:datalevel2)
\end{equation}
Note that the null model is the same in all three scenarios.

To obtain the posterior model probability for each model in every scenario, we first compute the Bayes factor of every model against the null model. This includes the trivial Bayes factor of the null model against the null model, resulting in ten Bayes factors in total. Next, we calculate the sum of all these Bayes factors and divide each Bayes factor by this sum. This results in the posterior model probabilities shown in Table 5.^[Note that it would be possible to calculate the posterior model probabilities directly from the marginal likelihoods. However, marginal likelihoods are often not computed explicitly in the model fitting process (e.g., when using the bayes_factor function in the `brms` package), so we believe that readers might find it easier to compute posterior model probabilities from Bayes factors.] All posterior model probabilities from the table and the posterior probability of the null model (see table note) sum to one.

\begin{table}[H]
\caption{Posterior Model Probabilities}
\begin{tabularx}{\textwidth}{XXXX}
\toprule
 & \multicolumn{1}{c}{Scenario 1} & \multicolumn{1}{c}{Scenario 2} & \multicolumn{1}{c}{Scenario 3}\\
\midrule
Digit model & $0.87*10^{-126}$ & $0.29*10^{-46}$ & $0.95$\\
Side model & $0.20*10^{-88}$ & $0.18*10^{-66}$ & $0.77*10^{-53}$ \\
Full model & $0.43*10^{-158}$ & $0.47*10^{-61}$ & $0.05$\\
\bottomrule
\end{tabularx}
\begin{tablenotes}[para]
\textit{Note.} Scenario 1: large individual differences. Scenario 2: small individual differences. Scenario 3: no individual differences. The posterior model probability of the null model is the same in every scenario and equals $0.50*10^{107}$. 
\end{tablenotes}
\end{table}

The table shows that the digit model without individual digit effects from scenario 3 has the highest posterior model probability, followed by the full model without individual digit and side effects (scenario 3). This means that the data are most likely under these models than the other eight models included in the comparison. In addition, the models with smaller prior expectations for individual differences (scenario 2), have higher posterior model probabilities than the models in scenario 1. Thus, decreasing our expectations of the size of the individual digit and side effects, or completely removing them, results in higher posterior model probabilities. Together with Figure \@ref(fig:randomparfigure), showing the small differences in individual effects (also illustrated by Haaf et al., 2019), this result supports the notion that side and digit effects are small and there is a lack in individual variability.

This example illustrates the benefits of comparing hierarchical models of varying complexity. While individual differences are a common phenomenon in psychology, there are instances where general effects are sufficient to describe the data adequately, or where only a random intercept, but no random slopes are necessary. As shown in Table 5, Bayesian model comparisons can help researchers to identify the model that best describes the data.

# Discussion

<!-- To conclude this tutorial on Bayesian hierarchical modeling in psychology, we will highlight the most important recommendations to keep in mind when applying a Bayesian hierarchical model.   -->

In this tutorial, we described model and prior specification, estimation, and model comparison for hierarchical models in the Bayesian framework. In the following, we want to present several key recommendations.

As prior distributions influence model estimation and model comparison, it is important to choose suitable priors. We believe that simulating data based on the chosen priors is an intuitive tool to find reasonable prior settings that take the existing knowledge in the specific model application context into account. We showed that it can be problematic to use default priors without checking their suitability. Therefore, we recommend performing prior predictive checks with the priors that will be applied before running the analysis.

We illustrated different methods to obtain Bayes factors. We want to emphasize the importance of checking the variability of the Bayes factor estimate, the Bayes factor sensitivity to prior distributions, and to ensure that the Bayes factor estimate is based on a sufficient number of posterior samples.

This tutorial showed the application of two R packages: `rstan` and `brms`. Which package to use depends mostly on the complexity of the model and your technical background. For the non-technical user, we recommend using `brms` because of its intuitive model specification. However, the default prior setup can be unsuitable for a specific research context and might need to be adapted. Additionally, users are limited to predefined parametrizations. `rstan` offers the most flexibility in model and prior specification. This flexibility comes with a cost: if the prior or the model is not well defined, results can be misleading.

## Further Recommendations

With this tutorial, we attempted to cover the most relevant topics that researchers encounter when starting with Bayesian hierarchical modeling. However, there are still many remaining issues and recommendations that we could not discuss. Here, we would like to point to further references in the literature if researchers want to continue mastering Bayesian hierarchical modeling.

First, we kept the discussion of the software packages and code fairly brief. We provide further code examples for the digit classification task in the Online Supplement. For different models, it may be useful to study additional tutorials and books. For `rstan`, we recommend the `Stan` manual [@stan2018a; @stan2018b]. For `brms` there are several useful tutorials [@R-brms_a; @R-brms_b; @burkner2019ordinal]. There are also several resources for Bayesian estimation with other samplers [@lee2013bayesian; @rouder2013hierarchical]. Two other useful tutorials for Bayesian modeling are Schad, Betancourt, and
Vasishth [-@schad2019toward] and Schad, Nicenboim, Bürkner, Betancourt, and Vasishth [-@schad2021workflow].

Second, we did not discuss study planning. When planning an experiment, researchers may want to determine the number of participants required to gain conclusive results. Bayesian analysis has the advantage that an optional stopping paradigm may be employed [@rouder2014optional; @schonbrodt2017sequential]. Schönbrodt and Wagenmakers [-@schonbrodt2018bayes] and Stefan and colleagues [-@stefan2019tutorial] introduced Bayes factor design analysis for fixed and sequential designs. However, these approaches to study planning have not yet been fully extended to a hierarchical setup [but see recent efforts by Vasishth et al., -@vasishth2021sample]. The extension to nested designs is not entirely trivial because, in addition to planning the number of participants, we also need to consider the number of trials [@rouder2018power]. While we think there remains some open questions on the topic, the tutorial by Stefan and colleagues [-@stefan2019tutorial] is a good place to start for applied researchers.

## Potential Technical Issues

One key advantage of Bayesian over frequentist hierarchical modeling is that complex modeling leads to fewer convergence issues in the Bayesian framework. This advantage, however, does not imply that there are no technical issues in the Bayesian framework. One possible technical issue is with the choice of priors when using `Stan` for estimation. The `Stan` algorithm is most efficient with specific prior distributions. For example, it may be beneficial to use a truncated $t$-distribution on standard deviation parameters instead of an inverse gamma distribution on variances [@gelman2006prior]. Using priors not recommended by the `Stan` team may lead to more issues with convergence. In this tutorial, we highlight that choosing priors based on their predictions on data is beneficial. We still think this is the best path for substantive researchers. If one wants more flexibility in choices of priors or has continuous issues with convergence with their chosen prior distributions, then it might be worth considering using `JAGS` instead of `Stan`. `JAGS` [@Plummer03jagsa] has a similar functionality and syntax as `Stan`. While `Stan` tends to be more efficient in the sampling phase, especially in case of correlated parameters [@hecht2021comparing], the `JAGS` algorithm could be more flexible and less affected by the choice of priors.

Another technical issue may arise with the estimation of Bayes factors. Here, we highlighted two approaches, the Savage-Dickey density ratio [@wagenmakers2010bayesian] and bridge sampling [@gronau2017tutorial]. We already highlighted the drawbacks of the SD density ratio. Yet, bridge sampling may also fail in some cases. Failures of the bridge sampling algorithm are identified by unstable Bayes factors across repeated sampling runs. There are improvements to the algorithm that are specifically developed for hierarchical models [@gronau2019simple]. Yet, it remains crucial to assess the stability of Bayes factor estimates.

## Living with Uncertainty

Let us reconsider our scenario from the introduction. The graduate student may have read our tutorial and went back to their supervisor to propose a Bayesian hierarchical analysis of their data. Compared to the initially considered ANOVA analysis, there are many small decisions to be made for this more extensive analysis. These small decisions tend to make researchers uncomfortable. What prior is the right one? Are there enough posterior samples?

In Bayesian modeling, we have to learn living with these uncertainties. In fact, we recommend embracing them as these small decisions. First, we at the very least are making these decisions deliberately. Second, we can check their impact on the results. Throughout the tutorial we highlighted several potential robustness checks, from prior predictions over convergence assessment, to sensitivity analysis for Bayes factors. These checks may help researchers to understand how robust their results are and when they break. Therefore, we hope the checks will support switching from simple procedures to Bayesian hierarchical modeling. Because ultimately, the results from such an analysis are much richer.

In summary, by offering a comparison of software packages, guidance on prior selection and default priors, and assessment of the Bayes factor as model comparison method, we hope that Bayesian hierarchical models will become available to a wider psychology public.

\newpage

# Acknowledgement

We would like to express our gratitude to the R developers [Version 3.6.1; @R-base] and the following R package developers:

- Document layout: *rmarkdown* [Version 1.16; @R-rmarkdown], *papaja* [Version 0.1.0.9842; @R-papaja], *knitr* [Version 1.25; @R-knitr], *kableExtra* [Version 1.1.0; @R-kableExtra]
- Data analysis and cleaning: *LaplacesDemon* [Version 16.1.4; @R-LaplacesDemon], *plyr* [Version 1.8.4; @R-dplyr; @R-plyr], *dplyr* [Version 0.8.3; @R-dplyr], *readr* [Version 1.3.1; @R-readr], *crch* [Version 1.0.4; @R-crch_a; @R-crch_b; @R-crch_c; @R-crch_d], *foreign* [Version 0.8.71; @R-foreign], *Matrix* [Version 1.2.17; @R-Matrix], *tidyr* [Version 1.0.0; @R-tidyr]
- Model fit and comparison: *rstan* [Version 2.19.2; @R-rstan], *brms* [Version 2.13.0; @R-brms_a; @R-brms_b], *BayesFactor* [Version 0.9.12.4.2; @R-BayesFactor], *bridgesampling* [Version 1.0.0; @R-bridgesampling]
- Visualization: *coda* [Version 0.19.3; @R-coda], *lattice* [Version 0.20.38; @R-lattice], *ggplot2* [Version 3.3.2; @R-ggplot2], *Rmisc* [Version 1.5; @R-Rmisc], *devtools* [Version 2.2.1; @R-devtools], *gghalves* [Version 0.1.0; @R-gghalves], *bayesplot* [Version 1.7.1; @R-bayesplot], *gridExtra* [Version 2.3; @R-gridExtra], *tibble* [Version 3.0.2; @R-tibble], *ggmcmc* [Version 1.4.1; @R-ggmcmc], *cowplot* [Version 1.0.0; @R-cowplot], *ggridges* [Version 0.5.1; @R-ggridges], *circlize* [Version 0.4.9; @R-circlize], *StanHeaders* [Version 2.19.0; @R-StanHeaders], *unikn* [Version 0.2.0; @R-unikn]
- Other: *Rcpp* [Version 1.0.3; @R-Rcpp_a; @R-Rcpp_b], *usethis* [Version 1.5.1; @R-usethis]

\newpage

# References
```{r create_r-references}
r_refs(file = "r-references.bib")
# cite_r("r-references.bib")  # to obtain package versions 
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs">

</div>
\endgroup

\newpage

### 2022-06-09 manuscript.tex
All documents except .rds

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  doc,floatsintext]{apa6}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Bayesian Hierarchical Modeling: An Introduction and Reassessment},
  pdfauthor={Myrthe Veenman1, Angelika Stefan1, \& Julia Haaf1},
  pdflang={en-EN},
  pdfkeywords={Multilevel, Tutorial, rstan, brms, Bayes factor},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother
\shorttitle{Bayesian Hierarchical Modeling}
\keywords{Multilevel, Tutorial, rstan, brms, Bayes factor}
\usepackage{csquotes}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{xcolor}
\definecolor{mypink}{RGB}{255, 230, 255}
\usepackage{todonotes}
\newcommand{\Julia}[1]{\todo[inline, color=mypink]{#1}}
\usepackage{graphicx}
\usepackage{float}
\usepackage{tabularx}
\usepackage{hyperref}
\ifxetex
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\else
  \usepackage[main=english]{babel}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\fi
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\title{Bayesian Hierarchical Modeling: An Introduction and Reassessment}
\author{Myrthe Veenman\textsuperscript{1}, Angelika Stefan\textsuperscript{1}, \& Julia Haaf\textsuperscript{1}}
\date{}

\authornote{

This paper is written in R Markdown. The R Markdown file, other R code, and additional information on this study can be found in the Online Supplement.

Correspondence concerning this article should be addressed to Myrthe Veenman, . E-mail: \href{mailto:myrthe.veenman@gmail.com}{\nolinkurl{myrthe.veenman@gmail.com}}

}

\affiliation{\vspace{0.5cm}\textsuperscript{1} University of Amsterdam}

\abstract{
With the recent development of easy-to-use tools for Bayesian analysis, psychologists have started to embrace Bayesian hierarchical modeling. Bayesian hierarchical models provide an intuitive account of inter- and intraindividual variability and are particularly suited for the evaluation of repeated-measures designs. Here, we provide guidance for model specificaton and interpretation in Bayesian hierarchical modeling and describe common pitfalls that can arise in the process of model fitting and evaluation. Our introduction gives particular emphasis to prior specification and prior sensitivity, as well as to the calculation of Bayes factors for model comparisons. We illustrate the use of state-of-the-art software programs \texttt{Stan} and \texttt{brms}. The result is an overview over best practices in Bayesian hierarchical modeling that, as we hope, will help psychologists in making the best use of Bayesian hierarchical modeling.
}

\begin{document}
\maketitle

\newpage

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Imagine a typical scenario in the life of an experimental psychology graduate student: You have just collected a large data set, either online or in the lab. Each participant has spent one valuable hour completing many trials per experimental condition. Now it is time to analyze the data, and your supervisor suggests to run a within-subjects ANOVA. This procedure requires you to first aggregate the data by condition and participant. Doing so reduces your data set from 10,000 rows to just 400. Reducing the data by aggregation makes you a bit queasy: It feels like your efforts in data collection and the participants' time are not valued this way. You know there needs to be a better way. And there is: Hierarchical modeling.

Just as in the scenario above, experiments in psychology are often implemented in a repeated-measures design where participants respond to several items, stimuli, or conditions. From a data analysis perspective, such a design implies that observations are nested within participants. The statistical and the psychological literature agree that the optimal analysis accounting for such a data structure is hierarchical modeling (e.g., Efron \& Morris, 1977; Lee, 2011; Rouder \& Lu, 2005). And yet, it is still not the norm to apply hierarchical modeling to data from psychological experiments. One reason for this is that hierarchical modeling is more difficult than familiar procedures such as ANOVA.

With recent software development, hierarchical modeling has become more and more accessible. In addition, many researchers have argued that hierarchical modeling is easier and the interpretation is more intuitive when it is done in the Bayesian statistical framework (e.g., Lynch, 2007; Rouder \& Lu, 2005; Rouder, Morey, \& Pratte, 2013).
However, there are challenges to the Bayesian way of hierarchical modeling (Rouder \& Lu, 2005). These challenges, such as the choice of priors, the need for model comparisons of highly complex models, and programming skills required to use state-of-the-art software solutions, can make researchers hesitant to use this approach. We believe that existing introductions to Bayesian hierarchical modeling do not adequately address these challenges. For example, they use outdated software (Rouder et al., 2013), do not cover model comparison (Rouder \& Lu, 2005), or do not discuss prior specification (Shiffrin, Lee, Kim, \& Wagenmakers, 2008). This paper aims to make Bayesian hierarchical modeling available to a wider public by providing:

\begin{itemize}
\tightlist
\item
  A comparison of and guidance for the use of two software packages,\\
\item
  Guidance on prior specification,
\item
  Practical instructions for Bayes factor model comparison.
\end{itemize}

Our tutorial is directed at researchers who have a basic understanding of Bayesian inference and parameter estimation and are planning to perform more complex Bayesian analyses.

The application of Bayesian hierarchical analysis will be illustrated using a digit classification task collected by Rouder, Lu, Speckman, Sun, and Jiang (2005). In an experimental task participants repeatedly indicated whether a target digit (i.e., 2, 3, 4, 6, 7, or 8) was greater or smaller than 5. The participants in the task performed multiple trials per condition (i.e., digit). Therefore, observations are nested in participants. Our analyses focus on two effects that can potentially occur in this digit classification task: The digit effect and the side effect. According to the symbolic distance hypothesis (Moyer \& Landauer, 1967) processing of numbers is analogue. If numbers are closer on the number line, we may confuse them more, and comparing them is more difficult. In line with this hypothesis, the digit effect postulates that response times (RTs) are slower when digits are closer to 5. Additionally, we may hypothesize that RTs are affected by whether the target digit was smaller or greater than 5, yet it is unclear which side of 5 would lead to faster or slower RTs. The digit classification task is a standard task that usually would be analyzed in a 3 (digit) by 2 (side) within-subjects ANOVA. In this tutorial, we show how inferences can be improved using a hierarchical Bayesian model instead.

\begin{figure}[H]

\includegraphics[width=0.95\linewidth]{Images/Fig1_meanrtandhierplot_v2} \hfill{}

\caption{Panel A shows the response times in seconds for two conditions of the number classification task (i.e., digits 3 and 4). The grey lines represent the individual effects, and the blue line the aggregated effect. Panel B shows how individual estimates are shrunken towards the general trend. Sample represents the observed effect per participant in seconds, while Model represents the estimated effects per participant. }\label{fig:figure1}
\end{figure}

\hypertarget{why-multilevel-modeling}{%
\subsection{Why Multilevel Modeling?}\label{why-multilevel-modeling}}

Before diving into hierarchical modeling, let us consider two alternative approaches: Analyzing aggregated data and analyzing the data separately for each participant. Analyses of aggregated data such as the ANOVA from the scenario above examine a general effect that is assumed to be consistent across subjects. For instance, when investigating whether there is an effect of digit on RT, the mean RT for each participant is computed for each digit-condition, and a repeated-measures ANOVA is conducted on these aggregated scores. Analyses of aggregated data cannot investigate whether a trend is consistent across participants. Therefore, when this option is applied and individual variability is ignored, the general trend can be an inaccurate representation of the true general trend (Rouder, Morey, \& Pratte, 2013; Haaf \& Rouder, 2017; van Doorn, Aust, Haaf, Stefan, and Wagenmakers, 2021). This is shown in Panel A of Figure \ref{fig:figure1} where some of the individual trends (i.e., slopes of the grey lines) differ from the general trend (i.e., slope of the blue line). If every participant takes part in each condition, the person variability affects all conditions, thus inducing a correlation across conditions. For example, participants in a symbolic distance effect experiment repeatedly respond to target digits inducing a correlation between the responses to the target digits, such as between target digit 3 and target digit 4. Therefore, the observations are not independent as we assume if we perform analysis on aggregated data. As a result, the type I error rate of a test of the effect increases: there is a higher probability of concluding that there is an effect if there is no effect (Rouder \& Lu, 2005).

A second alternative analysis method is to conduct an analysis on each participant's data independently. For example, in the digit classification task, we would estimate one digit effect per person (cf.~the grey lines in Figure 1A), but no overall effect (cf.~the blue line in Figure 1A). Therefore, when analyzing participant data independently, it is hard to draw any general conclusions.

In hierarchical analyses, the general or overall effect is called fixed effect, and the individual-specific deviations from this effect are called random effects (Rouder \& Lu, 2005). We can also estimate individual effects, not as deviations from an overarching effect, but as effects for each individual, as obtained from a per-participant analysis. In this tutorial, we will use the terms general effect, individual deviation, and individual effects instead of fixed and random effects, because we believe these terms are better suited for the interpretation of our analyses.

If estimated for each participant independently, observed individual effects are quite variable because they are perturbed by sample noise. In hierarchical modeling, these individual effects are optimally and automatically corrected towards the general trend (Efron \& Morris, 1977). This phenomenon, also called ``shrinkage,'' is illustrated in Panel B of Figure \ref{fig:figure1}. The rightmost observation in the figure is in sample quite large compared to the other observed effects. The estimated effect (green point) is drastically corrected towards the mean. Conversely, the general trend is only influenced slightly by individuals showing a highly divergent trend. By adding the individual variability to the model, we no longer assume independence between observations, but the overall effect in the model leaves the possibility to generalize the effect. Therefore, hierarchical analysis can be considered a middle ground between an independent analysis and an analysis of aggregated data.

The Bayesian framework offers an intuitive approach for hierarchical modeling (Rouder \& Lu, 2005; Lynch, 2007). The flexibility of the Bayesian approach facilitates the implementation of complex hierarchical models (Rouder, Morey, \& Pratte, 2013). It also allows for the inclusion of prior knowledge into the analysis, as well as the monitoring of evidence for specific hypotheses as the data accumulate (Wagenmakers et al., 2018). A complete discussion of the Bayesion approach is beyond the scope of this manuscript. We refer the interested reader to Etz and Vandekerckhove (2018) and Wagenmakers and colleagues (2018).

\hypertarget{model-specification}{%
\section{Model Specification}\label{model-specification}}

In the following section, we specify two models that can be used for our application example: a normal and a log-normal model.\footnote{Online Supplement A provides the parameterization for models of aggregated and individual data, to illustrate the difference with hierarchical models.} We are aware that many psychologists do not engage with equations on a daily basis. In this section, however, we try to accessibly summarize the key considerations of model specification. We encourage everyone to read on and not skip the section. Additionally, we provide a non-technical verbal summary.

Throughout the section, we assume that each observation \(Y_{ijk}\) is the response time from participant \(i\) in trial \(k\) of condition \(j\). The condition \(j\) is the digit presented to the participant, and can take the values 2, 3, 4, 6, 7, and 8. We further assume that the digit side (\(x\)) is recorded using effect coding where a value of -0.5 indicates that the digit is smaller than 5, and a value of 0.5 indicates that the digit is larger than 5. Additionally, we assume that the digits presented in the trial are dummy-coded in the data using four variables (\(u\), \(v\), \(w\), \(z\)). If digits 2 and 8 were presented all four variables are zero. The dummy variables represent the digits 3 (\(u\)), 4 (\(v\)), 6 (\(w\)), and 7 (\(z\)), and each take the value 1 if the respective digit was presented in the trial, and 0 if it was not.

\hypertarget{normal-model}{%
\subsection{Normal Model}\label{normal-model}}

In the normal model, the observation \(Y_{ijk}\) is assumed to come from a normal distribution. The mean of this normal distribution is predicted by the person-specific intercept \(\gamma\), the effect of the side (i.e., smaller or greater than 5, \(\beta\)) and the effect of the distance from 5 (i.e., symbolic distance effect, \(\delta\)'s). The variance is represented by \(\sigma^2\). This is captured in the following equation:

\begin{equation}
Y_{ijk} \sim \text{Normal}(\gamma_{i} + x_j \beta_{i} + u_j \delta_{7i} + v_j \delta_{6 i} + w_j \delta_{4i} + z_j \delta_{3i}, \sigma^2), \label{eq:datalevel}
\end{equation}

where

\begin{align*}
&
\begin{aligned}
x_j = 
\begin{cases}
\frac{1}{2} & j < 5\\
- \frac{1}{2} & j > 5, \\
\end{cases}
\end{aligned}
&
\begin{aligned}
u_j = 
\begin{cases}
1 & j = 7\\
0 & \text{Otherwise},
\end{cases}
\end{aligned}
&
\begin{aligned}
v_j = 
\begin{cases}
1 & j = 6\\
0 & \text{Otherwise},
\end{cases}
\end{aligned}
\end{align*}

\begin{align*}
&
\begin{aligned}
w_j = 
\begin{cases}
1 & j = 4\\
0 & \text{Otherwise},
\end{cases}
\end{aligned}
&
\begin{aligned}
z_j = 
\begin{cases}
1 & j = 3\\
0 & \text{Otherwise}.
\end{cases}
\end{aligned}
\end{align*}
The model equation shows that, for instance, for participant \emph{i} = 1 and digit \emph{j} = 7, the mean RT is modeled by: \(\gamma_{i = 1} - \frac{1}{2} \beta_{i = 1} + \delta_{7,i = 1}\). If all \(\delta\)s are positive, then the model equation indicates that individuals respond fastest when the digits 2 and 8 are presented, as in this situation there are no digit effects in the equation.

The intercept \(\gamma_i\), the side effect \(\beta\), and the digit effects \(\delta\) are person-specific. This means that for these parameters, the model will produce one estimate for each participant. The multilevel model assumes that these individual-level effects in turn follow a normal distribution:

\begin{equation}
\begin{aligned}
\gamma_{i} &\sim \text{Normal}(\mu_{\gamma}, \sigma^2_{\gamma}), \label{eq:fixedvsrandom}\\
\beta_{i} &\sim \text{Normal}(\mu_{\beta}, \sigma^2_{\beta}), \\
\delta_{7i} &\sim \text{Normal}(\mu_{\delta_{7}}, \sigma^2_{\delta_{7}}), \\
\delta_{6i} &\sim \text{Normal}(\mu_{\delta_{6}}, \sigma^2_{\delta_{6}}), \\
\delta_{4i} &\sim \text{Normal}(\mu_{\delta_{4}}, \sigma^2_{\delta_{4}}),\\
\delta_{3i} &\sim \text{Normal}(\mu_{\delta_{3}}, \sigma^2_{\delta_{3}}).\\
\end{aligned}
\end{equation}
The means (\(\mu_{\gamma}\), \(\mu_{\beta}\), \(\mu_{\delta}\)) of these normal distributions represent the general expected effect, for example, \(\mu_{\beta}\) represents the expected effect of digit side across individuals. The variances (\(\sigma^2_{\gamma}\), \(\sigma^2_{\beta}\), \(\sigma^2_{\delta}\)) represent the individual variation. Therefore, the individual level depends on another level, the overall effect level, containing \(\mu_{\gamma}\), \(\mu_{\beta}\), \(\mu_{\delta}\), \(\sigma^2_{\gamma}\), \(\sigma^2_{\beta}\), and \(\sigma^2_{\delta}\).

\hypertarget{log-normal-model}{%
\subsection{Log-normal Model}\label{log-normal-model}}

A normal model may not be appropriate in case of more complex theories, hypotheses, or data structures. For example, the data in the example are response times. The normal model, however, can yield negative values, which seems very problematic. In addition, RT distributions are right-skewed. A better representation of the data at hand is therefore given by a log-normal model (Schramm \& Rouder, 2019). This means that Equation \eqref{eq:datalevel} has to be adjusted:

\begin{equation}
Y_{ijk} \sim \text{LogNormal}(\gamma_{i} + x_j \beta_{i} + u_j \delta_{7i} + v_j \delta_{6i} + w_j \delta_{4i} + z_j \delta_{3i}, \sigma^2).  
\end{equation}
Note that the only change in the model formulation is the change of the likelihood function from a normal distribution to a log-normal distribution. The location parameter of the log-normal distribution is described through the same combination of parameters as the mean of the normal model. However, it has a different interpretation. For example, for a location parameter of \(\mu = 0\) and a scale parameter of \(\sigma^2 = 1\), the mode of the log-normal distribution is \(0.368\), and the mean is \(1.65\).

\hypertarget{summary}{%
\subsection{Summary}\label{summary}}

In summary, the models are placed directly on the raw RT data. The only difference between the normal and the log-normal models is the likelihood distribution. The normal model just assumes a symmetric normal distibution that can take on any value. The log-normal model assumes a log-normal distribution, which is restricted to positive values and has a right skew.

Both the normal and the log-normal models contain the following parameters: \(\gamma_i\) corresponds to the participant's overall response time, \(\beta_i\) corresponds to the participant's side effect, and the four \(\delta_{\cdot i}\) correspond to the effects between the digits furthest away from 5 (2 or 8) and the other digits. These person-specific effects all come from parent distributions. This results in the hierarchical structure shown in Figure \ref{fig:hierarchyplot}. The figure highlights the relationship between the parameters and the data.

In Bayesian modeling, prior distributions are needed to complete model specification. In the next section, we will take you through the process of setting these priors, followed by sections on Bayesian estimation and hypothesis testing for both the normal and log-normal model. In every section, we will first explain the normal model, followed by an illustration of the log-normal model.

\begin{figure}[H]

\includegraphics[width=0.7\linewidth]{Images/Fig2_Hierarchical structure} \hfill{}

\caption{Hierarchical structure of the normal and lognormal models for the number classification task.}\label{fig:hierarchyplot}
\end{figure}

\hypertarget{prior-settings}{%
\section{Prior Settings}\label{prior-settings}}

Bayesian analysis requires the specification of prior distributions. Prior distributions are probability distributions on model parameters that specify beliefs about the relative plausibility of parameter values before seeing the data (Wagenmakers et al., 2018). These prior beliefs can then be updated with the data to obtain the posterior beliefs about the model parameters, following Bayes' rule (Jeffreys, 1961).

For our models, priors are needed on the general effects (\(\mu_{\gamma}\), \(\mu_{\beta}\), \(\mu_{\delta}\)) and on the variances (\(\sigma_{\gamma}^2\), \(\sigma_{\beta}^2\), \(\sigma_{\delta}^2\), and \(\sigma^2\)), as shown in Equation \eqref{eq:fixedvsrandom}. For each of these parameters, we specify our beliefs as a probability distribution. For example, if we expect small side effects in the digit classification example, the prior distribution on \(\mu_{\beta}\) should have its peak close to zero. Large values for the variance parameters indicate large individual differences, whereas values close to zero indicate that effects are similar across individuals (Haaf \& Rouder, 2017). If we expect little variability of side effects between people, the prior assigned to \(\sigma_{\beta}\) should also have its peak close to zero. The width of prior distributions expresses how certain we are about the parameters before seeing the data. For example, if we want to express that we are certain that the side effect is zero, we can formulate a prior that allocates all probability mass to a parameter value of zero.

\hypertarget{fear-of-commitment}{%
\subsection{Fear of Commitment}\label{fear-of-commitment}}

In many cases substantive researchers fear committing to prior settings because they are unsure about their choices for the type and settings of the distribution. Therefore, they tend to choose wide priors to indicate little prior knowledge about the parameter values (Aczel et al., 2018). These include default priors offered by programming packages, such as \texttt{rstan} (Stan Development Team, 2019a; illustrated in Online Supplement B) and \texttt{brms} (Bürkner, 2017, 2018). Wide priors are priors that allocate (approximately) equal probability density to a great range of values, such as a uniform distribution without bounds (Gelman, 2006). However, wide priors can be problematic. First, since the distribution is spread out across a wide range of values, the probability density on any specific parameter value is very low. Therefore, any effect is unlikely under the prior as a wide range of effects is deemed plausible. This influences model comparisons: the support for the null hypothesis becomes increasingly large. This is also referred to as the Jeffreys-Lindley paradox (Lindley, 1957).

Another important issue is that wide priors are often improper, that is, the probability density does not integrate to one (Hobert \& Casella, 1996). An example of an improper prior is a uniform prior distribution ranging from minus infinity to infinity. In case of improper priors, an estimation of the general and individual effects cannot be provided. However, even proper priors can result in improper posteriors (Hobert \& Casella, 1996). Although estimation should not be possible in this situation, statistical software might still provide results without notifying the user that the posterior distribution does not exist. If users are not aware of this problem, this can result in misleading conclusions. Therefore, it is important to check if \emph{(a)} the priors are proper and \emph{(b)} the priors result in proper posteriors. There is no fail-safe method that guarantees proper posteriors for hierarchical models. The best approach is to choose well-reasoned informative priors for the model, for instance, by using the proposed four-step procedure below.

\hypertarget{procedure}{%
\subsection{Procedure}\label{procedure}}

We propose four steps to choose suitable priors for Bayesian hierarchical models. We first present the steps and then apply them to choose suitable priors for the parameters \(\mu_{\gamma}\), \(\mu_{\beta}\), \(\mu_{\delta}\), \(\sigma_{\gamma}^2\), \(\sigma_{\beta}^2\), \(\sigma_{\delta}^2\) and \(\sigma^2\) in our example.

\begin{figure}[H]

{\centering \includegraphics{manuscript_files/figure-latex/distributionsteps-1}

}

\caption{Prior specification steps.}\label{fig:distributionsteps}
\end{figure}

\hypertarget{step-1-distribution-family}{%
\subsection{Step 1: Distribution Family}\label{step-1-distribution-family}}

The first step in specifying a prior distribution is to find a suitable distribution family. For instance, the prior plausibility of parameter values can be expressed through a normal, an inverse gamma, or a Student's \emph{t}-distribution. These distributional families have different shapes and a different support, as illustrated in Figure \ref{fig:distritypeplot}. For example, the inverse gamma distribution is limited to non-negative parameter values and has a positive skew. The normal and Student's \emph{t}-distribution both have support over the real line, however, the Student's \emph{t}-distribution has wider tails, that is, it assigns a higher plausibility to values that are on the extremes of the distribution.

The support of the prior distribution should cover all theoretically possible parameter values. For example, if a parameter cannot have negative values, such as a standard deviation, the inverse gamma distribution would be a suitable choice. Conversely, if a parameter can take positive and negative values, the normal and Student \emph{t}-distribution would be suitable.

Note that although distributions within a distributional family share the same functional form, they can differ a lot in their shape and location. The exact appearance of a prior distribution is controlled by a set of hyperparameters, for example, the mean and variance for the location and scale of a normal prior. In the next section, we will give some guidance for specifying these hyperparameters.

\begin{figure}[H]

\includegraphics[width=1\linewidth]{Images/Fig4_distrtypefigure_v3} \hfill{}

\caption{Common probability density distributions that may be used as priors. Panel A shows normal distributions varying in mean and standard deviation. Panel B shows inverse gamma distributions that differ in the shape and scale parameter. Panel C shows Student's \textit{t}-distributions that differ in the location and scale parameter. For all three \textit{t}-distributions, the degrees of freedom are set to 3.}\label{fig:distritypeplot}
\end{figure}

\hypertarget{step-2-set-hyperparameters}{%
\subsection{Step 2: Set Hyperparameters}\label{step-2-set-hyperparameters}}

Location, scale, and sometimes shape of the selected distribution can be adjusted by choosing adequate hyperparameters. The goal is to adjust the prior to reflect which parameter values we deem most plausible and how much uncertainty there is. For instance, panel A of Figure \ref{fig:distritypeplot} illustrates, that increasing the mean in the normal distribution leads to a shift of location: higher parameter values are deemed more plausible according to the pink distribution compared to the green distribution. Panel B shows how uncertainty can be specified, for example, by decreasing the shape of the inverse gamma distribution while keeping the scale constant: the orange distribution assigns probability to a wider range of values than the green distribution.

For this step, it is important to consider the scale on which variables are measured. For instance, RTs could be measured in seconds or in milliseconds. This means that if a researcher expects a response time of one second, they will formulate a prior distribution with a mean of 1 if RT is measured in seconds, or formulate a prior distribution with a mean of 1000 if RT is measured in milliseconds. Similarly, the measurement scale will also affect the width of the specified prior.

\hypertarget{step-3-prior-predictions}{%
\subsection{Step 3: Prior Predictions}\label{step-3-prior-predictions}}

A key advantage of model specification with priors is that we can obtain preditions on what the data would look like that are generated from the model. These prior predictions allow us to make the implications of our specified priors more concrete. Researchers can look at relevant statistic that summarize the simulated data and decide if they correspond to reasonable observations for their specific research context. If the simulated data are in line with the researchers' expectations, the priors can be used for the analysis. If the simulated data do not match the expectations, the priors have to be adjusted (Step 4).

\begin{figure}[H]

\includegraphics[width=1\linewidth]{Images/Fig5_priorpredfigure_v2} \hfill{}

\caption{Side and digit effects from predicted data. Violins represent the distribution of predicted effects across 1000 simulation runs. The blue points are the average effects across simultions. Panel A shows the prior predictions for the normal model based on the default priors of \textit{brms}. Panel B shows the prior predictions for the normal model based on the selected priors. Panel C shows the prior predictions for the lognormal model based on the priors chosen for the normal model. Panel D shows the prior predictions for the lognormal model based on the selected priors.}\label{fig:priorpredfigurepanel}
\end{figure}

The simulation proceeds as follows: First, one random draw from each prior distribution is taken. This is plugged into the individual effect distributions (Equation (2)). Then, the likelihood of the model, such as specified in Equations (1) and (3), is used to simulate observations per condition. This means that for \emph{n} participants, \emph{n} random values for the model parameters are generated based on the priors, which will result in the \emph{k} observations per \emph{j} conditions. This process is repeated \emph{m} times, where \emph{m} is the number of simulations drawn. Note that the more simulations are conducted, the closer one gets to the full picture because enough samples from the prior distribution are needed. Finally, we can visualize the simulated data. Figure \ref{fig:priorpredfigurepanel} shows prior predictions for the key effects from the number classification task. For this task, we mainly care about the side effect and digit effects corresponding to the \(\beta\)s and the \(\delta\)s from the model. Here, the prior predictions for the average side effect and the average digit effect across participants are depicted.

For step 3, the intention is not to compare the actual data of interest to the data predicted by the priors. Instead, the researchers' \emph{expectations} of plausible data are compared to the prior predictions. Therefore, prior predictive checks are conducted \emph{before} looking at the actual data. If the prior predictions match the expectations, the priors can be used for the analysis. If they do not match the expectations, the priors need to be adjusted.

\hypertarget{step-4-adjust}{%
\subsection{Step 4: Adjust}\label{step-4-adjust}}

Based on the results of the prior predictions, it might be necessary to adjust the priors. Typically, the focus of adjustments will be on finding adequate parameter values for the prior distribution, that is, on repeating step 2 and the following steps of the prior specification process.

In the following two sections, we will provide a walk-through of all four steps of prior specification for our two models of the symbolic distance effect.

\hypertarget{normal-model-1}{%
\subsection{Normal Model}\label{normal-model-1}}

In the symbolic distance model, we have to specify priors for the parameters \(\mu_{\gamma}\), \(\mu_{\beta}\), \(\mu_{\delta}\), \(\sigma^2_{\gamma}\), \(\sigma^2_{\beta}\), and \(\sigma^2_{\delta}\). The parameters \(\mu_{\gamma}\), \(\mu_{\beta}\), and \(\mu_{\delta}\) can be positive or negative. All variance terms (\(\sigma_{\gamma}^2\), \(\sigma_{\beta}^2\), \(\sigma_{\delta}^2\), and \(\sigma^2\)) are constrained to be positive. If there is no information from previous studies, researchers might be inclined to use the default priors in \texttt{brms} as a starting point.

In \texttt{brms}, different default priors are set on different types of parameters (Bürkner, 2017). The prior on the group-level regression weights (\(\mu_{\beta}\), \(\mu_{\delta}\)) is a normal distribution with a mean of 0 and a standard deviation of 1. The prior on the intercept \(\mu_{\gamma}\) is a scaled and shifted Student's \emph{t}-distribution with 3 degrees of freedom, a location parameter of 1, and a scale parameter of 10. \texttt{brms} specifies priors on standard deviations (\(\sqrt{\sigma^2}\)) instead of variances. For the normal model, this means that priors are set on \(\sigma_{\gamma}\), \(\sigma_{\beta}\), \(\sigma_{\delta}\) and \(\sigma\), instead of on \(\sigma_{\gamma}^2\), \(\sigma_{\beta}^2\), \(\sigma_{\delta}^2\) and \(\sigma^2\). The default prior on the standard deviations is a central, scaled Student's \emph{t}-distribution with 3 degrees of freedom and a scale parameter of 10 that is truncated at zero. To summarize, the default priors would be set the following way:

\begin{equation}
\begin{aligned}
\mu_{\gamma} &\sim \text{Student's t}(3, 1, 10), \label{eq:chosenpriors1} \\ 
\mu_{\beta}, \mu_{\delta} &\sim \text{Normal}(0, 1), \\ 
\sigma, \sigma_{\gamma}, \sigma_{\beta}, \sigma_{\delta} &\sim \text{Truncated-Student's t}_{+}(3, 0, 10). 
\end{aligned}
\end{equation}
The prior distributions \texttt{brms} specifies on the standard deviations can be transformed to prior distributions on variances. Prior distributions on variances have a smaller scale than priors on standard deviation. Note, however, that the transformation of the prior distributions is not as simple as the transformation of the parameter itself. In our Online Supplement, we provide R-code with a numeric solution to the transformation. For the \texttt{brms} priors, we obtain the following implied prior distributions on variances:

\begin{equation}
\begin{aligned}
\sigma_{\gamma}^2, \sigma_{\beta}^2, \sigma_{\delta}^2, \sigma^2 &\sim \text{Student's t}(3, 0, 1.48). 
\end{aligned}
\end{equation}

The priors are visualized on the right side of Figure \ref{fig:priorvisualizationbrms2}. Note that these prior distributions are on the group-level parameters and represent the uncertainty about the general effect. The uncertainty about individual effects is represented by the marginal prior distribution on the left. This distribution takes the prior variability of the mean and variance as well as the implied individual variability into account.

\begin{figure}[H]

\includegraphics{manuscript_files/figure-latex/priorvisualizationbrms2-1} \hfill{}

\caption{Visualization of \textit{brms} default prior settings for the normal model. On the right, the prior distributions on general effects parameters are shown. On the left, the implied distribution of the individual effects for every parameter based on these priors are presented. }\label{fig:priorvisualizationbrms2}
\end{figure}

We will continue with step 3 of the prior specification process and check what the expected data would look like according to the selected priors. The results of the prior prediction for the symbolic distance effect with the priors from Equation \eqref{eq:chosenpriors1} with the number of participants \(n = 10\), the number of trials per condition \(k = 60\), and the number of simulations \(m = 1000\) are shown in panel A of Figure \ref{fig:priorpredfigurepanel}. Code for the simulations can be found in our Online Supplement.

The figure shows the difference in RTs between digits for every repetition averaged across subjects (i.e., 1000 grey dots per effect). The blue dots represent the mean effect. Using the default prior from \texttt{brms}, the mean effects are expected to be around zero, with a mean effect of zero. However, the range around this mean is very wide, ranging from -10 to 10 second effects. Most of the trials are in a smaller range of -5 to 5 second effects. This means that according to our priors, we would expect the effects to fall in this range. For this task, where participants have to respond as fast as possible, such large RT effects are unlikely. Therefore, it seems reasonable to adjust the priors.

\begin{figure}[H]

\includegraphics{manuscript_files/figure-latex/priorvisualization2-1} \hfill{}

\caption{Visualization of the adjusted prior settings for the normal model. On the right, the prior distributions on general effects parameters are shown. On the left, the implied distribution of the individual effects for every parameter based on these priors are presented. }\label{fig:priorvisualization2}
\end{figure}

The normal distribution on \(\mu_{\beta}\) and \(\mu_{\delta}\) seems reasonable if we assume that a two-sided hypothesis about the symbolic distance effect is tested. However, we expect less variance around this mean. Therefore, we decrease the setting for the variation. The wide tails of the Student's \emph{t}-distribution could be one of the reasons for the high range in variation. We change the prior for \(\mu_{\gamma}\) to a normal distribution that has more narrow tails. As RTs cannot be negative, we adjust the prior to be positive only by truncating the normal distribution using a lower bound of zero. Finally, we would like to use a more common conjugate prior for the variances to ensure obtaining a proper posterior distribution.\footnote{Conjugate priors result in posterior distributions that belong to the same distribution family as the prior distribution (Fink, 1997). For a normal likelihood, these are the normal distribution for the mean and the inverse gamma distribution for the variance.} Therefore, we choose an inverse gamma distribution. We can also be more specific about the expected RTs. Based on our knowledge of these types of tasks (e.g., Haaf \& Rouder, 2017), it seems highly plausible that participants will respond within a second or two. We would expect effects in the range of 10 milliseconds to 100 milliseconds. We adjust our priors using these new distribution types and settings, evaluate the prior predictions (shown in panel B of Figure \ref{fig:priorpredfigurepanel}) and arrive at the following priors:\footnote{Because response times cannot be negative the prior for the overall baseline \(\mu_\gamma\) is a truncated normal with a lower bound of zero.}

\begin{equation}
\begin{aligned}
\mu_{\gamma} &\sim \text{Truncated-Normal}_{+}(0.5, 1), \label{eq:chosenpriors2var} \\
\mu_{\beta} &\sim \text{Normal}(0, 0.09), \\
\mu_{\delta} &\sim \text{Normal}(0, 0.09), \\
\sigma^2 &\sim \text{Inverse-Gamma}(3, 0.7),  \\
\sigma^2_{\gamma} &\sim \text{Inverse-Gamma}(3, 0.7), \\   
\sigma^2_{\beta} &\sim \text{Inverse-Gamma}(3, 0.5), \\   
\sigma^2_{\delta} &\sim \text{Inverse-Gamma}(3, 0.5).    
\end{aligned}
\end{equation}
Setting equivalent prior distributions on the standard deviations instead of on the variances results in the following priors:

\begin{equation}
\begin{aligned}
\mu_{\gamma} &\sim \text{Truncated-Normal}_{+}(0.5, 1), \label{eq:chosenpriors2sd} \\
\mu_{\beta} &\sim \text{Normal}(0, 0.3), \\ 
\mu_{\delta} &\sim \text{Normal}(0, 0.3), \\  
\sigma &\sim \text{Inverse-Gamma}(13.8, 6.3),  \\ 
\sigma_{\gamma} &\sim \text{Inverse-Gamma}(13.8, 6.3), \\   
\sigma_{\beta} &\sim \text{Inverse-Gamma}(13.8, 5.3), \\    
\sigma_{\delta} &\sim \text{Inverse-Gamma}(13.8, 5.3).     
\end{aligned}
\end{equation}
The adjusted priors are visualized in Figure \ref{fig:priorvisualization2}. The figure shows that the adjusted priors assign their highest probability density to a smaller range of values and are, therefore, more specific than \texttt{brms}'s default priors presented in Figure \ref{fig:priorvisualizationbrms2}. In addition, the shape of the distribution is different for the individual variation, resulting in a higher probability for values closer to zero. Panel B of Figure \ref{fig:priorpredfigurepanel} shows the predictions of the normal model with adjusted prior distributions. Most effects now range between -0.5 and 0.5 seconds. This aligns better with our expectations based on existing literature, but still allows for a considerable uncertainty about the true size of the effects.

\hypertarget{log-normal-model-1}{%
\subsection{Log-normal model}\label{log-normal-model-1}}

We follow the same procedure for prior specification in the log-normal model. As before, we prefer to apply conjugate priors. The conjugate priors that we applied to the normal model are also conjugate priors for the log-normal model. As a starting point, we can therefore use the adjusted priors of the normal model and perform a prior prediction, shown in panel C of Figure \ref{fig:priorpredfigurepanel}. These priors result in a slightly higher variance in RTs than we would expect, and, importantly, the mean response time differences are not centered around zero in all conditions. This pattern is inconsistent with our expectations, which means that prior distributions for the log-normal model need to be adjusted.

The transformation of the prior distributions is not as simple as the transformation of the parameter itself. Panel C shows that with the same priors as for the normal model, we obtain different data for the log-normal model compared to the normal model (panel B). Therefore, we adjusted the priors until the simulated data matched our expectations and looked like panel B. This yields the following priors that result in the prior prediction shown in panel D of Figure \ref{fig:priorpredfigurepanel}:\footnote{For the overall baseline \(\mu_\gamma\) we adjusted the prior based on our expectation of the expected overall response times. A mean parameter of -0.5 for the log-normal likelihood corresponds to a mean response time of 1.}

\begin{equation}
\begin{aligned}
\mu_{\gamma} &\sim \text{Normal}(-0.5, 1), \label{eq:chosenpriorslogvar} \\
\mu_{\beta} &\sim \text{Normal}(0, 0.005), \\
\mu_{\delta} &\sim \text{Normal}(0, 0.005), \\
\sigma^2 &\sim \text{Inverse-Gamma}(3, 0.3),  \\
\sigma^2_{\gamma} &\sim \text{Inverse-Gamma}(3, 0.3), \\   
\sigma^2_{\beta} &\sim \text{Inverse-Gamma}(3, 0.01), \\   
\sigma^2_{\delta} &\sim \text{Inverse-Gamma}(3, 0.01).    
\end{aligned}
\end{equation}
Setting the same prior distribution on the standard deviations instead of on the variances yields the following priors:

\begin{equation}
\begin{aligned}
\mu_{\gamma} &\sim \text{Normal}(-0.5, 1), \label{eq:chosenpriorslogsd} \\
\mu_{\beta} &\sim \text{Normal}(0, 0.07), \\
\mu_{\delta} &\sim \text{Normal}(0, 0.07), \\
\sigma &\sim \text{Inverse-Gamma}(13.8, 4.1),  \\
\sigma_{\gamma} &\sim \text{Inverse-Gamma}(13.8, 4.1), \\   
\sigma_{\beta} &\sim \text{Inverse-Gamma}(13.8, 0.7), \\   
\sigma_{\delta} &\sim \text{Inverse-Gamma}(13.8, 0.7).    
\end{aligned}
\end{equation}

\hypertarget{prior-sensitivity}{%
\subsection{Prior sensitivity}\label{prior-sensitivity}}

In the previous sections, we decided on one set of prior distributions for each of the proposed models. However, researchers may find it difficult to settle on a single set of prior distributions in practice. This raises the question: What happens if you cannot decide on one prior? If several possibilities seem plausible, it is possible to use them all separately for the analysis and investigate their influence on the analysis results. This is called a prior sensitivity analysis (Roos, Martins, Held, \& Rue, 2015). Generally, when it comes to prior specification, it is important to be \emph{transparent} and to provide a justification for the selected prior distributions (Stefan, Katsimpokis, Gronau, \& Wagenmakers, 2021). Prior distributions are rarely inadmissible. However, as we demonstrated earlier, their usefulness in constraining model predictions can vary. Therefore, providing a justification of prior distributions in a paper can help readers gauge the generative quality of a presented model. When multiple justifiable prior distributions are applied to the same analysis, it is important to be transparent about this to avoid the impression of cherry-picking priors. If priors are chosen for their best fit to the data, this can lead to overfitting and spurious results. Prior sensitivity analyses and transparency about the prior specification process more generally, can therefore help to increase trust in the research results.

\hypertarget{bayesian-parameter-estimation}{%
\section{Bayesian Parameter Estimation}\label{bayesian-parameter-estimation}}

In Bayesian statistics, parameter estimates are obtained from the posterior distribution. The posterior distribution captures the uncertainty about the parameter after seeing the data, \(\bf{Y}\). The mean of this distribution can be used as a point estimate for the parameter. The posterior distribution is calculated by updating the prior distribution using Bayes rule. For the symbolic distance effect, we have a model with many parameters, so the posterior distribution is also multidimensional. Instead of writing out all the parameters, we replace all the model parameters with the vector \(\bm{\theta}\), resulting in the joint posterior:

\begin{equation}
\underbrace{P(\bm{\theta}|\bf{Y})}_{\text{Posterior}} = \underbrace{P(\bm{\theta})}_{\text{Prior}}\times\underbrace{\frac{P(\bf{Y}|\bm{\theta})}{P(\bf{Y})}}_{\text{Updating factor}}.
\label{eq:Bayesrule}
\end{equation}

Typically, we are not interested in the distribution of all parameters together, but in the posterior of one parameter independent of all the others. This posterior distribution is called the marginal posterior, and it is computed by integrating out all other parameters of the joint posterior. The marginal posterior distribution for the side effect \(\mu_{\beta}\), for instance, is:

\begin{equation}
P(\mu_{\beta}| \bf{Y}) = \int P(\bf{Y} | \mu_{\beta}, \bm{\theta})P(\mu_{\beta})P(\bm{\theta})) d\bm{\theta}  
\end{equation}
where \(\bm{\theta}\) now represents all parameters except for \(\mu_{\beta}\). With an increasing number of parameters and an increasingly complex model structure, it is no longer possible to find analytical solutions for the multidimensional integrals involved in the posterior distributions. A solution to this issue is to sample from the joint posterior distribution. Using algorithms such as Markov chain Monte Carlo sampling (MCMC; for an introduction see Van Ravenzwaaij, Cassey, \& Brown, 2018), Gibbs sampling (Chib, 1995) or Hamiltonian Monte Carlo sampling (Betancourt \& Girolami, 2015). It is out of the scope of this paper to go into detail on the algorithms, but we refer the interested reader to a tutorial paper on MCMC by van Ravenzwaaij et al. (2018), a conceptual introduction to Hamiltonian Monte Carlo by Betancourt (2018), and a summary of samplers by Green et al. (2015).

An approximation for the marginal posterior distribution can be obtained by assessing the empirical distribution of MCMC samples for the parameter of interest disregarding the other parameters. Since the frequency of combinations of samples in the joint posterior is determined by their posterior plausibility, this procedure is equivalent to a multidimensional integration across all other parameters.

Every sampling algorithm requires the specification of the number of samples (also called iterations) that are drawn from the posterior distribution. Furthermore, all MCMC sampling methods have in common that results depend (somewhat) on chosen starting values, that is, initial parameter values for which the unnormalized posterior (i.e., likelihood times prior) is evaluated. This dependency can be assessed if the sampling algorithm is repeated several times (i.e., chains) with different starting values. Since the first few iterations of a chain---often called warm-up or burn-in---are somewhat dependent on the starting value, these samples are typically discarded.

Before interpreting the estimation results, it has to be checked whether the posterior distribution of the parameters has converged, meaning that a stationary posterior distribution has been reached (Vehtari, Gelman, Simpson, Carpenter, \& Bürkner, 2021). We will explain these convergence checks in more detail in the section Model Diagnostics.

\hypertarget{software}{%
\subsection{Software}\label{software}}

There are many different programs available to fit Bayesian hierarchical models, that is, to obtain samples from the joint posterior distribution of parameters. In this tutorial, we focus on two of the most commonly used \texttt{R} packages, \texttt{rstan} (Stan Development Team, 2019a) and \texttt{brms} (Bürkner, 2017, 2018). We assume that readers are somewhat familiar with \texttt{R}. The package \texttt{rstan} provides the most modern algorithms and flexibility in model set-up. The package \texttt{brms} was created as an overlay to \texttt{rstan} and is based on the well-known \texttt{lme4} syntax (Bates, Mächler, Bolker, \& Walker, 2015). The goal of \texttt{brms} is to ease the transition to Bayesian hierarchical modeling for novice users of Bayesian statistics. The packages work similarly for the normal and log-normal model. In this section, we explain how a model can be fitted in each package using code snippets. In addition, we offer a separate Rmarkdown file for each package in our Online Supplement that explains in more depth how to analyze the full symbolic distance effect model with the specific package.

\hypertarget{rstan}{%
\subsubsection{Rstan}\label{rstan}}

\texttt{Rstan} allows the application of \texttt{Stan} (Carpenter et al., 2017), a probabilistic programming language, in \texttt{R}. The package uses a version of the Hamiltonian Monte Carlo (HMC) algorithm, No-U-turn sampler (NUTS; Hoffman \& Gelman, 2014), for sampling. \texttt{Rstan} requires two files: a \texttt{.stan} file containing the model specification and an R script containing the code for the model fitting. In the \texttt{.stan} file, the parameters, priors, and the likelihood are specified. An example \texttt{.stan} file of a simple, non-hierarchical normal model, where the variance is already known and the mean has to be estimated, is shown below (adapted example from Nicenboim, Schad, \& Vasishth, 2021, Chapter 10). The \texttt{.stan} file for the symbolic distance effect, including explanations, can be found in the Online Supplement. The file is divided into three sections: \texttt{data}, \texttt{parameters}, and \texttt{model}.

\scriptsize

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{data \{}
\NormalTok{  int}\SpecialCharTok{\textless{}}\NormalTok{lower }\OtherTok{=} \DecValTok{1}\SpecialCharTok{\textgreater{}}\NormalTok{ N;             }\SpecialCharTok{/}\ErrorTok{/}\NormalTok{ Total number of trials }
\NormalTok{  vector[N] y;                  }\SpecialCharTok{/}\ErrorTok{/}\NormalTok{ Score }\ControlFlowTok{in}\NormalTok{ each trial }
\NormalTok{  real}\SpecialCharTok{\textless{}}\NormalTok{lower }\OtherTok{=} \DecValTok{0}\SpecialCharTok{\textgreater{}}\NormalTok{ sigma;        }\SpecialCharTok{/}\ErrorTok{/}\NormalTok{ The variance of the normal distribution }
\NormalTok{\}}

\NormalTok{parameters \{}
\NormalTok{  real mu; }
\NormalTok{\}}

\NormalTok{model \{}
  \SpecialCharTok{/}\ErrorTok{/}\NormalTok{ Priors}\SpecialCharTok{:} 
\NormalTok{  target }\SpecialCharTok{+}\ErrorTok{=} \FunctionTok{normal\_lpdf}\NormalTok{(mu }\SpecialCharTok{|} \DecValTok{0}\NormalTok{, }\DecValTok{20}\NormalTok{);}
  \SpecialCharTok{/}\ErrorTok{/}\NormalTok{ Likelihood}\SpecialCharTok{:} 
  \ControlFlowTok{for}\NormalTok{(i }\ControlFlowTok{in} \DecValTok{1}\SpecialCharTok{:}\NormalTok{N)}
\NormalTok{    target }\SpecialCharTok{+}\ErrorTok{=} \FunctionTok{normal\_lpdf}\NormalTok{(y[i] }\SpecialCharTok{|}\NormalTok{ mu, sigma);}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

\normalsize

The \texttt{data} section specifies the variables in the dataset, as well as constant values in the model. For the symbolic distance effect model, the \texttt{data} section contains the total number of observations, the participant index variable, the response time variable in the dataset, the condition for every response time (side of the digit and the digit indicator, i.e., variables \(x_j\), \(u_j\), \(v_j\), \(w_j\) and \(z_j\) from \eqref{eq:datalevel}) and the prior specification according to Equations \eqref{eq:chosenpriors2var} and \eqref{eq:chosenpriorslogvar}. The \texttt{parameters} section specifies all parameters that are estimated. These are the general and individual effects. The last section, \texttt{model}, contains the likelihood function and priors, as specified earlier in this article.

In the corresponding \texttt{R} file, the model is fitted to the data using the \texttt{stan} function, as shown below. In this function, it is possible to specify the number of iterations, chains and warm-up samples in the Monte Carlo sampling procedure. Next to these specifications, the function requires a directory path to the \texttt{.stan} file and an \texttt{R} object of type list containing the data (here called \texttt{myData\_list}). Every element of the \texttt{data} section in the \texttt{.stan} file needs to be specified with the same name in the data object.
~

\scriptsize

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{model\_fit }\OtherTok{\textless{}{-}} \FunctionTok{stan}\NormalTok{(}\AttributeTok{file =} \StringTok{"./myModel.stan"}\NormalTok{,  }\CommentTok{\# Stan file with model  }
                  \AttributeTok{data =}\NormalTok{ myData\_list,       }\CommentTok{\# List with observed data and constants}
                  \AttributeTok{iter =} \DecValTok{4000}\NormalTok{,              }\CommentTok{\# Number (Nr.) of iterations per chain}
                  \AttributeTok{chains =} \DecValTok{4}\NormalTok{,               }\CommentTok{\# Nr. of chains }
                  \AttributeTok{warmup =} \DecValTok{1000}\NormalTok{)            }\CommentTok{\# Nr. iterations for warmup per chain}
\end{Highlighting}
\end{Shaded}

\normalsize

The package \texttt{rstan} offers a lot of flexibility. It is possible to specify the likelihood and priors in the exact way as is wished. However, this comes with the price of some technical knowledge. The commented code for fitting the full symbolic distance effect model using \texttt{rstan}, can be found in \href{https://www.dropbox.com/s/06f6vubmnhnm326/Appendix-C--Fit-BHM-in-rstan.pdf?dl=0}{Online Supplement C: \emph{Bayesian Hierarchical Modeling in Rstan}}.

\hypertarget{brms}{%
\subsubsection{Brms}\label{brms}}

The \texttt{R} package \texttt{brms} also allows fitting Bayesian models using \texttt{Stan}. However, model specification and model fitting can be achieved using a much simplified \texttt{R}-script. Model specification is based on a formula syntax that is similar to the notation in the popular \texttt{lme4} package for frequentist hierarchical modeling (Bates, Mächler, Bolker, \& Walker, 2015). The formula is an object that specifies the dependent variable as a function of the independent variables. The \texttt{\textasciitilde{}} (tilde) separates the dependent variable (\texttt{rt}) on the left side from the independent variables on the right side. The formula distinguishes between general effects and individual deviations. The first part of the equation on the predictor side of the formula shown below (i.e., \texttt{1\ +\ side\ +\ dif1\ +\ dif2\ +\ dif3\ +\ dif4}) represents the general effects. The \texttt{1} represents the intercept, in the case of the symbolic distance effect \(\mu_{\gamma}\), \texttt{side}, and \texttt{dif1} to \texttt{dif4} represent the general effects of side and distance (\(\mu_{\beta}\) and \(\mu_{\delta}\)). Within the round brackets individual deviations and grouping variables are specified. In this case the grouping variable (\texttt{ind}) indicates that each participant has an individual effect. In other cases effects might vary per item, block, or even the combination of participant and item. The double bar (\texttt{\textbar{}\textbar{}}) indicates that the correlation between the parameters should not be modeled. When using a single ba (\texttt{\textbar{}}) then correlations across all individual effects (intercept, side and distance) are modeled. This approach, however, is a bit more difficult to interpret and would also not correspond to the implemented \texttt{Stan} model.

\scriptsize

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{formula }\OtherTok{\textless{}{-}}\NormalTok{ rt }\SpecialCharTok{\textasciitilde{}} \DecValTok{1} \SpecialCharTok{+}\NormalTok{ side }\SpecialCharTok{+}\NormalTok{ dif1 }\SpecialCharTok{+}\NormalTok{ dif2 }\SpecialCharTok{+}\NormalTok{ dif3 }\SpecialCharTok{+}\NormalTok{ dif4 }\SpecialCharTok{+}   \CommentTok{\# General effects }
\NormalTok{          (}\DecValTok{1} \SpecialCharTok{+}\NormalTok{ side }\SpecialCharTok{+}\NormalTok{ dif1 }\SpecialCharTok{+}\NormalTok{ dif2 }\SpecialCharTok{+}\NormalTok{ dif3 }\SpecialCharTok{+}\NormalTok{ dif4 }\SpecialCharTok{||}\NormalTok{ ind)  }\CommentTok{\# Individual effects }

\NormalTok{model\_fit }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}\AttributeTok{formula =}\NormalTok{ formula,                      }\CommentTok{\# Model formula }
                 \AttributeTok{data =}\NormalTok{ myData\_dataframe,                }\CommentTok{\# Dataset (parameters, observations) }
                 \AttributeTok{family =} \FunctionTok{gaussian}\NormalTok{(),                    }\CommentTok{\# Family, distribution dependent variable }
                 \AttributeTok{prior =}\NormalTok{ priorsmodel,                    }\CommentTok{\# Specification of the priors }
                 \AttributeTok{warmup =} \DecValTok{1000}\NormalTok{,                          }\CommentTok{\# Nr. iterations for warmup (per chain)}
                 \AttributeTok{iter =} \DecValTok{4000}\NormalTok{,                            }\CommentTok{\# Nr. iterations (per chain)}
                 \AttributeTok{chains =} \DecValTok{4}\NormalTok{,                             }\CommentTok{\# Nr. chains }
                 \AttributeTok{core =} \DecValTok{4}\NormalTok{)                               }\CommentTok{\# Nr. of cores for parallel estimation}
\end{Highlighting}
\end{Shaded}

\normalsize The \texttt{brm} function requires a data frame (here called \texttt{myData\_dataframe}) that contains all variables mentioned in the formula. The argument \texttt{family} in the \texttt{brm} function specifies the exponential family of distributions according to which the dependent variable is assumed to be distributed. For the normal model, this argument takes the value \texttt{gaussian}; for the log-normal model, it needs to be specified as \texttt{lognormal}.

Finally, we specify the priors. In \texttt{brms}, priors can be set on the general effects represented by \texttt{class\ =\ b} and on the individual variation represented by \texttt{class\ =\ sd}. It is important to note that priors can only be set on the specific parameters that are part of the brms model specification. For example, \texttt{brms} only allows setting priors on standard deviations and not on variances.
Therefore, for the normal model of the symbolic distance effect, we will use the priors for standard deviations specified in Equation \eqref{eq:chosenpriors2sd}. To set the priors, we use the \texttt{set\_prior} function, as shown below. In this function, first, the distribution is specified, using the \texttt{Stan} programming language. Then, the type of parameter to which the prior applies is defined by \texttt{class}. The parameter to which this prior applies is specified by \texttt{coef}. For the priors on the individual effects, an additional element has to be specified, namely, the grouping variable (i.e., \texttt{group}). For example, the prior distributions on the general side effect, \(\mu_{\beta}\), and on its variability between individuals, \(\sigma_{\beta}\), can be specified as follows:
~

\scriptsize

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{set\_prior}\NormalTok{(}\StringTok{"normal(0,0.3)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"b"}\NormalTok{, }\AttributeTok{coef=}\StringTok{"side"}\NormalTok{)  }\CommentTok{\# general effect }
\FunctionTok{set\_prior}\NormalTok{(}\StringTok{"inv\_gamma(13.8, 5.3)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"sd"}\NormalTok{,       }\CommentTok{\# individual variation }
          \AttributeTok{coef=}\StringTok{"side"}\NormalTok{, }\AttributeTok{group =} \StringTok{"ind"}\NormalTok{) }
\end{Highlighting}
\end{Shaded}

\normalsize

\scriptsize

\normalsize If you do not specify a prior for a certain parameter, the default prior as explained in the prior section will be applied. It is possible to save all priors in an R \texttt{data.frame} and provide these in the \texttt{brm} function, this and the full code with explanation for fitting the full symbolic distance effect model using \texttt{brms} is shown in our \href{https://www.dropbox.com/s/mthphl3ce2kk0xx/Appendix-D--Fit-BHM-in-brms.pdf?dl=0}{Online Supplement D: \emph{Bayesian Hierarchical Modeling in brms}}.\footnote{In our example of the Number Classification Task we are applying a truncated normal prior distribution for only one parameter. This requires a slightly different \texttt{brms} formula and prior specification, explained in the Online Supplement.}

\hypertarget{estimation-of-the-number-classification-task}{%
\section{Estimation of the Number Classification Task}\label{estimation-of-the-number-classification-task}}

\hypertarget{data}{%
\subsection{Data}\label{data}}

For our example, we will use the data set by Rouder et al.~(2005). The data are available under a free license on GitHub \url{https://github.com/PerceptionCognitionLab/data0/tree/master/lexDec-dist5}. The data set contains observations from 54 participants who took part in a number classification task. Each participant performed 360 trials in six blocks. For the analysis, only the correct responses are included. To account for practice effects, the first 25 trials of the first block are removed from the analysis. The trials after the breaks between the blocks are also removed from the analysis to correct for lack of concentration. In addition, trials with very fast RTs (\textless{} 250 MS) and very slow RTs (\textgreater{} 2000 MS) are excluded. Two participants were excluded because they did not respond properly to the task. Similar criteria were applied by Haaf, Klaassen, and Rouder (2019). This results in 52 participants and a total of 17031 observations. In the following two sections we will describe the model fitting and parameter estimation procedure for the normal and log-normal model, respectively.

\hypertarget{normal-model-2}{%
\subsection{Normal Model}\label{normal-model-2}}

\begin{figure}[H]

\includegraphics[width=1\linewidth]{Images/Fig8_traceplots3pnew_v2} \hfill{}

\caption{Panel A and B show the trace plot of the digit parameter $\mu_{\delta_{6}}$ for both packages. Panel C shows the estimated $\hat{R}$ for all individual effects per parameter group. The stars represent the $\hat{R}$ of the general effects. Panel D illustrates the variation in the estimated number of effective samples per parameter group. The dashed lines represent the total number of iterations for each package. The stars represent the number of effective samples of the general effects.}\label{fig:traceplots3p}
\end{figure}

\hypertarget{model-fitting}{%
\subsubsection{Model Fitting}\label{model-fitting}}

The model is fitted using the packages \texttt{rstan} and \texttt{brms}. For \texttt{rstan}, four chains are run with 4000 iterations per chain, of which 1000 serve as warm-up. This results in a total of 12000 samples from the joint posterior distribution. For \texttt{brms}, four chains are run with 6000 iterations per chain, of which 1000 serve as warm-up. This results in a total of 20000 draws from the joint posterior distribution.

\hypertarget{model-diagnostics}{%
\subsubsection{Model Diagnostics}\label{model-diagnostics}}

Before we can interpret the results, we have to evaluate whether they are reliable. Specifically, it has to be checked whether the sampling chains of the parameters have converged, meaning that a stationary posterior distribution has been reached (Vehtari, Gelman, Simpson, Carpenter, \& Bürkner, 2021). If convergence is not achieved, the posterior estimates may change substantially if the sampling algorithm was run with more iterations or different starting values. The most common convergence measures are trace plots and the \(\hat{R}\) statistic. We will also discuss an additional check provided by \texttt{rstan} and \texttt{brms}, namely, the number of effective samples. For a critical evaluation of currently used convergence statistics, we recommend the interested reader to consult Vehtari et al.~(2019).

\hypertarget{trace-plot}{%
\paragraph{Trace plot}\label{trace-plot}}

Trace plots show the sampled parameter value at each iteration. If the distributions have converged, the plot should look like a hairy caterpillar (Lee \& Wagenmakers, 2013, p. 99): the iterations move up and down, but have much overlap in the middle. If multiple chains are used, iterations are shown per chain. In this case, the chains should overlap. If the posterior distribution did not converge, one chain could, for instance, be found on top of the figure, while all other chains are on the bottom (for an illustration see Vehtari, Gelman, Simpson, Carpenter, \& Bürkner, 2019). Or, perhaps the first half of the chain is on the top of the graph and the second half on the bottom. The trace plot does not include burn-in samples, since these will not be used for parameter estimation.

When inspecting the trace plots of the symbolic distance effect parameters in the normal model, the posterior distributions of the parameters appear to have converged as the iterations come together and look similar to a caterpillar. The trace plot of one of the parameters, \(\delta_{6, i}\), is shown in panel A for \texttt{rstan} and panel B for \texttt{brms} of Figure \ref{fig:traceplots3p}. The trace plot in panel B shows less overlap in the middle, indicating that \texttt{brms} had a bit more dependency within chains. This might be a sign of slight issues. However, not that much to be too concerned. Trace plots for other parameters can be found in the Online Supplement.

\hypertarget{hatbmr}{%
\paragraph{\texorpdfstring{\(\hat{\bm{R}}\)}{\textbackslash hat\{\textbackslash bm\{R\}\}}}\label{hatbmr}}

Next to a visual inspection of convergence, a numerical inspection is usually performed as well. A common approach is to check \(\hat{R}\), also called the Gelman-Rubin Diagnostic (Gelman \& Rubin, 1992). \(\hat{R}\) is the ratio of between-chain variance and within-chain variance (Sorensen \& Vasishth, 2015). If the chains diverge, the between-chain variance will be higher than the within-chain variance, resulting in a \(\hat{R}\) greater than 1. This indicates that the chains have not converged. There are different types of \(\hat{R}\), such as the split- and rank-based \(\hat{R}\) (Gelman et al., 2013; Vehtari, Gelman, Simpson, Carpenter, \& Bürkner, 2019). Even though researchers tend to trust convergence statistics more than visual inspection, these statistics are also not fool-proof. Therefore, the development of better convergence statistics is a current topic in Bayesian analysis methods.

The \(\hat{R}\) values for all individual and general symbolic distance effect parameters computed by \texttt{rstan} and \texttt{brms} are shown in panel C of Figure \ref{fig:traceplots3p}. The figure shows that \(\hat{R}\)'s are not substantially greater than 1. Therefore, it seems that the distributions have converged.

\hypertarget{number-of-effective-samples}{%
\paragraph{Number of Effective Samples}\label{number-of-effective-samples}}

Another popular convergence statistic is the number of effective samples. This statistic is concerned with the dependence of posterior samples. In MCMC algorithms, the samples are to some degree dependent on one another: the parameter values at iteration \emph{i} are similar to the parameter values at iteration \emph{i-1}. The number of effective samples represents the estimated total number of \emph{independent} draws from the posterior for every model parameter (Stan Development Team, 2018a, sec. 15.4). If the number of effective samples is low, this indicates a high dependency of iterations in the sampling procedure. Therefore, the number of effective samples can be seen as a measure of the amount of new information about the posterior distribution that is provided by the total number of samples drawn from the posterior. A common rule for interpretation is that the number should be close to or even greater than the number of total iterations (Carpenter, 2018; Vehtari, Gelman, Simpson, Carpenter, \& Bürkner, 2019).

For the symbolic distance effect, the estimated number of effective samples is provided in panel D of Figure \ref{fig:traceplots3p}. For the package \texttt{rstan} the number is generally high, much higher than the actual number of samples depicted by the green dashed line. The \texttt{brms} package shows much more variation. The lowest effective sample sizes can be found for the intercept parameters. Table 1 also displays the number of effective samples for the general effects parameters in the \texttt{rstan} model. If the number of effective samples is too low, \texttt{brms} and \texttt{rstan} will return a warning. Here, this was not case, and all other diagnostics were good, so we continue with the interpretation of the estimates.

\hypertarget{estimation-results}{%
\subsubsection{Estimation Results}\label{estimation-results}}

Considering good/acceptable convergence diagnostics, we can now interpret the estimation results. First, we present the general effects, then we visualize the individual variability of these effects.

\begin{figure}[H]

\includegraphics[width=1\linewidth]{Images/Fig9_posteriormu_v2} \hfill{}

\caption{The posterior distributions for the general effects estimated by the two packages. The middle line within the distributions represents the posterior mean. The shaded area within the distributions represent 95\% credible intervals.}\label{fig:modelposteriorsmufigure3}
\end{figure}

\hypertarget{general-effects}{%
\paragraph{General Effects}\label{general-effects}}

For the symbolic distance effect, there are two main questions we want to answer. First, we want to know whether there is a digit effect and, second, we want to know whether there is a side effect. Therefore, the key model parameters for general effects are \(\mu_{\beta}\), \(\mu_{\delta_{3}}\), \(\mu_{\delta_4}\), \(\mu_{\delta_6}\), and \(\mu_{\delta_{7}}\). In the following section, we will focus our reporting on these parameters.

The posterior distributions for the focal parameters in the normal model are presented in Figure \ref{fig:modelposteriorsmufigure3}. The posterior means estimated by \texttt{rstan} and \texttt{brms} can be found in the Online Supplement. The credible interval displayed in the figure is the Bayesian version of a confidence interval and provides a measure of uncertainty about the parameter value (Wagenmakers et al., 2018).

The parameter estimates produced by the two packages look very similar. The \texttt{brms} estimates seem to be slightly higher with slightly more narrow posterior distributions compared to \texttt{rstan}. All parameters are presented on a scale of seconds. As expected, the effects are small. For instance, \(\mu_{\delta_{7}}\) has a posterior mean of 0.012, corresponding to a 11 millisecond effect.

For the digit effects, positive parameter estimates indicate that the response is slower for the non-baseline digits, that is, digits that are closer to 5. The parameters \(\mu_{\delta_{6}}\) and \(\mu_{\delta_{4}}\) are distributed around higher values and their credible intervals do not contain zero. Note that this result is consistent with the symbolic distance hypothesis that the digits 4 and 6 have the highest response times because they are closest to the 5, the comparison value. The 95\% credible intervals of \(\mu_{\delta_{7}}\) and \(\mu_{\delta_{3}}\) contain zero. Therefore, it is plausible that the response times do not differ substantially from the baseline condition. The posterior distribution of the side effect \(\mu_{\beta}\) is centered around zero with a posterior mean of -0.007 and a 95\% credible interval ranging from -0.046 to 0.031. This indicates a small to non-existent side effect.

\hypertarget{individual-effects}{%
\paragraph{Individual effects}\label{individual-effects}}

\begin{table}[h]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:inddevtable}Posterior Variance, Lower and Upper Bound of the 95\% Credible Interval, Number of Effective Samples, and $\hat{R}$ of the Variance Parameters as estimated by Rstan.}

\begin{tabular}{lccccc}
\toprule
Parameters & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{Lower Bound} & \multicolumn{1}{c}{Upper Bound} & \multicolumn{1}{c}{$n_{eff}$} & \multicolumn{1}{c}{$\hat{R}$}\\
\midrule
$\sigma_{\gamma}^2$ & 0.034 & 0.024 & 0.049 & 17,556.041 & 1.000\\
$\sigma_{\beta}^2$ & 0.020 & 0.014 & 0.029 & 20,115.358 & 1.000\\
$\sigma_{\delta_{7}}^2$ & 0.020 & 0.014 & 0.029 & 20,340.216 & 1.000\\
$\sigma_{\delta_{6}}^2$ & 0.020 & 0.014 & 0.029 & 19,716.061 & 1.000\\
$\sigma_{\delta_{4}}^2$ & 0.020 & 0.014 & 0.029 & 18,989.012 & 1.000\\
$\sigma_{\delta_{3}}^2$ & 0.020 & 0.014 & 0.029 & 18,357.620 & 1.000\\
\bottomrule
\end{tabular}

\end{threeparttable}
\end{center}

\end{table}

Thanks to the hierarchical structure of the model, we can investigate whether the general effects hold for all the individuals by inspecting the individual variation and the individual parameter estimates. The individual variation is represented by the variance or standard deviation parameters in the model. The summaries of the posterior distributions of the variance parameters as estimated by \texttt{rstan} are shown in Table \ref{tab:inddevtable}. The estimates for the standard deviations by \texttt{brms} are smaller compared to \texttt{rstan} and can be found in Online Supplement D.

Posterior means of the variances for the digit and side effects are all very close to 0.020 which corresponds to a standard deviation of 0.141 seconds, that is, 141 milliseconds. Additionally, the 95\% credible intervals do not contain zero. If we consider the effect of digit 7 wit an 11 millisecond general effect this is a substantial amount of individual variation. Therefore, we conclude that there are individual differences in the symbolic distance effect.

\begin{figure}[H]

\includegraphics[width=1\linewidth]{Images/Fig10_randompar_v2} \hfill{}

\caption{The posterior means for individual effect parameters with 95\% credible intervals as estimated by \textit{rstan}, shown in increasing order. The dashed line represents the general posterior mean. Pink intervals contain zero, blue interval do not contain zero.}\label{fig:randomparfigure}
\end{figure}

Figure \ref{fig:randomparfigure} shows the posterior means and 95\% credible intervals for the individual effects of all 52 participants. Estimates are printed in increasing order, from lowest to highest. Pink intervals indicate that the 95\% credible interval contains zero, while blue intervals are either entirely above or below zero. A corresponding figure with individual effect estimates for \texttt{brms} can be found in Online Supplement D.

The largest variation across participants can be seen for the intercept parameter. This means that participants vary considerably in their baseline response times. There is also considerable individual variation in the digit effects (\(\delta\)'s). However, most of the intervals contain zero which implies that the effects might be small or non-existent. The variability is most pronounced for the effect parameters of the digits close to 5. For these digits, there are also more intervals that do not contain zero, indicating that an effect might be present on the subject level. The posterior mean of the individual side effect (\(\beta_{i}\)) varies between -0.05 and 0.05. Again, all but four 95\% credible intervals include zero indicating uncertainty about the existence of an effect for anyone. The credible intervals of the side effect and digit effects \(\delta_{7}\) and \(\delta_{3}\) indicate that different subjects may show opposite effects (Haaf \& Rouder, 2019).

Finally, Figure \ref{fig:modelestimatesfigure3} shows how the relationship between individuals' digit effects.\footnote{The design and code for this figure are by Langen (2020).} The posterior means of individual effects are represented by the dots. The right side of each plot represents the distribution of the individual estimates. Note that this distribution is not the posterior distribution of the general effect. It rather corresponds to the left side of Figure 6. The two panels of Figure \ref{fig:modelestimatesfigure3} show that the estimates are almost identical between \texttt{Stan} and \texttt{brms}. This means that differences in the parametrization between the packages did not influence the results of the parameter estimation.

\begin{figure}[H]

\includegraphics[width=0.75\linewidth]{Images/Fig11_modelestimatefigure2_onlyrstan_v2} \hfill{}

\caption{Model estimates for digit effect parameters in \textit{rstan}. The points represent the mean parameter estimates for each individual. The violin plots on the right side show the variance of the individual parameter estimates. The size of effects is not very constistent within a person.}\label{fig:modelestimatesfigure3}
\end{figure}

\hypertarget{log-normal-model-2}{%
\subsection{Log-normal Model}\label{log-normal-model-2}}

\hypertarget{model-estimation}{%
\subsubsection{Model Estimation}\label{model-estimation}}

The model is fitted using the packages \texttt{rstan} and \texttt{brms}. For \texttt{rstan}, we use the priors as specified in Equation \eqref{eq:chosenpriorslogvar}. For \texttt{brms}, we use the priors in Equation \eqref{eq:chosenpriorslogsd}. The rest of the settings are equivalent for both packages. We ran four chains with 4000 iterations per chain, including 1000 warm-up iterations. This results in a total of 12000 samples from the posterior distribution.

\hypertarget{model-diagnostics-1}{%
\subsubsection{Model Diagnostics}\label{model-diagnostics-1}}

Before interpreting the results, we have to check if the posterior distributions of the parameters have converged. The trace plot of \(\mu_{\delta_{6}}\) is shown panel A for \texttt{rstan} and B for \texttt{brms} of Figure \ref{fig:traceplotlog}. Trace plots for all other parameters can be found in the Online Supplement. All plots look like a hairy caterpillar. Therefore, they indicate that the posterior distributions of the parameters have converged. \(\hat{R}\) values shown in Figure \ref{fig:traceplotlog}C for all parameters are all close to 1 and most of the number of effective samples shown in Figure \ref{fig:traceplotlog}D exceed the total number of iterations. The number of effective samples for the general effects are comparatively low but still sufficient. In combination with the other diagnostics, we conclude that the posterior distributions of the parameters have converged and we proceed with the interpretation of the results.

\begin{figure}[H]

\includegraphics[width=1\linewidth]{Images/Fig12_traceplots3plog_v2} \hfill{}

\caption{Panel A and B show the trace plot of the digit parameter \(\\mu_{\\delta_{6}}\) in the log-normal model for both packages. Panel C shows the estimated \(\\hat{R}\) for all individual effects per parameter group. The stars represent the \(\\hat{R}\) of the general effects. Panel D illustrates the variation in the estimated number of effective samples per parameter group. The dashed lines represent the total number of iterations for each package. The stars represent the number of effective samples of the general effects.}\label{fig:traceplotlog}
\end{figure}

\hypertarget{results}{%
\subsubsection{Results}\label{results}}

\hypertarget{general-effects-1}{%
\paragraph{General effects}\label{general-effects-1}}

We inspect the posterior distributions for the general effect parameters presented in Figure \ref{fig:postplotlog}.

\begin{figure}[H]

\includegraphics[width=1\linewidth]{Images/Fig13_posteriormulog_v2} \hfill{}

\caption{The posterior distributions for the general effects in the log-normal model estimated by the two packages. The middle line within the distributions represents the posterior mean. The shaded area within the distributions represent 95\% credible intervals.}\label{fig:postplotlog}
\end{figure}

We note that the values of the parameter estimates for the log-normal model cannot be interpreted in the same way as in normal model. The reason is that the mean parameter of the log-normal distribution does not correspond to its expected value. To better understand the size of an effect that corresponds to the parameter values, some calculations are required. For instance, if we would like to know the median difference in response times for digit 3 and 2, we calculate the difference between the exponential of the RT estimate for the condition digit = 3 and the exponential of the RT estimate for the condition digit = 2:

\begin{equation}
\begin{aligned}
&e^{(\mu_{\gamma} + x_j \mu_{\beta} + w_j \mu_{\delta_{3}})} - e^{(\mu_{\gamma} +  x_j \mu_{\beta})} \\
&= e^{(-.56 + .5 \times (-.01) + 1 \times 0.03)} - e^{(-.56 + .5 \times (-.01))} \\ 
&= 0.017 
\end{aligned}
\end{equation}
This value corresponds to the median estimate of \(\mu_{\delta_{3}}\) from the normal model. Calculating the mean estimate of \(\mu_{\delta_{3}}\) would be similar, but more complicated as the mean of a log-normal distribution also depends on its variance.

One property of these parameter estimates we can immediately interpret is their sign (i.e., positive, negative or zero). A positive estimate indicates that the RT increases, whereas a negative effect indicates a decrease. When inspecting the general digit effect parameters, they all appear to positively influence the RT. This is in line with our expectations since the baseline RT in the model is set equal to the RT for the digits furthest away from 5 and the symbolic distance effect postulates that RT should increase as the digits become closer to five. The side effect is close to zero, with a posterior mean of -0.008 and a 95\% credible interval from -0.022 to 0.007. Thus, the general effect of side seems small or negligible.

\hypertarget{individual-effects-1}{%
\paragraph{Individual effects}\label{individual-effects-1}}

Next, we inspect the individual deviations from the general effects. Table \ref{tab:inddevtablelog} presents the estimates of the variance parameters by \texttt{rstan}. Results for brms can be found in the Online Supplement.

\begin{table}[h]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:inddevtablelog}Posterior Variance, Lower and Upper Bound of the 95\% Credible Interval, the Number of Effective Samples, and the $\hat{R}$ of the Variance Parameters as estimated by Rstan.}

\begin{tabular}{lccccc}
\toprule
Parameters & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{Lower Bound} & \multicolumn{1}{c}{Upper Bound} & \multicolumn{1}{c}{$n_{eff}$} & \multicolumn{1}{c}{$\hat{R}$}\\
\midrule
$\sigma_{\gamma}^2$ & 0.030 & 0.021 & 0.043 & 15,617.383 & 1.000\\
$\sigma_{\beta}^2$ & 0.001 & 0.001 & 0.002 & 7,815.681 & 1.000\\
$\sigma_{\delta_{7}}^2$ & 0.001 & 0.001 & 0.002 & 6,805.297 & 1.000\\
$\sigma_{\delta_{6}}^2$ & 0.001 & 0.001 & 0.002 & 7,325.299 & 1.000\\
$\sigma_{\delta_{4}}^2$ & 0.001 & 0.001 & 0.002 & 6,237.497 & 1.000\\
$\sigma_{\delta_{3}}^2$ & 0.001 & 0.001 & 0.002 & 6,841.198 & 1.001\\
\bottomrule
\end{tabular}

\end{threeparttable}
\end{center}

\end{table}

The estimated parameter values for the variance terms are very small, but different from zero. This indicates that individuals may vary slightly with respect to the general effect parameters. The estimates for the standard deviations by \texttt{brms} are similar compared to the estimates by \texttt{rstan}, but slightly smaller. The exact estimates by \texttt{brms} can be found in Online Supplement F.
Figure \ref{fig:randomparplotlog} shows the individual estimates for the digit and side effects. The pink lines indicate that the 95\% credible interval of the individual effect contains zero. The estimated general effect is represented by the dashed line. For \(\beta\), \(\delta_{7}\), and \(\delta_{3}\), most 95\% credible intervals contain zero. For \(\delta_{6}\) and \(\delta_{4}\), most credible intervals are entirely larger than zero, indicating an increase in RT for most participants.

\begin{figure}[H]

\includegraphics[width=1\linewidth]{Images/Fig14_randomparlog_v2} \hfill{}

\caption{The posterior means for individual effect parameters in the log-normal model with 95\% credible intervals as estimated by \textit{rstan}, shown in increasing order. The dashed line represents the general posterior mean. Pink intervals contain zero, blue interval do not contain zero.}\label{fig:randomparplotlog}
\end{figure}

We can further inspect the individual estimates for the digit effects in Figure \ref{fig:modeldigitestplotlog}. The figure shows the individual estimates for the digit effects with the distribution of the variability of the individual estimates (not the posterior distribution of the general effect). The figure illustrates that there is considerable variance in the point estimates of the individual digit effects. However, compared to the normal model there is considerably more hierarchical shrinkage.

\hypertarget{model-comparison}{%
\section{Model Comparison}\label{model-comparison}}

The symbolic distance effect postulates that response times (RTs) are slower when digits are closer to 5 (digit effects), and that RTs may be influenced by whether the target digit is smaller or greater than 5 (side effect). Using Bayesian estimation, we are not able to directly test these hypotheses. Credible intervals often contain zero, but they always also contain lots of other values that have a chance of being the ``true'' parameter value, making them unsuitable to test hypotheses. Therefore, our goal for the following section is to set up a Bayesian hypothesis test for the parameters of interest in our model. Specifically, we will achieve this by comparing the predictive accuracy of a null model where we set one or some parameters to zero, and and effects models where these parameters are unconstrained.

\begin{figure}[H]

\includegraphics[width=0.75\linewidth]{Images/Fig15_inddiflogplot_onlyrstan_v2} \hfill{}

\caption{Model estimates for digit effect parameters in the log-normal model in \textit{rstan}. The points represent the mean parameter estimates for each individual. The violin plots on the right side show the variance of the individual parameter estimates.}\label{fig:modeldigitestplotlog}
\end{figure}

\hypertarget{bayes-factor}{%
\subsection{Bayes factor}\label{bayes-factor}}

The Bayes factor (BF) is a measure of how well one model predicts the data compared to another (Jeffreys, 1961; Kass \& Raftery, 1995; Rouder \& Morey, 2018). The Bayes factor is defined as the probability of the data under one model compared to the probability of the data under another model. For instance, the Bayes factor comparing a null model (\(\mathcal{M}_0\)) and a full model (\(\mathcal{M}_{\text{F}}\)) can be obtained by:

\begin{equation}
\text{BF}_{0\text{F}} = \frac{P(\bm{Y} | \mathcal{M}_0)}{P(\bm{Y} | \mathcal{M}_{\text{F}})}. 
\end{equation}

The Bayes factor can also be interpreted as the relative evidence of two competing models. Bayes factors have a few very convenient characteristics. One of them is that Bayes factors are transitive, that is, evidence for the full model over the null model can be obtained by flipping numerator and denominatior in equation 13. For a full overview of the advantages of Bayesian inference with Bayes factor, see Wagenmakers et al. (2018).

Our models consist of several parameters. Therefore, computing the Bayes factor of, for example, the normal model versus the log-normal model becomes somewhat difficult. Let the vector of all parameters of the normal model be \(\bm{\theta_\text{n}}\) and the vector of all parameters for the log-normal model be \(\bm{\theta_{\text{l}}}\). With the prior distributions for the two models, \(f(\bm{\theta_\text{n}})\) and \(f(\bm{\theta_\text{l}})\), the Bayes factor can be calculated by the following ratio of marginal likelihoods:

\begin{equation}
\text{BF}_{\text{nl}} = \frac{\int_{-\infty}^{\infty}P(\bm{Y} | \mathcal{M}_\text{n}, \bm{\theta_{\text{n}}})f(\bm{\theta_{\text{n}}})d\bm{\theta_{\text{n}}}}{\int_{-\infty}^{\infty}P(\bm{Y} | \mathcal{M}_\text{l}, \bm{\theta_{\text{l}}})f(\bm{\theta_{\text{l}}})d\bm{\theta_{\text{l}}}}. 
\end{equation}

Imagine this results in \(\text{BF}_{\text{nl}} = 100\). This means that the data are 100 times more likely under the normal model than under the log-normal model, clear evidence in favor of the normal model. But how do we know whether a Bayes factor of 100-to-1, or 10-to-1 or 3-to-1. As a reminder, Bayes factors are the relative predictive accuracy of two models for the observed data. Therefore, Bayes factors are odds or ratios. Odds themselves are directly interpretable without the need for decisions or cutoffs. For example, if a presidential candidate is favored 10-to-1 over another one, then these are just the odds, and it is beside the point whether these are large or not. According to this line of thought, Rouder et al. (2018) argue that researchers should take a similar approach when interpreting and reporting Bayes factors. While there exist rules of thumb to categorize Bayes factors verbally, we suggest researchers use their research context to form substantive conclusions based on Bayes factors, not arbitrary rules of thumb.

Same as the posterior distribution, Bayes factors can rarely be computed analytically due to the multidimensional integrals involved. The difficulty to compute Bayes factors is one of the major drawbacks of Bayesian inference. However, several algorithms are available to estimate Bayes factor from posterior samples. In the following section we will present two methods to obtain Bayes factors for the symbolic distance effect.

\hypertarget{savage-dickey-density-ratio}{%
\subsubsection{Savage-Dickey Density Ratio}\label{savage-dickey-density-ratio}}

One method to compute Bayes factors is the Savage-Dickey density ratio (SDD-ratio; Dickey \& Lientz, 1970; Morey, Rouder, Pratte, \& Speckman, 2011; Wagenmakers, Lodewyckx, Kuriyal, \& Grasman, 2010). This approach can be readily used to estimate the Bayes factor between two nested models that differ in only one parameter. A prime example for such a case is comparing an effect model with a null model. For example, we may compare a model with an overall side effect (effect model) to a model with no overall side effect (null model). In this case, the Bayes factor in favor of the side effect can be approximated by the ratio of the posterior density of the side effect and the prior density of the side effect at the point zero. More generally, the Savage Dickey approach works whenever one model fixes a parameter to a constant and another model assigns a prior distribution to the same parameter.

Figure \ref{fig:sdratioplotall} shows a visualization of the SDD-ratio for the side parameter in the normal model of the symbolic distance effect. The posterior distribution is depicted in orange, the prior in green. The Bayes factor can be computed as a ratio of the posterior and prior density at the parameter value of zero, as depicted by the dashed line connecting the two points on the density functions.

\begin{figure}[H]

\includegraphics[width=0.6\linewidth]{Images/Fig16_sdrplotside_v2} \hfill{}

\caption{The prior and posterior distribution of the side parameter estimated by the \textit{brms} package. The dots represent the height of the distributions at zero. The ratio of these dots indicates the Savage-Dickey density ratio. The Bayes factor represents the evidence for the hypothesis that the effect equals zero. A Bayes factor above one indicates that the evidence is in favor of the null hypothesis.}\label{fig:sdratioplotall}
\end{figure}

With \texttt{rstan}, the SDD-ratio can be computed by fitting a density function to the posterior distribution obtained from the MCMC samples, and comparing the value of this density function at \(\mu_{\beta} = 0\) to the value of the prior distribution at \(\mu_{\beta} = 0\). This is illustrated in the code block below. First, we extract the posterior samples for the parameter \(\mu_{\beta}\), the general effect of side. We apply the \texttt{logspline} function (Kooperberg, 2019) to the samples to obtain an estimate of the log-density of the posterior distribution. The density at the zero is calculated with the \texttt{dlogspline} function. Finally, we compare the estimated density of the posterior to the density of the prior.

\scriptsize

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(logspline)}
\CommentTok{\# Get posterior samples for the side parameter mu2}
\NormalTok{samples\_side }\OtherTok{\textless{}{-}}\NormalTok{ rstan}\SpecialCharTok{::}\FunctionTok{extract}\NormalTok{(model\_fit)}\SpecialCharTok{$}\NormalTok{mu2   }
\NormalTok{log.posterior }\OtherTok{\textless{}{-}} \FunctionTok{logspline}\NormalTok{(samples\_side)                 }\CommentTok{\# estimate of log{-}density post. distr. }
\NormalTok{posterior\_at\_zero }\OtherTok{\textless{}{-}} \FunctionTok{dlogspline}\NormalTok{(}\DecValTok{0}\NormalTok{, log.posterior)        }\CommentTok{\# compute posterior density }
\NormalTok{prior\_at\_zero }\OtherTok{\textless{}{-}} \FunctionTok{dnorm}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\FloatTok{0.3}\NormalTok{)                        }\CommentTok{\# computer prior density }
\NormalTok{BF01 }\OtherTok{\textless{}{-}}\NormalTok{ posterior\_at\_zero}\SpecialCharTok{/}\NormalTok{prior\_at\_zero                  }\CommentTok{\# compare estimated density posterior to prior at point 0 }
\end{Highlighting}
\end{Shaded}

\normalsize

In \texttt{brms}, the SDD-ratio can be computed using the built-in \texttt{hypothesis} function, as shown in the code snippet below. The resulting Bayes factor, also depicted in Figure \ref{fig:sdratioplotall}, is \(BF_{01} = 18.28\), indicating evidence for the null model and against a side effect.
~

\scriptsize

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# We will use the model fit from before, specified as model\_fit}
\NormalTok{hypothesis\_fit }\OtherTok{\textless{}{-}} \FunctionTok{hypothesis}\NormalTok{(model\_fit,    }\CommentTok{\# Object containing model fit }
                             \StringTok{"mu2 = 0"}\NormalTok{)   }\CommentTok{\# The parameter we want to test with the null hypothesis }
\end{Highlighting}
\end{Shaded}

\normalsize

The Savage-Dickey approach has a number of drawbacks (see, for example, Heck, 2019). First, one issue is the quality of approximation for the SSD-ratio. If either the prior or, even more problematically, the posterior have a really low density at the test value (in this case zero), it becomes difficult to estimate the SSD-ratio from the samples. This issue often occurs when the effect is large, moving the posterior distribution away from zero. In many cases there might be no samples around zero, which means that the density estimate will be very inaccurate. Therefore, we would recommend only to use the SSD-ratio if you are confident that enough posterior samples are available from the tails of the distribution. One way of ensuring that is to drastically increase the number of drawn samples.

Second, the Savage-Dickey approach only applies for a limited set of to-be-compared models. In the example used here, we compared a model where the overall side effect is allowed to vary with a model where it is zero. Both models, however, still allow for individual variability of the side effect. For the null model, individual variability implies that some individuals are expected to have an effect favoring larger numbers and other individuals are expected to have an effect favoring smaller numbers, but, on average, both groups perfectly balance out at zero. To us, this model seems nonsensical, and, more importantly, it is not what researchers have in mind if they want to test a model without side effect. We would prefer to use a null model where none of the participants have a side effect. Yet, the Savage-Dickey density ratio is not available for such a model compared to the effects model.

\hypertarget{bridge-sampling}{%
\subsubsection{Bridge Sampling}\label{bridge-sampling}}

Bridge sampling offers a more flexible way to obtain Bayes factors (Kass \& Raftery, 1995). It allows the computation of Bayes factors for comparing models that differ in more than one parameter. Specifically, bridge sampling is a sampling-based algorithmic method to obtain the marginal likelihood of the data under a given model from the posterior samples (Gronau et al., 2017; Meng \& Wong, 1996).

Using bridge sampling we are able to estimate Bayes factors between any models of interest for the classification task data. We identify four models of interest:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Full model: This is the model we have been using throughout this paper. It states that there is an effect of side and an effect of digits for all participants.
\item
  Side model: This model specifies that participants show a side effect but no digit effect. All digit parameters \(\delta_{3}\) to \(\delta_{7}\) are set to zero.
\item
  Digit model: This model states that there are no side effects but that participants exhibit digits effects. All side parameters \(\beta\) are set to zero.
\item
  Null model: In this model, there is no side or digit effect on the RT. The digit parameters \(\delta_{3}\) to \(\delta_{7}\) and side parameters \(\beta\) are set to zero.
\end{enumerate}

Note that models models 2-4 are nested in the most complex full model, and the null model is nested in models 3 and 4, but models 3 and 4 are not nested. Additionally, the models differ by several parameters. For example, compared to the full model the Digit model restricts the side parameter for all participants to zero. Using bridge sampling, we can estimate the marginal likelihood for each model separately, and then compute Bayes factors as their ratios (Equation 14).

To obtain the marginal likelihood for a specific model in \texttt{rstan} with bridge sampling (Gronau et al., 2017), the \texttt{bridgesampling} package can be used (Gronau, Singmann, \& Wagenmakers, 2020). To obtain the Bayes factor comparing the side model and the full model, we first have to fit both models. Then, we use the \texttt{bridge\_sampler} function from the \texttt{bridgesampling} package to obtain the marginal likelihood of the data under each of the models. Finally, the models are compared using the \texttt{bf} function from the \texttt{bridgesampling} package.
~

\scriptsize

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# We will use the model fit of both models (called sidemodel and fullmodel)}
\CommentTok{\# (how to perform this is illustrated in the model estimation section)}

\CommentTok{\# Bridge sampling }
\FunctionTok{library}\NormalTok{(bridgesampling)}
\NormalTok{ML\_Hside }\OtherTok{\textless{}{-}} \FunctionTok{bridge\_sampler}\NormalTok{(sidemodel)  }\CommentTok{\# compute the marginal likelihood under the side model }
\NormalTok{ML\_Hfull }\OtherTok{\textless{}{-}} \FunctionTok{bridge\_sampler}\NormalTok{(fullmodel)  }\CommentTok{\# compute the marginal likelihood under the full model }

\CommentTok{\# Obtain Bayes factor (in this way evidence in favor of the side hypothesis)}
\NormalTok{bridgesampling}\SpecialCharTok{::}\FunctionTok{bf}\NormalTok{(ML\_Hside, ML\_Hfull)}
\end{Highlighting}
\end{Shaded}

\normalsize

In the \texttt{brms} package, Bayes factors based on bridge sampling can be obtained with the function \texttt{bayes\_factor} as shown in the code snippet below. The function takes two fitted models as input. Therefore, both models first have to be fitted as explained in Model Estimation section of this manuscript.

\scriptsize

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{bayes\_factor}\NormalTok{(fitsidemodel, fitfullmodel)}\SpecialCharTok{$}\NormalTok{bf}
\end{Highlighting}
\end{Shaded}

\normalsize

Even though bridge sampling is a flexible method for obtaining Bayes factors, it is still an estimation method based on posterior samples. Therefore, as with any Bayes factor estimation method, we advice to take two precautions with bridge sampling. First, it is always wise to increase the number of iterations when performing model comparison. The reason is that convergence of estimation of posterior distributions and convergence of estimation of the marginal likelihood can differ. Therefore, a rule of thumb is to use ten times the number of iterations than we would usually use for estimation. The second precaution is to conduct a stability analysis. Both the posterior samples from the model estimation procedure and the estimate of the marginal likelihood from the bridge sampling can be variable. Therefore, stability may best be assessed by repeating the entire process of fitting each model and applying the bridge sampling function several times.

We assess the stability of the obtained Bayes factor by repeating the estimating process ten times with \texttt{rstan} and \texttt{brms}, respectively. The results are shown in Figure \ref{fig:bsbfestimatesbrmsrstan}. The figure shows the evidence in favor of the null, digit and side model against the full model. All Bayes factors are larger than one. Therefore, it can be concluded that the data are more likely under less complex models. There is almost no variation in the computed Bayes factors per model, indicating that the Bayes factor estimate is stable. However, the Bayes factor estimates in \texttt{rstan} generally show stronger evidence than Bayes factor estimates based on \texttt{brms}. It is important to note that this difference is not due to the bridge sampling function, but a result of a difference in parametrization. However, for both packages, the conclusion is the same: the data are least likely to have occurred under the full model. In case the estimates are unstable, it is recommended to further increase the number of posterior samples.

\begin{figure}[H]

\includegraphics[width=0.7\linewidth]{Images/Fig17_normalbfplot01} \hfill{}

\caption{The BF estimates of the other models (0; null, side and digit model) against the full model (1). The green triangles represent the estimates using \textit{rstan}. The orange dots represent the \textit{brms} estimates. The dashed line represents a $\text{BF}_{01}$ of 1. }\label{fig:bsbfestimatesbrmsrstan}
\end{figure}

Finally, it is time to answer our research question: Is there a symbolic distance effect? When considering all Bayes factor estimates from Figure \ref{fig:sdratioplotall} and \ref{fig:bsbfestimatesbrmsrstan}, the null model (\(\text{BF}_{max} = 232 * 10^{105}\) for \texttt{rstan}) performs best compared to the side model (\(\text{BF}_{max} = 458 * 10^{68}\) for \texttt{rstan}), the digit model (\(\text{BF}_{max} = 269 * 10^{31}\) for \texttt{rstan}). Therefore, the data from Rouder et al.~(2005) show evidence against the symbolic distance effect.

\hypertarget{sensitivity-analysis}{%
\subsection{Sensitivity Analysis}\label{sensitivity-analysis}}

The full model including the side and digit effects performs worst compared to all other models in the Bayes factor model comparison. However, we did find non-zero digit effects in the model estimation. How is this possible?

A possible explanation is a lack of individual variability in the digit and side effects. The full model assumes that participants vary in the digit and side effect, and we specified our expectations about this size of the variability in the prior. However, if there is a lack of individual variability in the effects, the full model is punished for the complexity added by the individual variation (illustrated by Rouder et al., 2021). To investigate this idea, we can reformulate the prior distributions on the variance parameters to indicate a smaller individual variation or, even more extreme, remove these individual effects from the model entirely. Then, we can calculate the posterior model probabilities for all specified models.

Specifically, we will compare models from three scenarios: scenario 1 specifies large individual differences equivalent to Equation \eqref{eq:chosenpriors2var}. For the models in scenario 2, we adjust the priors on \(\sigma^2_{\beta}\) and \(\sigma^2_{\delta}\) in the following way:

\begin{equation}
\begin{aligned}
\sigma^2_{\beta} &\sim \text{Inverse-Gamma}(3, 0.05), \\   
\sigma^2_{\delta} &\sim \text{Inverse-Gamma}(3, 0.05). 
\end{aligned}
\end{equation}
The smaller value for the scale parameter of the inverse gamma distribution (i.e., 0.05 instead of 0.5) results in a shift of the peak of the prior towards smaller values for the individual variance. This means that the variation of the individual effects is expected to be much smaller than before.

In scenario 3, we remove individual variability from the side, digit, and full model. This results in the following full model, from which the side and digit model follow logically:

\begin{equation}
Y_{ijk} \sim \text{Normal}(\gamma_{i} + x_j \mu_\beta + u_j \mu_{\delta{7}} + v_j \mu_{\delta{6}} + w_j \mu_{\delta{4}} + z_j \mu_{\delta{3}}, \sigma^2). \label{eq:datalevel2}
\end{equation}
Note that the null model is the same in all three scenarios.

To obtain the posterior model probability for each model in every scenario, we first compute the Bayes factor of every model against the null model. This includes the trivial Bayes factor of the null model against the null model, resulting in ten Bayes factors in total. Next, we calculate the sum of all these Bayes factors and divide each Bayes factor by this sum. This results in the posterior model probabilities shown in Table 5.\footnote{Note that it would be possible to calculate the posterior model probabilities directly from the marginal likelihoods. However, marginal likelihoods are often not computed explicitly in the model fitting process (e.g., when using the bayes\_factor function in the \texttt{brms} package), so we believe that readers might find it easier to compute posterior model probabilities from Bayes factors.} All posterior model probabilities from the table and the posterior probability of the null model (see table note) sum to one.

\begin{table}[H]
\caption{Posterior Model Probabilities}
\begin{tabularx}{\textwidth}{XXXX}
\toprule
 & \multicolumn{1}{c}{Scenario 1} & \multicolumn{1}{c}{Scenario 2} & \multicolumn{1}{c}{Scenario 3}\\
\midrule
Digit model & $0.87*10^{-126}$ & $0.29*10^{-46}$ & $0.95$\\
Side model & $0.20*10^{-88}$ & $0.18*10^{-66}$ & $0.77*10^{-53}$ \\
Full model & $0.43*10^{-158}$ & $0.47*10^{-61}$ & $0.05$\\
\bottomrule
\end{tabularx}
\begin{tablenotes}[para]
\textit{Note.} Scenario 1: large individual differences. Scenario 2: small individual differences. Scenario 3: no individual differences. The posterior model probability of the null model is the same in every scenario and equals $0.50*10^{107}$. 
\end{tablenotes}
\end{table}

The table shows that the digit model without individual digit effects from scenario 3 has the highest posterior model probability, followed by the full model without individual digit and side effects (scenario 3). This means that the data are most likely under these models than the other eight models included in the comparison. In addition, the models with smaller prior expectations for individual differences (scenario 2), have higher posterior model probabilities than the models in scenario 1. Thus, decreasing our expectations of the size of the individual digit and side effects, or completely removing them, results in higher posterior model probabilities. Together with Figure \ref{fig:randomparfigure}, showing the small differences in individual effects (also illustrated by Haaf et al., 2019), this result supports the notion that side and digit effects are small and there is a lack in individual variability.

This example illustrates the benefits of comparing hierarchical models of varying complexity. While individual differences are a common phenomenon in psychology, there are instances where general effects are sufficient to describe the data adequately, or where only a random intercept, but no random slopes are necessary. As shown in Table 5, Bayesian model comparisons can help researchers to identify the model that best describes the data.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

In this tutorial, we described model and prior specification, estimation, and model comparison for hierarchical models in the Bayesian framework. In the following, we want to present several key recommendations.

As prior distributions influence model estimation and model comparison, it is important to choose suitable priors. We believe that simulating data based on the chosen priors is an intuitive tool to find reasonable prior settings that take the existing knowledge in the specific model application context into account. We showed that it can be problematic to use default priors without checking their suitability. Therefore, we recommend performing prior predictive checks with the priors that will be applied before running the analysis.

We illustrated different methods to obtain Bayes factors. We want to emphasize the importance of checking the variability of the Bayes factor estimate, the Bayes factor sensitivity to prior distributions, and to ensure that the Bayes factor estimate is based on a sufficient number of posterior samples.

This tutorial showed the application of two R packages: \texttt{rstan} and \texttt{brms}. Which package to use depends mostly on the complexity of the model and your technical background. For the non-technical user, we recommend using \texttt{brms} because of its intuitive model specification. However, the default prior setup can be unsuitable for a specific research context and might need to be adapted. Additionally, users are limited to predefined parametrizations. \texttt{rstan} offers the most flexibility in model and prior specification. This flexibility comes with a cost: if the prior or the model is not well defined, results can be misleading.

\hypertarget{further-recommendations}{%
\subsection{Further Recommendations}\label{further-recommendations}}

With this tutorial, we attempted to cover the most relevant topics that researchers encounter when starting with Bayesian hierarchical modeling. However, there are still many remaining issues and recommendations that we could not discuss. Here, we would like to point to further references in the literature if researchers want to continue mastering Bayesian hierarchical modeling.

First, we kept the discussion of the software packages and code fairly brief. We provide further code examples for the digit classification task in the Online Supplement. For different models, it may be useful to study additional tutorials and books. For \texttt{rstan}, we recommend the \texttt{Stan} manual (Stan Development Team, 2018b, 2018a). For \texttt{brms} there are several useful tutorials (Bürkner, 2017, 2018; Bürkner \& Vuorre, 2019). There are also several resources for Bayesian estimation with other samplers (Lee \& Wagenmakers, 2013; Rouder, Morey, \& Pratte, 2013). Two other useful tutorials for Bayesian modeling are Schad, Betancourt, and
Vasishth (2019) and Schad, Nicenboim, Bürkner, Betancourt, and Vasishth (2021).

Second, we did not discuss study planning. When planning an experiment, researchers may want to determine the number of participants required to gain conclusive results. Bayesian analysis has the advantage that an optional stopping paradigm may be employed (Rouder, 2014; Schönbrodt, Wagenmakers, Zehetleitner, \& Perugini, 2017). Schönbrodt and Wagenmakers (2018) and Stefan and colleagues (2019) introduced Bayes factor design analysis for fixed and sequential designs. However, these approaches to study planning have not yet been fully extended to a hierarchical setup (but see recent efforts by Vasishth et al., 2021). The extension to nested designs is not entirely trivial because, in addition to planning the number of participants, we also need to consider the number of trials (Rouder \& Haaf, 2018). While we think there remains some open questions on the topic, the tutorial by Stefan and colleagues (2019) is a good place to start for applied researchers.

\hypertarget{potential-technical-issues}{%
\subsection{Potential Technical Issues}\label{potential-technical-issues}}

One key advantage of Bayesian over frequentist hierarchical modeling is that complex modeling leads to fewer convergence issues in the Bayesian framework. This advantage, however, does not imply that there are no technical issues in the Bayesian framework. One possible technical issue is with the choice of priors when using \texttt{Stan} for estimation. The \texttt{Stan} algorithm is most efficient with specific prior distributions. For example, it may be beneficial to use a truncated \(t\)-distribution on standard deviation parameters instead of an inverse gamma distribution on variances (Gelman, 2006). Using priors not recommended by the \texttt{Stan} team may lead to more issues with convergence. In this tutorial, we highlight that choosing priors based on their predictions on data is beneficial. We still think this is the best path for substantive researchers. If one wants more flexibility in choices of priors or has continuous issues with convergence with their chosen prior distributions, then it might be worth considering using \texttt{JAGS} instead of \texttt{Stan}. \texttt{JAGS} (Plummer, 2003) has a similar functionality and syntax as \texttt{Stan}. While \texttt{Stan} tends to be more efficient in the sampling phase, especially in case of correlated parameters (Hecht, Weirich, \& Zitzmann, 2021), the \texttt{JAGS} algorithm could be more flexible and less affected by the choice of priors.

Another technical issue may arise with the estimation of Bayes factors. Here, we highlighted two approaches, the Savage-Dickey density ratio (Wagenmakers, Lodewyckx, Kuriyal, \& Grasman, 2010) and bridge sampling (Gronau et al., 2017). We already highlighted the drawbacks of the SD density ratio. Yet, bridge sampling may also fail in some cases. Failures of the bridge sampling algorithm are identified by unstable Bayes factors across repeated sampling runs. There are improvements to the algorithm that are specifically developed for hierarchical models (Gronau, Wagenmakers, Heck, \& Matzke, 2019). Yet, it remains crucial to assess the stability of Bayes factor estimates.

\hypertarget{living-with-uncertainty}{%
\subsection{Living with Uncertainty}\label{living-with-uncertainty}}

Let us reconsider our scenario from the introduction. The graduate student may have read our tutorial and went back to their supervisor to propose a Bayesian hierarchical analysis of their data. Compared to the initially considered ANOVA analysis, there are many small decisions to be made for this more extensive analysis. These small decisions tend to make researchers uncomfortable. What prior is the right one? Are there enough posterior samples?

In Bayesian modeling, we have to learn living with these uncertainties. In fact, we recommend embracing them as these small decisions. First, we at the very least are making these decisions deliberately. Second, we can check their impact on the results. Throughout the tutorial we highlighted several potential robustness checks, from prior predictions over convergence assessment, to sensitivity analysis for Bayes factors. These checks may help researchers to understand how robust their results are and when they break. Therefore, we hope the checks will support switching from simple procedures to Bayesian hierarchical modeling. Because ultimately, the results from such an analysis are much richer.

In summary, by offering a comparison of software packages, guidance on prior selection and default priors, and assessment of the Bayes factor as model comparison method, we hope that Bayesian hierarchical models will become available to a wider psychology public.

\newpage

\hypertarget{acknowledgement}{%
\section{Acknowledgement}\label{acknowledgement}}

We would like to express our gratitude to the R developers {[}Version 3.6.1; R Core Team (2019){]} and the following R package developers:

\begin{itemize}
\tightlist
\item
  Document layout: \emph{rmarkdown} {[}Version 1.16; Xie, Allaire, and Grolemund (2018){]}, \emph{papaja} {[}Version 0.1.0.9842; Aust and Barth (2018){]}, \emph{knitr} {[}Version 1.25; Xie (2015){]}, \emph{kableExtra} {[}Version 1.1.0; Zhu (2019){]}
\item
  Data analysis and cleaning: \emph{LaplacesDemon} {[}Version 16.1.4; Statisticat and LLC. (2020){]}, \emph{plyr} {[}Version 1.8.4; Wickham, François, Henry, and Müller (2019); Wickham (2011){]}, \emph{dplyr} {[}Version 0.8.3; Wickham, François, Henry, and Müller (2019){]}, \emph{readr} {[}Version 1.3.1; Wickham, Hester, and Francois (2018){]}, \emph{crch} {[}Version 1.0.4; Messner, Mayr, and Zeileis (2016); Messner, Mayr, and Zeileis (2017); Messner, Zeileis, Broecker, and Mayr (2014); Messner, Mayr, Zeileis, and Wilks (2014){]}, \emph{foreign} {[}Version 0.8.71; R Core Team (2018){]}, \emph{Matrix} {[}Version 1.2.17; Bates and Maechler (2019){]}, \emph{tidyr} {[}Version 1.0.0; Wickham and Henry (2019){]}
\item
  Model fit and comparison: \emph{rstan} {[}Version 2.19.2; Stan Development Team (2019a){]}, \emph{brms} {[}Version 2.13.0; Bürkner (2017); Bürkner (2018){]}, \emph{BayesFactor} {[}Version 0.9.12.4.2; Morey and Rouder (2018){]}, \emph{bridgesampling} {[}Version 1.0.0; Gronau, Singmann, and Wagenmakers (2020){]}
\item
  Visualization: \emph{coda} {[}Version 0.19.3; Plummer, Best, Cowles, and Vines (2006){]}, \emph{lattice} {[}Version 0.20.38; Sarkar (2008){]}, \emph{ggplot2} {[}Version 3.3.2; Wickham (2016){]}, \emph{Rmisc} {[}Version 1.5; Hope (2013){]}, \emph{devtools} {[}Version 2.2.1; Wickham, Hester, and Chang (2019){]}, \emph{gghalves} {[}Version 0.1.0; Tiedemann (2020){]}, \emph{bayesplot} {[}Version 1.7.1; Gabry, Simpson, Vehtari, Betancourt, and Gelman (2019){]}, \emph{gridExtra} {[}Version 2.3; Auguie (2017){]}, \emph{tibble} {[}Version 3.0.2; Müller and Wickham (2019){]}, \emph{ggmcmc} {[}Version 1.4.1; Fernández-i-Marín (2016){]}, \emph{cowplot} {[}Version 1.0.0; Wilke (2019){]}, \emph{ggridges} {[}Version 0.5.1; Wilke (2018){]}, \emph{circlize} {[}Version 0.4.9; Gu, Gu, Eils, Schlesner, and Brors (2014){]}, \emph{StanHeaders} {[}Version 2.19.0; Stan Development Team (2019b){]}, \emph{unikn} {[}Version 0.2.0; Neth and Gradwohl (2019){]}
\item
  Other: \emph{Rcpp} {[}Version 1.0.3; Eddelbuettel and François (2011); Eddelbuettel and Balamuta (2017){]}, \emph{usethis} {[}Version 1.5.1; Wickham and Bryan (2019){]}
\end{itemize}

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-aczel2018expert}{}%
Aczel, B., Hoekstra, R., Gelman, A., Wagenmakers, E.-J., Kluglist, I. G., Rouder, J. N., \ldots{} van Ravenzwaaij, D. (2018). Expert opinions on how to conduct and report {Bayesian} inference. doi:\href{https://doi.org/10.31234/osf.io/23m7f}{10.31234/osf.io/23m7f}

\leavevmode\hypertarget{ref-R-gridExtra}{}%
Auguie, B. (2017). \emph{gridExtra: Miscellaneous functions for "grid" graphics}. Retrieved from \url{https://CRAN.R-project.org/package=gridExtra}

\leavevmode\hypertarget{ref-R-papaja}{}%
Aust, F., \& Barth, M. (2018). \emph{{papaja}: {Create} {APA} manuscripts with {R Markdown}}. Retrieved from \url{https://github.com/crsh/papaja}

\leavevmode\hypertarget{ref-R-Matrix}{}%
Bates, D., \& Maechler, M. (2019). \emph{Matrix: Sparse and dense matrix classes and methods}. Retrieved from \url{https://CRAN.R-project.org/package=Matrix}

\leavevmode\hypertarget{ref-R-lme4}{}%
Bates, D., Mächler, M., Bolker, B., \& Walker, S. (2015). Fitting linear mixed-effects models using {lme4}. \emph{Journal of Statistical Software}, \emph{67}(1), 1--48. doi:\href{https://doi.org/10.18637/jss.v067.i01}{10.18637/jss.v067.i01}

\leavevmode\hypertarget{ref-betancourt2018conceptual}{}%
Betancourt, M. (2018). A {Conceptual Introduction} to {Hamiltonian Monte Carlo}. \emph{arXiv Preprint arXiv:1701.02434}.

\leavevmode\hypertarget{ref-betancourt2015hamiltonian}{}%
Betancourt, M., \& Girolami, M. (2015). {Hamiltonian Monte Carlo} for hierarchical models. \emph{Current Trends in {Bayesian} Methodology with Applications}, \emph{79}(30), 2--4.

\leavevmode\hypertarget{ref-R-brms_a}{}%
Bürkner, P.-C. (2017). {brms}: An {R} package for {Bayesian} multilevel models using {Stan}. \emph{Journal of Statistical Software}, \emph{80}(1), 1--28. doi:\href{https://doi.org/10.18637/jss.v080.i01}{10.18637/jss.v080.i01}

\leavevmode\hypertarget{ref-R-brms_b}{}%
Bürkner, P.-C. (2018). Advanced {Bayesian} multilevel modeling with the {R} package {brms}. \emph{The R Journal}, \emph{10}(1), 395--411. doi:\href{https://doi.org/10.32614/RJ-2018-017}{10.32614/RJ-2018-017}

\leavevmode\hypertarget{ref-burkner2019ordinal}{}%
Bürkner, P.-C., \& Vuorre, M. (2019). Ordinal regression models in psychology: A tutorial. \emph{Advances in Methods and Practices in Psychological Science}, \emph{2}(1), 77--101.

\leavevmode\hypertarget{ref-neff2018}{}%
Carpenter, B. (2018). We were measuring the speed of stan incorrectly -- it's faster than we thought in some cases due to antithetical sampling {[}blog post{]}. Retrieved from \url{https://statmodeling.stat.columbia.edu/2018/01/18/\%20measuring-speed-stan-incorrectly-faster-thought-cases-due-antithetical-sampling/}

\leavevmode\hypertarget{ref-carpenter2017stan}{}%
Carpenter, B., Gelman, A., Hoffman, M. D., Lee, D., Goodrich, B., Betancourt, M., \ldots{} Riddell, A. (2017). Stan: A probabilistic programming language. \emph{Journal of Statistical Software}, \emph{76}(1).

\leavevmode\hypertarget{ref-chib1995marginal}{}%
Chib, S. (1995). Marginal likelihood from the {Gibbs} output. \emph{Journal of the American Statistical Association}, \emph{90}, 1313--1321.

\leavevmode\hypertarget{ref-dickey1970weighted}{}%
Dickey, J. M., \& Lientz, B. P. (1970). The weighted likelihood ratio, sharp hypotheses about chances, the order of a {Markov} chain. \emph{The Annals of Mathematical Statistics}, \emph{41}, 214--226.

\leavevmode\hypertarget{ref-R-Rcpp_b}{}%
Eddelbuettel, D., \& Balamuta, J. J. (2017). {Extending extit{R} with extit{C++}: A Brief Introduction to extit{Rcpp}}. \emph{PeerJ Preprints}, \emph{5}, e3188v1. doi:\href{https://doi.org/10.7287/peerj.preprints.3188v1}{10.7287/peerj.preprints.3188v1}

\leavevmode\hypertarget{ref-R-Rcpp_a}{}%
Eddelbuettel, D., \& François, R. (2011). {Rcpp}: Seamless {R} and {C++} integration. \emph{Journal of Statistical Software}, \emph{40}(8), 1--18. doi:\href{https://doi.org/10.18637/jss.v040.i08}{10.18637/jss.v040.i08}

\leavevmode\hypertarget{ref-efron1977stein}{}%
Efron, B., \& Morris, C. (1977). Stein's paradox in statistics. \emph{Scientific American}, \emph{236}(5), 119--127.

\leavevmode\hypertarget{ref-etz2018introduction}{}%
Etz, A., \& Vandekerckhove, J. (2018). Introduction to {Bayesian} inference for psychology. \emph{Psychonomic Bulletin \& Review}, \emph{25}(1), 5--34. doi:\href{https://doi.org/10.3758/s13423-017-1262-3}{10.3758/s13423-017-1262-3}

\leavevmode\hypertarget{ref-R-ggmcmc}{}%
Fernández-i-Marín, X. (2016). {ggmcmc}: Analysis of {MCMC} samples and {B}ayesian inference. \emph{Journal of Statistical Software}, \emph{70}(9), 1--20. doi:\href{https://doi.org/10.18637/jss.v070.i09}{10.18637/jss.v070.i09}

\leavevmode\hypertarget{ref-Fink1997}{}%
Fink, D. (1997). A compendium of conjugate priors. https://www.johndcook.com/CompendiumOfConjugatePriors.pdf.

\leavevmode\hypertarget{ref-R-bayesplot}{}%
Gabry, J., Simpson, D., Vehtari, A., Betancourt, M., \& Gelman, A. (2019). Visualization in {Bayesian} workflow. \emph{Journal of the Royal Statistical Society}, \emph{182}, 389--402. doi:\href{https://doi.org/10.1111/rssa.12378}{10.1111/rssa.12378}

\leavevmode\hypertarget{ref-gelman2006prior}{}%
Gelman, A. (2006). Prior distributions for variance parameters in hierarchical models (comment on article by {Browne} and {Draper}). \emph{Bayesian Analysis}, \emph{1}(3), 515--534.

\leavevmode\hypertarget{ref-gelman2013bayesian}{}%
Gelman, A., Carlin, J. B., Stern, H. S., Dunson, D. B., Vehtari, A., \& Rubin, D. B. (2013). \emph{Bayesian data analysis}. CRC press.

\leavevmode\hypertarget{ref-gelman1992inference}{}%
Gelman, A., \& Rubin, D. B. (1992). Inference from iterative simulation using multiple sequences. \emph{Statistical Science}, \emph{7}(4), 457--472.

\leavevmode\hypertarget{ref-green2015bayesian}{}%
Green, P. J., Łatuszyński, K., Pereyra, M., \& Robert, C. P. (2015). Bayesian computation: A summary of the current state, and samples backwards and forwards. \emph{Statistics and Computing}, \emph{25}(4), 835--862.

\leavevmode\hypertarget{ref-gronau2017tutorial}{}%
Gronau, Q. F., Sarafoglou, A., Matzke, D., Ly, A., Boehm, U., Marsman, M., \ldots{} Steingroever, H. (2017). A tutorial on bridge sampling. \emph{Journal of Mathematical Psychology}, \emph{81}, 80--97.

\leavevmode\hypertarget{ref-R-bridgesampling}{}%
Gronau, Q. F., Singmann, H., \& Wagenmakers, E.-J. (2020). {bridgesampling}: An {R} package for estimating normalizing constants. \emph{Journal of Statistical Software}, \emph{92}(10), 1--29. doi:\href{https://doi.org/10.18637/jss.v092.i10}{10.18637/jss.v092.i10}

\leavevmode\hypertarget{ref-gronau2019simple}{}%
Gronau, Q. F., Wagenmakers, E.-J., Heck, D. W., \& Matzke, D. (2019). A simple method for comparing complex models: {Bayesian} model comparison for hierarchical multinomial processing tree models using {Warp}-III bridge sampling. \emph{Psychometrika}, \emph{84}(1), 261--284.

\leavevmode\hypertarget{ref-R-circlize}{}%
Gu, Z., Gu, L., Eils, R., Schlesner, M., \& Brors, B. (2014). Circlize implements and enhances circular visualization in {R}. \emph{Bioinformatics}, \emph{30}, 2811--2812.

\leavevmode\hypertarget{ref-haaf2019capturing}{}%
Haaf, J. M., Klaassen, F., \& Rouder, J. (2019). Capturing ordinal theoretical constraint in psychological science.

\leavevmode\hypertarget{ref-haaf2017developing}{}%
Haaf, J. M., \& Rouder, J. N. (2017). Developing constraint in {Bayesian} mixed models. \emph{Psychological Methods}, \emph{22}(4), 779.

\leavevmode\hypertarget{ref-haaf2019some}{}%
Haaf, J. M., \& Rouder, J. N. (2019). Some do and some don't? Accounting for variability of individual difference structures. \emph{Psychonomic Bulletin \& Review}, \emph{26}(3), 772--789.

\leavevmode\hypertarget{ref-hecht2021comparing}{}%
Hecht, M., Weirich, S., \& Zitzmann, S. (2021). Comparing the MCMC efficiency of JAGS and stan for the multi-level intercept-only model in the covariance-and mean-based and classic parametrization. \emph{Psych}, \emph{3}(4), 751--779.

\leavevmode\hypertarget{ref-heck2019caveat}{}%
Heck, D. W. (2019). A caveat on the {Savage--Dickey} density ratio: The case of computing {Bayes} factors for regression parameters. \emph{British Journal of Mathematical and Statistical Psychology}, \emph{72}(2), 316--333.

\leavevmode\hypertarget{ref-hobert1996effect}{}%
Hobert, J. P., \& Casella, G. (1996). The effect of improper priors on {Gibbs} sampling in hierarchical linear mixed models. \emph{Journal of the American Statistical Association}, \emph{91}(436), 1461--1473.

\leavevmode\hypertarget{ref-hoffman2014no}{}%
Hoffman, M. D., \& Gelman, A. (2014). The no-u-turn sampler: Adaptively setting path lengths in {Hamiltonian Monte Carlo}. \emph{J. Mach. Learn. Res.}, \emph{15}(1), 1593--1623.

\leavevmode\hypertarget{ref-R-Rmisc}{}%
Hope, R. M. (2013). \emph{Rmisc: Rmisc: Ryan miscellaneous}. Retrieved from \url{https://CRAN.R-project.org/package=Rmisc}

\leavevmode\hypertarget{ref-jeffreys1961theory}{}%
Jeffreys, H. (1961). Theory of probability.,(oxford university press: Oxford, UK).

\leavevmode\hypertarget{ref-kass1995bayes}{}%
Kass, R. E., \& Raftery, A. E. (1995). Bayes factors. \emph{Journal of the American Statistical Association}, \emph{90}(430), 773--795.

\leavevmode\hypertarget{ref-logspline}{}%
Kooperberg, C. (2019). \emph{Logspline: Routines for logspline density estimation}. Retrieved from \url{https://CRAN.R-project.org/package=logspline}

\leavevmode\hypertarget{ref-j_van_langen_2020_3715576}{}%
Langen, J. van. (2020). Open-visualizations in r and python (Version v.1.0.4). Zenodo. doi:\href{https://doi.org/10.5281/zenodo.3715576}{10.5281/zenodo.3715576}

\leavevmode\hypertarget{ref-lee2011cognitive}{}%
Lee, M. D. (2011). How cognitive modeling can benefit from hierarchical {Bayesian} models. \emph{Journal of Mathematical Psychology}, \emph{55}(1), 1--7.

\leavevmode\hypertarget{ref-lee2013bayesian}{}%
Lee, M., \& Wagenmakers, E.-J. (2013). Bayesian data analysis for cognitive science: A practical course. New York, NY: Cambridge University Press.

\leavevmode\hypertarget{ref-lindley1957statistical}{}%
Lindley, D. V. (1957). A statistical paradox. \emph{Biometrika}, \emph{44}(1/2), 187--192.

\leavevmode\hypertarget{ref-lynch2007introduction}{}%
Lynch, S. M. (2007). \emph{Introduction to applied bayesian statistics and estimation for social scientists}. Springer Science \& Business Media.

\leavevmode\hypertarget{ref-meng1996simulating}{}%
Meng, X.-L., \& Wong, W. H. (1996). Simulating ratios of normalizing constants via a simple identity: A theoretical exploration. \emph{Statistica Sinica}, 831--860.

\leavevmode\hypertarget{ref-R-crch_a}{}%
Messner, J. W., Mayr, G. J., \& Zeileis, A. (2016). Heteroscedastic censored and truncated regression with {crch}. \emph{The R Journal}, \emph{8}(1), 173--181. Retrieved from \url{https://journal.r-project.org/archive/2016-1/messner-mayr-zeileis.pdf}

\leavevmode\hypertarget{ref-R-crch_b}{}%
Messner, J. W., Mayr, G. J., \& Zeileis, A. (2017). Non-homogeneous boosting for predictor selection in ensemble post-processing. \emph{Monthly Weather Review}, \emph{145}(1), 137--147. doi:\href{https://doi.org/10.1175/MWR-D-16-0088.1}{10.1175/MWR-D-16-0088.1}

\leavevmode\hypertarget{ref-R-crch_d}{}%
Messner, J. W., Mayr, G. J., Zeileis, A., \& Wilks, D. S. (2014). Heteroscedastic extended logistic regression for postprocessing of ensemble guidance. \emph{Monthly Weather Review}, \emph{142}(1), 448--456. doi:\href{https://doi.org/10.1175/MWR-D-13-00271.1}{10.1175/MWR-D-13-00271.1}

\leavevmode\hypertarget{ref-R-crch_c}{}%
Messner, J. W., Zeileis, A., Broecker, J., \& Mayr, G. J. (2014). Probabilistic wind power forecasts with an inverse power curve transformation and censored regression. \emph{Wind Energy}, \emph{17}(11), 1753--1766. doi:\href{https://doi.org/10.1002/we.1666}{10.1002/we.1666}

\leavevmode\hypertarget{ref-R-BayesFactor}{}%
Morey, R. D., \& Rouder, J. N. (2018). \emph{BayesFactor: Computation of bayes factors for common designs}. Retrieved from \url{https://CRAN.R-project.org/package=BayesFactor}

\leavevmode\hypertarget{ref-morey2011using}{}%
Morey, R. D., Rouder, J. N., Pratte, M. S., \& Speckman, P. L. (2011). Using MCMC chain outputs to efficiently estimate bayes factors. \emph{Journal of Mathematical Psychology}, \emph{55}(5), 368--378.

\leavevmode\hypertarget{ref-moyer1967time}{}%
Moyer, R. S., \& Landauer, T. K. (1967). Time required for judgements of numerical inequality. \emph{Nature}, \emph{215}(5109), 1519--1520.

\leavevmode\hypertarget{ref-R-tibble}{}%
Müller, K., \& Wickham, H. (2019). \emph{Tibble: Simple data frames}. Retrieved from \url{https://CRAN.R-project.org/package=tibble}

\leavevmode\hypertarget{ref-R-unikn}{}%
Neth, H., \& Gradwohl, N. (2019). \emph{Unikn: Graphical elements of the university of konstanz's corporate design}. Konstanz, Germany: Social Psychology; Decision Sciences, University of Konstanz. Retrieved from \url{https://CRAN.R-project.org/package=unikn}

\leavevmode\hypertarget{ref-nicenboim2021introduction}{}%
Nicenboim, B., Schad, D., \& Vasishth, S. (2021). An introduction to {Bayesian} data analysis for cognitive science. Under contract with Chapman; Hall/CRC Statistics in the Social and~\ldots. Retrieved from \url{https://vasishth.github.io/bayescogsci/book/}

\leavevmode\hypertarget{ref-Plummer03jagsa}{}%
Plummer, M. (2003). JAGS: A program for analysis of bayesian graphical models using gibbs sampling.

\leavevmode\hypertarget{ref-R-coda}{}%
Plummer, M., Best, N., Cowles, K., \& Vines, K. (2006). CODA: Convergence diagnosis and output analysis for MCMC. \emph{R News}, \emph{6}(1), 7--11. Retrieved from \url{https://journal.r-project.org/archive/}

\leavevmode\hypertarget{ref-R-foreign}{}%
R Core Team. (2018). \emph{Foreign: Read data stored by 'minitab', 's', 'SAS', 'SPSS', 'stata', 'systat', 'weka', 'dBase', ...} Retrieved from \url{https://CRAN.R-project.org/package=foreign}

\leavevmode\hypertarget{ref-R-base}{}%
R Core Team. (2019). \emph{R: A language and environment for statistical computing}. Vienna, Austria: R Foundation for Statistical Computing. Retrieved from \url{https://www.R-project.org/}

\leavevmode\hypertarget{ref-roos2015sensitivity}{}%
Roos, M., Martins, T. G., Held, L., \& Rue, H. (2015). Sensitivity analysis for {Bayesian} hierarchical models. \emph{Bayesian Analysis}, \emph{10}(2), 321--349.

\leavevmode\hypertarget{ref-rouder2014optional}{}%
Rouder, J. N. (2014). Optional stopping: No problem for {Bayesians}. \emph{Psychonomic Bulletin \& Review}, \emph{21}(2), 301--308.

\leavevmode\hypertarget{ref-rouder2018power}{}%
Rouder, J. N., \& Haaf, J. M. (2018). Power, dominance, and constraint: A note on the appeal of different design traditions. \emph{Advances in Methods and Practices in Psychological Science}, \emph{1}(1), 19--26.

\leavevmode\hypertarget{ref-rouder2018theories}{}%
Rouder, J. N., Haaf, J. M., \& Aust, F. (2018). From theories to models to predictions: A {Bayesian} model comparison approach. \emph{Communication Monographs}, \emph{85}(1), 41--56.

\leavevmode\hypertarget{ref-rouder2005introduction}{}%
Rouder, J. N., \& Lu, J. (2005). An introduction to {Bayesian} hierarchical models with an application in the theory of signal detection. \emph{Psychonomic Bulletin \& Review}, \emph{12}(4), 573--604.

\leavevmode\hypertarget{ref-rouder2005hierarchical}{}%
Rouder, J. N., Lu, J., Speckman, P., Sun, D., \& Jiang, Y. (2005). A hierarchical model for estimating response time distributions. \emph{Psychonomic Bulletin \& Review}, \emph{12}(2), 195--223.

\leavevmode\hypertarget{ref-rouder2018teaching}{}%
Rouder, J. N., \& Morey, R. D. (2018). Teaching {Bayes}' theorem: Strength of evidence as predictive accuracy. \emph{The American Statistician}.

\leavevmode\hypertarget{ref-rouder2013hierarchical}{}%
Rouder, J. N., Morey, R. D., \& Pratte, M. S. (2013). Hierarchical {Bayesian} models. \emph{Practice}, \emph{1}(5), 10.

\leavevmode\hypertarget{ref-R-lattice}{}%
Sarkar, D. (2008). \emph{Lattice: Multivariate data visualization with r}. New York: Springer. Retrieved from \url{http://lmdvr.r-forge.r-project.org}

\leavevmode\hypertarget{ref-schad2019toward}{}%
Schad, D. J., Betancourt, M., \& Vasishth, S. (2019). Toward a principled {Bayesian} workflow in cognitive science. \emph{arXiv Preprint arXiv:1904.12765}.

\leavevmode\hypertarget{ref-schad2021workflow}{}%
Schad, D. J., Nicenboim, B., Bürkner, P.-C., Betancourt, M., \& Vasishth, S. (2021). Workflow techniques for the robust use of bayes factors. Retrieved from \url{http://arxiv.org/abs/2103.08744}

\leavevmode\hypertarget{ref-schonbrodt2018bayes}{}%
Schönbrodt, F. D., \& Wagenmakers, E.-J. (2018). Bayes factor design analysis: Planning for compelling evidence. \emph{Psychonomic Bulletin \& Review}, \emph{25}(1), 128--142.

\leavevmode\hypertarget{ref-schonbrodt2017sequential}{}%
Schönbrodt, F. D., Wagenmakers, E.-J., Zehetleitner, M., \& Perugini, M. (2017). Sequential hypothesis testing with {Bayes} factors: Efficiently testing mean differences. \emph{Psychological Methods}, \emph{22}(2), 322.

\leavevmode\hypertarget{ref-schramm2019reaction}{}%
Schramm, P., \& Rouder, J. (2019). Are reaction time transformations really beneficial?

\leavevmode\hypertarget{ref-shiffrin2008survey}{}%
Shiffrin, R. M., Lee, M. D., Kim, W., \& Wagenmakers, E.-J. (2008). A survey of model evaluation approaches with a tutorial on hierarchical {Bayesian} methods. \emph{Cognitive Science}, \emph{32}(8), 1248--1284.

\leavevmode\hypertarget{ref-sorensen2015bayesian}{}%
Sorensen, T., \& Vasishth, S. (2015). Bayesian linear mixed models using stan: A tutorial for psychologists, linguists, and cognitive scientists. \emph{arXiv Preprint arXiv:1506.06201}.

\leavevmode\hypertarget{ref-stan2018b}{}%
Stan Development Team. (2018a). Stan modeling language reference manual, version 2.18.0. Retrieved from \url{https://mc-stan.org/docs/2_18/reference-manual/}

\leavevmode\hypertarget{ref-stan2018a}{}%
Stan Development Team. (2018b). Stan modeling language users guide, version 2.18.0. Retrieved from \url{https://mc-stan.org/docs/2_18/stan-users-guide}

\leavevmode\hypertarget{ref-R-rstan}{}%
Stan Development Team. (2019a). {RStan}: The {R} interface to {Stan}. Retrieved from \url{http://mc-stan.org/}

\leavevmode\hypertarget{ref-R-StanHeaders}{}%
Stan Development Team. (2019b). {StanHeaders}: Headers for the {R} interface to {Stan}. Retrieved from \url{http://mc-stan.org/}

\leavevmode\hypertarget{ref-R-LaplacesDemon}{}%
Statisticat, \& LLC. (2020). \emph{LaplacesDemon: Complete environment for {Bayesian} inference}. Bayesian-Inference.com. Retrieved from \url{https://web.archive.org/web/20150206004624/http://www.bayesian-inference.com/software}

\leavevmode\hypertarget{ref-Stefan_etal2021}{}%
Stefan, A., Katsimpokis, D., Gronau, Q. F., \& Wagenmakers, E.-J. (2021). Expert agreement in prior elicitation and its effects on {Bayesian} inference. \emph{PsyArXiv Preprint}. doi:\href{https://doi.org/10.31234/osf.io/8xkqd}{10.31234/osf.io/8xkqd}

\leavevmode\hypertarget{ref-stefan2019tutorial}{}%
Stefan, A. M., Gronau, Q. F., Schönbrodt, F. D., \& Wagenmakers, E.-J. (2019). A tutorial on {Bayes} factor design analysis using an informed prior. \emph{Behavior Research Methods}, \emph{51}(3), 1042--1058.

\leavevmode\hypertarget{ref-R-gghalves}{}%
Tiedemann, F. (2020). \emph{Gghalves: Compose half-half plots using your favourite geoms}. Retrieved from \url{https://github.com/erocoar/gghalves}

\leavevmode\hypertarget{ref-van2021bayes}{}%
van Doorn, J., Aust, F., Haaf, J. M., Stefan, A., \& Wagenmakers, E.-J. (2021). Bayes factors for mixed models. \emph{PsyArXiv Preprint}. doi:\href{https://doi.org/10.31234/osf.io/y65h8}{10.31234/osf.io/y65h8}

\leavevmode\hypertarget{ref-van2018simple}{}%
Van Ravenzwaaij, D., Cassey, P., \& Brown, S. D. (2018). A simple introduction to {Markov Chain Monte--Carlo} sampling. \emph{Psychonomic Bulletin \& Review}, \emph{25}(1), 143--154.

\leavevmode\hypertarget{ref-vasishth2021sample}{}%
Vasishth, S., Yadav, H., Schad, D., \& Nicenboim, B. (2021). Sample size determination for bayesian hierarchical models commonly used in psycholinguistics.

\leavevmode\hypertarget{ref-vehtari2019rank}{}%
Vehtari, A., Gelman, A., Simpson, D., Carpenter, B., \& Bürkner, P.-C. (2019). \emph{arXiv Preprint arXiv:1903.08008}.

\leavevmode\hypertarget{ref-Vehtari_2021}{}%
Vehtari, A., Gelman, A., Simpson, D., Carpenter, B., \& Bürkner, P.-C. (2021). Rank-normalization, folding, and localization: An improved rˆ for assessing convergence of MCMC (with discussion). \emph{Bayesian Analysis}, \emph{16}(2). doi:\href{https://doi.org/10.1214/20-ba1221}{10.1214/20-ba1221}

\leavevmode\hypertarget{ref-wagenmakers2010bayesian}{}%
Wagenmakers, E.-J., Lodewyckx, T., Kuriyal, H., \& Grasman, R. (2010). Bayesian hypothesis testing for psychologists: A tutorial on the {Savage--Dickey} method. \emph{Cognitive Psychology}, \emph{60}(3), 158--189.

\leavevmode\hypertarget{ref-wagenmakers2018bayesian}{}%
Wagenmakers, E.-J., Marsman, M., Jamil, T., Ly, A., Verhagen, J., Love, J., \ldots{} others. (2018). Bayesian inference for psychology. Part i: Theoretical advantages and practical ramifications. \emph{Psychonomic Bulletin \& Review}, \emph{25}(1), 35--57.

\leavevmode\hypertarget{ref-R-plyr}{}%
Wickham, H. (2011). The split-apply-combine strategy for data analysis. \emph{Journal of Statistical Software}, \emph{40}(1), 1--29. Retrieved from \url{http://www.jstatsoft.org/v40/i01/}

\leavevmode\hypertarget{ref-R-ggplot2}{}%
Wickham, H. (2016). \emph{ggplot2: Elegant graphics for data analysis}. Springer-Verlag New York. Retrieved from \url{https://ggplot2.tidyverse.org}

\leavevmode\hypertarget{ref-R-usethis}{}%
Wickham, H., \& Bryan, J. (2019). \emph{Usethis: Automate package and project setup}. Retrieved from \url{https://CRAN.R-project.org/package=usethis}

\leavevmode\hypertarget{ref-R-dplyr}{}%
Wickham, H., François, R., Henry, L., \& Müller, K. (2019). \emph{Dplyr: A grammar of data manipulation}. Retrieved from \url{https://CRAN.R-project.org/package=dplyr}

\leavevmode\hypertarget{ref-R-tidyr}{}%
Wickham, H., \& Henry, L. (2019). \emph{Tidyr: Tidy messy data}. Retrieved from \url{https://CRAN.R-project.org/package=tidyr}

\leavevmode\hypertarget{ref-R-devtools}{}%
Wickham, H., Hester, J., \& Chang, W. (2019). \emph{Devtools: Tools to make developing r packages easier}. Retrieved from \url{https://CRAN.R-project.org/package=devtools}

\leavevmode\hypertarget{ref-R-readr}{}%
Wickham, H., Hester, J., \& Francois, R. (2018). \emph{Readr: Read rectangular text data}. Retrieved from \url{https://CRAN.R-project.org/package=readr}

\leavevmode\hypertarget{ref-R-ggridges}{}%
Wilke, C. O. (2018). \emph{Ggridges: Ridgeline plots in 'ggplot2'}. Retrieved from \url{https://CRAN.R-project.org/package=ggridges}

\leavevmode\hypertarget{ref-R-cowplot}{}%
Wilke, C. O. (2019). \emph{Cowplot: Streamlined plot theme and plot annotations for 'ggplot2'}. Retrieved from \url{https://CRAN.R-project.org/package=cowplot}

\leavevmode\hypertarget{ref-R-knitr}{}%
Xie, Y. (2015). \emph{Dynamic documents with {R} and knitr} (2nd ed.). Boca Raton, Florida: Chapman; Hall/CRC. Retrieved from \url{https://yihui.name/knitr/}

\leavevmode\hypertarget{ref-R-rmarkdown}{}%
Xie, Y., Allaire, J. J., \& Grolemund, G. (2018). \emph{R markdown: The definitive guide}. Boca Raton, Florida: Chapman; Hall/CRC. Retrieved from \url{https://bookdown.org/yihui/rmarkdown}

\leavevmode\hypertarget{ref-R-kableExtra}{}%
Zhu, H. (2019). \emph{kableExtra: Construct complex table with 'kable' and pipe syntax}. Retrieved from \url{https://CRAN.R-project.org/package=kableExtra}

\end{CSLReferences}

\endgroup

\newpage

\end{document}

### 2022-06-09 G - Trace Plots /savetraceplotstopdf.R

@@ -1,0 +1,61 @@
+# Rstan traceplots all parameters 
+# Load packages 
+library(rstan)
+library(brms)
+
+# Rstan 
+# Normal model 
+# Load normal model fit 
+hier_modelc_adj_v2 <- readRDS("/Users/myrtheveenman/hier_modelc_adj09042021_v2.rds")
+
+names(hier_modelc_adj_v2)  # all parameter names 
+# Plot figure 
+# For one traceplot 
+pdf("traceplots_normalmodel.pdf", onefile = TRUE)
+plot(hier_modelc_adj_v2, plotfun = "trace", pars = names(hier_modelc_adj_v2)[1])
+dev.off()
+ 
+# Save all traceplots in .pdf 
+pdf("traceplots_normalmodel_rstan.pdf", onefile = TRUE)
+for (i in 1:326){
+  print(plot(hier_modelc_adj_v2, plotfun = "trace", pars = names(hier_modelc_adj_v2)[i]))  # rstan 
+}
+dev.off()
+
+
+# Log-normal model 
+# Load log-normal model fit 
+hier_logmodelc_adj13012022 <- readRDS("/Users/myrtheveenman/Documents/GitHub/multilevel-tutorial-paper/R objects/rstan/non-linear model/hier_logmodelc_adj13012022.rds")
+
+# Save traceplots in .pdf 
+pdf("traceplots_lognormalmodel_rstan.pdf", onefile = TRUE)
+for (i in 1:326){
+  print(plot(hier_logmodelc_adj13012022, plotfun = "trace", pars = names(hier_logmodelc_adj13012022)[i]))   # rstan 
+}
+dev.off()
+
+# Brms 
+# Normal model 
+# Load normal model fit 
+model_fit <- readRDS(file = "/Users/myrtheveenman/Documents/GitHub/multilevel-tutorial-paper/R objects/brms/Truncated/BRMSmodelfit18_03012022.rds") 
+variables(model_fit)
+
+# Save traceplots in .pdf 
+pdf("traceplots_normalmodel_brms.pdf", onefile = TRUE)
+for(i in 1:326){
+  print(mcmc_plot(model_fit, variable = variables(model_fit)[i], type = "trace") + scale_color_manual(values=c("#E66101", "#998EC3", "#542788", "#F1A340")) + scale_x_continuous(breaks=c(0,1000,2000, 3000, 4000, 5000), 
[…]

### 2022-06-09 R code/Brms/Log-normal model/fitnonlinearhierarchicalmodel_brms.R

@@ -1,0 +1,136 @@
+# non linear hierarchical model in brms 
+
+formula3 <- rt2 ~ 1 + side + dif1 + dif2 + dif3 + dif4 +   # General effects 
+  (1 + side + dif1 + dif2 + dif3 + dif4 || sub1)
+
+indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
+colnames(indat)=c('sub','block','trial','stim','resp','rt','error')
+
+## Cleaning the data according to criteria discussed in paper 
+# (code retrieved from Julia Haaf:
+# https://github.com/PerceptionAndCognitionLab/bf-order/blob/public/papers/submission/R-scripts/ld5.R)
+clean=function()
+{
+  indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
+  colnames(indat)=c('sub','block','trial','stim','resp','rt','error')
+  
+  bad1=indat$sub%in%c(34,43)
+  bad2=indat$rt<250 | indat$rt>2000
+  bad3=indat$err==1
+  bad4=indat$block==0 & indat$trial<20
+  bad5=indat$trial==0
+  
+  bad=bad1 | bad2 | bad3 |bad4 |bad5
+  dat=indat[!bad,]
+  return(dat)
+}
+
+indat1 <- clean()  # Final dataset 
+
+### Response time from milliseconds to seconds 
+indat1$rt2 <- indat1$rt/1000
+
+### Add variable with information of the side of the digit to the list, for parameter beta    
+# j = condition, if j > 5 than x = -1/2, if j < 5 than x = 1/2 
+# indat1$stim 0 = 2, 1=3, 2=4, 3=6, 4=7, 5=8 
+# so if bigger than 2 than 1/2, else than -1/2 
+for (j in 1:nrow(indat1)){
+  if (indat1$stim[j] < 3) {
+    indat1$side[j] <- 1/2}
+  else indat1$side[j] <- -1/2
+}
+
+### Add variable with information about difference between digits to the list, for parameters delta's 
+#### Difference 8 and 7 
+for (j in 1:nrow(indat1)){ 
[…]

### 2022-06-09 R code/Brms/Normal model/brms_model_cleaned_trunc.R

@@ -1,0 +1,165 @@
+# New brms fit 
+
+library(brms)
+
+################################################################################
+## 1. Load data 
+################################################################################
+indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
+colnames(indat)=c('sub','block','trial','stim','resp','rt','error')
+
+## Cleaning the data according to criteria discussed in paper 
+clean=function()
+{
+  indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
+  colnames(indat)=c('sub','block','trial','stim','resp','rt','error')
+  
+  bad1=indat$sub%in%c(34,43)
+  bad2=indat$rt<250 | indat$rt>2000
+  bad3=indat$err==1
+  bad4=indat$block==0 & indat$trial<20
+  bad5=indat$trial==0
+  
+  bad=bad1 | bad2 | bad3 |bad4 |bad5
+  dat=indat[!bad,]
+  return(dat)
+}
+
+indat1 <- clean()  # Final dataset 
+
+## Data add necessary information 
+### Add participant number 
+rank(unique(indat1$sub))
+sub <- rep(1:N, each = J)
+true.theta[sub]
+
+# create variable inter for the intercept gamma 
+indat1$inter <- rep(1, nrow(indat1))
+
+for (j in 1:nrow(indat1)){
+  if (j == 1) {
+    indat1$sub1[j] <- 1}
+  else if (indat1$sub[j] == indat1$sub[j-1]) {
+    indat1$sub1[j] <- indat1$sub1[j-1]}
+  else indat1$sub1[j] <- indat1$sub1[j-1] + 1
+}
+
+### Add variable with information of the side of the digit to the list, for parameter beta    
+# j = condition, if j > 5 than x = -1/2, if j < 5 than x = 1/2 
+# indat1$stim 0 = 2, 1=3, 2=4, 3=6, 4=7, 5=8 
+# so if bigger than 2 than 1/2, else than -1/2 
+for (j in 1:nrow(indat1)){
+  if (indat1$stim[j] < 3) {
+    indat1$side[j] <- 1/2}
+  else indat1$side[j] <- -1/2
+}
+
+### Add variable with information about difference between digits to the list, for parameters delta's 
+#### Difference 8 and 7 
+for (j in 1:nrow(indat1)){ 
[…]

### 2022-06-09 R code/Figures/Fig1.R

@@ -1,0 +1,111 @@
+# Efron & Moris figure 
+# Model 
+# Load output BayesFactor package: 
+samplesindat5 <- readRDS("R objects/BF model to create figure 1B/BFmodelfit15.rds")
+meanbfindat <- colMeans(samplesindat5)  # now BF model output with seperate effect mu's and different r scale (all set to 1)
+
+# From data, from proposal r code 
+# Observations 
+indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
+colnames(indat)=c('sub','block','trial','stim','resp','rt','error')
+
+clean=function()
+{
+  indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
+  colnames(indat)=c('sub','block','trial','stim','resp','rt','error')
+  
+  bad1=indat$sub%in%c(34,43)
+  bad2=indat$rt<250 | indat$rt>2000
+  bad3=indat$err==1
+  bad4=indat$block==0 & indat$trial<20
+  bad5=indat$trial==0
+  
+  bad=bad1 | bad2 | bad3 |bad4 |bad5
+  dat=indat[!bad,]
+  return(dat)
+}
+
+
+# x axis digit 
+# y axis: mean RT pp 
+indat1 <- clean()
+dat <- indat1  # Save the dataframe in new object 
+
+sub <- as.integer(as.factor(dat$sub))
+I <- max(sub)
+
+deltabf2 <- meanbfindat[(3*(I+1) + 1)] + meanbfindat[(3*(I+1) + 2):(4*(I+1))]
+deltabf2  # digit two estimates 
+
+
+# Mean observations 
+meanrtindat <- aggregate(indat1$rt, list(indat1$stim, indat1$sub), mean)  # mean response time per condition per participant 
+meanrt <- aggregate(indat1$rt, list(indat1$stim), mean)
+for (i in 1:nrow(meanrtindat)){
+  if (meanrtindat$Group.1[i] == 0) {meanrtindat$Group.1[i] <- 2 
+  } else if(meanrtindat$Group.1[i] == 1) {meanrtindat$Group.1[i] <- 3 
+  } else if(meanrtindat$Group.1[i] == 2) {meanrtindat$Group.1[i] <- 4
[…]

### 2022-06-09 R code/Figures/Fig10_randomparfigure_trunc_v2.R

@@ -1,0 +1,345 @@
+# Individual effects from Rstan 
+# For rstan the output is saved in hier_modelc_adj_07012022 (adjust "myPath" below to your directory)
+hier_modelc_adj_07012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
+
+# For brms the output is saved in object brmmodel8_new03012022 (this is without saving the prior samples)
+brmmodel8_new03012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/brms/Posterior/brmmodel8_new03012022.rds")  
+
+library(rstan)
+library(brms)
+library(gridExtra)
+
+indeffplot <- function(hier_modelc_adj_v2, brmmodel8_v2){
+  
+  # Only gamma's 
+  gamnamesstan <- paste("gammai", "[",1:52,"]", sep="")
+  gamstan <- as.data.frame(summary(hier_modelc_adj_v2, pars = c( gamnamesstan))$summary)
+  gamstanmu <- as.data.frame(summary(hier_modelc_adj_v2, pars = c("mu"))$summary)
+  gamstan$parameters <- c(gamnamesstan)
+  
+  for (i in 1:nrow(gamstan)){  # 1 if CI contains zero, 0 otherwise 
+    if (gamstan$`2.5%`[i]*gamstan$`97.5%`[i] < 0) {
+      gamstan$zero2[i] <- "pink"
+    } else (gamstan$zero2[i] <- "blue") 
+  }
+  
+  
+  gamstan2 <- gamstan[order(gamstan$mean),]
+  gamstan2$order <- c(1:52)
+  
+  gammarandomplot <- ggplot(gamstan2, aes(x = order, y = mean, ymin = `2.5%`, ymax = `97.5%`)) + 
+    geom_linerange(aes(color = "blue"), color = "#00BFC4") + 
+    geom_point() + 
+    geom_hline(yintercept = gamstanmu$mean, linetype = "dashed", alpha = .3) + 
+    labs(title = expression(~gamma[~i]), y = "Estimation") + 
+    theme(legend.position = "none", 
+          axis.title.x=element_blank(),
+          axis.text.x=element_blank(),
[…]

### 2022-06-09 R code/Figures/Fig11_modelestimatesfigure_truncated_onlyrstan.R

@@ -1,0 +1,102 @@
+# Model estimates figure 
+# Load fits 
+# For rstan the output is saved in hier_modelc_adj_07012022 (adjust "myPath" below to your directory)
+hier_modelc_adj_07012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
+
+library(rstan)
+
+# Plot results using https://github.com/jorvlan/open-visualizations/blob/master/R/repmes_tutorial_R.pdf
+# Install packages for plot 
+packages <- c("plyr", "lattice", "ggplot2", "dplyr", "readr", "rmarkdown","Rmisc")
+if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
+  install.packages(setdiff(packages, rownames(installed.packages())))  
+}
+if (!require(devtools)) {
+  install.packages("devtools")
+}
+devtools::install_github('erocoar/gghalves')
+
+library("plyr")
+library("lattice")
+library("ggplot2")
+library("dplyr")
+library("readr")
+library("rmarkdown")
+library("Rmisc")
+library("devtools")
+library("gghalves")
+
+indvestviolin_onlyrstan <- function(hier_modelc_adj_v2, brmmodel8_v2){
+  # Rstan 
+  ## Delta's 
+  ## Create dataframe 
+  parameter_m7 <- rep(c("Delta 1", "Delta 2", "Delta 3", "Delta 4"), each = 52 )
+  
+  ### Obtain all individual estimates delta 
+  hier_modeld <- hier_modelc_adj_v2
+  delta1_summary <- summary(hier_modeld, pars = c("delta1"), probs = c(0.1, 0.9))$summary
+  
+  delta2_summary <- summary(hier_modeld, pars = c("delta2"), probs = c(0.1, 0.9))$summary
+  
+  delta3_summary <- summary(hier_modeld, pars = c("delta3"), probs = c(0.1, 0.9))$summary
+  
+  delta4_summary <- summary(hier_modeld, pars = c("delta4"), probs = c(0.1, 0.9))$summary
+  
+  parest_m7 <- rbind(delta1_summary, delta2_summary, delta3_summary, delta4_summary)
+  datafr_m7 <- data.frame(parameter_m7, parest_m7)
+  
+  ## Plot 
+  set.seed(321)
[…]

### 2022-06-09 R code/Figures/Fig11b_modelestimatesfigure_trunc_v2.R

@@ -1,0 +1,180 @@
+# Model estimates figure 
+# Load fits 
+# For rstan the output is saved in hier_modelc_adj_07012022 (adjust "myPath" below to your directory)
+hier_modelc_adj_07012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
+
+# For brms the output is saved in object brmmodel8_new03012022 (this is without saving the prior samples)
+brmmodel8_new03012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/brms/Posterior/brmmodel8_new03012022.rds")  
+
+library(rstan)
+library(brms)
+
+# Plot results using https://github.com/jorvlan/open-visualizations/blob/master/R/repmes_tutorial_R.pdf
+# Install packages for plot 
+packages <- c("plyr", "lattice", "ggplot2", "dplyr", "readr", "rmarkdown","Rmisc")
+if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
+  install.packages(setdiff(packages, rownames(installed.packages())))  
+}
+if (!require(devtools)) {
+  install.packages("devtools")
+}
+devtools::install_github('erocoar/gghalves')
+
+library("plyr")
+library("lattice")
+library("ggplot2")
+library("dplyr")
+library("readr")
+library("rmarkdown")
+library("Rmisc")
+library("devtools")
+library("gghalves")
+
+indvestviolin <- function(hier_modelc_adj_v2, brmmodel8_v2){
+  # Rstan 
+  ## Delta's 
+  ## Create dataframe 
+  parameter_m7 <- rep(c("Delta 1", "Delta 2", "Delta 3", "Delta 4"), each = 52 )
+  
+  ### Obtain all individual estimates delta 
+  hier_modeld <- hier_modelc_adj_v2
+  delta1_summary <- summary(hier_modeld, pars = c("delta1"), probs = c(0.1, 0.9))$summary
+  
+  delta2_summary <- summary(hier_modeld, pars = c("delta2"), probs = c(0.1, 0.9))$summary
+  
[…]

### 2022-06-09 R code/Figures/Fig12_chainrhatneff_v2.R

@@ -1,0 +1,287 @@
+# Figure 18 + 19 (log models)
+# Traceplots dif 2 (difference between 6 and 7) + rhat
+# Based on output each package of symbolic distance effect model 
+# Load packages
+library(brms)
+library(rstan)
+
+# Load data 
+# For rstan the output is saved in hier_logmodelc_adj13012022  (adjust "myPath" below to your directory)
+hier_logmodelc_adj13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/rstan/log-normal model/hier_logmodelc_adj13012022.rds")
+
+# For brms the output is saved in object brmmodel8log13012022 (this is without saving the prior samples)
+brmmodel8log13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/brms/log-normal/brmmodel8log13012022.rds")  
+
+tracerhatnefflog <- function(hier_logmodelc_adj30032021, brmmodel8log02042021){
+#Rstan
+trdif2rstan <- plot(hier_logmodelc_adj30032021, plotfun = "trace", pars = c("mu4"))  # rstan 
+# plot(hier_modelc_adj, plotfun = "trace", pars = c("mu4"), inc_warmup = TRUE)  # show warmup -- > ugly plot 
+
+# Extract legend 
+library(cowplot)
+legendtrace <- cowplot::get_legend(trdif2rstan)
+library(grid)
+grid.draw(legendtrace)
+
+# Obtain colors from rstanplot 
+unique(ggplot_build(trdif2rstan)$data[[1]]$colour)
+# "#E66101" "#998EC3" "#542788" "#F1A340"
+
+# Final rstan plot 
+trdif2rstan2 <- trdif2rstan + ylab(expression(~mu[~delta[~6]])) + 
+  theme(axis.title.y=element_text(angle = 0, 
+                                  vjust = 0.5, 
+                                  hjust = -0.5, 
+                                  margin = margin(0, 0.7, 0, 0, "cm"), size = 14), 
+        legend.position = "none", 
+        axis.line.x  = element_line(size = 0.6), 
+        axis.line.y  = element_line(size = 0.6), 
+        plot.margin = unit(c(0.3,0.1,0.3,0.05),"cm"), 
[…]

### 2022-06-09 R code/Figures/Fig13_posteriorlog_v2.R

@@ -1,0 +1,80 @@
+# Figure 13
+# Posterior distribution parameters - lognormal model 
+# For rstan the output is saved in hier_logmodelc_adj13012022 (adjust "myPath" below to your directory)
+hier_logmodelc_adj13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/rstan/log-normal model/hier_logmodelc_adj13012022.rds")
+
+# For brms the output is saved in object brmmodel8log13012022 (this is without saving the prior samples)
+brmmodel8log13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/brms/log-normal/brmmodel8log13012022.rds")  
+
+
+# Used this link: https://cran.r-project.org/web/packages/bayesplot/vignettes/plotting-mcmc-draws.html 
+library(tidybayes)
+library(bayesplot)
+library(brms)
+library(rstan)
+
+posteriorlogfunc <- function(hier_logmodelc_adj30032021, brmmodel8log02042021){
+  post8 <- as.array(brmmodel8log02042021)
+  dimnames(post8)
+  color_scheme_set("orange")
+  pairs8 <- mcmc_areas(
+    post8, 
+    pars = c("b_side", "b_dif1", "b_dif2", "b_dif3", "b_dif4"),
+    prob = 0.95, # 95% intervals
+    prob_outer = 0.99, # 99%
+    point_est = "mean"
+  )
+  # scale_x_continuous(breaks=c(1,2,3,4), labels=c(expression(~delta[~7-8]), expression(~delta[~6-7]), expression(~delta[~4-3]), expression(~delta[~3-2]))) +
+  
+  pairs82 <- pairs8 + 
+    theme_apa() + 
+    ggtitle("Brms") + 
+    xlim(-0.035, 0.085) + 
+    scale_x_continuous(breaks = c(-0.025, 0, 0.025, 0.05, 0.075, 0.1)) + 
+    scale_y_discrete(labels= c( 
+      expression(~mu[~beta]), 
+      expression(~mu[~delta[~7]]), 
+      expression(~mu[~delta[~6]]), 
+      expression(~mu[~delta[~4]]), 
+      expression(~mu[~delta[~3]]))) + 
+    theme(axis.title.y=element_blank(),
[…]

### 2022-06-09 R code/Figures/Fig14_randomparlog_v2.R

@@ -1,0 +1,364 @@
+# Figure 14 
+# Individual effects from Rstan 
+# Load data 
+# For rstan the output is saved in hier_logmodelc_adj13012022 (adjust "myPath" below to your directory)
+hier_logmodelc_adj13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/rstan/log-normal model/hier_logmodelc_adj13012022.rds")
+
+# For brms the output is saved in object brmmodel8log13012022 (this is without saving the prior samples)
+brmmodel8log13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/brms/log-normal/brmmodel8log13012022.rds")  
+
+library(rstan)
+library(brms)
+
+randomparlogfunc <- function(hier_logmodelc_adj30032021, brmmodel8log02042021){
+  # Documentation on how to retrieve objects from stan output: https://cran.r-project.org/web/packages/rstan/vignettes/stanfit-objects.html
+  theme_set(theme_apa(base_size = 9))
+  # Only gamma's 
+  gamnamesstan <- paste("gammai", "[",1:52,"]", sep="")
+  gamstan <- as.data.frame(summary(hier_logmodelc_adj30032021, pars = c( gamnamesstan))$summary)
+  gamstanmu <- as.data.frame(summary(hier_logmodelc_adj30032021, pars = c("mu"))$summary)
+  gamstan$parameters <- c(gamnamesstan)
+  
+  for (i in 1:nrow(gamstan)){  # 1 if CI contains zero, 0 otherwise 
+    if (gamstan$`2.5%`[i]*gamstan$`97.5%`[i] < 0) {
+      gamstan$zero2[i] <- "pink"
+    } else (gamstan$zero2[i] <- "blue") 
+  }
+  
+  
+  gamstan2 <- gamstan[order(gamstan$mean),]
+  gamstan2$order <- c(1:52)
+  
+  gammarandomplot <- ggplot(gamstan2, aes(x = order, y = mean, ymin = `2.5%`, ymax = `97.5%`)) + 
+    geom_linerange(aes(color = "blue"), color = "#00BFC4") + 
+    geom_point() + 
[…]

### 2022-06-09 R code/Figures/Fig15_lognormaldifindest_onlyrstan_v2.R

@@ -1,0 +1,104 @@
+# Figure 15 
+
+# Load data 
+# For rstan the output is saved in hier_logmodelc_adj13012022 (adjust "myPath" below to your directory)
+hier_logmodelc_adj13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/rstan/log-normal model/hier_logmodelc_adj13012022.rds")
+
+# For brms the output is saved in object brmmodel8log13012022 (this is without saving the prior samples)
+brmmodel8log13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/brms/log-normal/brmmodel8log13012022.rds")  
+
+library(brms)
+library(rstan)
+
+# Plot results using https://github.com/jorvlan/open-visualizations/blob/master/R/repmes_tutorial_R.pdf
+# Install packages for plot 
+packages <- c("plyr", "lattice", "ggplot2", "dplyr", "readr", "rmarkdown","Rmisc")
+if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
+  install.packages(setdiff(packages, rownames(installed.packages())))  
+}
+if (!require(devtools)) {
+  install.packages("devtools")
+}
+devtools::install_github('erocoar/gghalves')
+
+library("plyr")
+library("lattice")
+library("ggplot2")
+library("dplyr")
+library("readr")
+library("rmarkdown")
+library("Rmisc")
+library("devtools")
+library("gghalves")
+
+lognormaldicfindestfunc_onlyrstan <- function(hier_logmodelc_adj30032021, brmmodel8log02042021){
+  # Rstan 
+  parameter_m7 <- rep(c("Delta 1", "Delta 2", "Delta 3", "Delta 4"), each = 52 )
+  
+  # Obtain all individual estimates delta 
+  hier_modeld <- hier_logmodelc_adj30032021
+  delta1_summary <- summary(hier_modeld, pars = c("delta1"), probs = c(0.1, 0.9))$summary
+  
+  delta2_summary <- summary(hier_modeld, pars = c("delta2"), probs = c(0.1, 0.9))$summary
+  
+  delta3_summary <- summary(hier_modeld, pars = c("delta3"), probs = c(0.1, 0.9))$summary
+  
[…]

### 2022-06-09 R code/Figures/Fig15b_lognormaldifindest_v2.R

@@ -1,0 +1,176 @@
+# Figure 15b (not in manuscript) 
+
+# Load data 
+# For rstan the output is saved in hier_logmodelc_adj13012022 (adjust "myPath" below to your directory)
+hier_logmodelc_adj13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/rstan/log-normal model/hier_logmodelc_adj13012022.rds")
+
+# For brms the output is saved in object brmmodel8log13012022 (this is without saving the prior samples)
+brmmodel8log13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/brms/log-normal/brmmodel8log13012022.rds")  
+
+library(brms)
+library(rstan)
+
+# Plot results using https://github.com/jorvlan/open-visualizations/blob/master/R/repmes_tutorial_R.pdf
+# Install packages for plot 
+packages <- c("plyr", "lattice", "ggplot2", "dplyr", "readr", "rmarkdown","Rmisc")
+if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
+  install.packages(setdiff(packages, rownames(installed.packages())))  
+}
+if (!require(devtools)) {
+  install.packages("devtools")
+}
+devtools::install_github('erocoar/gghalves')
+
+library("plyr")
+library("lattice")
+library("ggplot2")
+library("dplyr")
+library("readr")
+library("rmarkdown")
+library("Rmisc")
+library("devtools")
+library("gghalves")
+
+lognormaldicfindestfunc <- function(hier_logmodelc_adj30032021, brmmodel8log02042021){
+  # Rstan 
+  parameter_m7 <- rep(c("Delta 1", "Delta 2", "Delta 3", "Delta 4"), each = 52 )
+  
+  # Obtain all individual estimates delta 
+  hier_modeld <- hier_logmodelc_adj30032021
+  delta1_summary <- summary(hier_modeld, pars = c("delta1"), probs = c(0.1, 0.9))$summary
+  
+  delta2_summary <- summary(hier_modeld, pars = c("delta2"), probs = c(0.1, 0.9))$summary
+  
[…]

### 2022-06-09 R code/Figures/Fig16_savagedickey_trunc.R

@@ -1,0 +1,191 @@
+# Load packages
+library(ggridges)
+library(brms)
+
+# Load brms output 
+# For brms the output is saved in object brmmodel8_new_bf_03012022 (this is without saving the prior samples)
+brmmodel8_new_bf_03012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/brms/BF/brmmodel8_new_bf_03012022.rds") 
+
+savagedickeyfigure <- function(brmmodel8bf){
+  # Obtain Savage-Dickey ratio 
+  # https://vuorre.netlify.app/post/2017/03/21/bayes-factors-with-brms/ 
+  # Beta 
+  b1brms <- hypothesis(brmmodel8bf, "a_side = 0")  # Evid.Ratio is Bayes factor BF01, have to set sample_prior = TRUE when fitting model 
+  b1brms$hypothesis$Evid.Ratio 
+  plotsdratiob1 <- plot(b1brms,plot = F, theme = theme_get())[[1]] 
+  plotsdratiob1 + xlim(-0.07, 0.07) 
+  
+  # Plot function 
+  savdicratioplot <- function(plotsdratio){  # put in plot.hypothesisbrms object 
+    plotsratio3data <- plotsdratio$data
+    
+    density(plotsratio3data$values[plotsratio3data$Type == "Posterior"])
+    xpriorsd <- c(density(plotsratio3data$values[plotsratio3data$Type == "Posterior"])[1]$x, density(plotsratio3data$values[plotsratio3data$Type == "Prior"])[1]$x)  # x 
+    ypriorsd <- c(density(plotsratio3data$values[plotsratio3data$Type == "Posterior"])[2]$y, density(plotsratio3data$values[plotsratio3data$Type == "Prior"])[2]$y)  # y 
+    xypriorsd <- data.frame(x = xpriorsd, y = ypriorsd, Type = rep(c("Posterior", "Prior"), each = 512))
+    maxprior <- max(density(plotsratio3data$values[plotsratio3data$Type == "Prior"])[2]$y)  
+    
+    
+    sdrdif3plot2 <- ggplot(xypriorsd, aes(x, y, color = Type, fill = Type)) + 
+      geom_line() + 
+      geom_area(position = "identity", alpha = 0.3) + 
[…]

### 2022-06-09 R code/Figures/Fig17_bridgesamplingresults_v2.R

@@ -1,0 +1,76 @@
+# BF rstan bridge-sampling 
+library(ggplot2)
+library(brms)
+library(rstan )
+
+# Load results
+# rstan
+resultsbsrstan <- read.csv("/myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/Model comparison/Bayes Factors per package/bssresults_07012022.csv")
+
+## Restructure 
+resultsbsrstan$package <- rep("Rstan", nrow(resultsbsrstan))
+resultsbsrstan <- data.frame(Fit4 = resultsbsrstan$Fit4, model = resultsbsrstan$model, package = resultsbsrstan$package)
+
+# brms 
+resultsbsbrms <- read.csv("/myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/Model comparison/Bayes Factors per package/bssresults_brms_15032022_10rep.csv")
+
+## Restructure 
+resultsbsbrms$package <- rep("Brms", nrow(resultsbsbrms))
+resultsbsbrms <- data.frame(Fit4 = resultsbsbrms$Fit1, model = resultsbsbrms$model, package = resultsbsbrms$package)
+
+# Combine rstan and brms results 
+resultsbf10 <- rbind(resultsbsrstan, resultsbsbrms)
+
+resultsbf10$BF01 <- 1/resultsbf10$Fit4
+
+level_order <- c('Null model', 'Side model', 'Digit model') 
+
+# Evidence of full model against alternative models 
+normalbfplot <- ggplot(resultsbf10[which(resultsbf10$model == "Null model" | 
+                                           resultsbf10$model == "Side model" |  
+                                           resultsbf10$model == "Digit model"),],
+                       aes(x = factor(model, level_order), y = Fit4, color = 
+                             package, shape = package, fill = package)) + 
+  geom_jitter(alpha = 0.5, position = position_jitterdodge(dodge.width = 0.5)) + 
+  # geom_violin(trim = FALSE) + 
+  scale_y_log10() + 
+  theme_classic() + 
+  labs(subtitle = "Linear model \n", 
+       x = "", y = "BF01") + 
+  scale_color_manual(name = "Package", 
+                     labels = c("Rstan", "Brms"),
+                     values = c("darkgreen", "darkorange")) + 
+  scale_shape_manual(name = "Package", 
+                     labels = c("Rstan", "Brms"),
+                     values = c(17, 19)) + 
[…]

### 2022-06-09 R code/Figures/Fig8_traceplotsfigure_trunc_v2.R

@@ -1,0 +1,438 @@
+# Traceplots dif 2 (difference between 6 and 7) + rhat
+# Based on output each package of symbolic distance effect model 
+# For rstan the output is saved in hier_modelc_adj_07012022 (adjust "myPath" below to your directory)
+hier_modelc_adj_07012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
+
+# For brms the output is saved in object brmmodel8_new03012022 (this is without saving the prior samples)
+brmmodel8_new03012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/brms/Posterior/brmmodel8_new03012022.rds")  
+
+tracerhatneff <- function(hier_modelc_adj_v2, brmmodel8_v2){
+  #Rstan
+  library(rstan)
+  trdif2rstan <- plot(hier_modelc_adj_v2, plotfun = "trace", pars = c("mu4"))  # rstan 
+  # plot(hier_modelc_adj, plotfun = "trace", pars = c("mu4"), inc_warmup = TRUE)  # show warmup -- > ugly plot 
+  
+  trdif2rstan <- trdif2rstan + labs(color='Chain') 
+  
+  # Extract legend 
+  library(cowplot)
+  legendtrace <- cowplot::get_legend(trdif2rstan)
+  library(grid)
+  grid.draw(legendtrace)
+  
+  # Obtain colors from rstanplot 
+  unique(ggplot_build(trdif2rstan)$data[[1]]$colour)
+  # "#E66101" "#998EC3" "#542788" "#F1A340"
+  
+  # Final rstan plot 
+  trdif2rstan2 <- trdif2rstan + ylab(expression(~mu[~delta[~6]])) + 
+    theme(axis.title.y=element_text(angle = 0, 
+                                    vjust = 0.5, 
+                                    hjust = -0.5, 
+                                    margin = margin(0, 0.7, 0, 0, "cm"), size = 14), 
+          legend.position = "none", 
+          axis.line.x  = element_line(size = 0.6), 
+          axis.line.y  = element_line(size = 0.6), 
+          plot.margin = unit(c(0.3,0.1,0.3,0.05),"cm"), 
[…]

### 2022-06-09 R code/Figures/Fig9_posteriorfigure_trunc_v2.R

@@ -1,0 +1,81 @@
+# Posterior distribution parameters 
+# For rstan the output is saved in hier_modelc_adj_07012022 (adjust "myPath" below to your directory)
+hier_modelc_adj_07012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
+
+# For brms the output is saved in object brmmodel8_new03012022 (this is without saving the prior samples)
+brmmodel8_new03012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/brms/Posterior/brmmodel8_new03012022.rds")  
+
+# Used this link: https://cran.r-project.org/web/packages/bayesplot/vignettes/plotting-mcmc-draws.html 
+
+posterplot <- function(hier_modelc_adj_v2, brmmodel8_v2){
+  library(tidybayes)
+  library(bayesplot)
+  library(rstan)
+  library(brms)
+  library(ggplot2)
+  library(papaja)
+  
+  post8 <- as.array(brmmodel8_v2)
+  dimnames(post8)
+  color_scheme_set("orange")
+  pairs8 <- mcmc_areas(
+    post8, 
+    pars = c("b_a_side", "b_a_dif1", "b_a_dif2", "b_a_dif3", "b_a_dif4"),
+    prob = 0.95, # 80% intervals
+    prob_outer = 0.99, # 99%
+    point_est = "mean"
+  )
+  # scale_x_continuous(breaks=c(1,2,3,4), labels=c(expression(~delta[~7-8]), expression(~delta[~6-7]), expression(~delta[~4-3]), expression(~delta[~3-2]))) +
+  
+  pairs82 <- pairs8 + 
+    theme_apa() + 
+    ggtitle("Brms") + 
+    xlim(-0.035, 0.085) + 
+    scale_x_continuous(breaks = c(-0.025, 0, 0.025, 0.05, 0.075)) + 
+    scale_y_discrete(labels= c( 
+      expression(~mu[~beta]), 
+      expression(~mu[~delta[~7]]), 
+      expression(~mu[~delta[~6]]), 
+      expression(~mu[~delta[~4]]), 
+      expression(~mu[~delta[~3]]))) + 
+    theme(axis.title.y=element_blank(),
[…]

### 2022-06-09 R code/Model Comparison /Sensitivity analysis/modelcomparison_sensitivityanalysis_BF.R

@@ -1,0 +1,406 @@
+# Sensitivity analysis 
+# Based on rstan results 
+library(plyr)
+library(rstan)
+library(ggplot2)
+
+# Preparations 
+# Load datareal 
+indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
+colnames(indat)=c('sub','block','trial','stim','resp','rt','error')
+
+## Cleaning the data according to criteria discussed in paper 
+clean=function()
+{
+  indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
+  colnames(indat)=c('sub','block','trial','stim','resp','rt','error')
+  
+  bad1=indat$sub%in%c(34,43)
+  bad2=indat$rt<250 | indat$rt>2000
+  bad3=indat$err==1
+  bad4=indat$block==0 & indat$trial<20
+  bad5=indat$trial==0
+  
+  bad=bad1 | bad2 | bad3 |bad4 |bad5
+  dat=indat[!bad,]
+  return(dat)
+}
+
+indat1 <- clean()  # Final dataset 
+
+## Data into list and fill the list with necessary information 
+datareal <- list(N = length(unique(indat1$sub)))
+datareal$All <- nrow(indat1)  # number of total observations, adding to the list for the analysis  
+
+### Indicating which observation belongs to which person 
+rank(unique(indat1$sub))
+sub <- rep(1:N, each = J)
+true.theta[sub]
+
+for (j in 1:nrow(indat1)){
+  if (j == 1) {
+    indat1$sub1[j] <- 1}
+  else if (indat1$sub[j] == indat1$sub[j-1]) {
+    indat1$sub1[j] <- indat1$sub1[j-1]}
+  else indat1$sub1[j] <- indat1$sub1[j-1] + 1
+}
+
+
+datareal$group_inds <- indat1$sub1  # Add variable with indicator person to the list used for the analysi
+datareal$y <- indat1$rt/1000  # rt in seconds instead of miliseconds, added to the list for analysis  
+
+### Add variable with information of the side of the digit to the list, for parameter beta    
+# j = condition, if j > 5 than x = -1/2, if j < 5 than x = 1/2 
+# indat1$stim 0 = 2, 1=3, 2=4, 3=6, 4=7, 5=8 
[…]

### 2022-06-09 R code/Model Comparison /brms/BF_bridgesampling_brms_trunc_v2.R

@@ -1,0 +1,303 @@
+# BF bridge-sampling brms 
+
+# https://discourse.mc-stan.org/t/bayes-factor-using-brms/4469
+# http://paul-buerkner.github.io/brms/reference/bayes_factor.brmsfit.html 
+
+
+library(brms)
+
+################################################################################
+## 1. Load data 
+################################################################################
+indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
+colnames(indat)=c('sub','block','trial','stim','resp','rt','error')
+
+## Cleaning the data according to criteria discussed in paper 
+clean=function()
+{
+  indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
+  colnames(indat)=c('sub','block','trial','stim','resp','rt','error')
+  
+  bad1=indat$sub%in%c(34,43)
+  bad2=indat$rt<250 | indat$rt>2000
+  bad3=indat$err==1
+  bad4=indat$block==0 & indat$trial<20
+  bad5=indat$trial==0
+  
+  bad=bad1 | bad2 | bad3 |bad4 |bad5
+  dat=indat[!bad,]
+  return(dat)
+}
+
+indat1 <- clean()  # Final dataset 
+
+## Data add necessary information 
+### Add participant number 
+sub <- rep(1:N, each = J)
+
+indat1$inter <- rep(1, nrow(indat1))
+
+for (j in 1:nrow(indat1)){
+  if (j == 1) {
+    indat1$sub1[j] <- 1}
+  else if (indat1$sub[j] == indat1$sub[j-1]) {
+    indat1$sub1[j] <- indat1$sub1[j-1]}
+  else indat1$sub1[j] <- indat1$sub1[j-1] + 1
+}
+
+### Add variable with information of the side of the digit to the list, for parameter beta    
+# j = condition, if j > 5 than x = -1/2, if j < 5 than x = 1/2 
+# indat1$stim 0 = 2, 1=3, 2=4, 3=6, 4=7, 5=8 
+# so if bigger than 2 than 1/2, else than -1/2 
+for (j in 1:nrow(indat1)){
+  if (indat1$stim[j] < 3) {
+    indat1$side[j] <- 1/2}
+  else indat1$side[j] <- -1/2
+}
+
[…]

### 2022-06-09 R code/Model Comparison /rstan/Bayesfactorhiersmodelsbf_trunc.R

@@ -1,0 +1,282 @@
+# Bayes factor estimation bridge sampling 
+library(plyr)
+library(rstan)
+library(ggplot2)
+# Load datareal 
+indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
+colnames(indat)=c('sub','block','trial','stim','resp','rt','error')
+
+## Cleaning the data according to criteria discussed in paper 
+clean=function()
+{
+  indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
+  colnames(indat)=c('sub','block','trial','stim','resp','rt','error')
+  
+  bad1=indat$sub%in%c(34,43)
+  bad2=indat$rt<250 | indat$rt>2000
+  bad3=indat$err==1
+  bad4=indat$block==0 & indat$trial<20
+  bad5=indat$trial==0
+  
+  bad=bad1 | bad2 | bad3 |bad4 |bad5
+  dat=indat[!bad,]
+  return(dat)
+}
+
+indat1 <- clean()  # Final dataset 
+
+## Data into list and fill the list with necessary information 
+datareal <- list(N = length(unique(indat1$sub)))
+datareal$All <- nrow(indat1)  # number of total observations, adding to the list for the analysis  
+
+### Indicating which observation belongs to which person 
+rank(unique(indat1$sub))
+sub <- rep(1:N, each = J)
+true.theta[sub]
+
+for (j in 1:nrow(indat1)){
+  if (j == 1) {
+    indat1$sub1[j] <- 1}
+  else if (indat1$sub[j] == indat1$sub[j-1]) {
+    indat1$sub1[j] <- indat1$sub1[j-1]}
+  else indat1$sub1[j] <- indat1$sub1[j-1] + 1
+}
+
+
+datareal$group_inds <- indat1$sub1  # Add variable with indicator person to the list used for the analysi
+datareal$y <- indat1$rt/1000  # rt in seconds instead of miliseconds, added to the list for analysis  
+
+### Add variable with information of the side of the digit to the list, for parameter beta    
+# j = condition, if j > 5 than x = -1/2, if j < 5 than x = 1/2 
[…]

### 2022-06-09 R code/Prior transformation/Prior_Transformation 2.R

@@ -1,0 +1,302 @@
+# Inverse-gamma prior on standard deviation:
+
+alpha = 1.745
+beta = 0.02
+
+dinvg <- function(x, alpha, beta){
+  (beta^alpha)/gamma(alpha)*x^(-alpha-1)*exp(-beta*x^(-1))
+}
+
+# Sanity check: This gives the same results as LaplacesDemon::dinvgamma()
+
+dinvg(0.3, alpha, beta)
+dinvgamma(0.3, alpha, beta)
+
+# This prior implies for the variance (replace x by sqrt(x) in inverse gamma equation)
+dvar <- function(x, alpha, beta){
+  (beta^alpha)/gamma(alpha)*x^(-0.5*alpha-0.5)*exp(-beta*x^(-0.5))  
+}
+
+dvar1 <- function(x, alpha, beta){  # same as dvar 
+  (beta^alpha)/gamma(alpha)*(1/sqrt(x))^(alpha+1)*exp(-beta/sqrt(x))
+}
+
+# The implied distribution of the variance is not an inverse gamma distribution (?)
+# However, different inverse gamma distributions can be more or less similar
+
+x <- seq(0, 1, by=0.001)
+plot(x, dvar1(x, alpha, beta), type="l") # implied distribution on variance
+points(x, dinvg(x, 3, 0.3), type="l", col="red") # red distribution: prior on variance used in Stan code
+points(x, dinvg(x, 3, 0.0006), type="l", col="blue") # blue distribution: "new" prior on variance (see Myrthe's e-mail) that produces more similar results to brms
+
+# We can also do this the other way around and start with the prior on variance
+alpha = 3
+beta = 0.3
+
+dinvg <- function(x, alpha, beta){
+  (beta^alpha)/gamma(alpha)*x^(-alpha-1)*exp(-beta*x^(-1))
+} # this is the same as before, but now the inverse gamma prior is on the variance, so x is the variance here
+
+# This inverse-gamma prior on variance implies the following prior on standard deviation
+
+dsd <- function(x, alpha, beta){
+  (beta^alpha)/gamma(alpha)*x^(-2*alpha-2)*exp(-beta*x^(-2))
+}
+
+# Again, the implied distribution on the standard deviation is not an inverse gamma,
+# but inverse gamma distributions can be more or less similar to the implied prior
+
[…]

### 2022-06-09 R code/Prior transformation/priortransformFunction.R

@@ -1,0 +1,44 @@
+#' Function to transform priors on variance to priors on standard deviation and vice-versa
+#' @param shape Shape parameter (alpha) of original inverse-gamma distribution
+#' @param scale Scale parameter (beta) of original inverse-gamma distribution
+#' @param direction Either "VarToSD" or "SDtoVar" depending on what the original prior is
+#' @importFrom LaplacesDemon dnivgamma
+#' @importFrom stats
+
+priortransform <- function(shape, scale, direction){
+
+  stopifnot("direction must be 'VarToSD' or 'SDtoVar'" = direction %in% c("VarToSD", "SDtoVar"))
+
+  dvar <- function(x, alpha, beta){
+    (beta^alpha)/gamma(alpha)*x^(-0.5*alpha-0.5)*exp(-beta*x^(-0.5))
+  }
+
+  dsd <- function(x, alpha, beta){
+    (beta^alpha)/gamma(alpha)*x^(-2*alpha-2)*exp(-beta*x^(-2))
+  }
+
+  if(direction=="VarToSD"){
+    impliedprior <- function(x) dsd(x, alpha=shape, beta=scale)
+    meantrans <- sqrt(scale/(shape-1))
+  } else if (direction=="SDtoVar") {
+    impliedprior <- function(x) dvar(x, alpha=shape, beta=scale)
+    meantrans <- (scale/(shape-1))^2
+  }
+
+  impliedprior_normalized <- function(x) impliedprior(x)/integrate(impliedprior, lower=0, upper=Inf)$value
+
+  minfun <- function(par){
+    x <- seq(0.0001, meantrans*3, length.out = 10000)
+    s <- sum((dinvgamma(x, par[1], par[2])-impliedprior_normalized(x))^2) # minimize sum of squares
+    if(is.nan(s)) return(100000)
+    return(s)
+  }
+
+  opt <- stats::optim(par=c(2, meantrans), fn=minfun, lower=c(1e-10, 1e-10), method="L-BFGS-B")
+
+  return(list(shape_original = shape,
+              scale_original = scale,
+              shape_new = opt$par[1],
+              scale_new = opt$par[2]))
+
+}

### 2022-06-09 R code/Rstan/Linear model/BHM in rstan_cleaned.R

@@ -1,0 +1,148 @@
+# Fitting an Bayesian Hierarchical Model using Stan - An example 
+
+################################################################################
+## 1. Load data 
+################################################################################
+indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
+colnames(indat)=c('sub','block','trial','stim','resp','rt','error')
+
+## Cleaning the data according to criteria discussed in paper 
+clean=function()
+{
+  indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
+  colnames(indat)=c('sub','block','trial','stim','resp','rt','error')
+  
+  bad1=indat$sub%in%c(34,43)
+  bad2=indat$rt<250 | indat$rt>2000
+  bad3=indat$err==1
+  bad4=indat$block==0 & indat$trial<20
+  bad5=indat$trial==0
+  
+  bad=bad1 | bad2 | bad3 |bad4 |bad5
+  dat=indat[!bad,]
+  return(dat)
+}
+
+indat1 <- clean()  # Final dataset 
+
+## Data into list and fill the list with necessary information 
+datareal <- list(N = length(unique(indat1$sub)))
+datareal$All <- nrow(indat1)  # number of total observations, adding to the list for the analysis  
+
+### Indicating which observation belongs to which person 
+sub <- rep(1:N, each = J)
+
+# Need a variable with group number that are crhonical (that doesn't skip numbers)
+for (j in 1:nrow(indat1)){
+  if (j == 1) {
+    indat1$sub1[j] <- 1}
+  else if (indat1$sub[j] == indat1$sub[j-1]) {
+    indat1$sub1[j] <- indat1$sub1[j-1]}
+  else indat1$sub1[j] <- indat1$sub1[j-1] + 1
+}
+
+
+datareal$group_inds <- indat1$sub1  # Add variable with indicator person to the list used for the analysi
+datareal$y <- indat1$rt/1000  # rt in seconds instead of miliseconds, added to the list for analysis  
+
+### Add variable with information of the side of the digit to the list, for parameter beta    
[…]

### 2022-06-09 R code/Rstan/Non-linear model/rstan_lognormal_cleaned.R

@@ -1,0 +1,148 @@
+# Lognormal rstan cleaned 
+# Fitting an Bayesian Hierarchical Model using Stan - An example 
+
+################################################################################
+## 1. Load data 
+################################################################################
+indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
+colnames(indat)=c('sub','block','trial','stim','resp','rt','error')
+
+## Cleaning the data according to criteria discussed in paper 
+clean=function()
+{
+  indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
+  colnames(indat)=c('sub','block','trial','stim','resp','rt','error')
+  
+  bad1=indat$sub%in%c(34,43)
+  bad2=indat$rt<250 | indat$rt>2000
+  bad3=indat$err==1
+  bad4=indat$block==0 & indat$trial<20
+  bad5=indat$trial==0
+  
+  bad=bad1 | bad2 | bad3 |bad4 |bad5
+  dat=indat[!bad,]
+  return(dat)
+}
+
+indat1 <- clean()  # Final dataset 
+
+## Data into list and fill the list with necessary information 
+datareal <- list(N = length(unique(indat1$sub)))
+datareal$All <- nrow(indat1)  # number of total observations, adding to the list for the analysis  
+
+### Indicating which observation belongs to which person 
+sub <- rep(1:N, each = J)
+
+# Need a variable with group number that are crhonical (that doesn't skip numbers)
+for (j in 1:nrow(indat1)){
+  if (j == 1) {
+    indat1$sub1[j] <- 1}
+  else if (indat1$sub[j] == indat1$sub[j-1]) {
+    indat1$sub1[j] <- indat1$sub1[j-1]}
+  else indat1$sub1[j] <- indat1$sub1[j-1] + 1
+}
+
+
+datareal$group_inds <- indat1$sub1  # Add variable with indicator person to the list used for the analysi
+datareal$y <- indat1$rt/1000  # rt in seconds instead of miliseconds, added to the list for analysis  
+
[…]

### 2022-06-10 D - Tutorial Normal Model brms/Online Supplement D - Bayesian Hierarchical Modeling in brms.Rmd
Updates

@@
 
-We have already estimated the model and saved it in the R object `brmsmodelfit_final`.
+We have already estimated the model and saved it in the R object `brmmodel8_new03012022.rds`.
 In case you do not want to wait until the model has been fitted but continue immediately with the rest of this tutorial, you can load this R object that contains the model estimations.
@@
 ```{r loadmodelfit, eval = FALSE}
-model_fit <- readRDS(file = "myPath/brmsmodelfit_final.rds")
+model_fit <- readRDS(file = "myPath/brmmodel8_new03012022.rds")
 
 # For model comparison - sample_prior was set to "yes" 
-model_fit_bf <- readRDS("myPath/brmsmodelfit_BF_final.rds")
-
-```
-\normalsize
-
-\scriptsize 
-```{r loadmodelfits, show = FALSE}
-model_fit <- readRDS(file = "/Users/myrtheveenman/Documents/GitHub/multilevel-tutorial-paper/R objects/brms/Truncated/BRMSmodelfit18_03012022.rds")
+model_fit_bf <- readRDS("myPath/brmmodel8_new_bf_03012022l.rds")
+
+```
+\normalsize
+
+\scriptsize 
+```{r loadmodelfits, show = FALSE, echo = FALSE}
+setwd('..')
+model_fit <- readRDS(file = "R objects/brms/Truncated/BRMSmodelfit18_03012022.rds")
 
 # For model comparison - sample_prior was set to "yes" 
-model_fit_bf <- readRDS("/Users/myrtheveenman/Documents/GitHub/multilevel-tutorial-paper/R objects/brms/Truncated/BRMSmodelfit18_BF_03012022.rds")
+model_fit_bf <- readRDS("R objects/brms/Truncated/BRMSmodelfit18_BF_03012022.rds")
 

### 2022-06-10 F - Tutorial Log-Normal Model brms/Online Supplement F - Bayesian Hierarchical Modeling in brms - Log Model.Rmd
Updates

@@
 
-We have already estimated the model and saved it in the R object `brmslogmodelfit_final`.
+We have already estimated the model and saved it in the R object `brmmodel8log13012022.rds`.
 In case you do not want to wait until the model has been fitted but continue immediately with the rest of this tutorial, you can load this R object that contains the model estimations.
@@
 ```{r loadmodelfit, eval = FALSE}
-logmodel_fit <- readRDS(file = "myPath/brmslogmodelfit_final.rds")
+logmodel_fit <- readRDS(file = "myPath/brmmodel8log13012022.rds")
 
 # For model comparison - sample_prior was set to "yes" 
-logmodel_fit_bf <- readRDS("myPath/brmslogmodelfit_BF_final.rds")
-
-```
-\normalsize
-
-\scriptsize 
-```{r loadmodelfits, show = FALSE}
-logmodel_fit <- readRDS(file = "/Users/myrtheveenman/Documents/GitHub/multilevel-tutorial-paper/R objects/brms/non-linear/brmmodel8log13012022.rds")
+logmodel_fit_bf <- readRDS("myPath/BRMSmodelfit18_26052022_bf.rds")
+
+```
+\normalsize
+
+\scriptsize 
+```{r loadmodelfits, show = FALSE, echo = FALSE}
+setwd('..')
+logmodel_fit <- readRDS(file = "R objects/brms/log-normal/brmmodel8log13012022.rds")
 
 # For model comparison - sample_prior was set to "yes" 
-logmodel_fit_bf <- readRDS("/Users/myrtheveenman/Documents/GitHub/multilevel-tutorial-paper/R objects/brms/BF/brmmodel8log26052022_bf.rds")
+logmodel_fit_bf <- readRDS("R objects/brms/BF/brmmodel8log26052022_bf.rds")
 

### 2022-06-29 manuscript.Rmd
Small fixes

@@
 The summaries of the posterior distributions of the variance parameters as estimated by `rstan` are shown in Table \@ref(tab:inddevtable).
-The estimates for the standard deviations by `brms` are smaller compared to `rstan` and can be found in Online Supplement D.
+The estimates for the standard deviations by `brms` are smaller compared to `rstan` and can be found in Online Supplement D and H.
 
@@
 The trace plot of $\mu_{\delta_{6}}$ is shown panel A for `rstan` and B for `brms` of Figure \@ref(fig:traceplotlog).
-Trace plots for all other parameters can be found in the Online Supplement.
+Trace plots for all other parameters can be found in Online Supplement G.
 All plots look like a hairy caterpillar.
@@
 Table \@ref(tab:inddevtablelog) presents the estimates of the variance parameters by `rstan`.
-Results for brms can be found in the Online Supplement.
+Results for brms can be found in Online Supplement H.
 ```{r inddevtablelog, results = "asis", cache = TRUE}
@@
 The estimates for the standard deviations by `brms` are similar compared to the estimates by `rstan`, but slightly smaller.
-The exact estimates by `brms` can be found in Online Supplement F.
+The exact estimates by `brms` can be found in Online Supplement F and H.
 Figure \@ref(fig:randomparplotlog) shows the individual estimates for the digit and side effects.

### 2022-06-29 manuscript.tex
Small fixes

@@
 The summaries of the posterior distributions of the variance parameters as estimated by \texttt{rstan} are shown in Table \ref{tab:inddevtable}.
-The estimates for the standard deviations by \texttt{brms} are smaller compared to \texttt{rstan} and can be found in Online Supplement D.
+The estimates for the standard deviations by \texttt{brms} are smaller compared to \texttt{rstan} and can be found in Online Supplement D and H.
 
@@
 The trace plot of \(\mu_{\delta_{6}}\) is shown panel A for \texttt{rstan} and B for \texttt{brms} of Figure \ref{fig:traceplotlog}.
-Trace plots for all other parameters can be found in the Online Supplement.
+Trace plots for all other parameters can be found in Online Supplement G.
 All plots look like a hairy caterpillar.
@@
 Table \ref{tab:inddevtablelog} presents the estimates of the variance parameters by \texttt{rstan}.
-Results for brms can be found in the Online Supplement.
+Results for brms can be found in Online Supplement H.
 
@@
 The estimates for the standard deviations by \texttt{brms} are similar compared to the estimates by \texttt{rstan}, but slightly smaller.
-The exact estimates by \texttt{brms} can be found in Online Supplement F.
+The exact estimates by \texttt{brms} can be found in Online Supplement F and H.
 Figure \ref{fig:randomparplotlog} shows the individual estimates for the digit and side effects.

### 2022-06-29 C - Tutorial Normal Model Rstan/Online Supplement C - Bayesian Hierarchical Modeling in rstan.Rmd
Improvements

@@
 title             : "Online Supplement C: Bayesian Hierarchical Modeling in rstan"
-author            : "Myrthe Veenman" 
+author            : " " 
 bibliography      : ["r-references.bib"]
@@
 |    This online supplement provides a tutorial on Bayesian Hierarchical models (BHM) in `rstan` [@R-rstan].
-In addition, it contains information about the Number Classification Task that could not be included in the paper.
+In addition, it contains information about the digit classification task that could not be included in the paper.
 Only the fit of the full model will be shown as the fitting the others should follow logically from this tutorial.

### 2022-06-29 manuscript.Rmd
Improvements

@@
 
-```{r figure1, echo=FALSE, fig.align='left', fig.cap="Panel A shows the response times in seconds for two conditions of the number classification task (i.e., digits 3 and 4).
+```{r figure1, echo=FALSE, fig.align='left', fig.cap="Panel A shows the response times in seconds for two conditions of the digit classification task (i.e., digits 3 and 4).
 The grey lines represent the individual effects, and the blue line the aggregated effect.
@@
 If every participant takes part in each condition, the person variability affects all conditions, thus inducing a correlation across conditions.
-For example, participants in a symbolic distance effect experiment repeatedly respond to target digits inducing a correlation between the responses to the target digits, such as between target digit 3 and target digit 4.
+For example, participants in a digit classification task repeatedly respond to target digits inducing a correlation between the responses to the target digits, such as between target digit 3 and target digit 4.
 Therefore, the observations are not independent as we assume if we perform analysis on aggregated data.
@@
 
-```{r hierarchyplot, echo=FALSE, fig.cap="Hierarchical structure of the normal and lognormal models for the number classification task.", fig.align = "left", fig.height = 3, fig.width = 4, out.width="70%"}
+```{r hierarchyplot, echo=FALSE, fig.cap="Hierarchical structure of the normal and lognormal models for the digit classification task.", fig.align = "left", fig.height = 3, fig.width = 4, out.width="70%"}
 knitr::include_graphics("Images/Fig2_Hierarchical structure.png")
@@
 Finally, we can visualize the simulated data.
-Figure \@ref(fig:priorpredfigurepanel) shows prior predictions for the key effects from the number classification task.
+Figure \@ref(fig:priorpredfigurepanel) shows prior predictions for the key effects from the digit classification task.
 For this task, we mainly care about the side effect and digit effects corresponding to the $\beta$s and the $\delta$s from the model.
@@
 \normalsize If you do not specify a prior for a certain parameter, the default prior as explained in the prior section will be applied.
-It is possible to save all priors in an R `data.frame` and provide these in the `brm` function, this and the full code with explanation for fitting the full symbolic distance effect model using `brms` is shown in our [Online Supplement D: *Bayesian Hierarchical Modeling in brms*](https://www.dropbox.com/s/mthphl3ce2kk0xx/Appendix-D--Fit-BHM-in-brms.pdf?dl=0).^[In our example of the Number Classification Task we are applying a truncated normal prior distribution for only one parameter.
+It is possible to save all priors in an R `data.frame` and provide these in the `brm` function, this and the full code with explanation for fitting the full symbolic distance effect model using `brms` is shown in our [Online Supplement D: *Bayesian Hierarchical Modeling in brms*](https://www.dropbox.com/s/mthphl3ce2kk0xx/Appendix-D--Fit-BHM-in-brms.pdf?dl=0).^[In our example of the digit classification task we are applying a truncated normal prior distribution for only one parameter.
 This requires a slightly different `brms` formula and prior specification, explained in the Online Supplement.]
 
-# Estimation of the Number Classification Task
+# Estimation of the Digit Classification Task
 ## Data 
@@
 The data are available under a free license on GitHub [https://github.com/PerceptionCognitionLab/data0/tree/master/lexDec-dist5](https://github.com/PerceptionCognitionLab/data0/tree/master/lexDec-dist5).
-The data set contains observations from 54 participants who took part in a number classification task.
+The data set contains observations from 54 participants who took part in a digit classification task.
 Each participant performed 360 trials in six blocks.

### 2022-06-29 C - Tutorial Normal Model Rstan/Online Supplement C - Bayesian Hierarchical Modeling in rstan.Rmd
Improvements

@@
 - For the document layout: *rmarkdown* [Version 1.16; @R-rmarkdown], *papaja* [Version 0.1.0.9842; @R-papaja], *knitr* [Version 1.25; @R-knitr], *kableExtra* [Version 1.1.0; @R-kableExtra]
-- For data structuring: *LaplacesDemon* [Version 16.1.4; @R-LaplacesDemon], *plyr* [Version 1.8.4; @R-plyr], *dplyr* [Version 0.8.3; @R-dplyr], *readr* [Version 1.3.1; @R-readr]
+- For data structuring: *LaplacesDemon* [Version 16.1.4; @R-LaplacesDemon], *plyr* [Version 1.8.4; @R-plyr], *dplyr* [Version 0.8.3; @R-dplyr], *readr* [Version 1.3.1; @R-readr], *truncnorm* [Version 1.0.8; @R-truncnorm]
 - To fit the model: *rstan* [Version 2.19.2; @R-rstan]
-- For the visualization of the results: *lattice* [Version 0.20.38; @R-lattice], *ggplot2* [Version 3.3.2; @R-ggplot2], *Rmisc* [Version 1.5; @R-Rmisc], *devtools* [Version 2.2.1; @R-devtools], *gghalves* [Version 0.1.0; @R-gghalves], *bayesplot* [Version 1.7.1; @R-bayesplot], *brms* [Version 2.13.0; @R-brms_a; @R-brms_b], *gridExtra* [Version 2.3; @R-gridExtra], *tibble* [Version 3.0.2; @R-tibble]
+- For the visualization of the results: *lattice* [Version 0.20.38; @R-lattice], *ggplot2* [Version 3.3.2; @R-ggplot2], *Rmisc* [Version 1.5; @R-Rmisc], *devtools* [Version 2.2.1; @R-devtools], *gghalves* [Version 0.1.0; @R-gghalves], *bayesplot* [Version 1.7.1; @R-bayesplot], *brms* [Version 2.13.0; @R-brms_a; @R-brms_b], *gridExtra* [Version 2.3; @R-gridExtra], *ggridges* [Version 0.5.3; @R-ggridges], *ggbeeswarm* [Version 0.6.0; @R-ggbeeswarm], *tibble* [Version 3.0.2; @R-tibble]
 - For model comparison: *bridgesamping* [Version 1.0.0; @R-bridgesampling]

### 2022-06-29 D - Tutorial Normal Model brms/Online Supplement D - Bayesian Hierarchical Modeling in brms.Rmd
Improvements

@@
 title             : "Online Supplement D: Bayesian Hierarchical Modeling in brms"
-author            : "Myrthe Veenman"
+author            : " "
 bibliography      : ["r-references.bib"]
@@
 |    This online supplement provides a tutorial on Bayesian Hierarchical models (BHM) in `brms` [@R-brms_a; @R-brms_b].
-In addition, it contains information about the Number Classification Task that could not be included in the paper.
+In addition, it contains information about the digit classification task that could not be included in the paper.
 As `brms` runs on `Stan` [@carpenter2017stan], it is very similar to `rstan` [@R-rstan].
@@
 - For the document layout: *rmarkdown* [Version 1.16; @R-rmarkdown], *papaja* [Version 0.1.0.9842; @R-papaja], *knitr* [Version 1.25; @R-knitr], *kableExtra* [Version 1.1.0; @R-kableExtra]
-- For data structuring: *LaplacesDemon* [Version 16.1.4; @R-LaplacesDemon], *plyr* [Version 1.8.4; @R-plyr], *dplyr* [Version 0.8.3; @R-dplyr], *readr* [Version 1.3.1; @R-readr], *truncnorm* [Version 1.0.8; @truncnorm]
+- For data structuring: *LaplacesDemon* [Version 16.1.4; @R-LaplacesDemon], *plyr* [Version 1.8.4; @R-plyr], *dplyr* [Version 0.8.3; @R-dplyr], *readr* [Version 1.3.1; @R-readr], *truncnorm* [Version 1.0.8; @R-truncnorm]
 - To fit the model and for model comparison: *brms* [Version 2.13.0; @R-brms_a; @R-brms_b]
-- For the visualization of the results: *lattice* [Version 0.20.38; @R-lattice], *ggplot2* [Version 3.3.2; @R-ggplot2], *Rmisc* [Version 1.5; @R-Rmisc], *devtools* [Version 2.2.1; @R-devtools], *gghalves* [Version 0.1.0; @R-gghalves], *bayesplot* [Version 1.7.1; @R-bayesplot], *gridExtra* [Version 2.3; @R-gridExtra], *ggridges* [Version 0.5.3; @ggridges], *ggbeeswarm* [Version 0.6.0; @ggbeeswarm], *tibble* [Version 3.0.2; @R-tibble]
+- For the visualization of the results: *lattice* [Version 0.20.38; @R-lattice], *ggplot2* [Version 3.3.2; @R-ggplot2], *Rmisc* [Version 1.5; @R-Rmisc], *devtools* [Version 2.2.1; @R-devtools], *gghalves* [Version 0.1.0; @R-gghalves], *bayesplot* [Version 1.7.1; @R-bayesplot], *gridExtra* [Version 2.3; @R-gridExtra], *ggridges* [Version 0.5.3; @R-ggridges], *ggbeeswarm* [Version 0.6.0; @R-ggbeeswarm], *tibble* [Version 3.0.2; @R-tibble]
 

### 2022-06-29 E - Tutorial Log-Normal Model Rstan/Online Supplement E - Bayesian Hierarchical Modeling in rstan - Log Model.Rmd
Improvements

@@
 title             : "Online Supplement E: Bayesian Hierarchical Modeling in rstan - Log Model"
-author            : "Myrthe Veenman" 
+author            : " " 
 bibliography      : ["r-references.bib"]
@@
 |    This online supplement provides a tutorial on log-normal Bayesian Hierarchical models (BHM) in `rstan` [@R-rstan].
-In addition, it contains information about the Number Classification Task that could not be included in the paper.
+In addition, it contains information about the digit classification task that could not be included in the paper.
 Only the fit of the full model will be shown as the fitting the others should follow logically from this tutorial.
@@
 - For the document layout: *rmarkdown* [Version 1.16; @R-rmarkdown], *papaja* [Version 0.1.0.9842; @R-papaja], *knitr* [Version 1.25; @R-knitr], *kableExtra* [Version 1.1.0; @R-kableExtra]
-- For data structuring: *LaplacesDemon* [Version 16.1.4; @R-LaplacesDemon], *plyr* [Version 1.8.4; @R-plyr], *dplyr* [Version 0.8.3; @R-dplyr], *readr* [Version 1.3.1; @R-readr]
+- For data structuring: *LaplacesDemon* [Version 16.1.4; @R-LaplacesDemon], *plyr* [Version 1.8.4; @R-plyr], *dplyr* [Version 0.8.3; @R-dplyr], *readr* [Version 1.3.1; @R-readr], *truncnorm* [Version 1.0.8; @R-truncnorm]
 - To fit the model: *rstan* [Version 2.19.2; @R-rstan]
-- For the visualization of the results: *lattice* [Version 0.20.38; @R-lattice], *ggplot2* [Version 3.3.2; @R-ggplot2], *Rmisc* [Version 1.5; @R-Rmisc], *devtools* [Version 2.2.1; @R-devtools], *gghalves* [Version 0.1.0; @R-gghalves], *bayesplot* [Version 1.7.1; @R-bayesplot], *brms* [Version 2.13.0; @R-brms_a; @R-brms_b], *gridExtra* [Version 2.3; @R-gridExtra], *tibble* [Version 3.0.2; @R-tibble]
+- For the visualization of the results: *lattice* [Version 0.20.38; @R-lattice], *ggplot2* [Version 3.3.2; @R-ggplot2], *Rmisc* [Version 1.5; @R-Rmisc], *devtools* [Version 2.2.1; @R-devtools], *gghalves* [Version 0.1.0; @R-gghalves], *bayesplot* [Version 1.7.1; @R-bayesplot], *brms* [Version 2.13.0; @R-brms_a; @R-brms_b], *gridExtra* [Version 2.3; @R-gridExtra], *ggbeeswarm* [Version 0.6.0; @R-ggbeeswarm], *tibble* [Version 3.0.2; @R-tibble]
 - For model comparison: *bridgesamping* [Version 1.0.0; @R-bridgesampling]

### 2022-06-29 F - Tutorial Log-Normal Model brms/Online Supplement F - Bayesian Hierarchical Modeling in brms - Log Model.Rmd
Improvements

@@
 title             : "Online Supplement F: Bayesian Hierarchical Modeling in brms - Log Model"
-author            : "Myrthe Veenman"
+author            : " "
 bibliography      : ["r-references.bib"]
@@
 |    This online supplement provides a tutorial on log-normal Bayesian Hierarchical models (BHM) in `brms` [@R-brms_a; @R-brms_b].
-In addition, it contains information about the Number Classification Task that could not be included in the paper.
+In addition, it contains information about the digit classification task that could not be included in the paper.
 As `brms` runs on `Stan` [@carpenter2017stan], it is very similar to `rstan` [@R-rstan].
@@
 \scriptsize 
-```{r traceplotsd, }
+```{r traceplotsd, show = FALSE, echo = FALSE}
 # Show trace plot for one sd parameter 

### 2022-06-29 manuscript.Rmd
Improvements

@@
 - Document layout: *rmarkdown* [Version 1.16; @R-rmarkdown], *papaja* [Version 0.1.0.9842; @R-papaja], *knitr* [Version 1.25; @R-knitr], *kableExtra* [Version 1.1.0; @R-kableExtra]
-- Data analysis and cleaning: *LaplacesDemon* [Version 16.1.4; @R-LaplacesDemon], *plyr* [Version 1.8.4; @R-dplyr; @R-plyr], *dplyr* [Version 0.8.3; @R-dplyr], *readr* [Version 1.3.1; @R-readr], *crch* [Version 1.0.4; @R-crch_a; @R-crch_b; @R-crch_c; @R-crch_d], *foreign* [Version 0.8.71; @R-foreign], *Matrix* [Version 1.2.17; @R-Matrix], *tidyr* [Version 1.0.0; @R-tidyr]
+- Data analysis and cleaning: *LaplacesDemon* [Version 16.1.4; @R-LaplacesDemon], *plyr* [Version 1.8.4; @R-dplyr; @R-plyr], *dplyr* [Version 0.8.3; @R-dplyr], *readr* [Version 1.3.1; @R-readr], *crch* [Version 1.0.4; @R-crch_a; @R-crch_b; @R-crch_c; @R-crch_d], *foreign* [Version 0.8.71; @R-foreign], *Matrix* [Version 1.2.17; @R-Matrix], *tidyr* [Version 1.0.0; @R-tidyr], *truncnorm* [Version 1.0.8; @R-truncnorm]
 - Model fit and comparison: *rstan* [Version 2.19.2; @R-rstan], *brms* [Version 2.13.0; @R-brms_a; @R-brms_b], *BayesFactor* [Version 0.9.12.4.2; @R-BayesFactor], *bridgesampling* [Version 1.0.0; @R-bridgesampling]
-- Visualization: *coda* [Version 0.19.3; @R-coda], *lattice* [Version 0.20.38; @R-lattice], *ggplot2* [Version 3.3.2; @R-ggplot2], *Rmisc* [Version 1.5; @R-Rmisc], *devtools* [Version 2.2.1; @R-devtools], *gghalves* [Version 0.1.0; @R-gghalves], *bayesplot* [Version 1.7.1; @R-bayesplot], *gridExtra* [Version 2.3; @R-gridExtra], *tibble* [Version 3.0.2; @R-tibble], *ggmcmc* [Version 1.4.1; @R-ggmcmc], *cowplot* [Version 1.0.0; @R-cowplot], *ggridges* [Version 0.5.1; @R-ggridges], *circlize* [Version 0.4.9; @R-circlize], *StanHeaders* [Version 2.19.0; @R-StanHeaders], *unikn* [Version 0.2.0; @R-unikn]
+- Visualization: *coda* [Version 0.19.3; @R-coda], *lattice* [Version 0.20.38; @R-lattice], *ggplot2* [Version 3.3.2; @R-ggplot2], *Rmisc* [Version 1.5; @R-Rmisc], *devtools* [Version 2.2.1; @R-devtools], *gghalves* [Version 0.1.0; @R-gghalves], *bayesplot* [Version 1.7.1; @R-bayesplot], *gridExtra* [Version 2.3; @R-gridExtra], *tibble* [Version 3.0.2; @R-tibble], *ggmcmc* [Version 1.4.1; @R-ggmcmc], *cowplot* [Version 1.0.0; @R-cowplot], *ggridges* [Version 0.5.1; @R-ggridges], *circlize* [Version 0.4.9; @R-circlize], *StanHeaders* [Version 2.19.0; @R-StanHeaders], *unikn* [Version 0.2.0; @R-unikn], *ggbeeswarm* [Version 0.6.0; @R-ggbeeswarm]
 - Other: *Rcpp* [Version 1.0.3; @R-Rcpp_a; @R-Rcpp_b], *usethis* [Version 1.5.1; @R-usethis]

### 2022-06-29 manuscript.tex
Improvements

@@
 \item
-  A comparison of and guidance for the use of two software packages,\\
+  A comparison of and guidance for the use of two software packages,
 \item
@@
 
-\caption{Panel A shows the response times in seconds for two conditions of the number classification task (i.e., digits 3 and 4).
+\caption{Panel A shows the response times in seconds for two conditions of the digit classification task (i.e., digits 3 and 4).
 The grey lines represent the individual effects, and the blue line the aggregated effect.
@@
 
-\caption{Hierarchical structure of the normal and lognormal models for the number classification task.}\label{fig:hierarchyplot}
+\caption{Hierarchical structure of the normal and lognormal models for the digit classification task.}\label{fig:hierarchyplot}
 \end{figure}
@@
 Finally, we can visualize the simulated data.
-Figure \ref{fig:priorpredfigurepanel} shows prior predictions for the key effects from the number classification task.
+Figure \ref{fig:priorpredfigurepanel} shows prior predictions for the key effects from the digit classification task.
 For this task, we mainly care about the side effect and digit effects corresponding to the \(\beta\)s and the \(\delta\)s from the model.
@@
 \normalsize If you do not specify a prior for a certain parameter, the default prior as explained in the prior section will be applied.
-It is possible to save all priors in an R \texttt{data.frame} and provide these in the \texttt{brm} function, this and the full code with explanation for fitting the full symbolic distance effect model using \texttt{brms} is shown in our \href{https://www.dropbox.com/s/mthphl3ce2kk0xx/Appendix-D--Fit-BHM-in-brms.pdf?dl=0}{Online Supplement D: \emph{Bayesian Hierarchical Modeling in brms}}.\footnote{In our example of the Number Classification Task we are applying a truncated normal prior distribution for only one parameter.
+It is possible to save all priors in an R \texttt{data.frame} and provide these in the \texttt{brm} function, this and the full code with explanation for fitting the full symbolic distance effect model using \texttt{brms} is shown in our \href{https://www.dropbox.com/s/mthphl3ce2kk0xx/Appendix-D--Fit-BHM-in-brms.pdf?dl=0}{Online Supplement D: \emph{Bayesian Hierarchical Modeling in brms}}.\footnote{In our example of the digit classification task we are applying a truncated normal prior distribution for only one parameter.
 This requires a slightly different \texttt{brms} formula and prior specification, explained in the Online Supplement.}
 
-\hypertarget{estimation-of-the-number-classification-task}{%
-\section{Estimation of the Number Classification Task}\label{estimation-of-the-number-classification-task}}
+\hypertarget{estimation-of-the-digit-classification-task}{%
+\section{Estimation of the Digit Classification Task}\label{estimation-of-the-digit-classification-task}}
 
@@
 The data are available under a free license on GitHub \url{https://github.com/PerceptionCognitionLab/data0/tree/master/lexDec-dist5}.
-The data set contains observations from 54 participants who took part in a number classification task.
+The data set contains observations from 54 participants who took part in a digit classification task.
 Each participant performed 360 trials in six blocks.
@@
   Data analysis and cleaning: \emph{LaplacesDemon} {[}Version 16.1.4; Statisticat and LLC.
-(2020){]}, \emph{plyr} {[}Version 1.8.4; Wickham, François, Henry, and Müller (2019); Wickham (2011){]}, \emph{dplyr} {[}Version 0.8.3; Wickham, François, Henry, and Müller (2019){]}, \emph{readr} {[}Version 1.3.1; Wickham, Hester, and Francois (2018){]}, \emph{crch} {[}Version 1.0.4; Messner, Mayr, and Zeileis (2016); Messner, Mayr, and Zeileis (2017); Messner, Zeileis, Broecker, and Mayr (2014); Messner, Mayr, Zeileis, and Wilks (2014){]}, \emph{foreign} {[}Version 0.8.71; R Core Team (2018){]}, \emph{Matrix} {[}Version 1.2.17; Bates and Maechler (2019){]}, \emph{tidyr} {[}Version 1.0.0; Wickham and Henry (2019){]}
+(2020){]}, \emph{plyr} {[}Version 1.8.4; Wickham, François, Henry, and Müller (2019); Wickham (2011){]}, \emph{dplyr} {[}Version 0.8.3; Wickham, François, Henry, and Müller (2019){]}, \emph{readr} {[}Version 1.3.1; Wickham, Hester, and Francois (2018){]}, \emph{crch} {[}Version 1.0.4; Messner, Mayr, and Zeileis (2016); Messner, Mayr, and Zeileis (2017); Messner, Zeileis, Broecker, and Mayr (2014); Messner, Mayr, Zeileis, and Wilks (2014){]}, \emph{foreign} {[}Version 0.8.71; R Core Team (2018){]}, \emph{Matrix} {[}Version 1.2.17; Bates and Maechler (2019){]}, \emph{tidyr} {[}Version 1.0.0; Wickham and Henry (2019){]}, \emph{truncnorm} {[}Version 1.0.8; Mersmann, Trautmann, Steuer, and Bornkamp (2018){]}
 \item
@@
 \item
-  Visualization: \emph{coda} {[}Version 0.19.3; Plummer, Best, Cowles, and Vines (2006){]}, \emph{lattice} {[}Version 0.20.38; Sarkar (2008){]}, \emph{ggplot2} {[}Version 3.3.2; Wickham (2016){]}, \emph{Rmisc} {[}Version 1.5; Hope (2013){]}, \emph{devtools} {[}Version 2.2.1; Wickham, Hester, and Chang (2019){]}, \emph{gghalves} {[}Version 0.1.0; Tiedemann (2020){]}, \emph{bayesplot} {[}Version 1.7.1; Gabry, Simpson, Vehtari, Betancourt, and Gelman (2019){]}, \emph{gridExtra} {[}Version 2.3; Auguie (2017){]}, \emph{tibble} {[}Version 3.0.2; Müller and Wickham (2019){]}, \emph{ggmcmc} {[}Version 1.4.1; Fernández-i-Marín (2016){]}, \emph{cowplot} {[}Version 1.0.0; Wilke (2019){]}, \emph{ggridges} {[}Version 0.5.1; Wilke (2018){]}, \emph{circlize} {[}Version 0.4.9; Gu, Gu, Eils, Schlesner, and Brors (2014){]}, \emph{StanHeaders} {[}Version 2.19.0; Stan Development Team (2019b){]}, \emph{unikn} {[}Version 0.2.0; Neth and Gradwohl (2019){]}
+  Visualization: \emph{coda} {[}Version 0.19.3; Plummer, Best, Cowles, and Vines (2006){]}, \emph{lattice} {[}Version 0.20.38; Sarkar (2008){]}, \emph{ggplot2} {[}Version 3.3.2; Wickham (2016){]}, \emph{Rmisc} {[}Version 1.5; Hope (2013){]}, \emph{devtools} {[}Version 2.2.1; Wickham, Hester, and Chang (2019){]}, \emph{gghalves} {[}Version 0.1.0; Tiedemann (2020){]}, \emph{bayesplot} {[}Version 1.7.1; Gabry, Simpson, Vehtari, Betancourt, and Gelman (2019){]}, \emph{gridExtra} {[}Version 2.3; Auguie (2017){]}, \emph{tibble} {[}Version 3.0.2; Müller and Wickham (2019){]}, \emph{ggmcmc} {[}Version 1.4.1; Fernández-i-Marín (2016){]}, \emph{cowplot} {[}Version 1.0.0; Wilke (2019){]}, \emph{ggridges} {[}Version 0.5.1; Wilke (2018){]}, \emph{circlize} {[}Version 0.4.9; Gu, Gu, Eils, Schlesner, and Brors (2014){]}, \emph{StanHeaders} {[}Version 2.19.0; Stan Development Team (2019b){]}, \emph{unikn} {[}Version 0.2.0; Neth and Gradwohl (2019){]}, \emph{ggbeeswarm} {[}Version 0.6.0; Clarke and Sherrill-Mix (2017){]}
 \item
@@
 
+\leavevmode\hypertarget{ref-R-ggbeeswarm}{}%
+Clarke, E., \& Sherrill-Mix, S. (2017). \emph{Ggbeeswarm: Categorical scatter (violin point) plots}. Retrieved from \url{https://CRAN.R-project.org/package=ggbeeswarm}
+
 \leavevmode\hypertarget{ref-dickey1970weighted}{}%
@@
 Meng, X.-L., \& Wong, W. H. (1996). Simulating ratios of normalizing constants via a simple identity: A theoretical exploration. \emph{Statistica Sinica}, 831--860.
+
+\leavevmode\hypertarget{ref-R-truncnorm}{}%
+Mersmann, O., Trautmann, H., Steuer, D., \& Bornkamp, B. (2018). \emph{Truncnorm: Truncated normal distribution}. Retrieved from \url{https://CRAN.R-project.org/package=truncnorm}
 

### 2022-07-29 manuscript.Rmd
Added declarations

@@
 
-# Acknowledgement
+# Declarations
+
+## Funding
+
+A.S. is supported by the Research Talent Programme from the Dutch Research Council and the Dutch Ministry of Education, Culture and Science (NWO; grant number 406.18.556).
+
+## Conflicts of interest
+
+The authors have no conflicts of interest to declare.
+
+## Availability of data and materials
+
+We do not have ownership of the data used in this study, but the the data set is publicly available on [GitHub](https://github.com/PerceptionCognitionLab/data0/blob/master/lexDec-dist5/).
+The `.stan` models, results of the model fit, and model comparison are available on [our GitHub repository]().
+
+## Code availability
+
+The R code used for the analyses and figures in this manuscript, and `R` tutorials on BHMs are available on [our GitHub repository]().
+
+## Authors’ contributions
+
+M.V.
+participated in the design and coordination of this study, performed statistical analyses, and drafted the paper; A.S.
+participated in the design and coordination of this study, helped with the statistical analyses, interpretation of the results and writing the paper; J.H.
+(principal investigator) conceived the study, participated in the design and coordination of the study, helped with the statistical analyses, interpretation of results and writing the paper.
+All authors read and approved the final paper.
+
+## Acknowledgements
 

### 2022-08-17 C - Tutorial Normal Model Rstan/Online Supplement C - Bayesian Hierarchical Modeling in rstan.Rmd
Updates package versions

--- 2022-06-29 C - Tutorial Normal Model Rstan/Online Supplement C - Bayesian Hierarchical Modeling in rstan.Rmd
@@
 
-- For the document layout: *rmarkdown* [Version 1.16; @R-rmarkdown], *papaja* [Version 0.1.0.9842; @R-papaja], *knitr* [Version 1.25; @R-knitr], *kableExtra* [Version 1.1.0; @R-kableExtra]
-- For data structuring: *LaplacesDemon* [Version 16.1.4; @R-LaplacesDemon], *plyr* [Version 1.8.4; @R-plyr], *dplyr* [Version 0.8.3; @R-dplyr], *readr* [Version 1.3.1; @R-readr]
-- To fit the model: *rstan* [Version 2.19.2; @R-rstan]
-- For the visualization of the results: *lattice* [Version 0.20.38; @R-lattice], *ggplot2* [Version 3.3.2; @R-ggplot2], *Rmisc* [Version 1.5; @R-Rmisc], *devtools* [Version 2.2.1; @R-devtools], *gghalves* [Version 0.1.0; @R-gghalves], *bayesplot* [Version 1.7.1; @R-bayesplot], *brms* [Version 2.13.0; @R-brms_a; @R-brms_b], *gridExtra* [Version 2.3; @R-gridExtra], *tibble* [Version 3.0.2; @R-tibble]
-- For model comparison: *bridgesamping* [Version 1.0.0; @R-bridgesampling]
+- For the document layout: *rmarkdown* [Version 2.9; @R-rmarkdown], *papaja* [Version 0.1.0.9997; @R-papaja], *knitr* [Version 1.33; @R-knitr], *kableExtra* [Version 1.3.4; @R-kableExtra]
+- For data structuring: *LaplacesDemon* [Version 16.1.6; @R-LaplacesDemon], *plyr* [Version 1.8.6; @R-plyr], *dplyr* [Version 1.0.7; @R-dplyr], *readr* [Version 2.0.0; @R-readr], *truncnorm* [Version 1.0.8; @R-truncnorm]
+- To fit the model: *rstan* [Version 2.21.3; @R-rstan]
+- For the visualization of the results: *lattice* [Version 0.20.44; @R-lattice], *ggplot2* [Version 3.3.2; @R-ggplot2], *Rmisc* [Version 1.5; @R-Rmisc], *devtools* [Version 2.4.2; @R-devtools], *gghalves* [Version 0.1.1.9000; @R-gghalves], *bayesplot* [Version 1.8.1; @R-bayesplot], *brms* [Version 2.16.3; @R-brms_a; @R-brms_b], *gridExtra* [Version 2.3; @R-gridExtra], *ggridges* [Version 0.5.3; @R-ggridges], *ggbeeswarm* [Version 0.6.0; @R-ggbeeswarm], *tibble* [Version 3.1.6; @R-tibble]
+- For model comparison: *bridgesamping* [Version 1.1.2; @R-bridgesampling]
 

### 2022-08-17 D - Tutorial Normal Model brms/Online Supplement D - Bayesian Hierarchical Modeling in brms.Rmd
Updates package versions

@@
 
-- For the document layout: *rmarkdown* [Version 1.16; @R-rmarkdown], *papaja* [Version 0.1.0.9842; @R-papaja], *knitr* [Version 1.25; @R-knitr], *kableExtra* [Version 1.1.0; @R-kableExtra]
-- For data structuring: *LaplacesDemon* [Version 16.1.4; @R-LaplacesDemon], *plyr* [Version 1.8.4; @R-plyr], *dplyr* [Version 0.8.3; @R-dplyr], *readr* [Version 1.3.1; @R-readr], *truncnorm* [Version 1.0.8; @R-truncnorm]
-- To fit the model and for model comparison: *brms* [Version 2.13.0; @R-brms_a; @R-brms_b]
-- For the visualization of the results: *lattice* [Version 0.20.38; @R-lattice], *ggplot2* [Version 3.3.2; @R-ggplot2], *Rmisc* [Version 1.5; @R-Rmisc], *devtools* [Version 2.2.1; @R-devtools], *gghalves* [Version 0.1.0; @R-gghalves], *bayesplot* [Version 1.7.1; @R-bayesplot], *gridExtra* [Version 2.3; @R-gridExtra], *ggridges* [Version 0.5.3; @R-ggridges], *ggbeeswarm* [Version 0.6.0; @R-ggbeeswarm], *tibble* [Version 3.0.2; @R-tibble]
+- For the document layout: *rmarkdown* [Version 2.9; @R-rmarkdown], *papaja* [Version 0.1.0.9997; @R-papaja], *knitr* [Version 1.33; @R-knitr], *kableExtra* [Version 1.3.4; @R-kableExtra]
+- For data structuring: *LaplacesDemon* [Version 16.1.6; @R-LaplacesDemon], *plyr* [Version 1.8.6; @R-plyr], *dplyr* [Version 1.0.7; @R-dplyr], *readr* [Version 2.0.0; @R-readr], *truncnorm* [Version 1.0.8; @R-truncnorm]
+- To fit the model and for model comparison: *brms* [Version 2.16.3; @R-brms_a; @R-brms_b]
+- For the visualization of the results: *lattice* [Version 0.20.44; @R-lattice], *ggplot2* [Version 3.3.5; @R-ggplot2], *Rmisc* [Version 1.5; @R-Rmisc], *devtools* [Version 2.4.2; @R-devtools], *gghalves* [Version 0.1.1.9000; @R-gghalves], *bayesplot* [Version 1.8.1; @R-bayesplot], *gridExtra* [Version 2.3; @R-gridExtra], *ggridges* [Version 0.5.3; @R-ggridges], *ggbeeswarm* [Version 0.6.0; @R-ggbeeswarm], *tibble* [Version 3.1.6; @R-tibble]
 

### 2022-08-17 E - Tutorial Log-Normal Model Rstan/Online Supplement E - Bayesian Hierarchical Modeling in rstan - Log Model.Rmd
Updates package versions

@@
 
-- For the document layout: *rmarkdown* [Version 1.16; @R-rmarkdown], *papaja* [Version 0.1.0.9842; @R-papaja], *knitr* [Version 1.25; @R-knitr], *kableExtra* [Version 1.1.0; @R-kableExtra]
-- For data structuring: *LaplacesDemon* [Version 16.1.4; @R-LaplacesDemon], *plyr* [Version 1.8.4; @R-plyr], *dplyr* [Version 0.8.3; @R-dplyr], *readr* [Version 1.3.1; @R-readr], *truncnorm* [Version 1.0.8; @R-truncnorm]
-- To fit the model: *rstan* [Version 2.19.2; @R-rstan]
-- For the visualization of the results: *lattice* [Version 0.20.38; @R-lattice], *ggplot2* [Version 3.3.2; @R-ggplot2], *Rmisc* [Version 1.5; @R-Rmisc], *devtools* [Version 2.2.1; @R-devtools], *gghalves* [Version 0.1.0; @R-gghalves], *bayesplot* [Version 1.7.1; @R-bayesplot], *brms* [Version 2.13.0; @R-brms_a; @R-brms_b], *gridExtra* [Version 2.3; @R-gridExtra], *ggbeeswarm* [Version 0.6.0; @R-ggbeeswarm], *tibble* [Version 3.0.2; @R-tibble]
-- For model comparison: *bridgesamping* [Version 1.0.0; @R-bridgesampling]
+- For the document layout: *rmarkdown* [Version 2.9; @R-rmarkdown], *papaja* [Version 0.1.0.9997; @R-papaja], *knitr* [Version 1.33; @R-knitr], *kableExtra* [Version 1.3.4; @R-kableExtra]
+- For data structuring: *LaplacesDemon* [Version 16.1.6; @R-LaplacesDemon], *plyr* [Version 1.8.6; @R-plyr], *dplyr* [Version 1.0.7; @R-dplyr], *readr* [Version 2.0.0; @R-readr], *truncnorm* [Version 1.0.8; @R-truncnorm]
+- To fit the model: *rstan* [Version 2.21.3; @R-rstan]
+- For the visualization of the results: *lattice* [Version 0.20.44; @R-lattice], *ggplot2* [Version 3.3.5; @R-ggplot2], *Rmisc* [Version 1.5; @R-Rmisc], *devtools* [Version 2.4.2; @R-devtools], *gghalves* [Version 0.1.1.9000; @R-gghalves], *bayesplot* [Version 1.8.1; @R-bayesplot], *brms* [Version 2.16.3; @R-brms_a; @R-brms_b], *gridExtra* [Version 2.3; @R-gridExtra], *ggbeeswarm* [Version 0.6.0; @R-ggbeeswarm], *tibble* [Version 3.1.6; @R-tibble]
+- For model comparison: *bridgesamping* [Version 1.1.2; @R-bridgesampling]
 

### 2022-08-17 manuscript.Rmd
Update manuscript.Rmd
Updates package versions

@@
 
-A.S. is supported by the Research Talent Programme from the Dutch Research Council and the Dutch Ministry of Education, Culture and Science (NWO; grant number 406.18.556).
+This work was supported in part by an NWO Research Talent grant to A.M.S. (406.18.556) and a Veni grant from the NWO (VI.Veni.201G.019) to J.M.H.
 
@@
 
-The R code used for the analyses and figures in this manuscript, and `R` tutorials on BHMs are available on [our GitHub repository]().
+The R code used for the analyses and figures in this manuscript, and `R` tutorials on BHMs are available under an CC-BY-4.0 license on [our GitHub repository]().
 
@@
 
-- Document layout: *rmarkdown* [Version 1.16; @R-rmarkdown], *papaja* [Version 0.1.0.9842; @R-papaja], *knitr* [Version 1.25; @R-knitr], *kableExtra* [Version 1.1.0; @R-kableExtra]
-- Data analysis and cleaning: *LaplacesDemon* [Version 16.1.4; @R-LaplacesDemon], *plyr* [Version 1.8.4; @R-dplyr; @R-plyr], *dplyr* [Version 0.8.3; @R-dplyr], *readr* [Version 1.3.1; @R-readr], *crch* [Version 1.0.4; @R-crch_a; @R-crch_b; @R-crch_c; @R-crch_d], *foreign* [Version 0.8.71; @R-foreign], *Matrix* [Version 1.2.17; @R-Matrix], *tidyr* [Version 1.0.0; @R-tidyr], *truncnorm* [Version 1.0.8; @R-truncnorm]
-- Model fit and comparison: *rstan* [Version 2.19.2; @R-rstan], *brms* [Version 2.13.0; @R-brms_a; @R-brms_b], *BayesFactor* [Version 0.9.12.4.2; @R-BayesFactor], *bridgesampling* [Version 1.0.0; @R-bridgesampling]
-- Visualization: *coda* [Version 0.19.3; @R-coda], *lattice* [Version 0.20.38; @R-lattice], *ggplot2* [Version 3.3.2; @R-ggplot2], *Rmisc* [Version 1.5; @R-Rmisc], *devtools* [Version 2.2.1; @R-devtools], *gghalves* [Version 0.1.0; @R-gghalves], *bayesplot* [Version 1.7.1; @R-bayesplot], *gridExtra* [Version 2.3; @R-gridExtra], *tibble* [Version 3.0.2; @R-tibble], *ggmcmc* [Version 1.4.1; @R-ggmcmc], *cowplot* [Version 1.0.0; @R-cowplot], *ggridges* [Version 0.5.1; @R-ggridges], *circlize* [Version 0.4.9; @R-circlize], *StanHeaders* [Version 2.19.0; @R-StanHeaders], *unikn* [Version 0.2.0; @R-unikn], *ggbeeswarm* [Version 0.6.0; @R-ggbeeswarm]
-- Other: *Rcpp* [Version 1.0.3; @R-Rcpp_a; @R-Rcpp_b], *usethis* [Version 1.5.1; @R-usethis]
+- Document layout: *rmarkdown* [Version 2.9; @R-rmarkdown], *papaja* [Version 0.1.0.9997; @R-papaja], *knitr* [Version 1.33; @R-knitr], *kableExtra* [Version 1.3.4; @R-kableExtra]
+- Data analysis and cleaning: *LaplacesDemon* [Version 16.1.6; @R-LaplacesDemon], *plyr* [Version 1.8.6; @R-dplyr; @R-plyr], *dplyr* [Version 1.0.7; @R-dplyr], *readr* [Version 2.0.0; @R-readr], *crch* [Version 1.0.4; @R-crch_a; @R-crch_b; @R-crch_c; @R-crch_d], *foreign* [Version 0.8.81; @R-foreign], *Matrix* [Version 1.3.3; @R-Matrix], *tidyr* [Version 1.1.3; @R-tidyr], *truncnorm* [Version 1.0.8; @R-truncnorm]
+- Model fit and comparison: *rstan* [Version 2.21.3; @R-rstan], *brms* [Version 2.16.3; @R-brms_a; @R-brms_b], *BayesFactor* [Version 0.9.12.4.2; @R-BayesFactor], *bridgesampling* [Version 1.1.2; @R-bridgesampling]
+- Visualization: *coda* [Version 0.19.4; @R-coda], *lattice* [Version 0.20.44; @R-lattice], *ggplot2* [Version 3.3.5; @R-ggplot2], *Rmisc* [Version 1.5; @R-Rmisc], *devtools* [Version 2.4.2; @R-devtools], *gghalves* [Version 0.1.1.9000; @R-gghalves], *bayesplot* [Version 1.8.1; @R-bayesplot], *gridExtra* [Version 2.3; @R-gridExtra], *tibble* [Version 3.1.6; @R-tibble], *ggmcmc* [Version 1.5.1.1; @R-ggmcmc], *cowplot* [Version 1.1.1; @R-cowplot], *ggridges* [Version 0.5.3; @R-ggridges], *circlize* [Version 0.4.13; @R-circlize], *StanHeaders* [Version 2.21.0.7; @R-StanHeaders], *unikn* [Version 0.4.0; @R-unikn], *ggbeeswarm* [Version 0.6.0; @R-ggbeeswarm]
+- Other: *Rcpp* [Version 1.0.7; @R-Rcpp_a; @R-Rcpp_b], *usethis* [Version 2.0.1; @R-usethis]
 

### 2022-08-17 manuscript.tex
Updates package versions

@@
 
-\hypertarget{acknowledgement}{%
-\section{Acknowledgement}\label{acknowledgement}}
+\hypertarget{declarations}{%
+\section{Declarations}\label{declarations}}
+
+\hypertarget{funding}{%
+\subsection{Funding}\label{funding}}
+
+This work was supported in part by an NWO Research Talent grant to A.M.S. (406.18.556) and a Veni grant from the NWO (VI.Veni.201G.019) to J.M.H.
+
+\hypertarget{conflicts-of-interest}{%
+\subsection{Conflicts of interest}\label{conflicts-of-interest}}
+
+The authors have no conflicts of interest to declare.
+
+\hypertarget{availability-of-data-and-materials}{%
+\subsection{Availability of data and materials}\label{availability-of-data-and-materials}}
+
+We do not have ownership of the data used in this study, but the the data set is publicly available on \href{https://github.com/PerceptionCognitionLab/data0/blob/master/lexDec-dist5/}{GitHub}.
+The \texttt{.stan} models, results of the model fit, and model comparison are available on \href{}{our GitHub repository}.
+
+\hypertarget{code-availability}{%
+\subsection{Code availability}\label{code-availability}}
+
+The R code used for the analyses and figures in this manuscript, and \texttt{R} tutorials on BHMs are available under an CC-BY-4.0 license on \href{}{our GitHub repository}.
+
+\hypertarget{authors-contributions}{%
+\subsection{Authors' contributions}\label{authors-contributions}}
+
+M.V.
+participated in the design and coordination of this study, performed statistical analyses, and drafted the paper; A.S.
+participated in the design and coordination of this study, helped with the statistical analyses, interpretation of the results and writing the paper; J.H.
+(principal investigator) conceived the study, participated in the design and coordination of the study, helped with the statistical analyses, interpretation of results and writing the paper.
+All authors read and approved the final paper.
+
+\hypertarget{acknowledgements}{%
+\subsection{Acknowledgements}\label{acknowledgements}}
 
@@
 \item
-  Document layout: \emph{rmarkdown} {[}Version 1.16; Xie, Allaire, and Grolemund (2018){]}, \emph{papaja} {[}Version 0.1.0.9842; Aust and Barth (2018){]}, \emph{knitr} {[}Version 1.25; Xie (2015){]}, \emph{kableExtra} {[}Version 1.1.0; Zhu (2019){]}
+  Document layout: \emph{rmarkdown} {[}Version 2.9; Xie, Allaire, and Grolemund (2018){]}, \emph{papaja} {[}Version 0.1.0.9997; Aust and Barth (2018){]}, \emph{knitr} {[}Version 1.33; Xie (2015){]}, \emph{kableExtra} {[}Version 1.3.4; Zhu (2019){]}
 \item
-  Data analysis and cleaning: \emph{LaplacesDemon} {[}Version 16.1.4; Statisticat and LLC.
-(2020){]}, \emph{plyr} {[}Version 1.8.4; Wickham, François, Henry, and Müller (2019); Wickham (2011){]}, \emph{dplyr} {[}Version 0.8.3; Wickham, François, Henry, and Müller (2019){]}, \emph{readr} {[}Version 1.3.1; Wickham, Hester, and Francois (2018){]}, \emph{crch} {[}Version 1.0.4; Messner, Mayr, and Zeileis (2016); Messner, Mayr, and Zeileis (2017); Messner, Zeileis, Broecker, and Mayr (2014); Messner, Mayr, Zeileis, and Wilks (2014){]}, \emph{foreign} {[}Version 0.8.71; R Core Team (2018){]}, \emph{Matrix} {[}Version 1.2.17; Bates and Maechler (2019){]}, \emph{tidyr} {[}Version 1.0.0; Wickham and Henry (2019){]}, \emph{truncnorm} {[}Version 1.0.8; Mersmann, Trautmann, Steuer, and Bornkamp (2018){]}
+  Data analysis and cleaning: \emph{LaplacesDemon} {[}Version 16.1.6; Statisticat and LLC.
+(2020){]}, \emph{plyr} {[}Version 1.8.6; Wickham, François, Henry, and Müller (2019); Wickham (2011){]}, \emph{dplyr} {[}Version 1.0.7; Wickham, François, Henry, and Müller (2019){]}, \emph{readr} {[}Version 2.0.0; Wickham, Hester, and Francois (2018){]}, \emph{crch} {[}Version 1.0.4; Messner, Mayr, and Zeileis (2016); Messner, Mayr, and Zeileis (2017); Messner, Zeileis, Broecker, and Mayr (2014); Messner, Mayr, Zeileis, and Wilks (2014){]}, \emph{foreign} {[}Version 0.8.81; R Core Team (2018){]}, \emph{Matrix} {[}Version 1.3.3; Bates and Maechler (2019){]}, \emph{tidyr} {[}Version 1.1.3; Wickham and Henry (2019){]}, \emph{truncnorm} {[}Version 1.0.8; Mersmann, Trautmann, Steuer, and Bornkamp (2018){]}
 \item
-  Model fit and comparison: \emph{rstan} {[}Version 2.19.2; Stan Development Team (2019a){]}, \emph{brms} {[}Version 2.13.0; Bürkner (2017); Bürkner (2018){]}, \emph{BayesFactor} {[}Version 0.9.12.4.2; Morey and Rouder (2018){]}, \emph{bridgesampling} {[}Version 1.0.0; Gronau, Singmann, and Wagenmakers (2020){]}
+  Model fit and comparison: \emph{rstan} {[}Version 2.21.3; Stan Development Team (2019a){]}, \emph{brms} {[}Version 2.16.3; Bürkner (2017); Bürkner (2018){]}, \emph{BayesFactor} {[}Version 0.9.12.4.2; Morey and Rouder (2018){]}, \emph{bridgesampling} {[}Version 1.1.2; Gronau, Singmann, and Wagenmakers (2020){]}
 \item
-  Visualization: \emph{coda} {[}Version 0.19.3; Plummer, Best, Cowles, and Vines (2006){]}, \emph{lattice} {[}Version 0.20.38; Sarkar (2008){]}, \emph{ggplot2} {[}Version 3.3.2; Wickham (2016){]}, \emph{Rmisc} {[}Version 1.5; Hope (2013){]}, \emph{devtools} {[}Version 2.2.1; Wickham, Hester, and Chang (2019){]}, \emph{gghalves} {[}Version 0.1.0; Tiedemann (2020){]}, \emph{bayesplot} {[}Version 1.7.1; Gabry, Simpson, Vehtari, Betancourt, and Gelman (2019){]}, \emph{gridExtra} {[}Version 2.3; Auguie (2017){]}, \emph{tibble} {[}Version 3.0.2; Müller and Wickham (2019){]}, \emph{ggmcmc} {[}Version 1.4.1; Fernández-i-Marín (2016){]}, \emph{cowplot} {[}Version 1.0.0; Wilke (2019){]}, \emph{ggridges} {[}Version 0.5.1; Wilke (2018){]}, \emph{circlize} {[}Version 0.4.9; Gu, Gu, Eils, Schlesner, and Brors (2014){]}, \emph{StanHeaders} {[}Version 2.19.0; Stan Development Team (2019b){]}, \emph{unikn} {[}Version 0.2.0; Neth and Gradwohl (2019){]}, \emph{ggbeeswarm} {[}Version 0.6.0; Clarke and Sherrill-Mix (2017){]}
+  Visualization: \emph{coda} {[}Version 0.19.4; Plummer, Best, Cowles, and Vines (2006){]}, \emph{lattice} {[}Version 0.20.44; Sarkar (2008){]}, \emph{ggplot2} {[}Version 3.3.5; Wickham (2016){]}, \emph{Rmisc} {[}Version 1.5; Hope (2013){]}, \emph{devtools} {[}Version 2.4.2; Wickham, Hester, and Chang (2019){]}, \emph{gghalves} {[}Version 0.1.1.9000; Tiedemann (2020){]}, \emph{bayesplot} {[}Version 1.8.1; Gabry, Simpson, Vehtari, Betancourt, and Gelman (2019){]}, \emph{gridExtra} {[}Version 2.3; Auguie (2017){]}, \emph{tibble} {[}Version 3.1.6; Müller and Wickham (2019){]}, \emph{ggmcmc} {[}Version 1.5.1.1; Fernández-i-Marín (2016){]}, \emph{cowplot} {[}Version 1.1.1; Wilke (2019){]}, \emph{ggridges} {[}Version 0.5.3; Wilke (2018){]}, \emph{circlize} {[}Version 0.4.13; Gu, Gu, Eils, Schlesner, and Brors (2014){]}, \emph{StanHeaders} {[}Version 2.21.0.7; Stan Development Team (2019b){]}, \emph{unikn} {[}Version 0.4.0; Neth and Gradwohl (2019){]}, \emph{ggbeeswarm} {[}Version 0.6.0; Clarke and Sherrill-Mix (2017){]}
 \item
-  Other: \emph{Rcpp} {[}Version 1.0.3; Eddelbuettel and François (2011); Eddelbuettel and Balamuta (2017){]}, \emph{usethis} {[}Version 1.5.1; Wickham and Bryan (2019){]}
+  Other: \emph{Rcpp} {[}Version 1.0.7; Eddelbuettel and François (2011); Eddelbuettel and Balamuta (2017){]}, \emph{usethis} {[}Version 2.0.1; Wickham and Bryan (2019){]}
 \end{itemize}

### 2022-08-17 manuscript.Rmd
Add links to manuscript

@@
 
-  This paper is written in R Markdown. The R Markdown file, other R code, and additional information on this study can be found in the Online Supplement.
+  This paper is written in R Markdown.
+The R Markdown file, other R code, and additional information on this study can be found in the [Online Supplement](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment).
 
@@
 # Model Specification 
-In the following section, we specify two models that can be used for our application example: a normal and a log-normal model.^[Online Supplement A provides the parameterization for models of aggregated and individual data, to illustrate the difference with hierarchical models.
+In the following section, we specify two models that can be used for our application example: a normal and a log-normal model.^[[Online Supplement A](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/blob/main/A%20-%20Model%20Parameterization%20/Online-Supplement-A---Model-Parameterization-for-Aggregated-and-Individual-Models.pdf) provides the parameterization for models of aggregated and individual data, to illustrate the difference with hierarchical models.
 ] We are aware that many psychologists do not engage with equations on a daily basis.
@@
 Therefore, they tend to choose wide priors to indicate little prior knowledge about the parameter values [@aczel2018expert].
-These include default priors offered by programming packages, such as `rstan` (Stan Development Team, 2019a; illustrated in Online Supplement B) and `brms` [@R-brms_a; @R-brms_b].
+These include default priors offered by programming packages, such as `rstan` (Stan Development Team, 2019a; illustrated in [Online Supplement B](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/blob/main/B%20-%20Default%20Priors%20Rstan/Online-Supplement-B---Default-Priors-Rstan.pdf)) and `brms` [@R-brms_a; @R-brms_b].
 Wide priors are priors that allocate (approximately) equal probability density to a great range of values, such as a uniform distribution without bounds [@gelman2006prior].
@@
 Note, however, that the transformation of the prior distributions is not as simple as the transformation of the parameter itself.
-In our Online Supplement, we provide R-code with a numeric solution to the transformation.
+In our [Online Supplement](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/tree/main/R%20code/Prior%20transformation), we provide R-code with a numeric solution to the transformation.
 For the `brms` priors, we obtain the following implied prior distributions on variances:
@@
 The results of the prior prediction for the symbolic distance effect with the priors from Equation \@ref(eq:chosenpriors1) with the number of participants $n = 10$, the number of trials per condition $k = 60$, and the number of simulations $m = 1000$ are shown in panel A of Figure \@ref(fig:priorpredfigurepanel).
-Code for the simulations can be found in our Online Supplement.
+Code for the simulations can be found in our [Online Supplement](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment).
 
@@
 In this section, we explain how a model can be fitted in each package using code snippets.
-In addition, we offer a separate Rmarkdown file for each package in our Online Supplement that explains in more depth how to analyze the full symbolic distance effect model with the specific package.
+In addition, we offer a separate Rmarkdown file for each package in our [Online Supplement](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment) that explains in more depth how to analyze the full symbolic distance effect model with the specific package.
 
@@
 An example `.stan` file of a simple, non-hierarchical normal model, where the variance is already known and the mean has to be estimated, is shown below [adapted example from @nicenboim2021introduction, Chapter 10].
-The `.stan` file for the symbolic distance effect, including explanations, can be found in the Online Supplement.
+The `.stan` file for the symbolic distance effect, including explanations, can be found in the [Online Supplement](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/blob/main/R%20objects/rstan/normal%20model/myModel.stan).
 The file is divided into three sections: `data`, `parameters`, and `model`.
@@
 However, this comes with the price of some technical knowledge.
-The commented code for fitting the full symbolic distance effect model using `rstan`, can be found in [Online Supplement C: *Bayesian Hierarchical Modeling in Rstan*](https://www.dropbox.com/s/06f6vubmnhnm326/Appendix-C--Fit-BHM-in-rstan.pdf?dl=0).
+The commented code for fitting the full symbolic distance effect model using `rstan`, can be found in [Online Supplement C: *Bayesian Hierarchical Modeling in Rstan*](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/blob/main/C%20-%20Tutorial%20Normal%20Model%20Rstan/Online-Supplement-C---Bayesian-Hierarchical-Modeling-in-rstan.pdf).
 
@@
 \normalsize If you do not specify a prior for a certain parameter, the default prior as explained in the prior section will be applied.
-It is possible to save all priors in an R `data.frame` and provide these in the `brm` function, this and the full code with explanation for fitting the full symbolic distance effect model using `brms` is shown in our [Online Supplement D: *Bayesian Hierarchical Modeling in brms*](https://www.dropbox.com/s/mthphl3ce2kk0xx/Appendix-D--Fit-BHM-in-brms.pdf?dl=0).^[In our example of the digit classification task we are applying a truncated normal prior distribution for only one parameter.
+It is possible to save all priors in an R `data.frame` and provide these in the `brm` function, this and the full code with explanation for fitting the full symbolic distance effect model using `brms` is shown in our [Online Supplement D: *Bayesian Hierarchical Modeling in brms*](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/blob/main/D%20-%20Tutorial%20Normal%20Model%20brms/Online-Supplement-D---Bayesian-Hierarchical-Modeling-in-brms.pdf).^[In our example of the digit classification task we are applying a truncated normal prior distribution for only one parameter.
 This requires a slightly different `brms` formula and prior specification, explained in the Online Supplement.]
@@
 However, not that much to be too concerned.
-Trace plots for other parameters can be found in the Online Supplement.
+Trace plots for other parameters can be found in
+[Online Supplement G](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/tree/main/G%20-%20Trace%20Plots%20).
 
@@
 The posterior distributions for the focal parameters in the normal model are presented in Figure \@ref(fig:modelposteriorsmufigure3).
-The posterior means estimated by `rstan` and `brms` can be found in the Online Supplement.
+The posterior means estimated by `rstan` and `brms` can be found in [Online Supplement H](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/tree/main/H%20-%20Posterior%20Distributions%20Tables).
 The credible interval displayed in the figure is the Bayesian version of a confidence interval and provides a measure of uncertainty about the parameter value [@wagenmakers2018bayesian].
@@
 The summaries of the posterior distributions of the variance parameters as estimated by `rstan` are shown in Table \@ref(tab:inddevtable).
-The estimates for the standard deviations by `brms` are smaller compared to `rstan` and can be found in Online Supplement D and H.
+The estimates for the standard deviations by `brms` are smaller compared to `rstan` and can be found in [Online Supplements D and H](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment).
 
@@
 Pink intervals indicate that the 95% credible interval contains zero, while blue intervals are either entirely above or below zero.
-A corresponding figure with individual effect estimates for `brms` can be found in Online Supplement D.
+A corresponding figure with individual effect estimates for `brms` can be found in [Online Supplement D](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/blob/main/D%20-%20Tutorial%20Normal%20Model%20brms/Online-Supplement-D---Bayesian-Hierarchical-Modeling-in-brms.pdf).
 
@@
 The trace plot of $\mu_{\delta_{6}}$ is shown panel A for `rstan` and B for `brms` of Figure \@ref(fig:traceplotlog).
-Trace plots for all other parameters can be found in Online Supplement G.
+Trace plots for all other parameters can be found in [Online Supplement G](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/tree/main/G%20-%20Trace%20Plots%20).
 All plots look like a hairy caterpillar.
@@
 Table \@ref(tab:inddevtablelog) presents the estimates of the variance parameters by `rstan`.
-Results for brms can be found in Online Supplement H.
+Results for brms can be found in [Online Supplement H](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/blob/main/H%20-%20Posterior%20Distributions%20Tables/Online-Supplement-H---Tables-with-Information-on-Posterior-Distributions.pdf).
 ```{r inddevtablelog, results = "asis", cache = TRUE}
@@
 The estimates for the standard deviations by `brms` are similar compared to the estimates by `rstan`, but slightly smaller.
-The exact estimates by `brms` can be found in Online Supplement F and H.
+The exact estimates by `brms` can be found in [Online Supplements F and H](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment).
 Figure \@ref(fig:randomparplotlog) shows the individual estimates for the digit and side effects.
@@
 First, we kept the discussion of the software packages and code fairly brief.
-We provide further code examples for the digit classification task in the Online Supplement.
+We provide further code examples for the digit classification task in the [Online Supplement](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment).
 For different models, it may be useful to study additional tutorials and books.

### 2022-08-17 manuscript.tex
Add links to manuscript

@@
 
-This paper is written in R Markdown. The R Markdown file, other R code, and additional information on this study can be found in the Online Supplement.
+This paper is written in R Markdown.
+The R Markdown file, other R code, and additional information on this study can be found in the \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment}{Online Supplement}.
 
@@
 
-In the following section, we specify two models that can be used for our application example: a normal and a log-normal model.\footnote{Online Supplement A provides the parameterization for models of aggregated and individual data, to illustrate the difference with hierarchical models.} We are aware that many psychologists do not engage with equations on a daily basis.
+In the following section, we specify two models that can be used for our application example: a normal and a log-normal model.\footnote{\href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/blob/main/A\%20-\%20Model\%20Parameterization\%20/Online-Supplement-A---Model-Parameterization-for-Aggregated-and-Individual-Models.pdf}{Online Supplement A} provides the parameterization for models of aggregated and individual data, to illustrate the difference with hierarchical models.} We are aware that many psychologists do not engage with equations on a daily basis.
 In this section, however, we try to accessibly summarize the key considerations of model specification.
@@
 Therefore, they tend to choose wide priors to indicate little prior knowledge about the parameter values (Aczel et al., 2018).
-These include default priors offered by programming packages, such as \texttt{rstan} (Stan Development Team, 2019a; illustrated in Online Supplement B) and \texttt{brms} (Bürkner, 2017, 2018).
+These include default priors offered by programming packages, such as \texttt{rstan} (Stan Development Team, 2019a; illustrated in \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/blob/main/B\%20-\%20Default\%20Priors\%20Rstan/Online-Supplement-B---Default-Priors-Rstan.pdf}{Online Supplement B}) and \texttt{brms} (Bürkner, 2017, 2018).
 Wide priors are priors that allocate (approximately) equal probability density to a great range of values, such as a uniform distribution without bounds (Gelman, 2006).
@@
 Note, however, that the transformation of the prior distributions is not as simple as the transformation of the parameter itself.
-In our Online Supplement, we provide R-code with a numeric solution to the transformation.
+In our \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/tree/main/R\%20code/Prior\%20transformation}{Online Supplement}, we provide R-code with a numeric solution to the transformation.
 For the \texttt{brms} priors, we obtain the following implied prior distributions on variances:
@@
 The results of the prior prediction for the symbolic distance effect with the priors from Equation \eqref{eq:chosenpriors1} with the number of participants \(n = 10\), the number of trials per condition \(k = 60\), and the number of simulations \(m = 1000\) are shown in panel A of Figure \ref{fig:priorpredfigurepanel}.
-Code for the simulations can be found in our Online Supplement.
+Code for the simulations can be found in our \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment}{Online Supplement}.
 
@@
 In this section, we explain how a model can be fitted in each package using code snippets.
-In addition, we offer a separate Rmarkdown file for each package in our Online Supplement that explains in more depth how to analyze the full symbolic distance effect model with the specific package.
+In addition, we offer a separate Rmarkdown file for each package in our \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment}{Online Supplement} that explains in more depth how to analyze the full symbolic distance effect model with the specific package.
 
@@
 An example \texttt{.stan} file of a simple, non-hierarchical normal model, where the variance is already known and the mean has to be estimated, is shown below (adapted example from Nicenboim, Schad, \& Vasishth, 2021, Chapter 10).
-The \texttt{.stan} file for the symbolic distance effect, including explanations, can be found in the Online Supplement.
+The \texttt{.stan} file for the symbolic distance effect, including explanations, can be found in the \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/blob/main/R\%20objects/rstan/normal\%20model/myModel.stan}{Online Supplement}.
 The file is divided into three sections: \texttt{data}, \texttt{parameters}, and \texttt{model}.
@@
 However, this comes with the price of some technical knowledge.
-The commented code for fitting the full symbolic distance effect model using \texttt{rstan}, can be found in \href{https://www.dropbox.com/s/06f6vubmnhnm326/Appendix-C--Fit-BHM-in-rstan.pdf?dl=0}{Online Supplement C: \emph{Bayesian Hierarchical Modeling in Rstan}}.
+The commented code for fitting the full symbolic distance effect model using \texttt{rstan}, can be found in \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/blob/main/C\%20-\%20Tutorial\%20Normal\%20Model\%20Rstan/Online-Supplement-C---Bayesian-Hierarchical-Modeling-in-rstan.pdf}{Online Supplement C: \emph{Bayesian Hierarchical Modeling in Rstan}}.
 
@@
 \normalsize If you do not specify a prior for a certain parameter, the default prior as explained in the prior section will be applied.
-It is possible to save all priors in an R \texttt{data.frame} and provide these in the \texttt{brm} function, this and the full code with explanation for fitting the full symbolic distance effect model using \texttt{brms} is shown in our \href{https://www.dropbox.com/s/mthphl3ce2kk0xx/Appendix-D--Fit-BHM-in-brms.pdf?dl=0}{Online Supplement D: \emph{Bayesian Hierarchical Modeling in brms}}.\footnote{In our example of the digit classification task we are applying a truncated normal prior distribution for only one parameter.
+It is possible to save all priors in an R \texttt{data.frame} and provide these in the \texttt{brm} function, this and the full code with explanation for fitting the full symbolic distance effect model using \texttt{brms} is shown in our \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/blob/main/D\%20-\%20Tutorial\%20Normal\%20Model\%20brms/Online-Supplement-D---Bayesian-Hierarchical-Modeling-in-brms.pdf}{Online Supplement D: \emph{Bayesian Hierarchical Modeling in brms}}.\footnote{In our example of the digit classification task we are applying a truncated normal prior distribution for only one parameter.
 This requires a slightly different \texttt{brms} formula and prior specification, explained in the Online Supplement.}
@@
 However, not that much to be too concerned.
-Trace plots for other parameters can be found in the Online Supplement.
+Trace plots for other parameters can be found in
+\href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/tree/main/G\%20-\%20Trace\%20Plots\%20}{Online Supplement G}.
 
@@
 The posterior distributions for the focal parameters in the normal model are presented in Figure \ref{fig:modelposteriorsmufigure3}.
-The posterior means estimated by \texttt{rstan} and \texttt{brms} can be found in the Online Supplement.
+The posterior means estimated by \texttt{rstan} and \texttt{brms} can be found in \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/tree/main/H\%20-\%20Posterior\%20Distributions\%20Tables}{Online Supplement H}.
 The credible interval displayed in the figure is the Bayesian version of a confidence interval and provides a measure of uncertainty about the parameter value (Wagenmakers et al., 2018).
@@
 The summaries of the posterior distributions of the variance parameters as estimated by \texttt{rstan} are shown in Table \ref{tab:inddevtable}.
-The estimates for the standard deviations by \texttt{brms} are smaller compared to \texttt{rstan} and can be found in Online Supplement D and H.
+The estimates for the standard deviations by \texttt{brms} are smaller compared to \texttt{rstan} and can be found in \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment}{Online Supplements D and H}.
 
@@
 Pink intervals indicate that the 95\% credible interval contains zero, while blue intervals are either entirely above or below zero.
-A corresponding figure with individual effect estimates for \texttt{brms} can be found in Online Supplement D.
+A corresponding figure with individual effect estimates for \texttt{brms} can be found in \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/blob/main/D\%20-\%20Tutorial\%20Normal\%20Model\%20brms/Online-Supplement-D---Bayesian-Hierarchical-Modeling-in-brms.pdf}{Online Supplement D}.
 
@@
 The trace plot of \(\mu_{\delta_{6}}\) is shown panel A for \texttt{rstan} and B for \texttt{brms} of Figure \ref{fig:traceplotlog}.
-Trace plots for all other parameters can be found in Online Supplement G.
+Trace plots for all other parameters can be found in \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/tree/main/G\%20-\%20Trace\%20Plots\%20}{Online Supplement G}.
 All plots look like a hairy caterpillar.
@@
 Table \ref{tab:inddevtablelog} presents the estimates of the variance parameters by \texttt{rstan}.
-Results for brms can be found in Online Supplement H.
+Results for brms can be found in \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/blob/main/H\%20-\%20Posterior\%20Distributions\%20Tables/Online-Supplement-H---Tables-with-Information-on-Posterior-Distributions.pdf}{Online Supplement H}.
 
@@
 The estimates for the standard deviations by \texttt{brms} are similar compared to the estimates by \texttt{rstan}, but slightly smaller.
-The exact estimates by \texttt{brms} can be found in Online Supplement F and H.
+The exact estimates by \texttt{brms} can be found in \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment}{Online Supplements F and H}.
 Figure \ref{fig:randomparplotlog} shows the individual estimates for the digit and side effects.
@@
 First, we kept the discussion of the software packages and code fairly brief.
-We provide further code examples for the digit classification task in the Online Supplement.
+We provide further code examples for the digit classification task in the \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment}{Online Supplement}.
 For different models, it may be useful to study additional tutorials and books.

### 2022-08-17 manuscript.Rmd
Small updates manuscript

@@
   - name          : "Myrthe Veenman"
-    affiliation   : "1"
+    affiliation   : "2"
     corresponding : yes    # Define only one corresponding author
-    address       : " "
+    address       : "Wassenaarseweg 52, Leiden"
     email         : "myrthe.veenman@gmail.com"
@@
     institution   : "University of Amsterdam"
+  - id            : "2"
+    institution   : "Leiden University"
 
@@
 We do not have ownership of the data used in this study, but the the data set is publicly available on [GitHub](https://github.com/PerceptionCognitionLab/data0/blob/master/lexDec-dist5/).
-The `.stan` models, results of the model fit, and model comparison are available on [our GitHub repository]().
+The `.stan` models, results of the model fit, and model comparison are available on [our GitHub repository](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment).
 
@@
 
-The R code used for the analyses and figures in this manuscript, and `R` tutorials on BHMs are available under an CC-BY-4.0 license on [our GitHub repository]().
+The R code used for the analyses and figures in this manuscript, and `R` tutorials on BHMs are available on [our GitHub repository](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment).
 
@@
 M.V.
-participated in the design and coordination of this study, performed statistical analyses, and drafted the paper; A.S.
-participated in the design and coordination of this study, helped with the statistical analyses, interpretation of the results and writing the paper; J.H.
+participated in the design and coordination of this study, performed statistical analyses, and drafted the paper; A.M.S.
+participated in the design and coordination of this study, helped with the statistical analyses, interpretation of the results and writing the paper; J.M.H.
 (principal investigator) conceived the study, participated in the design and coordination of the study, helped with the statistical analyses, interpretation of results and writing the paper.

### 2022-08-17 manuscript.tex
Small updates manuscript

@@
   pdftitle={Bayesian Hierarchical Modeling: An Introduction and Reassessment},
-  pdfauthor={Myrthe Veenman1, Angelika Stefan1, \& Julia Haaf1},
+  pdfauthor={Myrthe Veenman2, Angelika Stefan1, \& Julia Haaf1},
   pdflang={en-EN},
@@
 \title{Bayesian Hierarchical Modeling: An Introduction and Reassessment}
-\author{Myrthe Veenman\textsuperscript{1}, Angelika Stefan\textsuperscript{1}, \& Julia Haaf\textsuperscript{1}}
+\author{Myrthe Veenman\textsuperscript{2}, Angelika Stefan\textsuperscript{1}, \& Julia Haaf\textsuperscript{1}}
 \date{}
@@
 
-Correspondence concerning this article should be addressed to Myrthe Veenman, . E-mail: \href{mailto:myrthe.veenman@gmail.com}{\nolinkurl{myrthe.veenman@gmail.com}}
+Correspondence concerning this article should be addressed to Myrthe Veenman, Wassenaarseweg 52, Leiden. E-mail: \href{mailto:myrthe.veenman@gmail.com}{\nolinkurl{myrthe.veenman@gmail.com}}
 
@@
 
-\affiliation{\vspace{0.5cm}\textsuperscript{1} University of Amsterdam}
+\affiliation{\vspace{0.5cm}\textsuperscript{1} University of Amsterdam\\\textsuperscript{2} Leiden University}
 
@@
 We do not have ownership of the data used in this study, but the the data set is publicly available on \href{https://github.com/PerceptionCognitionLab/data0/blob/master/lexDec-dist5/}{GitHub}.
-The \texttt{.stan} models, results of the model fit, and model comparison are available on \href{}{our GitHub repository}.
+The \texttt{.stan} models, results of the model fit, and model comparison are available on \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment}{our GitHub repository}.
 
@@
 
-The R code used for the analyses and figures in this manuscript, and \texttt{R} tutorials on BHMs are available under an CC-BY-4.0 license on \href{}{our GitHub repository}.
+The R code used for the analyses and figures in this manuscript, and \texttt{R} tutorials on BHMs are available on \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment}{our GitHub repository}.
 

### 2022-08-19 manuscript.Rmd
small updates

@@
   This paper is written in R Markdown.
-The R Markdown file, other R code, and additional information on this study can be found in the [Online Supplement](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment).
+The R Markdown file, other R code, and additional information on this study can be found in the [Online Supplement: https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment).
 
@@
 In hierarchical modeling, these individual effects are optimally and automatically corrected towards the general trend [@efron1977stein].
-This phenomenon, also called "shrinkage", is illustrated in Panel B of Figure \@ref(fig:figure1).
+This phenomenon, also called `shrinkage', is illustrated in Panel B of Figure \@ref(fig:figure1).
 The rightmost observation in the figure is in sample quite large compared to the other observed effects.
@@
 
-```{r hierarchyplot, echo=FALSE, fig.cap="Hierarchical structure of the normal and lognormal models for the digit classification task.", fig.align = "left", fig.height = 3, fig.width = 4, out.width="70%"}
+```{r hierarchyplot, echo=FALSE, fig.cap="Hierarchical structure of the normal and log-normal models for the digit classification task.", fig.align = "left", fig.height = 3, fig.width = 4, out.width="70%"}
 knitr::include_graphics("Images/Fig2_Hierarchical structure.png")
@@
 Panel B shows the prior predictions for the normal model based on the selected priors.
-Panel C shows the prior predictions for the lognormal model based on the priors chosen for the normal model.
-Panel D shows the prior predictions for the lognormal model based on the selected priors.", fig.align = "left", fig.height = 3, fig.width = 4, out.width="100%"}
+Panel C shows the prior predictions for the log-normal model based on the priors chosen for the normal model.
+Panel D shows the prior predictions for the log-normal model based on the selected priors.", fig.align = "left", fig.height = 3, fig.width = 4, out.width="100%"}
 knitr::include_graphics("Images/Fig5_priorpredfigure_v2.png")
@@
 In this section, we explain how a model can be fitted in each package using code snippets.
-In addition, we offer a separate Rmarkdown file for each package in our [Online Supplement](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment) that explains in more depth how to analyze the full symbolic distance effect model with the specific package.
+In addition, we offer a separate R Markdown file for each package in our [Online Supplement](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment) that explains in more depth how to analyze the full symbolic distance effect model with the specific package.
 
@@
 (2005).
-The data are available under a free license on GitHub [https://github.com/PerceptionCognitionLab/data0/tree/master/lexDec-dist5](https://github.com/PerceptionCognitionLab/data0/tree/master/lexDec-dist5).
+The data are available under a free license on GitHub [https://github.com/PerceptionCognitionLab/data0/tree/master/lexDec-dist5]((https://github.com/PerceptionCognitionLab/data0/tree/master/lexDec-dist5)).
 The data set contains observations from 54 participants who took part in a digit classification task.
@@
 
+\newpage
+
 ## Normal Model
@@
 ### Estimation Results
-Considering good/acceptable convergence diagnostics, we can now interpret the estimation results. First, we present the general effects, then we visualize the individual variability of these effects.
+Considering acceptable to good convergence diagnostics, we can now interpret the estimation results.
+First, we present the general effects, then we visualize the individual variability of these effects.
 
@@
 
-We do not have ownership of the data used in this study, but the the data set is publicly available on [GitHub](https://github.com/PerceptionCognitionLab/data0/blob/master/lexDec-dist5/).
-The `.stan` models, results of the model fit, and model comparison are available on [our GitHub repository](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment).
+We do not have ownership of the data used in this study, but the the data set is publicly available on [GitHub (https://github.com/PerceptionCognitionLab/data0/blob/master/lexDec-dist5/)](https://github.com/PerceptionCognitionLab/data0/blob/master/lexDec-dist5/).
+The `.stan` models, results of the model fit, and model comparison are available on [our GitHub repository: https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment).
 
@@
 
-The R code used for the analyses and figures in this manuscript, and `R` tutorials on BHMs are available on [our GitHub repository](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment).
+The R code used for the analyses and figures in this manuscript, and `R` tutorials on BHMs are available on [our GitHub repository: https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment).
 
@@
 \setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in}
+\setlength{\leftskip}{0.2in}
 

### 2022-08-19 manuscript.tex
small updates

@@
 This paper is written in R Markdown.
-The R Markdown file, other R code, and additional information on this study can be found in the \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment}{Online Supplement}.
+The R Markdown file, other R code, and additional information on this study can be found in the \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment}{Online Supplement: https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment}.
 
@@
 In hierarchical modeling, these individual effects are optimally and automatically corrected towards the general trend (Efron \& Morris, 1977).
-This phenomenon, also called ``shrinkage,'' is illustrated in Panel B of Figure \ref{fig:figure1}.
+This phenomenon, also called `shrinkage', is illustrated in Panel B of Figure \ref{fig:figure1}.
 The rightmost observation in the figure is in sample quite large compared to the other observed effects.
@@
 
-\caption{Hierarchical structure of the normal and lognormal models for the digit classification task.}\label{fig:hierarchyplot}
+\caption{Hierarchical structure of the normal and log-normal models for the digit classification task.}\label{fig:hierarchyplot}
 \end{figure}
@@
 Panel B shows the prior predictions for the normal model based on the selected priors.
-Panel C shows the prior predictions for the lognormal model based on the priors chosen for the normal model.
-Panel D shows the prior predictions for the lognormal model based on the selected priors.}\label{fig:priorpredfigurepanel}
+Panel C shows the prior predictions for the log-normal model based on the priors chosen for the normal model.
+Panel D shows the prior predictions for the log-normal model based on the selected priors.}\label{fig:priorpredfigurepanel}
 \end{figure}
@@
 In this section, we explain how a model can be fitted in each package using code snippets.
-In addition, we offer a separate Rmarkdown file for each package in our \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment}{Online Supplement} that explains in more depth how to analyze the full symbolic distance effect model with the specific package.
+In addition, we offer a separate R Markdown file for each package in our \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment}{Online Supplement} that explains in more depth how to analyze the full symbolic distance effect model with the specific package.
 
@@
 For our example, we will use the data set by Rouder et al.~(2005).
-The data are available under a free license on GitHub \url{https://github.com/PerceptionCognitionLab/data0/tree/master/lexDec-dist5}.
+The data are available under a free license on GitHub \href{(https://github.com/PerceptionCognitionLab/data0/tree/master/lexDec-dist5)}{https://github.com/PerceptionCognitionLab/data0/tree/master/lexDec-dist5}.
 The data set contains observations from 54 participants who took part in a digit classification task.
@@
 
+\newpage
+
 \hypertarget{normal-model-2}{%
@@
 
-Considering good/acceptable convergence diagnostics, we can now interpret the estimation results. First, we present the general effects, then we visualize the individual variability of these effects.
+Considering acceptable to good convergence diagnostics, we can now interpret the estimation results.
+First, we present the general effects, then we visualize the individual variability of these effects.
 
@@
 
-We do not have ownership of the data used in this study, but the the data set is publicly available on \href{https://github.com/PerceptionCognitionLab/data0/blob/master/lexDec-dist5/}{GitHub}.
-The \texttt{.stan} models, results of the model fit, and model comparison are available on \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment}{our GitHub repository}.
+We do not have ownership of the data used in this study, but the the data set is publicly available on \href{https://github.com/PerceptionCognitionLab/data0/blob/master/lexDec-dist5/}{GitHub (https://github.com/PerceptionCognitionLab/data0/blob/master/lexDec-dist5/)}.
+The \texttt{.stan} models, results of the model fit, and model comparison are available on \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment}{our GitHub repository: https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment}.
 
@@
 
-The R code used for the analyses and figures in this manuscript, and \texttt{R} tutorials on BHMs are available on \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment}{our GitHub repository}.
+The R code used for the analyses and figures in this manuscript, and \texttt{R} tutorials on BHMs are available on \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment}{our GitHub repository: https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment}.
 

### 2022-08-19 manuscript.Rmd
Small updates

@@
   - name          : "Myrthe Veenman"
-    affiliation   : "2"
+    affiliation   : "1"
     corresponding : yes    # Define only one corresponding author
@@
   - name          : "Angelika Stefan"
-    affiliation   : "1"
+    affiliation   : "2"
     corresponding : no    # Define only one corresponding author
   - name          : "Julia Haaf"
-    affiliation   : "1"
+    affiliation   : "2"
     corresponding : no    # Define only one corresponding author
@@
   - id            : "1"
+    institution   : "Leiden University"
+  - id            : "2"
     institution   : "University of Amsterdam"
-  - id            : "2"
-    institution   : "Leiden University"
 

### 2022-08-19 manuscript.tex
Small updates

@@
   pdftitle={Bayesian Hierarchical Modeling: An Introduction and Reassessment},
-  pdfauthor={Myrthe Veenman2, Angelika Stefan1, \& Julia Haaf1},
+  pdfauthor={Myrthe Veenman1, Angelika Stefan2, \& Julia Haaf2},
   pdflang={en-EN},
@@
 \title{Bayesian Hierarchical Modeling: An Introduction and Reassessment}
-\author{Myrthe Veenman\textsuperscript{2}, Angelika Stefan\textsuperscript{1}, \& Julia Haaf\textsuperscript{1}}
+\author{Myrthe Veenman\textsuperscript{1}, Angelika Stefan\textsuperscript{2}, \& Julia Haaf\textsuperscript{2}}
 \date{}
@@
 
-\affiliation{\vspace{0.5cm}\textsuperscript{1} University of Amsterdam\\\textsuperscript{2} Leiden University}
+\affiliation{\vspace{0.5cm}\textsuperscript{1} Leiden University\\\textsuperscript{2} University of Amsterdam}
 

### 2022-08-22 manuscript.Rmd
Update manuscript.Rmd

@@
     email         : "myrthe.veenman@gmail.com"
-  - name          : "Angelika Stefan"
+  - name          : "Angelika M. Stefan"
     affiliation   : "2"
     corresponding : no    # Define only one corresponding author
-  - name          : "Julia Haaf"
+  - name          : "Julia M. Haaf"
     affiliation   : "2"

### 2022-10-26 manuscript.Rmd
Updates writing + models on correlation random effects

@@
 Additionally, we may hypothesize that RTs are affected by whether the target digit was smaller or greater than 5, yet it is unclear which side of 5 would lead to faster or slower RTs.
-The digit classification task is a standard task that usually would be analyzed in a 3 (digit) by 2 (side) within-subjects ANOVA.
-In this tutorial, we show how inferences can be improved using a hierarchical Bayesian model instead.
+\textcolor{blue}{The digit classification task is a standard task that usually would be analyzed in a 3 (digit, closest to 5, further away from 5, furthest away from 5) by 2 (side, greater or samller than 5) within-subjects ANOVA.} In this tutorial, we show how inferences can be improved using a hierarchical Bayesian model instead.
 
@@
 
-```{r figure1code, echo=FALSE, fig.align='left', fig.cap="Response times in milliseconds in two conditions (i.e., digit 3 and 4).
+```{r figure1code, echo=FALSE, fig.align='left', fig.cap="Average response times in milliseconds in two conditions (i.e., digit 3 and 4).
 The grey lines represent the individual response time effects, and the blue line the mean aggregated response time effect.", fig.height=4.5, fig.width=6, eval = FALSE}
@@
 
-```{r figure1, echo=FALSE, fig.align='left', fig.cap="Panel A shows the response times in seconds for two conditions of the digit classification task (i.e., digits 3 and 4).
-The grey lines represent the individual effects, and the blue line the aggregated effect.
+```{r figure1, echo=FALSE, fig.align='left', fig.cap="Panel A shows the average response times in seconds for two conditions of the digit classification task (i.e., digits 3 and 4).
+The grey lines represent the average within individual effects, and the blue line the aggregated effect over all individuals.
 Panel B shows how individual estimates are shrunken towards the general trend.
@@
 \end{align*}
-The model equation shows that, for instance, for participant *i* = 1 and digit *j* = 7, the mean RT is modeled by: $\gamma_{i = 1} - \frac{1}{2} \beta_{i = 1} + \delta_{7,i = 1}$.
+\textcolor{blue}{The model equation shows that, for instance, for participant $i = 1$ and digit $j = 7$, with $\gamma_{i = 1}$ equal to $0.5$, $\beta_{i = 1}$ equal to $0.2$, $\delta_{7, i = 1}$ and $\delta_{3, i = 1}$ equal to $0.25$, and $\delta_{6, i = 1}$ and $\delta_{4, i = 1}$ equal to $0.3$, the mean RT is modeled by: $0.5 - 0.1 +  0.25$.
 If all $\delta$s are positive, then the model equation indicates that individuals respond fastest when the digits 2 and 8 are presented, as in this situation there are no digit effects in the equation.
+However, for digit $j = 8$, the mean RT is modeled by $0.5 - 0.1$ in this scenario, while for digit $j = 2$, the mean RT is modeled by $0.5 + 0.1$.}
 
@@
 The `R` package `brms` also allows fitting Bayesian models using `Stan`.
-However, model specification and model fitting can be achieved using a much simplified `R`-script.
+However, model specification and model fitting can be achieved using a much simplified `R`-script \textcolor{blue}{alone}.
 Model specification is based on a formula syntax that is similar to the notation in the popular `lme4` package for frequentist hierarchical modeling [@R-lme4].
@@
 This means that the data are 100 times more likely under the normal model than under the log-normal model, clear evidence in favor of the normal model.
-But how do we know whether a Bayes factor of 100-to-1, or 10-to-1 or 3-to-1.
+\textcolor{blue}{But how do we know whether a Bayes factor of 100-to-1, or 10-to-1 or 3-to-1 fulfills?}.
 As a reminder, Bayes factors are the relative predictive accuracy of two models for the observed data.
@@
 Finally, we compare the estimated density of the posterior to the density of the prior.
+\textcolor{blue}{This process is explained in further detail by Van Ravenzwaaij and Etz} [-@van2021simulation].
 
@@
 
-Note that models models 2-4 are nested in the most complex full model, and the null model is nested in models 3 and 4, but models 3 and 4 are not nested.
+\textcolor{blue}{Note that models models 2-4 are nested in the most complex full model, and the null model is nested in models 2 and 3, but models 2 and 3 are not nested}.
 Additionally, the models differ by several parameters.
@@
 Next, we calculate the sum of all these Bayes factors and divide each Bayes factor by this sum.
-This results in the posterior model probabilities shown in Table 5.^[Note that it would be possible to calculate the posterior model probabilities directly from the marginal likelihoods.
+This results in the posterior model probabilities shown in Table \@ref(tab:postmodprob).^[Note that it would be possible to calculate the posterior model probabilities directly from the marginal likelihoods.
 However, marginal likelihoods are often not computed explicitly in the model fitting process (e.g., when using the bayes_factor function in the `brms` package), so we believe that readers might find it easier to compute posterior model probabilities from Bayes factors.] All posterior model probabilities from the table and the posterior probability of the null model (see table note) sum to one.
@@
 \caption{Posterior Model Probabilities}
+\label{tab:postmodprob}
 \begin{tabularx}{\textwidth}{XXXX}
@@
 While individual differences are a common phenomenon in psychology, there are instances where general effects are sufficient to describe the data adequately, or where only a random intercept, but no random slopes are necessary.
-As shown in Table 5, Bayesian model comparisons can help researchers to identify the model that best describes the data.
+As shown in Table \@ref(tab:postmodprob), Bayesian model comparisons can help researchers to identify the model that best describes the data.
 

### 2022-10-26 manuscript.tex
Updates writing + models on correlation random effects

@@
   pdftitle={Bayesian Hierarchical Modeling: An Introduction and Reassessment},
-  pdfauthor={Myrthe Veenman1, Angelika Stefan2, \& Julia Haaf2},
+  pdfauthor={Myrthe Veenman1, Angelika M. Stefan2, \& Julia M. Haaf2},
   pdflang={en-EN},
@@
 \title{Bayesian Hierarchical Modeling: An Introduction and Reassessment}
-\author{Myrthe Veenman\textsuperscript{1}, Angelika Stefan\textsuperscript{2}, \& Julia Haaf\textsuperscript{2}}
+\author{Myrthe Veenman\textsuperscript{1}, Angelika M. Stefan\textsuperscript{2}, \& Julia M. Haaf\textsuperscript{2}}
 \date{}
@@
 
-\caption{Panel A shows the response times in seconds for two conditions of the digit classification task (i.e., digits 3 and 4).
-The grey lines represent the individual effects, and the blue line the aggregated effect.
+\caption{Panel A shows the average response times in seconds for two conditions of the digit classification task (i.e., digits 3 and 4).
+The grey lines represent the average within individual effects, and the blue line the aggregated effect over all individuals.
 Panel B shows how individual estimates are shrunken towards the general trend.
@@
 \end{align*}
-The model equation shows that, for instance, for participant \emph{i} = 1 and digit \emph{j} = 7, the mean RT is modeled by: \(\gamma_{i = 1} - \frac{1}{2} \beta_{i = 1} + \delta_{7,i = 1}\).
-If all \(\delta\)s are positive, then the model equation indicates that individuals respond fastest when the digits 2 and 8 are presented, as in this situation there are no digit effects in the equation.
+\textcolor{blue}{The model equation shows that, for instance, for participant $i = 1$ and digit $j = 7$, with $\gamma_{i = 1}$ equal to $0.5$, $\beta_{i = 1}$ equal to $0.2$, $\delta_{7, i = 1}$ and $\delta_{3, i = 1}$ equal to $0.25$, and $\delta_{6, i = 1}$ and $\delta_{4, i = 1}$ equal to $0.3$, the mean RT is modeled by: $0.5 - 0.1 +  0.25$.
+If all $\delta$s are positive, then the model equation indicates that individuals respond fastest when the digits 2 and 8 are presented, as in this situation there are no digit effects in the equation.
+However, for digit $j = 8$, the mean RT is modeled by $0.5 - 0.1$ in this scenario, while for digit $j = 2$, the mean RT is modeled by $0.5 + 0.1$.}
 
@@
 The \texttt{R} package \texttt{brms} also allows fitting Bayesian models using \texttt{Stan}.
-However, model specification and model fitting can be achieved using a much simplified \texttt{R}-script.
+However, model specification and model fitting can be achieved using a much simplified \texttt{R}-script \textcolor{blue}{alone}.
 Model specification is based on a formula syntax that is similar to the notation in the popular \texttt{lme4} package for frequentist hierarchical modeling (Bates, Mächler, Bolker, \& Walker, 2015).
@@
 Finally, we compare the estimated density of the posterior to the density of the prior.
+\textcolor{blue}{This process is explained in further detail by Van Ravenzwaaij and Etz} (2021).
 
@@
 Next, we calculate the sum of all these Bayes factors and divide each Bayes factor by this sum.
-This results in the posterior model probabilities shown in Table 5.\footnote{Note that it would be possible to calculate the posterior model probabilities directly from the marginal likelihoods.
+This results in the posterior model probabilities shown in Table \ref{tab:postmodprob}.\footnote{Note that it would be possible to calculate the posterior model probabilities directly from the marginal likelihoods.
 However, marginal likelihoods are often not computed explicitly in the model fitting process (e.g., when using the bayes\_factor function in the \texttt{brms} package), so we believe that readers might find it easier to compute posterior model probabilities from Bayes factors.} All posterior model probabilities from the table and the posterior probability of the null model (see table note) sum to one.
@@
 While individual differences are a common phenomenon in psychology, there are instances where general effects are sufficient to describe the data adequately, or where only a random intercept, but no random slopes are necessary.
-As shown in Table 5, Bayesian model comparisons can help researchers to identify the model that best describes the data.
+As shown in Table \ref{tab:postmodprob}, Bayesian model comparisons can help researchers to identify the model that best describes the data.
 
@@
 
+\leavevmode\hypertarget{ref-van2021simulation}{}%
+Ravenzwaaij, D.
+van, \& Etz, A.
+(2021).
+Simulation studies as a tool to understand bayes factors.
+\emph{Advances in Methods and Practices in Psychological Science}, \emph{4}(1), 2515245920972624.
+doi:\href{https://doi.org/10.1177/2515245920972624}{10.1177/2515245920972624}
+
 \leavevmode\hypertarget{ref-roos2015sensitivity}{}%

### 2022-10-26 BHM_correlation/Brms/brms_correlation.R

@@ -1,0 +1,337 @@
+# New brms fit correlation
+
+library(brms)
+
+################################################################################
+## 1. Load data 
+################################################################################
+indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
+colnames(indat)=c('sub','block','trial','stim','resp','rt','error')
+
+## Cleaning the data according to criteria discussed in paper 
+clean=function()
+{
+  indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
+  colnames(indat)=c('sub','block','trial','stim','resp','rt','error')
+  
+  bad1=indat$sub%in%c(34,43)
+  bad2=indat$rt<250 | indat$rt>2000
+  bad3=indat$err==1
+  bad4=indat$block==0 & indat$trial<20
+  bad5=indat$trial==0
+  
+  bad=bad1 | bad2 | bad3 |bad4 |bad5
+  dat=indat[!bad,]
+  return(dat)
+}
+
+indat1 <- clean()  # Final dataset 
+
+## Data add necessary information 
+### Add participant number 
+rank(unique(indat1$sub))
+sub <- rep(1:N, each = J)
+true.theta[sub]
+
+# create variable inter for the intercept gamma 
+indat1$inter <- rep(1, nrow(indat1))
+
+for (j in 1:nrow(indat1)){
+  if (j == 1) {
+    indat1$sub1[j] <- 1}
+  else if (indat1$sub[j] == indat1$sub[j-1]) {
+    indat1$sub1[j] <- indat1$sub1[j-1]}
+  else indat1$sub1[j] <- indat1$sub1[j-1] + 1
+}
+
+### Add variable with information of the side of the digit to the list, for parameter beta    
+# j = condition, if j > 5 than x = -1/2, if j < 5 than x = 1/2 
+# indat1$stim 0 = 2, 1=3, 2=4, 3=6, 4=7, 5=8 
+# so if bigger than 2 than 1/2, else than -1/2 
+for (j in 1:nrow(indat1)){
+  if (indat1$stim[j] < 3) {
+    indat1$side[j] <- 1/2}
+  else indat1$side[j] <- -1/2
+}
+
+### Add variable with information about difference between digits to the list, for parameters delta's 
+#### Difference 8 and 7 
+for (j in 1:nrow(indat1)){ 
[…]

### 2022-10-26 BHM_correlation/Rstan/BHM_rstan_correlation.R

@@ -1,0 +1,224 @@
+# Estimation rstan multilevel with correlation between random effects 
+
+# Fitting an Bayesian Hierarchical Model using Stan - An example 
+
+################################################################################
+## 1. Load data 
+################################################################################
+indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
+colnames(indat)=c('sub','block','trial','stim','resp','rt','error')
+
+## Cleaning the data according to criteria discussed in paper 
+clean=function()
+{
+  indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
+  colnames(indat)=c('sub','block','trial','stim','resp','rt','error')
+  
+  bad1=indat$sub%in%c(34,43)
+  bad2=indat$rt<250 | indat$rt>2000
+  bad3=indat$err==1
+  bad4=indat$block==0 & indat$trial<20
+  bad5=indat$trial==0
+  
+  bad=bad1 | bad2 | bad3 |bad4 |bad5
+  dat=indat[!bad,]
+  return(dat)
+}
+
+indat1 <- clean()  # Final dataset 
+
+## Data into list and fill the list with necessary information 
+datareal <- list(N = length(unique(indat1$sub)))
+datareal$All <- nrow(indat1)  # number of total observations, adding to the list for the analysis  
+
+### Indicating which observation belongs to which person 
+sub <- rep(1:N, each = J)
+
+# Need a variable with group number that are crhonical (that doesn't skip numbers)
+for (j in 1:nrow(indat1)){
+  if (j == 1) {
+    indat1$sub1[j] <- 1}
+  else if (indat1$sub[j] == indat1$sub[j-1]) {
+    indat1$sub1[j] <- indat1$sub1[j-1]}
+  else indat1$sub1[j] <- indat1$sub1[j-1] + 1
+}
+
+
+datareal$group_inds <- indat1$sub1  # Add variable with indicator person to the list used for the analysi
+datareal$y <- indat1$rt/1000  # rt in seconds instead of miliseconds, added to the list for analysis  
+
+### Add variable with information of the side of the digit to the list, for parameter beta    
+# j = condition, if j > 5 than x = -1/2, if j < 5 than x = 1/2 
[…]

### 2022-10-26 BHM_correlation/Brms/brms_correlation.R
Update brms_correlation.R

@@ -127,7 +127,8 @@ priorsmodel9_cor <- c(set_prior("normal(0.5,1)", class = "b", lb = 0, nlpar = "b
                   set_prior("inv_gamma(13.8,5.3)", class = "sd", coef="dif3", group = "sub1",  nlpar = "a"),
                   set_prior("normal(0,0.3)", class = "b", coef="dif4",  nlpar = "a"),  
                   set_prior("inv_gamma(13.8,5.3)", class = "sd", coef="dif4", group = "sub1",  nlpar = "a"),
-                  set_prior("inv_gamma(13.8, 6.3)", class = "sigma"))
+                  set_prior("inv_gamma(13.8, 6.3)", class = "sigma"), 
+                  set_prior("lkj_corr_cholesky(1)", class = "L"))
 
 # Fit Model 
 bform_cor <- brms::bf(rt2 ~ a + b, nl = TRUE, 

### 2022-11-13 manuscript.Rmd
Logmodel comparison

@@
   - name          : "Julia M. Haaf"
-    affiliation   : "2"
+    affiliation   : "3"
     corresponding : no    # Define only one corresponding author
@@
   - id            : "2" 
+    institution   : "Universität der Bundeswehr München"
+  - id            : "3"
     institution   : "University of Amsterdam"
@@
 
-In hierarchical analyses, the general or overall effect is called fixed effect, and the individual-specific deviations from this effect are called random effects (Rouder & Lu, 2005).
+\textcolor{blue}{In hierarchical analyses, the general or overall effect is sometimes called fixed effect, and the individual-specific deviations from this effect are in that case referred to as random effects (Rouder \& Lu, 2005).
 We can also estimate individual effects, not as deviations from an overarching effect, but as effects for each individual, as obtained from a per-participant analysis.
-In this tutorial, we will use the terms general effect, individual deviation, and individual effects instead of fixed and random effects, because we believe these terms are better suited for the interpretation of our analyses.
+Therefore, more common terms that suit the interpretation of the analyses better are general effect, individual deviation, and individual effects instead of fixed and random effects.
+We will use the terms general effect, individual deviation, and individual effect throughout this tutorial.}
 
@@
 It rather corresponds to the left side of Figure 6.
-The two panels of Figure \@ref(fig:modelestimatesfigure3) show that the estimates are almost identical between `Stan` and `brms`.
-This means that differences in the parametrization between the packages did not influence the results of the parameter estimation.
 
@@
 \begin{equation}
-\text{BF}_{\text{nl}} = \frac{\int_{-\infty}^{\infty}P(\bm{Y} | \mathcal{M}_\text{n}, \bm{\theta_{\text{n}}})f(\bm{\theta_{\text{n}}})d\bm{\theta_{\text{n}}}}{\int_{-\infty}^{\infty}P(\bm{Y} | \mathcal{M}_\text{l}, \bm{\theta_{\text{l}}})f(\bm{\theta_{\text{l}}})d\bm{\theta_{\text{l}}}}.
+\text{BF}_{\text{nl}} = \frac{P(\bm{Y} | \mathcal{M}_n)}{P(\bm{Y} | \mathcal{M}_{\text{l}}) = \frac{\int_{-\infty}^{\infty}P(\bm{Y} | \mathcal{M}_\text{n}, \bm{\theta_{\text{n}}})f(\bm{\theta_{\text{n}}})d\bm{\theta_{\text{n}}}}{\int_{-\infty}^{\infty}P(\bm{Y} | \mathcal{M}_\text{l}, \bm{\theta_{\text{l}}})f(\bm{\theta_{\text{l}}})d\bm{\theta_{\text{l}}}}.
 \end{equation}
@@
 For the non-technical user, we recommend using `brms` because of its intuitive model specification.
-However, the default prior setup can be unsuitable for a specific research context and might need to be adapted.
+However, the default prior setup can be unsuitable for a specific research context and (virtually) always needs to be adapted.
 Additionally, users are limited to predefined parametrizations.

### 2022-11-13 manuscript.tex
Logmodel comparison

@@
 
-In hierarchical analyses, the general or overall effect is called fixed effect, and the individual-specific deviations from this effect are called random effects (Rouder \& Lu, 2005).
+\textcolor{blue}{In hierarchical analyses, the general or overall effect is sometimes called fixed effect, and the individual-specific deviations from this effect are in that case referred to random effects (Rouder \& Lu, 2005).
 We can also estimate individual effects, not as deviations from an overarching effect, but as effects for each individual, as obtained from a per-participant analysis.
-In this tutorial, we will use the terms general effect, individual deviation, and individual effects instead of fixed and random effects, because we believe these terms are better suited for the interpretation of our analyses.
+Therefore, more common terms that suit the interpretation of the analyses better are general effect, individual deviation, and individual effects instead of fixed and random effects.
+We will use the terms general effect, individual deviation, and individual effect throughout this tutorial.}
 
@@
 It rather corresponds to the left side of Figure 6.
-The two panels of Figure \ref{fig:modelestimatesfigure3} show that the estimates are almost identical between \texttt{Stan} and \texttt{brms}.
-This means that differences in the parametrization between the packages did not influence the results of the parameter estimation.
 
@@
 For the non-technical user, we recommend using \texttt{brms} because of its intuitive model specification.
-However, the default prior setup can be unsuitable for a specific research context and might need to be adapted.
+However, the default prior setup can be unsuitable for a specific research context and (virtually) always needs to be adapted.
 Additionally, users are limited to predefined parametrizations.

### 2022-11-13 BHM_correlation/Brms/brms_correlation.R

@@ -137,7 +137,7 @@ bform_cor <- brms::bf(rt2 ~ a + b, nl = TRUE,
 )
 
 # Adjust 'myPath' to your directory 
-brmmodel9_new21102022 <- brm(formula = bform_cor,         # Model formula 
+brmmodel9_new26102022 <- brm(formula = bform_cor,         # Model formula 
                              data = indat1,              # Data frame with variables + dependent variable 
                              family = gaussian(),        # Response distribution 
                              prior = priorsmodel9_cor,       # Priors 
@@ -149,6 +149,10 @@ brmmodel9_new21102022 <- brm(formula = bform_cor,         # Model formula
                              #control = list(adapt_delta = 0.95),  # Control sampler's behavior, this avoided the problem of convergement 
                              # Finally we can save the model fit to a file 
                              file = "/myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R code/BRMSmodelfit19_21102022")
+
+# !!! Check this prior: # student_t(3, 0, 2.5)    sd                           a                default
+# student_t(3, 0, 2.5)    sd                           b                default
+
 
 # Warning messages brmmodel9_new21102022:
 #   1: The largest R-hat is 1.74, indicating chains have not mixed.

### 2022-11-13 BHM_correlation/Rstan/BHM_rstan_correlation.R

@@ -179,13 +179,71 @@ saveRDS(hier_modelc_cor_25102022, "hier_modelc_cor_25102022.rds")
 
 
 # With the different code 
-hier_modelc_cor2_25102022 <- stan(file = "/Users/myrtheveenman/myModel_cor2.stan", 
+hier_modelc_cor2_12112022 <- stan(file = "/Users/myrtheveenman/Documents/GitHub/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/BHM_correlation/Rstan/myModel_cor2.stan", 
                                  data = datareal, 
-                                 iter = 6000, chains = 4,
-                                 control = list(max_treedepth = 15, adapt_delta = 0.95),
-                                 warmup = 1000, cores = 4)
+                                 iter = 10000, chains = 4,
+                                 control = list(max_treedepth = 18, adapt_delta = 0.97),
+                                 warmup = 3000, cores = 4)
 
-saveRDS(hier_modelc_cor2_25102022, "hier_modelc_cor2_25102022.rds")
+saveRDS(hier_modelc_cor2_12112022, "hier_modelc_cor2_12112022.rds")
+
+# Warning messages with 10000 iterations, 3000 warmup, control = list(max_treedepth = 18, adapt_delta = 0.97):
+#   1: There were 1023 divergent transitions after warmup. See
+# https://mc-stan.org/misc/warnings.html#divergent-transitions-after-warmup
+# to find out why this is a problem and how to eliminate them. 
+# 2: There were 4 chains where the estimated Bayesian Fraction of Missing Information was low. See
+# https://mc-stan.org/misc/warnings.html#bfmi-low 
+# 3: Examine the pairs() plot to diagnose sampling problems
+# 
+# 4: Bulk Effective Samples Size (ESS) is too low, indicating posterior means and medians may be unreliable.
+# Running the chains for more iterations may help. See
+# https://mc-stan.org/misc/warnings.html#bulk-ess 
+# 5: Tail Effective Samples Size (ESS) is too low, indicating posterior variances and tail quantiles may be unreliable.
+# Running the chains for more iterations may help. See
+# https://mc-stan.org/misc/warnings.html#tail-ess 
+
+# When running pairs()
+# the following parameters were dropped because they are constant
+# L[1,1] L[1,2] L[1,3] L[2,3] L[1,4] L[2,4] L[3,4] L[1,5] L[2,5] L[3,5] L[4,5] L[1,6] L[2,6] L[3,6] L[4,6] L[5,6]
+
+
[…]

### 2022-11-13 R code/Figures/Fig10_randomparfigure_trunc_ordered_per_individuals_rev.R

@@ -1,0 +1,346 @@
+# Individual effects from Rstan 
+# Order on participant 
+# For rstan the output is saved in hier_modelc_adj_07012022 (adjust "myPath" below to your directory)
+hier_modelc_adj_07012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
+
+# For brms the output is saved in object brmmodel8_new03012022 (this is without saving the prior samples)
+brmmodel8_new03012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/brms/Posterior/brmmodel8_new03012022.rds")  
+
+library(rstan)
+library(brms)
+library(gridExtra)
+
+indeffplot_orderedind <- function(hier_modelc_adj_v2, brmmodel8_v2){
+  
+  # Only gamma's 
+  gamnamesstan <- paste("gammai", "[",1:52,"]", sep="")
+  gamstan <- as.data.frame(summary(hier_modelc_adj_v2, pars = c( gamnamesstan))$summary)
+  gamstanmu <- as.data.frame(summary(hier_modelc_adj_v2, pars = c("mu"))$summary)
+  gamstan$parameters <- c(gamnamesstan)
+  
+  for (i in 1:nrow(gamstan)){  # 1 if CI contains zero, 0 otherwise 
+    if (gamstan$`2.5%`[i]*gamstan$`97.5%`[i] < 0) {
+      gamstan$zero2[i] <- "pink"
+    } else (gamstan$zero2[i] <- "blue") 
+  }
+  
+  
+  gamstan2 <- gamstan[order(gamstan$mean),]
+  gamstan$order <- c(1:52)
+  
+  gammarandomplot <- ggplot(gamstan, aes(x = order, y = mean, ymin = `2.5%`, ymax = `97.5%`)) + 
+    geom_linerange(aes(color = "blue"), color = "#00BFC4") + 
+    geom_point() + 
+    geom_hline(yintercept = gamstanmu$mean, linetype = "dashed", alpha = .3) + 
+    labs(title = expression(~gamma[~i]), y = "Estimation") + 
+    theme(legend.position = "none", 
+          axis.title.x=element_blank(),
[…]

### 2022-11-13 rstan_modelcomparison_normalvslognormalmodel.R
Create rstan_modelcomparison_normalvslognormalmodel.R

@@ -1,0 +1,21 @@
+# Normal vs log normal BF comparison 
+
+# Two digit models scenario 3 (without individual differences) 
+
+
+normal_fit3 <- stan(file = "/Users/myrtheveenman/Documents/GitHub/multilevel-tutorial-paper/R objects/rstan/linear model/rstan_anova_digit_trunc.stan", data = datareal, 
+             iter = 4000, chains = 4, warmup = 1000, cores = 4)
+
+saveRDS(normal_fit3, "modelcomparison_normal_fit3.stan")
+
+lognormal_fit3 <- stan(file = "/Users/myrtheveenman/Documents/GitHub/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/rstan/log-normal model/rstan_anova_digit_log.stan", data = datareal, 
+             iter = 4000, chains = 4, warmup = 1000, cores = 4)
+
+saveRDS(lognormal_fit3, "modelcomparison_normal_fit3.stan")
+
+
+bs_normal3 <- bridge_sampler(normal_fit3)
+bs_lognormal3 <- bridge_sampler(lognormal_fit3)
+
+bridgesampling::bf(bs_normal3, bs_lognormal3)[1]$bf
+# Estimated Bayes factor in favor of bs_normal3 over bs_lognormal3: 0.00000 

### 2022-11-14 manuscript.tex
Clarification code

@@
   pdftitle={Bayesian Hierarchical Modeling: An Introduction and Reassessment},
-  pdfauthor={Myrthe Veenman1, Angelika M. Stefan2, \& Julia M. Haaf2},
+  pdfauthor={Myrthe Veenman1, Angelika M. Stefan2, \& Julia M. Haaf3},
   pdflang={en-EN},
@@
 \title{Bayesian Hierarchical Modeling: An Introduction and Reassessment}
-\author{Myrthe Veenman\textsuperscript{1}, Angelika M. Stefan\textsuperscript{2}, \& Julia M. Haaf\textsuperscript{2}}
+\author{Myrthe Veenman\textsuperscript{1}, Angelika M. Stefan\textsuperscript{2}, \& Julia M. Haaf\textsuperscript{3}}
 \date{}
@@
 
-\affiliation{\vspace{0.5cm}\textsuperscript{1} Leiden University\\\textsuperscript{2} University of Amsterdam}
+\affiliation{\vspace{0.5cm}\textsuperscript{1} Leiden University\\\textsuperscript{2} Universität der Bundeswehr München\\\textsuperscript{3} University of Amsterdam}
 
@@
 
-\textcolor{blue}{In hierarchical analyses, the general or overall effect is sometimes called fixed effect, and the individual-specific deviations from this effect are in that case referred to random effects (Rouder \& Lu, 2005).
+\textcolor{blue}{In hierarchical analyses, the general or overall effect is sometimes called fixed effect, and the individual-specific deviations from this effect are in that case referred to as random effects (Rouder \& Lu, 2005).
 We can also estimate individual effects, not as deviations from an overarching effect, but as effects for each individual, as obtained from a per-participant analysis.
@@
 \begin{equation}
-\text{BF}_{\text{nl}} = \frac{\int_{-\infty}^{\infty}P(\bm{Y} | \mathcal{M}_\text{n}, \bm{\theta_{\text{n}}})f(\bm{\theta_{\text{n}}})d\bm{\theta_{\text{n}}}}{\int_{-\infty}^{\infty}P(\bm{Y} | \mathcal{M}_\text{l}, \bm{\theta_{\text{l}}})f(\bm{\theta_{\text{l}}})d\bm{\theta_{\text{l}}}}.
+\text{BF}_{\text{nl}} = \frac{P(\bm{Y} | \mathcal{M}_n)}{P(\bm{Y} | \mathcal{M}_{\text{l}})} = \frac{\int_{-\infty}^{\infty}P(\bm{Y} | \mathcal{M}_\text{n}, \bm{\theta_{\text{n}}})f(\bm{\theta_{\text{n}}})d\bm{\theta_{\text{n}}}}{\int_{-\infty}^{\infty}P(\bm{Y} | \mathcal{M}_\text{l}, \bm{\theta_{\text{l}}})f(\bm{\theta_{\text{l}}})d\bm{\theta_{\text{l}}}}.
 \end{equation}

### 2022-11-14 BHM_correlation/Rstan/BHM_rstan_correlation.R

@@ -151,13 +151,13 @@ datareal$K <- 6  # intercept + side effect + digit effects (4)
 ################################################################################
 library(rstan) 
 # Adjust "myPath" below to your directory  
-hier_modelc_cor_25102022 <- stan(file = "/Users/myrtheveenman/myModel_cor.stan", 
-                                 data = datareal, 
-                                 iter = 4000, chains = 4,
-                                 # , control = list(max_treedepth = 15, adapt_delta = 0.95)
-                                 warmup = 1000, cores = 4)
-
-saveRDS(hier_modelc_cor_25102022, "hier_modelc_cor_25102022.rds")
+# hier_modelc_cor_25102022 <- stan(file = "/Users/myrtheveenman/myModel_cor.stan", 
+#                                  data = datareal, 
+#                                  iter = 4000, chains = 4,
+#                                  # , control = list(max_treedepth = 15, adapt_delta = 0.95)
+#                                  warmup = 1000, cores = 4)
+# 
+# saveRDS(hier_modelc_cor_25102022, "hier_modelc_cor_25102022.rds")
 
 # Warning messages:
 #   1: There were 3157 divergent transitions after warmup. See
@@ -178,7 +178,7 @@ saveRDS(hier_modelc_cor_25102022, "hier_modelc_cor_25102022.rds")
 # https://mc-stan.org/misc/warnings.html#tail-ess 
 
 
-# With the different code 
+# With the different code - this is the one to work on 
 hier_modelc_cor2_12112022 <- stan(file = "/Users/myrtheveenman/Documents/GitHub/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/BHM_correlation/Rstan/myModel_cor2.stan", 
                                  data = datareal, 
                                  iter = 10000, chains = 4,

### 2022-11-15 manuscript.Rmd
Clarification code
sensitivity log normal

@@
 \begin{equation}
-\text{BF}_{\text{nl}} = \frac{P(\bm{Y} | \mathcal{M}_n)}{P(\bm{Y} | \mathcal{M}_{\text{l}}) = \frac{\int_{-\infty}^{\infty}P(\bm{Y} | \mathcal{M}_\text{n}, \bm{\theta_{\text{n}}})f(\bm{\theta_{\text{n}}})d\bm{\theta_{\text{n}}}}{\int_{-\infty}^{\infty}P(\bm{Y} | \mathcal{M}_\text{l}, \bm{\theta_{\text{l}}})f(\bm{\theta_{\text{l}}})d\bm{\theta_{\text{l}}}}.
+\text{BF}_{\text{nl}} = \frac{P(\bm{Y} | \mathcal{M}_n)}{P(\bm{Y} | \mathcal{M}_{\text{l}})} = \frac{\int_{-\infty}^{\infty}P(\bm{Y} | \mathcal{M}_\text{n}, \bm{\theta_{\text{n}}})f(\bm{\theta_{\text{n}}})d\bm{\theta_{\text{n}}}}{\int_{-\infty}^{\infty}P(\bm{Y} | \mathcal{M}_\text{l}, \bm{\theta_{\text{l}}})f(\bm{\theta_{\text{l}}})d\bm{\theta_{\text{l}}}}.
 \end{equation}
@@
 \begin{table}[H]
-\caption{Posterior Model Probabilities}
+\caption{Posterior Model Probabilities Normal Models}
 \label{tab:postmodprob}
@@
 Scenario 3: no individual differences.
-The posterior model probability of the null model is the same in every scenario and equals $0.50*10^{107}$.
+The posterior model probability of the null model is the same in every scenario and equals $0.50*10^{-107}$.
 \end{tablenotes}
@@
 
+### Log-normal Model
+We can take a similar approach for the log-normal model to compare the hierarchical models on varying complexity.
+We evaluate their performance in two scenarios that correspond to scenario 1 and 3 for the normal model.
+This means that in scenario 1 for the log-normal model, we estimated a null, digit, side and full model based on the prior specified in Equation, specifying large individual differences.
+For scenario 2, we remove individual variability from the side, digit, and full model.
+This resulted in the following posterior model probabilities shown in Table \@ref(tab:postmodproblog).
+
+\begin{table}[H]
+\caption{Posterior Model Probabilities Log-normal Models}
+\label{tab:postmodproblog}
+\begin{tabularx}{\textwidth}{XXXX}
+\toprule
+ & \multicolumn{1}{l}{Scenario 1} & \multicolumn{1}{l}{Scenario 3}\\
+\midrule
+Digit model & $26.61 * 10^9$ & $48.55 * 10^{15}$\\
+Side model & $0.25 * 10^{-49}$  & $0.50 * 10^{-52}$ \\
+Full model & $11.02 * 10^7$ &  $96.84 * 10^{14}$\\
+\bottomrule
+\end{tabularx}
+\begin{tablenotes}[para]
+\textit{Note.} Scenario 1: large individual differences.
+Scenario 2: no individual differences.
+The posterior model probability of the null model is the same in every scenario and equals $0.86*10^{-51}$.
+\end{tablenotes}
+\end{table}
+
+The table shows that the digit model without individual digit effects from scenario 2 has the highest posterior model probability.
+This is in line with our findings for the normal model.
+Finally, we can compare the normal model with the highest marginal likelihood to the log-normal with the highest marginal likelihood.
+The normal digit model from scenario 3 to the log-normal model from scenario 2.
+The log marginal likelihood for the normal digit model without individual differences was $6441.04$ while the log marginal likelihood for the log-normal digit model without individual differences was $10637.45$.
+This indicates that the Bayes factor is in favor of the log-normal digit model over the normal digit model.
+
 This example illustrates the benefits of comparing hierarchical models of varying complexity.
 While individual differences are a common phenomenon in psychology, there are instances where general effects are sufficient to describe the data adequately, or where only a random intercept, but no random slopes are necessary.
-As shown in Table \@ref(tab:postmodprob), Bayesian model comparisons can help researchers to identify the model that best describes the data.
+As shown in Tables \@ref(tab:postmodprob) and \@ref(tab:postmodproblog), Bayesian model comparisons can help researchers to identify the model that best describes the data.
 

### 2022-11-15 manuscript.tex
sensitivity log normal

@@
 
+\hypertarget{log-normal-model-3}{%
+\subsubsection{Log-normal Model}\label{log-normal-model-3}}
+
+We can take a similar approach for the log-normal model to compare the hierarchical models on varying complexity.
+We evaluate their performance in two scenarios that correspond to scenario 1 and 3 for the normal model.
+This means that in scenario 1 for the log-normal model, we estimated a null, digit, side and full model based on the prior specified in Equation, specifying large individual differences.
+For scenario 2, we remove individual variability from the side, digit, and full model.
+This resulted in the following posterior model probabilities shown in Table \ref{tab:postmodproblog}.
+
+\begin{table}[H]
+\caption{Posterior Model Probabilities Log-normal Models}
+\label{tab:postmodproblog}
+\begin{tabularx}{\textwidth}{XXXX}
+\toprule
+ & \multicolumn{1}{l}{Scenario 1} & \multicolumn{1}{c}{Scenario 3}\\
+\midrule
+Digit model & $26.61 * 10^9$ & $48.55 * 10^{15}$\\
+Side model & $0.25 * 10^{-49}$  & $0.50 * 10^{-52}$ \\
+Full model & $11.02 * 10^7$ &  $96.84 * 10^{14}$\\
+\bottomrule
+\end{tabularx}
+\begin{tablenotes}[para]
+\textit{Note.} Scenario 1: large individual differences.
+Scenario 2: no individual differences.
+The posterior model probability of the null model is the same in every scenario and equals $0.86*10^{-51}$.
+\end{tablenotes}
+\end{table}
+
+The table shows that the digit model without individual digit effects from scenario 2 has the highest posterior model probability.
+This is in line with our findings for the normal model.
+Finally, we can compare the normal model with the highest marginal likelihood to the log-normal with the highest marginal likelihood.
+The normal digit model from scenario 3 to the log-normal model from scenario 2.
+The log marginal likelihood for the normal digit model without individual differences was \(6441.04\) while the log marginal likelihood for the log-normal digit model without individual differences was \(10637.45\).
+This indicates that the Bayes factor is in favor of the log-normal digit model over the normal digit model.
+
 This example illustrates the benefits of comparing hierarchical models of varying complexity.
 While individual differences are a common phenomenon in psychology, there are instances where general effects are sufficient to describe the data adequately, or where only a random intercept, but no random slopes are necessary.
-As shown in Table \ref{tab:postmodprob}, Bayesian model comparisons can help researchers to identify the model that best describes the data.
+As shown in Tables \ref{tab:postmodprob} and \ref{tab:postmodproblog}, Bayesian model comparisons can help researchers to identify the model that best describes the data.
 

### 2022-11-15 R code/Figures/Fig10_randomparfigure_trunc_ordered_per_individuals_rev.R

@@ -344,3 +344,337 @@ indeffplot_orderedind <- function(hier_modelc_adj_v2, brmmodel8_v2){
 indeffplot_orderedind(hier_modelc_adj_07012022, brmmodel8_new03012022)
 
 
+indeffplot_orderedind2 <- function(hier_modelc_adj_v2, brmmodel8_v2){
+  
+  # Only gamma's 
+  gamnamesstan <- paste("gammai", "[",1:52,"]", sep="")
+  gamstan <- as.data.frame(summary(hier_modelc_adj_v2, pars = c( gamnamesstan))$summary)
+  gamstanmu <- as.data.frame(summary(hier_modelc_adj_v2, pars = c("mu"))$summary)
+  gamstan$parameters <- c(gamnamesstan)
+  
+  for (i in 1:nrow(gamstan)){  # 1 if CI contains zero, 0 otherwise 
+    if (gamstan$`2.5%`[i]*gamstan$`97.5%`[i] < 0) {
+      gamstan$zero2[i] <- "pink"
+    } else (gamstan$zero2[i] <- "blue") 
+  }
+  
+  
+  gamstan2 <- gamstan[order(gamstan$mean),]
+  gamstan2$order <- c(1:52)
+  
+  gammarandomplot <- ggplot(gamstan2, aes(x = order, y = mean, ymin = `2.5%`, ymax = `97.5%`)) + 
+    geom_linerange(aes(color = "blue"), color = "#00BFC4") + 
+    geom_point() + 
+    geom_hline(yintercept = gamstanmu$mean, linetype = "dashed", alpha = .3) + 
+    labs(title = expression(~gamma[~i]), y = "Estimation") + 
+    theme(legend.position = "none", 
+          axis.title.x=element_blank(),
+          axis.text.x=element_blank(),
+          axis.ticks.x=element_blank())
+  # geom_hline(yintercept = 0) + because nothing zero 
+  
+  # Beta  
+  # for rstan you get estimates for parameters, so not  deviation so don't have to add them 
+  betnamesstan <- paste("betai", "[",1:52,"]", sep="")
+  betstan <- as.data.frame(summary(hier_modelc_adj_v2, pars = c( betnamesstan))$summary)
+  betstanmu <- as.data.frame(summary(hier_modelc_adj_v2, pars = c("mu2"))$summary)
[…]

### 2022-11-15 rstan_modelcomparison_normalvslognormalmodel.R

@@ -1,12 +1,174 @@
 # Normal vs log normal BF comparison 
 
-# Two digit models scenario 3 (without individual differences) 
 
+library(plyr)
+library(rstan)
+library(ggplot2)
+library(bridgesampling)
+
+# Preparations 
+# Load datareal 
+indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
+colnames(indat)=c('sub','block','trial','stim','resp','rt','error')
+
+## Cleaning the data according to criteria discussed in paper 
+clean=function()
+{
+  indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
+  colnames(indat)=c('sub','block','trial','stim','resp','rt','error')
+  
+  bad1=indat$sub%in%c(34,43)
+  bad2=indat$rt<250 | indat$rt>2000
+  bad3=indat$err==1
+  bad4=indat$block==0 & indat$trial<20
+  bad5=indat$trial==0
+  
+  bad=bad1 | bad2 | bad3 |bad4 |bad5
+  dat=indat[!bad,]
+  return(dat)
+}
+
+indat1 <- clean()  # Final dataset 
+
+## Data into list and fill the list with necessary information 
+datareal <- list(N = length(unique(indat1$sub)))
+datareal$All <- nrow(indat1)  # number of total observations, adding to the list for the analysis  
+
+### Indicating which observation belongs to which person 
+rank(unique(indat1$sub))
+sub <- rep(1:N, each = J)
+true.theta[sub]
+
+for (j in 1:nrow(indat1)){
+  if (j == 1) {
+    indat1$sub1[j] <- 1}
+  else if (indat1$sub[j] == indat1$sub[j-1]) {
+    indat1$sub1[j] <- indat1$sub1[j-1]}
+  else indat1$sub1[j] <- indat1$sub1[j-1] + 1
+}
+
+
+datareal$group_inds <- indat1$sub1  # Add variable with indicator person to the list used for the analysi
+datareal$y <- indat1$rt/1000  # rt in seconds instead of miliseconds, added to the list for analysis  
+
[…]

### 2022-11-30 manuscript.Rmd
Revisions

@@
    - \newcommand{\Julia}[1]{\todo[inline, color=mypink]{#1}}
+   - \newcommand{\Revision}[1]{\textcolor{blue}{#1}}
    - \usepackage{graphicx}
@@
 Additionally, we may hypothesize that RTs are affected by whether the target digit was smaller or greater than 5, yet it is unclear which side of 5 would lead to faster or slower RTs.
-\textcolor{blue}{The digit classification task is a standard task that usually would be analyzed in a 3 (digit, closest to 5, further away from 5, furthest away from 5) by 2 (side, greater or samller than 5) within-subjects ANOVA.} In this tutorial, we show how inferences can be improved using a hierarchical Bayesian model instead.
+\textcolor{blue}{The digit classification task is a standard task that usually would be analyzed in a 3 (digit, closest to 5, further away from 5, furthest away from 5) by 2 (side, greater or smaller than 5) within-subjects ANOVA.} In this tutorial, we show how inferences can be improved using a hierarchical Bayesian model instead.
 
@@
 
-\textcolor{blue}{In hierarchical analyses, the general or overall effect is sometimes called fixed effect, and the individual-specific deviations from this effect are in that case referred to as random effects (Rouder \& Lu, 2005).
+\Revision{In hierarchical analyses, the general or overall effect is sometimes called fixed effect, and the individual-specific deviations from this effect are in that case referred to as random effects (Rouder \& Lu, 2005).
 We can also estimate individual effects, not as deviations from an overarching effect, but as effects for each individual, as obtained from a per-participant analysis.
@@
 For example, the data in the example are response times.
-The normal model, however, can yield negative values, which seems very problematic.
-In addition, RT distributions are right-skewed.
+\Revision{Response times cannot be negative.
+The normal model, however, can yield negative values.
+When using the normal model for response times, we are assigning probability to negative values that are not possible.
+This seems very problematic.} In addition, RT distributions are right-skewed.
 A better representation of the data at hand is therefore given by a log-normal model [@schramm2019reaction].
@@
 Note that the only change in the model formulation is the change of the likelihood function from a normal distribution to a log-normal distribution.
-The location parameter of the log-normal distribution is described through the same combination of parameters as the mean of the normal model.
+\Revision{The log-normal distribution only assigns probability to positive values, and is, just as the RT distributions, right-skewed.
+Therefore, the log-normal distribution is a better match for the RT data.} The location parameter of the log-normal distribution is described through the same combination of parameters as the mean of the normal model.
 However, it has a different interpretation.
@@
 The only difference between the normal and the log-normal models is the likelihood distribution.
-The normal model just assumes a symmetric normal distibution that can take on any value.
+The normal model just assumes a symmetric normal distribution that can take on any value.
 The log-normal model assumes a log-normal distribution, which is restricted to positive values and has a right skew.
@@
 As before, we prefer to apply conjugate priors.
-The conjugate priors that we applied to the normal model are also conjugate priors for the log-normal model.
-As a starting point, we can therefore use the adjusted priors of the normal model and perform a prior prediction, shown in panel C of Figure \@ref(fig:priorpredfigurepanel).
+\Revision{The conjugate priors that we applied to the normal model are also conjugate priors for the log-normal model [@fink1997compendium].
+For instance, with a normal distribution as prior for the mean of the log-normal model, we assume that the mean of the parameter can be either negative, zero, or positive, and assign a higher probability to values closer to zero.
+Applying this conjugate prior, the posterior of the mean in the log-normal model will result in a normal distribution as well.}
+As a starting point, we can use the adjusted priors of the normal model and perform a prior prediction, shown in panel C of Figure \@ref(fig:priorpredfigurepanel).
 These priors result in a slightly higher variance in RTs than we would expect, and, importantly, the mean response time differences are not centered around zero in all conditions.
@@
 However, compared to the normal model there is considerably more hierarchical shrinkage.
+\Revision{This means that in the log-normal model, the individual effects are corrected more towards the group mean, compared to the normal model.
+Or, the group mean in the log-normal model is not affected as much by individuals showing a divergent trend compared to the normal model.
+Whether we should prefer a model with more shrinkage depends on the theoretical considerations.
+If more individual variation is expected, less shrinkage would be preferred.
+However, the advantage of more shrinkage is that outliers do not influence the estimation of the effect as much.
+}
 
@@
 
-\textcolor{blue}{Note that models models 2-4 are nested in the most complex full model, and the null model is nested in models 2 and 3, but models 2 and 3 are not nested}.
+\Revision{Note that models models 2-4 are nested in the most complex full model, and the null model is nested in models 2 and 3, but models 2 and 3 are not nested}.
 Additionally, the models differ by several parameters.
@@
 Vasishth [-@schad2019toward] and Schad, Nicenboim, Bürkner, Betancourt, and Vasishth [-@schad2021workflow].
+\Revision{There are also other R software packages for Bayesian multilevel modeling besides `brms` and `rstan`, such as `BayesFactor`[@R-BayesFactor] and `rstanarm`[@rstanarm2020], that we did not illustrate in this tutorial.
+However, most of these other R software packages are a bit more limited, for instance, in model specification and prior settings.}
 

### 2022-11-30 manuscript.tex
Revisions

@@
 \newcommand{\Julia}[1]{\todo[inline, color=mypink]{#1}}
+\newcommand{\Revision}[1]{\textcolor{blue}{#1}}
 \usepackage{graphicx}
@@
 For example, the data in the example are response times.
-The normal model, however, can yield negative values, which seems very problematic.
-In addition, RT distributions are right-skewed.
+\textcolor{blue}{Response times cannot be negative.
+The normal model, however, can yield negative values.
+When using the normal model for response times, we are assigning probability to negative values that are not possible.
+This seems very problematic.} In addition, RT distributions are right-skewed.
 A better representation of the data at hand is therefore given by a log-normal model (Schramm \& Rouder, 2019).
@@
 Note that the only change in the model formulation is the change of the likelihood function from a normal distribution to a log-normal distribution.
-The location parameter of the log-normal distribution is described through the same combination of parameters as the mean of the normal model.
+\textcolor{blue}{The log-normal distribution only assigns probability to positive values, and is, just as the RT distributions, right-skewed.
+Therefore, the log-normal distribution is a better match for the RT data.} The location parameter of the log-normal distribution is described through the same combination of parameters as the mean of the normal model.
 However, it has a different interpretation.
@@
 As before, we prefer to apply conjugate priors.
-The conjugate priors that we applied to the normal model are also conjugate priors for the log-normal model.
-As a starting point, we can therefore use the adjusted priors of the normal model and perform a prior prediction, shown in panel C of Figure \ref{fig:priorpredfigurepanel}.
+\textcolor{blue}{The conjugate priors that we applied to the normal model are also conjugate priors for the log-normal model [@fink1997compendium].
+For instance, with a normal distribution as prior for the mean of the log-normal model, we assume that the mean of the parameter can be either negative, zero, or positive, and assign a higher probability to values closer to zero.
+Applying this conjugate prior, the posterior of the mean in the log-normal model will result in a normal distribution as well.}
+As a starting point, we can use the adjusted priors of the normal model and perform a prior prediction, shown in panel C of Figure \ref{fig:priorpredfigurepanel}.
 These priors result in a slightly higher variance in RTs than we would expect, and, importantly, the mean response time differences are not centered around zero in all conditions.
@@
 However, compared to the normal model there is considerably more hierarchical shrinkage.
+\textcolor{blue}{This means that in the log-normal model, the individual effects are corrected more towards the group mean, compared to the normal model.
+Or, the group mean in the log-normal model is not affected as much by individuals showing a divergent trend compared to the normal model.
+Whether we should prefer a model with more shrinkage depends on the theoretical considerations.
+If more individual variation is expected, less shrinkage would be preferred.
+However, the advantage of more shrinkage is that outliers do not influence the estimation of the effect as much.
+}
 
@@
 \toprule
- & \multicolumn{1}{l}{Scenario 1} & \multicolumn{1}{c}{Scenario 3}\\
+ & \multicolumn{1}{l}{Scenario 1} & \multicolumn{1}{l}{Scenario 3}\\
 \midrule
@@
 Vasishth (2019) and Schad, Nicenboim, Bürkner, Betancourt, and Vasishth (2021).
+\textcolor{blue}{There are also other R software packages for Bayesian multilevel modeling besides `brms` and `rstan`, such as `BayesFactor`[@R-BayesFactor] and `rstanarm`[@rstanarm2020], that we did not illustrate in this tutorial.
+However, most of these other R software packages are a bit more limited, for instance, in model specification and prior settings.}
 

### 2022-12-18 manuscript.Rmd
Updates revision

@@
 We can also estimate individual effects, not as deviations from an overarching effect, but as effects for each individual, as obtained from a per-participant analysis.
-Therefore, more common terms that suit the interpretation of the analyses better are general effect, individual deviation, and individual effects instead of fixed and random effects.
-We will use the terms general effect, individual deviation, and individual effect throughout this tutorial.}
+Throughout this tutorial, we will use the terms that suit the interpretation of our analyses, namely general effect, individual deviation, and individual effect.}
 
@@
 ## Step 3: Prior Predictions 
-A key advantage of model specification with priors is that we can obtain preditions on what the data would look like that are generated from the model.
+A key advantage of model specification with priors is that we can obtain preditions on what the data would look like that are generated from the model \Revision{[@etz2018bayesian; @lee2018determining]}.
 These prior predictions allow us to make the implications of our specified priors more concrete.
@@
 ^[Conjugate priors result in posterior distributions that belong to the same distribution family as the prior distribution [@Fink1997].
-For a normal likelihood, these are the normal distribution for the mean and the inverse gamma distribution for the variance.] Therefore, we choose an inverse gamma distribution.
+For a normal likelihood, these are the normal distribution for the mean and the inverse gamma distribution for the variance.]\Revision{However, in `rstan`, non-conjugate priors are commonly applied [e.g., @gelman2006prior]} Therefore, we choose an inverse gamma distribution.
 We can also be more specific about the expected RTs.
@@
 The credible interval displayed in the figure is the Bayesian version of a confidence interval and provides a measure of uncertainty about the parameter value [@wagenmakers2018bayesian].
+\Revision{"Without considering the prior, a posterior credible interval does not directly allow one to conclude that there is evidence for or against an effect." It merely tells us }
 
@@
 The 95% credible intervals of $\mu_{\delta_{7}}$ and $\mu_{\delta_{3}}$ contain zero.
-Therefore, it is plausible that the response times do not differ substantially from the baseline condition.
-The posterior distribution of the side effect $\mu_{\beta}$ is centered around zero with a posterior mean of `r printnum(fitrstanhiertable3["mu2", "Mean"], digits = 3)` and a 95% credible interval ranging from `r printnum(fitrstanhiertable3["mu2", "Lower Bound"], digits = 3)` to `r printnum(fitrstanhiertable3["mu2", "Upper Bound"], digits = 3)`.
-This indicates a small to non-existent side effect.
+\Revision{Therefore, it is plausible that the response times do not differ substantially from the baseline condition.- is this one necessary to adjust?} The posterior distribution of the side effect $\mu_{\beta}$ is centered around zero with a posterior mean of `r printnum(fitrstanhiertable3["mu2", "Mean"], digits = 3)` and a 95% credible interval ranging from `r printnum(fitrstanhiertable3["mu2", "Lower Bound"], digits = 3)` to `r printnum(fitrstanhiertable3["mu2", "Upper Bound"], digits = 3)`.
+\Revision{This indicates a small to non-existent side effect.
+-change}
 
@@
 
-Finally, Figure \@ref(fig:modelestimatesfigure3) shows how the relationship between individuals' digit effects.^[The design and code for this figure are by @j_van_langen_2020_3715576.] The posterior means of individual effects are represented by the dots.
+\Revision{The individual variation can also be visualized in a different way with focusing on the relation between the individual estimates.} Figure \@ref(fig:modelestimatesfigure3) shows how the relationship between individuals' digit effects.^[The design and code for this figure are by @j_van_langen_2020_3715576.] The posterior means of individual effects are represented by the dots.
 The right side of each plot represents the distribution of the individual estimates.

### 2022-12-18 manuscript.tex
Updates revision

@@
 
-Finally, Figure \ref{fig:modelestimatesfigure3} shows how the relationship between individuals' digit effects.\footnote{The design and code for this figure are by Langen (2020).} The posterior means of individual effects are represented by the dots.
+The individual variation can also be visualized in slightly different way, while focusing on the relation between the individual estimates.
+Figure \ref{fig:modelestimatesfigure3} shows how the relationship between individuals' digit effects.\footnote{The design and code for this figure are by Langen (2020).} The posterior means of individual effects are represented by the dots.
 The right side of each plot represents the distribution of the individual estimates.

### 2022-12-18 BHM_correlation/Rstan/BHM_rstan_correlation.R

@@ -181,19 +181,39 @@ library(rstan)
 
 
 # With the different code - this is the one to work on 
-hier_modelc_cor2_12112022 <- stan(file = "myModel_cor2.stan", 
+hier_modelc_cor2_06122022 <- stan(file = "/Users/myrtheveenman/Documents/GitHub/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/BHM_correlation/Rstan/myModel_cor2.stan", 
                                  data = datareal, 
-                                 iter = 2000, chains = 4,
+                                 iter = 5000, chains = 4,
                               #  control = list(max_treedepth = 15, adapt_delta = 0.97),
-                                 warmup = 1000, cores = 4)
+                                 warmup = 2500, cores = 4)
 
 saveRDS(hier_modelc_cor2_12112022, "hier_modelc_cor2_12112022.rds")
 
-summary(hier_modelc_cor2_12112022, pars = c("beta", "sigma_p", "sigma", "L"))$summary
-traceplot(hier_modelc_cor2_12112022, pars = "beta")
-traceplot(hier_modelc_cor2_12112022, pars = "sigma_p")
-traceplot(hier_modelc_cor2_12112022, pars = "L")
-traceplot(hier_modelc_cor2_12112022, pars = "sigma")
+summary(hier_modelc_cor2_06122022, pars = c("beta", "sigma_p", "sigma", "L"))$summary
+rstan::traceplot(hier_modelc_cor2_06122022, pars = "beta")
+rstan::traceplot(hier_modelc_cor2_06122022, pars = "sigma_p")
+rstan::traceplot(hier_modelc_cor2_06122022, pars = "L")
+rstan::traceplot(hier_modelc_cor2_06122022, pars = "sigma")
+
+# Warning messages 5000 iterations 2500 warmup:
+#   1: There were 358 divergent transitions after warmup. See
+# https://mc-stan.org/misc/warnings.html#divergent-transitions-after-warmup
+# to find out why this is a problem and how to eliminate them.
+# 2: There were 2506 transitions after warmup that exceeded the maximum treedepth. Increase max_treedepth above 10. See
+# https://mc-stan.org/misc/warnings.html#maximum-treedepth-exceeded
+# 3: There were 4 chains where the estimated Bayesian Fraction of Missing Information was low. See
+# https://mc-stan.org/misc/warnings.html#bfmi-low
+# 4: Examine the pairs() plot to diagnose sampling problems
+# 
[…]

### 2023-04-05 manuscript.Rmd
Update manuscript.Rmd

@@
 Note that the only change in the model formulation is the change of the likelihood function from a normal distribution to a log-normal distribution.
-\Revision{The log-normal distribution only assigns probability to positive values, and is, just as the RT distributions, right-skewed.
+\Revision{The likelihood function provides the likelihood of parameter values in the form of probability densities.
+The log-normal distribution only assigns probability to positive data values, and is, just as the RT distributions, right-skewed.
 Therefore, the log-normal distribution is a better match for the RT data.} The location parameter of the log-normal distribution is described through the same combination of parameters as the mean of the normal model.
@@
 Another important issue is that wide priors are often improper, that is, the probability density does not integrate to one [@hobert1996effect].
-An example of an improper prior is a uniform prior distribution ranging from minus infinity to infinity.
+\Revision{The integral is used as an approximation of the sum for continuous parameters to calculate the probability density.} An example of an improper prior is a uniform prior distribution ranging from minus infinity to infinity.
 In case of improper priors, an estimation of the general and individual effects cannot be provided.
@@
 ## Step 3: Prior Predictions 
-A key advantage of model specification with priors is that we can obtain preditions on what the data would look like that are generated from the model \Revision{[@etz2018bayesian; @lee2018determining]}.
+A key advantage of model specification with priors is that we can obtain predictions on what the data would look like that are generated from the model \Revision{[@etz2018bayesian; @lee2018determining]}.
 These prior predictions allow us to make the implications of our specified priors more concrete.
@@
 The credible interval displayed in the figure is the Bayesian version of a confidence interval and provides a measure of uncertainty about the parameter value [@wagenmakers2018bayesian].
-\Revision{"Without considering the prior, a posterior credible interval does not directly allow one to conclude that there is evidence for or against an effect." It merely tells us }
+\Revision{In contrast to the confidence interval, the credible interval provides the probability of a range of values for an effect.
+However, considering the prior and under the assumption that there is an effect, the credible interval does not provide evidence to conclude whether there is an effect [Add Wagenmakers - principle of predictable irrelevance].
+}
 

### 2023-04-06 BHM_correlation/Rstan/BHM_rstan_correlation.R
changing model with correlation to non-centered parameterization

@@ -128,6 +128,7 @@ dd3 <- 3  # mean for variance of delta, adjusted
 f3 <- 0.5   # variance for variance of delta, adjusted 
 
 
+
 datareal$a <- a 
 datareal$b <- b 
 datareal$c <- c 
@@ -181,19 +182,19 @@ library(rstan)
 
 
 # With the different code - this is the one to work on 
-hier_modelc_cor2_12112022 <- stan(file = "myModel_cor2.stan", 
+hier_modelc_cor_06042023 <- stan(file = "myModel_cor_julia.stan", 
                                  data = datareal, 
-                                 iter = 2000, chains = 4,
+                                 iter = 2500, chains = 4,
                               #  control = list(max_treedepth = 15, adapt_delta = 0.97),
                                  warmup = 1000, cores = 4)
 
-saveRDS(hier_modelc_cor2_12112022, "hier_modelc_cor2_12112022.rds")
+saveRDS(hier_modelc_cor_06042023, "hier_modelc_cor_06042023.rds")
 
-summary(hier_modelc_cor2_12112022, pars = c("beta", "sigma_p", "sigma", "L"))$summary
-traceplot(hier_modelc_cor2_12112022, pars = "beta")
-traceplot(hier_modelc_cor2_12112022, pars = "sigma_p")
-traceplot(hier_modelc_cor2_12112022, pars = "L")
-traceplot(hier_modelc_cor2_12112022, pars = "sigma")
+summary(hier_modelc_cor_06042023, pars = c("beta", "s2_p", "sigma", "Omega"))$summary
+traceplot(hier_modelc_cor_06042023, pars = "beta")
+traceplot(hier_modelc_cor_06042023, pars = "s2_p")
+traceplot(hier_modelc_cor_06042023, pars = "L")
+traceplot(hier_modelc_cor_06042023, pars = "sigma")
 
 # Warning messages with 10000 iterations, 3000 warmup, control = list(max_treedepth = 18, adapt_delta = 0.97):
 #   1: There were 1023 divergent transitions after warmup. See

### 2023-04-06 propriety_section.Rmd
Adding sections with revisions in separate document

--- 2022-11-30 manuscript.Rmd
@@
    - \usepackage{hyperref}
+   - \newenvironment{CSLReferences}{}{\par}
 
@@
 
-```{r setup, include = FALSE}
-library(rmarkdown)
-library("papaja")
-library(knitr)
-library(plyr)
-library(dplyr)
-library(ggplot2)
-library(rstan)
-library(brms)
-library(LaplacesDemon)
-library(gridExtra)
-library(ggridges)
-library(BayesFactor)
-library(crch)
-library(foreign)
-library(tibble)
-library(bridgesampling)
-library(cowplot)
-library(kableExtra)
-library(readr)
-library(lattice)
-library(Rmisc)
-library(devtools)
-library(gghalves)
-library(bayesplot)
-library(ggmcmc)
-library(lme4)
-library(logspline)
-library(truncnorm)
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
 ```
 
-```{r analysis-preferences}
-# Seed for random number generation
-set.seed(42)
-knitr::opts_chunk$set(cache.extra = knitr::rand_seed, fig.pos = "H")
+### Section in the Intro about out tutorial:
+
+Our tutorial is directed at researchers who have a basic \Revision{conceptual} understanding of Bayesian inference and parameter estimation and are planning to perform more complex Bayesian analyses \Revision{of psychological phenomena.
+We hope to achieve a balance of detail and overview---while we want to provide guidance for a full Bayesian workflow (similar to Gelman et al.
+and Schad et al.) we don't shy away from in-depth explanations for the most curious scientists.
+To ensure that a broad audience is able to follow the tutorial, we limit and carefully explain equations.
+However, rudimentary knowledge of calculus is advantageous to fully understand Bayesian analysis.}
+
+### Section about prior propriety
+
+Another important issue is that wide priors are often improper probability distributions [@hobert1996effect].
+\Revision{A prior is proper if two conditions hold: 1.
+All values of the probability distribution are positive definite, and 2.
+the probability distribution sums to one for discrete data or integrates to one for continuous data.
+An example of a proper and improper probability distribution are provided in Figure XXX.
+The figure shows two hypothetical priors for parameter $\theta$.
+The first condition can be assessed by checking the y-values---both priors have only positive definite values as function output (0.1, 0.2, and 0.4).
+To assess the second condition, we need to sum up the function values.
+Panel A shows a proper distribution where the values sum to 1; panel B shows an improper distribution where the values sum to 1.2.}
+
+```{r, echo = F, fig.width=7, fig.height=3.5, fig.align='center', fig.cap="Two hypothetical priors for a discrete parameter $\\theta$.
+A.
+A proper probability distribution that sums to one.
+B.
+An improper probability distribution that sums to 1.2."}
+layout(matrix(1:2, nrow = 1))
+par(mgp = c(2, .7, 0))
+x <- 1:5
+y <- c(12, 6, 6, 3, 3)
+y <- y / sum(y)
+
+plot(x, y, pch = 19, col = "gray40", axes = F, ylim = c(0, 1), ylab = "Probability", xlab = expression(theta), main = "A. Proper", xlim = c(.5, 5.5))
+axis(1, x)
+axis(2, c(0, unique(y), 1), las = 2)
+for(i in x) lines(rep(x[i], 2), c(0, y[i]), col = "gray40", lwd = 2)
+
+par(mgp = c(2, .7, 0))
+y[3] <- .4
+
+plot(x, y, pch = 19, col = "gray40", axes = F, ylim = c(0, 1), xlim = c(.5, 5.5), ylab = "Probability", xlab = expression(theta), main = "B. Improper")
+axis(1, x)
+axis(2, c(0, unique(y), 1), las = 2)
+for(i in x) lines(rep(x[i], 2), c(0, y[i]), col = "gray40", lwd = 2)
 ```
 
-\newpage
+A \Revision{more common} example of an improper prior is a uniform prior distribution ranging from minus infinity to infinity.
+In case of improper priors, \Revision{it is often impossible to obtain correct estimation of general and individual effects in hierarchical models because the resulting posteriors will again be improper probability distributions.} However, even proper priors can result in improper posteriors [@hobert1996effect].
+Although estimation should not be possible in this situation, statistical software might still provide results without notifying the user that the posterior distribution does not exist.
+If users are not aware of the impropriety problem, this can result in misleading conclusions.
+Therefore, it is important to check if *(a)* the priors are proper and *(b)* the priors result in proper posteriors.
+\Revision{Unfortunately,} there is no fail-safe method that guarantees proper posteriors for hierarchical models.
+The best approach is to choose well-reasoned informative priors for the model, for instance, by using the proposed four-step procedure below.
+\Revision{Priors that are proper and not too diffuse will in most cases lead to proper posteriors.}
 
-# Introduction
-
-Imagine a typical scenario in the life of an experimental psychology graduate student: You have just collected a large data set, either online or in the lab.
-Each participant has spent one valuable hour completing many trials per experimental condition.
-Now it is time to analyze the data, and your supervisor suggests to run a within-subjects ANOVA.
-This procedure requires you to first aggregate the data by condition and participant.
-Doing so reduces your data set from 10,000 rows to just 400.
-Reducing the data by aggregation makes you a bit queasy: It feels like your efforts in data collection and the participants' time are not valued this way.
-You know there needs to be a better way.
-And there is: Hierarchical modeling.
-
-Just as in the scenario above, experiments in psychology are often implemented in a repeated-measures design where participants respond to several items, stimuli, or conditions.
-From a data analysis perspective, such a design implies that observations are nested within participants.
-The statistical and the psychological literature agree that the optimal analysis accounting for such a data structure is hierarchical modeling [e.g., @efron1977stein; Lee, -@lee2011cognitive; Rouder & Lu, -@rouder2005introduction].
-And yet, it is still not the norm to apply hierarchical modeling to data from psychological experiments.
-One reason for this is that hierarchical modeling is more difficult than familiar procedures such as ANOVA.
-
-With recent software development, hierarchical modeling has become more and more accessible.
-In addition, many researchers have argued that hierarchical modeling is easier and the interpretation is more intuitive when it is done in the Bayesian statistical framework [e.g., Lynch, -@lynch2007introduction; Rouder & Lu, 2005; Rouder, Morey, & Pratte, -@rouder2013hierarchical].
-However, there are challenges to the Bayesian way of hierarchical modeling (Rouder & Lu, 2005).
-These challenges, such as the choice of priors, the need for model comparisons of highly complex models, and programming skills required to use state-of-the-art software solutions, can make researchers hesitant to use this approach.
-We believe that existing introductions to Bayesian hierarchical modeling do not adequately address these challenges.
-For example, they use outdated software (Rouder et al., 2013), do not cover model comparison (Rouder & Lu, 2005), or do not discuss prior specification [@shiffrin2008survey].
-This paper aims to make Bayesian hierarchical modeling available to a wider public by providing:
-
-- A comparison of and guidance for the use of two software packages,
-- Guidance on prior specification, 
-- Practical instructions for Bayes factor model comparison.
-
-Our tutorial is directed at researchers who have a basic understanding of Bayesian inference and parameter estimation and are planning to perform more complex Bayesian analyses.
-
-The application of Bayesian hierarchical analysis will be illustrated using a digit classification task collected by @rouder2005hierarchical.
-In an experimental task participants repeatedly indicated whether a target digit (i.e., 2, 3, 4, 6, 7, or 8) was greater or smaller than 5.
-The participants in the task performed multiple trials per condition (i.e., digit).
-Therefore, observations are nested in participants.
-Our analyses focus on two effects that can potentially occur in this digit classification task: The digit effect and the side effect.
-According to the symbolic distance hypothesis [Moyer & Landauer, -@moyer1967time] processing of numbers is analogue.
-If numbers are closer on the number line, we may confuse them more, and comparing them is more difficult.
-In line with this hypothesis, the digit effect postulates that response times (RTs) are slower when digits are closer to 5.
-Additionally, we may hypothesize that RTs are affected by whether the target digit was smaller or greater than 5, yet it is unclear which side of 5 would lead to faster or slower RTs.
-\textcolor{blue}{The digit classification task is a standard task that usually would be analyzed in a 3 (digit, closest to 5, further away from 5, furthest away from 5) by 2 (side, greater or smaller than 5) within-subjects ANOVA.} In this tutorial, we show how inferences can be improved using a hierarchical Bayesian model instead.
-
-```{r data, include = FALSE, eval = FALSE}
-# maybe add object here where I save the link 
-indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
-colnames(indat)=c('sub','block','trial','stim','resp','rt','error')
-
-hist(indat$rt[indat$sub==0])
-table(indat$rt[indat$sub==0])
-nrow(indat[indat$sub==0,])
-
-unique(indat$stim)
-for (i in unique(indat$stim)){
-hist(indat$rt[indat$stim==i])
-}
-
-clean=function()
-{
-  indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
-  colnames(indat)=c('sub','block','trial','stim','resp','rt','error')
-
-  bad1=indat$sub%in%c(34,43)
-  bad2=indat$rt<250 | indat$rt>2000
-  bad3=indat$err==1
-  bad4=indat$block==0 & indat$trial<20
-  bad5=indat$trial==0
-
-  bad=bad1 | bad2 | bad3 |bad4 |bad5
-  dat=indat[!bad,]
-  return(dat)
-}
-
-# x axis digit 
-# y axis: mean RT pp 
-indat1 <- clean()
-
-meanrtindat <- aggregate(indat1$rt, list(indat1$stim, indat1$sub), mean)  # mean response time per condition per participant 
-meanrt <- aggregate(indat1$rt, list(indat1$stim), mean)
-for (i in 1:nrow(meanrtindat)){
-  if (meanrtindat$Group.1[i] == 0) {meanrtindat$Group.1[i] <- 2 
-  } else if(meanrtindat$Group.1[i] == 1) {meanrtindat$Group.1[i] <- 3 
-  } else if(meanrtindat$Group.1[i] == 2) {meanrtindat$Group.1[i] <- 4
-  } else if(meanrtindat$Group.1[i] == 3) {meanrtindat$Group.1[i] <- 6 
-  } else if(meanrtindat$Group.1[i] == 4) {meanrtindat$Group.1[i] <- 7 
-  } else {meanrtindat$Group.1[i] <- 8} 
-}
-
-for (i in 1:nrow(meanrt)){
-  if (meanrt$Group.1[i] == 0) {meanrt$Group.1[i] <- 2 
-  } else if(meanrt$Group.1[i] == 1) {meanrt$Group.1[i] <- 3 
-  } else if(meanrt$Group.1[i] == 2) {meanrt$Group.1[i] <- 4
-  } else if(meanrt$Group.1[i] == 3) {meanrt$Group.1[i] <- 6 
-  } else if(meanrt$Group.1[i] == 4) {meanrt$Group.1[i] <- 7 
-  } else {meanrt$Group.1[i] <- 8} 
-}
-```
-
-```{r figure1code, echo=FALSE, fig.align='left', fig.cap="Average response times in milliseconds in two conditions (i.e., digit 3 and 4).
-The grey lines represent the individual response time effects, and the blue line the mean aggregated response time effect.", fig.height=4.5, fig.width=6, eval = FALSE}
-theme_set(theme_apa(base_size = 9))
-meanrtplot <- ggplot(meanrtindat, aes(y = x, x = as.factor(Group.1))) + geom_line(aes(group = Group.2), alpha = .15) + 
-  geom_line(data = meanrt, aes(group = 1), color='blue') + geom_point(data = meanrt, aes(group = 1), color='blue') + 
-  labs(x = "Digit", y = "Response Time (Ms)") + theme_bw() + theme(axis.line = element_line(colour = "black"), panel.border = element_blank(), panel.background = element_blank())
-
-# meanrtplot
-
-# Difference 3 and 4 
-dif34indat <- meanrtindat[which(meanrtindat$Group.1 == 3 | meanrtindat$Group.1 == 4),]
-meandif34 <- meanrt[which(meanrt$Group.1 == 3 | meanrt$Group.1 == 4),]
-
-meanrtplot2 <- ggplot(dif34indat, aes(y = x, x = as.factor(Group.1))) + geom_line(aes(group = Group.2), alpha = .15) + geom_point(alpha = .15) + 
-  geom_line(data = meandif34, aes(group = 1), color='blue', size = 1) + geom_point(data = meandif34, aes(group = 1), color='blue', shape = 21, size = 2, stroke = 1) + 
-  labs(x = "Digit", y = "Response Time (Ms)") + theme_bw() + theme(axis.line = element_line(colour = "black"), panel.border = element_blank(), panel.background = element_blank())
-
-meanrtplot2 
-```
-
-```{r figure1, echo=FALSE, fig.align='left', fig.cap="Panel A shows the average response times in seconds for two conditions of the digit classification task (i.e., digits 3 and 4).
-The grey lines represent the average within individual effects, and the blue line the aggregated effect over all individuals.
-Panel B shows how individual estimates are shrunken towards the general trend.
-Sample represents the observed effect per participant in seconds, while Model represents the estimated effects per participant.
-", fig.height=3, fig.width=6, out.width ="95%"}
-knitr::include_graphics("Images/Fig1_meanrtandhierplot_v2.png")
-```
-
-```{r figure1b, echo=FALSE, fig.align='left', fig.cap="RTs per condition (i.e., digit).
-The grey lines represent the individual RTs, and the blue line the mean aggregated response time.", fig.height=4, fig.width=6, eval = FALSE}
-
-meanrtindat$Group.3 <- ifelse(meanrtindat$Group.1 > 5, 1, 0)
-meanside <- with(meanrtindat, tapply(x, Group.3, mean))
-meanside2 <- as.data.frame(t(meanside))
-meanside3 <- data.frame(y = c(meanside2$`0`, meanside2$`1`), type = c("Smaller", "Larger"))
-
-indside <- with(meanrtindat, tapply(x, list(Group.3, Group.2), mean))
-indside2 <- as.data.frame(t(indside))
-indside3 <- data.frame(y = c(indside2$`0`,indside2$`1`), type = rep(c("Smaller", "Larger"), each = nrow(indside2)), group = rep(1:52, 2))
-
-# matplot(indside, type = "b", lty = 1, col = adjustcolor(1, .3), pch = 20
-#         , ylim = c(300, 1200), xlim = c(0.8, 2.2), xaxt = "n")
-# axis(1, 1:2, labels = c("smaller", "larger"))
-# lines(meanside, lwd = 2, type = "b", col = "darkred")
-
-level_order <- c("Smaller", "Larger")
-theme_set(theme_apa(base_size = 9))
-meanrtplot2 <- ggplot(indside3, aes(y = y, x = factor(type, level_order))) + 
-  geom_line(aes(group = group), alpha = .15) + geom_point() + 
-  geom_line(data = meanside3, aes(group = 1), color='blue', size = 1) + 
-  geom_point(data = meanside3, 
-             aes(group = 1), color='blue', shape = 21, size = 2, stroke = 1) + 
-  labs(x = "", y = "Response Time (Ms)") + 
-  theme_bw() + 
-  theme(axis.line = element_line(colour = "black"), 
-        panel.border = element_blank(), 
-        panel.background = element_blank())
-
-# meanrtplot2 
-```
-
-## Why Multilevel Modeling?
-
-Before diving into hierarchical modeling, let us consider two alternative approaches: Analyzing aggregated data and analyzing the data separately for each participant.
-Analyses of aggregated data such as the ANOVA from the scenario above examine a general effect that is assumed to be consistent across subjects.
-For instance, when investigating whether there is an effect of digit on RT, the mean RT for each participant is computed for each digit-condition, and a repeated-measures ANOVA is conducted on these aggregated scores.
-Analyses of aggregated data cannot investigate whether a trend is consistent across participants.
-Therefore, when this option is applied and individual variability is ignored, the general trend can be an inaccurate representation of the true general trend [Rouder, Morey, & Pratte, 2013; Haaf & Rouder, 2017; van Doorn, Aust, Haaf, Stefan, and Wagenmakers, -@van2021bayes].
-This is shown in Panel A of Figure \@ref(fig:figure1) where some of the individual trends (i.e., slopes of the grey lines) differ from the general trend (i.e., slope of the blue line).
-If every participant takes part in each condition, the person variability affects all conditions, thus inducing a correlation across conditions.
-For example, participants in a digit classification task repeatedly respond to target digits inducing a correlation between the responses to the target digits, such as between target digit 3 and target digit 4.
-Therefore, the observations are not independent as we assume if we perform analysis on aggregated data.
-As a result, the type I error rate of a test of the effect increases: there is a higher probability of concluding that there is an effect if there is no effect (Rouder & Lu, 2005).
-
-A second alternative analysis method is to conduct an analysis on each participant's data independently.
-For example, in the digit classification task, we would estimate one digit effect per person (cf.
-the grey lines in Figure 1A), but no overall effect (cf.
-the blue line in Figure 1A).
-Therefore, when analyzing participant data independently, it is hard to draw any general conclusions.
-
-\Revision{In hierarchical analyses, the general or overall effect is sometimes called fixed effect, and the individual-specific deviations from this effect are in that case referred to as random effects (Rouder \& Lu, 2005).
-We can also estimate individual effects, not as deviations from an overarching effect, but as effects for each individual, as obtained from a per-participant analysis.
-Therefore, more common terms that suit the interpretation of the analyses better are general effect, individual deviation, and individual effects instead of fixed and random effects.
-We will use the terms general effect, individual deviation, and individual effect throughout this tutorial.}
-
-If estimated for each participant independently, observed individual effects are quite variable because they are perturbed by sample noise.
-In hierarchical modeling, these individual effects are optimally and automatically corrected towards the general trend [@efron1977stein].
-This phenomenon, also called `shrinkage', is illustrated in Panel B of Figure \@ref(fig:figure1).
-The rightmost observation in the figure is in sample quite large compared to the other observed effects.
-The estimated effect (green point) is drastically corrected towards the mean.
-Conversely, the general trend is only influenced slightly by individuals showing a highly divergent trend.
-By adding the individual variability to the model, we no longer assume independence between observations, but the overall effect in the model leaves the possibility to generalize the effect.
-Therefore, hierarchical analysis can be considered a middle ground between an independent analysis and an analysis of aggregated data.
-```{r figure2, echo=FALSE, eval = FALSE, fig.cap="Sample represents the observed response time in milliseconds per individual, while model represents the individual response time estimates.
-Individual estimates are shrunken towards the general trend.
-", fig.align = "left", fig.height = 2, fig.width = 3, out.width="70%"}
-knitr::include_graphics("Images/hierplotefmor.png")
-```
-
-The Bayesian framework offers an intuitive approach for hierarchical modeling (Rouder & Lu, 2005; Lynch, 2007).
-The flexibility of the Bayesian approach facilitates the implementation of complex hierarchical models (Rouder, Morey, & Pratte, 2013).
-It also allows for the inclusion of prior knowledge into the analysis, as well as the monitoring of evidence for specific hypotheses as the data accumulate [@wagenmakers2018bayesian].
-A complete discussion of the Bayesion approach is beyond the scope of this manuscript.
-We refer the interested reader to Etz and Vandekerckhove [-@etz2018introduction] and Wagenmakers and colleagues [-@wagenmakers2018bayesian].
-
-# Model Specification 
-In the following section, we specify two models that can be used for our application example: a normal and a log-normal model.^[[Online Supplement A](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/blob/main/A%20-%20Model%20Parameterization%20/Online-Supplement-A---Model-Parameterization-for-Aggregated-and-Individual-Models.pdf) provides the parameterization for models of aggregated and individual data, to illustrate the difference with hierarchical models.
-] We are aware that many psychologists do not engage with equations on a daily basis.
-In this section, however, we try to accessibly summarize the key considerations of model specification.
-We encourage everyone to read on and not skip the section.
-Additionally, we provide a non-technical verbal summary.
-
-Throughout the section, we assume that each observation $Y_{ijk}$ is the response time from participant $i$ in trial $k$ of condition $j$.
-The condition $j$ is the digit presented to the participant, and can take the values 2, 3, 4, 6, 7, and 8.
-We further assume that the digit side ($x$) is recorded using effect coding where a value of -0.5 indicates that the digit is smaller than 5, and a value of 0.5 indicates that the digit is larger than 5.
-Additionally, we assume that the digits presented in the trial are dummy-coded in the data using four variables ($u$, $v$, $w$, $z$).
-If digits 2 and 8 were presented all four variables are zero.
-The dummy variables represent the digits 3 ($u$), 4 ($v$), 6 ($w$), and 7 ($z$), and each take the value 1 if the respective digit was presented in the trial, and 0 if it was not.
-
-## Normal Model
-
-In the normal model, the observation $Y_{ijk}$ is assumed to come from a normal distribution.
-The mean of this normal distribution is predicted by the person-specific intercept $\gamma$, the effect of the side (i.e., smaller or greater than 5, $\beta$) and the effect of the distance from 5 (i.e., symbolic distance effect, $\delta$’s).
-The variance is represented by $\sigma^2$.
-This is captured in the following equation:
-
-\begin{equation}
-Y_{ijk} \sim \text{Normal}(\gamma_{i} + x_j \beta_{i} + u_j \delta_{7i} + v_j \delta_{6 i} + w_j \delta_{4i} + z_j \delta_{3i}, \sigma^2), (\#eq:datalevel)
-\end{equation}
-
-where
-
-<!-- \[ -->
-<!-- x_j =  -->
-<!-- \begin{cases} -->
-<!-- \frac{1}{2} & j < 5\\ -->
-<!-- - \frac{1}{2} & j > 5, \\ -->
-<!-- \end{cases} -->
-
-<!-- \; \; \; -->
-
-<!-- \[ -->
-<!-- u_j =  -->
-<!-- \begin{cases} -->
-<!-- 1 & j = 7\\ -->
-<!-- 0 & \text{Otherwise}, -->
-<!-- \end{cases} -->
-
-<!-- \; \; \; -->
-
-<!-- v_j =  -->
-<!-- \begin{cases} -->
-<!-- 1 & j = 6\\ -->
-<!-- 0 & \text{Otherwise}, -->
-<!-- \end{cases} -->
-<!-- \] -->
-
-\begin{align*}
-&
-\begin{aligned}
-x_j = 
-\begin{cases}
-\frac{1}{2} & j < 5\\
-- \frac{1}{2} & j > 5, \\
-\end{cases}
-\end{aligned}
-&
-\begin{aligned}
-u_j = 
-\begin{cases}
-1 & j = 7\\
-0 & \text{Otherwise},
-\end{cases}
-\end{aligned}
-&
-\begin{aligned}
-v_j = 
-\begin{cases}
-1 & j = 6\\
-0 & \text{Otherwise},
-\end{cases}
-\end{aligned}
-\end{align*}
-
-<!-- \[ -->
-<!-- w_j =  -->
-<!-- \begin{cases} -->
-<!-- 1 & j = 4\\ -->
-<!-- 0 & \text{Otherwise}, -->
-<!-- \end{cases} -->
-
-<!-- \; \; \; -->
-
-<!-- z_j =  -->
-<!-- \begin{cases} -->
-<!-- 1 & j = 3\\ -->
-<!-- 0 & \text{Otherwise}. -->
-<!-- \end{cases} -->
-<!-- \] -->
-
-\begin{align*}
-&
-\begin{aligned}
-w_j = 
-\begin{cases}
-1 & j = 4\\
-0 & \text{Otherwise},
-\end{cases}
-\end{aligned}
-&
-\begin{aligned}
-z_j = 
-\begin{cases}
-1 & j = 3\\
-0 & \text{Otherwise}.
-\end{cases}
-\end{aligned}
-\end{align*}
-\textcolor{blue}{The model equation shows that, for instance, for participant $i = 1$ and digit $j = 7$, with $\gamma_{i = 1}$ equal to $0.5$, $\beta_{i = 1}$ equal to $0.2$, $\delta_{7, i = 1}$ and $\delta_{3, i = 1}$ equal to $0.25$, and $\delta_{6, i = 1}$ and $\delta_{4, i = 1}$ equal to $0.3$, the mean RT is modeled by: $0.5 - 0.1 +  0.25$.
-If all $\delta$s are positive, then the model equation indicates that individuals respond fastest when the digits 2 and 8 are presented, as in this situation there are no digit effects in the equation.
-However, for digit $j = 8$, the mean RT is modeled by $0.5 - 0.1$ in this scenario, while for digit $j = 2$, the mean RT is modeled by $0.5 + 0.1$.}
-
-The intercept $\gamma_i$, the side effect $\beta$, and the digit effects $\delta$ are person-specific.
-This means that for these parameters, the model will produce one estimate for each participant.
-The multilevel model assumes that these individual-level effects in turn follow a normal distribution:
-
-\begin{equation}
-\begin{aligned}
-\gamma_{i} &\sim \text{Normal}(\mu_{\gamma}, \sigma^2_{\gamma}), (\#eq:fixedvsrandom)\\
-\beta_{i} &\sim \text{Normal}(\mu_{\beta}, \sigma^2_{\beta}), \\
-\delta_{7i} &\sim \text{Normal}(\mu_{\delta_{7}}, \sigma^2_{\delta_{7}}), \\
-\delta_{6i} &\sim \text{Normal}(\mu_{\delta_{6}}, \sigma^2_{\delta_{6}}), \\
-\delta_{4i} &\sim \text{Normal}(\mu_{\delta_{4}}, \sigma^2_{\delta_{4}}),\\
-\delta_{3i} &\sim \text{Normal}(\mu_{\delta_{3}}, \sigma^2_{\delta_{3}}).\\
-\end{aligned}
-\end{equation}
-The means ($\mu_{\gamma}$, $\mu_{\beta}$, $\mu_{\delta}$) of these normal distributions represent the general expected effect, for example, $\mu_{\beta}$ represents the expected effect of digit side across individuals.
-The variances ($\sigma^2_{\gamma}$, $\sigma^2_{\beta}$, $\sigma^2_{\delta}$) represent the individual variation.
-Therefore, the individual level depends on another level, the overall effect level, containing $\mu_{\gamma}$, $\mu_{\beta}$, $\mu_{\delta}$, $\sigma^2_{\gamma}$, $\sigma^2_{\beta}$, and $\sigma^2_{\delta}$.
-
-## Log-normal Model 
-A normal model may not be appropriate in case of more complex theories, hypotheses, or data structures.
-For example, the data in the example are response times.
-\Revision{Response times cannot be negative.
-The normal model, however, can yield negative values.
-When using the normal model for response times, we are assigning probability to negative values that are not possible.
-This seems very problematic.} In addition, RT distributions are right-skewed.
-A better representation of the data at hand is therefore given by a log-normal model [@schramm2019reaction].
-This means that Equation \@ref(eq:datalevel) has to be adjusted:
-
-\begin{equation}
-Y_{ijk} \sim \text{LogNormal}(\gamma_{i} + x_j \beta_{i} + u_j \delta_{7i} + v_j \delta_{6i} + w_j \delta_{4i} + z_j \delta_{3i}, \sigma^2).  
-\end{equation}
-Note that the only change in the model formulation is the change of the likelihood function from a normal distribution to a log-normal distribution.
-\Revision{The log-normal distribution only assigns probability to positive values, and is, just as the RT distributions, right-skewed.
-Therefore, the log-normal distribution is a better match for the RT data.} The location parameter of the log-normal distribution is described through the same combination of parameters as the mean of the normal model.
-However, it has a different interpretation.
-For example, for a location parameter of $\mu = 0$ and a scale parameter of $\sigma^2 = 1$, the mode of the log-normal distribution is $0.368$, and the mean is $1.65$.
-
-## Summary
-
-In summary, the models are placed directly on the raw RT data.
-The only difference between the normal and the log-normal models is the likelihood distribution.
-The normal model just assumes a symmetric normal distribution that can take on any value.
-The log-normal model assumes a log-normal distribution, which is restricted to positive values and has a right skew.
-
-Both the normal and the log-normal models contain the following parameters: $\gamma_i$ corresponds to the participant's overall response time, $\beta_i$ corresponds to the participant's side effect, and the four $\delta_{\cdot i}$ correspond to the effects between the digits furthest away from 5 (2 or 8) and the other digits.
-These person-specific effects all come from parent distributions.
-This results in the hierarchical structure shown in Figure \@ref(fig:hierarchyplot).
-The figure highlights the relationship between the parameters and the data.
-
-In Bayesian modeling, prior distributions are needed to complete model specification.
-In the next section, we will take you through the process of setting these priors, followed by sections on Bayesian estimation and hypothesis testing for both the normal and log-normal model.
-In every section, we will first explain the normal model, followed by an illustration of the log-normal model.
-
-```{r hierarchyplot, echo=FALSE, fig.cap="Hierarchical structure of the normal and log-normal models for the digit classification task.", fig.align = "left", fig.height = 3, fig.width = 4, out.width="70%"}
-knitr::include_graphics("Images/Fig2_Hierarchical structure.png")
-```
-
-# Prior Settings 
-Bayesian analysis requires the specification of prior distributions.
-Prior distributions are probability distributions on model parameters that specify beliefs about the relative plausibility of parameter values before seeing the data [@wagenmakers2018bayesian].
-These prior beliefs can then be updated with the data to obtain the posterior beliefs about the model parameters, following Bayes' rule [@jeffreys1961theory].
-
-For our models, priors are needed on the general effects ($\mu_{\gamma}$, $\mu_{\beta}$, $\mu_{\delta}$) and on the variances ($\sigma_{\gamma}^2$, $\sigma_{\beta}^2$, $\sigma_{\delta}^2$, and $\sigma^2$), as shown in Equation \@ref(eq:fixedvsrandom).
-For each of these parameters, we specify our beliefs as a probability distribution.
-For example, if we expect small side effects in the digit classification example, the prior distribution on $\mu_{\beta}$ should have its peak close to zero.
-Large values for the variance parameters indicate large individual differences, whereas values close to zero indicate that effects are similar across individuals [@haaf2017developing].
-If we expect little variability of side effects between people, the prior assigned to $\sigma_{\beta}$ should also have its peak close to zero.
-The width of prior distributions expresses how certain we are about the parameters before seeing the data.
-For example, if we want to express that we are certain that the side effect is zero, we can formulate a prior that allocates all probability mass to a parameter value of zero.
-
-## Fear of Commitment  
-In many cases substantive researchers fear committing to prior settings because they are unsure about their choices for the type and settings of the distribution.
-Therefore, they tend to choose wide priors to indicate little prior knowledge about the parameter values [@aczel2018expert].
-These include default priors offered by programming packages, such as `rstan` (Stan Development Team, 2019a; illustrated in [Online Supplement B](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/blob/main/B%20-%20Default%20Priors%20Rstan/Online-Supplement-B---Default-Priors-Rstan.pdf)) and `brms` [@R-brms_a; @R-brms_b].
-Wide priors are priors that allocate (approximately) equal probability density to a great range of values, such as a uniform distribution without bounds [@gelman2006prior].
-However, wide priors can be problematic.
-First, since the distribution is spread out across a wide range of values, the probability density on any specific parameter value is very low.
-Therefore, any effect is unlikely under the prior as a wide range of effects is deemed plausible.
-This influences model comparisons: the support for the null hypothesis becomes increasingly large.
-This is also referred to as the Jeffreys-Lindley paradox [@lindley1957statistical].
-
-Another important issue is that wide priors are often improper, that is, the probability density does not integrate to one [@hobert1996effect].
-An example of an improper prior is a uniform prior distribution ranging from minus infinity to infinity.
-In case of improper priors, an estimation of the general and individual effects cannot be provided.
-However, even proper priors can result in improper posteriors [@hobert1996effect].
-Although estimation should not be possible in this situation, statistical software might still provide results without notifying the user that the posterior distribution does not exist.
-If users are not aware of this problem, this can result in misleading conclusions.
-Therefore, it is important to check if *(a)* the priors are proper and *(b)* the priors result in proper posteriors.
-There is no fail-safe method that guarantees proper posteriors for hierarchical models.
-The best approach is to choose well-reasoned informative priors for the model, for instance, by using the proposed four-step procedure below.
-
-## Procedure 
-We propose four steps to choose suitable priors for Bayesian hierarchical models.
-We first present the steps and then apply them to choose suitable priors for the parameters $\mu_{\gamma}$, $\mu_{\beta}$, $\mu_{\delta}$, $\sigma_{\gamma}^2$, $\sigma_{\beta}^2$, $\sigma_{\delta}^2$ and $\sigma^2$ in our example.
-
-```{r distributionsteps, echo=FALSE, fig.cap = "Prior specification steps.", fig.align='center', warning=FALSE, message=FALSE, fig.height=3, fig.width=4, cache = TRUE}
-# https://jokergoo.github.io/circlize_book/book/graphics.html 
-library(circlize)
-library(unikn)   # https://cran.r-project.org/web/packages/unikn/vignettes/colors.html
-
-# circos.initialize(letters[1:4], xlim = c(0, 1))
-# col = rand_color(4)
-# tail = c("point", "normal", "point", "normal")
-# circos.track(ylim = c(0, 1), panel.fun = function(x, y) {
-#     circos.arrow(x1 = 0, x2 = 1, y = 0.5, width = 0.4, 
-#         arrow.head.width = 0.6, arrow.head.length = ux(1, "cm"), 
-#         col = col[CELL_META$sector.numeric.index], 
-#         tail = tail[CELL_META$sector.numeric.index])
-# }, bg.border = NA, track.height = 0.4)
-
-circos.clear()
-cell_cycle = data.frame(phase = c("1.
-Distribution family", "2.
-Distribution settings", "3.
-Prior predictions", "4.
-Adjust"), levels = c("Distribution type", "Distribution settings", "Prior predictions", "Adjust"),
-                        hour = c(8, 8, 8, 8))
-color = c(NA)
-circos.par(start.degree = 90)
-circos.initialize(cell_cycle$phase, xlim = cbind(rep(0, 4), cell_cycle$hour))
-circos.track(ylim = c(0, 1), panel.fun = function(x, y) {
-    circos.arrow(CELL_META$xlim[1], CELL_META$xlim[2], 
-        arrow.head.width = CELL_META$yrange*0.8, arrow.head.length = ux(0.5, "cm"),
-        col = color[CELL_META$sector.numeric.index])
-    circos.text(CELL_META$xcenter, CELL_META$ycenter, CELL_META$sector.index, 
-        facing = "bending.inside", niceFacing = TRUE, cex = 0.6)
-}, bg.border = NA, track.height = 0.4)
-```
-
-## Step 1: Distribution Family  
-The first step in specifying a prior distribution is to find a suitable distribution family.
-For instance, the prior plausibility of parameter values can be expressed through a normal, an inverse gamma, or a Student's *t*-distribution.
-These distributional families have different shapes and a different support, as illustrated in Figure \@ref(fig:distritypeplot).
-For example, the inverse gamma distribution is limited to non-negative parameter values and has a positive skew.
-The normal and Student's *t*-distribution both have support over the real line, however, the Student's *t*-distribution has wider tails, that is, it assigns a higher plausibility to values that are on the extremes of the distribution.
-
-The support of the prior distribution should cover all theoretically possible parameter values.
-For example, if a parameter cannot have negative values, such as a standard deviation, the inverse gamma distribution would be a suitable choice.
-Conversely, if a parameter can take positive and negative values, the normal and Student *t*-distribution would be suitable.
-
-Note that although distributions within a distributional family share the same functional form, they can differ a lot in their shape and location.
-The exact appearance of a prior distribution is controlled by a set of hyperparameters, for example, the mean and variance for the location and scale of a normal prior.
-In the next section, we will give some guidance for specifying these hyperparameters.
-
-```{r distritypeplot, echo=FALSE, fig.cap="Common probability density distributions that may be used as priors.
-Panel A shows normal distributions varying in mean and standard deviation.
-Panel B shows inverse gamma distributions that differ in the shape and scale parameter.
-Panel C shows Student's \\textit{t}-distributions that differ in the location and scale parameter.
-For all three \\textit{t}-distributions, the degrees of freedom are set to 3.", fig.align = "left", fig.height = 8, fig.width = 12, out.width="100%"}
-knitr::include_graphics("Images/Fig4_distrtypefigure_v3.png")
-```
-
-## Step 2: Set Hyperparameters  
-Location, scale, and sometimes shape of the selected distribution can be adjusted by choosing adequate hyperparameters.
-The goal is to adjust the prior to reflect which parameter values we deem most plausible and how much uncertainty there is.
-For instance, panel A of Figure \@ref(fig:distritypeplot) illustrates, that increasing the mean in the normal distribution leads to a shift of location: higher parameter values are deemed more plausible according to the pink distribution compared to the green distribution.
-Panel B shows how uncertainty can be specified, for example, by decreasing the shape of the inverse gamma distribution while keeping the scale constant: the orange distribution assigns probability to a wider range of values than the green distribution.
-
-For this step, it is important to consider the scale on which variables are measured.
-For instance, RTs could be measured in seconds or in milliseconds.
-This means that if a researcher expects a response time of one second, they will formulate a prior distribution with a mean of 1 if RT is measured in seconds, or formulate a prior distribution with a mean of 1000 if RT is measured in milliseconds.
-Similarly, the measurement scale will also affect the width of the specified prior.
-
-## Step 3: Prior Predictions 
-A key advantage of model specification with priors is that we can obtain preditions on what the data would look like that are generated from the model.
-These prior predictions allow us to make the implications of our specified priors more concrete.
-Researchers can look at relevant statistic that summarize the simulated data and decide if they correspond to reasonable observations for their specific research context.
-If the simulated data are in line with the researchers' expectations, the priors can be used for the analysis.
-If the simulated data do not match the expectations, the priors have to be adjusted (Step 4).
-
-```{r priorpredfigurepanel, echo=FALSE, fig.cap="Side and digit effects from predicted data.
-Violins represent the distribution of predicted effects across 1000 simulation runs.
-The blue points are the average effects across simultions.
-Panel A shows the prior predictions for the normal model based on the default priors of \\textit{brms}.
-Panel B shows the prior predictions for the normal model based on the selected priors.
-Panel C shows the prior predictions for the log-normal model based on the priors chosen for the normal model.
-Panel D shows the prior predictions for the log-normal model based on the selected priors.", fig.align = "left", fig.height = 3, fig.width = 4, out.width="100%"}
-knitr::include_graphics("Images/Fig5_priorpredfigure_v2.png")
-```
-
-The simulation proceeds as follows: First, one random draw from each prior distribution is taken.
-This is plugged into the individual effect distributions (Equation (2)).
-Then, the likelihood of the model, such as specified in Equations (1) and (3), is used to simulate observations per condition.
-This means that for *n* participants, *n* random values for the model parameters are generated based on the priors, which will result in the *k* observations per *j* conditions.
-This process is repeated *m* times, where *m* is the number of simulations drawn.
-Note that the more simulations are conducted, the closer one gets to the full picture because enough samples from the prior distribution are needed.
-Finally, we can visualize the simulated data.
-Figure \@ref(fig:priorpredfigurepanel) shows prior predictions for the key effects from the digit classification task.
-For this task, we mainly care about the side effect and digit effects corresponding to the $\beta$s and the $\delta$s from the model.
-Here, the prior predictions for the average side effect and the average digit effect across participants are depicted.
-
-For step 3, the intention is not to compare the actual data of interest to the data predicted by the priors.
-Instead, the researchers' *expectations* of plausible data are compared to the prior predictions.
-Therefore, prior predictive checks are conducted *before* looking at the actual data.
-If the prior predictions match the expectations, the priors can be used for the analysis.
-If they do not match the expectations, the priors need to be adjusted.
-
-## Step 4: Adjust 
-Based on the results of the prior predictions, it might be necessary to adjust the priors.
-Typically, the focus of adjustments will be on finding adequate parameter values for the prior distribution, that is, on repeating step 2 and the following steps of the prior specification process.
-
-In the following two sections, we will provide a walk-through of all four steps of prior specification for our two models of the symbolic distance effect.
-
-## Normal Model
-
-In the symbolic distance model, we have to specify priors for the parameters $\mu_{\gamma}$, $\mu_{\beta}$, $\mu_{\delta}$, $\sigma^2_{\gamma}$, $\sigma^2_{\beta}$, and $\sigma^2_{\delta}$.
-The parameters $\mu_{\gamma}$, $\mu_{\beta}$, and $\mu_{\delta}$ can be positive or negative.
-All variance terms ($\sigma_{\gamma}^2$, $\sigma_{\beta}^2$, $\sigma_{\delta}^2$, and $\sigma^2$) are constrained to be positive.
-If there is no information from previous studies, researchers might be inclined to use the default priors in `brms` as a starting point.
-
-In `brms`, different default priors are set on different types of parameters [@R-brms_a].
-The prior on the group-level regression weights ($\mu_{\beta}$, $\mu_{\delta}$) is a normal distribution with a mean of 0 and a standard deviation of 1.
-The prior on the intercept $\mu_{\gamma}$ is a scaled and shifted Student's *t*-distribution with 3 degrees of freedom, a location parameter of 1, and a scale parameter of 10.
-`brms` specifies priors on standard deviations ($\sqrt{\sigma^2}$) instead of variances.
-For the normal model, this means that priors are set on $\sigma_{\gamma}$, $\sigma_{\beta}$, $\sigma_{\delta}$ and $\sigma$, instead of on $\sigma_{\gamma}^2$, $\sigma_{\beta}^2$, $\sigma_{\delta}^2$ and $\sigma^2$.
-The default prior on the standard deviations is a central, scaled Student's *t*-distribution with 3 degrees of freedom and a scale parameter of 10 that is truncated at zero.
-To summarize, the default priors would be set the following way:
-
-\begin{equation}
-\begin{aligned}
-\mu_{\gamma} &\sim \text{Student's t}(3, 1, 10), (\#eq:chosenpriors1) \\ 
-\mu_{\beta}, \mu_{\delta} &\sim \text{Normal}(0, 1), \\ 
-\sigma, \sigma_{\gamma}, \sigma_{\beta}, \sigma_{\delta} &\sim \text{Truncated-Student's t}_{+}(3, 0, 10). 
-\end{aligned}
-\end{equation}
-The prior distributions `brms` specifies on the standard deviations can be transformed to prior distributions on variances.
-Prior distributions on variances have a smaller scale than priors on standard deviation.
-Note, however, that the transformation of the prior distributions is not as simple as the transformation of the parameter itself.
-In our [Online Supplement](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/tree/main/R%20code/Prior%20transformation), we provide R-code with a numeric solution to the transformation.
-For the `brms` priors, we obtain the following implied prior distributions on variances:
-
-\begin{equation}
-\begin{aligned}
-\sigma_{\gamma}^2, \sigma_{\beta}^2, \sigma_{\delta}^2, \sigma^2 &\sim \text{Student's t}(3, 0, 1.48). 
-\end{aligned}
-\end{equation}
-
-The priors are visualized on the right side of Figure \@ref(fig:priorvisualizationbrms2).
-Note that these prior distributions are on the group-level parameters and represent the uncertainty about the general effect.
-The uncertainty about individual effects is represented by the marginal prior distribution on the left.
-This distribution takes the prior variability of the mean and variance as well as the implied individual variability into account.
-
-```{r priorvisualizationbrms2, fig.cap="Visualization of \\textit{brms} default prior settings for the normal model.
-On the right, the prior distributions on general effects parameters are shown.
-On the left, the implied distribution of the individual effects for every parameter based on these priors are presented.
-", fig.align = "left", message=FALSE, warning=FALSE, cache = TRUE}
-# library(crch)  for truncated student's t
-
-# Plots 
-x <- seq(-3, 3, by=.01)
-# color codes: https://flaviocopes.com/rgb-color-codes/
-
-# Instead do this, take integral  
-Nplotst <- 1000000
-mugammaplotbrms <- rst(Nplotst, nu = 3, mu = 1, sigma = 10)
-vargammaplotbrms <- rtt(Nplotst, df = 3, location = 0, scale = 10, left = 0, right = Inf)  # should be truncated 
-betaplot2brms <- rnorm(Nplotst, mugammaplotbrms, sqrt(vargammaplotbrms))
-
-pgamma2databrms <- data.frame(betaplot2brms)
-pgamma2brms <- ggplot(pgamma2databrms, aes(x = betaplot2brms)) + geom_density_line(color = "#0072B2", fill = "#0072B2", alpha = 0.1, position = "identity") + ylab("Density") + xlab("") + labs(title = expression(~gamma[i])) + theme(plot.title = element_text(size = 14)) + xlim(-20, 20)
-
-parametersgammabrms <- c("Mean" = "#D95F02", "Variance" = "#1B9E77")
-
-ptogethergamma2brms <- ggplot(data = data.frame(x = c(-20, 20)), aes(x)) +
-  stat_function(fun = dst, n = 1001, args = list(nu = 3, mu = 1, sigma = 10), aes(color = "Mean", fill = "Mean"), geom = "area", alpha = 0.3) + ylab("") + xlab("") + 
-  stat_function(fun = dtt, n = 1001, args = list(df = 3, location = 0, scale = 10, left = 0, right = Inf), geom = "area", aes(color = "Variance", fill = "Variance"), alpha = 0.3, position = "identity") + geom_hline(yintercept=0, colour="white", size=1) + scale_color_manual(name = "Prior on", values = parametersgammabrms, aesthetics = c("colour", "fill"), labels = c(expression(~mu[~gamma]), expression(~sigma[~gamma]))) +
-  theme(legend.text=element_text(size = 9), legend.title = element_text(size = 7), legend.key.width=unit(0.45,"cm"), legend.key.height=unit(0.5,"cm"))
-
-# Beta 
-mubetaplotbrms <- rnorm(Nplotst, 0, 1)
-varbetaplotbrms <- rtt(Nplotst, df = 3, location = 0, scale = 10, left = 0, right = Inf)  
-betaplot22brms <- rnorm(Nplotst, mubetaplotbrms, sqrt(varbetaplotbrms))
-
-pbeta2databrms <- data.frame(betaplot22brms)
-pbeta2brms <- ggplot(pbeta2databrms, aes(x = betaplot22brms)) + geom_density_line(color = "#0072B2", fill = "#0072B2", alpha = 0.1, position = "identity") + ylab("Density") + xlab("") + labs(title = expression(~beta["i"])) + theme(plot.title = element_text(size = 14)) + xlim(-10, 10)
-
-ptogetherbeta2brms <- ggplot(data = data.frame(x = c(-20, 20)), aes(x)) +
-  stat_function(fun = dnorm, n = 1001, args = list(0, 1), aes(color = "Mean", fill = "Mean"), geom = "area", alpha = 0.3) + ylab("") + xlab("") + 
-  stat_function(fun = dtt, n = 1001, args = list(df = 3, location = 0, scale = 10, left = 0, right = Inf), geom = "area", aes(color = "Variance", fill = "Variance"), alpha = 0.3, position = "identity") + geom_hline(yintercept=0, colour="white", size=1) + scale_color_manual(name = "Prior on", values = parametersgammabrms, aesthetics = c("colour", "fill"), labels = c(expression(~mu[~beta]), expression(~sigma[~beta]))) +
-  theme(legend.text=element_text(size = 9), legend.title = element_text(size = 7), legend.key.width=unit(0.45,"cm"), legend.key.height=unit(0.5,"cm"))
-
-# Delta's
-
-mudeltaplotbrms <- rnorm(Nplotst, 0, 1)
-vardeltaplotbrms <- rtt(Nplotst, df = 3, location = 0, scale = 10, left = 0, right = Inf)  
-deltaplot2brms <- rnorm(Nplotst, mudeltaplotbrms, sqrt(vardeltaplotbrms))
-
-pdelta2databrms <- data.frame(deltaplot2brms)
-pdelta2brms <- ggplot(pdelta2databrms, aes(x = deltaplot2brms)) + geom_density_line(color = "#0072B2", fill = "#0072B2", alpha = 0.1, position = "identity") + ylab("Density") + xlab("Individual effect") + labs(title = expression(~delta[i])) + theme(plot.title = element_text(size = 14)) + xlim(-10, 10)
-
-ptogetherdelta2brms <- ggplot(data = data.frame(x = c(-20, 20)), aes(x)) +
-  stat_function(fun = dnorm, n = 1001, args = list(0, 1), aes(color = "Mean", fill = "Mean"), geom = "area", alpha = 0.3) + ylab("") + xlab("Parameter value") + 
-  stat_function(fun = dtt, n = 1001, args = list(df = 3, location = 0, scale = 10, left = 0, right = Inf)  , geom = "area", aes(color = "Variance", fill = "Variance"), alpha = 0.3, position = "identity") + geom_hline(yintercept=0, colour="white", size=1) + scale_color_manual(name = "Prior on", values = parametersgammabrms, aesthetics = c("colour", "fill"), labels = c(expression(~mu[~delta]), expression(~sigma[~delta]))) +
-  theme(legend.text=element_text(size = 9), legend.title = element_text(size = 7), legend.key.width=unit(0.45,"cm"), legend.key.height=unit(0.5,"cm"))
-
-theme_set(theme_apa(base_size = 9))
-#ptogether
-
-# Final figure 
-grid.arrange(pgamma2brms, ptogethergamma2brms, pbeta2brms, ptogetherbeta2brms, pdelta2brms, ptogetherdelta2brms, nrow = 3, ncol = 2, widths = c(0.4, 0.6))
-```
-
-We will continue with step 3 of the prior specification process and check what the expected data would look like according to the selected priors.
-The results of the prior prediction for the symbolic distance effect with the priors from Equation \@ref(eq:chosenpriors1) with the number of participants $n = 10$, the number of trials per condition $k = 60$, and the number of simulations $m = 1000$ are shown in panel A of Figure \@ref(fig:priorpredfigurepanel).
-Code for the simulations can be found in our [Online Supplement](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment).
-
-The figure shows the difference in RTs between digits for every repetition averaged across subjects (i.e., 1000 grey dots per effect).
-The blue dots represent the mean effect.
-Using the default prior from `brms`, the mean effects are expected to be around zero, with a mean effect of zero.
-However, the range around this mean is very wide, ranging from -10 to 10 second effects.
-Most of the trials are in a smaller range of -5 to 5 second effects.
-This means that according to our priors, we would expect the effects to fall in this range.
-For this task, where participants have to respond as fast as possible, such large RT effects are unlikely.
-Therefore, it seems reasonable to adjust the priors.
-
-```{r priorvisualization2, fig.cap="Visualization of the adjusted prior settings for the normal model.
-On the right, the prior distributions on general effects parameters are shown.
-On the left, the implied distribution of the individual effects for every parameter based on these priors are presented.
-", fig.align = "left", fig.height = 6, fig.width = 8, warning=FALSE, message=FALSE, cache = TRUE}
-### Set the priors 
-a <- 0.5  # mean for mu, mu is the mean of theta  
-b <- 1  # variance for mu 
-c <- 3  # mean for sigma 
-d <- .7  # variance for sigma, sigma is the variance of observations  
-e <- 3  # mean for g, g is the variance of theta   
-f <- .7  # variance for g
-
-#### Priors for beta 
-a2 <- 0  # mean for mean of beta
-b2 <- 0.09  # variance for mean of beta, used to be 1 
-e2 <- 3  # mean for variance of beta, used to be .7 
-f2 <- 0.5  # variance for variance of beta, used to be .2.
-
-#### Prior for deltas, same for every delta 
-a3 <- 0  # mean for mean of delta 
-b3 <- 0.09  # sd for mean of delta, 0.5, 0.3  
-dd3 <- 3  # mean for variance of delta 
-f3 <- 0.5   # variance for variance of delta
-
-# Plots 
-x <- seq(-3, 3, by=.01)
-# color codes: https://flaviocopes.com/rgb-color-codes/
-
-## Gamma 
-mu <- rnorm(10, a, b)
-g <- rinvgamma(10, e, f)
-gammaplot <- dnorm(x, mu, g)  # I think this is not possible. Get a value for mu and g for each x --> 601
-
-# Instead do this, take integral  
-Nplot <- 1000000
-mugammaplot <- rtruncnorm(Nplot, a = 0, mean = 0.5, sd = 1)
-vargammaplot <- rinvgamma(Nplot, e, f)
-betaplot2 <- rnorm(Nplot, mugammaplot, sqrt(vargammaplot))
-# plot(density(betaplot2, na.rm = TRUE))
-
-pgamma2data <- data.frame(betaplot2)
-pgamma2 <- ggplot(pgamma2data, aes(x = betaplot2)) + geom_density_line(color = "#0072B2", fill = "#0072B2", alpha = 0.1, position = "identity") + ylab("Density") + xlab("") + labs(title = expression(~gamma[i])) + theme(plot.title = element_text(size = 14)) + xlim(-10, 10)
-
-parametersgamma <- c("Mean" = "#D95F02", "Variance" = "#1B9E77")
-
-ptogethergamma <- ggplot(data = data.frame(x = c(-5, 5)), aes(x)) +
-  stat_function(fun = dnorm, n = 1001, args = list(mean = a, sd = b), geom = "area", color = "#D95F02", fill = "#D95F02", alpha = 0.3) + ylab("") +
-  stat_function(fun = dinvgamma, n = 1001, args = list(shape = e, scale = f), geom = "area", color = "#1B9E77", fill = "#1B9E77", alpha = 0.3, position = "identity") + geom_hline(yintercept=0, colour="white", size=1) + scale_color_manual(name = "Prior parameters gamma", values = parametersgamma) +
-  scale_fill_manual(name = "Bar",values = parametersgamma)
-
-ptogethergamma2 <- ggplot(data = data.frame(x = c(-5, 5)), aes(x)) +
-  stat_function(fun = dtruncnorm, n = 1001, args = list(a = 0, mean = 0.5, sd = 1), aes(color = "Mean", fill = "Mean"), geom = "area", alpha = 0.3) + ylab("") + xlab("") + 
-  stat_function(fun = dinvgamma, n = 1001, args = list(shape = e, scale = f), geom = "area", aes(color = "Variance", fill = "Variance"), alpha = 0.3, position = "identity") + geom_hline(yintercept=0, colour="white", size=1) + scale_color_manual(name = "Prior on", values = parametersgamma, aesthetics = c("colour", "fill"), labels = c(expression(~mu[~gamma]), expression(~sigma[~gamma]^2))) +
-  theme(legend.text=element_text(size = 9), legend.title = element_text(size = 7), legend.key.width=unit(0.45,"cm"), legend.key.height=unit(0.5,"cm")) + xlim(-1, 1)
-
-# Beta 
-mu2 <- dnorm(x, a2, b2)
-g2 <- dinvgamma(x, e2, f2)
-betaplot <- dnorm(x, mu2, g2)
-
-mubetaplot <- rnorm(Nplot, a2, b2)
-varbetaplot <- rnorm(Nplot, e2, f2)
-betaplot22 <- rnorm(Nplot, mubetaplot, sqrt(varbetaplot))
-# plot(density(betaplot22, na.rm = TRUE))
-
-pbeta2data <- data.frame(betaplot22)
-pbeta2 <- ggplot(pbeta2data, aes(x = betaplot22)) + geom_density_line(color = "#0072B2", fill = "#0072B2", alpha = 0.1, position = "identity") + ylab("Density") + xlab("") + labs(title = expression(~beta["i"])) + theme(plot.title = element_text(size = 14)) + xlim(-10, 10)
-
-ptogetherbeta2 <- ggplot(data = data.frame(x = c(-5, 5)), aes(x)) +
-  stat_function(fun = dnorm, n = 1001, args = list(mean = a2, sd = b2), aes(color = "Mean", fill = "Mean"), geom = "area", alpha = 0.3) + ylab("") + xlab("") + 
-  stat_function(fun = dinvgamma, n = 1001, args = list(shape = e2, scale = f2), geom = "area", aes(color = "Variance", fill = "Variance"), alpha = 0.3, position = "identity") + geom_hline(yintercept=0, colour="white", size=1) + scale_color_manual(name = "Prior on", values = parametersgamma, aesthetics = c("colour", "fill"), labels = c(expression(~mu[~beta]), expression(~sigma[~beta]^2))) +
-  theme(legend.text=element_text(size = 9), legend.title = element_text(size = 7), legend.key.width=unit(0.45,"cm"), legend.key.height=unit(0.5,"cm")) + xlim(-1, 1)
-
-# Delta's 
-mu3 <- dnorm(x, a3, b3)
-g3 <- dinvgamma(x, dd3, f3)
-deltaplot <- dnorm(x, mu3, g3)
-
-mudeltaplot <- rnorm(Nplot, a3, b3)
-vardeltaplot <- rnorm(Nplot, dd3, f3)
-deltaplot2 <- rnorm(Nplot, mudeltaplot, sqrt(vardeltaplot))
-
-pdelta2data <- data.frame(deltaplot2)
-pdelta2 <- ggplot(pdelta2data, aes(x = deltaplot2)) + geom_density_line(color = "#0072B2", fill = "#0072B2", alpha = 0.1, position = "identity") + ylab("Density") + xlab("Individual Effect") + labs(title = expression(~delta[i])) + theme(plot.title = element_text(size = 14)) + xlim(-10, 10)
-
-ptogetherdelta2 <- ggplot(data = data.frame(x = c(-5, 5)), aes(x)) +
-  stat_function(fun = dnorm, n = 1001, args = list(mean = a3, sd = b3), aes(color = "Mean", fill = "Mean"), geom = "area", alpha = 0.3) + ylab("") + xlab("Parameter Value") + 
-  stat_function(fun = dinvgamma, n = 1001, args = list(shape = dd3, scale = f3), geom = "area", aes(color = "Variance", fill = "Variance"), alpha = 0.3, position = "identity") + geom_hline(yintercept=0, colour="white", size=1) + scale_color_manual(name = "Prior on", values = parametersgamma, aesthetics = c("colour", "fill"), labels = c(expression(~mu[~delta]), expression(~sigma[~delta]^2))) +
-  theme(legend.text=element_text(size = 9), legend.title = element_text(size = 7), legend.key.width=unit(0.45,"cm"), legend.key.height=unit(0.5,"cm")) + xlim(-1, 1)
-
-theme_set(theme_apa(base_size = 9))
-# ptogether
-
-# Final figure 
-grid.arrange(pgamma2, ptogethergamma2, pbeta2, ptogetherbeta2, pdelta2, ptogetherdelta2, nrow = 3, ncol = 2, widths = c(0.4, 0.6))
-
-```
-
-The normal distribution on $\mu_{\beta}$ and $\mu_{\delta}$ seems reasonable if we assume that a two-sided hypothesis about the symbolic distance effect is tested.
-However, we expect less variance around this mean.
-Therefore, we decrease the setting for the variation.
-The wide tails of the Student's *t*-distribution could be one of the reasons for the high range in variation.
-We change the prior for $\mu_{\gamma}$ to a normal distribution that has more narrow tails.
-As RTs cannot be negative, we adjust the prior to be positive only by truncating the normal distribution using a lower bound of zero.
-Finally, we would like to use a more common conjugate prior for the variances to ensure obtaining a proper posterior distribution.
-^[Conjugate priors result in posterior distributions that belong to the same distribution family as the prior distribution [@Fink1997].
-For a normal likelihood, these are the normal distribution for the mean and the inverse gamma distribution for the variance.] Therefore, we choose an inverse gamma distribution.
-We can also be more specific about the expected RTs.
-Based on our knowledge of these types of tasks [e.g., @haaf2017developing], it seems highly plausible that participants will respond within a second or two.
-We would expect effects in the range of 10 milliseconds to 100 milliseconds.
-We adjust our priors using these new distribution types and settings, evaluate the prior predictions (shown in panel B of Figure \@ref(fig:priorpredfigurepanel)) and arrive at the following priors:^[Because response times cannot be negative the prior for the overall baseline $\mu_\gamma$ is a truncated normal with a lower bound of zero.]
-
-\begin{equation}
-\begin{aligned}
-\mu_{\gamma} &\sim \text{Truncated-Normal}_{+}(0.5, 1), (\#eq:chosenpriors2var) \\
-\mu_{\beta} &\sim \text{Normal}(0, 0.09), \\
-\mu_{\delta} &\sim \text{Normal}(0, 0.09), \\
-\sigma^2 &\sim \text{Inverse-Gamma}(3, 0.7),  \\
-\sigma^2_{\gamma} &\sim \text{Inverse-Gamma}(3, 0.7), \\   
-\sigma^2_{\beta} &\sim \text{Inverse-Gamma}(3, 0.5), \\   
-\sigma^2_{\delta} &\sim \text{Inverse-Gamma}(3, 0.5).    
-\end{aligned}
-\end{equation}
-Setting equivalent prior distributions on the standard deviations instead of on the variances results in the following priors:
-
-\begin{equation}
-\begin{aligned}
-\mu_{\gamma} &\sim \text{Truncated-Normal}_{+}(0.5, 1), (\#eq:chosenpriors2sd) \\
-\mu_{\beta} &\sim \text{Normal}(0, 0.3), \\ 
-\mu_{\delta} &\sim \text{Normal}(0, 0.3), \\  
-\sigma &\sim \text{Inverse-Gamma}(13.8, 6.3),  \\ 
-\sigma_{\gamma} &\sim \text{Inverse-Gamma}(13.8, 6.3), \\   
-\sigma_{\beta} &\sim \text{Inverse-Gamma}(13.8, 5.3), \\    
-\sigma_{\delta} &\sim \text{Inverse-Gamma}(13.8, 5.3).     
-\end{aligned}
-\end{equation}
-The adjusted priors are visualized in Figure \@ref(fig:priorvisualization2).
-The figure shows that the adjusted priors assign their highest probability density to a smaller range of values and are, therefore, more specific than `brms`'s default priors presented in Figure \@ref(fig:priorvisualizationbrms2).
-In addition, the shape of the distribution is different for the individual variation, resulting in a higher probability for values closer to zero.
-Panel B of Figure \@ref(fig:priorpredfigurepanel) shows the predictions of the normal model with adjusted prior distributions.
-Most effects now range between -0.5 and 0.5 seconds.
-This aligns better with our expectations based on existing literature, but still allows for a considerable uncertainty about the true size of the effects.
-
-## Log-normal model
-
-We follow the same procedure for prior specification in the log-normal model.
-As before, we prefer to apply conjugate priors.
-\Revision{The conjugate priors that we applied to the normal model are also conjugate priors for the log-normal model [@fink1997compendium].
-For instance, with a normal distribution as prior for the mean of the log-normal model, we assume that the mean of the parameter can be either negative, zero, or positive, and assign a higher probability to values closer to zero.
-Applying this conjugate prior, the posterior of the mean in the log-normal model will result in a normal distribution as well.}
-As a starting point, we can use the adjusted priors of the normal model and perform a prior prediction, shown in panel C of Figure \@ref(fig:priorpredfigurepanel).
-These priors result in a slightly higher variance in RTs than we would expect, and, importantly, the mean response time differences are not centered around zero in all conditions.
-This pattern is inconsistent with our expectations, which means that prior distributions for the log-normal model need to be adjusted.
-
-```{r priorpredfigurepanellog, echo=FALSE, fig.cap="Predicted mean RT difference between digits per repetition.
-The blue line represents the aggregated trend.
-Panel A represents the prior prediction based on the priors of the normal model.
-Panel B shows the prior prediction based on the selected prior for the log-normal model.", fig.align = "left", fig.height = 3, fig.width = 4, out.width="100%", eval = FALSE}
-knitr::include_graphics("Images/priorpredfigure_lognormal_v2.png")
-```
-The transformation of the prior distributions is not as simple as the transformation of the parameter itself.
-Panel C shows that with the same priors as for the normal model, we obtain different data for the log-normal model compared to the normal model (panel B).
-Therefore, we adjusted the priors until the simulated data matched our expectations and looked like panel B.
-This yields the following priors that result in the prior prediction shown in panel D of Figure \@ref(fig:priorpredfigurepanel):^[For the overall baseline $\mu_\gamma$ we adjusted the prior based on our expectation of the expected overall response times.
-A mean parameter of -0.5 for the log-normal likelihood corresponds to a mean response time of 1.]
-
-\begin{equation}
-\begin{aligned}
-\mu_{\gamma} &\sim \text{Normal}(-0.5, 1), (\#eq:chosenpriorslogvar) \\
-\mu_{\beta} &\sim \text{Normal}(0, 0.005), \\
-\mu_{\delta} &\sim \text{Normal}(0, 0.005), \\
-\sigma^2 &\sim \text{Inverse-Gamma}(3, 0.3),  \\
-\sigma^2_{\gamma} &\sim \text{Inverse-Gamma}(3, 0.3), \\   
-\sigma^2_{\beta} &\sim \text{Inverse-Gamma}(3, 0.01), \\   
-\sigma^2_{\delta} &\sim \text{Inverse-Gamma}(3, 0.01).    
-\end{aligned}
-\end{equation}
-Setting the same prior distribution on the standard deviations instead of on the variances yields the following priors:
-
-\begin{equation}
-\begin{aligned}
-\mu_{\gamma} &\sim \text{Normal}(-0.5, 1), (\#eq:chosenpriorslogsd) \\
-\mu_{\beta} &\sim \text{Normal}(0, 0.07), \\
-\mu_{\delta} &\sim \text{Normal}(0, 0.07), \\
-\sigma &\sim \text{Inverse-Gamma}(13.8, 4.1),  \\
-\sigma_{\gamma} &\sim \text{Inverse-Gamma}(13.8, 4.1), \\   
-\sigma_{\beta} &\sim \text{Inverse-Gamma}(13.8, 0.7), \\   
-\sigma_{\delta} &\sim \text{Inverse-Gamma}(13.8, 0.7).    
-\end{aligned}
-\end{equation}
-
-## Prior sensitivity 
-In the previous sections, we decided on one set of prior distributions for each of the proposed models.
-However, researchers may find it difficult to settle on a single set of prior distributions in practice.
-This raises the question: What happens if you cannot decide on one prior?
-If several possibilities seem plausible, it is possible to use them all separately for the analysis and investigate their influence on the analysis results.
-This is called a prior sensitivity analysis [@roos2015sensitivity].
-Generally, when it comes to prior specification, it is important to be *transparent* and to provide a justification for the selected prior distributions [@Stefan_etal2021].
-Prior distributions are rarely inadmissible.
-However, as we demonstrated earlier, their usefulness in constraining model predictions can vary.
-Therefore, providing a justification of prior distributions in a paper can help readers gauge the generative quality of a presented model.
-When multiple justifiable prior distributions are applied to the same analysis, it is important to be transparent about this to avoid the impression of cherry-picking priors.
-If priors are chosen for their best fit to the data, this can lead to overfitting and spurious results.
-Prior sensitivity analyses and transparency about the prior specification process more generally, can therefore help to increase trust in the research results.
-
-# Bayesian Parameter Estimation
-
-In Bayesian statistics, parameter estimates are obtained from the posterior distribution.
-The posterior distribution captures the uncertainty about the parameter after seeing the data, $\bf{Y}$.
-The mean of this distribution can be used as a point estimate for the parameter.
-The posterior distribution is calculated by updating the prior distribution using Bayes rule.
-For the symbolic distance effect, we have a model with many parameters, so the posterior distribution is also multidimensional.
-Instead of writing out all the parameters, we replace all the model parameters with the vector $\bm{\theta}$, resulting in the joint posterior:
-
-\begin{equation}
-\underbrace{P(\bm{\theta}|\bf{Y})}_{\text{Posterior}} = \underbrace{P(\bm{\theta})}_{\text{Prior}}\times\underbrace{\frac{P(\bf{Y}|\bm{\theta})}{P(\bf{Y})}}_{\text{Updating factor}}.
-(\#eq:Bayesrule)
-\end{equation}
-
-Typically, we are not interested in the distribution of all parameters together, but in the posterior of one parameter independent of all the others.
-This posterior distribution is called the marginal posterior, and it is computed by integrating out all other parameters of the joint posterior.
-The marginal posterior distribution for the side effect $\mu_{\beta}$, for instance, is:
-
-\begin{equation}
-P(\mu_{\beta}| \bf{Y}) = \int P(\bf{Y} | \mu_{\beta}, \bm{\theta})P(\mu_{\beta})P(\bm{\theta})) d\bm{\theta}  
-\end{equation}
-where $\bm{\theta}$ now represents all parameters except for $\mu_{\beta}$.
-With an increasing number of parameters and an increasingly complex model structure, it is no longer possible to find analytical solutions for the multidimensional integrals involved in the posterior distributions.
-A solution to this issue is to sample from the joint posterior distribution.
-Using algorithms such as Markov chain Monte Carlo sampling [MCMC; for an introduction see  @van2018simple], Gibbs sampling [@chib1995marginal] or Hamiltonian Monte Carlo sampling [@betancourt2015hamiltonian].
-It is out of the scope of this paper to go into detail on the algorithms, but we refer the interested reader to a tutorial paper on MCMC by van Ravenzwaaij et al.
-[-@van2018simple], a conceptual introduction to Hamiltonian Monte Carlo by Betancourt [-@betancourt2018conceptual], and a summary of samplers by Green et al.
-[-@green2015bayesian].
-
-An approximation for the marginal posterior distribution can be obtained by assessing the empirical distribution of MCMC samples for the parameter of interest disregarding the other parameters.
-Since the frequency of combinations of samples in the joint posterior is determined by their posterior plausibility, this procedure is equivalent to a multidimensional integration across all other parameters.
-
-Every sampling algorithm requires the specification of the number of samples (also called iterations) that are drawn from the posterior distribution.
-Furthermore, all MCMC sampling methods have in common that results depend (somewhat) on chosen starting values, that is, initial parameter values for which the unnormalized posterior (i.e., likelihood times prior) is evaluated.
-This dependency can be assessed if the sampling algorithm is repeated several times (i.e., chains) with different starting values.
-Since the first few iterations of a chain---often called warm-up or burn-in---are somewhat dependent on the starting value, these samples are typically discarded.
-
-Before interpreting the estimation results, it has to be checked whether the posterior distribution of the parameters has converged, meaning that a stationary posterior distribution has been reached [@Vehtari_2021].
-We will explain these convergence checks in more detail in the section Model Diagnostics.
-
-## Software 
-There are many different programs available to fit Bayesian hierarchical models, that is, to obtain samples from the joint posterior distribution of parameters.
-In this tutorial, we focus on two of the most commonly used `R` packages, `rstan` [@R-rstan] and `brms` [@R-brms_a; @R-brms_b].
-We assume that readers are somewhat familiar with `R`.
-The package `rstan` provides the most modern algorithms and flexibility in model set-up.
-The package `brms` was created as an overlay to `rstan` and is based on the well-known `lme4` syntax [@R-lme4].
-The goal of `brms` is to ease the transition to Bayesian hierarchical modeling for novice users of Bayesian statistics.
-The packages work similarly for the normal and log-normal model.
-In this section, we explain how a model can be fitted in each package using code snippets.
-In addition, we offer a separate R Markdown file for each package in our [Online Supplement](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment) that explains in more depth how to analyze the full symbolic distance effect model with the specific package.
-
-### Rstan
-
-`Rstan` allows the application of `Stan` [@carpenter2017stan], a probabilistic programming language, in `R`.
-The package uses a version of the Hamiltonian Monte Carlo (HMC) algorithm, No-U-turn sampler [NUTS; Hoffman & Gelman, -@hoffman2014no], for sampling.
-`Rstan` requires two files: a `.stan` file containing the model specification and an R script containing the code for the model fitting.
-In the `.stan` file, the parameters, priors, and the likelihood are specified.
-An example `.stan` file of a simple, non-hierarchical normal model, where the variance is already known and the mean has to be estimated, is shown below [adapted example from @nicenboim2021introduction, Chapter 10].
-The `.stan` file for the symbolic distance effect, including explanations, can be found in the [Online Supplement](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/blob/main/R%20objects/rstan/normal%20model/myModel.stan).
-The file is divided into three sections: `data`, `parameters`, and `model`.
-
-\scriptsize
-```{r stanfileexample, eval = FALSE, echo = TRUE}
-data {
-  int<lower = 1> N;             // Total number of trials 
-  vector[N] y;                  // Score in each trial 
-  real<lower = 0> sigma;        // The variance of the normal distribution 
-}
-
-parameters {
-  real mu; 
-}
-
-model {
-  // Priors: 
-  target += normal_lpdf(mu | 0, 20);
-  // Likelihood: 
-  for(i in 1:N)
-    target += normal_lpdf(y[i] | mu, sigma);
-}
-```
-\normalsize
-
-The `data` section specifies the variables in the dataset, as well as constant values in the model.
-For the symbolic distance effect model, the `data` section contains the total number of observations, the participant index variable, the response time variable in the dataset, the condition for every response time (side of the digit and the digit indicator, i.e., variables $x_j$, $u_j$, $v_j$, $w_j$ and $z_j$ from \@ref(eq:datalevel)) and the prior specification according to Equations \@ref(eq:chosenpriors2var) and \@ref(eq:chosenpriorslogvar).
-The `parameters` section specifies all parameters that are estimated.
-These are the general and individual effects.
-The last section, `model`, contains the likelihood function and priors, as specified earlier in this article.
-
-In the corresponding `R` file, the model is fitted to the data using the `stan` function, as shown below.
-In this function, it is possible to specify the number of iterations, chains and warm-up samples in the Monte Carlo sampling procedure.
-Next to these specifications, the function requires a directory path to the `.stan` file and an `R` object of type list containing the data (here called `myData_list`).
-Every element of the `data` section in the `.stan` file needs to be specified with the same name in the data object.
-&nbsp;
-
-\scriptsize
-```{r stanfitexample, eval = FALSE, echo = TRUE}
-model_fit <- stan(file = "./myModel.stan",  # Stan file with model  
-                  data = myData_list,       # List with observed data and constants
-                  iter = 4000,              # Number (Nr.) of iterations per chain
-                  chains = 4,               # Nr. of chains 
-                  warmup = 1000)            # Nr. iterations for warmup per chain
-```
-\normalsize
-
-The package `rstan` offers a lot of flexibility.
-It is possible to specify the likelihood and priors in the exact way as is wished.
-However, this comes with the price of some technical knowledge.
-The commented code for fitting the full symbolic distance effect model using `rstan`, can be found in [Online Supplement C: *Bayesian Hierarchical Modeling in Rstan*](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/blob/main/C%20-%20Tutorial%20Normal%20Model%20Rstan/Online-Supplement-C---Bayesian-Hierarchical-Modeling-in-rstan.pdf).
-
-### Brms
-
-The `R` package `brms` also allows fitting Bayesian models using `Stan`.
-However, model specification and model fitting can be achieved using a much simplified `R`-script \textcolor{blue}{alone}.
-Model specification is based on a formula syntax that is similar to the notation in the popular `lme4` package for frequentist hierarchical modeling [@R-lme4].
-The formula is an object that specifies the dependent variable as a function of the independent variables.
-The `~` (tilde) separates the dependent variable (`rt`) on the left side from the independent variables on the right side.
-The formula distinguishes between general effects and individual deviations.
-The first part of the equation on the predictor side of the formula shown below (i.e., `1 + side + dif1 + dif2 + dif3 + dif4`) represents the general effects.
-The `1` represents the intercept, in the case of the symbolic distance effect $\mu_{\gamma}$, `side`, and `dif1` to `dif4` represent the general effects of side and distance ($\mu_{\beta}$ and $\mu_{\delta}$).
-Within the round brackets individual deviations and grouping variables are specified.
-In this case the grouping variable (`ind`) indicates that each participant has an individual effect.
-In other cases effects might vary per item, block, or even the combination of participant and item.
-The double bar (`||`) indicates that the correlation between the parameters should not be modeled.
-When using a single ba (`|`) then correlations across all individual effects (intercept, side and distance) are modeled.
-This approach, however, is a bit more difficult to interpret and would also not correspond to the implemented `Stan` model.
-
-\scriptsize
-```{r brmsfitexample, eval = FALSE, echo = TRUE}
-formula <- rt ~ 1 + side + dif1 + dif2 + dif3 + dif4 +   # General effects 
-          (1 + side + dif1 + dif2 + dif3 + dif4 || ind)  # Individual effects
-
-model_fit <- brm(formula = formula,                      # Model formula 
-                 data = myData_dataframe,                # Dataset (parameters, observations) 
-                 family = gaussian(),                    # Family, distribution dependent variable 
-                 prior = priorsmodel,                    # Specification of the priors 
-                 warmup = 1000,                          # Nr. iterations for warmup (per chain)
-                 iter = 4000,                            # Nr. iterations (per chain)
-                 chains = 4,                             # Nr. chains 
-                 core = 4)                               # Nr. of cores for parallel estimation
-```
-
-```{r formulafit, eval = FALSE, echo = FALSE}
-formula <- rt ~ 1 + side + dif1 + dif2 + dif3 + dif4 +    # General effects 
-          (1 + side + dif1 + dif2 + dif3 + dif4 || ind)   # Individual effects
-
-```
-
-\normalsize The `brm` function requires a data frame (here called `myData_dataframe`) that contains all variables mentioned in the formula.
-The argument `family` in the `brm` function specifies the exponential family of distributions according to which the dependent variable is assumed to be distributed.
-For the normal model, this argument takes the value `gaussian`; for the log-normal model, it needs to be specified as `lognormal`.
-
-Finally, we specify the priors.
-In `brms`, priors can be set on the general effects represented by `class = b` and on the individual variation represented by `class = sd`.
-It is important to note that priors can only be set on the specific parameters that are part of the brms model specification.
-For example, `brms` only allows setting priors on standard deviations and not on variances.
-Therefore, for the normal model of the symbolic distance effect, we will use the priors for standard deviations specified in Equation \@ref(eq:chosenpriors2sd).
-To set the priors, we use the `set_prior` function, as shown below.
-In this function, first, the distribution is specified, using the `Stan` programming language.
-Then, the type of parameter to which the prior applies is defined by `class`.
-The parameter to which this prior applies is specified by `coef`.
-For the priors on the individual effects, an additional element has to be specified, namely, the grouping variable (i.e., `group`).
-For example, the prior distributions on the general side effect, $\mu_{\beta}$, and on its variability between individuals, $\sigma_{\beta}$, can be specified as follows:
-&nbsp;
-
-\scriptsize 
-```{r priorsillusbrms2, eval = FALSE, echo = TRUE}
-set_prior("normal(0,0.3)", class = "b", coef="side")  # general effect 
-set_prior("inv_gamma(13.8, 5.3)", class = "sd",       # individual variation 
-          coef="side", group = "ind") 
-```
-\normalsize
-
-\scriptsize 
-```{r priorsillusbrms, eval = FALSE, echo = FALSE}
-priorsmodel <- c(set_prior("normal(0,1)", class = "Intercept"),  
-                 set_prior("inv_gamma(13.75136, 4.102063)", class = "sd", coef="Intercept", group = "sub1"), 
-                 set_prior("normal(0,0.3)", class = "b", coef="xj"),  
-                 set_prior("inv_gamma(13.75134,5.295736)", class = "sd", coef="xj", group = "sub1"), 
-                 set_prior("normal(0,0.3)", class = "b", coef="dif1"), 
-                 set_prior("inv_gamma(13.75134,5.295736)", class = "sd", coef="dif1", group = "sub1"),
-                 set_prior("normal(0,0.3)", class = "b", coef="dif2"),  
-                 set_prior("inv_gamma(13.75134,5.295736)", class = "sd", coef="dif2", group = "sub1"),
-                 set_prior("normal(0,0.3)", class = "b", coef="dif3"), 
-                 set_prior("inv_gamma(13.75134,5.295736)", class = "sd", coef="dif3", group = "sub1"),
-                 set_prior("normal(0,0.3)", class = "b", coef="dif4"),  
-                 set_prior("inv_gamma(13.75134,5.295736)", class = "sd", coef="dif4", group = "sub1"),
-                 set_prior("inv_gamma(13.75136, 4.102063)", class = "sigma"))
-
-```
-
-\normalsize If you do not specify a prior for a certain parameter, the default prior as explained in the prior section will be applied.
-It is possible to save all priors in an R `data.frame` and provide these in the `brm` function, this and the full code with explanation for fitting the full symbolic distance effect model using `brms` is shown in our [Online Supplement D: *Bayesian Hierarchical Modeling in brms*](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/blob/main/D%20-%20Tutorial%20Normal%20Model%20brms/Online-Supplement-D---Bayesian-Hierarchical-Modeling-in-brms.pdf).^[In our example of the digit classification task we are applying a truncated normal prior distribution for only one parameter.
-This requires a slightly different `brms` formula and prior specification, explained in the Online Supplement.]
-
-# Estimation of the Digit Classification Task
-## Data 
-For our example, we will use the data set by Rouder et al.
-(2005).
-The data are available under a free license on GitHub [https://github.com/PerceptionCognitionLab/data0/tree/master/lexDec-dist5]((https://github.com/PerceptionCognitionLab/data0/tree/master/lexDec-dist5)).
-The data set contains observations from 54 participants who took part in a digit classification task.
-Each participant performed 360 trials in six blocks.
-For the analysis, only the correct responses are included.
-To account for practice effects, the first 25 trials of the first block are removed from the analysis.
-The trials after the breaks between the blocks are also removed from the analysis to correct for lack of concentration.
-In addition, trials with very fast RTs (< 250 MS) and very slow RTs (> 2000 MS) are excluded.
-Two participants were excluded because they did not respond properly to the task.
-Similar criteria were applied by @haaf2019capturing.
-This results in 52 participants and a total of 17031 observations.
-In the following two sections we will describe the model fitting and parameter estimation procedure for the normal and log-normal model, respectively.
-
-\newpage
-
-## Normal Model
-
-```{r traceplots3p, echo=FALSE, fig.align = "left", fig.height = 11, fig.width = 4, out.width="100%", fig.cap="Panel A and B show the trace plot of the digit parameter $\\mu_{\\delta_{6}}$ for both packages.
-Panel C shows the estimated $\\hat{R}$ for all individual effects per parameter group.
-The stars represent the $\\hat{R}$ of the general effects.
-Panel D illustrates the variation in the estimated number of effective samples per parameter group.
-The dashed lines represent the total number of iterations for each package.
-The stars represent the number of effective samples of the general effects."}
-knitr::include_graphics("Images/Fig8_traceplots3pnew_v2.png")
-```
-
-### Model Fitting 
-The model is fitted using the packages `rstan` and `brms`.
-For `rstan`, four chains are run with 4000 iterations per chain, of which 1000 serve as warm-up.
-This results in a total of 12000 samples from the joint posterior distribution.
-For `brms`, four chains are run with 6000 iterations per chain, of which 1000 serve as warm-up.
-This results in a total of 20000 draws from the joint posterior distribution.
-
-### Model Diagnostics   
-Before we can interpret the results, we have to evaluate whether they are reliable.
-Specifically, it has to be checked whether the sampling chains of the parameters have converged, meaning that a stationary posterior distribution has been reached [@Vehtari_2021].
-If convergence is not achieved, the posterior estimates may change substantially if the sampling algorithm was run with more iterations or different starting values.
-The most common convergence measures are trace plots and the $\hat{R}$ statistic.
-We will also discuss an additional check provided by `rstan` and `brms`, namely, the number of effective samples.
-For a critical evaluation of currently used convergence statistics, we recommend the interested reader to consult Vehtari et al.
-(2019).
-
-#### Trace plot
-
-Trace plots show the sampled parameter value at each iteration.
-If the distributions have converged, the plot should look like a hairy caterpillar [Lee & Wagenmakers, -@lee2013bayesian, p.
-99]: the iterations move up and down, but have much overlap in the middle.
-If multiple chains are used, iterations are shown per chain.
-In this case, the chains should overlap.
-If the posterior distribution did not converge, one chain could, for instance, be found on top of the figure, while all other chains are on the bottom [for an illustration see @vehtari2019rank].
-Or, perhaps the first half of the chain is on the top of the graph and the second half on the bottom.
-The trace plot does not include burn-in samples, since these will not be used for parameter estimation.
-
-When inspecting the trace plots of the symbolic distance effect parameters in the normal model, the posterior distributions of the parameters appear to have converged as the iterations come together and look similar to a caterpillar.
-The trace plot of one of the parameters, $\delta_{6, i}$, is shown in panel A for `rstan` and panel B for `brms` of Figure \@ref(fig:traceplots3p).
-The trace plot in panel B shows less overlap in the middle, indicating that `brms` had a bit more dependency within chains.
-This might be a sign of slight issues.
-However, not that much to be too concerned.
-Trace plots for other parameters can be found in
-[Online Supplement G](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/tree/main/G%20-%20Trace%20Plots%20).
-
-#### $\hat{\bm{R}}$
-Next to a visual inspection of convergence, a numerical inspection is usually performed as well.
-A common approach is to check $\hat{R}$, also called the Gelman-Rubin Diagnostic [@gelman1992inference].
-$\hat{R}$ is the ratio of between-chain variance and within-chain variance [@sorensen2015bayesian].
-If the chains diverge, the between-chain variance will be higher than the within-chain variance, resulting in a $\hat{R}$ greater than 1.
-This indicates that the chains have not converged.
-There are different types of $\hat{R}$, such as the split- and rank-based $\hat{R}$ [@gelman2013bayesian; @vehtari2019rank].
-Even though researchers tend to trust convergence statistics more than visual inspection, these statistics are also not fool-proof.
-Therefore, the development of better convergence statistics is a current topic in Bayesian analysis methods.
-
-The $\hat{R}$ values for all individual and general symbolic distance effect parameters computed by `rstan` and `brms` are shown in panel C of Figure \@ref(fig:traceplots3p).
-The figure shows that $\hat{R}$'s are not substantially greater than 1.
-Therefore, it seems that the distributions have converged.
-
-#### Number of Effective Samples 
-Another popular convergence statistic is the number of effective samples.
-This statistic is concerned with the dependence of posterior samples.
-In MCMC algorithms, the samples are to some degree dependent on one another: the parameter values at iteration *i* are similar to the parameter values at iteration *i-1*.
-The number of effective samples represents the estimated total number of *independent* draws from the posterior for every model parameter [@stan2018b, section 15.4].
-If the number of effective samples is low, this indicates a high dependency of iterations in the sampling procedure.
-Therefore, the number of effective samples can be seen as a measure of the amount of new information about the posterior distribution that is provided by the total number of samples drawn from the posterior.
-A common rule for interpretation is that the number should be close to or even greater than the number of total iterations [@vehtari2019rank; @neff2018].
-
-For the symbolic distance effect, the estimated number of effective samples is provided in panel D of Figure \@ref(fig:traceplots3p).
-For the package `rstan` the number is generally high, much higher than the actual number of samples depicted by the green dashed line.
-The `brms` package shows much more variation.
-The lowest effective sample sizes can be found for the intercept parameters.
-Table 1 also displays the number of effective samples for the general effects parameters in the `rstan` model.
-If the number of effective samples is too low, `brms` and `rstan` will return a warning.
-Here, this was not case, and all other diagnostics were good, so we continue with the interpretation of the estimates.
-
-### Estimation Results
-Considering acceptable to good convergence diagnostics, we can now interpret the estimation results.
-First, we present the general effects, then we visualize the individual variability of these effects.
-
-```{r modelposteriorsmufigure3, echo=FALSE, fig.cap="The posterior distributions for the general effects estimated by the two packages.
-The middle line within the distributions represents the posterior mean.
-The shaded area within the distributions represent 95\\% credible intervals.", fig.align = "left", fig.height = 11, fig.width = 1, out.width="100%"}
-knitr::include_graphics("Images/Fig9_posteriormu_v2.png")
-```
-
-#### General Effects 
-For the symbolic distance effect, there are two main questions we want to answer.
-First, we want to know whether there is a digit effect and, second, we want to know whether there is a side effect.
-Therefore, the key model parameters for general effects are $\mu_{\beta}$, $\mu_{\delta_{3}}$, $\mu_{\delta_4}$, $\mu_{\delta_6}$, and $\mu_{\delta_{7}}$.
-In the following section, we will focus our reporting on these parameters.
-
-The posterior distributions for the focal parameters in the normal model are presented in Figure \@ref(fig:modelposteriorsmufigure3).
-The posterior means estimated by `rstan` and `brms` can be found in [Online Supplement H](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/tree/main/H%20-%20Posterior%20Distributions%20Tables).
-The credible interval displayed in the figure is the Bayesian version of a confidence interval and provides a measure of uncertainty about the parameter value [@wagenmakers2018bayesian].
-
-```{r geneftable, results = "hide", cache = TRUE, warning = FALSE}
-# example - version 1 with SE 
-rstanfithier8 <- readRDS(file = "R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
-fitrstanhiertable <- summary(rstanfithier8, pars = c("mu", "mu2", "mu3", "mu4", "mu5", "mu6"))$summary[,c("mean", "se_mean", "2.5%", "97.5%", "n_eff", "Rhat")]
-# as.data.frame(fitrstanhiertable)
-fitrstanhiertable2 <- as.data.frame(fitrstanhiertable)
-rownamesfitrstanhiertable <-  c("$\\mu_{\\gamma}$","$\\mu_{\\beta}$", "$\\mu_{\\delta_{7}}$", "$\\mu_{\\delta_{6}}$", "$\\mu_{\\delta_{4}}$", "$\\mu_{\\delta_{3}}$")
-colnames(fitrstanhiertable2) <- c("Mean", "SE", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")
-
-fitrstanhiertable3 <- add_column(fitrstanhiertable2, Parameters = rownamesfitrstanhiertable, .before = "Mean")
-#apa_table(fitrstanhiertable3, row.names = FALSE, caption = "Posterior Mean, Standar Error (SE) of the Mean, Lower and Upper Bound of the 95\\% Credible Interval, the Number of Effective Samples, and the $\\hat{R}$ of the General Effect Parameters as estimated by Rstan.", align = c("l", "c", "c", "c", "c", "c", "c"), digits = 3, escape = FALSE)
-
-# example - version 2 without SE 
-rstanfithier8 <- readRDS(file = "R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
-fitrstanhiertable <- summary(rstanfithier8, pars = c("mu", "mu2", "mu3", "mu4", "mu5", "mu6"))$summary[,c("mean", "2.5%", "97.5%", "n_eff", "Rhat")]
-# as.data.frame(fitrstanhiertable)
-fitrstanhiertable2 <- as.data.frame(fitrstanhiertable)
-rownamesfitrstanhiertable <-  c("$\\mu_{\\gamma}$","$\\mu_{\\beta}$", "$\\mu_{\\delta_{7}}$", "$\\mu_{\\delta_{6}}$", "$\\mu_{\\delta_{4}}$", "$\\mu_{\\delta_{3}}$")
-colnames(fitrstanhiertable2) <- c("Mean", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")
-
-fitrstanhiertable3 <- add_column(fitrstanhiertable2, Parameters = rownamesfitrstanhiertable, .before = "Mean")
-apa_table(fitrstanhiertable3, row.names = FALSE, caption = "Posterior Mean, Lower and Upper Bound of the 95\\% Credible Interval, Number of Effective Samples, and $\\hat{R}$ of the General Effect Parameters as estimated by rstan.", align = c("l", "c", "c", "c", "c", "c"), digits = 3, placement = "h", escape = FALSE)
-```
-The parameter estimates produced by the two packages look very similar.
-The `brms` estimates seem to be slightly higher with slightly more narrow posterior distributions compared to `rstan`.
-All parameters are presented on a scale of seconds.
-As expected, the effects are small.
-For instance, $\mu_{\delta_{7}}$ has a posterior mean of `r printnum(fitrstanhiertable3["mu3", "Mean"], digits = 3)`, corresponding to a 11 millisecond effect.
-
-For the digit effects, positive parameter estimates indicate that the response is slower for the non-baseline digits, that is, digits that are closer to 5.
-The parameters $\mu_{\delta_{6}}$ and $\mu_{\delta_{4}}$ are distributed around higher values and their credible intervals do not contain zero.
-Note that this result is consistent with the symbolic distance hypothesis that the digits 4 and 6 have the highest response times because they are closest to the 5, the comparison value.
-The 95% credible intervals of $\mu_{\delta_{7}}$ and $\mu_{\delta_{3}}$ contain zero.
-Therefore, it is plausible that the response times do not differ substantially from the baseline condition.
-The posterior distribution of the side effect $\mu_{\beta}$ is centered around zero with a posterior mean of `r printnum(fitrstanhiertable3["mu2", "Mean"], digits = 3)` and a 95% credible interval ranging from `r printnum(fitrstanhiertable3["mu2", "Lower Bound"], digits = 3)` to `r printnum(fitrstanhiertable3["mu2", "Upper Bound"], digits = 3)`.
-This indicates a small to non-existent side effect.
-
-#### Individual effects
-
-```{r inddevtable, results = "asis", cache = TRUE}
-# example - With SE 
-# rstanfithier8 <- readRDS(file = "R objects/rstanmodelfit_final_kopie.rds")
-fitrstanhiertable2 <- summary(rstanfithier8, pars = c("g", "g2", "g3", "g4", "g5", "g6"))$summary[,c("mean", "se_mean", "2.5%", "97.5%", "n_eff", "Rhat")]
-# as.data.frame(fitrstanhiertable)
-fitrstanhiertable22 <- as.data.frame(fitrstanhiertable2)
-rownamesfitrstanhiertable2 <-  c("$\\sigma_{\\gamma}^2$","$\\sigma_{\\beta}^2$", "$\\sigma_{\\delta_{7}}^2$", "$\\sigma_{\\delta_{6}}^2$", "$\\sigma_{\\delta_{4}}^2$", "$\\sigma_{\\delta_{3}}^2$")
-colnames(fitrstanhiertable22) <- c("Mean", "SE", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")
-
-fitrstanhiertable32 <- add_column(fitrstanhiertable22, Parameters = rownamesfitrstanhiertable2, .before = "Mean")
-# apa_table(fitrstanhiertable32, row.names = FALSE, caption = "Posterior Variance, Standar Error (SE) of the Variance, Lower and Upper Bound of the 95\\% Credible Interval, the Number of Effective Samples, and the $\\hat{R}$ of the General Effect Parameters as estimated by Rstan.", align = c("l", "c", "c", "c", "c", "c", "c"), digits = 3, escape = FALSE)
-
-# example - without SE 
-# rstanfithier8 <- readRDS(file = "R objects/rstanmodelfit_final_kopie.rds")
-fitrstanhiertable2 <- summary(rstanfithier8, pars = c("g", "g2", "g3", "g4", "g5", "g6"))$summary[,c("mean", "2.5%", "97.5%", "n_eff", "Rhat")]
-# as.data.frame(fitrstanhiertable)
-fitrstanhiertable22 <- as.data.frame(fitrstanhiertable2)
-rownamesfitrstanhiertable2 <-  c("$\\sigma_{\\gamma}^2$","$\\sigma_{\\beta}^2$", "$\\sigma_{\\delta_{7}}^2$", "$\\sigma_{\\delta_{6}}^2$", "$\\sigma_{\\delta_{4}}^2$", "$\\sigma_{\\delta_{3}}^2$")
-colnames(fitrstanhiertable22) <- c("Mean", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")
-
-fitrstanhiertable32 <- add_column(fitrstanhiertable22, Parameters = rownamesfitrstanhiertable2, .before = "Mean")
-apa_table(fitrstanhiertable32, row.names = FALSE, caption = "Posterior Variance, Lower and Upper Bound of the 95\\% Credible Interval, Number of Effective Samples, and $\\hat{R}$ of the Variance Parameters as estimated by Rstan.", align = c("l", "c", "c", "c", "c", "c"), digits = 3, placement = "h", escape = FALSE)
-```
-Thanks to the hierarchical structure of the model, we can investigate whether the general effects hold for all the individuals by inspecting the individual variation and the individual parameter estimates.
-The individual variation is represented by the variance or standard deviation parameters in the model.
-The summaries of the posterior distributions of the variance parameters as estimated by `rstan` are shown in Table \@ref(tab:inddevtable).
-The estimates for the standard deviations by `brms` are smaller compared to `rstan` and can be found in [Online Supplements D and H](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment).
-
-Posterior means of the variances for the digit and side effects are all very close to `r printnum(fitrstanhiertable32["g3", "Mean"], digits = 3)` which corresponds to a standard deviation of `r printnum(sqrt(fitrstanhiertable32["g3", "Mean"]), digits = 3)` seconds, that is, 141 milliseconds.
-Additionally, the 95% credible intervals do not contain zero.
-If we consider the effect of digit 7 wit an 11 millisecond general effect this is a substantial amount of individual variation.
-Therefore, we conclude that there are individual differences in the symbolic distance effect.
-
-```{r randomparfigure, echo=FALSE, fig.cap="The posterior means for individual effect parameters with 95\\% credible intervals as estimated by \\textit{rstan}, shown in increasing order.
-The dashed line represents the general posterior mean.
-Pink intervals contain zero, blue interval do not contain zero.", fig.align = "left", fig.height = 1, fig.width = 0.5, out.width="100%"}
-knitr::include_graphics("Images/Fig10_randompar_v2.png")
-
-```
-
-Figure \@ref(fig:randomparfigure) shows the posterior means and 95% credible intervals for the individual effects of all 52 participants.
-Estimates are printed in increasing order, from lowest to highest.
-Pink intervals indicate that the 95% credible interval contains zero, while blue intervals are either entirely above or below zero.
-A corresponding figure with individual effect estimates for `brms` can be found in [Online Supplement D](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/blob/main/D%20-%20Tutorial%20Normal%20Model%20brms/Online-Supplement-D---Bayesian-Hierarchical-Modeling-in-brms.pdf).
-
-The largest variation across participants can be seen for the intercept parameter.
-This means that participants vary considerably in their baseline response times.
-There is also considerable individual variation in the digit effects ($\delta$'s).
-However, most of the intervals contain zero which implies that the effects might be small or non-existent.
-The variability is most pronounced for the effect parameters of the digits close to 5.
-For these digits, there are also more intervals that do not contain zero, indicating that an effect might be present on the subject level.
-The posterior mean of the individual side effect ($\beta_{i}$) varies between -0.05 and 0.05.
-Again, all but four 95% credible intervals include zero indicating uncertainty about the existence of an effect for anyone.
-The credible intervals of the side effect and digit effects $\delta_{7}$ and $\delta_{3}$ indicate that different subjects may show opposite effects [@haaf2019some].
-
-Finally, Figure \@ref(fig:modelestimatesfigure3) shows how the relationship between individuals' digit effects.^[The design and code for this figure are by @j_van_langen_2020_3715576.] The posterior means of individual effects are represented by the dots.
-The right side of each plot represents the distribution of the individual estimates.
-Note that this distribution is not the posterior distribution of the general effect.
-It rather corresponds to the left side of Figure 6.
-
-```{r modelestimatesfigure3, echo=FALSE, fig.cap="Model estimates for digit effect parameters in \\textit{rstan}.
-The points represent the mean parameter estimates for each individual.
-The violin plots on the right side show the variance of the individual parameter estimates.
-The size of effects is not very constistent within a person.", fig.align = "left", fig.height = 11, fig.width = 1, out.width="75%"}
-knitr::include_graphics("Images/Fig11_modelestimatefigure2_onlyrstan_v2.png")
-```
-
-## Log-normal Model 
-### Model Estimation 
-The model is fitted using the packages `rstan` and `brms`.
-For `rstan`, we use the priors as specified in Equation \@ref(eq:chosenpriorslogvar).
-For `brms`, we use the priors in Equation \@ref(eq:chosenpriorslogsd).
-The rest of the settings are equivalent for both packages.
-We ran four chains with 4000 iterations per chain, including 1000 warm-up iterations.
-This results in a total of 12000 samples from the posterior distribution.
-
-### Model Diagnostics 
-Before interpreting the results, we have to check if the posterior distributions of the parameters have converged.
-The trace plot of $\mu_{\delta_{6}}$ is shown panel A for `rstan` and B for `brms` of Figure \@ref(fig:traceplotlog).
-Trace plots for all other parameters can be found in [Online Supplement G](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/tree/main/G%20-%20Trace%20Plots%20).
-All plots look like a hairy caterpillar.
-Therefore, they indicate that the posterior distributions of the parameters have converged.
-$\hat{R}$ values shown in Figure \@ref(fig:traceplotlog)C for all parameters are all close to 1 and most of the number of effective samples shown in Figure \@ref(fig:traceplotlog)D exceed the total number of iterations.
-The number of effective samples for the general effects are comparatively low but still sufficient.
-In combination with the other diagnostics, we conclude that the posterior distributions of the parameters have converged and we proceed with the interpretation of the results.
-
-(ref:captiontraceplotlog) Panel A and B show the trace plot of the digit parameter $\\mu_{\\delta_{6}}$ in the log-normal model for both packages.
-Panel C shows the estimated $\\hat{R}$ for all individual effects per parameter group.
-The stars represent the $\\hat{R}$ of the general effects.
-Panel D illustrates the variation in the estimated number of effective samples per parameter group.
-The dashed lines represent the total number of iterations for each package.
-The stars represent the number of effective samples of the general effects.
-
-```{r traceplotlog, echo=FALSE, fig.cap="(ref:captiontraceplotlog)", fig.align = "left", fig.height = 11, fig.width = 4, out.width="100%"}
-knitr::include_graphics("Images/Fig12_traceplots3plog_v2.png")
-```
-
-### Results
-#### General effects  
-We inspect the posterior distributions for the general effect parameters presented in Figure \@ref(fig:postplotlog). 
-```{r postplotlog, echo=FALSE, fig.cap="The posterior distributions for the general effects in the log-normal model estimated by the two packages.
-The middle line within the distributions represents the posterior mean.
-The shaded area within the distributions represent 95\\% credible intervals.", fig.align = "left", fig.height = 11, fig.width = 1, out.width="100%"}
-knitr::include_graphics("Images/Fig13_posteriormulog_v2.png")
-```
-
-```{r geneftablelog, results = "hide", cache = TRUE, warning = F}
-# example - with SE 
-rstanfithier8log <- readRDS(file = "R objects/rstan/log-normal model/hier_logmodelc_adj13012022.rds")
-fitrstanhiertablelog <- summary(rstanfithier8log, pars = c("mu", "mu2", "mu3", "mu4", "mu5", "mu6"))$summary[,c("mean", "se_mean", "2.5%", "97.5%", "n_eff", "Rhat")]
-# as.data.frame(fitrstanhiertable)
-fitrstanhiertable2log <- as.data.frame(fitrstanhiertablelog)
-rownamesfitrstanhiertablelog <-  c("$\\mu_{\\gamma}$","$\\mu_{\\beta}$", "$\\mu_{\\delta_{7}}$", "$\\mu_{\\delta_{6}}$", "$\\mu_{\\delta_{4}}$", "$\\mu_{\\delta_{3}}$")
-colnames(fitrstanhiertable2log) <- c("Mean", "SE", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")
-
-fitrstanhiertable3log <- add_column(fitrstanhiertable2log, Parameters = rownamesfitrstanhiertablelog, .before = "Mean")
-# apa_table(fitrstanhiertable3log, row.names = FALSE, caption = "Posterior Mean, Standar Error (SE) of the Mean, Lower and Upper Bound of the 95\\% Credible Interval, the Number of Effective Samples, and the $\\hat{R}$ of the General Effect Parameters as estimated by Rstan.", align = c("l", "c", "c", "c", "c", "c", "c"), digits = 3, escape = FALSE)
-
-# example - without SE 
-rstanfithier8log <- readRDS(file = "R objects/rstan/log-normal model/hier_logmodelc_adj13012022.rds")
-fitrstanhiertablelog <- summary(rstanfithier8log, pars = c("mu", "mu2", "mu3", "mu4", "mu5", "mu6"))$summary[,c("mean", "2.5%", "97.5%", "n_eff", "Rhat")]
-# as.data.frame(fitrstanhiertable)
-fitrstanhiertable2log <- as.data.frame(fitrstanhiertablelog)
-rownamesfitrstanhiertablelog <-  c("$\\mu_{\\gamma}$","$\\mu_{\\beta}$", "$\\mu_{\\delta_{7}}$", "$\\mu_{\\delta_{6}}$", "$\\mu_{\\delta_{4}}$", "$\\mu_{\\delta_{3}}$")
-colnames(fitrstanhiertable2log) <- c("Mean", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")
-
-fitrstanhiertable3log <- add_column(fitrstanhiertable2log, Parameters = rownamesfitrstanhiertablelog, .before = "Mean")
-apa_table(fitrstanhiertable3log, row.names = FALSE, caption = "Posterior Mean, Lower and Upper Bound of the 95\\% Credible Interval, Number of Effective Samples, and $\\hat{R}$ of the General Effect Parameters as estimated by Rstan.", align = c("l", "c", "c", "c", "c", "c"), digits = 3, placement = "h", escape = FALSE)
-```
-
-```{r modelfitlog, eval = FALSE}
-# Based on model estimates 
-> exp(-0.56 - (1/2)*0.01 + 0.08)
-[1] 0.6156972
-> exp(-0.56 - (1/2)*0.01 + 0.03)
-[1] 0.5856693
-> 0.6156972 - 0.5856693
-[1] 0.0300279
-> exp(-.56 + .5  (-.01) + .03) - exp(-.56 + .5  (-.01))
-Error: attempt to apply non-function
-> exp(-.56 + .5 * (-.01) + .03) - exp(-.56 + .5  * (-.01))  # difference 3 and 2 
-[1] 0.01730914
-> exp(-.56 + .5 * (-.01) + .08) - exp(-.56 + .5  * (-.01))  # difference 4 and 2? 
-[1] 0.04733705
-```
-We note that the values of the parameter estimates for the log-normal model cannot be interpreted in the same way as in normal model.
-The reason is that the mean parameter of the log-normal distribution does not correspond to its expected value.
-To better understand the size of an effect that corresponds to the parameter values, some calculations are required.
-For instance, if we would like to know the median difference in response times for digit 3 and 2, we calculate the difference between the exponential of the RT estimate for the condition digit = 3 and the exponential of the RT estimate for the condition digit = 2:
-
-\begin{equation}
-\begin{aligned}
-&e^{(\mu_{\gamma} + x_j \mu_{\beta} + w_j \mu_{\delta_{3}})} - e^{(\mu_{\gamma} +  x_j \mu_{\beta})} \\
-&= e^{(-.56 + .5 \times (-.01) + 1 \times 0.03)} - e^{(-.56 + .5 \times (-.01))} \\ 
-&= 0.017 
-\end{aligned}
-\end{equation}
-This value corresponds to the median estimate of $\mu_{\delta_{3}}$ from the normal model.
-Calculating the mean estimate of $\mu_{\delta_{3}}$ would be similar, but more complicated as the mean of a log-normal distribution also depends on its variance.
-
-One property of these parameter estimates we can immediately interpret is their sign (i.e., positive, negative or zero).
-A positive estimate indicates that the RT increases, whereas a negative effect indicates a decrease.
-When inspecting the general digit effect parameters, they all appear to positively influence the RT.
-This is in line with our expectations since the baseline RT in the model is set equal to the RT for the digits furthest away from 5 and the symbolic distance effect postulates that RT should increase as the digits become closer to five.
-The side effect is close to zero, with a posterior mean of `r printnum(fitrstanhiertable3log["mu2", "Mean"], digits = 3)` and a 95% credible interval from `r printnum(fitrstanhiertable3log["mu2", "Lower Bound"], digits = 3)` to `r printnum(fitrstanhiertable3log["mu2", "Upper Bound"], digits = 3)`.
-Thus, the general effect of side seems small or negligible.
-
-#### Individual effects 
-Next, we inspect the individual deviations from the general effects.
-Table \@ref(tab:inddevtablelog) presents the estimates of the variance parameters by `rstan`.
-Results for brms can be found in [Online Supplement H](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/blob/main/H%20-%20Posterior%20Distributions%20Tables/Online-Supplement-H---Tables-with-Information-on-Posterior-Distributions.pdf).
-```{r inddevtablelog, results = "asis", cache = TRUE}
-# example - with SE 
-# rstanfithier8log <- readRDS(file = "R objects/hier_logmodelc_adj.rds")
-fitrstanhiertable2log <- summary(rstanfithier8log, pars = c("g", "g2", "g3", "g4", "g5", "g6"))$summary[,c("mean", "se_mean", "2.5%", "97.5%", "n_eff", "Rhat")]
-# as.data.frame(fitrstanhiertable)
-fitrstanhiertable2log2 <- as.data.frame(fitrstanhiertable2log)
-rownamesfitrstanhiertable2log2 <-  c("$\\sigma_{\\gamma}^2$","$\\sigma_{\\beta}^2$", "$\\sigma_{\\delta_{7}}^2$", "$\\sigma_{\\delta_{6}}^2$", "$\\sigma_{\\delta_{4}}^2$", "$\\sigma_{\\delta_{3}}^2$")
-colnames(fitrstanhiertable2log2) <- c("Mean", "SE", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")
-
-fitrstanhiertable32log <- add_column(fitrstanhiertable2log2, Parameters = rownamesfitrstanhiertable2log2, .before = "Mean")
-# apa_table(fitrstanhiertable32log, row.names = FALSE, caption = "Posterior Variance, Lower and Upper Bound of the 95\\% Credible Interval, Number of Effective Samples, and $\\hat{R}$ of the General Effect Parameters as estimated by Rstan", align = c("l", "c", "c", "c", "c", "c", "c"), digits = 3, escape = FALSE)
-
-# example - without SE 
-# rstanfithier8log <- readRDS(file = "R objects/hier_logmodelc_adj.rds")
-fitrstanhiertable2log <- summary(rstanfithier8log, pars = c("g", "g2", "g3", "g4", "g5", "g6"))$summary[,c("mean", "2.5%", "97.5%", "n_eff", "Rhat")]
-# as.data.frame(fitrstanhiertable)
-fitrstanhiertable2log2 <- as.data.frame(fitrstanhiertable2log)
-rownamesfitrstanhiertable2log2 <-  c("$\\sigma_{\\gamma}^2$","$\\sigma_{\\beta}^2$", "$\\sigma_{\\delta_{7}}^2$", "$\\sigma_{\\delta_{6}}^2$", "$\\sigma_{\\delta_{4}}^2$", "$\\sigma_{\\delta_{3}}^2$")
-colnames(fitrstanhiertable2log2) <- c("Mean", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")
-
-fitrstanhiertable32log <- add_column(fitrstanhiertable2log2, Parameters = rownamesfitrstanhiertable2log2, .before = "Mean")
-apa_table(fitrstanhiertable32log, row.names = FALSE, caption = "Posterior Variance, Lower and Upper Bound of the 95\\% Credible Interval, the Number of Effective Samples, and the $\\hat{R}$ of the Variance Parameters as estimated by Rstan.", align = c("l", "c", "c", "c", "c", "c"), digits = 3, placement = "h", escape = FALSE)
-```
-The estimated parameter values for the variance terms are very small, but different from zero.
-This indicates that individuals may vary slightly with respect to the general effect parameters.
-The estimates for the standard deviations by `brms` are similar compared to the estimates by `rstan`, but slightly smaller.
-The exact estimates by `brms` can be found in [Online Supplements F and H](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment).
-Figure \@ref(fig:randomparplotlog) shows the individual estimates for the digit and side effects.
-The pink lines indicate that the 95% credible interval of the individual effect contains zero.
-The estimated general effect is represented by the dashed line.
-For $\beta$, $\delta_{7}$, and $\delta_{3}$, most 95% credible intervals contain zero.
-For $\delta_{6}$ and $\delta_{4}$, most credible intervals are entirely larger than zero, indicating an increase in RT for most participants.
-
-```{r randomparplotlog, echo=FALSE, fig.cap="The posterior means for individual effect parameters in the log-normal model with 95\\% credible intervals as estimated by \\textit{rstan}, shown in increasing order.
-The dashed line represents the general posterior mean.
-Pink intervals contain zero, blue interval do not contain zero.", fig.align = "left", fig.height = 1, fig.width = 0.5, out.width="100%"}
-knitr::include_graphics("Images/Fig14_randomparlog_v2.png")
-```
-We can further inspect the individual estimates for the digit effects in Figure \@ref(fig:modeldigitestplotlog).
-The figure shows the individual estimates for the digit effects with the distribution of the variability of the individual estimates (not the posterior distribution of the general effect).
-The figure illustrates that there is considerable variance in the point estimates of the individual digit effects.
-However, compared to the normal model there is considerably more hierarchical shrinkage.
-\Revision{This means that in the log-normal model, the individual effects are corrected more towards the group mean, compared to the normal model.
-Or, the group mean in the log-normal model is not affected as much by individuals showing a divergent trend compared to the normal model.
-Whether we should prefer a model with more shrinkage depends on the theoretical considerations.
-If more individual variation is expected, less shrinkage would be preferred.
-However, the advantage of more shrinkage is that outliers do not influence the estimation of the effect as much.
-}
-
-# Model Comparison 
-The symbolic distance effect postulates that response times (RTs) are slower when digits are closer to 5 (digit effects), and that RTs may be influenced by whether the target digit is smaller or greater than 5 (side effect).
-Using Bayesian estimation, we are not able to directly test these hypotheses.
-Credible intervals often contain zero, but they always also contain lots of other values that have a chance of being the "true" parameter value, making them unsuitable to test hypotheses.
-Therefore, our goal for the following section is to set up a Bayesian hypothesis test for the parameters of interest in our model.
-Specifically, we will achieve this by comparing the predictive accuracy of a null model where we set one or some parameters to zero, and and effects models where these parameters are unconstrained.
-
-```{r modeldigitestplotlog, echo=FALSE, fig.cap="Model estimates for digit effect parameters in the log-normal model in \\textit{rstan}.
-The points represent the mean parameter estimates for each individual.
-The violin plots on the right side show the variance of the individual parameter estimates.", fig.align = "left", fig.height = 11, fig.width = 1, out.width="75%"}
-knitr::include_graphics("Images/Fig15_inddiflogplot_onlyrstan_v2.png")
-```
-
-## Bayes factor
-The Bayes factor (BF) is a measure of how well one model predicts the data compared to another [@jeffreys1961theory; @kass1995bayes; @rouder2018teaching].
-The Bayes factor is defined as the probability of the data under one model compared to the probability of the data under another model.
-For instance, the Bayes factor comparing a null model ($\mathcal{M}_0$) and a full model ($\mathcal{M}_{\text{F}}$) can be obtained by:
-
-\begin{equation}
-\text{BF}_{0\text{F}} = \frac{P(\bm{Y} | \mathcal{M}_0)}{P(\bm{Y} | \mathcal{M}_{\text{F}})}. 
-\end{equation}
-
-The Bayes factor can also be interpreted as the relative evidence of two competing models.
-Bayes factors have a few very convenient characteristics.
-One of them is that Bayes factors are transitive, that is, evidence for the full model over the null model can be obtained by flipping numerator and denominatior in equation 13.
-For a full overview of the advantages of Bayesian inference with Bayes factor, see Wagenmakers et al.
-[-@wagenmakers2018bayesian].
+### Section on marginal likelihood STILL WORK IN PROGRESS
 
@@
 
-Imagine this results in $\text{BF}_{\text{nl}} = 100$.
-This means that the data are 100 times more likely under the normal model than under the log-normal model, clear evidence in favor of the normal model.
-\textcolor{blue}{But how do we know whether a Bayes factor of 100-to-1, or 10-to-1 or 3-to-1 fulfills?}.
-As a reminder, Bayes factors are the relative predictive accuracy of two models for the observed data.
-Therefore, Bayes factors are odds or ratios.
-Odds themselves are directly interpretable without the need for decisions or cutoffs.
-For example, if a presidential candidate is favored 10-to-1 over another one, then these are just the odds, and it is beside the point whether these are large or not.
-According to this line of thought, Rouder et al.
-[-@rouder2018theories] argue that researchers should take a similar approach when interpreting and reporting Bayes factors.
-While there exist rules of thumb to categorize Bayes factors verbally, we suggest researchers use their research context to form substantive conclusions based on Bayes factors, not arbitrary rules of thumb.
+References to include:
 
-Same as the posterior distribution, Bayes factors can rarely be computed analytically due to the multidimensional integrals involved.
-The difficulty to compute Bayes factors is one of the major drawbacks of Bayesian inference.
-However, several algorithms are available to estimate Bayes factor from posterior samples.
-In the following section we will present two methods to obtain Bayes factors for the symbolic distance effect.
+Gelman, A., Vehtari, A., Simpson, D., Margossian, C. C., Carpenter, B., Yao, Y., ... & Modrák, M. (2020). Bayesian workflow. arXiv preprint arXiv:2011.01808.
 
-### Savage-Dickey Density Ratio  
-One method to compute Bayes factors is the Savage-Dickey density ratio [SDD-ratio; Dickey & Lientz, -@dickey1970weighted; Morey, Rouder, Pratte, & Speckman, -@morey2011using; Wagenmakers, Lodewyckx, Kuriyal, & Grasman, -@wagenmakers2010bayesian].
-This approach can be readily used to estimate the Bayes factor between two nested models that differ in only one parameter.
-A prime example for such a case is comparing an effect model with a null model.
-For example, we may compare a model with an overall side effect (effect model) to a model with no overall side effect (null model).
-In this case, the Bayes factor in favor of the side effect can be approximated by the ratio of the posterior density of the side effect and the prior density of the side effect at the point zero.
-More generally, the Savage Dickey approach works whenever one model fixes a parameter to a constant and another model assigns a prior distribution to the same parameter.
-
-Figure \@ref(fig:sdratioplotall) shows a visualization of the SDD-ratio for the side parameter in the normal model of the symbolic distance effect.
-The posterior distribution is depicted in orange, the prior in green.
-The Bayes factor can be computed as a ratio of the posterior and prior density at the parameter value of zero, as depicted by the dashed line connecting the two points on the density functions.
-
-```{r sdratioplotall, echo=FALSE, fig.cap="The prior and posterior distribution of the side parameter estimated by the \\textit{brms} package.
-The dots represent the height of the distributions at zero.
-The ratio of these dots indicates the Savage-Dickey density ratio.
-The Bayes factor represents the evidence for the hypothesis that the effect equals zero.
-A Bayes factor above one indicates that the evidence is in favor of the null hypothesis.", fig.align = "left", fig.height = 6, fig.width = 4, out.width="60%"}
-knitr::include_graphics("Images/Fig16_sdrplotside_v2.png")
-```
-With `rstan`, the SDD-ratio can be computed by fitting a density function to the posterior distribution obtained from the MCMC samples, and comparing the value of this density function at $\mu_{\beta} = 0$ to the value of the prior distribution at $\mu_{\beta} = 0$.
-This is illustrated in the code block below.
-First, we extract the posterior samples for the parameter $\mu_{\beta}$, the general effect of side.
-We apply the `logspline` function [@logspline] to the samples to obtain an estimate of the log-density of the posterior distribution.
-The density at the zero is calculated with the `dlogspline` function.
-Finally, we compare the estimated density of the posterior to the density of the prior.
-\textcolor{blue}{This process is explained in further detail by Van Ravenzwaaij and Etz} [-@van2021simulation].
-
-\scriptsize
-```{r codesnippetbfrstansdratio, eval=FALSE, echo=TRUE}
-library(logspline)
-# Get posterior samples for the side parameter mu2
-samples_side <- rstan::extract(model_fit)$mu2   
-log.posterior <- logspline(samples_side)                 # estimate of log-density post. distr. 
-posterior_at_zero <- dlogspline(0, log.posterior)        # compute posterior density 
-prior_at_zero <- dnorm(0, 0, 0.3)                        # computer prior density 
-BF01 <- posterior_at_zero/prior_at_zero                  # compare estimated density posterior to prior at point 0 
-```
-\normalsize
-In `brms`, the SDD-ratio can be computed using the built-in `hypothesis` function, as shown in the code snippet below.
-The resulting Bayes factor, also depicted in Figure \@ref(fig:sdratioplotall), is $BF_{01} = 18.28$, indicating evidence for the null model and against a side effect.
-&nbsp;
-
-\scriptsize
-```{r codesnippetbfbrms, eval=FALSE, echo=TRUE}
-# We will use the model fit from before, specified as model_fit
-hypothesis_fit <- hypothesis(model_fit,    # Object containing model fit 
-                             "mu2 = 0")   # The parameter we want to test with the null hypothesis 
-```
-\normalsize
-
-The Savage-Dickey approach has a number of drawbacks [see, for example, Heck, -@heck2019caveat].
-First, one issue is the quality of approximation for the SSD-ratio.
-If either the prior or, even more problematically, the posterior have a really low density at the test value (in this case zero), it becomes difficult to estimate the SSD-ratio from the samples.
-This issue often occurs when the effect is large, moving the posterior distribution away from zero.
-In many cases there might be no samples around zero, which means that the density estimate will be very inaccurate.
-Therefore, we would recommend only to use the SSD-ratio if you are confident that enough posterior samples are available from the tails of the distribution.
-One way of ensuring that is to drastically increase the number of drawn samples.
-
-Second, the Savage-Dickey approach only applies for a limited set of to-be-compared models.
-In the example used here, we compared a model where the overall side effect is allowed to vary with a model where it is zero.
-Both models, however, still allow for individual variability of the side effect.
-For the null model, individual variability implies that some individuals are expected to have an effect favoring larger numbers and other individuals are expected to have an effect favoring smaller numbers, but, on average, both groups perfectly balance out at zero.
-To us, this model seems nonsensical, and, more importantly, it is not what researchers have in mind if they want to test a model without side effect.
-We would prefer to use a null model where none of the participants have a side effect.
-Yet, the Savage-Dickey density ratio is not available for such a model compared to the effects model.
-
-### Bridge Sampling  
-Bridge sampling offers a more flexible way to obtain Bayes factors [@kass1995bayes].
-It allows the computation of Bayes factors for comparing models that differ in more than one parameter.
-Specifically, bridge sampling is a sampling-based algorithmic method to obtain the marginal likelihood of the data under a given model from the posterior samples [@meng1996simulating; @gronau2017tutorial].
-
-Using bridge sampling we are able to estimate Bayes factors between any models of interest for the classification task data. We identify four models of interest:
-
-1. Full model: This is the model we have been using throughout this paper. It states that there is an effect of side and an effect of digits for all participants. 
-2. Side model: This model specifies that participants show a side effect but no digit effect. All digit parameters $\delta_{3}$ to $\delta_{7}$ are set to zero. 
-3. Digit model: This model states that there are no side effects but that participants exhibit digits effects. All side parameters $\beta$ are set to zero. 
-4. Null model: In this model, there is no side or digit effect on the RT. The digit parameters $\delta_{3}$ to $\delta_{7}$ and side parameters $\beta$ are set to zero.
-
-\Revision{Note that models models 2-4 are nested in the most complex full model, and the null model is nested in models 2 and 3, but models 2 and 3 are not nested}.
-Additionally, the models differ by several parameters.
-For example, compared to the full model the Digit model restricts the side parameter for all participants to zero.
-Using bridge sampling, we can estimate the marginal likelihood for each model separately, and then compute Bayes factors as their ratios (Equation 14).
-
-To obtain the marginal likelihood for a specific model in `rstan` with bridge sampling [@gronau2017tutorial], the `bridgesampling` package can be used [@R-bridgesampling].
-To obtain the Bayes factor comparing the side model and the full model, we first have to fit both models.
-Then, we use the `bridge_sampler` function from the `bridgesampling` package to obtain the marginal likelihood of the data under each of the models.
-Finally, the models are compared using the `bf` function from the `bridgesampling` package.
-&nbsp;
-
-\scriptsize
-```{r codesnippetrstanestBF, eval = FALSE, echo = TRUE}
-
-# We will use the model fit of both models (called sidemodel and fullmodel)
-# (how to perform this is illustrated in the model estimation section)
-
-# Bridge sampling 
-library(bridgesampling)
-ML_Hside <- bridge_sampler(sidemodel)  # compute the marginal likelihood under the side model 
-ML_Hfull <- bridge_sampler(fullmodel)  # compute the marginal likelihood under the full model
-
-# Obtain Bayes factor (in this way evidence in favor of the side hypothesis)
-bridgesampling::bf(ML_Hside, ML_Hfull)
-
-```
-\normalsize
-In the `brms` package, Bayes factors based on bridge sampling can be obtained with the function `bayes_factor` as shown in the code snippet below.
-The function takes two fitted models as input.
-Therefore, both models first have to be fitted as explained in Model Estimation section of this manuscript.
-
-\scriptsize
-```{r codesnippetbrmsestBF, eval = FALSE, echo = TRUE}
-bayes_factor(fitsidemodel, fitfullmodel)$bf
-```
-\normalsize
-
-Even though bridge sampling is a flexible method for obtaining Bayes factors, it is still an estimation method based on posterior samples.
-Therefore, as with any Bayes factor estimation method, we advice to take two precautions with bridge sampling.
-First, it is always wise to increase the number of iterations when performing model comparison.
-The reason is that convergence of estimation of posterior distributions and convergence of estimation of the marginal likelihood can differ.
-Therefore, a rule of thumb is to use ten times the number of iterations than we would usually use for estimation.
-The second precaution is to conduct a stability analysis.
-Both the posterior samples from the model estimation procedure and the estimate of the marginal likelihood from the bridge sampling can be variable.
-Therefore, stability may best be assessed by repeating the entire process of fitting each model and applying the bridge sampling function several times.
-
-We assess the stability of the obtained Bayes factor by repeating the estimating process ten times with `rstan` and `brms`, respectively.
-The results are shown in Figure \@ref(fig:bsbfestimatesbrmsrstan).
-The figure shows the evidence in favor of the null, digit and side model against the full model.
-All Bayes factors are larger than one.
-Therefore, it can be concluded that the data are more likely under less complex models.
-There is almost no variation in the computed Bayes factors per model, indicating that the Bayes factor estimate is stable.
-However, the Bayes factor estimates in `rstan` generally show stronger evidence than Bayes factor estimates based on `brms`.
-It is important to note that this difference is not due to the bridge sampling function, but a result of a difference in parametrization.
-However, for both packages, the conclusion is the same: the data are least likely to have occurred under the full model.
-In case the estimates are unstable, it is recommended to further increase the number of posterior samples.
-
-```{r bsbfestimatesbrmsrstan, echo=FALSE, fig.cap="The BF estimates of the other models (0; null, side and digit model) against the full model (1).
-The green triangles represent the estimates using \\textit{rstan}.
-The orange dots represent the \\textit{brms} estimates.
-The dashed line represents a $\\text{BF}_{01}$ of 1.
-", fig.align = "left", fig.height = 11, fig.width = 4, out.width="70%"}
-knitr::include_graphics("Images/Fig17_normalbfplot01.png")
-```
-Finally, it is time to answer our research question: Is there a symbolic distance effect?
-When considering all Bayes factor estimates from Figure \@ref(fig:sdratioplotall) and \@ref(fig:bsbfestimatesbrmsrstan), the null model ($\text{BF}_{max} = 232 * 10^{105}$ for `rstan`) performs best compared to the side model ($\text{BF}_{max} = 458 * 10^{68}$ for `rstan`), the digit model ($\text{BF}_{max} = 269 * 10^{31}$ for `rstan`).
-Therefore, the data from Rouder et al.
-(2005) show evidence against the symbolic distance effect.
-
-## Sensitivity Analysis 
-The full model including the side and digit effects performs worst compared to all other models in the Bayes factor model comparison.
-However, we did find non-zero digit effects in the model estimation.
-How is this possible?
-
-A possible explanation is a lack of individual variability in the digit and side effects.
-The full model assumes that participants vary in the digit and side effect, and we specified our expectations about this size of the variability in the prior.
-However, if there is a lack of individual variability in the effects, the full model is punished for the complexity added by the individual variation (illustrated by Rouder et al., 2021).
-To investigate this idea, we can reformulate the prior distributions on the variance parameters to indicate a smaller individual variation or, even more extreme, remove these individual effects from the model entirely.
-Then, we can calculate the posterior model probabilities for all specified models.
-
-Specifically, we will compare models from three scenarios: scenario 1 specifies large individual differences equivalent to Equation \@ref(eq:chosenpriors2var).
-For the models in scenario 2, we adjust the priors on $\sigma^2_{\beta}$ and $\sigma^2_{\delta}$ in the following way:
-
-\begin{equation}
-\begin{aligned}
-\sigma^2_{\beta} &\sim \text{Inverse-Gamma}(3, 0.05), \\   
-\sigma^2_{\delta} &\sim \text{Inverse-Gamma}(3, 0.05). 
-\end{aligned}
-\end{equation}
-The smaller value for the scale parameter of the inverse gamma distribution (i.e., 0.05 instead of 0.5) results in a shift of the peak of the prior towards smaller values for the individual variance.
-This means that the variation of the individual effects is expected to be much smaller than before.
-
-In scenario 3, we remove individual variability from the side, digit, and full model. This results in the following full model, from which the side and digit model follow logically:
-
-\begin{equation}
-Y_{ijk} \sim \text{Normal}(\gamma_{i} + x_j \mu_\beta + u_j \mu_{\delta{7}} + v_j \mu_{\delta{6}} + w_j \mu_{\delta{4}} + z_j \mu_{\delta{3}}, \sigma^2). (\#eq:datalevel2)
-\end{equation}
-Note that the null model is the same in all three scenarios.
-
-To obtain the posterior model probability for each model in every scenario, we first compute the Bayes factor of every model against the null model.
-This includes the trivial Bayes factor of the null model against the null model, resulting in ten Bayes factors in total.
-Next, we calculate the sum of all these Bayes factors and divide each Bayes factor by this sum.
-This results in the posterior model probabilities shown in Table \@ref(tab:postmodprob).^[Note that it would be possible to calculate the posterior model probabilities directly from the marginal likelihoods.
-However, marginal likelihoods are often not computed explicitly in the model fitting process (e.g., when using the bayes_factor function in the `brms` package), so we believe that readers might find it easier to compute posterior model probabilities from Bayes factors.] All posterior model probabilities from the table and the posterior probability of the null model (see table note) sum to one.
-
-\begin{table}[H]
-\caption{Posterior Model Probabilities Normal Models}
-\label{tab:postmodprob}
-\begin{tabularx}{\textwidth}{XXXX}
-\toprule
- & \multicolumn{1}{c}{Scenario 1} & \multicolumn{1}{c}{Scenario 2} & \multicolumn{1}{c}{Scenario 3}\\
-\midrule
-Digit model & $0.87*10^{-126}$ & $0.29*10^{-46}$ & $0.95$\\
-Side model & $0.20*10^{-88}$ & $0.18*10^{-66}$ & $0.77*10^{-53}$ \\
-Full model & $0.43*10^{-158}$ & $0.47*10^{-61}$ & $0.05$\\
-\bottomrule
-\end{tabularx}
-\begin{tablenotes}[para]
-\textit{Note.} Scenario 1: large individual differences.
-Scenario 2: small individual differences.
-Scenario 3: no individual differences.
-The posterior model probability of the null model is the same in every scenario and equals $0.50*10^{-107}$.
-\end{tablenotes}
-\end{table}
-
-The table shows that the digit model without individual digit effects from scenario 3 has the highest posterior model probability, followed by the full model without individual digit and side effects (scenario 3).
-This means that the data are most likely under these models than the other eight models included in the comparison.
-In addition, the models with smaller prior expectations for individual differences (scenario 2), have higher posterior model probabilities than the models in scenario 1.
-Thus, decreasing our expectations of the size of the individual digit and side effects, or completely removing them, results in higher posterior model probabilities.
-Together with Figure \@ref(fig:randomparfigure), showing the small differences in individual effects (also illustrated by Haaf et al., 2019), this result supports the notion that side and digit effects are small and there is a lack in individual variability.
-
-### Log-normal Model
-We can take a similar approach for the log-normal model to compare the hierarchical models on varying complexity.
-We evaluate their performance in two scenarios that correspond to scenario 1 and 3 for the normal model.
-This means that in scenario 1 for the log-normal model, we estimated a null, digit, side and full model based on the prior specified in Equation, specifying large individual differences.
-For scenario 2, we remove individual variability from the side, digit, and full model.
-This resulted in the following posterior model probabilities shown in Table \@ref(tab:postmodproblog).
-
-\begin{table}[H]
-\caption{Posterior Model Probabilities Log-normal Models}
-\label{tab:postmodproblog}
-\begin{tabularx}{\textwidth}{XXXX}
-\toprule
- & \multicolumn{1}{l}{Scenario 1} & \multicolumn{1}{l}{Scenario 3}\\
-\midrule
-Digit model & $26.61 * 10^9$ & $48.55 * 10^{15}$\\
-Side model & $0.25 * 10^{-49}$  & $0.50 * 10^{-52}$ \\
-Full model & $11.02 * 10^7$ &  $96.84 * 10^{14}$\\
-\bottomrule
-\end{tabularx}
-\begin{tablenotes}[para]
-\textit{Note.} Scenario 1: large individual differences.
-Scenario 2: no individual differences.
-The posterior model probability of the null model is the same in every scenario and equals $0.86*10^{-51}$.
-\end{tablenotes}
-\end{table}
-
-The table shows that the digit model without individual digit effects from scenario 2 has the highest posterior model probability.
-This is in line with our findings for the normal model.
-Finally, we can compare the normal model with the highest marginal likelihood to the log-normal with the highest marginal likelihood.
-The normal digit model from scenario 3 to the log-normal model from scenario 2.
-The log marginal likelihood for the normal digit model without individual differences was $6441.04$ while the log marginal likelihood for the log-normal digit model without individual differences was $10637.45$.
-This indicates that the Bayes factor is in favor of the log-normal digit model over the normal digit model.
-
-This example illustrates the benefits of comparing hierarchical models of varying complexity.
-While individual differences are a common phenomenon in psychology, there are instances where general effects are sufficient to describe the data adequately, or where only a random intercept, but no random slopes are necessary.
-As shown in Tables \@ref(tab:postmodprob) and \@ref(tab:postmodproblog), Bayesian model comparisons can help researchers to identify the model that best describes the data.
-
-# Discussion
-
-<!-- To conclude this tutorial on Bayesian hierarchical modeling in psychology, we will highlight the most important recommendations to keep in mind when applying a Bayesian hierarchical model.   -->
-
-In this tutorial, we described model and prior specification, estimation, and model comparison for hierarchical models in the Bayesian framework.
-In the following, we want to present several key recommendations.
-
-As prior distributions influence model estimation and model comparison, it is important to choose suitable priors.
-We believe that simulating data based on the chosen priors is an intuitive tool to find reasonable prior settings that take the existing knowledge in the specific model application context into account.
-We showed that it can be problematic to use default priors without checking their suitability.
-Therefore, we recommend performing prior predictive checks with the priors that will be applied before running the analysis.
-
-We illustrated different methods to obtain Bayes factors.
-We want to emphasize the importance of checking the variability of the Bayes factor estimate, the Bayes factor sensitivity to prior distributions, and to ensure that the Bayes factor estimate is based on a sufficient number of posterior samples.
-
-This tutorial showed the application of two R packages: `rstan` and `brms`.
-Which package to use depends mostly on the complexity of the model and your technical background.
-For the non-technical user, we recommend using `brms` because of its intuitive model specification.
-However, the default prior setup can be unsuitable for a specific research context and (virtually) always needs to be adapted.
-Additionally, users are limited to predefined parametrizations.
-`rstan` offers the most flexibility in model and prior specification.
-This flexibility comes with a cost: if the prior or the model is not well defined, results can be misleading.
-
-## Further Recommendations
-
-With this tutorial, we attempted to cover the most relevant topics that researchers encounter when starting with Bayesian hierarchical modeling.
-However, there are still many remaining issues and recommendations that we could not discuss.
-Here, we would like to point to further references in the literature if researchers want to continue mastering Bayesian hierarchical modeling.
-
-First, we kept the discussion of the software packages and code fairly brief.
-We provide further code examples for the digit classification task in the [Online Supplement](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment).
-For different models, it may be useful to study additional tutorials and books.
-For `rstan`, we recommend the `Stan` manual [@stan2018a; @stan2018b].
-For `brms` there are several useful tutorials [@R-brms_a; @R-brms_b; @burkner2019ordinal].
-There are also several resources for Bayesian estimation with other samplers [@lee2013bayesian; @rouder2013hierarchical].
-Two other useful tutorials for Bayesian modeling are Schad, Betancourt, and
-Vasishth [-@schad2019toward] and Schad, Nicenboim, Bürkner, Betancourt, and Vasishth [-@schad2021workflow].
-\Revision{There are also other R software packages for Bayesian multilevel modeling besides `brms` and `rstan`, such as `BayesFactor`[@R-BayesFactor] and `rstanarm`[@rstanarm2020], that we did not illustrate in this tutorial.
-However, most of these other R software packages are a bit more limited, for instance, in model specification and prior settings.}
-
-Second, we did not discuss study planning.
-When planning an experiment, researchers may want to determine the number of participants required to gain conclusive results.
-Bayesian analysis has the advantage that an optional stopping paradigm may be employed [@rouder2014optional; @schonbrodt2017sequential].
-Schönbrodt and Wagenmakers [-@schonbrodt2018bayes] and Stefan and colleagues [-@stefan2019tutorial] introduced Bayes factor design analysis for fixed and sequential designs.
-However, these approaches to study planning have not yet been fully extended to a hierarchical setup [but see recent efforts by Vasishth et al., -@vasishth2021sample].
-The extension to nested designs is not entirely trivial because, in addition to planning the number of participants, we also need to consider the number of trials [@rouder2018power].
-While we think there remains some open questions on the topic, the tutorial by Stefan and colleagues [-@stefan2019tutorial] is a good place to start for applied researchers.
-
-## Potential Technical Issues
-
-One key advantage of Bayesian over frequentist hierarchical modeling is that complex modeling leads to fewer convergence issues in the Bayesian framework.
-This advantage, however, does not imply that there are no technical issues in the Bayesian framework.
-One possible technical issue is with the choice of priors when using `Stan` for estimation.
-The `Stan` algorithm is most efficient with specific prior distributions.
-For example, it may be beneficial to use a truncated $t$-distribution on standard deviation parameters instead of an inverse gamma distribution on variances [@gelman2006prior].
-Using priors not recommended by the `Stan` team may lead to more issues with convergence.
-In this tutorial, we highlight that choosing priors based on their predictions on data is beneficial.
-We still think this is the best path for substantive researchers.
-If one wants more flexibility in choices of priors or has continuous issues with convergence with their chosen prior distributions, then it might be worth considering using `JAGS` instead of `Stan`.
-`JAGS` [@Plummer03jagsa] has a similar functionality and syntax as `Stan`.
-While `Stan` tends to be more efficient in the sampling phase, especially in case of correlated parameters [@hecht2021comparing], the `JAGS` algorithm could be more flexible and less affected by the choice of priors.
-
-Another technical issue may arise with the estimation of Bayes factors.
-Here, we highlighted two approaches, the Savage-Dickey density ratio [@wagenmakers2010bayesian] and bridge sampling [@gronau2017tutorial].
-We already highlighted the drawbacks of the SD density ratio.
-Yet, bridge sampling may also fail in some cases.
-Failures of the bridge sampling algorithm are identified by unstable Bayes factors across repeated sampling runs.
-There are improvements to the algorithm that are specifically developed for hierarchical models [@gronau2019simple].
-Yet, it remains crucial to assess the stability of Bayes factor estimates.
-
-## Living with Uncertainty
-
-Let us reconsider our scenario from the introduction.
-The graduate student may have read our tutorial and went back to their supervisor to propose a Bayesian hierarchical analysis of their data.
-Compared to the initially considered ANOVA analysis, there are many small decisions to be made for this more extensive analysis.
-These small decisions tend to make researchers uncomfortable.
-What prior is the right one?
-Are there enough posterior samples?
-
-In Bayesian modeling, we have to learn living with these uncertainties.
-In fact, we recommend embracing them as these small decisions.
-First, we at the very least are making these decisions deliberately.
-Second, we can check their impact on the results.
-Throughout the tutorial we highlighted several potential robustness checks, from prior predictions over convergence assessment, to sensitivity analysis for Bayes factors.
-These checks may help researchers to understand how robust their results are and when they break.
-Therefore, we hope the checks will support switching from simple procedures to Bayesian hierarchical modeling.
-Because ultimately, the results from such an analysis are much richer.
-
-In summary, by offering a comparison of software packages, guidance on prior selection and default priors, and assessment of the Bayes factor as model comparison method, we hope that Bayesian hierarchical models will become available to a wider psychology public.
-
-\newpage
-
-# Declarations
-
-## Funding
-
-This work was supported in part by an NWO Research Talent grant to A.M.S. (406.18.556) and a Veni grant from the NWO (VI.Veni.201G.019) to J.M.H.
-
-## Conflicts of interest
-
-The authors have no conflicts of interest to declare.
-
-## Availability of data and materials
-
-We do not have ownership of the data used in this study, but the the data set is publicly available on [GitHub (https://github.com/PerceptionCognitionLab/data0/blob/master/lexDec-dist5/)](https://github.com/PerceptionCognitionLab/data0/blob/master/lexDec-dist5/).
-The `.stan` models, results of the model fit, and model comparison are available on [our GitHub repository: https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment).
-
-## Code availability
-
-The R code used for the analyses and figures in this manuscript, and `R` tutorials on BHMs are available on [our GitHub repository: https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment).
-
-## Authors’ contributions
-
-M.V.
-participated in the design and coordination of this study, performed statistical analyses, and drafted the paper; A.M.S.
-participated in the design and coordination of this study, helped with the statistical analyses, interpretation of the results and writing the paper; J.M.H.
-(principal investigator) conceived the study, participated in the design and coordination of the study, helped with the statistical analyses, interpretation of results and writing the paper.
-All authors read and approved the final paper.
-
-## Acknowledgements
-
-We would like to express our gratitude to the R developers [Version 3.6.1; @R-base] and the following R package developers:
-
-- Document layout: *rmarkdown* [Version 2.9; @R-rmarkdown], *papaja* [Version 0.1.0.9997; @R-papaja], *knitr* [Version 1.33; @R-knitr], *kableExtra* [Version 1.3.4; @R-kableExtra]
-- Data analysis and cleaning: *LaplacesDemon* [Version 16.1.6; @R-LaplacesDemon], *plyr* [Version 1.8.6; @R-dplyr; @R-plyr], *dplyr* [Version 1.0.7; @R-dplyr], *readr* [Version 2.0.0; @R-readr], *crch* [Version 1.0.4; @R-crch_a; @R-crch_b; @R-crch_c; @R-crch_d], *foreign* [Version 0.8.81; @R-foreign], *Matrix* [Version 1.3.3; @R-Matrix], *tidyr* [Version 1.1.3; @R-tidyr], *truncnorm* [Version 1.0.8; @R-truncnorm]
-- Model fit and comparison: *rstan* [Version 2.21.3; @R-rstan], *brms* [Version 2.16.3; @R-brms_a; @R-brms_b], *BayesFactor* [Version 0.9.12.4.2; @R-BayesFactor], *bridgesampling* [Version 1.1.2; @R-bridgesampling]
-- Visualization: *coda* [Version 0.19.4; @R-coda], *lattice* [Version 0.20.44; @R-lattice], *ggplot2* [Version 3.3.5; @R-ggplot2], *Rmisc* [Version 1.5; @R-Rmisc], *devtools* [Version 2.4.2; @R-devtools], *gghalves* [Version 0.1.1.9000; @R-gghalves], *bayesplot* [Version 1.8.1; @R-bayesplot], *gridExtra* [Version 2.3; @R-gridExtra], *tibble* [Version 3.1.6; @R-tibble], *ggmcmc* [Version 1.5.1.1; @R-ggmcmc], *cowplot* [Version 1.1.1; @R-cowplot], *ggridges* [Version 0.5.3; @R-ggridges], *circlize* [Version 0.4.13; @R-circlize], *StanHeaders* [Version 2.21.0.7; @R-StanHeaders], *unikn* [Version 0.4.0; @R-unikn], *ggbeeswarm* [Version 0.6.0; @R-ggbeeswarm]
-- Other: *Rcpp* [Version 1.0.7; @R-Rcpp_a; @R-Rcpp_b], *usethis* [Version 2.0.1; @R-usethis]
-
-\newpage
-
-# References
-```{r create_r-references}
-r_refs(file = "r-references.bib")
-# cite_r("r-references.bib")  # to obtain package versions 
-```
-
-\begingroup
-\setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.2in}
-
-<div id = "refs">
-
-</div>
-\endgroup
-
-\newpage
+Schad, D. J., Betancourt, M., & Vasishth, S. (2021). Toward a principled Bayesian workflow in cognitive science. Psychological methods, 26(1), 103.

### 2023-04-06 manuscript.Rmd
Updates revision

@@
 For example, participants in a digit classification task repeatedly respond to target digits inducing a correlation between the responses to the target digits, such as between target digit 3 and target digit 4.
-Therefore, the observations are not independent as we assume if we perform analysis on aggregated data.
-As a result, the type I error rate of a test of the effect increases: there is a higher probability of concluding that there is an effect if there is no effect (Rouder & Lu, 2005).
+\Revision{Therefore, analyses on aggregated data do not appropriately account for the interdependence between observations [@van2021bayes; @van2023bayes].} As a result, the type I error rate of a test of the effect increases: there is a higher probability of concluding that there is an effect if there is no effect (Rouder & Lu, 2005).
 
@@
 The model is fitted using the packages `rstan` and `brms`.
-For `rstan`, four chains are run with 4000 iterations per chain, of which 1000 serve as warm-up.
+\Revision{We used different settings for `rstan` and `brms` due to the difference between the models for these packages.
+Although we tried to make the models as similar as possible, the models are not identical.
+This is partly due to difference in prior specification (i.e., specifying prior on the variances versus standard deviations).} For `rstan`, four chains are run with 4000 iterations per chain, of which 1000 serve as warm-up.
 This results in a total of 12000 samples from the joint posterior distribution.
@@
 Here, this was not case, and all other diagnostics were good, so we continue with the interpretation of the estimates.
+\Revision{The difference between the model diagnostics of `rstan` and `brms` can be partly explained by the difference in model parametrization, such as the difference in the prior specification (i.e., specifying priors on variances versus priors on standard deviations).
+Although we tried to make the models as similar as possible, they are not identical.}
 
@@
 For the digit effects, positive parameter estimates indicate that the response is slower for the non-baseline digits, that is, digits that are closer to 5.
-The parameters $\mu_{\delta_{6}}$ and $\mu_{\delta_{4}}$ are distributed around higher values and their credible intervals do not contain zero.
+The parameters $\mu_{\delta_{6}}$ and $\mu_{\delta_{4}}$ are distributed around higher values.
 Note that this result is consistent with the symbolic distance hypothesis that the digits 4 and 6 have the highest response times because they are closest to the 5, the comparison value.
-The 95% credible intervals of $\mu_{\delta_{7}}$ and $\mu_{\delta_{3}}$ contain zero.
-\Revision{Therefore, it is plausible that the response times do not differ substantially from the baseline condition.- is this one necessary to adjust?} The posterior distribution of the side effect $\mu_{\beta}$ is centered around zero with a posterior mean of `r printnum(fitrstanhiertable3["mu2", "Mean"], digits = 3)` and a 95% credible interval ranging from `r printnum(fitrstanhiertable3["mu2", "Lower Bound"], digits = 3)` to `r printnum(fitrstanhiertable3["mu2", "Upper Bound"], digits = 3)`.
-\Revision{This indicates a small to non-existent side effect.
--change}
+The posterior distribution of the side effect $\mu_{\beta}$ is centered around zero with a posterior mean of `r printnum(fitrstanhiertable3["mu2", "Mean"], digits = 3)` and a 95% credible interval ranging from `r printnum(fitrstanhiertable3["mu2", "Lower Bound"], digits = 3)` to `r printnum(fitrstanhiertable3["mu2", "Upper Bound"], digits = 3)`.
+\Revision{Note that the credible interval overlaps with zero.
+Commonly, this is misinterpreted as an absence of an effect.
+However, in combination with the mean estimate, we can conclude that the effect is small.}
 
@@
 There is also considerable individual variation in the digit effects ($\delta$'s).
-However, most of the intervals contain zero which implies that the effects might be small or non-existent.
-The variability is most pronounced for the effect parameters of the digits close to 5.
+\Revision{However, most of the intervals contain zero which implies that the effects might be small or non-existent.} The variability is most pronounced for the effect parameters of the digits close to 5.
 For these digits, there are also more intervals that do not contain zero, indicating that an effect might be present on the subject level.
@@
 The violin plots on the right side show the variance of the individual parameter estimates.
-The size of effects is not very constistent within a person.", fig.align = "left", fig.height = 11, fig.width = 1, out.width="75%"}
+", fig.align = "left", fig.height = 11, fig.width = 1, out.width="75%"}
 knitr::include_graphics("Images/Fig11_modelestimatefigure2_onlyrstan_v2.png")
@@
 Calculating the mean estimate of $\mu_{\delta_{3}}$ would be similar, but more complicated as the mean of a log-normal distribution also depends on its variance.
+\Revision{This approach can also be used iteratively.
+This means that the approach above is applied to every iteration in the chain.
+This results in a posterior distribution of the effect quantifying the posterior uncertainty.}
 
@@
 The side effect is close to zero, with a posterior mean of `r printnum(fitrstanhiertable3log["mu2", "Mean"], digits = 3)` and a 95% credible interval from `r printnum(fitrstanhiertable3log["mu2", "Lower Bound"], digits = 3)` to `r printnum(fitrstanhiertable3log["mu2", "Upper Bound"], digits = 3)`.
-Thus, the general effect of side seems small or negligible.
+\Revision{This indicates that the general effect of side is small.}
 
@@
 It is important to note that this difference is not due to the bridge sampling function, but a result of a difference in parametrization.
-However, for both packages, the conclusion is the same: the data are least likely to have occurred under the full model.
+\Revision{The models for `rstan` and `brms` are similar but not identical, partly due to the difference in prior specification (i.e., variances versus standard deviations).} However, for both packages, the conclusion is the same: the data are least likely to have occurred under the full model.
 In case the estimates are unstable, it is recommended to further increase the number of posterior samples.
@@
 To obtain the posterior model probability for each model in every scenario, we first compute the Bayes factor of every model against the null model.
-This includes the trivial Bayes factor of the null model against the null model, resulting in ten Bayes factors in total.
-Next, we calculate the sum of all these Bayes factors and divide each Bayes factor by this sum.
+This includes the trivial Bayes factor of the null model against the null model.
+\Revision{This setup results in 10 models, and, therefore, 10 Bayes factors.
+We assume that all models are equally likely, resulting in a prior model probability for each model of 1/10.} Next, we calculate the sum of all these Bayes factors and divide each Bayes factor by this sum.
 This results in the posterior model probabilities shown in Table \@ref(tab:postmodprob).^[Note that it would be possible to calculate the posterior model probabilities directly from the marginal likelihoods.
@@
 In addition, the models with smaller prior expectations for individual differences (scenario 2), have higher posterior model probabilities than the models in scenario 1.
-Thus, decreasing our expectations of the size of the individual digit and side effects, or completely removing them, results in higher posterior model probabilities.
-Together with Figure \@ref(fig:randomparfigure), showing the small differences in individual effects (also illustrated by Haaf et al., 2019), this result supports the notion that side and digit effects are small and there is a lack in individual variability.
+\Revision{Thus, decreasing the variability of the size of the individual digit and side effects (scenario 2), or completely removing the variance (scenario 3), results in higher posterior model probabilities.} Together with Figure \@ref(fig:randomparfigure), showing the small differences in individual effects (also illustrated by Haaf et al., 2019), this result supports the notion that side and digit effects are small and there is a lack in individual variability.
 
@@
 
-## Potential Technical Issues
+\Revision{## Technical Possibilities are Endless}
 
@@
 We already highlighted the drawbacks of the SD density ratio.
-Yet, bridge sampling may also fail in some cases.
+\Revision{Bayes factor estimates can also be biased.
+However, bridge sampling has shown to be relatively unbiased [@sarafoglou2021evaluating].} Yet, bridge sampling may also fail in some cases.
 Failures of the bridge sampling algorithm are identified by unstable Bayes factors across repeated sampling runs.

### 2023-05-07 manuscript.Rmd
Updates for revision

@@
 
-Our tutorial is directed at researchers who have a basic understanding of Bayesian inference and parameter estimation and are planning to perform more complex Bayesian analyses.
+Our tutorial is directed at researchers who have a basic \Revision{conceptual} understanding of Bayesian inference and parameter estimation and are planning to perform more complex Bayesian analyses \Revision{of psychological phenomena.
+We hope to achieve a balance of detail and overview---while we want to provide guidance for a full Bayesian workflow, similar to Gelman et al.
+[-@gelman2020] and Schad et al.[-@schad2019toward; -@schad2021workflow], we don't shy away from in-depth explanations for the most curious scientists.
+To ensure that a broad audience is able to follow the tutorial, we limit and carefully explain equations.
+However, rudimentary knowledge of calculus is advantageous to fully understand Bayesian analysis.}
 
@@
 
-Another important issue is that wide priors are often improper, that is, the probability density does not integrate to one [@hobert1996effect].
-\Revision{The integral is used as an approximation of the sum for continuous parameters to calculate the probability density.} An example of an improper prior is a uniform prior distribution ranging from minus infinity to infinity.
-In case of improper priors, an estimation of the general and individual effects cannot be provided.
-However, even proper priors can result in improper posteriors [@hobert1996effect].
+Another important issue is that wide priors are often improper probability distributions [@hobert1996effect].
+\Revision{A prior is proper if two conditions hold: 1.
+All values of the probability distribution are positive definite, and 2.
+the probability distribution sums to one for discrete data or integrates to one for continuous data.
+An example of a proper and improper probability distribution are provided in Figure \@ref(fig:properfigure).
+The figure shows two hypothetical priors for parameter $\theta$.
+The first condition can be assessed by checking the y-values---both priors have only positive definite values as function output (0.1, 0.2, and 0.4).
+To assess the second condition, we need to sum up the function values.
+Panel A shows a proper distribution where the values sum to 1; panel B shows an improper distribution where the values sum to 1.2.}
+
+```{r properfigure, echo = F, fig.width=7, fig.height=3.5, fig.align='center', fig.cap="Two hypothetical priors for a discrete parameter $\\theta$.
+A.
+A proper probability distribution that sums to one.
+B.
+An improper probability distribution that sums to 1.2."}
+layout(matrix(1:2, nrow = 1))
+par(mgp = c(2, .7, 0))
+x <- 1:5
+y <- c(12, 6, 6, 3, 3)
+y <- y / sum(y)
+
+plot(x, y, pch = 19, col = "gray40", axes = F, ylim = c(0, 1), ylab = "Probability", xlab = expression(theta), main = "A. Proper", xlim = c(.5, 5.5))
+axis(1, x)
+axis(2, c(0, unique(y), 1), las = 2)
+for(i in x) lines(rep(x[i], 2), c(0, y[i]), col = "gray40", lwd = 2)
+
+par(mgp = c(2, .7, 0))
+y[3] <- .4
+
+plot(x, y, pch = 19, col = "gray40", axes = F, ylim = c(0, 1), xlim = c(.5, 5.5), ylab = "Probability", xlab = expression(theta), main = "B. Improper")
+axis(1, x)
+axis(2, c(0, unique(y), 1), las = 2)
+for(i in x) lines(rep(x[i], 2), c(0, y[i]), col = "gray40", lwd = 2)
+```
+
+A \Revision{more common} example of an improper prior is a uniform prior distribution ranging from minus infinity to infinity.
+In case of improper priors, \Revision{it is often impossible to obtain correct estimation of general and individual effects in hierarchical models because the resulting posteriors will again be improper probability distributions.} However, even proper priors can result in improper posteriors [@hobert1996effect].
 Although estimation should not be possible in this situation, statistical software might still provide results without notifying the user that the posterior distribution does not exist.
-If users are not aware of this problem, this can result in misleading conclusions.
+If users are not aware of the impropriety problem, this can result in misleading conclusions.
 Therefore, it is important to check if *(a)* the priors are proper and *(b)* the priors result in proper posteriors.
-There is no fail-safe method that guarantees proper posteriors for hierarchical models.
+\Revision{Unfortunately,} there is no fail-safe method that guarantees proper posteriors for hierarchical models.
 The best approach is to choose well-reasoned informative priors for the model, for instance, by using the proposed four-step procedure below.
+\Revision{Priors that are proper and not too diffuse will in most cases lead to proper posteriors.}
 
@@
 \end{equation}
-The prior distributions `brms` specifies on the standard deviations can be transformed to prior distributions on variances.
-Prior distributions on variances have a smaller scale than priors on standard deviation.
+The prior distributions `brms` specifies on the standard deviations can be transformed to prior distributions on variances ($var$).
+Prior distributions on variances have a smaller scale than priors on standard deviation ($sd$).
 Note, however, that the transformation of the prior distributions is not as simple as the transformation of the parameter itself.
-In our [Online Supplement](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/tree/main/R%20code/Prior%20transformation), we provide R-code with a numeric solution to the transformation.
+\Revision{If we formulate a prior on standard deviations, there is always an implied prior on the variances as well.
+If we think of the distribution as $f(x)$, $x$ can either be defined as $\sqrt{(var)}$, if we formulated the prior on the standard deviation, or $x$ can be defined as ${sd}^2$ if we formulated the prior on the variance.
+From this, we can derive the implied distribution $f(var)$ or $f(sd)$.
+Notably, this is not the same as calculating $\sqrt{(f(x))}$ or $f(x)^2$.
+To transform the priors on standard deviations to variances, we try to find the parameters that match the prior distribution to the implied distribution.} In our [Online Supplement](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/tree/main/R%20code/Prior%20transformation), we provide R-code with a numeric solution to the transformation.
 For the `brms` priors, we obtain the following implied prior distributions on variances:
@@
 
-\Revision{## Technical Possibilities are Endless}
+## \Revision{Technical Possibilities are Endless}
 

### 2023-05-07 manuscript.tex
Updates for revision

@@
 
-Our tutorial is directed at researchers who have a basic understanding of Bayesian inference and parameter estimation and are planning to perform more complex Bayesian analyses.
+Our tutorial is directed at researchers who have a basic \textcolor{blue}{conceptual} understanding of Bayesian inference and parameter estimation and are planning to perform more complex Bayesian analyses \textcolor{blue}{of psychological phenomena.
+We hope to achieve a balance of detail and overview---while we want to provide guidance for a full Bayesian workflow, similar to Gelman et al.
+[-@gelman2020] and Schad et al.[-@schad2019toward; -@schad2021workflow], we don't shy away from in-depth explanations for the most curious scientists.
+To ensure that a broad audience is able to follow the tutorial, we limit and carefully explain equations.
+However, rudimentary knowledge of calculus is advantageous to fully understand Bayesian analysis.}
 
@@
 For example, participants in a digit classification task repeatedly respond to target digits inducing a correlation between the responses to the target digits, such as between target digit 3 and target digit 4.
-Therefore, the observations are not independent as we assume if we perform analysis on aggregated data.
-As a result, the type I error rate of a test of the effect increases: there is a higher probability of concluding that there is an effect if there is no effect (Rouder \& Lu, 2005).
+\textcolor{blue}{Therefore, analyses on aggregated data do not appropriately account for the interdependence between observations [@van2021bayes; @van2023bayes].} As a result, the type I error rate of a test of the effect increases: there is a higher probability of concluding that there is an effect if there is no effect (Rouder \& Lu, 2005).
 
@@
 Note that the only change in the model formulation is the change of the likelihood function from a normal distribution to a log-normal distribution.
-\textcolor{blue}{The log-normal distribution only assigns probability to positive values, and is, just as the RT distributions, right-skewed.
+\textcolor{blue}{The likelihood function provides the likelihood of parameter values in the form of probability densities.
+The log-normal distribution only assigns probability to positive data values, and is, just as the RT distributions, right-skewed.
 Therefore, the log-normal distribution is a better match for the RT data.} The location parameter of the log-normal distribution is described through the same combination of parameters as the mean of the normal model.
@@
 
-Another important issue is that wide priors are often improper, that is, the probability density does not integrate to one (Hobert \& Casella, 1996).
-An example of an improper prior is a uniform prior distribution ranging from minus infinity to infinity.
-In case of improper priors, an estimation of the general and individual effects cannot be provided.
-However, even proper priors can result in improper posteriors (Hobert \& Casella, 1996).
+Another important issue is that wide priors are often improper probability distributions (Hobert \& Casella, 1996).
+\textcolor{blue}{A prior is proper if two conditions hold: 1.
+All values of the probability distribution are positive definite, and 2.
+the probability distribution sums to one for discrete data or integrates to one for continuous data.
+An example of a proper and improper probability distribution are provided in Figure \\ref{fig:properfigure}.
+The figure shows two hypothetical priors for parameter $\theta$.
+The first condition can be assessed by checking the y-values---both priors have only positive definite values as function output (0.1, 0.2, and 0.4).
+To assess the second condition, we need to sum up the function values.
+Panel A shows a proper distribution where the values sum to 1; panel B shows an improper distribution where the values sum to 1.2.}
+
+\begin{figure}[H]
+
+{\centering \includegraphics{manuscript_files/figure-latex/properfigure-1}
+
+}
+
+\caption{Two hypothetical priors for a discrete parameter $\theta$.
+A.
+A proper probability distribution that sums to one.
+B.
+An improper probability distribution that sums to 1.2.}\label{fig:properfigure}
+\end{figure}
+
+A \textcolor{blue}{more common} example of an improper prior is a uniform prior distribution ranging from minus infinity to infinity.
+In case of improper priors, \textcolor{blue}{it is often impossible to obtain correct estimation of general and individual effects in hierarchical models because the resulting posteriors will again be improper probability distributions.} However, even proper priors can result in improper posteriors (Hobert \& Casella, 1996).
 Although estimation should not be possible in this situation, statistical software might still provide results without notifying the user that the posterior distribution does not exist.
-If users are not aware of this problem, this can result in misleading conclusions.
+If users are not aware of the impropriety problem, this can result in misleading conclusions.
 Therefore, it is important to check if \emph{(a)} the priors are proper and \emph{(b)} the priors result in proper posteriors.
-There is no fail-safe method that guarantees proper posteriors for hierarchical models.
+\textcolor{blue}{Unfortunately,} there is no fail-safe method that guarantees proper posteriors for hierarchical models.
 The best approach is to choose well-reasoned informative priors for the model, for instance, by using the proposed four-step procedure below.
+\textcolor{blue}{Priors that are proper and not too diffuse will in most cases lead to proper posteriors.}
 
@@
 
-A key advantage of model specification with priors is that we can obtain preditions on what the data would look like that are generated from the model.
+A key advantage of model specification with priors is that we can obtain predictions on what the data would look like that are generated from the model \textcolor{blue}{[@etz2018bayesian; @lee2018determining]}.
 These prior predictions allow us to make the implications of our specified priors more concrete.
@@
 \end{equation}
-The prior distributions \texttt{brms} specifies on the standard deviations can be transformed to prior distributions on variances.
-Prior distributions on variances have a smaller scale than priors on standard deviation.
+The prior distributions \texttt{brms} specifies on the standard deviations can be transformed to prior distributions on variances (\(var\)).
+Prior distributions on variances have a smaller scale than priors on standard deviation (\(sd\)).
 Note, however, that the transformation of the prior distributions is not as simple as the transformation of the parameter itself.
-In our \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/tree/main/R\%20code/Prior\%20transformation}{Online Supplement}, we provide R-code with a numeric solution to the transformation.
+\textcolor{blue}{If we formulate a prior on standard deviations, there is always an implied prior on the variances as well.
+If we think of the distribution as $f(x)$, $x$ can either be defined as $\sqrt{(var)}$, if we formulated the prior on the standard deviation, or $x$ can be defined as $(sd^2)$ if we formulated the prior on the variance.
+From this, we can derive the implied distribution $f(var)$ or $f(sd)$.
+Notably, this is not the same as calculating $\sqrt{(f(x))}$ or $f(x)^2$.
+To transform the priors on standard deviations to variances, we try to find the parameters that make the prior distribution match the implied distribution.} In our \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/tree/main/R\%20code/Prior\%20transformation}{Online Supplement}, we provide R-code with a numeric solution to the transformation.
 For the \texttt{brms} priors, we obtain the following implied prior distributions on variances:
@@
 Finally, we would like to use a more common conjugate prior for the variances to ensure obtaining a proper posterior distribution.\footnote{Conjugate priors result in posterior distributions that belong to the same distribution family as the prior distribution (Fink, 1997).
-For a normal likelihood, these are the normal distribution for the mean and the inverse gamma distribution for the variance.} Therefore, we choose an inverse gamma distribution.
+For a normal likelihood, these are the normal distribution for the mean and the inverse gamma distribution for the variance.}\textcolor{blue}{However, in `rstan`, non-conjugate priors are commonly applied [e.g., @gelman2006prior]} Therefore, we choose an inverse gamma distribution.
 We can also be more specific about the expected RTs.
@@
 The model is fitted using the packages \texttt{rstan} and \texttt{brms}.
-For \texttt{rstan}, four chains are run with 4000 iterations per chain, of which 1000 serve as warm-up.
+\textcolor{blue}{We used different settings for `rstan` and `brms` due to the difference between the models for these packages.
+Although we tried to make the models as similar as possible, the models are not identical.
+This is partly due to difference in prior specification (i.e., specifying prior on the variances versus standard deviations).} For \texttt{rstan}, four chains are run with 4000 iterations per chain, of which 1000 serve as warm-up.
 This results in a total of 12000 samples from the joint posterior distribution.
@@
 Here, this was not case, and all other diagnostics were good, so we continue with the interpretation of the estimates.
+\textcolor{blue}{The difference between the model diagnostics of `rstan` and `brms` can be partly explained by the difference in model parametrization, such as the difference in the prior specification (i.e., specifying priors on variances versus priors on standard deviations).
+Although we tried to make the models as similar as possible, they are not identical.}
 
@@
 The credible interval displayed in the figure is the Bayesian version of a confidence interval and provides a measure of uncertainty about the parameter value (Wagenmakers et al., 2018).
+\textcolor{blue}{In contrast to the confidence interval, the credible interval provides the probability of a range of values for an effect.
+However, considering the prior and under the assumption that there is an effect, the credible interval does not provide evidence to conclude whether there is an effect [Add Wagenmakers - principle of predictable irrelevance].
+}
 
@@
 For the digit effects, positive parameter estimates indicate that the response is slower for the non-baseline digits, that is, digits that are closer to 5.
-The parameters \(\mu_{\delta_{6}}\) and \(\mu_{\delta_{4}}\) are distributed around higher values and their credible intervals do not contain zero.
+The parameters \(\mu_{\delta_{6}}\) and \(\mu_{\delta_{4}}\) are distributed around higher values.
 Note that this result is consistent with the symbolic distance hypothesis that the digits 4 and 6 have the highest response times because they are closest to the 5, the comparison value.
-The 95\% credible intervals of \(\mu_{\delta_{7}}\) and \(\mu_{\delta_{3}}\) contain zero.
-Therefore, it is plausible that the response times do not differ substantially from the baseline condition.
 The posterior distribution of the side effect \(\mu_{\beta}\) is centered around zero with a posterior mean of -0.007 and a 95\% credible interval ranging from -0.046 to 0.031.
-This indicates a small to non-existent side effect.
+\textcolor{blue}{Note that the credible interval overlaps with zero.
+Commonly, this is misinterpreted as an absence of an effect.
+However, in combination with the mean estimate, we can conclude that the effect is small.}
 
@@
 There is also considerable individual variation in the digit effects (\(\delta\)'s).
-However, most of the intervals contain zero which implies that the effects might be small or non-existent.
-The variability is most pronounced for the effect parameters of the digits close to 5.
+\textcolor{blue}{However, most of the intervals contain zero which implies that the effects might be small or non-existent.} The variability is most pronounced for the effect parameters of the digits close to 5.
 For these digits, there are also more intervals that do not contain zero, indicating that an effect might be present on the subject level.
@@
 
-The individual variation can also be visualized in slightly different way, while focusing on the relation between the individual estimates.
-Figure \ref{fig:modelestimatesfigure3} shows how the relationship between individuals' digit effects.\footnote{The design and code for this figure are by Langen (2020).} The posterior means of individual effects are represented by the dots.
+\textcolor{blue}{The individual variation can also be visualized in a different way with focusing on the relation between the individual estimates.} Figure \ref{fig:modelestimatesfigure3} shows how the relationship between individuals' digit effects.\footnote{The design and code for this figure are by Langen (2020).} The posterior means of individual effects are represented by the dots.
 The right side of each plot represents the distribution of the individual estimates.
@@
 The violin plots on the right side show the variance of the individual parameter estimates.
-The size of effects is not very constistent within a person.}\label{fig:modelestimatesfigure3}
+}\label{fig:modelestimatesfigure3}
 \end{figure}
@@
 Calculating the mean estimate of \(\mu_{\delta_{3}}\) would be similar, but more complicated as the mean of a log-normal distribution also depends on its variance.
+\textcolor{blue}{This approach can also be used iteratively.
+This means that the approach above is applied to every iteration in the chain.
+This results in a posterior distribution of the effect quantifying the posterior uncertainty.}
 
@@
 The side effect is close to zero, with a posterior mean of -0.008 and a 95\% credible interval from -0.022 to 0.007.
-Thus, the general effect of side seems small or negligible.
+\textcolor{blue}{This indicates that the general effect of side is small.}
 
@@
 It is important to note that this difference is not due to the bridge sampling function, but a result of a difference in parametrization.
-However, for both packages, the conclusion is the same: the data are least likely to have occurred under the full model.
+\textcolor{blue}{The models for `rstan` and `brms` are similar but not identical, partly due to the difference in prior specification (i.e., variances versus standard deviations).} However, for both packages, the conclusion is the same: the data are least likely to have occurred under the full model.
 In case the estimates are unstable, it is recommended to further increase the number of posterior samples.
@@
 To obtain the posterior model probability for each model in every scenario, we first compute the Bayes factor of every model against the null model.
-This includes the trivial Bayes factor of the null model against the null model, resulting in ten Bayes factors in total.
-Next, we calculate the sum of all these Bayes factors and divide each Bayes factor by this sum.
+This includes the trivial Bayes factor of the null model against the null model.
+\textcolor{blue}{This setup results in 10 models, and, therefore, 10 Bayes factors.
+We assume that all models are equally likely, resulting in a prior model probability for each model of 1/10.} Next, we calculate the sum of all these Bayes factors and divide each Bayes factor by this sum.
 This results in the posterior model probabilities shown in Table \ref{tab:postmodprob}.\footnote{Note that it would be possible to calculate the posterior model probabilities directly from the marginal likelihoods.
@@
 In addition, the models with smaller prior expectations for individual differences (scenario 2), have higher posterior model probabilities than the models in scenario 1.
-Thus, decreasing our expectations of the size of the individual digit and side effects, or completely removing them, results in higher posterior model probabilities.
-Together with Figure \ref{fig:randomparfigure}, showing the small differences in individual effects (also illustrated by Haaf et al., 2019), this result supports the notion that side and digit effects are small and there is a lack in individual variability.
+\textcolor{blue}{Thus, decreasing the variability of the size of the individual digit and side effects (scenario 2), or completely removing the variance (scenario 3), results in higher posterior model probabilities.} Together with Figure \ref{fig:randomparfigure}, showing the small differences in individual effects (also illustrated by Haaf et al., 2019), this result supports the notion that side and digit effects are small and there is a lack in individual variability.
 
@@
 Two other useful tutorials for Bayesian modeling are Schad, Betancourt, and
-Vasishth (2019) and Schad, Nicenboim, Bürkner, Betancourt, and Vasishth (2021).
+Vasishth (\textbf{schad2019toward?}) and Schad, Nicenboim, Bürkner, Betancourt, and Vasishth (2021).
 \textcolor{blue}{There are also other R software packages for Bayesian multilevel modeling besides `brms` and `rstan`, such as `BayesFactor`[@R-BayesFactor] and `rstanarm`[@rstanarm2020], that we did not illustrate in this tutorial.
@@
 
-\hypertarget{potential-technical-issues}{%
-\subsection{Potential Technical Issues}\label{potential-technical-issues}}
+\hypertarget{section}{%
+\subsection{\texorpdfstring{\textcolor{blue}{Technical Possibilities are Endless}}{}}\label{section}}
 
@@
 We already highlighted the drawbacks of the SD density ratio.
-Yet, bridge sampling may also fail in some cases.
+\textcolor{blue}{Bayes factor estimates can also be biased.
+However, bridge sampling has shown to be relatively unbiased [@sarafoglou2021evaluating].} Yet, bridge sampling may also fail in some cases.
 Failures of the bridge sampling algorithm are identified by unstable Bayes factors across repeated sampling runs.
@@
 
+\leavevmode\hypertarget{ref-van2021bayes}{}%
+Doorn, J. van, Aust, F., Haaf, J. M., Stefan, A. M., \& Wagenmakers, E.-J. (2021). Bayes factors for mixed models. \emph{Computational Brain \& Behavior}, 1--13.
+
 \leavevmode\hypertarget{ref-R-Rcpp_b}{}%
@@
 
-\leavevmode\hypertarget{ref-schad2019toward}{}%
-Schad, D. J., Betancourt, M., \& Vasishth, S. (2019). Toward a principled {Bayesian} workflow in cognitive science. \emph{arXiv Preprint arXiv:1904.12765}.
-
 \leavevmode\hypertarget{ref-schad2021workflow}{}%
@@
 
-\leavevmode\hypertarget{ref-van2021bayes}{}%
-van Doorn, J., Aust, F., Haaf, J.
-M., Stefan, A., \& Wagenmakers, E.-J.
-(2021).
-Bayes factors for mixed models.
-\emph{PsyArXiv Preprint}.
-doi:\href{https://doi.org/10.31234/osf.io/y65h8}{10.31234/osf.io/y65h8}
-
 \leavevmode\hypertarget{ref-van2018simple}{}%

### 2023-05-08 manuscript.Rmd
Updates revision

@@
 This indicates that the chains have not converged.
+\Revision{The current adhered criteria is that $\hat{R}$ should not exceed 1.01 }[@vehtari2019rank].
 There are different types of $\hat{R}$, such as the split- and rank-based $\hat{R}$ [@gelman2013bayesian; @vehtari2019rank].
 Even though researchers tend to trust convergence statistics more than visual inspection, these statistics are also not fool-proof.
-Therefore, the development of better convergence statistics is a current topic in Bayesian analysis methods.
+\Revision{For instance, in hierarchical modeling, with many parameters, it is more likely that parameters will exceed the threshold while there are no issues with the parameter estimation.
+It is, therefore, advisable to evaluate parameters carefully.} The development of better convergence statistics is a current topic in Bayesian analysis methods.
 
 The $\hat{R}$ values for all individual and general symbolic distance effect parameters computed by `rstan` and `brms` are shown in panel C of Figure \@ref(fig:traceplots3p).
-The figure shows that $\hat{R}$'s are not substantially greater than 1.
+The figure shows that $\hat{R}$'s are not substantially greater than \Revision{1.01}.
 Therefore, it seems that the distributions have converged.
@@
 #### Number of Effective Samples 
-Another popular convergence statistic is the number of effective samples.
+Another popular convergence statistic is the number of effective samples (ESS).
 This statistic is concerned with the dependence of posterior samples.
@@
 Therefore, the number of effective samples can be seen as a measure of the amount of new information about the posterior distribution that is provided by the total number of samples drawn from the posterior.
-A common rule for interpretation is that the number should be close to or even greater than the number of total iterations [@vehtari2019rank; @neff2018].
+\Revision{A common rule for interpretation is that the number of effective samples should equal at least 100 per chain} [@vehtari2019rank].
+\Revision{In the case of four chains, this would mean a value of 400 or higher.
+The higher the value the better.
+There are different ways to estimate the number of effective samples.
+Bulk-ESS looks, for instance, at number of effective samples at the centre of the posterior distribution, while Tail-ESS evaluates the tails of the posterior distribution ().
+In this tutorial, ...
+.}
 

### 2023-05-08 manuscript.tex
Updates revision

@@
 \textcolor{blue}{If we formulate a prior on standard deviations, there is always an implied prior on the variances as well.
-If we think of the distribution as $f(x)$, $x$ can either be defined as $\sqrt{(var)}$, if we formulated the prior on the standard deviation, or $x$ can be defined as $(sd^2)$ if we formulated the prior on the variance.
+If we think of the distribution as $f(x)$, $x$ can either be defined as $\sqrt{(var)}$, if we formulated the prior on the standard deviation, or $x$ can be defined as ${sd}^2$ if we formulated the prior on the variance.
 From this, we can derive the implied distribution $f(var)$ or $f(sd)$.
 Notably, this is not the same as calculating $\sqrt{(f(x))}$ or $f(x)^2$.
-To transform the priors on standard deviations to variances, we try to find the parameters that make the prior distribution match the implied distribution.} In our \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/tree/main/R\%20code/Prior\%20transformation}{Online Supplement}, we provide R-code with a numeric solution to the transformation.
+To transform the priors on standard deviations to variances, we try to find the parameters that match the prior distribution to the implied distribution.} In our \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/tree/main/R\%20code/Prior\%20transformation}{Online Supplement}, we provide R-code with a numeric solution to the transformation.
 For the \texttt{brms} priors, we obtain the following implied prior distributions on variances:
@@
 This indicates that the chains have not converged.
+\textcolor{blue}{The current adhered criteria is that $\hat{R}$ should not exceed 1.01 }(Vehtari, Gelman, Simpson, Carpenter, \& Bürkner, 2019).
 There are different types of \(\hat{R}\), such as the split- and rank-based \(\hat{R}\) (Gelman et al., 2013; Vehtari, Gelman, Simpson, Carpenter, \& Bürkner, 2019).
 Even though researchers tend to trust convergence statistics more than visual inspection, these statistics are also not fool-proof.
-Therefore, the development of better convergence statistics is a current topic in Bayesian analysis methods.
+\textcolor{blue}{For instance, in hierarchical modeling, with many parameters, it is more likely that parameters will exceed the threshold while there are no issues with the parameter estimation.
+It is, therefore, advisable to evaluate separate parameters carefully.} The development of better convergence statistics is a current topic in Bayesian analysis methods.
 
 The \(\hat{R}\) values for all individual and general symbolic distance effect parameters computed by \texttt{rstan} and \texttt{brms} are shown in panel C of Figure \ref{fig:traceplots3p}.
-The figure shows that \(\hat{R}\)'s are not substantially greater than 1.
+The figure shows that \(\hat{R}\)'s are not substantially greater than \textcolor{blue}{1.01}.
 Therefore, it seems that the distributions have converged.
@@
 
-Another popular convergence statistic is the number of effective samples.
+Another popular convergence statistic is the number of effective samples (ESS).
 This statistic is concerned with the dependence of posterior samples.
@@
 Therefore, the number of effective samples can be seen as a measure of the amount of new information about the posterior distribution that is provided by the total number of samples drawn from the posterior.
-A common rule for interpretation is that the number should be close to or even greater than the number of total iterations (Carpenter, 2018; Vehtari, Gelman, Simpson, Carpenter, \& Bürkner, 2019).
+\textcolor{blue}{A common rule for interpretation is that the number should be at least 100 per chain.}(Vehtari, Gelman, Simpson, Carpenter, \& Bürkner, 2019) \textcolor{blue}{In case of four chains, this would mean a value of 400 or higher.
+The higher the value the better.
+There are different ways to estimate the number of effective samples.
+Bulk-ESS looks, for instance, at number of effective samples at the centre of the posterior distribution, while Tail-ESS evaluates the tails of the posterior distribution ().
+In this tutorial, ...
+.}
 
@@
 
-\leavevmode\hypertarget{ref-neff2018}{}%
-Carpenter, B.
-(2018).
-We were measuring the speed of stan incorrectly -- it's faster than we thought in some cases due to antithetical sampling {[}blog post{]}.
-Retrieved from \url{https://statmodeling.stat.columbia.edu/2018/01/18/\%20measuring-speed-stan-incorrectly-faster-thought-cases-due-antithetical-sampling/}
-
 \leavevmode\hypertarget{ref-carpenter2017stan}{}%

### 2023-05-08 manuscript.Rmd
Updates revision (including new figure means se)

@@
 The higher the value the better.
-There are different ways to estimate the number of effective samples.
-Bulk-ESS looks, for instance, at number of effective samples at the centre of the posterior distribution, while Tail-ESS evaluates the tails of the posterior distribution ().
-In this tutorial, ...
-.}
+To get a better understanding of convergence, Vehtari and colleagues (2021) recommended evaluating different quantities of the posterior distribution instead, using so-called Bulk- and Tail-ESS.
+Bulk-ESS evaluates the center of the posterior distribution, while Tail-ESS evaluates the tails of the posterior distribution.
+For this tutorial, we focus on the basic calculation of the number of effective samples suggested by Gelman et al.
+(2013) and improved by Vehtari et al.
+(2021).}
 
@@
 
-```{r modelposteriorsmufigure3, echo=FALSE, fig.cap="The posterior distributions for the general effects estimated by the two packages.
+```{r modelposteriorsmufigure3, echo=FALSE, fig.cap="Panel A shows the empirical mean RTs per digit.
+The standard error is represented by bars around the dots.
+Panel B and C show the posterior distributions for the general effects estimated by the two packages.
 The middle line within the distributions represents the posterior mean.
 The shaded area within the distributions represent 95\\% credible intervals.", fig.align = "left", fig.height = 11, fig.width = 1, out.width="100%"}
-knitr::include_graphics("Images/Fig9_posteriormu_v2.png")
+knitr::include_graphics("Images/Fig9_posteriormu_v3.png")
 ```

### 2023-05-08 manuscript.tex
Updates revision (including new figure means se)

@@
 
-Before interpreting the estimation results, it has to be checked whether the posterior distribution of the parameters has converged, meaning that a stationary posterior distribution has been reached (Vehtari, Gelman, Simpson, Carpenter, \& Bürkner, 2021).
+Before interpreting the estimation results, it has to be checked whether the posterior distribution of the parameters has converged, meaning that a stationary posterior distribution has been reached (Vehtari, Gelman, Simpson, Carpenter, \& Bürkner, 2021a).
 We will explain these convergence checks in more detail in the section Model Diagnostics.
@@
 Before we can interpret the results, we have to evaluate whether they are reliable.
-Specifically, it has to be checked whether the sampling chains of the parameters have converged, meaning that a stationary posterior distribution has been reached (Vehtari, Gelman, Simpson, Carpenter, \& Bürkner, 2021).
+Specifically, it has to be checked whether the sampling chains of the parameters have converged, meaning that a stationary posterior distribution has been reached (Vehtari, Gelman, Simpson, Carpenter, \& Bürkner, 2021a).
 If convergence is not achieved, the posterior estimates may change substantially if the sampling algorithm was run with more iterations or different starting values.
@@
 In this case, the chains should overlap.
-If the posterior distribution did not converge, one chain could, for instance, be found on top of the figure, while all other chains are on the bottom (for an illustration see Vehtari, Gelman, Simpson, Carpenter, \& Bürkner, 2019).
+If the posterior distribution did not converge, one chain could, for instance, be found on top of the figure, while all other chains are on the bottom (for an illustration see Vehtari, Gelman, Simpson, Carpenter, \& Bürkner, 2021b).
 Or, perhaps the first half of the chain is on the top of the graph and the second half on the bottom.
@@
 This indicates that the chains have not converged.
-\textcolor{blue}{The current adhered criteria is that $\hat{R}$ should not exceed 1.01 }(Vehtari, Gelman, Simpson, Carpenter, \& Bürkner, 2019).
-There are different types of \(\hat{R}\), such as the split- and rank-based \(\hat{R}\) (Gelman et al., 2013; Vehtari, Gelman, Simpson, Carpenter, \& Bürkner, 2019).
+\textcolor{blue}{The current adhered criteria is that $\hat{R}$ should not exceed 1.01 }(Vehtari, Gelman, Simpson, Carpenter, \& Bürkner, 2021b).
+There are different types of \(\hat{R}\), such as the split- and rank-based \(\hat{R}\) (Gelman et al., 2013; Vehtari, Gelman, Simpson, Carpenter, \& Bürkner, 2021b).
 Even though researchers tend to trust convergence statistics more than visual inspection, these statistics are also not fool-proof.
 \textcolor{blue}{For instance, in hierarchical modeling, with many parameters, it is more likely that parameters will exceed the threshold while there are no issues with the parameter estimation.
-It is, therefore, advisable to evaluate separate parameters carefully.} The development of better convergence statistics is a current topic in Bayesian analysis methods.
+It is, therefore, advisable to evaluate parameters carefully.} The development of better convergence statistics is a current topic in Bayesian analysis methods.
 
@@
 Therefore, the number of effective samples can be seen as a measure of the amount of new information about the posterior distribution that is provided by the total number of samples drawn from the posterior.
-\textcolor{blue}{A common rule for interpretation is that the number should be at least 100 per chain.}(Vehtari, Gelman, Simpson, Carpenter, \& Bürkner, 2019) \textcolor{blue}{In case of four chains, this would mean a value of 400 or higher.
+\textcolor{blue}{A common rule for interpretation is that the number of effective samples should equal at least 100 per chain} (Vehtari, Gelman, Simpson, Carpenter, \& Bürkner, 2021b).
+\textcolor{blue}{In the case of four chains, this would mean a value of 400 or higher.
 The higher the value the better.
-There are different ways to estimate the number of effective samples.
-Bulk-ESS looks, for instance, at number of effective samples at the centre of the posterior distribution, while Tail-ESS evaluates the tails of the posterior distribution ().
-In this tutorial, ...
-.}
+To get a better understanding of convergence, Vehtari and colleagues (2021) recommended evaluating different quantities of the posterior distribution instead, using so-called Bulk- and Tail-ESS.
+Bulk-ESS evaluates the center of the posterior distribution, while Tail-ESS evaluates the tails of the posterior distribution.
+For this tutorial, we focus on the basic calculation of the number of effective samples suggested by Gelman et al.
+(2013) and improved by Vehtari et al.
+(2021).}
 
@@
 
-\includegraphics[width=1\linewidth]{Images/Fig9_posteriormu_v2} \hfill{}
-
-\caption{The posterior distributions for the general effects estimated by the two packages.
+\includegraphics[width=1\linewidth]{Images/Fig9_posteriormu_v3} \hfill{}
+
+\caption{Panel A shows the empirical mean RT per digit.
+The standard error is represented by bars around the dots.
+Panel B and C show the posterior distributions for the general effects estimated by the two packages.
 The middle line within the distributions represents the posterior mean.
@@
 
-\leavevmode\hypertarget{ref-vehtari2019rank}{}%
-Vehtari, A., Gelman, A., Simpson, D., Carpenter, B., \& Bürkner, P.-C. (2019). \emph{arXiv Preprint arXiv:1903.08008}.
-
 \leavevmode\hypertarget{ref-Vehtari_2021}{}%
 Vehtari, A., Gelman, A., Simpson, D., Carpenter, B., \& Bürkner, P.-C.
-(2021).
-Rank-normalization, folding, and localization: An improved rˆ for assessing convergence of MCMC (with discussion).
+(2021a).
+Rank-normalization, folding, and localization: An improved r{}for assessing convergence of MCMC (with discussion).
 \emph{Bayesian Analysis}, \emph{16}(2).
 doi:\href{https://doi.org/10.1214/20-ba1221}{10.1214/20-ba1221}
+
+\leavevmode\hypertarget{ref-vehtari2019rank}{}%
+Vehtari, A., Gelman, A., Simpson, D., Carpenter, B., \& Bürkner, P.-C.
+(2021b).
+Rank-normalization, folding, and localization: An improved rhat for assessing convergence of MCMC (with discussion).
+\emph{Bayesian Analysis}, \emph{16}(2), 667--718.
 

### 2023-05-09 manuscript.Rmd
Updates added references

@@
 Our tutorial is directed at researchers who have a basic \Revision{conceptual} understanding of Bayesian inference and parameter estimation and are planning to perform more complex Bayesian analyses \Revision{of psychological phenomena.
-We hope to achieve a balance of detail and overview---while we want to provide guidance for a full Bayesian workflow, similar to Gelman et al.
-[-@gelman2020] and Schad et al.[-@schad2019toward; -@schad2021workflow], we don't shy away from in-depth explanations for the most curious scientists.
+We hope to achieve a balance of detail and overview---while we want to provide guidance for a full Bayesian workflow, similar to Gelman et al.} [-@gelman2020] \Revision{and Schad et al.}[-@schad2021toward; -@schad2021workflow]\Revision{, we don't shy away from in-depth explanations for the most curious scientists.
 To ensure that a broad audience is able to follow the tutorial, we limit and carefully explain equations.
-However, rudimentary knowledge of calculus is advantageous to fully understand Bayesian analysis.}
+However, rudimentary knowledge of calculus is advantageous to fully understand Bayesian analysis.
+}
 
@@
 
-The Bayesian framework offers an intuitive approach for hierarchical modeling (Rouder & Lu, 2005; Lynch, 2007).
+\Revision{There are several tutorials for readers interested in the frequentist approach of hierarchical model estimation} [e.g., @singmann2019introduction; @west2022linear].
+\Revision{However, it has been argued that} the Bayesian framework offers a more intuitive approach for hierarchical modeling (Rouder & Lu, 2005; Lynch, 2007).
 The flexibility of the Bayesian approach facilitates the implementation of complex hierarchical models (Rouder, Morey, & Pratte, 2013).
@@
 \Revision{In contrast to the confidence interval, the credible interval provides the probability of a range of values for an effect.
-However, considering the prior and under the assumption that there is an effect, the credible interval does not provide evidence to conclude whether there is an effect [Add Wagenmakers - principle of predictable irrelevance].
-}
+However, considering the prior and under the assumption that there is an effect, the credible interval does not provide evidence to conclude whether there is an effect} [@wagenmakers2020].
 
@@
 Two other useful tutorials for Bayesian modeling are Schad, Betancourt, and
-Vasishth [-@schad2019toward] and Schad, Nicenboim, Bürkner, Betancourt, and Vasishth [-@schad2021workflow].
+Vasishth [-@schad2021toward] and Schad, Nicenboim, Bürkner, Betancourt, and Vasishth [-@schad2021workflow].
 \Revision{There are also other R software packages for Bayesian multilevel modeling besides `brms` and `rstan`, such as `BayesFactor`[@R-BayesFactor] and `rstanarm`[@rstanarm2020], that we did not illustrate in this tutorial.

### 2023-05-09 manuscript.tex
Updates added references

@@
 Our tutorial is directed at researchers who have a basic \textcolor{blue}{conceptual} understanding of Bayesian inference and parameter estimation and are planning to perform more complex Bayesian analyses \textcolor{blue}{of psychological phenomena.
-We hope to achieve a balance of detail and overview---while we want to provide guidance for a full Bayesian workflow, similar to Gelman et al.
-[-@gelman2020] and Schad et al.[-@schad2019toward; -@schad2021workflow], we don't shy away from in-depth explanations for the most curious scientists.
+We hope to achieve a balance of detail and overview---while we want to provide guidance for a full Bayesian workflow, similar to Gelman et al.} (\textbf{gelman2020?}) \textcolor{blue}{and Schad et al.}(2021; 2021)\textcolor{blue}{, we don't shy away from in-depth explanations for the most curious scientists.
 To ensure that a broad audience is able to follow the tutorial, we limit and carefully explain equations.
-However, rudimentary knowledge of calculus is advantageous to fully understand Bayesian analysis.}
+However, rudimentary knowledge of calculus is advantageous to fully understand Bayesian analysis.
+}
 
@@
 
-The Bayesian framework offers an intuitive approach for hierarchical modeling (Rouder \& Lu, 2005; Lynch, 2007).
+\textcolor{blue}{There are several tutorials for readers interested in the frequentist approach of hierarchical model estimation} (e.g., Singmann \& Kellen, 2019; West, Welch, \& Galecki, 2022).
+\textcolor{blue}{However, it has been argued that} the Bayesian framework offers a more intuitive approach for hierarchical modeling (Rouder \& Lu, 2005; Lynch, 2007).
 The flexibility of the Bayesian approach facilitates the implementation of complex hierarchical models (Rouder, Morey, \& Pratte, 2013).
@@
 
-\caption{Panel A shows the empirical mean RT per digit.
+\caption{Panel A shows the empirical mean RTs per digit.
 The standard error is represented by bars around the dots.
@@
 \textcolor{blue}{In contrast to the confidence interval, the credible interval provides the probability of a range of values for an effect.
-However, considering the prior and under the assumption that there is an effect, the credible interval does not provide evidence to conclude whether there is an effect [Add Wagenmakers - principle of predictable irrelevance].
-}
+However, considering the prior and under the assumption that there is an effect, the credible interval does not provide evidence to conclude whether there is an effect} (\textbf{wagenmakers2020?}).
 
@@
 Two other useful tutorials for Bayesian modeling are Schad, Betancourt, and
-Vasishth (\textbf{schad2019toward?}) and Schad, Nicenboim, Bürkner, Betancourt, and Vasishth (2021).
+Vasishth (2021) and Schad, Nicenboim, Bürkner, Betancourt, and Vasishth (2021).
 \textcolor{blue}{There are also other R software packages for Bayesian multilevel modeling besides `brms` and `rstan`, such as `BayesFactor`[@R-BayesFactor] and `rstanarm`[@rstanarm2020], that we did not illustrate in this tutorial.
@@
 
+\leavevmode\hypertarget{ref-schad2021toward}{}%
+Schad, D.
+J., Betancourt, M., \& Vasishth, S.
+(2021).
+Toward a principled bayesian workflow in cognitive science.
+\emph{Psychological Methods}, \emph{26}(1), 103.
+doi:\href{https://doi.org/10.1037/met0000275}{10.1037/met0000275}
+
 \leavevmode\hypertarget{ref-schad2021workflow}{}%
@@
 \emph{Cognitive Science}, \emph{32}(8), 1248--1284.
+
+\leavevmode\hypertarget{ref-singmann2019introduction}{}%
+Singmann, H., \& Kellen, D. (2019). An introduction to mixed models for experimental psychology. In \emph{New methods in cognitive psychology} (pp. 4--31). Routledge.
 
@@
 \emph{Psychonomic Bulletin \& Review}, \emph{25}(1), 35--57.
+doi:\href{https://doi.org/10.3758/s13423-017-1343-3}{10.3758/s13423-017-1343-3}
+
+\leavevmode\hypertarget{ref-west2022linear}{}%
+West, B. T., Welch, K. B., \& Galecki, A. T. (2022). \emph{Linear mixed models: A practical guide using statistical software}. Crc Press.
 

### 2023-05-09 manuscript.Rmd
Updates revision (added references + change names supplements)

@@
 Our tutorial is directed at researchers who have a basic \Revision{conceptual} understanding of Bayesian inference and parameter estimation and are planning to perform more complex Bayesian analyses \Revision{of psychological phenomena.
-We hope to achieve a balance of detail and overview---while we want to provide guidance for a full Bayesian workflow, similar to Gelman et al.} [-@gelman2020] \Revision{and Schad et al.}[-@schad2021toward; -@schad2021workflow]\Revision{, we don't shy away from in-depth explanations for the most curious scientists.
+We hope to achieve a balance of detail and overview---while we want to provide guidance for a full Bayesian workflow, similar to Gelman et al.} [-@gelman2020bayesian] \Revision{and Schad et al.}[-@schad2021toward; -@schad2021workflow]\Revision{, we don't shy away from in-depth explanations for the most curious scientists.
 To ensure that a broad audience is able to follow the tutorial, we limit and carefully explain equations.
@@
 Notably, this is not the same as calculating $\sqrt{(f(x))}$ or $f(x)^2$.
-To transform the priors on standard deviations to variances, we try to find the parameters that match the prior distribution to the implied distribution.} In our [Online Supplement](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/tree/main/R%20code/Prior%20transformation), we provide R-code with a numeric solution to the transformation.
+To transform the priors on standard deviations to variances, we try to find the parameters that match the prior distribution to the implied distribution.} In [Online Supplement J](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/tree/main/R%20code/Prior%20transformation), we provide R-code with a numeric solution to the transformation.
 For the `brms` priors, we obtain the following implied prior distributions on variances:
@@
 The results of the prior prediction for the symbolic distance effect with the priors from Equation \@ref(eq:chosenpriors1) with the number of participants $n = 10$, the number of trials per condition $k = 60$, and the number of simulations $m = 1000$ are shown in panel A of Figure \@ref(fig:priorpredfigurepanel).
-Code for the simulations can be found in our [Online Supplement](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment).
+Code for the simulations can be found in [Online Supplement J](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/blob/main/R%20code/Figures/Fig5_priorpredictions_trunc_v3.R).
 
@@
 In this section, we explain how a model can be fitted in each package using code snippets.
-In addition, we offer a separate R Markdown file for each package in our [Online Supplement](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment) that explains in more depth how to analyze the full symbolic distance effect model with the specific package.
+In addition, we offer separate R Markdown files for each package in [Online Supplements C-F](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment) that explains in more depth how to analyze the full symbolic distance effect model with the specific package.
 
@@
 An example `.stan` file of a simple, non-hierarchical normal model, where the variance is already known and the mean has to be estimated, is shown below [adapted example from @nicenboim2021introduction, Chapter 10].
-The `.stan` file for the symbolic distance effect, including explanations, can be found in the [Online Supplement](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/blob/main/R%20objects/rstan/normal%20model/myModel.stan).
+The `.stan` file for the symbolic distance effect, including explanations, can be found in [Online Supplement K](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/blob/main/R%20objects/rstan/normal%20model/myModel.stan).
 The file is divided into three sections: `data`, `parameters`, and `model`.
@@
 \Revision{In contrast to the confidence interval, the credible interval provides the probability of a range of values for an effect.
-However, considering the prior and under the assumption that there is an effect, the credible interval does not provide evidence to conclude whether there is an effect} [@wagenmakers2020].
+However, considering the prior and under the assumption that there is an effect, the credible interval does not provide evidence to conclude whether there is an effect} [@wagenmakers2020principle].
 

### 2023-05-09 manuscript.tex
Updates revision (added references + change names supplements)

@@
 Our tutorial is directed at researchers who have a basic \textcolor{blue}{conceptual} understanding of Bayesian inference and parameter estimation and are planning to perform more complex Bayesian analyses \textcolor{blue}{of psychological phenomena.
-We hope to achieve a balance of detail and overview---while we want to provide guidance for a full Bayesian workflow, similar to Gelman et al.} (\textbf{gelman2020?}) \textcolor{blue}{and Schad et al.}(2021; 2021)\textcolor{blue}{, we don't shy away from in-depth explanations for the most curious scientists.
+We hope to achieve a balance of detail and overview---while we want to provide guidance for a full Bayesian workflow, similar to Gelman et al.} (2020) \textcolor{blue}{and Schad et al.}(2021; 2022)\textcolor{blue}{, we don't shy away from in-depth explanations for the most curious scientists.
 To ensure that a broad audience is able to follow the tutorial, we limit and carefully explain equations.
@@
 This is called a prior sensitivity analysis (Roos, Martins, Held, \& Rue, 2015).
-Generally, when it comes to prior specification, it is important to be \emph{transparent} and to provide a justification for the selected prior distributions (Stefan, Katsimpokis, Gronau, \& Wagenmakers, 2021).
+Generally, when it comes to prior specification, it is important to be \emph{transparent} and to provide a justification for the selected prior distributions (Stefan, Katsimpokis, Gronau, \& Wagenmakers, 2022).
 Prior distributions are rarely inadmissible.
@@
 \textcolor{blue}{In contrast to the confidence interval, the credible interval provides the probability of a range of values for an effect.
-However, considering the prior and under the assumption that there is an effect, the credible interval does not provide evidence to conclude whether there is an effect} (\textbf{wagenmakers2020?}).
+However, considering the prior and under the assumption that there is an effect, the credible interval does not provide evidence to conclude whether there is an effect} (Wagenmakers, Lee, Rouder, \& Morey, 2020).
 
@@
 Two other useful tutorials for Bayesian modeling are Schad, Betancourt, and
-Vasishth (2021) and Schad, Nicenboim, Bürkner, Betancourt, and Vasishth (2021).
+Vasishth (2021) and Schad, Nicenboim, Bürkner, Betancourt, and Vasishth (2022).
 \textcolor{blue}{There are also other R software packages for Bayesian multilevel modeling besides `brms` and `rstan`, such as `BayesFactor`[@R-BayesFactor] and `rstanarm`[@rstanarm2020], that we did not illustrate in this tutorial.
@@
 Schönbrodt and Wagenmakers (2018) and Stefan and colleagues (2019) introduced Bayes factor design analysis for fixed and sequential designs.
-However, these approaches to study planning have not yet been fully extended to a hierarchical setup (but see recent efforts by Vasishth et al., 2021).
+However, these approaches to study planning have not yet been fully extended to a hierarchical setup (but see recent efforts by Vasishth et al., 2023).
 The extension to nested designs is not entirely trivial because, in addition to planning the number of participants, we also need to consider the number of trials (Rouder \& Haaf, 2018).
@@
 \leavevmode\hypertarget{ref-burkner2019ordinal}{}%
-Bürkner, P.-C., \& Vuorre, M. (2019). Ordinal regression models in psychology: A tutorial. \emph{Advances in Methods and Practices in Psychological Science}, \emph{2}(1), 77--101.
+Bürkner, P.-C., \& Vuorre, M.
+(2019).
+Ordinal regression models in psychology: A tutorial.
+\emph{Advances in Methods and Practices in Psychological Science}, \emph{2}(1), 77--101.
+doi:\href{https://doi.org/10.1177/251524591882319}{10.1177/251524591882319}
 
@@
 \leavevmode\hypertarget{ref-van2021bayes}{}%
-Doorn, J. van, Aust, F., Haaf, J. M., Stefan, A. M., \& Wagenmakers, E.-J. (2021). Bayes factors for mixed models. \emph{Computational Brain \& Behavior}, 1--13.
+Doorn, J.
+van, Aust, F., Haaf, J.
+M., Stefan, A.
+M., \& Wagenmakers, E.-J.
+(2021).
+Bayes factors for mixed models.
+\emph{Computational Brain \& Behavior}, 1--13.
+doi:\href{https://doi.org/10.1007/s42113-021-00113-2}{10.1007/s42113-021-00113-2}
 
@@
 \leavevmode\hypertarget{ref-gelman2006prior}{}%
-Gelman, A. (2006). Prior distributions for variance parameters in hierarchical models (comment on article by {Browne} and {Draper}). \emph{Bayesian Analysis}, \emph{1}(3), 515--534.
+Gelman, A.
+(2006).
+Prior distributions for variance parameters in hierarchical models (comment on article by {Browne} and {Draper}).
+\emph{Bayesian Analysis}, \emph{1}(3), 515--534.
+doi:\href{https://doi.org/10.1214/06-BA117A}{10.1214/06-BA117A}
 
@@
 Gelman, A., \& Rubin, D. B. (1992). Inference from iterative simulation using multiple sequences. \emph{Statistical Science}, \emph{7}(4), 457--472.
+
+\leavevmode\hypertarget{ref-gelman2020bayesian}{}%
+Gelman, A., Vehtari, A., Simpson, D., Margossian, C. C., Carpenter, B., Yao, Y., \ldots{} Modrák, M. (2020). Bayesian workflow. \emph{arXiv Preprint arXiv:2011.01808}.
 
@@
 \emph{Statistics and Computing}, \emph{25}(4), 835--862.
+doi:\href{https://doi.org/10.1007/s11222-015-9574-5}{10.1007/s11222-015-9574-5}
 
@@
 \emph{Journal of Mathematical Psychology}, \emph{81}, 80--97.
+doi:\href{https://doi.org/10.1016/j.jmp.2017.09.005}{10.1016/j.jmp.2017.09.005}
 
@@
 \emph{Psychometrika}, \emph{84}(1), 261--284.
+doi:\href{https://doi.org/10.1007/s11336-018-9648-3}{10.1007/s11336-018-9648-3}
 
@@
 \leavevmode\hypertarget{ref-haaf2019capturing}{}%
-Haaf, J. M., Klaassen, F., \& Rouder, J. (2019). Capturing ordinal theoretical constraint in psychological science.
+Haaf, J. M., Klaassen, F., \& Rouder, J. (2019). Capturing ordinal theoretical constraint in psychological science. doi:\href{https://doi.org/10.31234/osf.io/a4xu9}{10.31234/osf.io/a4xu9}
 
 \leavevmode\hypertarget{ref-haaf2017developing}{}%
-Haaf, J. M., \& Rouder, J. N. (2017). Developing constraint in {Bayesian} mixed models. \emph{Psychological Methods}, \emph{22}(4), 779.
+Haaf, J.
+M., \& Rouder, J.
+N.
+(2017).
+Developing constraint in {Bayesian} mixed models.
+\emph{Psychological Methods}, \emph{22}(4), 779.
+doi:\href{https://doi.org/10.1037/met0000156}{10.1037/met0000156}
 
 \leavevmode\hypertarget{ref-haaf2019some}{}%
-Haaf, J. M., \& Rouder, J. N. (2019). Some do and some don't? Accounting for variability of individual difference structures. \emph{Psychonomic Bulletin \& Review}, \emph{26}(3), 772--789.
+Haaf, J.
+M., \& Rouder, J.
+N.
+(2019).
+Some do and some don't?
+Accounting for variability of individual difference structures.
+\emph{Psychonomic Bulletin \& Review}, \emph{26}(3), 772--789.
+doi:\href{https://doi.org/10.3758/s13423-018-1522-x}{10.3758/s13423-018-1522-x}
 
@@
 \emph{Psych}, \emph{3}(4), 751--779.
+doi:\href{https://doi.org/10.3390/psych3040048}{10.3390/psych3040048}
 
@@
 \emph{British Journal of Mathematical and Statistical Psychology}, \emph{72}(2), 316--333.
+doi:\href{https://doi.org/10.1111/bmsp.12150}{10.1111/bmsp.12150}
 
@@
 \emph{Journal of the American Statistical Association}, \emph{91}(436), 1461--1473.
+doi:\href{https://doi.org/10.2307/2291572}{10.2307/2291572}
 
@@
 \leavevmode\hypertarget{ref-lee2011cognitive}{}%
-Lee, M. D. (2011). How cognitive modeling can benefit from hierarchical {Bayesian} models. \emph{Journal of Mathematical Psychology}, \emph{55}(1), 1--7.
+Lee, M.
+D.
+(2011).
+How cognitive modeling can benefit from hierarchical {Bayesian} models.
+\emph{Journal of Mathematical Psychology}, \emph{55}(1), 1--7.
+doi:\href{https://doi.org/10.1016/j.jmp.2010.08.013}{10.1016/j.jmp.2010.08.013}
 
@@
 \emph{Journal of Mathematical Psychology}, \emph{55}(5), 368--378.
+doi:\href{https://doi.org/10.1016/j.jmp.2011.06.004}{10.1016/j.jmp.2011.06.004}
 
 \leavevmode\hypertarget{ref-moyer1967time}{}%
-Moyer, R. S., \& Landauer, T. K. (1967). Time required for judgements of numerical inequality. \emph{Nature}, \emph{215}(5109), 1519--1520.
+Moyer, R.
+S., \& Landauer, T.
+K.
+(1967).
+Time required for judgements of numerical inequality.
+\emph{Nature}, \emph{215}(5109), 1519--1520.
+doi:\href{https://doi.org/10.1038/2151519a0}{10.1038/2151519a0}
 
@@
 \leavevmode\hypertarget{ref-roos2015sensitivity}{}%
-Roos, M., Martins, T. G., Held, L., \& Rue, H. (2015). Sensitivity analysis for {Bayesian} hierarchical models. \emph{Bayesian Analysis}, \emph{10}(2), 321--349.
+Roos, M., Martins, T.
+G., Held, L., \& Rue, H.
+(2015).
+Sensitivity analysis for {Bayesian} hierarchical models.
+\emph{Bayesian Analysis}, \emph{10}(2), 321--349.
+doi:\href{https://doi.org/10.1214/14-BA909}{10.1214/14-BA909}
 
 \leavevmode\hypertarget{ref-rouder2014optional}{}%
-Rouder, J. N. (2014). Optional stopping: No problem for {Bayesians}. \emph{Psychonomic Bulletin \& Review}, \emph{21}(2), 301--308.
+Rouder, J.
+N.
+(2014).
+Optional stopping: No problem for {Bayesians}.
+\emph{Psychonomic Bulletin \& Review}, \emph{21}(2), 301--308.
+doi:\href{https://doi.org/10.3758/s13423-014-0595-4}{10.3758/s13423-014-0595-4}
 
@@
 \emph{Advances in Methods and Practices in Psychological Science}, \emph{1}(1), 19--26.
+doi:\href{https://doi.org/10.1177/2515245917745058}{10.1177/2515245917745058}
 
 \leavevmode\hypertarget{ref-rouder2018theories}{}%
-Rouder, J. N., Haaf, J. M., \& Aust, F. (2018). From theories to models to predictions: A {Bayesian} model comparison approach. \emph{Communication Monographs}, \emph{85}(1), 41--56.
+Rouder, J.
+N., Haaf, J.
+M., \& Aust, F.
+(2018).
+From theories to models to predictions: A {Bayesian} model comparison approach.
+\emph{Communication Monographs}, \emph{85}(1), 41--56.
+doi:\href{https://doi.org/10.1080/03637751.2017.1394581}{10.1080/03637751.2017.1394581}
 
 \leavevmode\hypertarget{ref-rouder2005introduction}{}%
-Rouder, J. N., \& Lu, J. (2005). An introduction to {Bayesian} hierarchical models with an application in the theory of signal detection. \emph{Psychonomic Bulletin \& Review}, \emph{12}(4), 573--604.
+Rouder, J.
+N., \& Lu, J.
+(2005).
+An introduction to {Bayesian} hierarchical models with an application in the theory of signal detection.
+\emph{Psychonomic Bulletin \& Review}, \emph{12}(4), 573--604.
+doi:\href{https://doi.org/10.3758/BF03196750}{10.3758/BF03196750}
 
 \leavevmode\hypertarget{ref-rouder2005hierarchical}{}%
-Rouder, J. N., Lu, J., Speckman, P., Sun, D., \& Jiang, Y. (2005). A hierarchical model for estimating response time distributions. \emph{Psychonomic Bulletin \& Review}, \emph{12}(2), 195--223.
+Rouder, J.
+N., Lu, J., Speckman, P., Sun, D., \& Jiang, Y.
+(2005).
+A hierarchical model for estimating response time distributions.
+\emph{Psychonomic Bulletin \& Review}, \emph{12}(2), 195--223.
+doi:\href{https://doi.org/10.3758/BF03257252}{10.3758/BF03257252}
 
 \leavevmode\hypertarget{ref-rouder2018teaching}{}%
-Rouder, J. N., \& Morey, R. D. (2018). Teaching {Bayes}' theorem: Strength of evidence as predictive accuracy. \emph{The American Statistician}.
+Rouder, J.
+N., \& Morey, R.
+D.
+(2018).
+Teaching {Bayes}' theorem: Strength of evidence as predictive accuracy.
+\emph{The American Statistician}, \emph{73}(2), 186--190.
+doi:\url{https://doi.org/10.1080/00031305.2017.1341334}
 
@@
 \leavevmode\hypertarget{ref-schad2021workflow}{}%
-Schad, D. J., Nicenboim, B., Bürkner, P.-C., Betancourt, M., \& Vasishth, S. (2021). Workflow techniques for the robust use of bayes factors. Retrieved from \url{http://arxiv.org/abs/2103.08744}
+Schad, D.
+J., Nicenboim, B., Bürkner, P.-C., Betancourt, M., \& Vasishth, S.
+(2022).
+Workflow techniques for the robust use of bayes factors.
+\emph{Psychological Methods}.
+American Psychological Association.
+doi:\href{https://doi.org/10.1037/met0000472}{10.1037/met0000472}
 
 \leavevmode\hypertarget{ref-schonbrodt2018bayes}{}%
-Schönbrodt, F. D., \& Wagenmakers, E.-J. (2018). Bayes factor design analysis: Planning for compelling evidence. \emph{Psychonomic Bulletin \& Review}, \emph{25}(1), 128--142.
+Schönbrodt, F.
+D., \& Wagenmakers, E.-J.
+(2018).
+Bayes factor design analysis: Planning for compelling evidence.
+\emph{Psychonomic Bulletin \& Review}, \emph{25}(1), 128--142.
+doi:\href{https://doi.org/10.3758/s13423-017-1230-y}{10.3758/s13423-017-1230-y}
 
@@
 \emph{Psychological Methods}, \emph{22}(2), 322.
+doi:\href{https://doi.org/10.1037/met0000061}{10.1037/met0000061}
 
 \leavevmode\hypertarget{ref-schramm2019reaction}{}%
-Schramm, P., \& Rouder, J. (2019). Are reaction time transformations really beneficial?
+Schramm, P., \& Rouder, J. (2019). Are reaction time transformations really beneficial? doi:\href{https://doi.org/10.31234/osf.io/9ksa6}{10.31234/osf.io/9ksa6}
 
@@
 \emph{Cognitive Science}, \emph{32}(8), 1248--1284.
+doi:\href{https://doi.org/10.1080/03640210802414826}{10.1080/03640210802414826}
 
 \leavevmode\hypertarget{ref-singmann2019introduction}{}%
-Singmann, H., \& Kellen, D. (2019). An introduction to mixed models for experimental psychology. In \emph{New methods in cognitive psychology} (pp. 4--31). Routledge.
+Singmann, H., \& Kellen, D.
+(2019).
+An introduction to mixed models for experimental psychology.
+In \emph{New methods in cognitive psychology} (pp.
+4--31).
+Routledge.
+doi:\href{https://doi.org/10.4324/9780429318405-2}{10.4324/9780429318405-2}
 
@@
 Retrieved from \url{https://web.archive.org/web/20150206004624/http://www.bayesian-inference.com/software}
-
-\leavevmode\hypertarget{ref-Stefan_etal2021}{}%
-Stefan, A., Katsimpokis, D., Gronau, Q.
-F., \& Wagenmakers, E.-J.
-(2021).
-Expert agreement in prior elicitation and its effects on {Bayesian} inference.
-\emph{PsyArXiv Preprint}.
-doi:\href{https://doi.org/10.31234/osf.io/8xkqd}{10.31234/osf.io/8xkqd}
 
@@
 \emph{Behavior Research Methods}, \emph{51}(3), 1042--1058.
+doi:\href{https://doi.org/10.3758/s13428-018-01189-8}{10.3758/s13428-018-01189-8}
+
+\leavevmode\hypertarget{ref-Stefan_etal2021}{}%
+Stefan, A.
+M., Katsimpokis, D., Gronau, Q.
+F., \& Wagenmakers, E.-J.
+(2022).
+Expert agreement in prior elicitation and its effects on bayesian inference.
+\emph{Psychonomic Bulletin \& Review}, \emph{29}(5), 1776--1794.
+doi:\href{https://doi.org/10.3758/s13423-022-02074-4}{10.3758/s13423-022-02074-4}
 
@@
 \leavevmode\hypertarget{ref-van2018simple}{}%
-Van Ravenzwaaij, D., Cassey, P., \& Brown, S. D. (2018). A simple introduction to {Markov Chain Monte--Carlo} sampling. \emph{Psychonomic Bulletin \& Review}, \emph{25}(1), 143--154.
+Van Ravenzwaaij, D., Cassey, P., \& Brown, S.
+D.
+(2018).
+A simple introduction to {Markov Chain Monte--Carlo} sampling.
+\emph{Psychonomic Bulletin \& Review}, \emph{25}(1), 143--154.
+doi:\href{https://doi.org/10.3758/s13423-016-1015-8}{10.3758/s13423-016-1015-8}
 
 \leavevmode\hypertarget{ref-vasishth2021sample}{}%
-Vasishth, S., Yadav, H., Schad, D., \& Nicenboim, B. (2021). Sample size determination for bayesian hierarchical models commonly used in psycholinguistics.
+Vasishth, S., Yadav, H., Schad, D.
+J., \& Nicenboim, B.
+(2023).
+Sample size determination for bayesian hierarchical models commonly used in psycholinguistics.
+\emph{Computational Brain \& Behavior}, \emph{6}(1), 102--126.
+doi:\href{https://doi.org/10.1007/s42113-021-00125-y}{10.1007/s42113-021-00125-y}
 
@@
 \emph{Bayesian Analysis}, \emph{16}(2), 667--718.
+doi:\href{https://doi.org/10.1214/20-BA1221}{10.1214/20-BA1221}
+
+\leavevmode\hypertarget{ref-wagenmakers2020principle}{}%
+Wagenmakers, E.-J., Lee, M.
+D., Rouder, J.
+N., \& Morey, R.
+D.
+(2020).
+The principle of predictive irrelevance or why intervals should not be used for model comparison featuring a point null hypothesis.
+\emph{The Theory of Statistics in Psychology: Applications, Use, and Misunderstandings}, 111--129.
+doi:\href{https://doi.org/10.1007/978-3-030-48043-1_8}{10.1007/978-3-030-48043-1\_8}
 
@@
 \emph{Cognitive Psychology}, \emph{60}(3), 158--189.
+doi:\href{https://doi.org/10.1016/j.cogpsych.2009.12.001}{10.1016/j.cogpsych.2009.12.001}
 

### 2023-05-09 manuscript.Rmd
Updates

@@
 ", fig.height=3, fig.width=6, out.width ="95%"}
-knitr::include_graphics("Images/Fig1_meanrtandhierplot_v2.png")
+knitr::include_graphics("I - Images/Fig1_meanrtandhierplot_v2.png")
 ```
@@
 ", fig.align = "left", fig.height = 2, fig.width = 3, out.width="70%"}
-knitr::include_graphics("Images/hierplotefmor.png")
+knitr::include_graphics("I - Images/hierplotefmor.png")
 ```
@@
 ```{r hierarchyplot, echo=FALSE, fig.cap="Hierarchical structure of the normal and log-normal models for the digit classification task.", fig.align = "left", fig.height = 3, fig.width = 4, out.width="70%"}
-knitr::include_graphics("Images/Fig2_Hierarchical structure.png")
+knitr::include_graphics("I - Images/Fig2_Hierarchical structure.png")
 ```
@@
 For all three \\textit{t}-distributions, the degrees of freedom are set to 3.", fig.align = "left", fig.height = 8, fig.width = 12, out.width="100%"}
-knitr::include_graphics("Images/Fig4_distrtypefigure_v3.png")
+knitr::include_graphics("I - Images/Fig4_distrtypefigure_v3.png")
 ```
@@
 Panel D shows the prior predictions for the log-normal model based on the selected priors.", fig.align = "left", fig.height = 3, fig.width = 4, out.width="100%"}
-knitr::include_graphics("Images/Fig5_priorpredfigure_v2.png")
+knitr::include_graphics("I - Images/Fig5_priorpredfigure_v2.png")
 ```
@@
 ^[Conjugate priors result in posterior distributions that belong to the same distribution family as the prior distribution [@Fink1997].
-For a normal likelihood, these are the normal distribution for the mean and the inverse gamma distribution for the variance.]\Revision{However, in `rstan`, non-conjugate priors are commonly applied [e.g., @gelman2006prior]} Therefore, we choose an inverse gamma distribution.
+For a normal likelihood, these are the normal distribution for the mean and the inverse gamma distribution for the variance.] ^[\Revision{However, in rstan, non-conjugate priors are commonly applied} [e.g., @gelman2006prior].] Therefore, we choose an inverse gamma distribution.
 We can also be more specific about the expected RTs.
@@
 As before, we prefer to apply conjugate priors.
-\Revision{The conjugate priors that we applied to the normal model are also conjugate priors for the log-normal model [@fink1997compendium].
-For instance, with a normal distribution as prior for the mean of the log-normal model, we assume that the mean of the parameter can be either negative, zero, or positive, and assign a higher probability to values closer to zero.
+\Revision{The conjugate priors that we applied to the normal model are also conjugate priors for the log-normal model} [@fink1997compendium].
+\Revision{For instance, with a normal distribution as prior for the mean of the log-normal model, we assume that the mean of the parameter can be either negative, zero, or positive, and assign a higher probability to values closer to zero.
 Applying this conjugate prior, the posterior of the mean in the log-normal model will result in a normal distribution as well.}
@@
 Panel B shows the prior prediction based on the selected prior for the log-normal model.", fig.align = "left", fig.height = 3, fig.width = 4, out.width="100%", eval = FALSE}
-knitr::include_graphics("Images/priorpredfigure_lognormal_v2.png")
+knitr::include_graphics("I - Images/priorpredfigure_lognormal_v2.png")
 ```
@@
 The stars represent the number of effective samples of the general effects."}
-knitr::include_graphics("Images/Fig8_traceplots3pnew_v2.png")
+knitr::include_graphics("I - Images/Fig8_traceplots3pnew_v2.png")
 ```
@@
 Although we tried to make the models as similar as possible, the models are not identical.
-This is partly due to difference in prior specification (i.e., specifying prior on the variances versus standard deviations).} For `rstan`, four chains are run with 4000 iterations per chain, of which 1000 serve as warm-up.
+This is partly due to differences in prior specification (i.e., specifying prior on the variances versus standard deviations).} For `rstan`, four chains are run with 4000 iterations per chain, of which 1000 serve as warm-up.
 This results in a total of 12000 samples from the joint posterior distribution.
@@
 The shaded area within the distributions represent 95\\% credible intervals.", fig.align = "left", fig.height = 11, fig.width = 1, out.width="100%"}
-knitr::include_graphics("Images/Fig9_posteriormu_v3.png")
+knitr::include_graphics("I - Images/Fig9_posteriormu_v3.png")
 ```
@@
 Pink intervals contain zero, blue interval do not contain zero.", fig.align = "left", fig.height = 1, fig.width = 0.5, out.width="100%"}
-knitr::include_graphics("Images/Fig10_randompar_v2.png")
+knitr::include_graphics("I - Images/Fig10_randompar_v2.png")
 
@@
 ", fig.align = "left", fig.height = 11, fig.width = 1, out.width="75%"}
-knitr::include_graphics("Images/Fig11_modelestimatefigure2_onlyrstan_v2.png")
+knitr::include_graphics("I - Images/Fig11_modelestimatefigure2_onlyrstan_v2.png")
 ```
@@
 ```{r traceplotlog, echo=FALSE, fig.cap="(ref:captiontraceplotlog)", fig.align = "left", fig.height = 11, fig.width = 4, out.width="100%"}
-knitr::include_graphics("Images/Fig12_traceplots3plog_v2.png")
+knitr::include_graphics("I - Images/Fig12_traceplots3plog_v2.png")
 ```
@@
 The shaded area within the distributions represent 95\\% credible intervals.", fig.align = "left", fig.height = 11, fig.width = 1, out.width="100%"}
-knitr::include_graphics("Images/Fig13_posteriormulog_v2.png")
+knitr::include_graphics("I - Images/Fig13_posteriormulog_v2.png")
 ```
@@
 Pink intervals contain zero, blue interval do not contain zero.", fig.align = "left", fig.height = 1, fig.width = 0.5, out.width="100%"}
-knitr::include_graphics("Images/Fig14_randomparlog_v2.png")
+knitr::include_graphics("I - Images/Fig14_randomparlog_v2.png")
 ```
@@
 The violin plots on the right side show the variance of the individual parameter estimates.", fig.align = "left", fig.height = 11, fig.width = 1, out.width="75%"}
-knitr::include_graphics("Images/Fig15_inddiflogplot_onlyrstan_v2.png")
+knitr::include_graphics("I - Images/Fig15_inddiflogplot_onlyrstan_v2.png")
 ```
@@
 Bayes factors have a few very convenient characteristics.
-One of them is that Bayes factors are transitive, that is, evidence for the full model over the null model can be obtained by flipping numerator and denominatior in equation 13.
+One of them is that Bayes factors are transitive, that is, evidence for the full model over the null model can be obtained by flipping numerator and denominator in equation 13.
 For a full overview of the advantages of Bayesian inference with Bayes factor, see Wagenmakers et al.
@@
 This means that the data are 100 times more likely under the normal model than under the log-normal model, clear evidence in favor of the normal model.
-\textcolor{blue}{But how do we know whether a Bayes factor of 100-to-1, or 10-to-1 or 3-to-1 fulfills?}.
+\Revision{But how do we know whether a Bayes factor of 100-to-1, or 10-to-1 or 3-to-1 fulfills?}.
 As a reminder, Bayes factors are the relative predictive accuracy of two models for the observed data.
@@
 A Bayes factor above one indicates that the evidence is in favor of the null hypothesis.", fig.align = "left", fig.height = 6, fig.width = 4, out.width="60%"}
-knitr::include_graphics("Images/Fig16_sdrplotside_v2.png")
+knitr::include_graphics("I - Images/Fig16_sdrplotside_v2.png")
 ```
@@
 ", fig.align = "left", fig.height = 11, fig.width = 4, out.width="70%"}
-knitr::include_graphics("Images/Fig17_normalbfplot01.png")
+knitr::include_graphics("I - Images/Fig17_normalbfplot01.png")
 ```
@@
 
-### Log-normal Model
-We can take a similar approach for the log-normal model to compare the hierarchical models on varying complexity.
+### \Revision{Log-normal Model}
+\Revision{We can take a similar approach for the log-normal model to compare the hierarchical models on varying complexity.
 We evaluate their performance in two scenarios that correspond to scenario 1 and 3 for the normal model.
@@
 For scenario 2, we remove individual variability from the side, digit, and full model.
-This resulted in the following posterior model probabilities shown in Table \@ref(tab:postmodproblog).
+This resulted in the following posterior model probabilities shown in Table} \@ref(tab:postmodproblog).
 
@@
 
-The table shows that the digit model without individual digit effects from scenario 2 has the highest posterior model probability.
+\Revision{The table shows that the digit model without individual digit effects from scenario 2 has the highest posterior model probability.
 This is in line with our findings for the normal model.
@@
 The log marginal likelihood for the normal digit model without individual differences was $6441.04$ while the log marginal likelihood for the log-normal digit model without individual differences was $10637.45$.
-This indicates that the Bayes factor is in favor of the log-normal digit model over the normal digit model.
-
-This example illustrates the benefits of comparing hierarchical models of varying complexity.
+This indicates that the Bayes factor is in favor of the log-normal digit model over the normal digit model.}
+
+The examples illustrate the benefits of comparing hierarchical models of varying complexity.
 While individual differences are a common phenomenon in psychology, there are instances where general effects are sufficient to describe the data adequately, or where only a random intercept, but no random slopes are necessary.
-As shown in Tables \@ref(tab:postmodprob) and \@ref(tab:postmodproblog), Bayesian model comparisons can help researchers to identify the model that best describes the data.
+As shown in Tables} \@ref(tab:postmodprob) and \@ref(tab:postmodproblog), Bayesian model comparisons can help researchers to identify the model that best describes the data.
 
@@
 For the non-technical user, we recommend using `brms` because of its intuitive model specification.
-However, the default prior setup can be unsuitable for a specific research context and (virtually) always needs to be adapted.
+However, the default prior setup can be unsuitable for a specific research context and \Revision{(virtually) always needs} to be adapted.
 Additionally, users are limited to predefined parametrizations.

### 2023-05-09 manuscript.tex
Updates

@@
 
-\includegraphics[width=0.95\linewidth]{Images/Fig1_meanrtandhierplot_v2} \hfill{}
+\includegraphics[width=0.95\linewidth]{I - Images/Fig1_meanrtandhierplot_v2} \hfill{}
 
@@
 
-\includegraphics[width=0.7\linewidth]{Images/Fig2_Hierarchical structure} \hfill{}
+\includegraphics[width=0.7\linewidth]{I - Images/Fig2_Hierarchical structure} \hfill{}
 
@@
 
-\includegraphics[width=1\linewidth]{Images/Fig4_distrtypefigure_v3} \hfill{}
+\includegraphics[width=1\linewidth]{I - Images/Fig4_distrtypefigure_v3} \hfill{}
 
@@
 
-\includegraphics[width=1\linewidth]{Images/Fig5_priorpredfigure_v2} \hfill{}
+\includegraphics[width=1\linewidth]{I - Images/Fig5_priorpredfigure_v2} \hfill{}
 
@@
 Notably, this is not the same as calculating $\sqrt{(f(x))}$ or $f(x)^2$.
-To transform the priors on standard deviations to variances, we try to find the parameters that match the prior distribution to the implied distribution.} In our \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/tree/main/R\%20code/Prior\%20transformation}{Online Supplement}, we provide R-code with a numeric solution to the transformation.
+To transform the priors on standard deviations to variances, we try to find the parameters that match the prior distribution to the implied distribution.} In \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/tree/main/R\%20code/Prior\%20transformation}{Online Supplement J}, we provide R-code with a numeric solution to the transformation.
 For the \texttt{brms} priors, we obtain the following implied prior distributions on variances:
@@
 The results of the prior prediction for the symbolic distance effect with the priors from Equation \eqref{eq:chosenpriors1} with the number of participants \(n = 10\), the number of trials per condition \(k = 60\), and the number of simulations \(m = 1000\) are shown in panel A of Figure \ref{fig:priorpredfigurepanel}.
-Code for the simulations can be found in our \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment}{Online Supplement}.
+Code for the simulations can be found in \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/blob/main/R\%20code/Figures/Fig5_priorpredictions_trunc_v3.R}{Online Supplement J}.
 
@@
 As RTs cannot be negative, we adjust the prior to be positive only by truncating the normal distribution using a lower bound of zero.
-Finally, we would like to use a more common conjugate prior for the variances to ensure obtaining a proper posterior distribution.\footnote{Conjugate priors result in posterior distributions that belong to the same distribution family as the prior distribution (Fink, 1997).
-For a normal likelihood, these are the normal distribution for the mean and the inverse gamma distribution for the variance.}\textcolor{blue}{However, in `rstan`, non-conjugate priors are commonly applied [e.g., @gelman2006prior]} Therefore, we choose an inverse gamma distribution.
+Finally, we would like to use a more common conjugate prior for the variances to ensure obtaining a proper posterior distribution.\footnote{Conjugate priors result in posterior distributions that belong to the same distribution family as the prior distribution (Fink, 1997a).
+For a normal likelihood, these are the normal distribution for the mean and the inverse gamma distribution for the variance.} \footnote{\textcolor{blue}{However, in `rstan`, non-conjugate priors are commonly applied} (e.g., Gelman, 2006).} Therefore, we choose an inverse gamma distribution.
 We can also be more specific about the expected RTs.
@@
 As before, we prefer to apply conjugate priors.
-\textcolor{blue}{The conjugate priors that we applied to the normal model are also conjugate priors for the log-normal model [@fink1997compendium].
-For instance, with a normal distribution as prior for the mean of the log-normal model, we assume that the mean of the parameter can be either negative, zero, or positive, and assign a higher probability to values closer to zero.
+\textcolor{blue}{The conjugate priors that we applied to the normal model are also conjugate priors for the log-normal model} (Fink, 1997b).
+\textcolor{blue}{For instance, with a normal distribution as prior for the mean of the log-normal model, we assume that the mean of the parameter can be either negative, zero, or positive, and assign a higher probability to values closer to zero.
 Applying this conjugate prior, the posterior of the mean in the log-normal model will result in a normal distribution as well.}
@@
 In this section, we explain how a model can be fitted in each package using code snippets.
-In addition, we offer a separate R Markdown file for each package in our \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment}{Online Supplement} that explains in more depth how to analyze the full symbolic distance effect model with the specific package.
+In addition, we offer separate R Markdown files for each package in \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment}{Online Supplements C-F} that explains in more depth how to analyze the full symbolic distance effect model with the specific package.
 
@@
 An example \texttt{.stan} file of a simple, non-hierarchical normal model, where the variance is already known and the mean has to be estimated, is shown below (adapted example from Nicenboim, Schad, \& Vasishth, 2021, Chapter 10).
-The \texttt{.stan} file for the symbolic distance effect, including explanations, can be found in the \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/blob/main/R\%20objects/rstan/normal\%20model/myModel.stan}{Online Supplement}.
+The \texttt{.stan} file for the symbolic distance effect, including explanations, can be found in \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/blob/main/R\%20objects/rstan/normal\%20model/myModel.stan}{Online Supplement K}.
 The file is divided into three sections: \texttt{data}, \texttt{parameters}, and \texttt{model}.
@@
 
-\includegraphics[width=1\linewidth]{Images/Fig8_traceplots3pnew_v2} \hfill{}
+\includegraphics[width=1\linewidth]{I - Images/Fig8_traceplots3pnew_v2} \hfill{}
 
@@
 Although we tried to make the models as similar as possible, the models are not identical.
-This is partly due to difference in prior specification (i.e., specifying prior on the variances versus standard deviations).} For \texttt{rstan}, four chains are run with 4000 iterations per chain, of which 1000 serve as warm-up.
+This is partly due to differences in prior specification (i.e., specifying prior on the variances versus standard deviations).} For \texttt{rstan}, four chains are run with 4000 iterations per chain, of which 1000 serve as warm-up.
 This results in a total of 12000 samples from the joint posterior distribution.
@@
 
-\includegraphics[width=1\linewidth]{Images/Fig9_posteriormu_v3} \hfill{}
+\includegraphics[width=1\linewidth]{I - Images/Fig9_posteriormu_v3} \hfill{}
 
@@
 
-\includegraphics[width=1\linewidth]{Images/Fig10_randompar_v2} \hfill{}
+\includegraphics[width=1\linewidth]{I - Images/Fig10_randompar_v2} \hfill{}
 
@@
 
-\includegraphics[width=0.75\linewidth]{Images/Fig11_modelestimatefigure2_onlyrstan_v2} \hfill{}
+\includegraphics[width=0.75\linewidth]{I - Images/Fig11_modelestimatefigure2_onlyrstan_v2} \hfill{}
 
@@
 
-\includegraphics[width=1\linewidth]{Images/Fig12_traceplots3plog_v2} \hfill{}
+\includegraphics[width=1\linewidth]{I - Images/Fig12_traceplots3plog_v2} \hfill{}
 
@@
 
-\includegraphics[width=1\linewidth]{Images/Fig13_posteriormulog_v2} \hfill{}
+\includegraphics[width=1\linewidth]{I - Images/Fig13_posteriormulog_v2} \hfill{}
 
@@
 
-\includegraphics[width=1\linewidth]{Images/Fig14_randomparlog_v2} \hfill{}
+\includegraphics[width=1\linewidth]{I - Images/Fig14_randomparlog_v2} \hfill{}
 
@@
 
-\includegraphics[width=0.75\linewidth]{Images/Fig15_inddiflogplot_onlyrstan_v2} \hfill{}
+\includegraphics[width=0.75\linewidth]{I - Images/Fig15_inddiflogplot_onlyrstan_v2} \hfill{}
 
@@
 
-\includegraphics[width=0.6\linewidth]{Images/Fig16_sdrplotside_v2} \hfill{}
+\includegraphics[width=0.6\linewidth]{I - Images/Fig16_sdrplotside_v2} \hfill{}
 
@@
 
-\includegraphics[width=0.7\linewidth]{Images/Fig17_normalbfplot01} \hfill{}
+\includegraphics[width=0.7\linewidth]{I - Images/Fig17_normalbfplot01} \hfill{}
 
@@
 
-\hypertarget{log-normal-model-3}{%
-\subsubsection{Log-normal Model}\label{log-normal-model-3}}
-
-We can take a similar approach for the log-normal model to compare the hierarchical models on varying complexity.
+\hypertarget{section}{%
+\subsubsection{\texorpdfstring{\textcolor{blue}{Log-normal Model}}{}}\label{section}}
+
+\textcolor{blue}{We can take a similar approach for the log-normal model to compare the hierarchical models on varying complexity.
 We evaluate their performance in two scenarios that correspond to scenario 1 and 3 for the normal model.
@@
 For scenario 2, we remove individual variability from the side, digit, and full model.
-This resulted in the following posterior model probabilities shown in Table \ref{tab:postmodproblog}.
+This resulted in the following posterior model probabilities shown in Table} \ref{tab:postmodproblog}.
 
@@
 
-The table shows that the digit model without individual digit effects from scenario 2 has the highest posterior model probability.
+\textcolor{blue}{The table shows that the digit model without individual digit effects from scenario 2 has the highest posterior model probability.
 This is in line with our findings for the normal model.
@@
 The normal digit model from scenario 3 to the log-normal model from scenario 2.
-The log marginal likelihood for the normal digit model without individual differences was \(6441.04\) while the log marginal likelihood for the log-normal digit model without individual differences was \(10637.45\).
-This indicates that the Bayes factor is in favor of the log-normal digit model over the normal digit model.
-
-This example illustrates the benefits of comparing hierarchical models of varying complexity.
+The log marginal likelihood for the normal digit model without individual differences was $6441.04$ while the log marginal likelihood for the log-normal digit model without individual differences was $10637.45$.
+This indicates that the Bayes factor is in favor of the log-normal digit model over the normal digit model.}
+
+The examples illustrate the benefits of comparing hierarchical models of varying complexity.
 While individual differences are a common phenomenon in psychology, there are instances where general effects are sufficient to describe the data adequately, or where only a random intercept, but no random slopes are necessary.
-As shown in Tables \ref{tab:postmodprob} and \ref{tab:postmodproblog}, Bayesian model comparisons can help researchers to identify the model that best describes the data.
+As shown in Tables\} \ref{tab:postmodprob} and \ref{tab:postmodproblog}, Bayesian model comparisons can help researchers to identify the model that best describes the data.
 
@@
 For the non-technical user, we recommend using \texttt{brms} because of its intuitive model specification.
-However, the default prior setup can be unsuitable for a specific research context and (virtually) always needs to be adapted.
+However, the default prior setup can be unsuitable for a specific research context and \textcolor{blue}{(virtually) always needs} to be adapted.
 Additionally, users are limited to predefined parametrizations.
@@
 
-\hypertarget{section}{%
-\subsection{\texorpdfstring{\textcolor{blue}{Technical Possibilities are Endless}}{}}\label{section}}
+\hypertarget{section-1}{%
+\subsection{\texorpdfstring{\textcolor{blue}{Technical Possibilities are Endless}}{}}\label{section-1}}
 
@@
 \leavevmode\hypertarget{ref-Fink1997}{}%
-Fink, D. (1997). A compendium of conjugate priors. https://www.johndcook.com/CompendiumOfConjugatePriors.pdf.
+Fink, D. (1997a). A compendium of conjugate priors. https://www.johndcook.com/CompendiumOfConjugatePriors.pdf.
+
+\leavevmode\hypertarget{ref-fink1997compendium}{}%
+Fink, D. (1997b). A compendium of conjugate priors.
 

### 2023-05-10 manuscript.Rmd
Updates

@@
 For example, participants in a digit classification task repeatedly respond to target digits inducing a correlation between the responses to the target digits, such as between target digit 3 and target digit 4.
-\Revision{Therefore, analyses on aggregated data do not appropriately account for the interdependence between observations [@van2021bayes; @van2023bayes].} As a result, the type I error rate of a test of the effect increases: there is a higher probability of concluding that there is an effect if there is no effect (Rouder & Lu, 2005).
+\Revision{Therefore, analyses on aggregated data do not appropriately account for the interdependence between observations} [@van2021bayes; @van2023bayes].
+As a result, the type I error rate of a test of the effect increases: there is a higher probability of concluding that there is an effect if there is no effect (Rouder & Lu, 2005).
 
@@
 When using the normal model for response times, we are assigning probability to negative values that are not possible.
-This seems very problematic.} In addition, RT distributions are right-skewed.
+This could be problematic.} In addition, RT distributions are right-skewed.
 A better representation of the data at hand is therefore given by a log-normal model [@schramm2019reaction].
@@
 The figure highlights the relationship between the parameters and the data.
+
+\Revision{Up until now, we specified models without a correlation between the individual effects.
+This means, for instance, that we do not expect that the estimations for the individual effects $\delta_{7, i}$ and $\delta_{3, i}$ to be related.
+However, in cases where we would expect the estimation of individual effects to be related, we could estimate the correlation coefficient between individual effects.
+This would add complexity to the model.
+For the sake of the tutorial, we will focus on the models without individual effect correlation, but we illustrate the estimation of the model with correlation between the individual effects in} [Online Supplement C and D](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment).
 
@@
 
-Another important issue is that wide priors are often improper probability distributions [@hobert1996effect].
+Another important issue is that wide priors are often \Revision{improper probability distributions} [@hobert1996effect].
 \Revision{A prior is proper if two conditions hold: 1.
@@
 the probability distribution sums to one for discrete data or integrates to one for continuous data.
-An example of a proper and improper probability distribution are provided in Figure \@ref(fig:properfigure).
-The figure shows two hypothetical priors for parameter $\theta$.
+An example of a proper and improper probability distribution are provided in Figure} \@ref(fig:properfigure).
+\Revision{The figure shows two hypothetical priors for parameter $\theta$.
 The first condition can be assessed by checking the y-values---both priors have only positive definite values as function output (0.1, 0.2, and 0.4).
@@
 ## Step 3: Prior Predictions 
-A key advantage of model specification with priors is that we can obtain predictions on what the data would look like that are generated from the model \Revision{[@etz2018bayesian; @lee2018determining]}.
+A key advantage of model specification with priors is that we can obtain predictions on what the data would look like that are generated from the model [@etz2018bayesian; @lee2018determining].
 These prior predictions allow us to make the implications of our specified priors more concrete.
@@
 The model is fitted using the packages `rstan` and `brms`.
-\Revision{We used different settings for `rstan` and `brms` due to the difference between the models for these packages.
+\Revision{We used different settings for} `rstan` \Revision{and} `brms` \Revision{due to the difference between the models for these packages.
 Although we tried to make the models as similar as possible, the models are not identical.
@@
 Bulk-ESS evaluates the center of the posterior distribution, while Tail-ESS evaluates the tails of the posterior distribution.
-For this tutorial, we focus on the basic calculation of the number of effective samples suggested by Gelman et al.
-(2013) and improved by Vehtari et al.
-(2021).}
+For this tutorial, we focus on the basic calculation of the number of effective samples suggested by Gelman et al.} [-@gelman2013bayesian] \Revision{and improved by Vehtari et al.} [-@vehtari2019rank].
 
@@
 Here, this was not case, and all other diagnostics were good, so we continue with the interpretation of the estimates.
-\Revision{The difference between the model diagnostics of `rstan` and `brms` can be partly explained by the difference in model parametrization, such as the difference in the prior specification (i.e., specifying priors on variances versus priors on standard deviations).
+\Revision{The difference between the model diagnostics of} `rstan` \Revision{and} `brms` \Revision{can be partly explained by the difference in model parametrization, such as the difference in the prior specification (i.e., specifying priors on variances versus priors on standard deviations).
 Although we tried to make the models as similar as possible, they are not identical.}
@@
 This means that the data are 100 times more likely under the normal model than under the log-normal model, clear evidence in favor of the normal model.
-\Revision{But how do we know whether a Bayes factor of 100-to-1, or 10-to-1 or 3-to-1 fulfills?}.
-As a reminder, Bayes factors are the relative predictive accuracy of two models for the observed data.
+\Revision{But how do we know whether a Bayes factor of 100-to-1, or 10-to-1 or 3-to-1 fulfills?} As a reminder, Bayes factors are the relative predictive accuracy of two models for the observed data.
 Therefore, Bayes factors are odds or ratios.
@@
 In the following section we will present two methods to obtain Bayes factors for the symbolic distance effect.
+\Revision{The first method that we will discuss, Savage-Dickey density ratio, evaluates the general effects with the individual effects in the model, while the second method, bridge-sampling, evaluates the full model, the general and individual effects are considered together.
+We want to make you aware that there have been discussions on which method should be preferred without consensus} [@van2021bayes; @van2023bayes; @rouder2023principles; @singmann2023statistics].
+\Revision{Therefore, we will illustrate the use of the two approaches and discuss their potential.}
 
@@
 
-\Revision{Note that models models 2-4 are nested in the most complex full model, and the null model is nested in models 2 and 3, but models 2 and 3 are not nested}.
+\Revision{Note that models 2-4 are nested in the most complex full model, and the null model is nested in models 2 and 3, but models 2 and 3 are not nested}.
 Additionally, the models differ by several parameters.
@@
 It is important to note that this difference is not due to the bridge sampling function, but a result of a difference in parametrization.
-\Revision{The models for `rstan` and `brms` are similar but not identical, partly due to the difference in prior specification (i.e., variances versus standard deviations).} However, for both packages, the conclusion is the same: the data are least likely to have occurred under the full model.
+\Revision{The models for} `rstan` \Revision{and} `brms` \Revision{are similar but not identical, partly due to the difference in prior specification (i.e., variances versus standard deviations).} However, for both packages, the conclusion is the same: the data are least likely to have occurred under the full model.
 In case the estimates are unstable, it is recommended to further increase the number of posterior samples.
@@
 Vasishth [-@schad2021toward] and Schad, Nicenboim, Bürkner, Betancourt, and Vasishth [-@schad2021workflow].
-\Revision{There are also other R software packages for Bayesian multilevel modeling besides `brms` and `rstan`, such as `BayesFactor`[@R-BayesFactor] and `rstanarm`[@rstanarm2020], that we did not illustrate in this tutorial.
+\Revision{There are also other R software packages for Bayesian multilevel modeling besides} `brms` \Revision{and} `rstan`\Revision{, such as} `BayesFactor`[@R-BayesFactor] \Revision{and} `rstanarm`[@rstanarm2020]\Revision{, that we did not illustrate in this tutorial.
 However, most of these other R software packages are a bit more limited, for instance, in model specification and prior settings.}
@@
 \Revision{Bayes factor estimates can also be biased.
-However, bridge sampling has shown to be relatively unbiased [@sarafoglou2021evaluating].} Yet, bridge sampling may also fail in some cases.
+However, bridge sampling has shown to be relatively unbiased} [@sarafoglou2021evaluating].
+Yet, bridge sampling may also fail in some cases.
 Failures of the bridge sampling algorithm are identified by unstable Bayes factors across repeated sampling runs.

### 2023-05-10 manuscript.tex
Updates

@@
 For example, participants in a digit classification task repeatedly respond to target digits inducing a correlation between the responses to the target digits, such as between target digit 3 and target digit 4.
-\textcolor{blue}{Therefore, analyses on aggregated data do not appropriately account for the interdependence between observations [@van2021bayes; @van2023bayes].} As a result, the type I error rate of a test of the effect increases: there is a higher probability of concluding that there is an effect if there is no effect (Rouder \& Lu, 2005).
+\textcolor{blue}{Therefore, analyses on aggregated data do not appropriately account for the interdependence between observations} (Doorn, Aust, Haaf, Stefan, \& Wagenmakers, 2021, 2023).
+As a result, the type I error rate of a test of the effect increases: there is a higher probability of concluding that there is an effect if there is no effect (Rouder \& Lu, 2005).
 
@@
 When using the normal model for response times, we are assigning probability to negative values that are not possible.
-This seems very problematic.} In addition, RT distributions are right-skewed.
+This could be problematic.} In addition, RT distributions are right-skewed.
 A better representation of the data at hand is therefore given by a log-normal model (Schramm \& Rouder, 2019).
@@
 
+\textcolor{blue}{Up until now, we specified models without a correlation between the individual effects.
+This means, for instance, that we do not expect that the estimations for the individual effects $\delta_{7, i}$ and $\delta_{3, i}$ to be related.
+However, in cases where we would expect the estimation of individual effects to be related, we could estimate the correlation coefficient between individual effects.
+This would add complexity to the model.
+For the sake of the tutorial, we will focus on the models without individual effect correlation, but we illustrate the estimation of the model with correlation between the individual effects in} \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment}{Online Supplement C and D}.
+
 In Bayesian modeling, prior distributions are needed to complete model specification.
@@
 
-Another important issue is that wide priors are often improper probability distributions (Hobert \& Casella, 1996).
+Another important issue is that wide priors are often \textcolor{blue}{improper probability distributions} (Hobert \& Casella, 1996).
 \textcolor{blue}{A prior is proper if two conditions hold: 1.
@@
 the probability distribution sums to one for discrete data or integrates to one for continuous data.
-An example of a proper and improper probability distribution are provided in Figure \\ref{fig:properfigure}.
-The figure shows two hypothetical priors for parameter $\theta$.
+An example of a proper and improper probability distribution are provided in Figure} \ref{fig:properfigure}.
+\textcolor{blue}{The figure shows two hypothetical priors for parameter $\theta$.
 The first condition can be assessed by checking the y-values---both priors have only positive definite values as function output (0.1, 0.2, and 0.4).
@@
 
-A key advantage of model specification with priors is that we can obtain predictions on what the data would look like that are generated from the model \textcolor{blue}{[@etz2018bayesian; @lee2018determining]}.
+A key advantage of model specification with priors is that we can obtain predictions on what the data would look like that are generated from the model (Etz, Haaf, Rouder, \& Vandekerckhove, 2018; Lee \& Vanpaemel, 2018).
 These prior predictions allow us to make the implications of our specified priors more concrete.
@@
 Finally, we would like to use a more common conjugate prior for the variances to ensure obtaining a proper posterior distribution.\footnote{Conjugate priors result in posterior distributions that belong to the same distribution family as the prior distribution (Fink, 1997a).
-For a normal likelihood, these are the normal distribution for the mean and the inverse gamma distribution for the variance.} \footnote{\textcolor{blue}{However, in `rstan`, non-conjugate priors are commonly applied} (e.g., Gelman, 2006).} Therefore, we choose an inverse gamma distribution.
+For a normal likelihood, these are the normal distribution for the mean and the inverse gamma distribution for the variance.} \footnote{\textcolor{blue}{However, in rstan, non-conjugate priors are commonly applied} (e.g., Gelman, 2006).} Therefore, we choose an inverse gamma distribution.
 We can also be more specific about the expected RTs.
@@
 The model is fitted using the packages \texttt{rstan} and \texttt{brms}.
-\textcolor{blue}{We used different settings for `rstan` and `brms` due to the difference between the models for these packages.
+\textcolor{blue}{We used different settings for} \texttt{rstan} \textcolor{blue}{and} \texttt{brms} \textcolor{blue}{due to the difference between the models for these packages.
 Although we tried to make the models as similar as possible, the models are not identical.
@@
 Bulk-ESS evaluates the center of the posterior distribution, while Tail-ESS evaluates the tails of the posterior distribution.
-For this tutorial, we focus on the basic calculation of the number of effective samples suggested by Gelman et al.
-(2013) and improved by Vehtari et al.
-(2021).}
+For this tutorial, we focus on the basic calculation of the number of effective samples suggested by Gelman et al.} (2013) \textcolor{blue}{and improved by Vehtari et al.} (2021b).
 
@@
 Here, this was not case, and all other diagnostics were good, so we continue with the interpretation of the estimates.
-\textcolor{blue}{The difference between the model diagnostics of `rstan` and `brms` can be partly explained by the difference in model parametrization, such as the difference in the prior specification (i.e., specifying priors on variances versus priors on standard deviations).
+\textcolor{blue}{The difference between the model diagnostics of} \texttt{rstan} \textcolor{blue}{and} \texttt{brms} \textcolor{blue}{can be partly explained by the difference in model parametrization, such as the difference in the prior specification (i.e., specifying priors on variances versus priors on standard deviations).
 Although we tried to make the models as similar as possible, they are not identical.}
@@
 This means that the data are 100 times more likely under the normal model than under the log-normal model, clear evidence in favor of the normal model.
-\textcolor{blue}{But how do we know whether a Bayes factor of 100-to-1, or 10-to-1 or 3-to-1 fulfills?}.
-As a reminder, Bayes factors are the relative predictive accuracy of two models for the observed data.
+\textcolor{blue}{But how do we know whether a Bayes factor of 100-to-1, or 10-to-1 or 3-to-1 fulfills?} As a reminder, Bayes factors are the relative predictive accuracy of two models for the observed data.
 Therefore, Bayes factors are odds or ratios.
@@
 In the following section we will present two methods to obtain Bayes factors for the symbolic distance effect.
+\textcolor{blue}{The first method that we will discuss, Savage-Dickey density ratio, evaluates the general effects with the individual effects in the model, while the second method, bridge-sampling, evaluates the full model, the general and individual effects are considered together.
+We want to make you aware that there have been discussions on which method should be preferred without consensus} (Doorn, Aust, Haaf, Stefan, \& Wagenmakers, 2021, 2023; Rouder, Schnuerch, Haaf, \& Morey, 2023; Singmann et al., 2023).
+\textcolor{blue}{Therefore, we will illustrate the use of the two approaches and discuss their potential.}
 
@@
 
-\textcolor{blue}{Note that models models 2-4 are nested in the most complex full model, and the null model is nested in models 2 and 3, but models 2 and 3 are not nested}.
+\textcolor{blue}{Note that models 2-4 are nested in the most complex full model, and the null model is nested in models 2 and 3, but models 2 and 3 are not nested}.
 Additionally, the models differ by several parameters.
@@
 It is important to note that this difference is not due to the bridge sampling function, but a result of a difference in parametrization.
-\textcolor{blue}{The models for `rstan` and `brms` are similar but not identical, partly due to the difference in prior specification (i.e., variances versus standard deviations).} However, for both packages, the conclusion is the same: the data are least likely to have occurred under the full model.
+\textcolor{blue}{The models for} \texttt{rstan} \textcolor{blue}{and} \texttt{brms} \textcolor{blue}{are similar but not identical, partly due to the difference in prior specification (i.e., variances versus standard deviations).} However, for both packages, the conclusion is the same: the data are least likely to have occurred under the full model.
 In case the estimates are unstable, it is recommended to further increase the number of posterior samples.
@@
 Vasishth (2021) and Schad, Nicenboim, Bürkner, Betancourt, and Vasishth (2022).
-\textcolor{blue}{There are also other R software packages for Bayesian multilevel modeling besides `brms` and `rstan`, such as `BayesFactor`[@R-BayesFactor] and `rstanarm`[@rstanarm2020], that we did not illustrate in this tutorial.
+\textcolor{blue}{There are also other R software packages for Bayesian multilevel modeling besides} \texttt{brms} \textcolor{blue}{and} \texttt{rstan}\textcolor{blue}{, such as} \texttt{BayesFactor}(Morey \& Rouder, 2018) \textcolor{blue}{and} \texttt{rstanarm}(Goodrich, Gabry, Ali, \& Brilleman, 2020)\textcolor{blue}{, that we did not illustrate in this tutorial.
 However, most of these other R software packages are a bit more limited, for instance, in model specification and prior settings.}
@@
 \textcolor{blue}{Bayes factor estimates can also be biased.
-However, bridge sampling has shown to be relatively unbiased [@sarafoglou2021evaluating].} Yet, bridge sampling may also fail in some cases.
+However, bridge sampling has shown to be relatively unbiased} (Sarafoglou et al., 2021).
+Yet, bridge sampling may also fail in some cases.
 Failures of the bridge sampling algorithm are identified by unstable Bayes factors across repeated sampling runs.
@@
 
+\leavevmode\hypertarget{ref-van2023bayes}{}%
+Doorn, J.
+van, Aust, F., Haaf, J.
+M., Stefan, A.
+M., \& Wagenmakers, E.-J.
+(2023).
+Bayes factors for mixed models: Perspective on responses.
+\emph{Computational Brain \& Behavior}, 1--13.
+doi:\href{https://doi.org/10.1007/s42113-022-00158-x}{10.1007/s42113-022-00158-x}
+
 \leavevmode\hypertarget{ref-R-Rcpp_b}{}%
@@
 
+\leavevmode\hypertarget{ref-etz2018bayesian}{}%
+Etz, A., Haaf, J.
+M., Rouder, J.
+N., \& Vandekerckhove, J.
+(2018).
+Bayesian inference and testing any hypothesis you can specify.
+\emph{Advances in Methods and Practices in Psychological Science}, \emph{1}(2), 281--295.
+doi:\href{https://doi.org/10.1177/2515245918773087}{10.1177/2515245918773087}
+
 \leavevmode\hypertarget{ref-etz2018introduction}{}%
@@
 Gelman, A., Vehtari, A., Simpson, D., Margossian, C. C., Carpenter, B., Yao, Y., \ldots{} Modrák, M. (2020). Bayesian workflow. \emph{arXiv Preprint arXiv:2011.01808}.
+
+\leavevmode\hypertarget{ref-rstanarm2020}{}%
+Goodrich, B., Gabry, J., Ali, I., \& Brilleman, S. (2020). Rstanarm: {Bayesian} applied regression modeling via {Stan}. Retrieved from \url{https://mc-stan.org/rstanarm}
 
@@
 doi:\href{https://doi.org/10.1016/j.jmp.2010.08.013}{10.1016/j.jmp.2010.08.013}
+
+\leavevmode\hypertarget{ref-lee2018determining}{}%
+Lee, M.
+D., \& Vanpaemel, W.
+(2018).
+Determining informative priors for cognitive models.
+\emph{Psychonomic Bulletin \& Review}, \emph{25}(1), 114--127.
+doi:\href{https://doi.org/10.3758/s13423-017-1238-3}{10.3758/s13423-017-1238-3}
 
@@
 
+\leavevmode\hypertarget{ref-rouder2023principles}{}%
+Rouder, J.
+N., Schnuerch, M., Haaf, J.
+M., \& Morey, R.
+D.
+(2023).
+Principles of model specification in ANOVA designs.
+\emph{Computational Brain \& Behavior}, \emph{6}(1), 50--63.
+doi:\href{https://doi.org/10.1007/s42113-022-00132-7}{10.1007/s42113-022-00132-7}
+
+\leavevmode\hypertarget{ref-sarafoglou2021evaluating}{}%
+Sarafoglou, A., Haaf, J.
+M., Ly, A., Gronau, Q.
+F., Wagenmakers, E.-J., \& Marsman, M.
+(2021).
+Evaluating multinomial order restrictions with bridge sampling.
+\emph{Psychological Methods}.
+doi:\href{https://doi.org/10.1037/met0000411}{10.1037/met0000411}
+
 \leavevmode\hypertarget{ref-R-lattice}{}%
@@
 doi:\href{https://doi.org/10.4324/9780429318405-2}{10.4324/9780429318405-2}
+
+\leavevmode\hypertarget{ref-singmann2023statistics}{}%
+Singmann, H., Kellen, D., Cox, G.
+E., Chandramouli, S.
+H., Davis-Stober, C.
+P., Dunn, J.
+C., \ldots{} others.
+(2023).
+Statistics in the service of science: Don't let the tail wag the dog.
+\emph{Computational Brain \& Behavior}, \emph{6}(1), 64--83.
+doi:\href{https://doi.org/10.1007/s42113-022-00129-2}{10.1007/s42113-022-00129-2}
 

### 2023-06-23 manuscript.Rmd
Updates based on Angelika's comments

@@
 
+\Revision{It is clear that neither of the alternative approaches yield satisfactory results by themselves.
+Conceptually, a hierarchical analysis provides a compromise between the aggregated and independent approach by drawing benefits from each method.
+The motivation for hierarchical models, as well as their general mathematical structure, has been covered by several tutorial articles} [e.g., @singmann2019introduction; @west2022linear]\Revision{, as well as text books (e.g., XX, XX).}
+
 \Revision{In hierarchical analyses, the general or overall effect is sometimes called fixed effect, and the individual-specific deviations from this effect are in that case referred to as random effects (Rouder \& Lu, 2005).
@@
 
-\Revision{There are several tutorials for readers interested in the frequentist approach of hierarchical model estimation} [e.g., @singmann2019introduction; @west2022linear].
-\Revision{However, it has been argued that} the Bayesian framework offers a more intuitive approach for hierarchical modeling (Rouder & Lu, 2005; Lynch, 2007).
+The Bayesian framework offers a more intuitive approach for hierarchical modeling (Rouder & Lu, 2005; Lynch, 2007).
 The flexibility of the Bayesian approach facilitates the implementation of complex hierarchical models (Rouder, Morey, & Pratte, 2013).
@@
 Note that the only change in the model formulation is the change of the likelihood function from a normal distribution to a log-normal distribution.
-\Revision{The likelihood function provides the likelihood of parameter values in the form of probability densities.
+\Revision{The likelihood function describes the probability model that is assumed to be underlying the data.
+When the parameters of this probability model are known, the joint probability of the data can be inferred from the likelihood function.
 The log-normal distribution only assigns probability to positive data values, and is, just as the RT distributions, right-skewed.
@@
 \Revision{For instance, with a normal distribution as prior for the mean of the log-normal model, we assume that the mean of the parameter can be either negative, zero, or positive, and assign a higher probability to values closer to zero.
-Applying this conjugate prior, the posterior of the mean in the log-normal model will result in a normal distribution as well.}
+If the mean of the log-normal distribution is smaller than zero, this means that the peak of the log-normal distribution moves closer to zero, i.e., that response times are quicker but still larger than zero.
+When applying the normal distribution as prior, the posterior of the mean in the log-normal model will result in a normal distribution as well.}
 As a starting point, we can use the adjusted priors of the normal model and perform a prior prediction, shown in panel C of Figure \@ref(fig:priorpredfigurepanel).
@@
 The credible interval displayed in the figure is the Bayesian version of a confidence interval and provides a measure of uncertainty about the parameter value [@wagenmakers2018bayesian].
-\Revision{In contrast to the confidence interval, the credible interval provides the probability of a range of values for an effect.
-However, considering the prior and under the assumption that there is an effect, the credible interval does not provide evidence to conclude whether there is an effect} [@wagenmakers2020principle].
+\Revision{In contrast to the confidence interval, the credible interval provides the probability that a parameter lies within a certain range.
+However, the credible interval does not provide evidence to conclude whether an effect is zero or not }[@wagenmakers2020principle].
 
@@
 # example - version 1 with SE 
-rstanfithier8 <- readRDS(file = "R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
+rstanfithier8 <- readRDS(file = "K - R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
 fitrstanhiertable <- summary(rstanfithier8, pars = c("mu", "mu2", "mu3", "mu4", "mu5", "mu6"))$summary[,c("mean", "se_mean", "2.5%", "97.5%", "n_eff", "Rhat")]
@@
 # example - version 2 without SE 
-rstanfithier8 <- readRDS(file = "R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
+rstanfithier8 <- readRDS(file = "K - R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
 fitrstanhiertable <- summary(rstanfithier8, pars = c("mu", "mu2", "mu3", "mu4", "mu5", "mu6"))$summary[,c("mean", "2.5%", "97.5%", "n_eff", "Rhat")]
@@
 Commonly, this is misinterpreted as an absence of an effect.
-However, in combination with the mean estimate, we can conclude that the effect is small.}
+However, credible intervals should only be used for parameter estimation.
+In cases where the credible overlaps with zero,  the effect is likely small, but can be non-zero.}
 
@@
 
-\Revision{The individual variation can also be visualized in a different way with focusing on the relation between the individual estimates.} Figure \@ref(fig:modelestimatesfigure3) shows how the relationship between individuals' digit effects.^[The design and code for this figure are by @j_van_langen_2020_3715576.] The posterior means of individual effects are represented by the dots.
+\Revision{The individual variation can also be visualized differently, for example by connecting individual estimates for each parameter.
+This type of variation can reveal correlations between individual parameter estimates, for example, participants scoring high on one digit parameter may also score high on another digit parameter.} Figure \@ref(fig:modelestimatesfigure3) shows the relationship between individuals' digit effects.^[The design and code for this figure are by @j_van_langen_2020_3715576.] The posterior means of individual effects are represented by the dots.
 The right side of each plot represents the distribution of the individual estimates.
@@
 # example - with SE 
-rstanfithier8log <- readRDS(file = "R objects/rstan/log-normal model/hier_logmodelc_adj13012022.rds")
+rstanfithier8log <- readRDS(file = "K - R objects/rstan/log-normal model/hier_logmodelc_adj13012022.rds")
 fitrstanhiertablelog <- summary(rstanfithier8log, pars = c("mu", "mu2", "mu3", "mu4", "mu5", "mu6"))$summary[,c("mean", "se_mean", "2.5%", "97.5%", "n_eff", "Rhat")]
@@
 # example - without SE 
-rstanfithier8log <- readRDS(file = "R objects/rstan/log-normal model/hier_logmodelc_adj13012022.rds")
+rstanfithier8log <- readRDS(file = "K - R objects/rstan/log-normal model/hier_logmodelc_adj13012022.rds")
 fitrstanhiertablelog <- summary(rstanfithier8log, pars = c("mu", "mu2", "mu3", "mu4", "mu5", "mu6"))$summary[,c("mean", "2.5%", "97.5%", "n_eff", "Rhat")]
@@
 In the following section we will present two methods to obtain Bayes factors for the symbolic distance effect.
-\Revision{The first method that we will discuss, Savage-Dickey density ratio, evaluates the general effects with the individual effects in the model, while the second method, bridge-sampling, evaluates the full model, the general and individual effects are considered together.
-We want to make you aware that there have been discussions on which method should be preferred without consensus} [@van2021bayes; @van2023bayes; @rouder2023principles; @singmann2023statistics].
-\Revision{Therefore, we will illustrate the use of the two approaches and discuss their potential.}
+\Revision{The first method that we discuss, Savage-Dickey density ratio, tests the general effect while disregarding individual variation (similar to type III sums of squares in ANOVA models).
+This means that the compared models differ in only a single parameter, that is, the general effect parameter, and individual effects are present in both models.
+The Savage-Dickey density ratio is computationally efficient, but can only be used for tests of a single parameter value as described before.
+This is why we additionally introduce a second approach, bridge-sampling, that is less computationally efficient, but more versatile.
+In this manuscript, we will use bridge sampling to test the general effects and individual variation together, as well as to compare the predictive accuracy of the normal and log-normal model.
+There have been discussions on which method should be preferred without consensus} [@van2021bayes; @van2023bayes; @rouder2023principles; @singmann2023statistics].
+\Revision{Therefore, we illustrate the use of the two approaches and discuss their potential.}
 
@@
 The log marginal likelihood for the normal digit model without individual differences was $6441.04$ while the log marginal likelihood for the log-normal digit model without individual differences was $10637.45$.
-This indicates that the Bayes factor is in favor of the log-normal digit model over the normal digit model.}
+This indicates that the Bayes factor is strongly in favor of the log-normal digit model over the normal digit model.}
 

### 2023-06-23 manuscript.tex
Updates based on Angelika's comments

@@
 
+\textcolor{blue}{It is clear that neither of the alternative approaches yield satisfactory results by themselves.
+Conceptually, a hierarchical analysis provides a compromise between the aggregated and independent approach by drawing benefits from each method.
+The motivation for hierarchical models, as well as their general mathematical structure, has been covered by several tutorial articles} (e.g., Singmann \& Kellen, 2019; West, Welch, \& Galecki, 2022)\textcolor{blue}{, as well as text books (e.g., XX, XX).}
+
 \textcolor{blue}{In hierarchical analyses, the general or overall effect is sometimes called fixed effect, and the individual-specific deviations from this effect are in that case referred to as random effects (Rouder \& Lu, 2005).
@@
 
-\textcolor{blue}{There are several tutorials for readers interested in the frequentist approach of hierarchical model estimation} (e.g., Singmann \& Kellen, 2019; West, Welch, \& Galecki, 2022).
-\textcolor{blue}{However, it has been argued that} the Bayesian framework offers a more intuitive approach for hierarchical modeling (Rouder \& Lu, 2005; Lynch, 2007).
+The Bayesian framework offers a more intuitive approach for hierarchical modeling (Rouder \& Lu, 2005; Lynch, 2007).
 The flexibility of the Bayesian approach facilitates the implementation of complex hierarchical models (Rouder, Morey, \& Pratte, 2013).
@@
 Note that the only change in the model formulation is the change of the likelihood function from a normal distribution to a log-normal distribution.
-\textcolor{blue}{The likelihood function provides the likelihood of parameter values in the form of probability densities.
+\textcolor{blue}{The likelihood function describes the probability model that is assumed to be underlying the data.
+When the parameters of this probability model are known, the joint probability of the data can be inferred from the likelihood function.
 The log-normal distribution only assigns probability to positive data values, and is, just as the RT distributions, right-skewed.
@@
 \textcolor{blue}{For instance, with a normal distribution as prior for the mean of the log-normal model, we assume that the mean of the parameter can be either negative, zero, or positive, and assign a higher probability to values closer to zero.
-Applying this conjugate prior, the posterior of the mean in the log-normal model will result in a normal distribution as well.}
+If the mean of the log-normal distribution is smaller than zero, this means that the peak of the log-normal distribution moves closer to zero, i.e., that response times are quicker but still larger than zero.
+When applying the normal distribution as prior, the posterior of the mean in the log-normal model will result in a normal distribution as well.}
 As a starting point, we can use the adjusted priors of the normal model and perform a prior prediction, shown in panel C of Figure \ref{fig:priorpredfigurepanel}.
@@
 The credible interval displayed in the figure is the Bayesian version of a confidence interval and provides a measure of uncertainty about the parameter value (Wagenmakers et al., 2018).
-\textcolor{blue}{In contrast to the confidence interval, the credible interval provides the probability of a range of values for an effect.
-However, considering the prior and under the assumption that there is an effect, the credible interval does not provide evidence to conclude whether there is an effect} (Wagenmakers, Lee, Rouder, \& Morey, 2020).
+\textcolor{blue}{In contrast to the confidence interval, the credible interval provides the probability that a parameter lies within a certain range.
+However, the credible interval does not provide evidence to conclude whether an effect is zero or not }(Wagenmakers, Lee, Rouder, \& Morey, 2020).
 
@@
 
-\textcolor{blue}{The individual variation can also be visualized in a different way with focusing on the relation between the individual estimates.} Figure \ref{fig:modelestimatesfigure3} shows how the relationship between individuals' digit effects.\footnote{The design and code for this figure are by Langen (2020).} The posterior means of individual effects are represented by the dots.
+\textcolor{blue}{The individual variation can also be visualized differently, for example by connecting individual estimates for each parameter.
+This type of variation can reveal correlations between individual parameter estimates, for example, participants scoring high on one digit parameter may also score high on another digit parameter.} Figure \ref{fig:modelestimatesfigure3} shows the relationship between individuals' digit effects.\footnote{The design and code for this figure are by Langen (2020).} The posterior means of individual effects are represented by the dots.
 The right side of each plot represents the distribution of the individual estimates.
@@
 In the following section we will present two methods to obtain Bayes factors for the symbolic distance effect.
-\textcolor{blue}{The first method that we will discuss, Savage-Dickey density ratio, evaluates the general effects with the individual effects in the model, while the second method, bridge-sampling, evaluates the full model, the general and individual effects are considered together.
-We want to make you aware that there have been discussions on which method should be preferred without consensus} (Doorn, Aust, Haaf, Stefan, \& Wagenmakers, 2021, 2023; Rouder, Schnuerch, Haaf, \& Morey, 2023; Singmann et al., 2023).
-\textcolor{blue}{Therefore, we will illustrate the use of the two approaches and discuss their potential.}
+\textcolor{blue}{The first method that we discuss, Savage-Dickey density ratio, tests the general effect while disregarding individual variation (similar to type III sums of squares in ANOVA models).
+This means that the compared models differ in only a single parameter, that is, the general effect parameter, and individual effects are present in both models.
+The Savage-Dickey density ratio is computationally efficient, but can only be used for tests of a single parameter value as described before.
+This is why we additionally introduce a second approach, bridge-sampling, that is less computationally efficient, but more versatile.
+In this manuscript, we will use bridge sampling to test the general effects and individual variation together, as well as to compare the predictive accuracy of the normal and log-normal model.
+There have been discussions on which method should be preferred without consensus} (Doorn, Aust, Haaf, Stefan, \& Wagenmakers, 2021, 2023; Rouder, Schnuerch, Haaf, \& Morey, 2023; Singmann et al., 2023).
+\textcolor{blue}{Therefore, we illustrate the use of the two approaches and discuss their potential.}
 

### 2023-06-26 propriety_section.Rmd
adding marginal likelihood section

@@
 
-Our models consist of several parameters.
-Therefore, computing the Bayes factor of, for example, the normal model versus the log-normal model becomes somewhat difficult.
-Let the vector of all parameters of the normal model be $\bm{\theta_\text{n}}$ and the vector of all parameters for the log-normal model be $\bm{\theta_{\text{l}}}$.
-With the prior distributions for the two models, $f(\bm{\theta_\text{n}})$ and $f(\bm{\theta_\text{l}})$, the Bayes factor can be calculated by the following ratio of marginal likelihoods:
+# Model Comparison
+
+The symbolic distance hypothesis postulates that response times (RTs) are slower when digits are closer to 5 (digit effects), and that RTs may be influenced by whether the target digit is smaller or greater than 5 (side effect).
+Solely using Bayesian estimation, we are not able to directly test these hypotheses.
+Credible intervals often contain zero, but they always also contain lots of other values that have a chance of being the "true" parameter value, making posterior-based intervals unsuitable to test hypotheses.
+Therefore, our goal for the following section is to set up a Bayesian hypothesis test for the parameters of interest in our model.
+Specifically, we will achieve this by comparing the predictive accuracy of a null model where we set one or some parameters to zero, and and effects models where these parameters are unconstrained.
+\Revision{This approach is called Bayes factor model comparison.}
+
+## Bayes factor
+
+\Revision{Rather than submerging us in technical details or providing a formal justification of the Bayesian approach, we provide an intuition of Bayesian model comparison.
+A formal justification of Bayes factors is provided by @jeffreys1961theory, @kass1995bayes and @rouder2018teaching.
+The Bayes factor (BF) is a measure of how well one model predicts the data compared to another.}
+
+\Revision{Bayesian analysis allows for predictions of data from any model because models are fully specified with a probability distribution for the data and priors for all parameters.
+Let the vector of all parameters of a model be $\bm{\theta}$.
+With the prior distributions these parameters, $f(\bm{\theta})$, the prediction for data from the model can be computed by integrating over all parameters:}
 
 \begin{equation}
-\text{BF}_{\text{nl}} = \frac{P(\bm{Y} | \mathcal{M}_n)}{P(\bm{Y} | \mathcal{M}_{\text{l}})} = \frac{\int_{-\infty}^{\infty}P(\bm{Y} | \mathcal{M}_\text{n}, \bm{\theta_{\text{n}}})f(\bm{\theta_{\text{n}}})d\bm{\theta_{\text{n}}}}{\int_{-\infty}^{\infty}P(\bm{Y} | \mathcal{M}_\text{l}, \bm{\theta_{\text{l}}})f(\bm{\theta_{\text{l}}})d\bm{\theta_{\text{l}}}}.
+p(\bm{Y} | \mathcal{M}) = \int_{\bm{\Theta}} p(\bm{Y} | \mathcal{M}, \bm{\theta})f(\bm{\theta})d\bm{\theta}. 
 \end{equation}
+
+\Revision{Note that $\int_{\bm{\Theta}}$ refers to a multidimensional integral over the entire parameter space.
+As our models consist of many parameters computing these predictions for any one model becomes somewhat difficult.
+Once the integral is computed, the resulting probability distribution of the data, $p(\bm{Y} | \mathcal{M})$ is not a function of the model parameters anymore.
+This marginal probability distribution is therefore also referred to as the marginal likelihood.
+Before observing the data, the marginal likelihood serves as prediction for possible data; once the data are observed, the function tells us how well these specific observations were predicted by the model.
+The Bayes factor then is the ratio of marginal likelihoods of two models, or the relative predictive accuracy of $\mathcal{M}_1$ over $\mathcal{M}_2$:}
+
+\begin{equation}
+\text{BF}_{1,2} = \frac{P(\bm{Y} | \mathcal{M}_1)}{P(\bm{Y} | \mathcal{M}_2)}. 
+\end{equation}
+
+\Revision{There is a second popular interpretation of the Bayes factor as the relative evidence of two competing models.
+One of the key insights from Bayes' rule is that these to things, the relative predictive accuracy of two models for the data, and the relative evidence from the data for the two models, are one and the same [@Rouder:Morey:2018].}
+
+Bayes factors have a few very convenient characteristics.
+One of them is that Bayes factors are transitive, that is, evidence for the full model over the null model can be obtained by flipping numerator and denominator in equation 13.
+For a full overview of the advantages of Bayesian inference with Bayes factor, see Wagenmakers and colleagues [-@wagenmakers2018bayesian].
+
+\newpage
 

### 2023-06-28 manuscript.Rmd
Updates

@@
 To ensure that a broad audience is able to follow the tutorial, we limit and carefully explain equations.
-However, rudimentary knowledge of calculus is advantageous to fully understand Bayesian analysis.
+However, rudimentary knowledge of calculus is advantageous to fully understand the Bayesian analyses.
 }
@@
 Therefore, when this option is applied and individual variability is ignored, the general trend can be an inaccurate representation of the true general trend [Rouder, Morey, & Pratte, 2013; Haaf & Rouder, 2017; van Doorn, Aust, Haaf, Stefan, and Wagenmakers, -@van2021bayes].
-This is shown in Panel A of Figure \@ref(fig:figure1) where some of the individual trends (i.e., slopes of the grey lines) differ from the general trend (i.e., slope of the blue line).
+
+\Revision{A pattern where a general trend is not representing the individual trend is} shown in Panel A of Figure \@ref(fig:figure1) where some of the individual trends (i.e., slopes of the grey lines) differ from the general trend (i.e., slope of the blue line).
 If every participant takes part in each condition, the person variability affects all conditions, thus inducing a correlation across conditions.
 For example, participants in a digit classification task repeatedly respond to target digits inducing a correlation between the responses to the target digits, such as between target digit 3 and target digit 4.
-\Revision{Therefore, analyses on aggregated data do not appropriately account for the interdependence between observations} [@van2021bayes; @van2023bayes].
+\Revision{Therefore, analyses on aggregated data do not appropriately account for differential experiment effects} [@van2021bayes; @van2023bayes].
 As a result, the type I error rate of a test of the effect increases: there is a higher probability of concluding that there is an effect if there is no effect (Rouder & Lu, 2005).
@@
 We can also estimate individual effects, not as deviations from an overarching effect, but as effects for each individual, as obtained from a per-participant analysis.
-Throughout this tutorial, we will use the terms that suit the interpretation of our analyses, namely general effect, individual deviation, and individual effect.}
+Throughout this tutorial, we will use the terms general effect, individual deviation, and individual effect.}
 
@@
 \end{align*}
-\textcolor{blue}{The model equation shows that, for instance, for participant $i = 1$ and digit $j = 7$, with $\gamma_{i = 1}$ equal to $0.5$, $\beta_{i = 1}$ equal to $0.2$, $\delta_{7, i = 1}$ and $\delta_{3, i = 1}$ equal to $0.25$, and $\delta_{6, i = 1}$ and $\delta_{4, i = 1}$ equal to $0.3$, the mean RT is modeled by: $0.5 - 0.1 +  0.25$.
-If all $\delta$s are positive, then the model equation indicates that individuals respond fastest when the digits 2 and 8 are presented, as in this situation there are no digit effects in the equation.
-However, for digit $j = 8$, the mean RT is modeled by $0.5 - 0.1$ in this scenario, while for digit $j = 2$, the mean RT is modeled by $0.5 + 0.1$.}
+
+\Revision{The model equation enables us to compute the expected response time $Y_{ij}$ for a participant $i$ when presented with a specific probe.
+For example, if participant $i = 1$ is presented with digit $j = 7$, $x_j$ takes the value $-\frac{1}{2}$ and $u_j$ takes the value $1$.
+The parameters $v_j$, $w_j$, and $z_j$ take the value zero, such that the equation for the expected value reduces to: $\gamma_{i} + \frac{1}{2} \beta_{i} + \delta_{7,i}$.
+Let us assume that the participant 1's mean response time across all items is $\gamma_{1} = 0.5$, they respond generally faster to digits to the right of five, as expressed by $\beta_{1} = 0.2$, and they respond slower to digit 7, given by $\delta_{7,1} = 0.25$.
+Then their expected response time for the digit 7 can be calculated by $0.5 - 0.5 * 0.2 + 1 * 0.25$.}
 
@@
 \end{equation}
-Note that the only change in the model formulation is the change of the likelihood function from a normal distribution to a log-normal distribution.
-\Revision{The likelihood function describes the probability model that is assumed to be underlying the data.
-When the parameters of this probability model are known, the joint probability of the data can be inferred from the likelihood function.
+Note that the only change in the model formulation is the change from a normal distribution to a log-normal distribution.
+\Revision{The density function describes the probability model that is assumed to be underlying the data.
+When the parameters of this probability model are known, the joint probability of the data can be inferred from the \Revision{density} function.
 The log-normal distribution only assigns probability to positive data values, and is, just as the RT distributions, right-skewed.
@@
 In summary, the models are placed directly on the raw RT data.
-The only difference between the normal and the log-normal models is the likelihood distribution.
+The only difference between the normal and the log-normal models is the probability distribution.
 The normal model just assumes a symmetric normal distribution that can take on any value.
@@
 \Revision{A prior is proper if two conditions hold: 1.
-All values of the probability distribution are positive definite, and 2.
+All values of the probability distribution are equal or greater than zero, and 2.
 the probability distribution sums to one for discrete data or integrates to one for continuous data.
@@
 \Revision{The figure shows two hypothetical priors for parameter $\theta$.
-The first condition can be assessed by checking the y-values---both priors have only positive definite values as function output (0.1, 0.2, and 0.4).
+The first condition can be assessed by checking the y-values---both priors have only positive values as function output (0.1, 0.2, and 0.4).
 To assess the second condition, we need to sum up the function values.
@@
 This is plugged into the individual effect distributions (Equation (2)).
-Then, the likelihood of the model, such as specified in Equations (1) and (3), is used to simulate observations per condition.
+Then, the probability of the model, such as specified in Equations (1) and (3), is used to simulate observations per condition.
 This means that for *n* participants, *n* random values for the model parameters are generated based on the priors, which will result in the *k* observations per *j* conditions.
@@
 All variance terms ($\sigma_{\gamma}^2$, $\sigma_{\beta}^2$, $\sigma_{\delta}^2$, and $\sigma^2$) are constrained to be positive.
-If there is no information from previous studies, researchers might be inclined to use the default priors in `brms` as a starting point.
-
-In `brms`, different default priors are set on different types of parameters [@R-brms_a].
+If there is no information from previous studies, researchers might be inclined to use the default priors \Revision{of software packages, often reasonably uninformative priors,} as a starting point.
+
+\Revision{We take the reasonably uninformative priors as a starting point as well specifying priors on different types of parameters.
+These priors were previously used as} default priors in `brms` [@R-brms_a].
 The prior on the group-level regression weights ($\mu_{\beta}$, $\mu_{\delta}$) is a normal distribution with a mean of 0 and a standard deviation of 1.
@@
 \end{equation}
-The prior distributions `brms` specifies on the standard deviations can be transformed to prior distributions on variances ($var$).
+The prior distributions specified on the standard deviations can be transformed to prior distributions on variances ($var$).
 Prior distributions on variances have a smaller scale than priors on standard deviation ($sd$).
@@
 The blue dots represent the mean effect.
-Using the default prior from `brms`, the mean effects are expected to be around zero, with a mean effect of zero.
+Using the \Revision{reasonably uninformatieve priors}, the mean effects are expected to be around zero, with a mean effect of zero.
 However, the range around this mean is very wide, ranging from -10 to 10 second effects.
@@
 ^[Conjugate priors result in posterior distributions that belong to the same distribution family as the prior distribution [@Fink1997].
-For a normal likelihood, these are the normal distribution for the mean and the inverse gamma distribution for the variance.] ^[\Revision{However, in rstan, non-conjugate priors are commonly applied} [e.g., @gelman2006prior].] Therefore, we choose an inverse gamma distribution.
+For a normal model, these are the normal distribution for the mean and the inverse gamma distribution for the variance.] ^[\Revision{However, in rstan, non-conjugate priors are commonly applied} [e.g., @gelman2006prior].] Therefore, we choose an inverse gamma distribution.
 We can also be more specific about the expected RTs.
@@
 This yields the following priors that result in the prior prediction shown in panel D of Figure \@ref(fig:priorpredfigurepanel):^[For the overall baseline $\mu_\gamma$ we adjusted the prior based on our expectation of the expected overall response times.
-A mean parameter of -0.5 for the log-normal likelihood corresponds to a mean response time of 1.]
+A mean parameter of -0.5 for the log-normal distribution corresponds to a mean response time of 1.]
 
@@
 Every sampling algorithm requires the specification of the number of samples (also called iterations) that are drawn from the posterior distribution.
-Furthermore, all MCMC sampling methods have in common that results depend (somewhat) on chosen starting values, that is, initial parameter values for which the unnormalized posterior (i.e., likelihood times prior) is evaluated.
+Furthermore, all MCMC sampling methods have in common that results depend (somewhat) on chosen starting values, that is, initial parameter values that \Revision{are fed to the sampler in the first iteration}.
 This dependency can be assessed if the sampling algorithm is repeated several times (i.e., chains) with different starting values.
@@
 `Rstan` requires two files: a `.stan` file containing the model specification and an R script containing the code for the model fitting.
-In the `.stan` file, the parameters, priors, and the likelihood are specified.
+In the `.stan` file, the parameters, priors, and the \Revision{probability distribution of the data} are specified.
 An example `.stan` file of a simple, non-hierarchical normal model, where the variance is already known and the mean has to be estimated, is shown below [adapted example from @nicenboim2021introduction, Chapter 10].
-The `.stan` file for the symbolic distance effect, including explanations, can be found in [Online Supplement K](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/blob/main/R%20objects/rstan/normal%20model/myModel.stan).
 The file is divided into three sections: `data`, `parameters`, and `model`.
@@
   target += normal_lpdf(mu | 0, 20);
-  // Likelihood: 
+  // Probability model: 
   for(i in 1:N)
@@
 These are the general and individual effects.
-The last section, `model`, contains the likelihood function and priors, as specified earlier in this article.
+The last section, `model`, contains the probability function and priors, as specified earlier in this article.
+\Revision{Instead of using} `y[i] ~ normal(mu, sigma)` \Revision{to estimate parameters, such as the mean from the example, we used the} `target` \Revision{specification} [for a in-depth explanation see @nicenboim2021introduction, Chapter 10].
+\Revision{Both methods indicate in this case that} `y[i]` \Revision{is normally distributed with the mean} `mu` \Revision{and the variance} `sigma`.
+\Revision{However, the target method saves constant terms that are required for model comparison} [@R-bridgesampling] The `.stan` file for the symbolic distance effect, including explanations, can be found in [Online Supplement K](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/blob/main/R%20objects/rstan/normal%20model/myModel.stan).
 
@@
 The package `rstan` offers a lot of flexibility.
-It is possible to specify the likelihood and priors in the exact way as is wished.
+It is possible to specify the probability model and priors in the exact way as is wished.
 However, this comes with the price of some technical knowledge.
@@
 \Revision{The difference between the model diagnostics of} `rstan` \Revision{and} `brms` \Revision{can be partly explained by the difference in model parametrization, such as the difference in the prior specification (i.e., specifying priors on variances versus priors on standard deviations).
-Although we tried to make the models as similar as possible, they are not identical.}
+We equated the models where sensible which resulted in models that are fairly comparable.
+}
 
@@
 This means that the data are 100 times more likely under the normal model than under the log-normal model, clear evidence in favor of the normal model.
-\Revision{But how do we know whether a Bayes factor of 100-to-1, or 10-to-1 or 3-to-1 fulfills?} As a reminder, Bayes factors are the relative predictive accuracy of two models for the observed data.
+\Revision{But how do we know whether a Bayes factor of 100-to-1, or 10-to-1 or 3-to-1 is big enough?} As a reminder, Bayes factors are the relative predictive accuracy of two models for the observed data.
 Therefore, Bayes factors are odds or ratios.
@@
 However, the Bayes factor estimates in `rstan` generally show stronger evidence than Bayes factor estimates based on `brms`.
-It is important to note that this difference is not due to the bridge sampling function, but a result of a difference in parametrization.
-\Revision{The models for} `rstan` \Revision{and} `brms` \Revision{are similar but not identical, partly due to the difference in prior specification (i.e., variances versus standard deviations).} However, for both packages, the conclusion is the same: the data are least likely to have occurred under the full model.
+\Revision{It is important to note that this difference is not due to the bridge sampling function (i.e., there is a little to no variation of Bayes factor estimates within packages), but a result of a difference in parametrization.
+The models for} `rstan` \Revision{and} `brms` \Revision{are similar but not identical, partly due to the difference in prior specification (i.e., variances versus standard deviations).} However, for both packages, the conclusion is the same: the data are least likely to have occurred under the full model.
 In case the estimates are unstable, it is recommended to further increase the number of posterior samples.
@@
 Then, we can calculate the posterior model probabilities for all specified models.
+\Revision{Typically in sensitivity analyses, the sensitivity of the Bayes factor on the width of the prior on effect size is assessed} [e.g., @van2019jasp]\Revision{.
+However, in our case the prior distributions on the variance parameters are key to understand sensitivity.}
 
@@
 \Revision{There are also other R software packages for Bayesian multilevel modeling besides} `brms` \Revision{and} `rstan`\Revision{, such as} `BayesFactor`[@R-BayesFactor] \Revision{and} `rstanarm`[@rstanarm2020]\Revision{, that we did not illustrate in this tutorial.
-However, most of these other R software packages are a bit more limited, for instance, in model specification and prior settings.}
+However, most of these other R software packages are more limited, for instance, in model specification and prior settings.}
 
@@
 We already highlighted the drawbacks of the SD density ratio.
-\Revision{Bayes factor estimates can also be biased.
+\Revision{Bayes factor estimates can also be biased (i.e., display a consistent deviation from the true Bayes factor value).
 However, bridge sampling has shown to be relatively unbiased} [@sarafoglou2021evaluating].

### 2023-06-28 manuscript.tex
Updates

@@
 Therefore, when this option is applied and individual variability is ignored, the general trend can be an inaccurate representation of the true general trend (Rouder, Morey, \& Pratte, 2013; Haaf \& Rouder, 2017; van Doorn, Aust, Haaf, Stefan, and Wagenmakers, 2021).
-This is shown in Panel A of Figure \ref{fig:figure1} where some of the individual trends (i.e., slopes of the grey lines) differ from the general trend (i.e., slope of the blue line).
+
+\textcolor{blue}{A pattern where a general trend is not representing the individual trend is} shown in Panel A of Figure \ref{fig:figure1} where some of the individual trends (i.e., slopes of the grey lines) differ from the general trend (i.e., slope of the blue line).
 If every participant takes part in each condition, the person variability affects all conditions, thus inducing a correlation across conditions.
@@
 \end{align*}
-\textcolor{blue}{The model equation shows that, for instance, for participant $i = 1$ and digit $j = 7$, with $\gamma_{i = 1}$ equal to $0.5$, $\beta_{i = 1}$ equal to $0.2$, $\delta_{7, i = 1}$ and $\delta_{3, i = 1}$ equal to $0.25$, and $\delta_{6, i = 1}$ and $\delta_{4, i = 1}$ equal to $0.3$, the mean RT is modeled by: $0.5 - 0.1 +  0.25$.
-If all $\delta$s are positive, then the model equation indicates that individuals respond fastest when the digits 2 and 8 are presented, as in this situation there are no digit effects in the equation.
-However, for digit $j = 8$, the mean RT is modeled by $0.5 - 0.1$ in this scenario, while for digit $j = 2$, the mean RT is modeled by $0.5 + 0.1$.}
+
+\textcolor{blue}{The model equation enables us to compute the expected response time $Y_{ij}$ for a participant $i$ when presented with a specific probe.
+For example, if participant $i = 1$ is presented with digit $j = 7$, $x_j$ takes the value $-\frac{1}{2}$ and $u_j$ takes the value $1$.
+The parameters $v_j$, $w_j$, and $z_j$ take the value zero, such that the equation for the expected value reduces to: $\gamma_{i} + \frac{1}{2} \beta_{i} + \delta_{7,i}$.
+Let us assume that the participant 1's mean response time across all items is $\gamma_{1} = 0.5$, they respond generally faster to digits to the right of five, as expressed by $\beta_{1} = 0.2$, and they respond slower to digit 7, given by $\delta_{7,1} = 0.25$.
+Then their expected response time for the digit 7 can be calculated by $0.5 - 0.5 * 0.2 + 1 * 0.25$.}
 
@@
 \end{equation}
-Note that the only change in the model formulation is the change of the likelihood function from a normal distribution to a log-normal distribution.
-\textcolor{blue}{The likelihood function describes the probability model that is assumed to be underlying the data.
-When the parameters of this probability model are known, the joint probability of the data can be inferred from the likelihood function.
+Note that the only change in the model formulation is the change from a normal distribution to a log-normal distribution.
+\textcolor{blue}{The density function describes the probability model that is assumed to be underlying the data.
+When the parameters of this probability model are known, the joint probability of the data can be inferred from the \textcolor{blue}{density} function.
 The log-normal distribution only assigns probability to positive data values, and is, just as the RT distributions, right-skewed.
@@
 This is plugged into the individual effect distributions (Equation (2)).
-Then, the likelihood of the model, such as specified in Equations (1) and (3), is used to simulate observations per condition.
+Then, the probability of the model, such as specified in Equations (1) and (3), is used to simulate observations per condition.
 This means that for \emph{n} participants, \emph{n} random values for the model parameters are generated based on the priors, which will result in the \emph{k} observations per \emph{j} conditions.
@@
 All variance terms (\(\sigma_{\gamma}^2\), \(\sigma_{\beta}^2\), \(\sigma_{\delta}^2\), and \(\sigma^2\)) are constrained to be positive.
-If there is no information from previous studies, researchers might be inclined to use the default priors in \texttt{brms} as a starting point.
-
-In \texttt{brms}, different default priors are set on different types of parameters (Bürkner, 2017).
+If there is no information from previous studies, researchers might be inclined to use the default priors \textcolor{blue}{of software packages, often reasonably uninformative priors,} as a starting point.
+
+\textcolor{blue}{We take the reasonably uninformative priors as a starting point as well specifying priors on different types of parameters.
+These priors were previously used as} default priors in \texttt{brms} (Bürkner, 2017).
 The prior on the group-level regression weights (\(\mu_{\beta}\), \(\mu_{\delta}\)) is a normal distribution with a mean of 0 and a standard deviation of 1.
@@
 \end{equation}
-The prior distributions \texttt{brms} specifies on the standard deviations can be transformed to prior distributions on variances (\(var\)).
+The prior distributions specified on the standard deviations can be transformed to prior distributions on variances (\(var\)).
 Prior distributions on variances have a smaller scale than priors on standard deviation (\(sd\)).
@@
 The blue dots represent the mean effect.
-Using the default prior from \texttt{brms}, the mean effects are expected to be around zero, with a mean effect of zero.
+Using the \textcolor{blue}{reasonably uninformatieve priors}, the mean effects are expected to be around zero, with a mean effect of zero.
 However, the range around this mean is very wide, ranging from -10 to 10 second effects.
@@
 Finally, we would like to use a more common conjugate prior for the variances to ensure obtaining a proper posterior distribution.\footnote{Conjugate priors result in posterior distributions that belong to the same distribution family as the prior distribution (Fink, 1997a).
-For a normal likelihood, these are the normal distribution for the mean and the inverse gamma distribution for the variance.} \footnote{\textcolor{blue}{However, in rstan, non-conjugate priors are commonly applied} (e.g., Gelman, 2006).} Therefore, we choose an inverse gamma distribution.
+For a normal model, these are the normal distribution for the mean and the inverse gamma distribution for the variance.} \footnote{\textcolor{blue}{However, in rstan, non-conjugate priors are commonly applied} (e.g., Gelman, 2006).} Therefore, we choose an inverse gamma distribution.
 We can also be more specific about the expected RTs.
@@
 This yields the following priors that result in the prior prediction shown in panel D of Figure \ref{fig:priorpredfigurepanel}:\footnote{For the overall baseline \(\mu_\gamma\) we adjusted the prior based on our expectation of the expected overall response times.
-A mean parameter of -0.5 for the log-normal likelihood corresponds to a mean response time of 1.}
+A mean parameter of -0.5 for the log-normal distribution corresponds to a mean response time of 1.}
 
@@
 Every sampling algorithm requires the specification of the number of samples (also called iterations) that are drawn from the posterior distribution.
-Furthermore, all MCMC sampling methods have in common that results depend (somewhat) on chosen starting values, that is, initial parameter values for which the unnormalized posterior (i.e., likelihood times prior) is evaluated.
+Furthermore, all MCMC sampling methods have in common that results depend (somewhat) on chosen starting values, that is, initial parameter values that \textcolor{blue}{are fed to the sampler in the first iteration}.
 This dependency can be assessed if the sampling algorithm is repeated several times (i.e., chains) with different starting values.
@@
 \texttt{Rstan} requires two files: a \texttt{.stan} file containing the model specification and an R script containing the code for the model fitting.
-In the \texttt{.stan} file, the parameters, priors, and the likelihood are specified.
+In the \texttt{.stan} file, the parameters, priors, and the \textcolor{blue}{probability distribution of the data} are specified.
 An example \texttt{.stan} file of a simple, non-hierarchical normal model, where the variance is already known and the mean has to be estimated, is shown below (adapted example from Nicenboim, Schad, \& Vasishth, 2021, Chapter 10).
-The \texttt{.stan} file for the symbolic distance effect, including explanations, can be found in \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/blob/main/R\%20objects/rstan/normal\%20model/myModel.stan}{Online Supplement K}.
 The file is divided into three sections: \texttt{data}, \texttt{parameters}, and \texttt{model}.
@@
 \NormalTok{  target }\SpecialCharTok{+}\ErrorTok{=} \FunctionTok{normal\_lpdf}\NormalTok{(mu }\SpecialCharTok{|} \DecValTok{0}\NormalTok{, }\DecValTok{20}\NormalTok{);}
-  \SpecialCharTok{/}\ErrorTok{/}\NormalTok{ Likelihood}\SpecialCharTok{:} 
+  \SpecialCharTok{/}\ErrorTok{/}\NormalTok{ Probability model}\SpecialCharTok{:} 
   \ControlFlowTok{for}\NormalTok{(i }\ControlFlowTok{in} \DecValTok{1}\SpecialCharTok{:}\NormalTok{N)}
@@
 These are the general and individual effects.
-The last section, \texttt{model}, contains the likelihood function and priors, as specified earlier in this article.
+The last section, \texttt{model}, contains the probability function and priors, as specified earlier in this article.
+\textcolor{blue}{Instead of using} \texttt{y{[}i{]}\ \textasciitilde{}\ normal(mu,\ sigma)} \textcolor{blue}{to estimate parameters, such as the mean from the example, we used the} \texttt{target} \textcolor{blue}{specification} (for a in-depth explanation see Nicenboim, Schad, \& Vasishth, 2021, Chapter 10).
+\textcolor{blue}{Both methods indicate in this case that} \texttt{y{[}i{]}} \textcolor{blue}{is normally distributed with the mean} \texttt{mu} \textcolor{blue}{and the variance} \texttt{sigma}.
+\textcolor{blue}{However, the target method saves constant terms that are required for model comparison} (Gronau, Singmann, \& Wagenmakers, 2020) The \texttt{.stan} file for the symbolic distance effect, including explanations, can be found in \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/blob/main/R\%20objects/rstan/normal\%20model/myModel.stan}{Online Supplement K}.
 
@@
 The package \texttt{rstan} offers a lot of flexibility.
-It is possible to specify the likelihood and priors in the exact way as is wished.
+It is possible to specify the probability model and priors in the exact way as is wished.
 However, this comes with the price of some technical knowledge.
@@
 However, the Bayes factor estimates in \texttt{rstan} generally show stronger evidence than Bayes factor estimates based on \texttt{brms}.
-It is important to note that this difference is not due to the bridge sampling function, but a result of a difference in parametrization.
-\textcolor{blue}{The models for} \texttt{rstan} \textcolor{blue}{and} \texttt{brms} \textcolor{blue}{are similar but not identical, partly due to the difference in prior specification (i.e., variances versus standard deviations).} However, for both packages, the conclusion is the same: the data are least likely to have occurred under the full model.
+\textcolor{blue}{It is important to note that this difference is not due to the bridge sampling function (i.e., there is a little to no variation of Bayes factor estimates within packages), but a result of a difference in parametrization.
+The models for} \texttt{rstan} \textcolor{blue}{and} \texttt{brms} \textcolor{blue}{are similar but not identical, partly due to the difference in prior specification (i.e., variances versus standard deviations).} However, for both packages, the conclusion is the same: the data are least likely to have occurred under the full model.
 In case the estimates are unstable, it is recommended to further increase the number of posterior samples.
@@
 Then, we can calculate the posterior model probabilities for all specified models.
+\textcolor{blue}{Typically in sensitivity analyses, the sensitivity of the Bayes factor on the width of the prior on effect size is assessed} (e.g., van Doorn et al., 2021)\textcolor{blue}{.
+However, in our case the prior distributions on the variance parameters are key to understand sensitivity.}
 
@@
 We already highlighted the drawbacks of the SD density ratio.
-\textcolor{blue}{Bayes factor estimates can also be biased.
+\textcolor{blue}{Bayes factor estimates can also be biased (i.e., display a consistent deviation from the true Bayes factor value).
 However, bridge sampling has shown to be relatively unbiased} (Sarafoglou et al., 2021).
@@
 
+\leavevmode\hypertarget{ref-van2019jasp}{}%
+van Doorn, J., van den Bergh, D., Bohm, U., Dablander, F., Derks, K., Draws, T., \ldots{} Wagenmakers, E.-J.
+(2021).
+The JASP guidelines for conducting and reporting a bayesian analysis.
+\emph{Psychonomic Bulletin \& Review}, \emph{28}, 813--826.
+doi:\href{https://doi.org/10.3758/s13423-020-01798-5}{10.3758/s13423-020-01798-5}
+
 \leavevmode\hypertarget{ref-van2018simple}{}%

### 2023-06-28 J - R code/Rstan/Normal model/BHM in rstan_cleaned.R

@@ -138,7 +138,7 @@ datareal$f3 <- f3
 ################################################################################
 library(rstan) 
 # Adjust "myPath" below to your directory  
-hier_modelc_adj_07012022 <- stan(file = "/myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/rstan/log-normal model/rewritestanmodel3_trunc.stan", data = datareal, 
+hier_modelc_adj_07012022 <- stan(file = "/myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/rstan/log-normal model/rewritestanmodel3_trunc.stan", data = datareal, 
                         iter = 4000, chains = 4,
                         # , control = list(max_treedepth = 15, adapt_delta = 0.95)
                         warmup = 1000, cores = 4)

### 2023-06-28 manuscript.Rmd
Updates

@@
 
-\Revision{A pattern where a general trend is not representing the individual trend is} shown in Panel A of Figure \@ref(fig:figure1) where some of the individual trends (i.e., slopes of the grey lines) differ from the general trend (i.e., slope of the blue line).
-If every participant takes part in each condition, the person variability affects all conditions, thus inducing a correlation across conditions.
+\Revision{A pattern where the general trend does not reflect individual trends is shown in Panel A of Figure} \@ref(fig:figure1)\Revision{.
+The slopes of the grey lines reflect individual trends, and some of them differ considerably from the general trend shown as the slope of the blue line.} If every participant takes part in each condition, the person variability affects all conditions, thus inducing a correlation across conditions.
 For example, participants in a digit classification task repeatedly respond to target digits inducing a correlation between the responses to the target digits, such as between target digit 3 and target digit 4.
@@
 Conceptually, a hierarchical analysis provides a compromise between the aggregated and independent approach by drawing benefits from each method.
-The motivation for hierarchical models, as well as their general mathematical structure, has been covered by several tutorial articles} [e.g., @singmann2019introduction; @west2022linear]\Revision{, as well as text books (e.g., XX, XX).}
+The motivation for hierarchical models, as well as their general mathematical structure, has been covered by several (tutorial) articles} [e.g., @efron1977stein; @singmann2019introduction; @west2022linear]\Revision{, as well as text books} [e.g., @gelman2006data; @hox2017multilevel; @nicenboim2023introduction].
 
@@
 If digits 2 and 8 were presented all four variables are zero.
-The dummy variables represent the digits 3 ($u$), 4 ($v$), 6 ($w$), and 7 ($z$), and each take the value 1 if the respective digit was presented in the trial, and 0 if it was not.
+The dummy variables represent the digits 3 ($u$), 4 ($v$), 6 ($w$), and 7 ($z$), and each take the value 1 if the respective digit was presented in the trial, and 0 if it was not [for more on contrast coding, we refer to @pinheiro2006mixed; and @singmann2019introduction].
 
@@
 \end{equation}
-Note that the only change in the model formulation is the change from a normal distribution to a log-normal distribution.
-\Revision{The density function describes the probability model that is assumed to be underlying the data.
-When the parameters of this probability model are known, the joint probability of the data can be inferred from the \Revision{density} function.
+
+\Revision{Note that the only change in the model formulation is the change from a normal distribution to a log-normal distribution.
+The density function describes the probability model that is assumed to be underlying the data.
+When the parameters of this probability model are known, the joint probability of the data can be inferred from the density function.
 The log-normal distribution only assigns probability to positive data values, and is, just as the RT distributions, right-skewed.

### 2023-06-28 manuscript.tex
Updates

@@
 
-\textcolor{blue}{A pattern where a general trend is not representing the individual trend is} shown in Panel A of Figure \ref{fig:figure1} where some of the individual trends (i.e., slopes of the grey lines) differ from the general trend (i.e., slope of the blue line).
-If every participant takes part in each condition, the person variability affects all conditions, thus inducing a correlation across conditions.
+\textcolor{blue}{A pattern where the general trend does not reflect individual trends is shown in Panel A of Figure} \ref{fig:figure1}\textcolor{blue}{.
+The slopes of the grey lines reflect individual trends, and some of them differ considerably from the general trend shown as the slope of the blue line.} If every participant takes part in each condition, the person variability affects all conditions, thus inducing a correlation across conditions.
 For example, participants in a digit classification task repeatedly respond to target digits inducing a correlation between the responses to the target digits, such as between target digit 3 and target digit 4.
-\textcolor{blue}{Therefore, analyses on aggregated data do not appropriately account for the interdependence between observations} (Doorn, Aust, Haaf, Stefan, \& Wagenmakers, 2021, 2023).
+\textcolor{blue}{Therefore, analyses on aggregated data do not appropriately account for differential experiment effects} (Doorn, Aust, Haaf, Stefan, \& Wagenmakers, 2021, 2023).
 As a result, the type I error rate of a test of the effect increases: there is a higher probability of concluding that there is an effect if there is no effect (Rouder \& Lu, 2005).
@@
 Conceptually, a hierarchical analysis provides a compromise between the aggregated and independent approach by drawing benefits from each method.
-The motivation for hierarchical models, as well as their general mathematical structure, has been covered by several tutorial articles} (e.g., Singmann \& Kellen, 2019; West, Welch, \& Galecki, 2022)\textcolor{blue}{, as well as text books (e.g., XX, XX).}
+The motivation for hierarchical models, as well as their general mathematical structure, has been covered by several tutorial articles} (e.g., Singmann \& Kellen, 2019; West, Welch, \& Galecki, 2022)\textcolor{blue}{, as well as text books} (e.g., Gelman \& Hill, 2006; Hox, Moerbeek, \& Van de Schoot, 2017; Nicenboim, Schad, \& Vasishth, 2023).
 
@@
 If digits 2 and 8 were presented all four variables are zero.
-The dummy variables represent the digits 3 (\(u\)), 4 (\(v\)), 6 (\(w\)), and 7 (\(z\)), and each take the value 1 if the respective digit was presented in the trial, and 0 if it was not.
+The dummy variables represent the digits 3 (\(u\)), 4 (\(v\)), 6 (\(w\)), and 7 (\(z\)), and each take the value 1 if the respective digit was presented in the trial, and 0 if it was not (for more on contrast coding, we refer to Pinheiro \& Bates, 2006; and Singmann \& Kellen, 2019).
 
@@
 \end{equation}
-Note that the only change in the model formulation is the change from a normal distribution to a log-normal distribution.
-\textcolor{blue}{The density function describes the probability model that is assumed to be underlying the data.
-When the parameters of this probability model are known, the joint probability of the data can be inferred from the \textcolor{blue}{density} function.
+
+\textcolor{blue}{Note that the only change in the model formulation is the change from a normal distribution to a log-normal distribution.
+The density function describes the probability model that is assumed to be underlying the data.
+When the parameters of this probability model are known, the joint probability of the data can be inferred from the density function.
 The log-normal distribution only assigns probability to positive data values, and is, just as the RT distributions, right-skewed.
@@
 \textcolor{blue}{A prior is proper if two conditions hold: 1.
-All values of the probability distribution are positive definite, and 2.
+All values of the probability distribution are equal or greater than zero, and 2.
 the probability distribution sums to one for discrete data or integrates to one for continuous data.
@@
 \textcolor{blue}{The figure shows two hypothetical priors for parameter $\theta$.
-The first condition can be assessed by checking the y-values---both priors have only positive definite values as function output (0.1, 0.2, and 0.4).
+The first condition can be assessed by checking the y-values---both priors have only positive values as function output (0.1, 0.2, and 0.4).
 To assess the second condition, we need to sum up the function values.
@@
 \textcolor{blue}{The difference between the model diagnostics of} \texttt{rstan} \textcolor{blue}{and} \texttt{brms} \textcolor{blue}{can be partly explained by the difference in model parametrization, such as the difference in the prior specification (i.e., specifying priors on variances versus priors on standard deviations).
-Although we tried to make the models as similar as possible, they are not identical.}
+We equated the models where sensible which resulted in models that are fairly comparable.
+}
 
@@
 This means that the data are 100 times more likely under the normal model than under the log-normal model, clear evidence in favor of the normal model.
-\textcolor{blue}{But how do we know whether a Bayes factor of 100-to-1, or 10-to-1 or 3-to-1 fulfills?} As a reminder, Bayes factors are the relative predictive accuracy of two models for the observed data.
+\textcolor{blue}{But how do we know whether a Bayes factor of 100-to-1, or 10-to-1 or 3-to-1 is big enough?} As a reminder, Bayes factors are the relative predictive accuracy of two models for the observed data.
 Therefore, Bayes factors are odds or ratios.
@@
 \textcolor{blue}{There are also other R software packages for Bayesian multilevel modeling besides} \texttt{brms} \textcolor{blue}{and} \texttt{rstan}\textcolor{blue}{, such as} \texttt{BayesFactor}(Morey \& Rouder, 2018) \textcolor{blue}{and} \texttt{rstanarm}(Goodrich, Gabry, Ali, \& Brilleman, 2020)\textcolor{blue}{, that we did not illustrate in this tutorial.
-However, most of these other R software packages are a bit more limited, for instance, in model specification and prior settings.}
+However, most of these other R software packages are more limited, for instance, in model specification and prior settings.}
 
@@
 
+\leavevmode\hypertarget{ref-gelman2006data}{}%
+Gelman, A., \& Hill, J. (2006). \emph{Data analysis using regression and multilevel/hierarchical models}. Cambridge university press.
+
 \leavevmode\hypertarget{ref-gelman1992inference}{}%
@@
 Hope, R. M. (2013). \emph{Rmisc: Rmisc: Ryan miscellaneous}. Retrieved from \url{https://CRAN.R-project.org/package=Rmisc}
+
+\leavevmode\hypertarget{ref-hox2017multilevel}{}%
+Hox, J. J., Moerbeek, M., \& Van de Schoot, R. (2017). \emph{Multilevel analysis: Techniques and applications}. Routledge.
 
@@
 
+\leavevmode\hypertarget{ref-nicenboim2023introduction}{}%
+Nicenboim, B., Schad, D., \& Vasishth, S. (2023). \emph{An introduction to bayesian data analysis for cognitive science}. Chapman; Hall/CRC Statistics in the Social; Behavioral Sciences Series.
+
+\leavevmode\hypertarget{ref-pinheiro2006mixed}{}%
+Pinheiro, J., \& Bates, D. (2006). \emph{Mixed-effects models in s and s-PLUS}. Springer science \& business media.
+
 \leavevmode\hypertarget{ref-Plummer03jagsa}{}%

### 2023-06-28 propriety_section.Rmd
Updates

@@
 \Revision{Rather than submerging us in technical details or providing a formal justification of the Bayesian approach, we provide an intuition of Bayesian model comparison.
-A formal justification of Bayes factors is provided by @jeffreys1961theory, @kass1995bayes and @rouder2018teaching.
-The Bayes factor (BF) is a measure of how well one model predicts the data compared to another.}
+A formal justification of Bayes factors is provided by} @jeffreys1961theory, @kass1995bayes and @rouder2018teaching.
+\Revision{The Bayes factor (BF) is a measure of how well one model predicts the data compared to another.}
 

### 2023-06-28 J - R code/Brms/Normal model/brms_model_cleaned_trunc.R

@@ -147,7 +147,7 @@ brmmodel8_new03012022 <- brm(formula = bform,         # Model formula
                              core = 4,                   # Cores for parallel estimation 
                              #control = list(adapt_delta = 0.95),  # Control sampler's behavior, this avoided the problem of convergement 
                              # Finally we can save the model fit to a file 
-                             file = "/myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R code/BRMSmodelfit18_03012022")
+                             file = "/myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/J - R code/BRMSmodelfit18_03012022")
 
 # For BF 
 brmmodel8_new_bf_03012022 <- brm(formula = bform,         # Model formula 
@@ -161,5 +161,5 @@ brmmodel8_new_bf_03012022 <- brm(formula = bform,         # Model formula
                              core = 4,                   # Cores for parallel estimation 
                              #control = list(adapt_delta = 0.95),  # Control sampler's behavior, this avoided the problem of convergement 
                              # Finally we can save the model fit to a file 
-                             file = "/myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R code/BRMSmodelfit18_BF_03012022")
+                             file = "/myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/J - R code/BRMSmodelfit18_BF_03012022")
 

### 2023-06-28 manuscript.Rmd
Updates (BF section)

@@
 Specifically, we will achieve this by comparing the predictive accuracy of a null model where we set one or some parameters to zero, and and effects models where these parameters are unconstrained.
+\Revision{This approach is called Bayes factor model comparison.}
 
@@
 ## Bayes factor
-The Bayes factor (BF) is a measure of how well one model predicts the data compared to another [@jeffreys1961theory; @kass1995bayes; @rouder2018teaching].
-The Bayes factor is defined as the probability of the data under one model compared to the probability of the data under another model.
-For instance, the Bayes factor comparing a null model ($\mathcal{M}_0$) and a full model ($\mathcal{M}_{\text{F}}$) can be obtained by:
+\Revision{Rather than submerging us in technical details or providing a formal justification of the Bayesian approach, we provide an intuition of Bayesian model comparison.
+A formal justification of Bayes factors is provided by} @jeffreys1961theory, @kass1995bayes and @rouder2018teaching.
+\Revision{The Bayes factor (BF) is a measure of how well one model predicts the data compared to another.}
+
+\Revision{Bayesian analysis allows for predictions of data from any model because models are fully specified with a probability distribution for the data and priors for all parameters.
+Let the vector of all parameters of a model be $\bm{\theta}$.
+With the prior distributions of these parameters, $f(\bm{\theta})$, the prediction for data from the model can be computed by integrating over all parameters:}
 
 \begin{equation}
-\text{BF}_{0\text{F}} = \frac{P(\bm{Y} | \mathcal{M}_0)}{P(\bm{Y} | \mathcal{M}_{\text{F}})}. 
+p(\bm{Y} | \mathcal{M}) = \int_{\bm{\Theta}} p(\bm{Y} | \mathcal{M}, \bm{\theta})f(\bm{\theta})d\bm{\theta}. 
 \end{equation}
 
-The Bayes factor can also be interpreted as the relative evidence of two competing models.
+\Revision{Note that $\int_{\bm{\Theta}}$ refers to a multidimensional integral over the entire parameter space.
+As our models consist of many parameters, computing these predictions for any one model becomes somewhat difficult.
+Once the integral is computed, the resulting probability distribution of the data, $p(\bm{Y} | \mathcal{M})$ is not a function of the model parameters anymore.
+This marginal probability distribution is therefore also referred to as the marginal likelihood.
+Before observing the data, the marginal likelihood serves as prediction for possible data; once the data are observed, the function tells us how well these specific observations were predicted by the model.
+The Bayes factor then is the ratio of marginal likelihoods of two models, or the relative predictive accuracy of $\mathcal{M}_1$ over $\mathcal{M}_2$:}
+
+\begin{equation}
+\text{BF}_{1,2} = \frac{P(\bm{Y} | \mathcal{M}_1)}{P(\bm{Y} | \mathcal{M}_2)}. 
+\end{equation}
+
+\Revision{There is a second popular interpretation of the Bayes factor as the relative evidence of two competing models.
+One of the key insights from Bayes' rule is that these two things, the relative predictive accuracy of two models for the data, and the relative evidence from the data for the two models, are one and the same} [@rouder2018teaching].
+
 Bayes factors have a few very convenient characteristics.
 One of them is that Bayes factors are transitive, that is, evidence for the full model over the null model can be obtained by flipping numerator and denominator in equation 13.
-For a full overview of the advantages of Bayesian inference with Bayes factor, see Wagenmakers et al.
-[-@wagenmakers2018bayesian].
-
-Our models consist of several parameters.
-Therefore, computing the Bayes factor of, for example, the normal model versus the log-normal model becomes somewhat difficult.
-Let the vector of all parameters of the normal model be $\bm{\theta_\text{n}}$ and the vector of all parameters for the log-normal model be $\bm{\theta_{\text{l}}}$.
-With the prior distributions for the two models, $f(\bm{\theta_\text{n}})$ and $f(\bm{\theta_\text{l}})$, the Bayes factor can be calculated by the following ratio of marginal likelihoods:
-
-\begin{equation}
-\text{BF}_{\text{nl}} = \frac{P(\bm{Y} | \mathcal{M}_n)}{P(\bm{Y} | \mathcal{M}_{\text{l}})} = \frac{\int_{-\infty}^{\infty}P(\bm{Y} | \mathcal{M}_\text{n}, \bm{\theta_{\text{n}}})f(\bm{\theta_{\text{n}}})d\bm{\theta_{\text{n}}}}{\int_{-\infty}^{\infty}P(\bm{Y} | \mathcal{M}_\text{l}, \bm{\theta_{\text{l}}})f(\bm{\theta_{\text{l}}})d\bm{\theta_{\text{l}}}}.
-\end{equation}
-
-Imagine this results in $\text{BF}_{\text{nl}} = 100$.
+For a full overview of the advantages of Bayesian inference with Bayes factor, see Wagenmakers and colleagues [-@wagenmakers2018bayesian].
+
+Imagine that $\text{BF}_{1,2} = 100$.
 This means that the data are 100 times more likely under the normal model than under the log-normal model, clear evidence in favor of the normal model.
@@
 
-Same as the posterior distribution, Bayes factors can rarely be computed analytically due to the multidimensional integrals involved.
-The difficulty to compute Bayes factors is one of the major drawbacks of Bayesian inference.
+\Revision{Same as the posterior distribution, Bayes factors can rarely be computed analytically.} The difficulty to compute Bayes factors is one of the major drawbacks of Bayesian inference.
 However, several algorithms are available to estimate Bayes factor from posterior samples.

### 2023-06-28 manuscript.tex
Updates (BF section)

@@
 Conceptually, a hierarchical analysis provides a compromise between the aggregated and independent approach by drawing benefits from each method.
-The motivation for hierarchical models, as well as their general mathematical structure, has been covered by several tutorial articles} (e.g., Singmann \& Kellen, 2019; West, Welch, \& Galecki, 2022)\textcolor{blue}{, as well as text books} (e.g., Gelman \& Hill, 2006; Hox, Moerbeek, \& Van de Schoot, 2017; Nicenboim, Schad, \& Vasishth, 2023).
+The motivation for hierarchical models, as well as their general mathematical structure, has been covered by several (tutorial) articles} (e.g., Efron \& Morris, 1977; Singmann \& Kellen, 2019; West, Welch, \& Galecki, 2022)\textcolor{blue}{, as well as text books} (e.g., Gelman \& Hill, 2006; Hox, Moerbeek, \& Van de Schoot, 2017; Nicenboim, Schad, \& Vasishth, 2023).
 
@@
 Specifically, we will achieve this by comparing the predictive accuracy of a null model where we set one or some parameters to zero, and and effects models where these parameters are unconstrained.
+\textcolor{blue}{This approach is called Bayes factor model comparison.}
 
@@
 
-The Bayes factor (BF) is a measure of how well one model predicts the data compared to another (Jeffreys, 1961; Kass \& Raftery, 1995; Rouder \& Morey, 2018).
-The Bayes factor is defined as the probability of the data under one model compared to the probability of the data under another model.
-For instance, the Bayes factor comparing a null model (\(\mathcal{M}_0\)) and a full model (\(\mathcal{M}_{\text{F}}\)) can be obtained by:
+\textcolor{blue}{Rather than submerging us in technical details or providing a formal justification of the Bayesian approach, we provide an intuition of Bayesian model comparison.
+A formal justification of Bayes factors is provided by} Jeffreys (1961), Kass and Raftery (1995) and Rouder and Morey (2018).
+\textcolor{blue}{The Bayes factor (BF) is a measure of how well one model predicts the data compared to another.}
+
+\textcolor{blue}{Bayesian analysis allows for predictions of data from any model because models are fully specified with a probability distribution for the data and priors for all parameters.
+Let the vector of all parameters of a model be $\bm{\theta}$.
+With the prior distributions of these parameters, $f(\bm{\theta})$, the prediction for data from the model can be computed by integrating over all parameters:}
 
 \begin{equation}
-\text{BF}_{0\text{F}} = \frac{P(\bm{Y} | \mathcal{M}_0)}{P(\bm{Y} | \mathcal{M}_{\text{F}})}. 
+p(\bm{Y} | \mathcal{M}) = \int_{\bm{\Theta}} p(\bm{Y} | \mathcal{M}, \bm{\theta})f(\bm{\theta})d\bm{\theta}. 
 \end{equation}
 
-The Bayes factor can also be interpreted as the relative evidence of two competing models.
+\textcolor{blue}{Note that $\int_{\bm{\Theta}}$ refers to a multidimensional integral over the entire parameter space.
+As our models consist of many parameters, computing these predictions for any one model becomes somewhat difficult.
+Once the integral is computed, the resulting probability distribution of the data, $p(\bm{Y} | \mathcal{M})$ is not a function of the model parameters anymore.
+This marginal probability distribution is therefore also referred to as the marginal likelihood.
+Before observing the data, the marginal likelihood serves as prediction for possible data; once the data are observed, the function tells us how well these specific observations were predicted by the model.
+The Bayes factor then is the ratio of marginal likelihoods of two models, or the relative predictive accuracy of $\mathcal{M}_1$ over $\mathcal{M}_2$:}
+
+\begin{equation}
+\text{BF}_{1,2} = \frac{P(\bm{Y} | \mathcal{M}_1)}{P(\bm{Y} | \mathcal{M}_2)}. 
+\end{equation}
+
+\textcolor{blue}{There is a second popular interpretation of the Bayes factor as the relative evidence of two competing models.
+One of the key insights from Bayes' rule is that these two things, the relative predictive accuracy of two models for the data, and the relative evidence from the data for the two models, are one and the same} (Rouder \& Morey, 2018).
+
 Bayes factors have a few very convenient characteristics.
 One of them is that Bayes factors are transitive, that is, evidence for the full model over the null model can be obtained by flipping numerator and denominator in equation 13.
-For a full overview of the advantages of Bayesian inference with Bayes factor, see Wagenmakers et al.
-(2018).
-
-Our models consist of several parameters.
-Therefore, computing the Bayes factor of, for example, the normal model versus the log-normal model becomes somewhat difficult.
-Let the vector of all parameters of the normal model be \(\bm{\theta_\text{n}}\) and the vector of all parameters for the log-normal model be \(\bm{\theta_{\text{l}}}\).
-With the prior distributions for the two models, \(f(\bm{\theta_\text{n}})\) and \(f(\bm{\theta_\text{l}})\), the Bayes factor can be calculated by the following ratio of marginal likelihoods:
-
-\begin{equation}
-\text{BF}_{\text{nl}} = \frac{P(\bm{Y} | \mathcal{M}_n)}{P(\bm{Y} | \mathcal{M}_{\text{l}})} = \frac{\int_{-\infty}^{\infty}P(\bm{Y} | \mathcal{M}_\text{n}, \bm{\theta_{\text{n}}})f(\bm{\theta_{\text{n}}})d\bm{\theta_{\text{n}}}}{\int_{-\infty}^{\infty}P(\bm{Y} | \mathcal{M}_\text{l}, \bm{\theta_{\text{l}}})f(\bm{\theta_{\text{l}}})d\bm{\theta_{\text{l}}}}.
-\end{equation}
-
-Imagine this results in \(\text{BF}_{\text{nl}} = 100\).
+For a full overview of the advantages of Bayesian inference with Bayes factor, see Wagenmakers and colleagues (2018).
+
+Imagine that \(\text{BF}_{1,2} = 100\).
 This means that the data are 100 times more likely under the normal model than under the log-normal model, clear evidence in favor of the normal model.
@@
 
-Same as the posterior distribution, Bayes factors can rarely be computed analytically due to the multidimensional integrals involved.
-The difficulty to compute Bayes factors is one of the major drawbacks of Bayesian inference.
+\textcolor{blue}{Same as the posterior distribution, Bayes factors can rarely be computed analytically.} The difficulty to compute Bayes factors is one of the major drawbacks of Bayesian inference.
 However, several algorithms are available to estimate Bayes factor from posterior samples.

### 2023-06-28 C - Tutorial Normal Model Rstan/Online Supplement C - Bayesian Hierarchical Modeling in rstan.Rmd
Updates appendices

@@
 
+### Individual Effect Correlation
+
+Up until now, we discussed the normal model from the manuscript.
+This model assumes that there is no correlation between the individual effects.
+As stated in the manuscript, it is also possible to allow for the estimation of the correlation between individual effects.
+This means, for instance, that we would expect that the estimations for the individual effects $\delta_{7,i}$ and $\delta_{3,i}$ are related.
+To allow for correlation between individual effects, we would have to adjust the formula in the `stan` file and the priors.
+For the model with correlation terms, we use the file `myModel_cor.stan`.
+Notice that we apply a different conceptualization of the model using matrix notation.
+The effects are called `beta`.
+
+When including correlation between individual effects, we are estimating more parameters and have to specify priors on these parameters as well.
+In specific, we set a prior on the correlation matrices.
+These priors will be transformed for efficiency and set on so called Cholesky factors of the matrices by the following line of code `Lcorr ~ lkj_corr_cholesky(1)` [for more information on Cholesky factors see @stan2018a; section 1.13].
+
 ## Fit the Model
@@
 
+To estimate the model with the correlation between random effects, note that one would need to adjust the `file` argument to equal and `myModel_cor.stan`.
+
+Back to our model without the correlation term form the manuscript.
 We have already fitted the model and saved it in the R object `model_fit`.

### 2023-06-28 D - Tutorial Normal Model brms/Online Supplement D - Bayesian Hierarchical Modeling in brms.Rmd
Updates appendices

@@
 This approach, however, is a bit more difficult to interpret and would also not correspond to the implemented `Stan` model.
+We explain the estimation of the model with correlations across individual effects later on.
 &nbsp;
@@
 
+### Individual Effect Correlation
+
+Up until now, we discussed the normal model from the manuscript.
+This model assumes that there is no correlation between the individual effects.
+As stated in the manuscript, it is also possible to allow for the estimation of the correlation between individual effects.
+This means, for instance, that we would expect that the estimations for the individual effects $\delta_{7,i}$ and $\delta_{3,i}$ are related.
+To allow for correlation between individual effects, we would have to adjust the model formula.
+For the estimation of the digit classification task, we could use the following formula:
+
+\scriptsize 
+```{r formulafitindcor, eval = FALSE}
+bform_cor <- brms::bf(rt2 ~ a + b, nl = TRUE, 
+                  lf(a ~ 0 + side + dif1 + dif2 + dif3 + dif4 + (0 + side + dif1 + dif2 + dif3 + dif4 | ind), center = TRUE),  
+                  lf(b ~ 0 + inter + (0 + inter | ind), cmc = FALSE)
+)
+```
+\normalsize
+
+The difference is that instead of using the double bar `||`, we are now using a single bar `|` before `ind`, indicating that we allow individual effects to correlate.
+
+This means that we are estimating more parameters and therefore we have to specify priors on these parameters as well.
+In specific, we can set a prior on the correlation matrix called `class = "cor"`.
+These priors will be transformed for efficiency and set on so called Cholesky factors of the matrices [for more information see @stan2018a; section 1.13].
+In our example, we set a prior on the Cholesky factors directly, by defining `class = "L"`, shown in the last line of the code block:
+
+\scriptsize 
+```{r priorscor, eval = FALSE}
+priorsmodel_cor <- c(set_prior("normal(0.5,1)", class = "b", lb = 0, nlpar = "b"),  
+                  set_prior("inv_gamma(13.8, 6.3)", class = "sd", coef="inter", group = "sub1", nlpar = "b"), 
+                  set_prior("normal(0,0.3)", class = "b", coef="side",  nlpar = "a"),  
+                  set_prior("inv_gamma(13.8,5.3)", class = "sd", coef="side", group = "sub1",  nlpar = "a"), 
+                  set_prior("normal(0,0.3)", class = "b", coef="dif1",  nlpar = "a"), 
+                  set_prior("inv_gamma(13.8,5.3)", class = "sd", coef="dif1", group = "sub1",  nlpar = "a"),
+                  set_prior("normal(0,0.3)", class = "b", coef="dif2",  nlpar = "a"),  
+                  set_prior("inv_gamma(13.8,5.3)", class = "sd", coef="dif2", group = "sub1",  nlpar = "a"),
+                  set_prior("normal(0,0.3)", class = "b", coef="dif3",  nlpar = "a"), 
+                  set_prior("inv_gamma(13.8,5.3)", class = "sd", coef="dif3", group = "sub1",  nlpar = "a"),
+                  set_prior("normal(0,0.3)", class = "b", coef="dif4",  nlpar = "a"),  
+                  set_prior("inv_gamma(13.8,5.3)", class = "sd", coef="dif4", group = "sub1",  nlpar = "a"),
+                  set_prior("inv_gamma(13.8, 6.3)", class = "sigma"), 
+                  set_prior("lkj_corr_cholesky(1)", class = "L"))
+
+```
+\normalsize
+
 ## Fit the Model
@@
 
-We have already estimated the model and saved it in the R object `brmmodel8_new03012022.rds`.
+To fit the model with the correlation between random effects, note that one would need to adjust `formula` and `priorsmodel` to the objects where these are saved in (i.e., `bform_cor` and `priorsmodel_cor`).
+
+Back to our model without the correlation term form the manuscript.
+We have already estimated this model and saved it in the R object `brmmodel8_new03012022.rds`.
 In case you do not want to wait until the model has been fitted but continue immediately with the rest of this tutorial, you can load this R object that contains the model estimations.
@@
 setwd('..')
-model_fit <- readRDS(file = "R objects/brms/Truncated/BRMSmodelfit18_03012022.rds")
+model_fit <- readRDS(file = "K - R objects/brms/Truncated/BRMSmodelfit18_03012022.rds")
 
 # For model comparison - sample_prior was set to "yes" 
-model_fit_bf <- readRDS("R objects/brms/Truncated/BRMSmodelfit18_BF_03012022.rds")
+model_fit_bf <- readRDS("K - R objects/brms/Truncated/BRMSmodelfit18_BF_03012022.rds")
 

### 2023-06-28 BHM_correlation/Brms/brms_correlation.R

@@ -145,10 +145,11 @@ brmmodel9_new26102022 <- brm(formula = bform_cor,         # Model formula
                              warmup = 1000,              # Iterations used for warmup 
                              iter = 4000,                # Total iterations per chain 
                              chains = 4,                 # Number of chains 
-                             core = 4,                   # Cores for parallel estimation 
+                             core = 4#,                   # Cores for parallel estimation 
                              #control = list(adapt_delta = 0.95),  # Control sampler's behavior, this avoided the problem of convergement 
                              # Finally we can save the model fit to a file 
-                             file = "/myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R code/BRMSmodelfit19_21102022")
+                             #file = "/myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/J - R code/BRMSmodelfit19_21102022"
+                             )
 
 # !!! Check this prior: # student_t(3, 0, 2.5)    sd                           a                default
 # student_t(3, 0, 2.5)    sd                           b                default

### 2023-06-28 BHM_correlation/Rstan/BHM_rstan_correlation.R

@@ -157,9 +157,9 @@ library(rstan)
 
 hier_modelc_cor_06042023 <- stan(file = "myModel_cor_julia.stan", 
                                  data = datareal, 
-                                 iter = 2500, chains = 4,
+                                 iter = 4000, chains = 4,
                               #  control = list(max_treedepth = 15, adapt_delta = 0.97),
-                                 warmup = 2500, cores = 4)
+                                 warmup = 1000, cores = 4)
 
 saveRDS(hier_modelc_cor_06042023, "hier_modelc_cor_06042023.rds")
 

### 2023-06-29 C - Tutorial Normal Model Rstan/Online Supplement C - Bayesian Hierarchical Modeling in rstan.Rmd
Updates figure cor and appendices

--- 2022-08-17 C - Tutorial Normal Model Rstan/Online Supplement C - Bayesian Hierarchical Modeling in rstan.Rmd
@@
 
-|  The stan file `myModel.stan` is used to fit the full model and can be found under R objects > rstan > linear model.
+|  The stan file `myModel.stan` is used to fit the full model and can be found under *K - R objects > rstan > normal model*.
 It contains descriptions for every line of code.
@@
 
+### Individual Effect Correlation
+
+\textcolor{blue}{Up until now, we discussed the normal model from the manuscript.
+This model assumes that there is no correlation between the individual effects.
+As stated in the manuscript, it is also possible to allow for the estimation of the correlation between individual effects.
+This means, for instance, that we would expect that the estimates for the individual effects $\delta_{7,i}$ and $\delta_{3,i}$ are related.
+To allow for correlation between individual effects, we would have to adjust the formula in the `stan` file and the priors.
+For the model with correlation terms, we use the file `myModel_cor.stan` that can be found in the folder} *K - R objects > rstan > normal model > correlation*.
+\textcolor{Notice that we apply a different conceptualization of the model using matrix notation.
+The effects are called `beta` and the individual effects are called `beta_p`.
+Because we are using matrices, the effects are stored a bit differently than for the other model.
+For example, `beta[3]` represents the general digit effect $\delta_{7}$, while `beta_p[1,3]` represents the digit effect for participant 1, $\delta_{7,1}$.}
+
+\textcolor{blue}{When including correlation between individual effects, we are estimating more parameters and have to specify priors for these parameters as well.
+In specific, we set a prior on the correlation matrices.
+These priors will be transformed for efficiency and set on so called Cholesky factors of the matrices by the following line of code `Lcorr ~ lkj_corr_cholesky(1)` [for more information on Cholesky factors see} @stan2018a; \textcolor{blue}{section 1.13].
+}
+
 ## Fit the Model
@@
 library(rstan) 
-model_fit <- stan(file = "myPath/rewritestanmodel3_julia.stan", 
+model_fit <- stan(file = "myPath/myModel.stan", 
                   data = myData_list, 
@@
 setwd('..')
-model_fit <- readRDS(file = "R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
+model_fit <- readRDS(file = "K - R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
 ```
@@
 
+\textcolor{blue}{We can do the same for the model with correlation between the individual effects (see Figure 7) and see that the estimated effects are somewhat different.}
+
+```{r modeldigitestplotcor, echo=FALSE, fig.cap="Model estimates for digit effect parameters in the normal model with correlation between individual effects.
+The points represent the mean parameter estimates for each individual.
+The violin plots on the right side show the variance of the individual parameter estimates.", fig.align = "left", fig.height = 11, fig.width = 1, out.width="75%"}
+knitr::include_graphics("Fig12_modelestimatefigure2_onlyrstancor_v2.png")
+```
+
 ## Model comparison
@@
 To be able to use this package, it is important that the *.stan* file is written in a way that it saves the constants (i.e., `target += normal_lpdf(mu4 | a3, sqrt(b3));`).
-|    First, we fit all the models (except for the full model that we already fitted. 
+|    First, we fit all the models without correlation between individual effects (except for the full model that we already fitted. 
 &nbsp;

### 2023-06-29 D - Tutorial Normal Model brms/Online Supplement D - Bayesian Hierarchical Modeling in brms.Rmd
Updates figure cor and appendices

@@
 
-To fit the model with the correlation between random effects, note that one would need to adjust `formula` and `priorsmodel` to the objects where these are saved in (i.e., `bform_cor` and `priorsmodel_cor`).
+To estimate the model with the correlation between individual effects, note that one would need to adjust `formula` and `priorsmodel` to the objects where these are saved in (i.e., `bform_cor` and `priorsmodel_cor`).
 

### 2023-06-29 BHM_correlation/Brms/brms_correlation.R

@@ -137,19 +137,22 @@ bform_cor <- brms::bf(rt2 ~ a + b, nl = TRUE,
 )
 
 # Adjust 'myPath' to your directory 
-brmmodel9_new26102022 <- brm(formula = bform_cor,         # Model formula 
+brmmodel9_new28062023 <- brm(formula = bform_cor,         # Model formula 
                              data = indat1,              # Data frame with variables + dependent variable 
                              family = gaussian(),        # Response distribution 
                              prior = priorsmodel9_cor,       # Priors 
                              # sample_prior = TRUE, need to add this for model comparison  
-                             warmup = 1000,              # Iterations used for warmup 
-                             iter = 4000,                # Total iterations per chain 
+                             warmup = 2000,              # Iterations used for warmup 
+                             iter = 6000,                # Total iterations per chain 
                              chains = 4,                 # Number of chains 
                              core = 4#,                   # Cores for parallel estimation 
                              #control = list(adapt_delta = 0.95),  # Control sampler's behavior, this avoided the problem of convergement 
                              # Finally we can save the model fit to a file 
                              #file = "/myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/J - R code/BRMSmodelfit19_21102022"
                              )
+
+
+saveRDS(brmmodel9_new28062023, "brmmodel9_new28062023.rds")
 
 # !!! Check this prior: # student_t(3, 0, 2.5)    sd                           a                default
 # student_t(3, 0, 2.5)    sd                           b                default

### 2023-06-29 BHM_correlation/Rstan/BHM_rstan_correlation.R

@@ -155,13 +155,13 @@ datareal$K <- 6  # intercept + side effect + digit effects (4)
 library(rstan) 
 # Don't forget to set the correct directory for the file path
 
-hier_modelc_cor_06042023 <- stan(file = "myModel_cor_julia.stan", 
+hier_modelc_cor_29062023 <- stan(file = "myModel_cor_julia.stan", 
                                  data = datareal, 
                                  iter = 4000, chains = 4,
                               #  control = list(max_treedepth = 15, adapt_delta = 0.97),
                                  warmup = 1000, cores = 4)
 
-saveRDS(hier_modelc_cor_06042023, "hier_modelc_cor_06042023.rds")
+saveRDS(hier_modelc_cor_29062023, "hier_modelc_cor_29062023.rds")
 
 summary(hier_modelc_cor_06042023, pars = c("beta", "s2_p", "sigma", "Omega"))$summary
 traceplot(hier_modelc_cor_06042023, pars = "beta")

### 2023-06-29 C - Tutorial Normal Model Rstan/Online Supplement C - Bayesian Hierarchical Modeling in rstan.Rmd
Updates appendices

@@
   2.2 Data  
+  2.3 Individual Effect Correlation 
 3. Fit model 
@@
 We use the target specification as it saves the constants, necessary for model comparison later on.
-For example, `target += normal_lpdf(delta2 | mu4, sqrt(g4))` basically means that "delta2" is normally distributed with a mean of "mu4" and a standard deviation of "g4".
+For example, `target += normal_lpdf(delta2 | mu4, sqrt(g4))` means that "delta2" is normally distributed with a mean of "mu4" and a standard deviation of "g4".
 Note that we want to estimate the variances, not the standard deviations.
@@
 
-### Individual Effect Correlation
+### \textcolor{blue}{Individual Effect Correlation}
 
@@
 As stated in the manuscript, it is also possible to allow for the estimation of the correlation between individual effects.
-This means, for instance, that we would expect that the estimates for the individual effects $\delta_{7,i}$ and $\delta_{3,i}$ are related.
-To allow for correlation between individual effects, we would have to adjust the formula in the `stan` file and the priors.
-For the model with correlation terms, we use the file `myModel_cor.stan` that can be found in the folder} *K - R objects > rstan > normal model > correlation*.
-\textcolor{Notice that we apply a different conceptualization of the model using matrix notation.
-The effects are called `beta` and the individual effects are called `beta_p`.
-Because we are using matrices, the effects are stored a bit differently than for the other model.
-For example, `beta[3]` represents the general digit effect $\delta_{7}$, while `beta_p[1,3]` represents the digit effect for participant 1, $\delta_{7,1}$.}
+This means, for instance, that we would expect that the estimates for the individual effects} $\delta_{7,i}$ and $\delta_{3,i}$ \textcolor{blue}{are related.
+To allow for correlation between individual effects, we would have to adjust the formula in the} `stan` \textcolor{blue}{file and the priors.
+For the model with correlation terms, we use the file} `myModel_cor.stan` \textcolor{blue}{that can be found in the folder} *K - R objects > rstan > normal model > correlation*.
+\textcolor{blue}{Notice that we apply a different conceptualization of the model using matrix notation.
+The effects are called} `beta` \textcolor{blue}{and the individual effects are called} `beta_p`.
+\textcolor{blue}{Because we are using matrices, the effects are stored a bit differently than for the other model.
+For example,} `beta[3]` \textcolor{blue}{represents the general digit effect} $\delta_{7}$, \textcolor{blue}{while} `beta_p[1,3]` \textcolor{blue}{represents the digit effect for participant 1,} $\delta_{7,1}$.}
 
@@
 In specific, we set a prior on the correlation matrices.
-These priors will be transformed for efficiency and set on so called Cholesky factors of the matrices by the following line of code `Lcorr ~ lkj_corr_cholesky(1)` [for more information on Cholesky factors see} @stan2018a; \textcolor{blue}{section 1.13].
+These priors will be transformed for efficiency and set on so called Cholesky factors of the matrices by the following line of code} `Lcorr ~ lkj_corr_cholesky(1)` \textcolor{blue}{[for more information on Cholesky factors see} @stan2018a; \textcolor{blue}{section 1.13].
 }
@@
 
-To estimate the model with the correlation between random effects, note that one would need to adjust the `file` argument to equal and `myModel_cor.stan`.
-
-Back to our model without the correlation term form the manuscript.
-We have already fitted the model and saved it in the R object `model_fit`.
+\textcolor{blue}{To estimate the model with the correlation between individual effects, note that one would need to adjust the} `file` \textcolor{blue}{argument to equal} `myModel_cor.stan`.
+
+\textcolor{blue}{Back to our model without the correlation term form the manuscript.} We have already fitted the model and saved it in the R object `model_fit`.
 In case you do not want to wait until the model has been fitted but continue immediately with the rest of this tutorial, you can load this R object that contains the model estimates.
@@
   theme_classic() + 
-  coord_cartesian(xlim = c(0.75, 5), ylim = c(-0.05, 0.12))
+  coord_cartesian(xlim = c(0.75, 5), ylim = c(-0.1, 0.15))
 

### 2023-06-29 D - Tutorial Normal Model brms/Online Supplement D - Bayesian Hierarchical Modeling in brms.Rmd
Updates appendices

--- 2023-06-28 D - Tutorial Normal Model brms/Online Supplement D - Bayesian Hierarchical Modeling in brms.Rmd
@@
   2.4 Priors  
+  2.5 Individual Effect Correlation 
 3. Fit model 
@@
 
-### Individual Effect Correlation
-
-Up until now, we discussed the normal model from the manuscript.
+### \textcolor{blue}{Individual Effect Correlation}
+
+\textcolor{blue}{Up until now, we discussed the normal model from the manuscript.
 This model assumes that there is no correlation between the individual effects.
 As stated in the manuscript, it is also possible to allow for the estimation of the correlation between individual effects.
-This means, for instance, that we would expect that the estimations for the individual effects $\delta_{7,i}$ and $\delta_{3,i}$ are related.
+This means, for instance, that we would expect that the estimations for the individual effects} $\delta_{7,i}$ \textcolor{blue}{and} $\delta_{3,i}$ \textcolor{blue}{are related.
 To allow for correlation between individual effects, we would have to adjust the model formula.
-For the estimation of the digit classification task, we could use the following formula:
+For the estimation of the digit classification task, we could use the following formula:}
 
@@
 
-The difference is that instead of using the double bar `||`, we are now using a single bar `|` before `ind`, indicating that we allow individual effects to correlate.
-
-This means that we are estimating more parameters and therefore we have to specify priors on these parameters as well.
-In specific, we can set a prior on the correlation matrix called `class = "cor"`.
-These priors will be transformed for efficiency and set on so called Cholesky factors of the matrices [for more information see @stan2018a; section 1.13].
-In our example, we set a prior on the Cholesky factors directly, by defining `class = "L"`, shown in the last line of the code block:
+\textcolor{blue}{The difference is that instead of using the double bar} `||`\textcolor{blue}{, we are now using a single bar} `|` \textcolor{blue}{before} `ind`\textcolor{blue}{, indicating that we allow individual effects to correlate.}
+
+\textcolor{blue}{This means that we are estimating more parameters and therefore we have to specify priors on these parameters as well.
+In specific, we can set a prior on the correlation matrix called} `class = "cor"`\textcolor{blue}{.
+These priors will be transformed for efficiency and set on so called Cholesky factors of the matrices [for more information see} @stan2018a; \textcolor{blue}{section 1.13].
+In our example, we set a prior on the Cholesky factors directly, by defining} `class = "L"`\textcolor{blue}{, shown in the last line of the code block:}
 
@@
 
-To fit the model with the correlation between random effects, note that one would need to adjust `formula` and `priorsmodel` to the objects where these are saved in (i.e., `bform_cor` and `priorsmodel_cor`).
-
-Back to our model without the correlation term form the manuscript.
-We have already estimated this model and saved it in the R object `brmmodel8_new03012022.rds`.
+\textcolor{blue}{To estimate the model with the correlation between individual effects, note that one would need to adjust} `formula` \textcolor{blue}{and} `priorsmodel` \textcolor{blue}{to the objects where these are saved in (i.e.,} `bform_cor` \textcolor{blue}{and} `priorsmodel_cor`\textcolor{blue}{).}
+
+\textcolor{blue}{Back to our model without the correlation term form the manuscript.} We have already estimated this model and saved it in the R object `brmmodel8_new03012022.rds`.
 In case you do not want to wait until the model has been fitted but continue immediately with the rest of this tutorial, you can load this R object that contains the model estimations.
@@
 
+\textcolor{blue}{We can do the same for the model with correlation between the individual effects (see Figure 7) and see that the estimated effects are somewhat different.}
+
+```{r modeldigitestplotcor, echo=FALSE, fig.cap="Model estimates for digit effect parameters in the normal model with correlation between individual effects.
+The points represent the mean parameter estimates for each individual.
+The violin plots on the right side show the variance of the individual parameter estimates.", fig.align = "left", fig.height = 11, fig.width = 1, out.width="75%"}
+knitr::include_graphics("Fig12_modelestimatefigure2_onlybrms_cor.png")
+```
+
 ## Model comparison

### 2023-06-29 E - Tutorial Log-Normal Model Rstan/Online Supplement E - Bayesian Hierarchical Modeling in rstan - Log Model.Rmd
Updates cor model

@@
 
-|  The stan file `myLogModel.stan` is used to fit the full model and can be found under R objects > rstan > linear model.
+|  The stan file `myLogModel.stan` is used to fit the full model and can be found under K - R objects > rstan > linear model.
 It contains descriptions for every line of code.
@@
 setwd('..')
-logmodel_fit <- readRDS(file = "R objects/rstan/log-normal model/hier_logmodelc_adj13012022.rds")
+logmodel_fit <- readRDS(file = "K - R objects/rstan/log-normal model/hier_logmodelc_adj13012022.rds")
 ```

### 2023-06-29 F - Tutorial Log-Normal Model brms/Online Supplement F - Bayesian Hierarchical Modeling in brms - Log Model.Rmd
Updates cor model

@@
 setwd('..')
-logmodel_fit <- readRDS(file = "R objects/brms/log-normal/brmmodel8log13012022.rds")
+logmodel_fit <- readRDS(file = "K - R objects/brms/log-normal/brmmodel8log13012022.rds")
 
 # For model comparison - sample_prior was set to "yes" 
-logmodel_fit_bf <- readRDS("R objects/brms/BF/brmmodel8log26052022_bf.rds")
+logmodel_fit_bf <- readRDS("K - R objects/brms/BF/brmmodel8log26052022_bf.rds")
 

### 2023-06-29 H - Posterior Distributions Tables/Online Supplement H - Tables with Information on Posterior Distributions.Rmd
Updates cor model

@@
 setwd('..')
-model_fit <- readRDS(file = "R objects/brms/Truncated/BRMSmodelfit18_03012022.rds")
+model_fit <- readRDS(file = "K - R objects/brms/Truncated/BRMSmodelfit18_03012022.rds")
 
@@
 setwd('..')
-logmodel_fit <- readRDS(file = "R objects/brms/log-normal/brmmodel8log13012022.rds")
+logmodel_fit <- readRDS(file = "K - R objects/brms/log-normal/brmmodel8log13012022.rds")
 

### 2023-06-29 J - R code/Figures/Fig1.R

@@ -1,7 +1,7 @@
 # Efron & Moris figure 
 # Model 
 # Load output BayesFactor package: 
-samplesindat5 <- readRDS("R objects/BF model to create figure 1B/BFmodelfit15.rds")
+samplesindat5 <- readRDS("K - R objects/BF model to create figure 1B/BFmodelfit15.rds")
 meanbfindat <- colMeans(samplesindat5)  # now BF model output with seperate effect mu's and different r scale (all set to 1)
 
 # From data, from proposal r code 

### 2023-06-29 J - R code/Figures/Fig10_randomparfigure_trunc_ordered_per_individuals_rev.R

@@ -1,10 +1,10 @@
 # Individual effects from Rstan 
 # Order on participant 
 # For rstan the output is saved in hier_modelc_adj_07012022 (adjust "myPath" below to your directory)
-hier_modelc_adj_07012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
+hier_modelc_adj_07012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
 
 # For brms the output is saved in object brmmodel8_new03012022 (this is without saving the prior samples)
-brmmodel8_new03012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/brms/Posterior/brmmodel8_new03012022.rds")  
+brmmodel8_new03012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/brms/Posterior/brmmodel8_new03012022.rds")  
 
 library(rstan)
 library(brms)

### 2023-06-29 J - R code/Figures/Fig10_randomparfigure_trunc_v2.R

@@ -1,9 +1,9 @@
 # Individual effects from Rstan 
 # For rstan the output is saved in hier_modelc_adj_07012022 (adjust "myPath" below to your directory)
-hier_modelc_adj_07012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
+hier_modelc_adj_07012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
 
 # For brms the output is saved in object brmmodel8_new03012022 (this is without saving the prior samples)
-brmmodel8_new03012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/brms/Posterior/brmmodel8_new03012022.rds")  
+brmmodel8_new03012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/brms/Posterior/brmmodel8_new03012022.rds")  
 
 library(rstan)
 library(brms)

### 2023-06-29 J - R code/Figures/Fig11_modelestimatesfigure_truncated_onlyrstan.R

@@ -1,7 +1,7 @@
 # Model estimates figure 
 # Load fits 
 # For rstan the output is saved in hier_modelc_adj_07012022 (adjust "myPath" below to your directory)
-hier_modelc_adj_07012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
+hier_modelc_adj_07012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
 
 library(rstan)
 

### 2023-06-29 J - R code/Figures/Fig11b_modelestimatesfigure_trunc_v2.R

@@ -1,10 +1,10 @@
 # Model estimates figure 
 # Load fits 
 # For rstan the output is saved in hier_modelc_adj_07012022 (adjust "myPath" below to your directory)
-hier_modelc_adj_07012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
+hier_modelc_adj_07012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
 
 # For brms the output is saved in object brmmodel8_new03012022 (this is without saving the prior samples)
-brmmodel8_new03012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/brms/Posterior/brmmodel8_new03012022.rds")  
+brmmodel8_new03012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/brms/Posterior/brmmodel8_new03012022.rds")  
 
 library(rstan)
 library(brms)

### 2023-06-29 J - R code/Figures/Fig12_chainrhatneff_v2.R

@@ -7,10 +7,10 @@ library(rstan)
 
 # Load data 
 # For rstan the output is saved in hier_logmodelc_adj13012022  (adjust "myPath" below to your directory)
-hier_logmodelc_adj13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/rstan/log-normal model/hier_logmodelc_adj13012022.rds")
+hier_logmodelc_adj13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/rstan/log-normal model/hier_logmodelc_adj13012022.rds")
 
 # For brms the output is saved in object brmmodel8log13012022 (this is without saving the prior samples)
-brmmodel8log13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/brms/log-normal/brmmodel8log13012022.rds")  
+brmmodel8log13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/brms/log-normal/brmmodel8log13012022.rds")  
 
 tracerhatnefflog <- function(hier_logmodelc_adj30032021, brmmodel8log02042021){
 #Rstan

### 2023-06-29 J - R code/Figures/Fig12_indest_rstancorindv.R

@@ -1,0 +1,116 @@
+# Figure 
+
+# Model estimates figure 
+# Load fits 
+# For rstan the output is saved in hier_modelc_cor_06042023 (adjust "myPath" below to your directory)
+hier_modelc_cor_29062023 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/rstan/normal model/correlation term/hier_modelc_cor_29062023.rds")
+
+library(rstan)
+
+# Plot results using https://github.com/jorvlan/open-visualizations/blob/master/R/repmes_tutorial_R.pdf
+# Install packages for plot 
+packages <- c("plyr", "lattice", "ggplot2", "dplyr", "readr", "rmarkdown","Rmisc")
+if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
+  install.packages(setdiff(packages, rownames(installed.packages())))  
+}
+if (!require(devtools)) {
+  install.packages("devtools")
+}
+devtools::install_github('erocoar/gghalves')
+
+library("plyr")
+library("lattice")
+library("ggplot2")
+library("dplyr")
+library("readr")
+library("rmarkdown")
+library("Rmisc")
+library("devtools")
+library("gghalves")
+
+indvestviolin_onlyrstan <- function(hier_modelc_adj_v2, brmmodel8_v2){
+  # Rstan 
+  ## Delta's 
+  ## Create dataframe 
+  parameter_m7 <- rep(c("Delta 1", "Delta 2", "Delta 3", "Delta 4"), each = 52 )
+  
+  ### Obtain all individual estimates delta 
+  hier_modeld <- hier_modelc_adj_v2
+  allpar_summary <- summary(hier_modeld, pars = c("beta_p"), probs = c(0.1, 0.9))$summary
+  
+  del1cor <- seq(3, 309, by = 6)
+  del2cor <- seq(4, 310, by = 6)
+  del3cor <- seq(5, 311, by = 6)
+  del4cor <- seq(6, 312, by = 6)
+  
+  delta1_summary <- allpar_summary[del1cor,] 
+  
+  delta2_summary <- allpar_summary[del2cor,]
+  
+  delta3_summary <- allpar_summary[del3cor,]
+  
+  delta4_summary <- allpar_summary[del4cor,]
+  
[…]

### 2023-06-29 J - R code/Figures/Fig12_indvestbrms_cor.R

@@ -1,0 +1,85 @@
+# Brms ind estimates for model with cor between indv estimates 
+# For brms the output is saved in brmmodel9_new28062023.rds (adjust "myPath" below to your directory)
+brmmodel9_new28062023 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/brms/normal model/correlation/brmmodel9_new28062023.rds")
+  
+  
+
+fig12_brmscor <- function(model_fit){
+  ### Delta's individual estimates 
+  estimatesbrms <- coef(model_fit)$sub1
+  #### Delta 1 - difference 8 and 7 
+  del1estimatesbrms <- estimatesbrms[,,"a_dif1"]
+  
+  #### Delta 2 - difference 7 and 6 
+  del2estimatesbrms <- estimatesbrms[,,"a_dif2"]
+  
+  #### Delta 3 - difference 4 and 3
+  del3estimatesbrms <- estimatesbrms[,,"a_dif3"]
+  
+  #### Delta 4 - difference 3 and 2 
+  del4estimatesbrms <- estimatesbrms[,,"a_dif4"]
+  
+  #### create dataframe 
+  parameter_m7 <- rep(c("Delta 1", "Delta 2", "Delta 3", "Delta 4"), each = 52 )
+  
+  #### gamma_summary2 not using for plot 
+  #### beta_summary2 not using for plot 
+  parest_m7brms <- rbind(del1estimatesbrms[,1], del2estimatesbrms[,1], del3estimatesbrms[,1], del4estimatesbrms[,1])
+  parest_m7brms2 <- t(parest_m7brms)
+  parest_m7brms3 <- as.data.frame(parest_m7brms2)
+  parest_m7brms4 <- c(parest_m7brms3$V1, parest_m7brms3$V2, parest_m7brms3$V3, parest_m7brms3$V4)
+  datafr_m7brms <- data.frame(parameter_m7, parest_m7brms4)
+  
+  # Plot 
+  set.seed(321)
+  datafr_m7brms$x <- rep(c(1, 2, 3, 4), each = 52)
+  datafr_m7brms$xj <- jitter(datafr_m7brms$x, amount = 0.09)
+  datafr_m7brms$id <- rep(c(1:52), 4)
+  
+  plotmodelest_m7brms <- ggplot(data=datafr_m7brms, aes(y=parest_m7brms4)) +
+    
+    #Add geom_() objects
[…]

### 2023-06-29 J - R code/Figures/Fig13_posteriorlog_v2.R

@@ -1,10 +1,10 @@
 # Figure 13
 # Posterior distribution parameters - lognormal model 
 # For rstan the output is saved in hier_logmodelc_adj13012022 (adjust "myPath" below to your directory)
-hier_logmodelc_adj13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/rstan/log-normal model/hier_logmodelc_adj13012022.rds")
+hier_logmodelc_adj13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/rstan/log-normal model/hier_logmodelc_adj13012022.rds")
 
 # For brms the output is saved in object brmmodel8log13012022 (this is without saving the prior samples)
-brmmodel8log13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/brms/log-normal/brmmodel8log13012022.rds")  
+brmmodel8log13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/brms/log-normal/brmmodel8log13012022.rds")  
 
 
 # Used this link: https://cran.r-project.org/web/packages/bayesplot/vignettes/plotting-mcmc-draws.html 

### 2023-06-29 J - R code/Figures/Fig14_randomparlog_v2.R

@@ -2,10 +2,10 @@
 # Individual effects from Rstan 
 # Load data 
 # For rstan the output is saved in hier_logmodelc_adj13012022 (adjust "myPath" below to your directory)
-hier_logmodelc_adj13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/rstan/log-normal model/hier_logmodelc_adj13012022.rds")
+hier_logmodelc_adj13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/rstan/log-normal model/hier_logmodelc_adj13012022.rds")
 
 # For brms the output is saved in object brmmodel8log13012022 (this is without saving the prior samples)
-brmmodel8log13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/brms/log-normal/brmmodel8log13012022.rds")  
+brmmodel8log13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/brms/log-normal/brmmodel8log13012022.rds")  
 
 library(rstan)
 library(brms)

### 2023-06-29 J - R code/Figures/Fig15b_lognormaldifindest_v2.R

@@ -2,10 +2,10 @@
 
 # Load data 
 # For rstan the output is saved in hier_logmodelc_adj13012022 (adjust "myPath" below to your directory)
-hier_logmodelc_adj13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/rstan/log-normal model/hier_logmodelc_adj13012022.rds")
+hier_logmodelc_adj13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/rstan/log-normal model/hier_logmodelc_adj13012022.rds")
 
 # For brms the output is saved in object brmmodel8log13012022 (this is without saving the prior samples)
-brmmodel8log13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/brms/log-normal/brmmodel8log13012022.rds")  
+brmmodel8log13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/brms/log-normal/brmmodel8log13012022.rds")  
 
 library(brms)
 library(rstan)

### 2023-06-29 J - R code/Figures/Fig16_savagedickey_trunc.R

@@ -4,7 +4,7 @@ library(brms)
 
 # Load brms output 
 # For brms the output is saved in object brmmodel8_new_bf_03012022 (this is without saving the prior samples)
-brmmodel8_new_bf_03012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/brms/BF/brmmodel8_new_bf_03012022.rds") 
+brmmodel8_new_bf_03012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/brms/BF/brmmodel8_new_bf_03012022.rds") 
 
 savagedickeyfigure <- function(brmmodel8bf){
   # Obtain Savage-Dickey ratio 

### 2023-06-29 J - R code/Figures/Fig17_bridgesamplingresults_v2.R

@@ -5,7 +5,7 @@ library(rstan )
 
 # Load results
 # rstan
-resultsbsrstan <- read.csv("/myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/Model comparison/Bayes Factors per package/bssresults_07012022.csv")
+resultsbsrstan <- read.csv("/myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/Model comparison/Bayes Factors per package/bssresults_07012022.csv")
 
 ## Restructure 
 resultsbsrstan$package <- rep("Rstan", nrow(resultsbsrstan))

### 2023-06-29 J - R code/Figures/Fig8_traceplotsfigure_trunc_v2.R

@@ -1,10 +1,10 @@
 # Traceplots dif 2 (difference between 6 and 7) + rhat
 # Based on output each package of symbolic distance effect model 
 # For rstan the output is saved in hier_modelc_adj_07012022 (adjust "myPath" below to your directory)
-hier_modelc_adj_07012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
+hier_modelc_adj_07012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
 
 # For brms the output is saved in object brmmodel8_new03012022 (this is without saving the prior samples)
-brmmodel8_new03012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/brms/Posterior/brmmodel8_new03012022.rds")  
+brmmodel8_new03012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/brms/Posterior/brmmodel8_new03012022.rds")  
 
 tracerhatneff <- function(hier_modelc_adj_v2, brmmodel8_v2){
   #Rstan

### 2023-06-29 J - R code/Figures/Fig9_posteriorfigure_trunc_v2.R

@@ -1,9 +1,9 @@
 # Posterior distribution parameters 
 # For rstan the output is saved in hier_modelc_adj_07012022 (adjust "myPath" below to your directory)
-hier_modelc_adj_07012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
+hier_modelc_adj_07012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
 
 # For brms the output is saved in object brmmodel8_new03012022 (this is without saving the prior samples)
-brmmodel8_new03012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/brms/Posterior/brmmodel8_new03012022.rds")  
+brmmodel8_new03012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/brms/Posterior/brmmodel8_new03012022.rds")  
 
 # Used this link: https://cran.r-project.org/web/packages/bayesplot/vignettes/plotting-mcmc-draws.html 
 

### 2023-06-30 manuscript.Rmd
Updates (replaced and removed files)

@@
 \Revision{Up until now, we specified models without a correlation between the individual effects.
-This means, for instance, that we do not expect that the estimations for the individual effects $\delta_{7, i}$ and $\delta_{3, i}$ to be related.
+This means, for instance, that we do not expect the estimations for the individual effects $\delta_{7, i}$ and $\delta_{3, i}$ to be related.
 However, in cases where we would expect the estimation of individual effects to be related, we could estimate the correlation coefficient between individual effects.
@@
 For all three \\textit{t}-distributions, the degrees of freedom are set to 3.", fig.align = "left", fig.height = 8, fig.width = 12, out.width="100%"}
-knitr::include_graphics("I - Images/Fig4_distrtypefigure_v3.png")
+knitr::include_graphics("I - Images/Fig5_distrtypefigure_v3.png")
 ```
@@
 Panel D shows the prior predictions for the log-normal model based on the selected priors.", fig.align = "left", fig.height = 3, fig.width = 4, out.width="100%"}
-knitr::include_graphics("I - Images/Fig5_priorpredfigure_v2.png")
+knitr::include_graphics("I - Images/Fig6_priorpredfigure_v2.png")
 ```
@@
 The stars represent the number of effective samples of the general effects."}
-knitr::include_graphics("I - Images/Fig8_traceplots3pnew_v2.png")
+knitr::include_graphics("I - Images/Fig9_traceplots3pnew_v2.png")
 ```
@@
 The shaded area within the distributions represent 95\\% credible intervals.", fig.align = "left", fig.height = 11, fig.width = 1, out.width="100%"}
-knitr::include_graphics("I - Images/Fig9_posteriormu_v3.png")
+knitr::include_graphics("I - Images/Fig10_posteriormu_v3.png")
 ```
@@
 Pink intervals contain zero, blue interval do not contain zero.", fig.align = "left", fig.height = 1, fig.width = 0.5, out.width="100%"}
-knitr::include_graphics("I - Images/Fig10_randompar_v2.png")
+knitr::include_graphics("I - Images/Fig11_randompar_v2.png")
 
@@
 ", fig.align = "left", fig.height = 11, fig.width = 1, out.width="75%"}
-knitr::include_graphics("I - Images/Fig11_modelestimatefigure2_onlyrstan_v2.png")
+knitr::include_graphics("I - Images/Fig12_modelestimatefigure2_onlyrstan_v2.png")
 ```
@@
 ```{r traceplotlog, echo=FALSE, fig.cap="(ref:captiontraceplotlog)", fig.align = "left", fig.height = 11, fig.width = 4, out.width="100%"}
-knitr::include_graphics("I - Images/Fig12_traceplots3plog_v2.png")
+knitr::include_graphics("I - Images/Fig13_traceplots3plog_v2.png")
 ```
@@
 The shaded area within the distributions represent 95\\% credible intervals.", fig.align = "left", fig.height = 11, fig.width = 1, out.width="100%"}
-knitr::include_graphics("I - Images/Fig13_posteriormulog_v2.png")
+knitr::include_graphics("I - Images/Fig14_posteriormulog_v2.png")
 ```
@@
 Pink intervals contain zero, blue interval do not contain zero.", fig.align = "left", fig.height = 1, fig.width = 0.5, out.width="100%"}
-knitr::include_graphics("I - Images/Fig14_randomparlog_v2.png")
+knitr::include_graphics("I - Images/Fig15_randomparlog_v2.png")
 ```
@@
 The violin plots on the right side show the variance of the individual parameter estimates.", fig.align = "left", fig.height = 11, fig.width = 1, out.width="75%"}
-knitr::include_graphics("I - Images/Fig15_inddiflogplot_onlyrstan_v2.png")
+knitr::include_graphics("I - Images/Fig16_inddiflogplot_onlyrstan_v2.png")
 ```
@@
 A Bayes factor above one indicates that the evidence is in favor of the null hypothesis.", fig.align = "left", fig.height = 6, fig.width = 4, out.width="60%"}
-knitr::include_graphics("I - Images/Fig16_sdrplotside_v2.png")
+knitr::include_graphics("I - Images/Fig17_sdrplotside_v2.png")
 ```
@@
 ", fig.align = "left", fig.height = 11, fig.width = 4, out.width="70%"}
-knitr::include_graphics("I - Images/Fig17_normalbfplot01.png")
+knitr::include_graphics("I - Images/Fig18_normalbfplot01.png")
 ```
@@
 We evaluate their performance in two scenarios that correspond to scenario 1 and 3 for the normal model.
-This means that in scenario 1 for the log-normal model, we estimated a null, digit, side and full model based on the prior specified in Equation, specifying large individual differences.
+This means that in scenario 1 for the log-normal model, we estimated a null, digit, side and full model based on the prior specified in Equation 8, specifying large individual differences.
 For scenario 2, we remove individual variability from the side, digit, and full model.
@@
 While individual differences are a common phenomenon in psychology, there are instances where general effects are sufficient to describe the data adequately, or where only a random intercept, but no random slopes are necessary.
-As shown in Tables} \@ref(tab:postmodprob) and \@ref(tab:postmodproblog), Bayesian model comparisons can help researchers to identify the model that best describes the data.
+As shown in Tables \@ref(tab:postmodprob) and \@ref(tab:postmodproblog), Bayesian model comparisons can help researchers to identify the model that best describes the data.
 

### 2023-06-30 BHM_correlation/Brms/brms_correlation.R

@@ -1,346 +1,0 @@
-# New brms fit correlation
-
-library(brms)
-
-################################################################################
-## 1. Load data 
-################################################################################
-indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
-colnames(indat)=c('sub','block','trial','stim','resp','rt','error')
-
-## Cleaning the data according to criteria discussed in paper 
-clean=function()
-{
-  indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
-  colnames(indat)=c('sub','block','trial','stim','resp','rt','error')
-  
-  bad1=indat$sub%in%c(34,43)
-  bad2=indat$rt<250 | indat$rt>2000
-  bad3=indat$err==1
-  bad4=indat$block==0 & indat$trial<20
-  bad5=indat$trial==0
-  
-  bad=bad1 | bad2 | bad3 |bad4 |bad5
-  dat=indat[!bad,]
-  return(dat)
-}
-
-indat1 <- clean()  # Final dataset 
-
-## Data add necessary information 
-### Add participant number 
-rank(unique(indat1$sub))
-sub <- rep(1:N, each = J)
-true.theta[sub]
-
-# create variable inter for the intercept gamma 
-indat1$inter <- rep(1, nrow(indat1))
-
-for (j in 1:nrow(indat1)){
-  if (j == 1) {
-    indat1$sub1[j] <- 1}
-  else if (indat1$sub[j] == indat1$sub[j-1]) {
-    indat1$sub1[j] <- indat1$sub1[j-1]}
-  else indat1$sub1[j] <- indat1$sub1[j-1] + 1
-}
-
-### Add variable with information of the side of the digit to the list, for parameter beta    
-# j = condition, if j > 5 than x = -1/2, if j < 5 than x = 1/2 
-# indat1$stim 0 = 2, 1=3, 2=4, 3=6, 4=7, 5=8 
-# so if bigger than 2 than 1/2, else than -1/2 
-for (j in 1:nrow(indat1)){
-  if (indat1$stim[j] < 3) {
-    indat1$side[j] <- 1/2}
-  else indat1$side[j] <- -1/2
-}
-
-### Add variable with information about difference between digits to the list, for parameters delta's 
-#### Difference 8 and 7 
[…]

### 2023-06-30 BHM_correlation/Rstan/BHM_rstan_correlation.R

@@ -1,172 +1,0 @@
-# Estimation rstan multilevel with correlation between random effects 
-
-# Fitting an Bayesian Hierarchical Model using Stan - An example 
-
-################################################################################
-## 1. Load data 
-################################################################################
-indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
-colnames(indat)=c('sub','block','trial','stim','resp','rt','error')
-
-## Cleaning the data according to criteria discussed in paper 
-clean=function()
-{
-  indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
-  colnames(indat)=c('sub','block','trial','stim','resp','rt','error')
-  
-  bad1=indat$sub%in%c(34,43)
-  bad2=indat$rt<250 | indat$rt>2000
-  bad3=indat$err==1
-  bad4=indat$block==0 & indat$trial<20
-  bad5=indat$trial==0
-  
-  bad=bad1 | bad2 | bad3 |bad4 |bad5
-  dat=indat[!bad,]
-  return(dat)
-}
-
-indat1 <- clean()  # Final dataset 
-
-## Data into list and fill the list with necessary information 
-datareal <- list(N = length(unique(indat1$sub)))
-datareal$All <- nrow(indat1)  # number of total observations, adding to the list for the analysis  
-
-### Indicating which observation belongs to which person 
-sub <- as.numeric(factor(indat1$sub))
-
-# Need a variable with group number that are crhonical (that doesn't skip numbers)
-for (j in 1:nrow(indat1)){
-  if (j == 1) {
-    indat1$sub1[j] <- 1}
-  else if (indat1$sub[j] == indat1$sub[j-1]) {
-    indat1$sub1[j] <- indat1$sub1[j-1]}
-  else indat1$sub1[j] <- indat1$sub1[j-1] + 1
-}
-
-
-datareal$group_inds <- indat1$sub1  # Add variable with indicator person to the list used for the analysi
-datareal$y <- indat1$rt/1000  # rt in seconds instead of miliseconds, added to the list for analysis  
-
[…]

### 2023-06-30 J - R code/Brms/Normal model/correlation/brms_correlation.R

@@ -1,0 +1,166 @@
+# New brms fit correlation
+
+library(brms)
+
+################################################################################
+## 1. Load data 
+################################################################################
+indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
+colnames(indat)=c('sub','block','trial','stim','resp','rt','error')
+
+## Cleaning the data according to criteria discussed in paper 
+clean=function()
+{
+  indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
+  colnames(indat)=c('sub','block','trial','stim','resp','rt','error')
+  
+  bad1=indat$sub%in%c(34,43)
+  bad2=indat$rt<250 | indat$rt>2000
+  bad3=indat$err==1
+  bad4=indat$block==0 & indat$trial<20
+  bad5=indat$trial==0
+  
+  bad=bad1 | bad2 | bad3 |bad4 |bad5
+  dat=indat[!bad,]
+  return(dat)
+}
+
+indat1 <- clean()  # Final dataset 
+
+## Data add necessary information 
+### Add participant number 
+rank(unique(indat1$sub))
+sub <- rep(1:N, each = J)
+true.theta[sub]
+
+# create variable inter for the intercept gamma 
+indat1$inter <- rep(1, nrow(indat1))
+
+for (j in 1:nrow(indat1)){
+  if (j == 1) {
+    indat1$sub1[j] <- 1}
+  else if (indat1$sub[j] == indat1$sub[j-1]) {
+    indat1$sub1[j] <- indat1$sub1[j-1]}
+  else indat1$sub1[j] <- indat1$sub1[j-1] + 1
+}
+
+### Add variable with information of the side of the digit to the list, for parameter beta    
+# j = condition, if j > 5 than x = -1/2, if j < 5 than x = 1/2 
+# indat1$stim 0 = 2, 1=3, 2=4, 3=6, 4=7, 5=8 
+# so if bigger than 2 than 1/2, else than -1/2 
+for (j in 1:nrow(indat1)){
+  if (indat1$stim[j] < 3) {
+    indat1$side[j] <- 1/2}
+  else indat1$side[j] <- -1/2
+}
+
+### Add variable with information about difference between digits to the list, for parameters delta's 
+#### Difference 8 and 7 
+for (j in 1:nrow(indat1)){ 
[…]

### 2023-06-30 J - R code/Figures/Fig10_posteriorfigure_trunc_v2.R

@@ -1,0 +1,81 @@
+# Posterior distribution parameters 
+# For rstan the output is saved in hier_modelc_adj_07012022 (adjust "myPath" below to your directory)
+hier_modelc_adj_07012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
+
+# For brms the output is saved in object brmmodel8_new03012022 (this is without saving the prior samples)
+brmmodel8_new03012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/brms/Posterior/brmmodel8_new03012022.rds")  
+
+# Used this link: https://cran.r-project.org/web/packages/bayesplot/vignettes/plotting-mcmc-draws.html 
+
+posterplot <- function(hier_modelc_adj_v2, brmmodel8_v2){
+  library(tidybayes)
+  library(bayesplot)
+  library(rstan)
+  library(brms)
+  library(ggplot2)
+  library(papaja)
+  
+  post8 <- as.array(brmmodel8_v2)
+  dimnames(post8)
+  color_scheme_set("orange")
+  pairs8 <- mcmc_areas(
+    post8, 
+    pars = c("b_a_side", "b_a_dif1", "b_a_dif2", "b_a_dif3", "b_a_dif4"),
+    prob = 0.95, # 80% intervals
+    prob_outer = 0.99, # 99%
+    point_est = "mean"
+  )
+  # scale_x_continuous(breaks=c(1,2,3,4), labels=c(expression(~delta[~7-8]), expression(~delta[~6-7]), expression(~delta[~4-3]), expression(~delta[~3-2]))) +
+  
+  pairs82 <- pairs8 + 
+    theme_apa() + 
+    ggtitle("Brms") + 
+    xlim(-0.035, 0.085) + 
+    scale_x_continuous(breaks = c(-0.025, 0, 0.025, 0.05, 0.075)) + 
+    scale_y_discrete(labels= c( 
+      expression(~mu[~beta]), 
+      expression(~mu[~delta[~7]]), 
+      expression(~mu[~delta[~6]]), 
+      expression(~mu[~delta[~4]]), 
+      expression(~mu[~delta[~3]]))) + 
+    theme(axis.title.y=element_blank(),
[…]

### 2023-06-30 J - R code/Figures/Fig10_randomparfigure_trunc_ordered_per_individuals_rev.R

@@ -1,680 +1,0 @@
-# Individual effects from Rstan 
-# Order on participant 
-# For rstan the output is saved in hier_modelc_adj_07012022 (adjust "myPath" below to your directory)
-hier_modelc_adj_07012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
-
-# For brms the output is saved in object brmmodel8_new03012022 (this is without saving the prior samples)
-brmmodel8_new03012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/brms/Posterior/brmmodel8_new03012022.rds")  
-
-library(rstan)
-library(brms)
-library(gridExtra)
-
-indeffplot_orderedind <- function(hier_modelc_adj_v2, brmmodel8_v2){
-  
-  # Only gamma's 
-  gamnamesstan <- paste("gammai", "[",1:52,"]", sep="")
-  gamstan <- as.data.frame(summary(hier_modelc_adj_v2, pars = c( gamnamesstan))$summary)
-  gamstanmu <- as.data.frame(summary(hier_modelc_adj_v2, pars = c("mu"))$summary)
-  gamstan$parameters <- c(gamnamesstan)
-  
-  for (i in 1:nrow(gamstan)){  # 1 if CI contains zero, 0 otherwise 
-    if (gamstan$`2.5%`[i]*gamstan$`97.5%`[i] < 0) {
-      gamstan$zero2[i] <- "pink"
-    } else (gamstan$zero2[i] <- "blue") 
-  }
-  
-  
-  gamstan2 <- gamstan[order(gamstan$mean),]
-  gamstan$order <- c(1:52)
-  
-  gammarandomplot <- ggplot(gamstan, aes(x = order, y = mean, ymin = `2.5%`, ymax = `97.5%`)) + 
-    geom_linerange(aes(color = "blue"), color = "#00BFC4") + 
-    geom_point() + 
-    geom_hline(yintercept = gamstanmu$mean, linetype = "dashed", alpha = .3) + 
-    labs(title = expression(~gamma[~i]), y = "Estimation") + 
-    theme(legend.position = "none", 
-          axis.title.x=element_blank(),
[…]

### 2023-06-30 J - R code/Figures/Fig10_randomparfigure_trunc_v2.R

@@ -1,345 +1,0 @@
-# Individual effects from Rstan 
-# For rstan the output is saved in hier_modelc_adj_07012022 (adjust "myPath" below to your directory)
-hier_modelc_adj_07012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
-
-# For brms the output is saved in object brmmodel8_new03012022 (this is without saving the prior samples)
-brmmodel8_new03012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/brms/Posterior/brmmodel8_new03012022.rds")  
-
-library(rstan)
-library(brms)
-library(gridExtra)
-
-indeffplot <- function(hier_modelc_adj_v2, brmmodel8_v2){
-  
-  # Only gamma's 
-  gamnamesstan <- paste("gammai", "[",1:52,"]", sep="")
-  gamstan <- as.data.frame(summary(hier_modelc_adj_v2, pars = c( gamnamesstan))$summary)
-  gamstanmu <- as.data.frame(summary(hier_modelc_adj_v2, pars = c("mu"))$summary)
-  gamstan$parameters <- c(gamnamesstan)
-  
-  for (i in 1:nrow(gamstan)){  # 1 if CI contains zero, 0 otherwise 
-    if (gamstan$`2.5%`[i]*gamstan$`97.5%`[i] < 0) {
-      gamstan$zero2[i] <- "pink"
-    } else (gamstan$zero2[i] <- "blue") 
-  }
-  
-  
-  gamstan2 <- gamstan[order(gamstan$mean),]
-  gamstan2$order <- c(1:52)
-  
-  gammarandomplot <- ggplot(gamstan2, aes(x = order, y = mean, ymin = `2.5%`, ymax = `97.5%`)) + 
-    geom_linerange(aes(color = "blue"), color = "#00BFC4") + 
-    geom_point() + 
-    geom_hline(yintercept = gamstanmu$mean, linetype = "dashed", alpha = .3) + 
-    labs(title = expression(~gamma[~i]), y = "Estimation") + 
-    theme(legend.position = "none", 
-          axis.title.x=element_blank(),
[…]

### 2023-06-30 J - R code/Figures/Fig11_modelestimatesfigure_truncated_onlyrstan.R

@@ -1,102 +1,0 @@
-# Model estimates figure 
-# Load fits 
-# For rstan the output is saved in hier_modelc_adj_07012022 (adjust "myPath" below to your directory)
-hier_modelc_adj_07012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
-
-library(rstan)
-
-# Plot results using https://github.com/jorvlan/open-visualizations/blob/master/R/repmes_tutorial_R.pdf
-# Install packages for plot 
-packages <- c("plyr", "lattice", "ggplot2", "dplyr", "readr", "rmarkdown","Rmisc")
-if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
-  install.packages(setdiff(packages, rownames(installed.packages())))  
-}
-if (!require(devtools)) {
-  install.packages("devtools")
-}
-devtools::install_github('erocoar/gghalves')
-
-library("plyr")
-library("lattice")
-library("ggplot2")
-library("dplyr")
-library("readr")
-library("rmarkdown")
-library("Rmisc")
-library("devtools")
-library("gghalves")
-
-indvestviolin_onlyrstan <- function(hier_modelc_adj_v2, brmmodel8_v2){
-  # Rstan 
-  ## Delta's 
-  ## Create dataframe 
-  parameter_m7 <- rep(c("Delta 1", "Delta 2", "Delta 3", "Delta 4"), each = 52 )
-  
-  ### Obtain all individual estimates delta 
-  hier_modeld <- hier_modelc_adj_v2
-  delta1_summary <- summary(hier_modeld, pars = c("delta1"), probs = c(0.1, 0.9))$summary
-  
-  delta2_summary <- summary(hier_modeld, pars = c("delta2"), probs = c(0.1, 0.9))$summary
-  
-  delta3_summary <- summary(hier_modeld, pars = c("delta3"), probs = c(0.1, 0.9))$summary
-  
-  delta4_summary <- summary(hier_modeld, pars = c("delta4"), probs = c(0.1, 0.9))$summary
-  
-  parest_m7 <- rbind(delta1_summary, delta2_summary, delta3_summary, delta4_summary)
-  datafr_m7 <- data.frame(parameter_m7, parest_m7)
-  
-  ## Plot 
-  set.seed(321)
[…]

### 2023-06-30 J - R code/Figures/Fig11_randomparfigure_trunc_ordered_per_individuals_rev.R

@@ -1,0 +1,680 @@
+# Individual effects from Rstan 
+# Order on participant 
+# For rstan the output is saved in hier_modelc_adj_07012022 (adjust "myPath" below to your directory)
+hier_modelc_adj_07012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
+
+# For brms the output is saved in object brmmodel8_new03012022 (this is without saving the prior samples)
+brmmodel8_new03012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/brms/Posterior/brmmodel8_new03012022.rds")  
+
+library(rstan)
+library(brms)
+library(gridExtra)
+
+indeffplot_orderedind <- function(hier_modelc_adj_v2, brmmodel8_v2){
+  
+  # Only gamma's 
+  gamnamesstan <- paste("gammai", "[",1:52,"]", sep="")
+  gamstan <- as.data.frame(summary(hier_modelc_adj_v2, pars = c( gamnamesstan))$summary)
+  gamstanmu <- as.data.frame(summary(hier_modelc_adj_v2, pars = c("mu"))$summary)
+  gamstan$parameters <- c(gamnamesstan)
+  
+  for (i in 1:nrow(gamstan)){  # 1 if CI contains zero, 0 otherwise 
+    if (gamstan$`2.5%`[i]*gamstan$`97.5%`[i] < 0) {
+      gamstan$zero2[i] <- "pink"
+    } else (gamstan$zero2[i] <- "blue") 
+  }
+  
+  
+  gamstan2 <- gamstan[order(gamstan$mean),]
+  gamstan$order <- c(1:52)
+  
+  gammarandomplot <- ggplot(gamstan, aes(x = order, y = mean, ymin = `2.5%`, ymax = `97.5%`)) + 
+    geom_linerange(aes(color = "blue"), color = "#00BFC4") + 
+    geom_point() + 
+    geom_hline(yintercept = gamstanmu$mean, linetype = "dashed", alpha = .3) + 
+    labs(title = expression(~gamma[~i]), y = "Estimation") + 
+    theme(legend.position = "none", 
+          axis.title.x=element_blank(),
[…]

### 2023-06-30 J - R code/Figures/Fig11_randomparfigure_trunc_v2.R

@@ -1,0 +1,345 @@
+# Individual effects from Rstan 
+# For rstan the output is saved in hier_modelc_adj_07012022 (adjust "myPath" below to your directory)
+hier_modelc_adj_07012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
+
+# For brms the output is saved in object brmmodel8_new03012022 (this is without saving the prior samples)
+brmmodel8_new03012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/brms/Posterior/brmmodel8_new03012022.rds")  
+
+library(rstan)
+library(brms)
+library(gridExtra)
+
+indeffplot <- function(hier_modelc_adj_v2, brmmodel8_v2){
+  
+  # Only gamma's 
+  gamnamesstan <- paste("gammai", "[",1:52,"]", sep="")
+  gamstan <- as.data.frame(summary(hier_modelc_adj_v2, pars = c( gamnamesstan))$summary)
+  gamstanmu <- as.data.frame(summary(hier_modelc_adj_v2, pars = c("mu"))$summary)
+  gamstan$parameters <- c(gamnamesstan)
+  
+  for (i in 1:nrow(gamstan)){  # 1 if CI contains zero, 0 otherwise 
+    if (gamstan$`2.5%`[i]*gamstan$`97.5%`[i] < 0) {
+      gamstan$zero2[i] <- "pink"
+    } else (gamstan$zero2[i] <- "blue") 
+  }
+  
+  
+  gamstan2 <- gamstan[order(gamstan$mean),]
+  gamstan2$order <- c(1:52)
+  
+  gammarandomplot <- ggplot(gamstan2, aes(x = order, y = mean, ymin = `2.5%`, ymax = `97.5%`)) + 
+    geom_linerange(aes(color = "blue"), color = "#00BFC4") + 
+    geom_point() + 
+    geom_hline(yintercept = gamstanmu$mean, linetype = "dashed", alpha = .3) + 
+    labs(title = expression(~gamma[~i]), y = "Estimation") + 
+    theme(legend.position = "none", 
+          axis.title.x=element_blank(),
[…]

### 2023-06-30 J - R code/Figures/Fig11b_modelestimatesfigure_trunc_v2.R

@@ -1,180 +1,0 @@
-# Model estimates figure 
-# Load fits 
-# For rstan the output is saved in hier_modelc_adj_07012022 (adjust "myPath" below to your directory)
-hier_modelc_adj_07012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
-
-# For brms the output is saved in object brmmodel8_new03012022 (this is without saving the prior samples)
-brmmodel8_new03012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/brms/Posterior/brmmodel8_new03012022.rds")  
-
-library(rstan)
-library(brms)
-
-# Plot results using https://github.com/jorvlan/open-visualizations/blob/master/R/repmes_tutorial_R.pdf
-# Install packages for plot 
-packages <- c("plyr", "lattice", "ggplot2", "dplyr", "readr", "rmarkdown","Rmisc")
-if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
-  install.packages(setdiff(packages, rownames(installed.packages())))  
-}
-if (!require(devtools)) {
-  install.packages("devtools")
-}
-devtools::install_github('erocoar/gghalves')
-
-library("plyr")
-library("lattice")
-library("ggplot2")
-library("dplyr")
-library("readr")
-library("rmarkdown")
-library("Rmisc")
-library("devtools")
-library("gghalves")
-
-indvestviolin <- function(hier_modelc_adj_v2, brmmodel8_v2){
-  # Rstan 
-  ## Delta's 
-  ## Create dataframe 
-  parameter_m7 <- rep(c("Delta 1", "Delta 2", "Delta 3", "Delta 4"), each = 52 )
-  
-  ### Obtain all individual estimates delta 
-  hier_modeld <- hier_modelc_adj_v2
-  delta1_summary <- summary(hier_modeld, pars = c("delta1"), probs = c(0.1, 0.9))$summary
-  
-  delta2_summary <- summary(hier_modeld, pars = c("delta2"), probs = c(0.1, 0.9))$summary
-  
[…]

### 2023-06-30 J - R code/Figures/Fig12_chainrhatneff_v2.R

@@ -1,287 +1,0 @@
-# Figure 18 + 19 (log models)
-# Traceplots dif 2 (difference between 6 and 7) + rhat
-# Based on output each package of symbolic distance effect model 
-# Load packages
-library(brms)
-library(rstan)
-
-# Load data 
-# For rstan the output is saved in hier_logmodelc_adj13012022  (adjust "myPath" below to your directory)
-hier_logmodelc_adj13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/rstan/log-normal model/hier_logmodelc_adj13012022.rds")
-
-# For brms the output is saved in object brmmodel8log13012022 (this is without saving the prior samples)
-brmmodel8log13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/brms/log-normal/brmmodel8log13012022.rds")  
-
-tracerhatnefflog <- function(hier_logmodelc_adj30032021, brmmodel8log02042021){
-#Rstan
-trdif2rstan <- plot(hier_logmodelc_adj30032021, plotfun = "trace", pars = c("mu4"))  # rstan 
-# plot(hier_modelc_adj, plotfun = "trace", pars = c("mu4"), inc_warmup = TRUE)  # show warmup -- > ugly plot 
-
-# Extract legend 
-library(cowplot)
-legendtrace <- cowplot::get_legend(trdif2rstan)
-library(grid)
-grid.draw(legendtrace)
-
-# Obtain colors from rstanplot 
-unique(ggplot_build(trdif2rstan)$data[[1]]$colour)
-# "#E66101" "#998EC3" "#542788" "#F1A340"
-
-# Final rstan plot 
-trdif2rstan2 <- trdif2rstan + ylab(expression(~mu[~delta[~6]])) + 
-  theme(axis.title.y=element_text(angle = 0, 
-                                  vjust = 0.5, 
-                                  hjust = -0.5, 
-                                  margin = margin(0, 0.7, 0, 0, "cm"), size = 14), 
-        legend.position = "none", 
-        axis.line.x  = element_line(size = 0.6), 
-        axis.line.y  = element_line(size = 0.6), 
-        plot.margin = unit(c(0.3,0.1,0.3,0.05),"cm"), 
[…]

### 2023-06-30 J - R code/Figures/Fig12_modelestimatesfigure_truncated_onlyrstan.R

@@ -1,0 +1,102 @@
+# Model estimates figure 
+# Load fits 
+# For rstan the output is saved in hier_modelc_adj_07012022 (adjust "myPath" below to your directory)
+hier_modelc_adj_07012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
+
+library(rstan)
+
+# Plot results using https://github.com/jorvlan/open-visualizations/blob/master/R/repmes_tutorial_R.pdf
+# Install packages for plot 
+packages <- c("plyr", "lattice", "ggplot2", "dplyr", "readr", "rmarkdown","Rmisc")
+if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
+  install.packages(setdiff(packages, rownames(installed.packages())))  
+}
+if (!require(devtools)) {
+  install.packages("devtools")
+}
+devtools::install_github('erocoar/gghalves')
+
+library("plyr")
+library("lattice")
+library("ggplot2")
+library("dplyr")
+library("readr")
+library("rmarkdown")
+library("Rmisc")
+library("devtools")
+library("gghalves")
+
+indvestviolin_onlyrstan <- function(hier_modelc_adj_v2, brmmodel8_v2){
+  # Rstan 
+  ## Delta's 
+  ## Create dataframe 
+  parameter_m7 <- rep(c("Delta 1", "Delta 2", "Delta 3", "Delta 4"), each = 52 )
+  
+  ### Obtain all individual estimates delta 
+  hier_modeld <- hier_modelc_adj_v2
+  delta1_summary <- summary(hier_modeld, pars = c("delta1"), probs = c(0.1, 0.9))$summary
+  
+  delta2_summary <- summary(hier_modeld, pars = c("delta2"), probs = c(0.1, 0.9))$summary
+  
+  delta3_summary <- summary(hier_modeld, pars = c("delta3"), probs = c(0.1, 0.9))$summary
+  
+  delta4_summary <- summary(hier_modeld, pars = c("delta4"), probs = c(0.1, 0.9))$summary
+  
+  parest_m7 <- rbind(delta1_summary, delta2_summary, delta3_summary, delta4_summary)
+  datafr_m7 <- data.frame(parameter_m7, parest_m7)
+  
+  ## Plot 
+  set.seed(321)
[…]

### 2023-06-30 J - R code/Figures/Fig13_chainrhatneff_v2.R

@@ -1,0 +1,286 @@
+# Traceplots dif 2 (difference between 6 and 7) + rhat
+# Based on output each package of symbolic distance effect model 
+# Load packages
+library(brms)
+library(rstan)
+
+# Load data 
+# For rstan the output is saved in hier_logmodelc_adj13012022  (adjust "myPath" below to your directory)
+hier_logmodelc_adj13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/rstan/log-normal model/hier_logmodelc_adj13012022.rds")
+
+# For brms the output is saved in object brmmodel8log13012022 (this is without saving the prior samples)
+brmmodel8log13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/brms/log-normal/brmmodel8log13012022.rds")  
+
+tracerhatnefflog <- function(hier_logmodelc_adj30032021, brmmodel8log02042021){
+#Rstan
+trdif2rstan <- plot(hier_logmodelc_adj30032021, plotfun = "trace", pars = c("mu4"))  # rstan 
+# plot(hier_modelc_adj, plotfun = "trace", pars = c("mu4"), inc_warmup = TRUE)  # show warmup -- > ugly plot 
+
+# Extract legend 
+library(cowplot)
+legendtrace <- cowplot::get_legend(trdif2rstan)
+library(grid)
+grid.draw(legendtrace)
+
+# Obtain colors from rstanplot 
+unique(ggplot_build(trdif2rstan)$data[[1]]$colour)
+# "#E66101" "#998EC3" "#542788" "#F1A340"
+
+# Final rstan plot 
+trdif2rstan2 <- trdif2rstan + ylab(expression(~mu[~delta[~6]])) + 
+  theme(axis.title.y=element_text(angle = 0, 
+                                  vjust = 0.5, 
+                                  hjust = -0.5, 
+                                  margin = margin(0, 0.7, 0, 0, "cm"), size = 14), 
+        legend.position = "none", 
+        axis.line.x  = element_line(size = 0.6), 
+        axis.line.y  = element_line(size = 0.6), 
+        plot.margin = unit(c(0.3,0.1,0.3,0.05),"cm"), 
[…]

### 2023-06-30 J - R code/Figures/Fig13_posteriorlog_v2.R

@@ -1,80 +1,0 @@
-# Figure 13
-# Posterior distribution parameters - lognormal model 
-# For rstan the output is saved in hier_logmodelc_adj13012022 (adjust "myPath" below to your directory)
-hier_logmodelc_adj13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/rstan/log-normal model/hier_logmodelc_adj13012022.rds")
-
-# For brms the output is saved in object brmmodel8log13012022 (this is without saving the prior samples)
-brmmodel8log13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/brms/log-normal/brmmodel8log13012022.rds")  
-
-
-# Used this link: https://cran.r-project.org/web/packages/bayesplot/vignettes/plotting-mcmc-draws.html 
-library(tidybayes)
-library(bayesplot)
-library(brms)
-library(rstan)
-
-posteriorlogfunc <- function(hier_logmodelc_adj30032021, brmmodel8log02042021){
-  post8 <- as.array(brmmodel8log02042021)
-  dimnames(post8)
-  color_scheme_set("orange")
-  pairs8 <- mcmc_areas(
-    post8, 
-    pars = c("b_side", "b_dif1", "b_dif2", "b_dif3", "b_dif4"),
-    prob = 0.95, # 95% intervals
-    prob_outer = 0.99, # 99%
-    point_est = "mean"
-  )
-  # scale_x_continuous(breaks=c(1,2,3,4), labels=c(expression(~delta[~7-8]), expression(~delta[~6-7]), expression(~delta[~4-3]), expression(~delta[~3-2]))) +
-  
-  pairs82 <- pairs8 + 
-    theme_apa() + 
-    ggtitle("Brms") + 
-    xlim(-0.035, 0.085) + 
-    scale_x_continuous(breaks = c(-0.025, 0, 0.025, 0.05, 0.075, 0.1)) + 
-    scale_y_discrete(labels= c( 
-      expression(~mu[~beta]), 
-      expression(~mu[~delta[~7]]), 
-      expression(~mu[~delta[~6]]), 
-      expression(~mu[~delta[~4]]), 
-      expression(~mu[~delta[~3]]))) + 
-    theme(axis.title.y=element_blank(),
[…]

### 2023-06-30 J - R code/Figures/Fig14_posteriorlog_v2.R

@@ -1,0 +1,79 @@
+# Posterior distribution parameters - lognormal model 
+# For rstan the output is saved in hier_logmodelc_adj13012022 (adjust "myPath" below to your directory)
+hier_logmodelc_adj13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/rstan/log-normal model/hier_logmodelc_adj13012022.rds")
+
+# For brms the output is saved in object brmmodel8log13012022 (this is without saving the prior samples)
+brmmodel8log13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/brms/log-normal/brmmodel8log13012022.rds")  
+
+
+# Used this link: https://cran.r-project.org/web/packages/bayesplot/vignettes/plotting-mcmc-draws.html 
+library(tidybayes)
+library(bayesplot)
+library(brms)
+library(rstan)
+
+posteriorlogfunc <- function(hier_logmodelc_adj30032021, brmmodel8log02042021){
+  post8 <- as.array(brmmodel8log02042021)
+  dimnames(post8)
+  color_scheme_set("orange")
+  pairs8 <- mcmc_areas(
+    post8, 
+    pars = c("b_side", "b_dif1", "b_dif2", "b_dif3", "b_dif4"),
+    prob = 0.95, # 95% intervals
+    prob_outer = 0.99, # 99%
+    point_est = "mean"
+  )
+  # scale_x_continuous(breaks=c(1,2,3,4), labels=c(expression(~delta[~7-8]), expression(~delta[~6-7]), expression(~delta[~4-3]), expression(~delta[~3-2]))) +
+  
+  pairs82 <- pairs8 + 
+    theme_apa() + 
+    ggtitle("Brms") + 
+    xlim(-0.035, 0.085) + 
+    scale_x_continuous(breaks = c(-0.025, 0, 0.025, 0.05, 0.075, 0.1)) + 
+    scale_y_discrete(labels= c( 
+      expression(~mu[~beta]), 
+      expression(~mu[~delta[~7]]), 
+      expression(~mu[~delta[~6]]), 
+      expression(~mu[~delta[~4]]), 
+      expression(~mu[~delta[~3]]))) + 
+    theme(axis.title.y=element_blank(),
[…]

### 2023-06-30 J - R code/Figures/Fig14_randomparlog_v2.R

@@ -1,364 +1,0 @@
-# Figure 14 
-# Individual effects from Rstan 
-# Load data 
-# For rstan the output is saved in hier_logmodelc_adj13012022 (adjust "myPath" below to your directory)
-hier_logmodelc_adj13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/rstan/log-normal model/hier_logmodelc_adj13012022.rds")
-
-# For brms the output is saved in object brmmodel8log13012022 (this is without saving the prior samples)
-brmmodel8log13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/brms/log-normal/brmmodel8log13012022.rds")  
-
-library(rstan)
-library(brms)
-
-randomparlogfunc <- function(hier_logmodelc_adj30032021, brmmodel8log02042021){
-  # Documentation on how to retrieve objects from stan output: https://cran.r-project.org/web/packages/rstan/vignettes/stanfit-objects.html
-  theme_set(theme_apa(base_size = 9))
-  # Only gamma's 
-  gamnamesstan <- paste("gammai", "[",1:52,"]", sep="")
-  gamstan <- as.data.frame(summary(hier_logmodelc_adj30032021, pars = c( gamnamesstan))$summary)
-  gamstanmu <- as.data.frame(summary(hier_logmodelc_adj30032021, pars = c("mu"))$summary)
-  gamstan$parameters <- c(gamnamesstan)
-  
-  for (i in 1:nrow(gamstan)){  # 1 if CI contains zero, 0 otherwise 
-    if (gamstan$`2.5%`[i]*gamstan$`97.5%`[i] < 0) {
-      gamstan$zero2[i] <- "pink"
-    } else (gamstan$zero2[i] <- "blue") 
-  }
-  
-  
-  gamstan2 <- gamstan[order(gamstan$mean),]
-  gamstan2$order <- c(1:52)
-  
-  gammarandomplot <- ggplot(gamstan2, aes(x = order, y = mean, ymin = `2.5%`, ymax = `97.5%`)) + 
-    geom_linerange(aes(color = "blue"), color = "#00BFC4") + 
-    geom_point() + 
[…]

### 2023-06-30 J - R code/Figures/Fig15_lognormaldifindest_onlyrstan_v2.R

@@ -1,104 +1,0 @@
-# Figure 15 
-
-# Load data 
-# For rstan the output is saved in hier_logmodelc_adj13012022 (adjust "myPath" below to your directory)
-hier_logmodelc_adj13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/rstan/log-normal model/hier_logmodelc_adj13012022.rds")
-
-# For brms the output is saved in object brmmodel8log13012022 (this is without saving the prior samples)
-brmmodel8log13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/brms/log-normal/brmmodel8log13012022.rds")  
-
-library(brms)
-library(rstan)
-
-# Plot results using https://github.com/jorvlan/open-visualizations/blob/master/R/repmes_tutorial_R.pdf
-# Install packages for plot 
-packages <- c("plyr", "lattice", "ggplot2", "dplyr", "readr", "rmarkdown","Rmisc")
-if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
-  install.packages(setdiff(packages, rownames(installed.packages())))  
-}
-if (!require(devtools)) {
-  install.packages("devtools")
-}
-devtools::install_github('erocoar/gghalves')
-
-library("plyr")
-library("lattice")
-library("ggplot2")
-library("dplyr")
-library("readr")
-library("rmarkdown")
-library("Rmisc")
-library("devtools")
-library("gghalves")
-
-lognormaldicfindestfunc_onlyrstan <- function(hier_logmodelc_adj30032021, brmmodel8log02042021){
-  # Rstan 
-  parameter_m7 <- rep(c("Delta 1", "Delta 2", "Delta 3", "Delta 4"), each = 52 )
-  
-  # Obtain all individual estimates delta 
-  hier_modeld <- hier_logmodelc_adj30032021
-  delta1_summary <- summary(hier_modeld, pars = c("delta1"), probs = c(0.1, 0.9))$summary
-  
-  delta2_summary <- summary(hier_modeld, pars = c("delta2"), probs = c(0.1, 0.9))$summary
-  
-  delta3_summary <- summary(hier_modeld, pars = c("delta3"), probs = c(0.1, 0.9))$summary
-  
[…]

### 2023-06-30 J - R code/Figures/Fig15b_lognormaldifindest_v2.R

@@ -1,176 +1,0 @@
-# Figure 15b (not in manuscript) 
-
-# Load data 
-# For rstan the output is saved in hier_logmodelc_adj13012022 (adjust "myPath" below to your directory)
-hier_logmodelc_adj13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/rstan/log-normal model/hier_logmodelc_adj13012022.rds")
-
-# For brms the output is saved in object brmmodel8log13012022 (this is without saving the prior samples)
-brmmodel8log13012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/brms/log-normal/brmmodel8log13012022.rds")  
-
-library(brms)
-library(rstan)
-
-# Plot results using https://github.com/jorvlan/open-visualizations/blob/master/R/repmes_tutorial_R.pdf
-# Install packages for plot 
-packages <- c("plyr", "lattice", "ggplot2", "dplyr", "readr", "rmarkdown","Rmisc")
-if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
-  install.packages(setdiff(packages, rownames(installed.packages())))  
-}
-if (!require(devtools)) {
-  install.packages("devtools")
-}
-devtools::install_github('erocoar/gghalves')
-
-library("plyr")
-library("lattice")
-library("ggplot2")
-library("dplyr")
-library("readr")
-library("rmarkdown")
-library("Rmisc")
-library("devtools")
-library("gghalves")
-
-lognormaldicfindestfunc <- function(hier_logmodelc_adj30032021, brmmodel8log02042021){
-  # Rstan 
-  parameter_m7 <- rep(c("Delta 1", "Delta 2", "Delta 3", "Delta 4"), each = 52 )
-  
-  # Obtain all individual estimates delta 
-  hier_modeld <- hier_logmodelc_adj30032021
-  delta1_summary <- summary(hier_modeld, pars = c("delta1"), probs = c(0.1, 0.9))$summary
-  
-  delta2_summary <- summary(hier_modeld, pars = c("delta2"), probs = c(0.1, 0.9))$summary
-  
[…]

### 2023-06-30 J - R code/Figures/Fig16_savagedickey_trunc.R

@@ -1,191 +1,0 @@
-# Load packages
-library(ggridges)
-library(brms)
-
-# Load brms output 
-# For brms the output is saved in object brmmodel8_new_bf_03012022 (this is without saving the prior samples)
-brmmodel8_new_bf_03012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/brms/BF/brmmodel8_new_bf_03012022.rds") 
-
-savagedickeyfigure <- function(brmmodel8bf){
-  # Obtain Savage-Dickey ratio 
-  # https://vuorre.netlify.app/post/2017/03/21/bayes-factors-with-brms/ 
-  # Beta 
-  b1brms <- hypothesis(brmmodel8bf, "a_side = 0")  # Evid.Ratio is Bayes factor BF01, have to set sample_prior = TRUE when fitting model 
-  b1brms$hypothesis$Evid.Ratio 
-  plotsdratiob1 <- plot(b1brms,plot = F, theme = theme_get())[[1]] 
-  plotsdratiob1 + xlim(-0.07, 0.07) 
-  
-  # Plot function 
-  savdicratioplot <- function(plotsdratio){  # put in plot.hypothesisbrms object 
-    plotsratio3data <- plotsdratio$data
-    
-    density(plotsratio3data$values[plotsratio3data$Type == "Posterior"])
-    xpriorsd <- c(density(plotsratio3data$values[plotsratio3data$Type == "Posterior"])[1]$x, density(plotsratio3data$values[plotsratio3data$Type == "Prior"])[1]$x)  # x 
-    ypriorsd <- c(density(plotsratio3data$values[plotsratio3data$Type == "Posterior"])[2]$y, density(plotsratio3data$values[plotsratio3data$Type == "Prior"])[2]$y)  # y 
-    xypriorsd <- data.frame(x = xpriorsd, y = ypriorsd, Type = rep(c("Posterior", "Prior"), each = 512))
-    maxprior <- max(density(plotsratio3data$values[plotsratio3data$Type == "Prior"])[2]$y)  
-    
-    
-    sdrdif3plot2 <- ggplot(xypriorsd, aes(x, y, color = Type, fill = Type)) + 
-      geom_line() + 
-      geom_area(position = "identity", alpha = 0.3) + 
[…]

### 2023-06-30 J - R code/Figures/Fig17_bridgesamplingresults_v2.R

@@ -1,76 +1,0 @@
-# BF rstan bridge-sampling 
-library(ggplot2)
-library(brms)
-library(rstan )
-
-# Load results
-# rstan
-resultsbsrstan <- read.csv("/myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/Model comparison/Bayes Factors per package/bssresults_07012022.csv")
-
-## Restructure 
-resultsbsrstan$package <- rep("Rstan", nrow(resultsbsrstan))
-resultsbsrstan <- data.frame(Fit4 = resultsbsrstan$Fit4, model = resultsbsrstan$model, package = resultsbsrstan$package)
-
-# brms 
-resultsbsbrms <- read.csv("/myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/R objects/Model comparison/Bayes Factors per package/bssresults_brms_15032022_10rep.csv")
-
-## Restructure 
-resultsbsbrms$package <- rep("Brms", nrow(resultsbsbrms))
-resultsbsbrms <- data.frame(Fit4 = resultsbsbrms$Fit1, model = resultsbsbrms$model, package = resultsbsbrms$package)
-
-# Combine rstan and brms results 
-resultsbf10 <- rbind(resultsbsrstan, resultsbsbrms)
-
-resultsbf10$BF01 <- 1/resultsbf10$Fit4
-
-level_order <- c('Null model', 'Side model', 'Digit model') 
-
-# Evidence of full model against alternative models 
-normalbfplot <- ggplot(resultsbf10[which(resultsbf10$model == "Null model" | 
-                                           resultsbf10$model == "Side model" |  
-                                           resultsbf10$model == "Digit model"),],
-                       aes(x = factor(model, level_order), y = Fit4, color = 
-                             package, shape = package, fill = package)) + 
-  geom_jitter(alpha = 0.5, position = position_jitterdodge(dodge.width = 0.5)) + 
-  # geom_violin(trim = FALSE) + 
-  scale_y_log10() + 
-  theme_classic() + 
-  labs(subtitle = "Linear model \n", 
-       x = "", y = "BF01") + 
-  scale_color_manual(name = "Package", 
-                     labels = c("Rstan", "Brms"),
-                     values = c("darkgreen", "darkorange")) + 
-  scale_shape_manual(name = "Package", 
-                     labels = c("Rstan", "Brms"),
-                     values = c(17, 19)) + 
-  scale_fill_manual(name = "Package", 
[…]

### 2023-06-30 J - R code/Figures/Fig17_savagedickey_trunc.R

@@ -1,0 +1,191 @@
+# Load packages
+library(ggridges)
+library(brms)
+
+# Load brms output 
+# For brms the output is saved in object brmmodel8_new_bf_03012022 (this is without saving the prior samples)
+brmmodel8_new_bf_03012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/brms/BF/brmmodel8_new_bf_03012022.rds") 
+
+savagedickeyfigure <- function(brmmodel8bf){
+  # Obtain Savage-Dickey ratio 
+  # https://vuorre.netlify.app/post/2017/03/21/bayes-factors-with-brms/ 
+  # Beta 
+  b1brms <- hypothesis(brmmodel8bf, "a_side = 0")  # Evid.Ratio is Bayes factor BF01, have to set sample_prior = TRUE when fitting model 
+  b1brms$hypothesis$Evid.Ratio 
+  plotsdratiob1 <- plot(b1brms,plot = F, theme = theme_get())[[1]] 
+  plotsdratiob1 + xlim(-0.07, 0.07) 
+  
+  # Plot function 
+  savdicratioplot <- function(plotsdratio){  # put in plot.hypothesisbrms object 
+    plotsratio3data <- plotsdratio$data
+    
+    density(plotsratio3data$values[plotsratio3data$Type == "Posterior"])
+    xpriorsd <- c(density(plotsratio3data$values[plotsratio3data$Type == "Posterior"])[1]$x, density(plotsratio3data$values[plotsratio3data$Type == "Prior"])[1]$x)  # x 
+    ypriorsd <- c(density(plotsratio3data$values[plotsratio3data$Type == "Posterior"])[2]$y, density(plotsratio3data$values[plotsratio3data$Type == "Prior"])[2]$y)  # y 
+    xypriorsd <- data.frame(x = xpriorsd, y = ypriorsd, Type = rep(c("Posterior", "Prior"), each = 512))
+    maxprior <- max(density(plotsratio3data$values[plotsratio3data$Type == "Prior"])[2]$y)  
+    
+    
+    sdrdif3plot2 <- ggplot(xypriorsd, aes(x, y, color = Type, fill = Type)) + 
+      geom_line() + 
+      geom_area(position = "identity", alpha = 0.3) + 
[…]

### 2023-06-30 J - R code/Figures/Fig4_distributions_v2.R

@@ -1,134 +1,0 @@
-# Figure distributions
-library(ggplot2)
-
-x <- seq(-3, 3, by=.1)
-narrow <- dnorm(x, 0, .1)
-#plot(x, narrow, type = "l")
-wide <- dnorm(x, 0, 10)
-#plot(x, wide)
-
-pwide <- ggplot(data = data.frame(x = c(-3, 3)), aes(x)) +
-  stat_function(fun = dnorm, n = 101, args = list(mean = 0, sd = 10)) + ylab("Density") +
-  scale_y_continuous(breaks = NULL) 
-
-pnarrow <- ggplot(data = data.frame(x = c(-3, 3)), aes(x)) +
-  stat_function(fun = dnorm, n = 101, args = list(mean = 0, sd = 1)) + ylab("Density") +
-  scale_y_continuous(breaks = NULL)
-
-ptogether <- ggplot(data = data.frame(x = c(-5, 5)), aes(x)) +
-  stat_function(fun = dnorm, n = 101, args = list(mean = 0, sd = 4), linetype = 2) + ylab("Density") +
-  stat_function(fun = dnorm, n = 101, args = list(mean = 0, sd = 1))
-
-parametersnormaldistrplot <- c("Mean = 0, SD = 4" = "#D95F02", "Mean = 0, SD = 1" = "#1B9E77", "Mean = 1, SD = 0.5" = "#E7298A")
-
-normaldistriplot <- ggplot(data = data.frame(x = c(-5, 5)), aes(x)) +
-  stat_function(fun = dnorm, n = 1001, args = list(mean = 0, sd = 4), linetype = 2, geom = "area", alpha = 0.3, position = "identity", aes(color = "Mean = 0, SD = 4", fill = "Mean = 0, SD = 4")) +  
-  ylab("Density") +
-  stat_function(fun = dnorm, n = 1001, args = list(mean = 0, sd = 1), linetype = 3, geom = "area", alpha = 0.3, position = "identity", aes(color = "Mean = 0, SD = 1", fill = "Mean = 0, SD = 1")) + 
-  stat_function(fun = dnorm, n = 1001, args = list(mean = 1, sd = 0.5), geom = "area", alpha = 0.3, position = "identity", aes(color = "Mean = 1, SD = 0.5", fill = "Mean = 1, SD = 0.5")) + 
-  geom_hline(yintercept=0, colour="white", size=1) + 
[…]

### 2023-06-30 J - R code/Figures/Fig5_distributions_v2.R

@@ -1,0 +1,134 @@
+# Figure distributions
+library(ggplot2)
+
+x <- seq(-3, 3, by=.1)
+narrow <- dnorm(x, 0, .1)
+#plot(x, narrow, type = "l")
+wide <- dnorm(x, 0, 10)
+#plot(x, wide)
+
+pwide <- ggplot(data = data.frame(x = c(-3, 3)), aes(x)) +
+  stat_function(fun = dnorm, n = 101, args = list(mean = 0, sd = 10)) + ylab("Density") +
+  scale_y_continuous(breaks = NULL) 
+
+pnarrow <- ggplot(data = data.frame(x = c(-3, 3)), aes(x)) +
+  stat_function(fun = dnorm, n = 101, args = list(mean = 0, sd = 1)) + ylab("Density") +
+  scale_y_continuous(breaks = NULL)
+
+ptogether <- ggplot(data = data.frame(x = c(-5, 5)), aes(x)) +
+  stat_function(fun = dnorm, n = 101, args = list(mean = 0, sd = 4), linetype = 2) + ylab("Density") +
+  stat_function(fun = dnorm, n = 101, args = list(mean = 0, sd = 1))
+
+parametersnormaldistrplot <- c("Mean = 0, SD = 4" = "#D95F02", "Mean = 0, SD = 1" = "#1B9E77", "Mean = 1, SD = 0.5" = "#E7298A")
+
+normaldistriplot <- ggplot(data = data.frame(x = c(-5, 5)), aes(x)) +
+  stat_function(fun = dnorm, n = 1001, args = list(mean = 0, sd = 4), linetype = 2, geom = "area", alpha = 0.3, position = "identity", aes(color = "Mean = 0, SD = 4", fill = "Mean = 0, SD = 4")) +  
+  ylab("Density") +
+  stat_function(fun = dnorm, n = 1001, args = list(mean = 0, sd = 1), linetype = 3, geom = "area", alpha = 0.3, position = "identity", aes(color = "Mean = 0, SD = 1", fill = "Mean = 0, SD = 1")) + 
+  stat_function(fun = dnorm, n = 1001, args = list(mean = 1, sd = 0.5), geom = "area", alpha = 0.3, position = "identity", aes(color = "Mean = 1, SD = 0.5", fill = "Mean = 1, SD = 0.5")) + 
+  geom_hline(yintercept=0, colour="white", size=1) + 
[…]

### 2023-06-30 J - R code/Figures/Fig5_priorpredictions_trunc_v3.R

@@ -1,449 +1,0 @@
-# Prior predictions figure 
-library(truncnorm)
-library(crch)
-library(LaplacesDemon)
-library(ggplot2)
-library(ggbeeswarm)
-library(gridExtra)
-
-###################################################################################################
-# Preparation simulation
-###################################################################################################
-
-# Without for loop 
-set.seed(123)   
-
-Ipp <- 10  # number of participants  
-Jpp <- 6   # conditions: digit 2, 3, 4, 6, 7, and 8 
-Kpp <- 60  # 60 observations per condition per person 
-
-exppp <- rep(1:Ipp, each = Jpp * Kpp)  # which row belongs to which participant 
-condpp <- rep(0:5, Ipp, each = Kpp)    # to which condition row belongs  
-subpp <- rep(1:Kpp, Ipp * Jpp)         # to which observation per condition per parcticipant belongs row 
-
-# add parameters x, u, v, w, z 
-xjpp <- NA 
-ujpp <- NA 
-vjpp <- NA 
-wjpp <- NA 
-zjpp <- NA 
-# side effect, greater or smaller than 5  
-for (j in 1:length(condpp)){  # 3 = 6
-  if (condpp[j] < 3) {
-    xjpp[j] <- 1/2}
-  else xjpp[j] <- -1/2
-}
-
-### Add variable with information about difference between digits, for parameters delta's 
-#### Difference 8 and 7 
-for (j in 1:length(condpp)){ 
-  if (condpp[j] == 5) {  # 5 = 8 , 
-    ujpp[j] <- 0}
-  else if (condpp[j] == 4) {
-    ujpp[j] <- 1}
-  else ujpp[j] <- 0
-}
-
-# Difference 7 and 6 
-for (j in 1:length(condpp)){ 
-  if (condpp[j] == 4) {  # 4 = 7, 3 = 6  
-    vjpp[j] <- 0}
-  else if (condpp[j] == 3) {
-    vjpp[j] <- 1}
-  else vjpp[j] <- 0
-}
-
-# Difference 4 and 3 
-for (j in 1:length(condpp)){ 
-  if (condpp[j] == 1) {  # 1 = 3, 2 = 4 
-    wjpp[j] <- 0}
-  else if (condpp[j] == 2) {
-    wjpp[j] <- 1}
-  else wjpp[j] <- 0
-}
-
-# Difference 3 and 2 
-for (j in 1:length(condpp)){ 
-  if (condpp[j] == 0) {  # 0 = 2, 1 = 3
[…]

### 2023-06-30 J - R code/Figures/Fig6_priorpredictions_trunc_v3.R

@@ -1,0 +1,449 @@
+# Prior predictions figure 
+library(truncnorm)
+library(crch)
+library(LaplacesDemon)
+library(ggplot2)
+library(ggbeeswarm)
+library(gridExtra)
+
+###################################################################################################
+# Preparation simulation
+###################################################################################################
+
+# Without for loop 
+set.seed(123)   
+
+Ipp <- 10  # number of participants  
+Jpp <- 6   # conditions: digit 2, 3, 4, 6, 7, and 8 
+Kpp <- 60  # 60 observations per condition per person 
+
+exppp <- rep(1:Ipp, each = Jpp * Kpp)  # which row belongs to which participant 
+condpp <- rep(0:5, Ipp, each = Kpp)    # to which condition row belongs  
+subpp <- rep(1:Kpp, Ipp * Jpp)         # to which observation per condition per parcticipant belongs row 
+
+# add parameters x, u, v, w, z 
+xjpp <- NA 
+ujpp <- NA 
+vjpp <- NA 
+wjpp <- NA 
+zjpp <- NA 
+# side effect, greater or smaller than 5  
+for (j in 1:length(condpp)){  # 3 = 6
+  if (condpp[j] < 3) {
+    xjpp[j] <- 1/2}
+  else xjpp[j] <- -1/2
+}
+
+### Add variable with information about difference between digits, for parameters delta's 
+#### Difference 8 and 7 
+for (j in 1:length(condpp)){ 
+  if (condpp[j] == 5) {  # 5 = 8 , 
+    ujpp[j] <- 0}
+  else if (condpp[j] == 4) {
+    ujpp[j] <- 1}
+  else ujpp[j] <- 0
+}
+
+# Difference 7 and 6 
+for (j in 1:length(condpp)){ 
+  if (condpp[j] == 4) {  # 4 = 7, 3 = 6  
+    vjpp[j] <- 0}
+  else if (condpp[j] == 3) {
+    vjpp[j] <- 1}
+  else vjpp[j] <- 0
+}
+
+# Difference 4 and 3 
+for (j in 1:length(condpp)){ 
+  if (condpp[j] == 1) {  # 1 = 3, 2 = 4 
+    wjpp[j] <- 0}
+  else if (condpp[j] == 2) {
+    wjpp[j] <- 1}
+  else wjpp[j] <- 0
+}
+
+# Difference 3 and 2 
+for (j in 1:length(condpp)){ 
+  if (condpp[j] == 0) {  # 0 = 2, 1 = 3
[…]

### 2023-06-30 J - R code/Figures/Fig8_traceplotsfigure_trunc_v2.R

@@ -1,438 +1,0 @@
-# Traceplots dif 2 (difference between 6 and 7) + rhat
-# Based on output each package of symbolic distance effect model 
-# For rstan the output is saved in hier_modelc_adj_07012022 (adjust "myPath" below to your directory)
-hier_modelc_adj_07012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
-
-# For brms the output is saved in object brmmodel8_new03012022 (this is without saving the prior samples)
-brmmodel8_new03012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/brms/Posterior/brmmodel8_new03012022.rds")  
-
-tracerhatneff <- function(hier_modelc_adj_v2, brmmodel8_v2){
-  #Rstan
-  library(rstan)
-  trdif2rstan <- plot(hier_modelc_adj_v2, plotfun = "trace", pars = c("mu4"))  # rstan 
-  # plot(hier_modelc_adj, plotfun = "trace", pars = c("mu4"), inc_warmup = TRUE)  # show warmup -- > ugly plot 
-  
-  trdif2rstan <- trdif2rstan + labs(color='Chain') 
-  
-  # Extract legend 
-  library(cowplot)
-  legendtrace <- cowplot::get_legend(trdif2rstan)
-  library(grid)
-  grid.draw(legendtrace)
-  
-  # Obtain colors from rstanplot 
-  unique(ggplot_build(trdif2rstan)$data[[1]]$colour)
-  # "#E66101" "#998EC3" "#542788" "#F1A340"
-  
-  # Final rstan plot 
-  trdif2rstan2 <- trdif2rstan + ylab(expression(~mu[~delta[~6]])) + 
-    theme(axis.title.y=element_text(angle = 0, 
-                                    vjust = 0.5, 
-                                    hjust = -0.5, 
-                                    margin = margin(0, 0.7, 0, 0, "cm"), size = 14), 
-          legend.position = "none", 
-          axis.line.x  = element_line(size = 0.6), 
-          axis.line.y  = element_line(size = 0.6), 
-          plot.margin = unit(c(0.3,0.1,0.3,0.05),"cm"), 
[…]

### 2023-06-30 J - R code/Figures/Fig9_posteriorfigure_trunc_v2.R

@@ -1,81 +1,0 @@
-# Posterior distribution parameters 
-# For rstan the output is saved in hier_modelc_adj_07012022 (adjust "myPath" below to your directory)
-hier_modelc_adj_07012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
-
-# For brms the output is saved in object brmmodel8_new03012022 (this is without saving the prior samples)
-brmmodel8_new03012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/brms/Posterior/brmmodel8_new03012022.rds")  
-
-# Used this link: https://cran.r-project.org/web/packages/bayesplot/vignettes/plotting-mcmc-draws.html 
-
-posterplot <- function(hier_modelc_adj_v2, brmmodel8_v2){
-  library(tidybayes)
-  library(bayesplot)
-  library(rstan)
-  library(brms)
-  library(ggplot2)
-  library(papaja)
-  
-  post8 <- as.array(brmmodel8_v2)
-  dimnames(post8)
-  color_scheme_set("orange")
-  pairs8 <- mcmc_areas(
-    post8, 
-    pars = c("b_a_side", "b_a_dif1", "b_a_dif2", "b_a_dif3", "b_a_dif4"),
-    prob = 0.95, # 80% intervals
-    prob_outer = 0.99, # 99%
-    point_est = "mean"
-  )
-  # scale_x_continuous(breaks=c(1,2,3,4), labels=c(expression(~delta[~7-8]), expression(~delta[~6-7]), expression(~delta[~4-3]), expression(~delta[~3-2]))) +
-  
-  pairs82 <- pairs8 + 
-    theme_apa() + 
-    ggtitle("Brms") + 
-    xlim(-0.035, 0.085) + 
-    scale_x_continuous(breaks = c(-0.025, 0, 0.025, 0.05, 0.075)) + 
-    scale_y_discrete(labels= c( 
-      expression(~mu[~beta]), 
-      expression(~mu[~delta[~7]]), 
-      expression(~mu[~delta[~6]]), 
-      expression(~mu[~delta[~4]]), 
-      expression(~mu[~delta[~3]]))) + 
-    theme(axis.title.y=element_blank(),
[…]

### 2023-06-30 J - R code/Figures/Fig9_traceplotsfigure_trunc_v2.R

@@ -1,0 +1,438 @@
+# Traceplots dif 2 (difference between 6 and 7) + rhat
+# Based on output each package of symbolic distance effect model 
+# For rstan the output is saved in hier_modelc_adj_07012022 (adjust "myPath" below to your directory)
+hier_modelc_adj_07012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
+
+# For brms the output is saved in object brmmodel8_new03012022 (this is without saving the prior samples)
+brmmodel8_new03012022 <- readRDS("myPath/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/K - R objects/brms/Posterior/brmmodel8_new03012022.rds")  
+
+tracerhatneff <- function(hier_modelc_adj_v2, brmmodel8_v2){
+  #Rstan
+  library(rstan)
+  trdif2rstan <- plot(hier_modelc_adj_v2, plotfun = "trace", pars = c("mu4"))  # rstan 
+  # plot(hier_modelc_adj, plotfun = "trace", pars = c("mu4"), inc_warmup = TRUE)  # show warmup -- > ugly plot 
+  
+  trdif2rstan <- trdif2rstan + labs(color='Chain') 
+  
+  # Extract legend 
+  library(cowplot)
+  legendtrace <- cowplot::get_legend(trdif2rstan)
+  library(grid)
+  grid.draw(legendtrace)
+  
+  # Obtain colors from rstanplot 
+  unique(ggplot_build(trdif2rstan)$data[[1]]$colour)
+  # "#E66101" "#998EC3" "#542788" "#F1A340"
+  
+  # Final rstan plot 
+  trdif2rstan2 <- trdif2rstan + ylab(expression(~mu[~delta[~6]])) + 
+    theme(axis.title.y=element_text(angle = 0, 
+                                    vjust = 0.5, 
+                                    hjust = -0.5, 
+                                    margin = margin(0, 0.7, 0, 0, "cm"), size = 14), 
+          legend.position = "none", 
+          axis.line.x  = element_line(size = 0.6), 
+          axis.line.y  = element_line(size = 0.6), 
+          plot.margin = unit(c(0.3,0.1,0.3,0.05),"cm"), 
[…]

### 2023-06-30 Sensitivity_logmodel.R

@@ -1,323 +1,0 @@
-# Sensitivity analysis 
-# Based on rstan results 
-library(plyr)
-library(rstan)
-library(ggplot2)
-
-# Preparations 
-# Load datareal 
-indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
-colnames(indat)=c('sub','block','trial','stim','resp','rt','error')
-
-## Cleaning the data according to criteria discussed in paper 
-clean=function()
-{
-  indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
-  colnames(indat)=c('sub','block','trial','stim','resp','rt','error')
-  
-  bad1=indat$sub%in%c(34,43)
-  bad2=indat$rt<250 | indat$rt>2000
-  bad3=indat$err==1
-  bad4=indat$block==0 & indat$trial<20
-  bad5=indat$trial==0
-  
-  bad=bad1 | bad2 | bad3 |bad4 |bad5
-  dat=indat[!bad,]
-  return(dat)
-}
-
-indat1 <- clean()  # Final dataset 
-
-## Data into list and fill the list with necessary information 
-datareal <- list(N = length(unique(indat1$sub)))
-datareal$All <- nrow(indat1)  # number of total observations, adding to the list for the analysis  
-
-### Indicating which observation belongs to which person 
-rank(unique(indat1$sub))
-sub <- rep(1:N, each = J)
-true.theta[sub]
-
-for (j in 1:nrow(indat1)){
-  if (j == 1) {
-    indat1$sub1[j] <- 1}
-  else if (indat1$sub[j] == indat1$sub[j-1]) {
-    indat1$sub1[j] <- indat1$sub1[j-1]}
-  else indat1$sub1[j] <- indat1$sub1[j-1] + 1
-}
-
-
-datareal$group_inds <- indat1$sub1  # Add variable with indicator person to the list used for the analysi
-datareal$y <- indat1$rt/1000  # rt in seconds instead of miliseconds, added to the list for analysis  
-
-### Add variable with information of the side of the digit to the list, for parameter beta    
-# j = condition, if j > 5 than x = -1/2, if j < 5 than x = 1/2 
[…]

### 2023-06-30 rstan_modelcomparison_normalvslognormalmodel.R

@@ -1,188 +1,0 @@
-# Normal vs log normal BF comparison 
-
-
-library(plyr)
-library(rstan)
-library(ggplot2)
-library(bridgesampling)
-
-# Preparations 
-# Load datareal 
-indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
-colnames(indat)=c('sub','block','trial','stim','resp','rt','error')
-
-## Cleaning the data according to criteria discussed in paper 
-clean=function()
-{
-  indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
-  colnames(indat)=c('sub','block','trial','stim','resp','rt','error')
-  
-  bad1=indat$sub%in%c(34,43)
-  bad2=indat$rt<250 | indat$rt>2000
-  bad3=indat$err==1
-  bad4=indat$block==0 & indat$trial<20
-  bad5=indat$trial==0
-  
-  bad=bad1 | bad2 | bad3 |bad4 |bad5
-  dat=indat[!bad,]
-  return(dat)
-}
-
-indat1 <- clean()  # Final dataset 
-
-## Data into list and fill the list with necessary information 
-datareal <- list(N = length(unique(indat1$sub)))
-datareal$All <- nrow(indat1)  # number of total observations, adding to the list for the analysis  
-
-### Indicating which observation belongs to which person 
-rank(unique(indat1$sub))
-sub <- rep(1:N, each = J)
-true.theta[sub]
-
-for (j in 1:nrow(indat1)){
-  if (j == 1) {
-    indat1$sub1[j] <- 1}
-  else if (indat1$sub[j] == indat1$sub[j-1]) {
-    indat1$sub1[j] <- indat1$sub1[j-1]}
-  else indat1$sub1[j] <- indat1$sub1[j-1] + 1
-}
-
-
-datareal$group_inds <- indat1$sub1  # Add variable with indicator person to the list used for the analysi
-datareal$y <- indat1$rt/1000  # rt in seconds instead of miliseconds, added to the list for analysis  
-
-### Add variable with information of the side of the digit to the list, for parameter beta    
[…]

### 2023-06-30 manuscript.tex

@@
 \textcolor{blue}{Up until now, we specified models without a correlation between the individual effects.
-This means, for instance, that we do not expect that the estimations for the individual effects $\delta_{7, i}$ and $\delta_{3, i}$ to be related.
+This means, for instance, that we do not expect the estimations for the individual effects $\delta_{7, i}$ and $\delta_{3, i}$ to be related.
 However, in cases where we would expect the estimation of individual effects to be related, we could estimate the correlation coefficient between individual effects.
@@
 
-\includegraphics[width=1\linewidth]{I - Images/Fig4_distrtypefigure_v3} \hfill{}
+\includegraphics[width=1\linewidth]{I - Images/Fig5_distrtypefigure_v3} \hfill{}
 
@@
 
-\includegraphics[width=1\linewidth]{I - Images/Fig5_priorpredfigure_v2} \hfill{}
+\includegraphics[width=1\linewidth]{I - Images/Fig6_priorpredfigure_v2} \hfill{}
 
@@
 
-\includegraphics[width=1\linewidth]{I - Images/Fig8_traceplots3pnew_v2} \hfill{}
+\includegraphics[width=1\linewidth]{I - Images/Fig9_traceplots3pnew_v2} \hfill{}
 
@@
 
-\includegraphics[width=1\linewidth]{I - Images/Fig9_posteriormu_v3} \hfill{}
+\includegraphics[width=1\linewidth]{I - Images/Fig10_posteriormu_v3} \hfill{}
 
@@
 
-\includegraphics[width=1\linewidth]{I - Images/Fig10_randompar_v2} \hfill{}
+\includegraphics[width=1\linewidth]{I - Images/Fig11_randompar_v2} \hfill{}
 
@@
 
-\includegraphics[width=0.75\linewidth]{I - Images/Fig11_modelestimatefigure2_onlyrstan_v2} \hfill{}
+\includegraphics[width=0.75\linewidth]{I - Images/Fig12_modelestimatefigure2_onlyrstan_v2} \hfill{}
 
@@
 
-\includegraphics[width=1\linewidth]{I - Images/Fig12_traceplots3plog_v2} \hfill{}
+\includegraphics[width=1\linewidth]{I - Images/Fig13_traceplots3plog_v2} \hfill{}
 
@@
 
-\includegraphics[width=1\linewidth]{I - Images/Fig13_posteriormulog_v2} \hfill{}
+\includegraphics[width=1\linewidth]{I - Images/Fig14_posteriormulog_v2} \hfill{}
 
@@
 
-\includegraphics[width=1\linewidth]{I - Images/Fig14_randomparlog_v2} \hfill{}
+\includegraphics[width=1\linewidth]{I - Images/Fig15_randomparlog_v2} \hfill{}
 
@@
 
-\includegraphics[width=0.75\linewidth]{I - Images/Fig15_inddiflogplot_onlyrstan_v2} \hfill{}
+\includegraphics[width=0.75\linewidth]{I - Images/Fig16_inddiflogplot_onlyrstan_v2} \hfill{}
 
@@
 
-\includegraphics[width=0.6\linewidth]{I - Images/Fig16_sdrplotside_v2} \hfill{}
+\includegraphics[width=0.6\linewidth]{I - Images/Fig17_sdrplotside_v2} \hfill{}
 
@@
 
-\includegraphics[width=0.7\linewidth]{I - Images/Fig17_normalbfplot01} \hfill{}
+\includegraphics[width=0.7\linewidth]{I - Images/Fig18_normalbfplot01} \hfill{}
 
@@
 While individual differences are a common phenomenon in psychology, there are instances where general effects are sufficient to describe the data adequately, or where only a random intercept, but no random slopes are necessary.
-As shown in Tables\} \ref{tab:postmodprob} and \ref{tab:postmodproblog}, Bayesian model comparisons can help researchers to identify the model that best describes the data.
+As shown in Tables \ref{tab:postmodprob} and \ref{tab:postmodproblog}, Bayesian model comparisons can help researchers to identify the model that best describes the data.
 

### 2023-07-12 manuscript.Rmd
Remove blue color from manuscript

@@
    - \newcommand{\Julia}[1]{\todo[inline, color=mypink]{#1}}
-   - \newcommand{\Revision}[1]{\textcolor{blue}{#1}}
+   - \newcommand{\Revision}[1]{\textcolor{black}{#1}}
    - \usepackage{graphicx}
@@
 Additionally, we may hypothesize that RTs are affected by whether the target digit was smaller or greater than 5, yet it is unclear which side of 5 would lead to faster or slower RTs.
-\textcolor{blue}{The digit classification task is a standard task that usually would be analyzed in a 3 (digit, closest to 5, further away from 5, furthest away from 5) by 2 (side, greater or smaller than 5) within-subjects ANOVA.} In this tutorial, we show how inferences can be improved using a hierarchical Bayesian model instead.
+\Revision{The digit classification task is a standard task that usually would be analyzed in a 3 (digit, closest to 5, further away from 5, furthest away from 5) by 2 (side, greater or smaller than 5) within-subjects ANOVA.} In this tutorial, we show how inferences can be improved using a hierarchical Bayesian model instead.
 

### 2023-07-12 manuscript.tex
Remove blue color from manuscript

@@
 \newcommand{\Julia}[1]{\todo[inline, color=mypink]{#1}}
-\newcommand{\Revision}[1]{\textcolor{blue}{#1}}
+\newcommand{\Revision}[1]{\textcolor{black}{#1}}
 \usepackage{graphicx}
@@
 
-Our tutorial is directed at researchers who have a basic \textcolor{blue}{conceptual} understanding of Bayesian inference and parameter estimation and are planning to perform more complex Bayesian analyses \textcolor{blue}{of psychological phenomena.
-We hope to achieve a balance of detail and overview---while we want to provide guidance for a full Bayesian workflow, similar to Gelman et al.} (2020) \textcolor{blue}{and Schad et al.}(2021; 2022)\textcolor{blue}{, we don't shy away from in-depth explanations for the most curious scientists.
+Our tutorial is directed at researchers who have a basic \textcolor{black}{conceptual} understanding of Bayesian inference and parameter estimation and are planning to perform more complex Bayesian analyses \textcolor{black}{of psychological phenomena.
+We hope to achieve a balance of detail and overview---while we want to provide guidance for a full Bayesian workflow, similar to Gelman et al.} (2020) \textcolor{black}{and Schad et al.}(2021; 2022)\textcolor{black}{, we don't shy away from in-depth explanations for the most curious scientists.
 To ensure that a broad audience is able to follow the tutorial, we limit and carefully explain equations.
@@
 Additionally, we may hypothesize that RTs are affected by whether the target digit was smaller or greater than 5, yet it is unclear which side of 5 would lead to faster or slower RTs.
-\textcolor{blue}{The digit classification task is a standard task that usually would be analyzed in a 3 (digit, closest to 5, further away from 5, furthest away from 5) by 2 (side, greater or smaller than 5) within-subjects ANOVA.} In this tutorial, we show how inferences can be improved using a hierarchical Bayesian model instead.
+\textcolor{black}{The digit classification task is a standard task that usually would be analyzed in a 3 (digit, closest to 5, further away from 5, furthest away from 5) by 2 (side, greater or smaller than 5) within-subjects ANOVA.} In this tutorial, we show how inferences can be improved using a hierarchical Bayesian model instead.
 
@@
 
-\textcolor{blue}{A pattern where the general trend does not reflect individual trends is shown in Panel A of Figure} \ref{fig:figure1}\textcolor{blue}{.
+\textcolor{black}{A pattern where the general trend does not reflect individual trends is shown in Panel A of Figure} \ref{fig:figure1}\textcolor{black}{.
 The slopes of the grey lines reflect individual trends, and some of them differ considerably from the general trend shown as the slope of the blue line.} If every participant takes part in each condition, the person variability affects all conditions, thus inducing a correlation across conditions.
 For example, participants in a digit classification task repeatedly respond to target digits inducing a correlation between the responses to the target digits, such as between target digit 3 and target digit 4.
-\textcolor{blue}{Therefore, analyses on aggregated data do not appropriately account for differential experiment effects} (Doorn, Aust, Haaf, Stefan, \& Wagenmakers, 2021, 2023).
+\textcolor{black}{Therefore, analyses on aggregated data do not appropriately account for differential experiment effects} (Doorn, Aust, Haaf, Stefan, \& Wagenmakers, 2021, 2023).
 As a result, the type I error rate of a test of the effect increases: there is a higher probability of concluding that there is an effect if there is no effect (Rouder \& Lu, 2005).
@@
 
-\textcolor{blue}{It is clear that neither of the alternative approaches yield satisfactory results by themselves.
+\textcolor{black}{It is clear that neither of the alternative approaches yield satisfactory results by themselves.
 Conceptually, a hierarchical analysis provides a compromise between the aggregated and independent approach by drawing benefits from each method.
-The motivation for hierarchical models, as well as their general mathematical structure, has been covered by several (tutorial) articles} (e.g., Efron \& Morris, 1977; Singmann \& Kellen, 2019; West, Welch, \& Galecki, 2022)\textcolor{blue}{, as well as text books} (e.g., Gelman \& Hill, 2006; Hox, Moerbeek, \& Van de Schoot, 2017; Nicenboim, Schad, \& Vasishth, 2023).
-
-\textcolor{blue}{In hierarchical analyses, the general or overall effect is sometimes called fixed effect, and the individual-specific deviations from this effect are in that case referred to as random effects (Rouder \& Lu, 2005).
+The motivation for hierarchical models, as well as their general mathematical structure, has been covered by several (tutorial) articles} (e.g., Efron \& Morris, 1977; Singmann \& Kellen, 2019; West, Welch, \& Galecki, 2022)\textcolor{black}{, as well as text books} (e.g., Gelman \& Hill, 2006; Hox, Moerbeek, \& Van de Schoot, 2017; Nicenboim, Schad, \& Vasishth, 2023).
+
+\textcolor{black}{In hierarchical analyses, the general or overall effect is sometimes called fixed effect, and the individual-specific deviations from this effect are in that case referred to as random effects (Rouder \& Lu, 2005).
 We can also estimate individual effects, not as deviations from an overarching effect, but as effects for each individual, as obtained from a per-participant analysis.
@@
 
-\textcolor{blue}{The model equation enables us to compute the expected response time $Y_{ij}$ for a participant $i$ when presented with a specific probe.
+\textcolor{black}{The model equation enables us to compute the expected response time $Y_{ij}$ for a participant $i$ when presented with a specific probe.
 For example, if participant $i = 1$ is presented with digit $j = 7$, $x_j$ takes the value $-\frac{1}{2}$ and $u_j$ takes the value $1$.
@@
 For example, the data in the example are response times.
-\textcolor{blue}{Response times cannot be negative.
+\textcolor{black}{Response times cannot be negative.
 The normal model, however, can yield negative values.
@@
 
-\textcolor{blue}{Note that the only change in the model formulation is the change from a normal distribution to a log-normal distribution.
+\textcolor{black}{Note that the only change in the model formulation is the change from a normal distribution to a log-normal distribution.
 The density function describes the probability model that is assumed to be underlying the data.
@@
 
-\textcolor{blue}{Up until now, we specified models without a correlation between the individual effects.
+\textcolor{black}{Up until now, we specified models without a correlation between the individual effects.
 This means, for instance, that we do not expect the estimations for the individual effects $\delta_{7, i}$ and $\delta_{3, i}$ to be related.
@@
 
-Another important issue is that wide priors are often \textcolor{blue}{improper probability distributions} (Hobert \& Casella, 1996).
-\textcolor{blue}{A prior is proper if two conditions hold: 1.
+Another important issue is that wide priors are often \textcolor{black}{improper probability distributions} (Hobert \& Casella, 1996).
+\textcolor{black}{A prior is proper if two conditions hold: 1.
 All values of the probability distribution are equal or greater than zero, and 2.
@@
 An example of a proper and improper probability distribution are provided in Figure} \ref{fig:properfigure}.
-\textcolor{blue}{The figure shows two hypothetical priors for parameter $\theta$.
+\textcolor{black}{The figure shows two hypothetical priors for parameter $\theta$.
 The first condition can be assessed by checking the y-values---both priors have only positive values as function output (0.1, 0.2, and 0.4).
@@
 
-A \textcolor{blue}{more common} example of an improper prior is a uniform prior distribution ranging from minus infinity to infinity.
-In case of improper priors, \textcolor{blue}{it is often impossible to obtain correct estimation of general and individual effects in hierarchical models because the resulting posteriors will again be improper probability distributions.} However, even proper priors can result in improper posteriors (Hobert \& Casella, 1996).
+A \textcolor{black}{more common} example of an improper prior is a uniform prior distribution ranging from minus infinity to infinity.
+In case of improper priors, \textcolor{black}{it is often impossible to obtain correct estimation of general and individual effects in hierarchical models because the resulting posteriors will again be improper probability distributions.} However, even proper priors can result in improper posteriors (Hobert \& Casella, 1996).
 Although estimation should not be possible in this situation, statistical software might still provide results without notifying the user that the posterior distribution does not exist.
@@
 Therefore, it is important to check if \emph{(a)} the priors are proper and \emph{(b)} the priors result in proper posteriors.
-\textcolor{blue}{Unfortunately,} there is no fail-safe method that guarantees proper posteriors for hierarchical models.
+\textcolor{black}{Unfortunately,} there is no fail-safe method that guarantees proper posteriors for hierarchical models.
 The best approach is to choose well-reasoned informative priors for the model, for instance, by using the proposed four-step procedure below.
-\textcolor{blue}{Priors that are proper and not too diffuse will in most cases lead to proper posteriors.}
+\textcolor{black}{Priors that are proper and not too diffuse will in most cases lead to proper posteriors.}
 
@@
 All variance terms (\(\sigma_{\gamma}^2\), \(\sigma_{\beta}^2\), \(\sigma_{\delta}^2\), and \(\sigma^2\)) are constrained to be positive.
-If there is no information from previous studies, researchers might be inclined to use the default priors \textcolor{blue}{of software packages, often reasonably uninformative priors,} as a starting point.
-
-\textcolor{blue}{We take the reasonably uninformative priors as a starting point as well specifying priors on different types of parameters.
+If there is no information from previous studies, researchers might be inclined to use the default priors \textcolor{black}{of software packages, often reasonably uninformative priors,} as a starting point.
+
+\textcolor{black}{We take the reasonably uninformative priors as a starting point as well specifying priors on different types of parameters.
 These priors were previously used as} default priors in \texttt{brms} (Bürkner, 2017).
@@
 Note, however, that the transformation of the prior distributions is not as simple as the transformation of the parameter itself.
-\textcolor{blue}{If we formulate a prior on standard deviations, there is always an implied prior on the variances as well.
+\textcolor{black}{If we formulate a prior on standard deviations, there is always an implied prior on the variances as well.
 If we think of the distribution as $f(x)$, $x$ can either be defined as $\sqrt{(var)}$, if we formulated the prior on the standard deviation, or $x$ can be defined as ${sd}^2$ if we formulated the prior on the variance.
@@
 The blue dots represent the mean effect.
-Using the \textcolor{blue}{reasonably uninformatieve priors}, the mean effects are expected to be around zero, with a mean effect of zero.
+Using the \textcolor{black}{reasonably uninformatieve priors}, the mean effects are expected to be around zero, with a mean effect of zero.
 However, the range around this mean is very wide, ranging from -10 to 10 second effects.
@@
 Finally, we would like to use a more common conjugate prior for the variances to ensure obtaining a proper posterior distribution.\footnote{Conjugate priors result in posterior distributions that belong to the same distribution family as the prior distribution (Fink, 1997a).
-For a normal model, these are the normal distribution for the mean and the inverse gamma distribution for the variance.} \footnote{\textcolor{blue}{However, in rstan, non-conjugate priors are commonly applied} (e.g., Gelman, 2006).} Therefore, we choose an inverse gamma distribution.
+For a normal model, these are the normal distribution for the mean and the inverse gamma distribution for the variance.} \footnote{\textcolor{black}{However, in rstan, non-conjugate priors are commonly applied} (e.g., Gelman, 2006).} Therefore, we choose an inverse gamma distribution.
 We can also be more specific about the expected RTs.
@@
 As before, we prefer to apply conjugate priors.
-\textcolor{blue}{The conjugate priors that we applied to the normal model are also conjugate priors for the log-normal model} (Fink, 1997b).
-\textcolor{blue}{For instance, with a normal distribution as prior for the mean of the log-normal model, we assume that the mean of the parameter can be either negative, zero, or positive, and assign a higher probability to values closer to zero.
+\textcolor{black}{The conjugate priors that we applied to the normal model are also conjugate priors for the log-normal model} (Fink, 1997b).
+\textcolor{black}{For instance, with a normal distribution as prior for the mean of the log-normal model, we assume that the mean of the parameter can be either negative, zero, or positive, and assign a higher probability to values closer to zero.
 If the mean of the log-normal distribution is smaller than zero, this means that the peak of the log-normal distribution moves closer to zero, i.e., that response times are quicker but still larger than zero.
@@
 Every sampling algorithm requires the specification of the number of samples (also called iterations) that are drawn from the posterior distribution.
-Furthermore, all MCMC sampling methods have in common that results depend (somewhat) on chosen starting values, that is, initial parameter values that \textcolor{blue}{are fed to the sampler in the first iteration}.
+Furthermore, all MCMC sampling methods have in common that results depend (somewhat) on chosen starting values, that is, initial parameter values that \textcolor{black}{are fed to the sampler in the first iteration}.
 This dependency can be assessed if the sampling algorithm is repeated several times (i.e., chains) with different starting values.
@@
 \texttt{Rstan} requires two files: a \texttt{.stan} file containing the model specification and an R script containing the code for the model fitting.
-In the \texttt{.stan} file, the parameters, priors, and the \textcolor{blue}{probability distribution of the data} are specified.
+In the \texttt{.stan} file, the parameters, priors, and the \textcolor{black}{probability distribution of the data} are specified.
 An example \texttt{.stan} file of a simple, non-hierarchical normal model, where the variance is already known and the mean has to be estimated, is shown below (adapted example from Nicenboim, Schad, \& Vasishth, 2021, Chapter 10).
@@
 The last section, \texttt{model}, contains the probability function and priors, as specified earlier in this article.
-\textcolor{blue}{Instead of using} \texttt{y{[}i{]}\ \textasciitilde{}\ normal(mu,\ sigma)} \textcolor{blue}{to estimate parameters, such as the mean from the example, we used the} \texttt{target} \textcolor{blue}{specification} (for a in-depth explanation see Nicenboim, Schad, \& Vasishth, 2021, Chapter 10).
-\textcolor{blue}{Both methods indicate in this case that} \texttt{y{[}i{]}} \textcolor{blue}{is normally distributed with the mean} \texttt{mu} \textcolor{blue}{and the variance} \texttt{sigma}.
-\textcolor{blue}{However, the target method saves constant terms that are required for model comparison} (Gronau, Singmann, \& Wagenmakers, 2020) The \texttt{.stan} file for the symbolic distance effect, including explanations, can be found in \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/blob/main/R\%20objects/rstan/normal\%20model/myModel.stan}{Online Supplement K}.
+\textcolor{black}{Instead of using} \texttt{y{[}i{]}\ \textasciitilde{}\ normal(mu,\ sigma)} \textcolor{black}{to estimate parameters, such as the mean from the example, we used the} \texttt{target} \textcolor{black}{specification} (for a in-depth explanation see Nicenboim, Schad, \& Vasishth, 2021, Chapter 10).
+\textcolor{black}{Both methods indicate in this case that} \texttt{y{[}i{]}} \textcolor{black}{is normally distributed with the mean} \texttt{mu} \textcolor{black}{and the variance} \texttt{sigma}.
+\textcolor{black}{However, the target method saves constant terms that are required for model comparison} (Gronau, Singmann, \& Wagenmakers, 2020) The \texttt{.stan} file for the symbolic distance effect, including explanations, can be found in \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/blob/main/R\%20objects/rstan/normal\%20model/myModel.stan}{Online Supplement K}.
 
@@
 The model is fitted using the packages \texttt{rstan} and \texttt{brms}.
-\textcolor{blue}{We used different settings for} \texttt{rstan} \textcolor{blue}{and} \texttt{brms} \textcolor{blue}{due to the difference between the models for these packages.
+\textcolor{black}{We used different settings for} \texttt{rstan} \textcolor{black}{and} \texttt{brms} \textcolor{black}{due to the difference between the models for these packages.
 Although we tried to make the models as similar as possible, the models are not identical.
@@
 This indicates that the chains have not converged.
-\textcolor{blue}{The current adhered criteria is that $\hat{R}$ should not exceed 1.01 }(Vehtari, Gelman, Simpson, Carpenter, \& Bürkner, 2021b).
+\textcolor{black}{The current adhered criteria is that $\hat{R}$ should not exceed 1.01 }(Vehtari, Gelman, Simpson, Carpenter, \& Bürkner, 2021b).
 There are different types of \(\hat{R}\), such as the split- and rank-based \(\hat{R}\) (Gelman et al., 2013; Vehtari, Gelman, Simpson, Carpenter, \& Bürkner, 2021b).
 Even though researchers tend to trust convergence statistics more than visual inspection, these statistics are also not fool-proof.
-\textcolor{blue}{For instance, in hierarchical modeling, with many parameters, it is more likely that parameters will exceed the threshold while there are no issues with the parameter estimation.
+\textcolor{black}{For instance, in hierarchical modeling, with many parameters, it is more likely that parameters will exceed the threshold while there are no issues with the parameter estimation.
 It is, therefore, advisable to evaluate parameters carefully.} The development of better convergence statistics is a current topic in Bayesian analysis methods.
@@
 The \(\hat{R}\) values for all individual and general symbolic distance effect parameters computed by \texttt{rstan} and \texttt{brms} are shown in panel C of Figure \ref{fig:traceplots3p}.
-The figure shows that \(\hat{R}\)'s are not substantially greater than \textcolor{blue}{1.01}.
+The figure shows that \(\hat{R}\)'s are not substantially greater than \textcolor{black}{1.01}.
 Therefore, it seems that the distributions have converged.
@@
 Therefore, the number of effective samples can be seen as a measure of the amount of new information about the posterior distribution that is provided by the total number of samples drawn from the posterior.
-\textcolor{blue}{A common rule for interpretation is that the number of effective samples should equal at least 100 per chain} (Vehtari, Gelman, Simpson, Carpenter, \& Bürkner, 2021b).
-\textcolor{blue}{In the case of four chains, this would mean a value of 400 or higher.
+\textcolor{black}{A common rule for interpretation is that the number of effective samples should equal at least 100 per chain} (Vehtari, Gelman, Simpson, Carpenter, \& Bürkner, 2021b).
+\textcolor{black}{In the case of four chains, this would mean a value of 400 or higher.
 The higher the value the better.
@@
 Bulk-ESS evaluates the center of the posterior distribution, while Tail-ESS evaluates the tails of the posterior distribution.
-For this tutorial, we focus on the basic calculation of the number of effective samples suggested by Gelman et al.} (2013) \textcolor{blue}{and improved by Vehtari et al.} (2021b).
+For this tutorial, we focus on the basic calculation of the number of effective samples suggested by Gelman et al.} (2013) \textcolor{black}{and improved by Vehtari et al.} (2021b).
 
@@
 Here, this was not case, and all other diagnostics were good, so we continue with the interpretation of the estimates.
-\textcolor{blue}{The difference between the model diagnostics of} \texttt{rstan} \textcolor{blue}{and} \texttt{brms} \textcolor{blue}{can be partly explained by the difference in model parametrization, such as the difference in the prior specification (i.e., specifying priors on variances versus priors on standard deviations).
+\textcolor{black}{The difference between the model diagnostics of} \texttt{rstan} \textcolor{black}{and} \texttt{brms} \textcolor{black}{can be partly explained by the difference in model parametrization, such as the difference in the prior specification (i.e., specifying priors on variances versus priors on standard deviations).
 We equated the models where sensible which resulted in models that are fairly comparable.
@@
 The credible interval displayed in the figure is the Bayesian version of a confidence interval and provides a measure of uncertainty about the parameter value (Wagenmakers et al., 2018).
-\textcolor{blue}{In contrast to the confidence interval, the credible interval provides the probability that a parameter lies within a certain range.
+\textcolor{black}{In contrast to the confidence interval, the credible interval provides the probability that a parameter lies within a certain range.
 However, the credible interval does not provide evidence to conclude whether an effect is zero or not }(Wagenmakers, Lee, Rouder, \& Morey, 2020).
@@
 The posterior distribution of the side effect \(\mu_{\beta}\) is centered around zero with a posterior mean of -0.007 and a 95\% credible interval ranging from -0.046 to 0.031.
-\textcolor{blue}{Note that the credible interval overlaps with zero.
+\textcolor{black}{Note that the credible interval overlaps with zero.
 Commonly, this is misinterpreted as an absence of an effect.
@@
 There is also considerable individual variation in the digit effects (\(\delta\)'s).
-\textcolor{blue}{However, most of the intervals contain zero which implies that the effects might be small or non-existent.} The variability is most pronounced for the effect parameters of the digits close to 5.
+\textcolor{black}{However, most of the intervals contain zero which implies that the effects might be small or non-existent.} The variability is most pronounced for the effect parameters of the digits close to 5.
 For these digits, there are also more intervals that do not contain zero, indicating that an effect might be present on the subject level.
@@
 
-\textcolor{blue}{The individual variation can also be visualized differently, for example by connecting individual estimates for each parameter.
+\textcolor{black}{The individual variation can also be visualized differently, for example by connecting individual estimates for each parameter.
 This type of variation can reveal correlations between individual parameter estimates, for example, participants scoring high on one digit parameter may also score high on another digit parameter.} Figure \ref{fig:modelestimatesfigure3} shows the relationship between individuals' digit effects.\footnote{The design and code for this figure are by Langen (2020).} The posterior means of individual effects are represented by the dots.
@@
 Calculating the mean estimate of \(\mu_{\delta_{3}}\) would be similar, but more complicated as the mean of a log-normal distribution also depends on its variance.
-\textcolor{blue}{This approach can also be used iteratively.
+\textcolor{black}{This approach can also be used iteratively.
 This means that the approach above is applied to every iteration in the chain.
@@
 The side effect is close to zero, with a posterior mean of -0.008 and a 95\% credible interval from -0.022 to 0.007.
-\textcolor{blue}{This indicates that the general effect of side is small.}
+\textcolor{black}{This indicates that the general effect of side is small.}
 
@@
 However, compared to the normal model there is considerably more hierarchical shrinkage.
-\textcolor{blue}{This means that in the log-normal model, the individual effects are corrected more towards the group mean, compared to the normal model.
+\textcolor{black}{This means that in the log-normal model, the individual effects are corrected more towards the group mean, compared to the normal model.
 Or, the group mean in the log-normal model is not affected as much by individuals showing a divergent trend compared to the normal model.
@@
 Specifically, we will achieve this by comparing the predictive accuracy of a null model where we set one or some parameters to zero, and and effects models where these parameters are unconstrained.
-\textcolor{blue}{This approach is called Bayes factor model comparison.}
+\textcolor{black}{This approach is called Bayes factor model comparison.}
 
@@
 
-\textcolor{blue}{Rather than submerging us in technical details or providing a formal justification of the Bayesian approach, we provide an intuition of Bayesian model comparison.
+\textcolor{black}{Rather than submerging us in technical details or providing a formal justification of the Bayesian approach, we provide an intuition of Bayesian model comparison.
 A formal justification of Bayes factors is provided by} Jeffreys (1961), Kass and Raftery (1995) and Rouder and Morey (2018).
-\textcolor{blue}{The Bayes factor (BF) is a measure of how well one model predicts the data compared to another.}
-
-\textcolor{blue}{Bayesian analysis allows for predictions of data from any model because models are fully specified with a probability distribution for the data and priors for all parameters.
+\textcolor{black}{The Bayes factor (BF) is a measure of how well one model predicts the data compared to another.}
+
+\textcolor{black}{Bayesian analysis allows for predictions of data from any model because models are fully specified with a probability distribution for the data and priors for all parameters.
 Let the vector of all parameters of a model be $\bm{\theta}$.
@@
 
-\textcolor{blue}{Note that $\int_{\bm{\Theta}}$ refers to a multidimensional integral over the entire parameter space.
+\textcolor{black}{Note that $\int_{\bm{\Theta}}$ refers to a multidimensional integral over the entire parameter space.
 As our models consist of many parameters, computing these predictions for any one model becomes somewhat difficult.
@@
 
-\textcolor{blue}{There is a second popular interpretation of the Bayes factor as the relative evidence of two competing models.
+\textcolor{black}{There is a second popular interpretation of the Bayes factor as the relative evidence of two competing models.
 One of the key insights from Bayes' rule is that these two things, the relative predictive accuracy of two models for the data, and the relative evidence from the data for the two models, are one and the same} (Rouder \& Morey, 2018).
@@
 This means that the data are 100 times more likely under the normal model than under the log-normal model, clear evidence in favor of the normal model.
-\textcolor{blue}{But how do we know whether a Bayes factor of 100-to-1, or 10-to-1 or 3-to-1 is big enough?} As a reminder, Bayes factors are the relative predictive accuracy of two models for the observed data.
+\textcolor{black}{But how do we know whether a Bayes factor of 100-to-1, or 10-to-1 or 3-to-1 is big enough?} As a reminder, Bayes factors are the relative predictive accuracy of two models for the observed data.
 Therefore, Bayes factors are odds or ratios.
@@
 
-\textcolor{blue}{Same as the posterior distribution, Bayes factors can rarely be computed analytically.} The difficulty to compute Bayes factors is one of the major drawbacks of Bayesian inference.
+\textcolor{black}{Same as the posterior distribution, Bayes factors can rarely be computed analytically.} The difficulty to compute Bayes factors is one of the major drawbacks of Bayesian inference.
 However, several algorithms are available to estimate Bayes factor from posterior samples.
 In the following section we will present two methods to obtain Bayes factors for the symbolic distance effect.
-\textcolor{blue}{The first method that we discuss, Savage-Dickey density ratio, tests the general effect while disregarding individual variation (similar to type III sums of squares in ANOVA models).
+\textcolor{black}{The first method that we discuss, Savage-Dickey density ratio, tests the general effect while disregarding individual variation (similar to type III sums of squares in ANOVA models).
 This means that the compared models differ in only a single parameter, that is, the general effect parameter, and individual effects are present in both models.
@@
 There have been discussions on which method should be preferred without consensus} (Doorn, Aust, Haaf, Stefan, \& Wagenmakers, 2021, 2023; Rouder, Schnuerch, Haaf, \& Morey, 2023; Singmann et al., 2023).
-\textcolor{blue}{Therefore, we illustrate the use of the two approaches and discuss their potential.}
+\textcolor{black}{Therefore, we illustrate the use of the two approaches and discuss their potential.}
 
@@
 
-\textcolor{blue}{Note that models 2-4 are nested in the most complex full model, and the null model is nested in models 2 and 3, but models 2 and 3 are not nested}.
+\textcolor{black}{Note that models 2-4 are nested in the most complex full model, and the null model is nested in models 2 and 3, but models 2 and 3 are not nested}.
 Additionally, the models differ by several parameters.
@@
 However, the Bayes factor estimates in \texttt{rstan} generally show stronger evidence than Bayes factor estimates based on \texttt{brms}.
-\textcolor{blue}{It is important to note that this difference is not due to the bridge sampling function (i.e., there is a little to no variation of Bayes factor estimates within packages), but a result of a difference in parametrization.
-The models for} \texttt{rstan} \textcolor{blue}{and} \texttt{brms} \textcolor{blue}{are similar but not identical, partly due to the difference in prior specification (i.e., variances versus standard deviations).} However, for both packages, the conclusion is the same: the data are least likely to have occurred under the full model.
+\textcolor{black}{It is important to note that this difference is not due to the bridge sampling function (i.e., there is a little to no variation of Bayes factor estimates within packages), but a result of a difference in parametrization.
+The models for} \texttt{rstan} \textcolor{black}{and} \texttt{brms} \textcolor{black}{are similar but not identical, partly due to the difference in prior specification (i.e., variances versus standard deviations).} However, for both packages, the conclusion is the same: the data are least likely to have occurred under the full model.
 In case the estimates are unstable, it is recommended to further increase the number of posterior samples.
@@
 Then, we can calculate the posterior model probabilities for all specified models.
-\textcolor{blue}{Typically in sensitivity analyses, the sensitivity of the Bayes factor on the width of the prior on effect size is assessed} (e.g., van Doorn et al., 2021)\textcolor{blue}{.
+\textcolor{black}{Typically in sensitivity analyses, the sensitivity of the Bayes factor on the width of the prior on effect size is assessed} (e.g., van Doorn et al., 2021)\textcolor{black}{.
 However, in our case the prior distributions on the variance parameters are key to understand sensitivity.}
@@
 This includes the trivial Bayes factor of the null model against the null model.
-\textcolor{blue}{This setup results in 10 models, and, therefore, 10 Bayes factors.
+\textcolor{black}{This setup results in 10 models, and, therefore, 10 Bayes factors.
 We assume that all models are equally likely, resulting in a prior model probability for each model of 1/10.} Next, we calculate the sum of all these Bayes factors and divide each Bayes factor by this sum.
@@
 In addition, the models with smaller prior expectations for individual differences (scenario 2), have higher posterior model probabilities than the models in scenario 1.
-\textcolor{blue}{Thus, decreasing the variability of the size of the individual digit and side effects (scenario 2), or completely removing the variance (scenario 3), results in higher posterior model probabilities.} Together with Figure \ref{fig:randomparfigure}, showing the small differences in individual effects (also illustrated by Haaf et al., 2019), this result supports the notion that side and digit effects are small and there is a lack in individual variability.
+\textcolor{black}{Thus, decreasing the variability of the size of the individual digit and side effects (scenario 2), or completely removing the variance (scenario 3), results in higher posterior model probabilities.} Together with Figure \ref{fig:randomparfigure}, showing the small differences in individual effects (also illustrated by Haaf et al., 2019), this result supports the notion that side and digit effects are small and there is a lack in individual variability.
 
 \hypertarget{section}{%
-\subsubsection{\texorpdfstring{\textcolor{blue}{Log-normal Model}}{}}\label{section}}
-
-\textcolor{blue}{We can take a similar approach for the log-normal model to compare the hierarchical models on varying complexity.
+\subsubsection{\texorpdfstring{\textcolor{black}{Log-normal Model}}{}}\label{section}}
+
+\textcolor{black}{We can take a similar approach for the log-normal model to compare the hierarchical models on varying complexity.
 We evaluate their performance in two scenarios that correspond to scenario 1 and 3 for the normal model.
@@
 
-\textcolor{blue}{The table shows that the digit model without individual digit effects from scenario 2 has the highest posterior model probability.
+\textcolor{black}{The table shows that the digit model without individual digit effects from scenario 2 has the highest posterior model probability.
 This is in line with our findings for the normal model.
@@
 For the non-technical user, we recommend using \texttt{brms} because of its intuitive model specification.
-However, the default prior setup can be unsuitable for a specific research context and \textcolor{blue}{(virtually) always needs} to be adapted.
+However, the default prior setup can be unsuitable for a specific research context and \textcolor{black}{(virtually) always needs} to be adapted.
 Additionally, users are limited to predefined parametrizations.
@@
 Vasishth (2021) and Schad, Nicenboim, Bürkner, Betancourt, and Vasishth (2022).
-\textcolor{blue}{There are also other R software packages for Bayesian multilevel modeling besides} \texttt{brms} \textcolor{blue}{and} \texttt{rstan}\textcolor{blue}{, such as} \texttt{BayesFactor}(Morey \& Rouder, 2018) \textcolor{blue}{and} \texttt{rstanarm}(Goodrich, Gabry, Ali, \& Brilleman, 2020)\textcolor{blue}{, that we did not illustrate in this tutorial.
+\textcolor{black}{There are also other R software packages for Bayesian multilevel modeling besides} \texttt{brms} \textcolor{black}{and} \texttt{rstan}\textcolor{black}{, such as} \texttt{BayesFactor}(Morey \& Rouder, 2018) \textcolor{black}{and} \texttt{rstanarm}(Goodrich, Gabry, Ali, \& Brilleman, 2020)\textcolor{black}{, that we did not illustrate in this tutorial.
 However, most of these other R software packages are more limited, for instance, in model specification and prior settings.}
@@
 \hypertarget{section-1}{%
-\subsection{\texorpdfstring{\textcolor{blue}{Technical Possibilities are Endless}}{}}\label{section-1}}
+\subsection{\texorpdfstring{\textcolor{black}{Technical Possibilities are Endless}}{}}\label{section-1}}
 
@@
 We already highlighted the drawbacks of the SD density ratio.
-\textcolor{blue}{Bayes factor estimates can also be biased (i.e., display a consistent deviation from the true Bayes factor value).
+\textcolor{black}{Bayes factor estimates can also be biased (i.e., display a consistent deviation from the true Bayes factor value).
 However, bridge sampling has shown to be relatively unbiased} (Sarafoglou et al., 2021).

### 2023-07-12 manuscript.Rmd
Remove blue text

@@
 The `R` package `brms` also allows fitting Bayesian models using `Stan`.
-However, model specification and model fitting can be achieved using a much simplified `R`-script \textcolor{blue}{alone}.
+However, model specification and model fitting can be achieved using a much simplified `R`-script \Revision{alone}.
 Model specification is based on a formula syntax that is similar to the notation in the popular `lme4` package for frequentist hierarchical modeling [@R-lme4].
@@
 Finally, we compare the estimated density of the posterior to the density of the prior.
-\textcolor{blue}{This process is explained in further detail by Van Ravenzwaaij and Etz} [-@van2021simulation].
+\Revision{This process is explained in further detail by Van Ravenzwaaij and Etz} [-@van2021simulation].
 

### 2023-07-12 manuscript.tex
Remove blue text

@@
 The \texttt{R} package \texttt{brms} also allows fitting Bayesian models using \texttt{Stan}.
-However, model specification and model fitting can be achieved using a much simplified \texttt{R}-script \textcolor{blue}{alone}.
+However, model specification and model fitting can be achieved using a much simplified \texttt{R}-script \textcolor{black}{alone}.
 Model specification is based on a formula syntax that is similar to the notation in the popular \texttt{lme4} package for frequentist hierarchical modeling (Bates, Mächler, Bolker, \& Walker, 2015).
@@
 Finally, we compare the estimated density of the posterior to the density of the prior.
-\textcolor{blue}{This process is explained in further detail by Van Ravenzwaaij and Etz} (2021).
+\textcolor{black}{This process is explained in further detail by Van Ravenzwaaij and Etz} (2021).
 
